### 2011-07-13 doc/manuscript.tex
Manual is being edited.

%% LyX 1.6.7 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage[letterpaper]{geometry}
\geometry{verbose,tmargin=1in,bmargin=1in}
\setlength{\parskip}{\bigskipamount}
\setlength{\parindent}{0pt}
\usepackage{url}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage[authoryear]{natbib}
\usepackage{longtable}
\usepackage{epstopdf}
\doublespacing

\newcommand{\chck}{\textbf{[CHECK]}}
\newcommand{\addfig}{\textbf{[ADD FIGURE]}}
\newcommand{\tba}{\textbf{[TO BE ADDED]}}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}
%% A simple dot to overcome graphicx limitations
\newcommand{\lyxdot}{.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\newcommand{\lyxaddress}[1]{
\par {\raggedright #1
\vspace{1.4em}
\noindent\par}
}

\@ifundefined{showcaptionsetup}{}{%
 \PassOptionsToPackage{caption=false}{subfig}}
\usepackage{subfig}
\makeatother

\usepackage{babel}

\begin{document}
\renewcommand\refname{Literature Cited}

\raggedright

\subsubsection*{Title: }

Inferring two types of gene transfer using whole genomes of \emph{Streptococcus
pyogenes} and\emph{ S.\ dysgalactiae} subspecies \emph{equisimilis}

\subsubsection*{Authors:}

Sang Chul Choi$^{1}$, Matthew D. Rasmussen$^{1}$, Melissa Jane Hubisz$^{1}$, Ilan 
Gronau$^{1}$,
Michael J. Stanhope$^{2}$, and Adam Siepel$^{1}$

\subsubsection*{Affiliations:}

\lyxaddress{$^{1}$Department of Biological Statistics and Computational Biology,
Cornell University, Ithaca, NY 14853, USA.}

\lyxaddress{$^{2}$Department of Population Medicine and Diagnostic Sciences,
College of Veterinary Medicine, Cornell University, Ithaca, NY 14853,
USA.}

\subsubsection*{Author for Correspondence:}

Adam Siepel, Department of Biological Statistics and Computational
Biology, Cornell University, Ithaca, NY 14853, USA, Phone: 607-254-1157,
Fax: 607-255-4698, Email: acs4@cornell.edu\clearpage{}

\subsubsection*{Abbreviations:}

phosphotransferase system (PTS), ATP-binding cassette (ABC), Aminoacyl
tRNA synthetase (aaRS), ancestral recombination graph (ARG), group
A streptococci (GAS), group C streptococci (GCS), group G streptococci
(GGS), \emph{Streptococcus pyogenes} (SPY), \emph{S.\ dysgalactiae}
subspecies \emph{dysgalactiae} (SDD), \emph{S.\ dysgalactiae} subspecies
\emph{equisimilis} (SDE).

\subsubsection*{Keywords:}

bacterial recombination, gene duplication and transfer, ClonalOrigin,
ClonalFrame, Mowgli

\subsubsection*{Version: 0.65 (July 8, 2011) \clearpage{}}

\section*{Abstract}

Using two closely related species of \emph{Streptococcus pyogenes}
(SPY) that is a pervasive human-specific pathogen, and \emph{S.\ dysgalactiae}
ssp.\emph{\ equisimilis} (SDE) that is its closely related species
and an opportunistic human pathogen we investigated two types of gene
transfer: recombining and non-recombining
gene transfer. The former lateral process was studied
with the core genomes of \emph{Streptococcus} genomes, and the latter
process with the pan genomes as well as the core genomes. Two
complementary methods were employed to examine genes and their
fragments that could have been exchanged between SPY and SDE as efforts
of finding their roles in the formation of pathogenicity of the bacteria.
We revisited the problem of directionality of gene transfer between
SPY and SDE, and found evidence supporting that homologous
recombining gene transfer was polarized to be relatively from SPY to SDE based
on the core genomes.  Although we did not find evidence for directional
non-recombining gene transfer, virulence genes appeared to undergo
non-recombining gene transfer more than did non-virulence genes.
Recombining gene transfer was not found to be associated with
virulence genes.  This might indicate that additional genes obtained via
non-recombining gene transfer were more effective than gene fragments
transfer via bacterial recombination for SDE in the acquisition of virulence
genes.  While horizontal gene transfer studies traditionally 
focused on genes as a whole,
we rediscovered smaller gene fragments and their transfer
due to recombination between SPY and SDE in aminoacyl tRNA
synthetase (aaRS), and phosphotransferase systems (PTS).

\clearpage{}

\section{Introduction}

\textbf{Two lateral processes of gene transfer:} 
Foreign bacterial genes can be laterally
brought into a bacterial genome by two different processes 
\citep{Ochman2001,Lawrence2009}:
homologous recombining transfer and non-recombining transfer.
In the former process donor genes that are homologous to those in
a recipient species replace the indigenous genes. In the latter process
donor genes are integrated into a recipient species without replacement
of genes in the species. 
Although the discrimination between the two
lateral processes was noted a decade ago, methods of detecting horizontal
gene transfer did not appear to properly take the differences into
account. 
Proper discrimination of the two processes of gene transfer
could lead to better understanding of biological nature
of transferred genes. For example, 
bacterial lateral transfers of small fragments of ribosomal RNA genes 
were known to be 
due to homologous recombination \citep{Sneath1993,Yap1999} not non-recombining
gene transfer. 
Because traditionally non-recombining gene transfer in bacteria has been
referred to as horizontal gene transfer, association of ribosomal RNA genes with
just lateral transfer might be misleading.

\textbf{Gene transfer between SPY and SDE:} Through virulence genes
bacteria can infect humans
with a variety of diseases: e.g., food poisoning by \emph{Eschericha
Coli}, strep throat by \emph{Streptococcus pyrogenes}, tuberculosis
by \emph{Mycobacterium tuberculosis}, cholera by \emph{Vibrio cholera},
anthrax by \emph{Bacillus anthracis}, pneumonia by \emph{Streptococcus
pneumoniae} and \emph{Chlamydophila pneumoniae}, and influenza by
\emph{Haemophilus influenzae}, to name a few. \emph{S.\ pyogenes}
(SPY) belonging to group A streptococci (GAS) primarily infect humans
in the throat and skin \citep{Broyles2009}. It has high prevalence
throughout the world causing diseases ranging from mild illnesses
such as pharyngitis to severe invasive streptococcal diseases including
necrotizing fasciitis and toxic shock syndrome \citep{Cunningham2000a}.
A closely related species called\emph{ S.\ dysgalactiae} ssp.\emph{\ 
equisimilis}
(SDE) belonging to group C or G streptococci (GCS/GGS) used to be
generally regarded as a veterinary pathogen \citep{Vandamme1996}.
Researchers have been confirming that GCS/GGS can cause serious and
life-threatening streptococcal diseases to humans traditionally associated
with GAS \citep[e.g.,][]{Brandt2009}. Because SDE can cohabit with
SPY at the same ecological niche of tissue sites, and both may exhibit
overlap in the disease profile, the close proximity
may allow the two species to interact with each other by exchanging
parts of their genomes. It has been suspected that opportunistic pathogenicity
of SDE may have been facilitated by the very genetic interaction between
the two species. 
Although virulence genes were suspected to be acquired via non-recombining gene
transfer rather than recombining transfer \citep{Groisman1994}, systematic
approaches to elucidating relative effects of the two types of gene transfer 
rarely have been applied. This unfortunate situation motivated us to study the
closely related bacteria as efforts of better understanding relative 
association of gene transfer with virulence genes.
We hoped that understanding their genetic exchanges
would enhance our knowledge of mechanisms of their pathogenicity on
humans.

\textbf{Previous studies on gene transfer and its directionality between
SPY and SDE:} 
Because of the close interaction between SPY and SDE, 
possible donor and recipient relationship of the two species were noted before.
However, previous researches in gene transfer between
the two pyogeneic species of SPY and SDE were limited to small portion
of the bacterial genome, mostly focusing on several genes 
\citep{Kalia2004,Towers2004,Bessen2005}.
For example, \citet{Sachse2002} showed gene transfer from SDE to
SPY for the downstream of the genomic region of streptococcal pyrogenic
exotoxins \emph{speG}. \citet{Davies2007} investigated phage 3396
from SDE as a vehicle of gene transfer from SDE to SPY. \citet{Davies2009}
demonstrated that an integrative conjugative element of SDE allowed
gene transfer from SDE to other streptococci. On the contrary, \citet{Kalia2001}
attempted to show that the direction of gene transfer from SPY to
SDE was more dominant than the other direction, albeit unfortunately
appeared to be premature. Because their researches on direction of
gene flow were limited to only small portion of regions in the genomes,
it remains to be elusive to determine the genome-wide degrees to which
genomic segments flow between the two species and directions of thereof.

\textbf{Methods of studying gene transfer in bacteria:} 
Studies on horizontal gene transfer in prokaryotes have focused on genes in very
large sets of species \citep{Koonin2001} using a great deal of genomes
available.  These studies gave us insights into complex evolutionary histories
of bacteria and the Archaea.  As sequencing costs plummet \citep{Mardis2011},
bacterial genome sequencing produces genomes of high quality even within
bacterial species \citep{Tettelin2009a}.  With more individual genomes per
species we now can go beyond the broad study on gene transfer towards
fine-scaled history between very closely related prokaryotes.  For example,
\citet{Liu2009} studied gene transfer events in \emph{Lactobacillus bulgaricus}
and \emph{Streptococcus thermophilus} using the compositional approach developed
by \citet{Karlin2001}. \citet{Luo2011} investigated gene transfer events in
\emph{Eschericha Coli} using the embedded quartet decomposition analysis
developed by \citet{Zhaxybayeva2006}. \citet{Hamady2006} combined phylogenetic
and compositional approaches to detecting horizontal gene transfer.
\citet{Caro-Quintero2011} studied a number of \emph{Shewanella baltica} genomes
to identify the genetic elements that enabled the species to adapt to redox
gradients. While compositional approaches were effective in analyzing
genome-scale data, phylognetic approaches have allowed to understand
evolutionary forces of gene exchange between species
\citep[e.g.,][]{Bapteste2004,Simonson2005,Beiko2009} by comparing differences
between gene trees and a reference species tree.  However, many of the methods
did not have explicit models of homologous recombining gene transfer that was
one of important evolutionary processes in horizontal gene transfer.  
% Overall,
% the development of methods for downstream analyses using bacterial genomes
% appears to lag behind that of high-throughput sequencing technology
% \citep[e.g.,][]{Green2011}.  
Recently, a method of detecting recombination 
with whole bacterial genomes \citep{Didelot2010} was developed. The explicit
model of homologous recombination allowed them to study 
flows of genetic segment between \emph{Bacillus cereus} species in the core
genomes that shared by most of individuals. 
In addition to recombination acting on bacterial genomes, genome rearragement,
translocation, duplication, loss, and more complex processes would govern the
evolution of bacteria. These processes, especially gene duplication, loss, and
non-recombining gene transfer, would result in more complex relationship between
even closely related bacteria. There have been progresses in 
methods of inferring gene duplication, loss, and transfer
events for gene families \citep{Doyon2011}. 
We noted that
the first approach could be used to infer transfer of genomic segments
due to homologous recombination, and the second could do similarly
due to non-recombining gene transfer.

%\textbf{Last paragraph before methods:} 
With these
methods for detecting gene transfer we interrogated gene flows
between SPY and SDE by measuring the relative impact of
two different processes of gene transfer on the pathogenicity of SPY
and SDE. This was possible thanks to recent release of bacterial
genomes of SPY and SDE. 
We here provided evidence that non-recombining
gene transfer was more effective for acquisitions of virulence genes
than homologous recombining gene transfer in SPY and SDE. 
While recombination rates varied along
the genomes, a few hot spots of recombination were observed. We found
directional gene flows in homologous recombining gene transfer process from SPY
to SDE. 
We showcased gene fragments of the inferred homologous gene transfers with high
posterior probabilities, mostly appearing to move from SPY to SDE.

\clearpage{}

\section{Methods and Materials}

\subsection{Data preparation}
Two genomes of \emph{S.\ pyogenes} (SPY), and two of \emph{S.\ dysgalactiae}
ssp.\emph{\ equisimilis} (SDE) were downloaded from NCBI with the following 
accessions:
NC\_004070 {Beres2002}, NC\_008024 \citep{Beres2006}, NC\_012891
\citep{Shimomura2011}, and CP002215 \citep{Suzuki2011}. We refer to these 
samples as SPY1, SPY2, SDE1, and SDE2, respectively. One genome
of \emph{S.\ dysgalactiae} 
ssp.\emph{\ dysgalactiae} was provided
by \citet{Suzuki2011} with accession identifiers of SddyATCC27957, referred to 
here as SDD. We also downloaded the genome (NC\_012471) of 
\emph{S.\ equi} ssp.\emph{\ equi} strain 4047 referred to as SEE.

\subsection{ClonalOrigin}
We present a short summary of the statistical framework developed by
\citet{Didelot2007} and \citet{Didelot2010} for studying bacterial 
recombination. A detailed account was found in the two papers and in the user 
guide available at \url{http://code.google.com/p/clonalorigin}. 
The analysis required first identifying sufficiently long contiguous blocks of 
high confident alignment. Multiple alignments with the six genomes
led to a dramatic reduction in the core 
genome length compared with those without SEE genome.
Because blocks of long multiple sequence alignments
were needed in order to obtain reliable 
estimates of recombination rate, we omitted the SEE sample 
from the ClonalOrigin analysis. We used progressiveMauve v2.3.1 
\citep{Darling2004,Darling2010} with default options
to align the five genomes.
We then used stripSubsetLCBs, a program of the ClonalOrigin package
to identify 276 blocks of sequence alignments with a length threshold of 
1500 base pairs.  We excluded two of these
blocks because they contained long consecutive gaps. Using these blocks, we ran 
ClonalFrame v1.1 available at
\url{http://www.xavierdidelot.xtreemhost.com} to obtain an 
estimate of the species tree relating the five genomes. ClonalFrame also 
outputs estimates of additional parameters, such as the average recombinant 
tract length and the ratio between recombination and mutation rates, however, 
these were not used in the subsequent analysis. We used a single Markov chain 
Monte Carlo (MCMC) with 10000 burn-in generations and 10000 sampling 
generations to obtain the tree estimate with branch lengths, 
and seven additional independent runs to validate adequate 
convergence of the first.

The next step of the analysis involved running ClonalOrigin in order to infer 
recombination histories across the alignment blocks. ClonalOrigin performs MCMC 
sampling of population parameters and recombination histories, given the 
species tree. The population parameters include mutation rate, 
recombination rate, and average tract length of a recombinant edge. A 
recombination history is represented by a \emph{recombinant tree}, which is a 
species tree augmented with a set of \emph{recombinant edges}. 
Although \citet{Didelot2010} described the concept of a recombinant tree
in detail, we elected to use Figure \ref{fig:clonalorigin} 
to illustrate it in order to clarify
summarization procedures we employed of posterior samples of recombinant trees.
Each recombinant 
edge is defined by time and branch of origin, time and branch of destination, 
and genomic interval in which it affects. The branch of origin is referred 
to as the \emph{donor} and the destination branch is referred to as the 
\emph{recipient}. A recombinant tree can be viewed as a restricted version of 
an ancestral recombination graph (ARG), and as such it induces a local 
genealogy for each site in the genomic block it describes. ClonalOrigin could
jointly sample population parameters and 
recombinant trees according to the posterior probabilities defined by a model 
approximating the coalescence process with recombination.
% \texttt{MAYBE ELABORATE MORE ON RESTRICTIONS OF CLONAL-ORIGIN'S MODEL? 
% SC:I HAVE NOT THOUGHT ABOUT MORE RESTRICTIONS ON THE RECOMBINANT TREES.}

% ClonalOrigin is applied in two stages. In both stages, sampling is performed 
% separately for the different alignment blocks. In the first stage, full 
% sampling is performed to obtain block-specific estimates of the model 
% parameters integrating out the possible recombinant trees. The block-specific 
% estimates of mutation rate, recombination rate, and average tract length are 
% then combined using a weighted median scheme to obtain a global estimate for 
% each parameter. In this scheme, each estimate is weighted according to the 
% length of the block with which it is associated, down-weighing the less 
% reliable estimates obtained from short blocks. In the second stage of 
% ClonalOrigin, the model parameters are fixed to their global estimate across 
% all blocks, and only the recombinant tree is sampled. Later this section we 
% describe several summaries of this posterior sample that are used in our 
% analysis. We used ClonalOrigin subversion r19 (downloaded from 
% \url{http://clonalorigin.googlecode.com/svn/trunk}). For the first stage, we 
% used two independent chains with 1,000,000 burn-in generations and  10,000,000 
% sampling generations, sub-sampled every 100,000 generations. The first chain 
% was used to obtain the block-specific estimates, and the second chain was used 
% to validate adequate convergence of the first. For the second stage, we used a 
% single chain with 10,000,000 burn-in generations and  100,000,000 sampling 
% generations, sub-sampled every 100,000 generations.

Two stages of MCMC with ClonalOrigin
were employed as described in \citet{Didelot2010}. 
Both stages of MCMC were performed independently for each of the 274 alignment blocks. 
We used ClonalOrigin subversion r19 that had been downloaded from 
\url{http://clonalorigin.googlecode.com/svn/trunk}. 
In the first MCMC we obtained estimates of 
mutation rate, recombination rate, and recombinant tract length. 
Two independent chains were executed with $10^6$ burn-in generations and $10^7$ 
sampling generations, sub-sampled every $10^5$-th generations. One of the two chains
was used to obtain the block-wise mean estimates, and the other chain was used 
to validate convergence of the first stage of MCMC. 
The estimators of the three population parameters were medians weighted over
block lengths, which down-weighed less reliable estimates obtained from shorter
alignment blocks. 
The second stage of MCMC with ClonalOrigin was executed with the three population
parameters fixed to the estimated values in the first stage. 
Two independent chains of the second stage
of ClonalOrigin's MCMC were also executed with a total generation being
$1.1\times10^8$. The first $10^7$ generations of each chain were discarded
as burn-in. The remaining $10^8$ generations were sub-sampled
every $10^5$-th generation.
We now describe the method of summarizing
the posterior sample to examine aspects of recombinant trees from the second
stage of ClonalOrigin's MCMC.

\citet{Didelot2010} presented two procedures of summarizing recombinant trees by
viewing trees from two different angles: one from recombinant trees by
aggregating all of the sites, and another from site-wise recombinant edge
changes to the species tree. By following their approaches
we quantified more aspects of recombinant trees: from the view of recombinant
trees with aggregated sites we counted gene tree topologies, and from the view of
site-wise recombinant edge changes we quantified recombination intensity along
the genome. 
We first describe procedures of summarizing recombinant trees from
the first viewpoint.
To summarize the posterior samples of ClonalOrigin we examined
the distribution of topologies of local genealogies induced by the sampled 
recombinant trees. Recall that a recombinant tree induces a local genealogy for 
each site along the analyzed block. Figure \ref{fig:clonalorigin} shows six
local gene trees including one that is the same as the species tree from the
most left-hand side of the block of 100 base pairs to the right.
We expected the dominant topology across 
sampled genealogies to be concordant with that of the species tree. However, 
other frequently sampled topologies were expected to be potentially informative about 
common recombination events. There are 105 possible 
rooted topologies relating five taxa, which allowed us to count topologies
induced by a posterior sample of recombinant trees.
Another useful summary suggested by 
\citet{Didelot2010} was to classify recombinant edges by types defined according 
to donor/recipient branch pairs, and record the number of edges sampled for
each recombinant edge type. For instance, our species tree has nine branches
including rooting branch and 
81 branch pairs, out of which 60 are possible donor/recipient pairs 
by prohibiting recombinant edges from going backward in time. The posterior 
expected number of edges of each type can be compared to the prior expected 
number determined by the three population parameters fixed in the second stage of 
ClonalOrigin. This comparison was displayed in a heat map of 
Figure \ref{fig:Heatmap-of-recombination}, which indicated branch pairs that 
might experience an excess or depletion of recombination events across all
blocks.

For the second viewpoint of recombinant trees
\citet{Didelot2010} suggested an approach used to trace
recombination intensity along the genome. 
For each edge type and each site, we recorded the sampling frequency of a 
recombinant edge of the type at that site (e.g., Figure S\ref{fig:muts}).
Figure \ref{fig:clonalorigin}B illustrates the posterior probability that a site
experiences influx from a donor branch using a single recombinant tree.
% This sampling frequency is an 
% estimate of the posterior probability taking into consideration the 
% sequence data and the setting chosen for the model parameters in the second 
% stage of ClonalOrigin. 
Summing these posteriors over a certain set of edge 
types, we could obtain a measure of recombination intensity along the genome 
(Figure \ref{fig:clonalorigin}B). 
This allowed us to find genomic regions enriched for recombination events in general, 
or regions enriched for recombination events of certain types (e.g.,
between SPY and SDE).  We note that the recombination intensity measure is not 
normalized to the interval $[0,1]$. Rather giving probabilistic meanings to the
recombination intensity we interpreted it as an average number of recombinant
edge types per site.
The measure appeared to act as an 
effective indicator of recombination intensity 
(see section of \textbf{Simulation Study}). 
Our simulation study indicated that edges of certain types were better indicators of 
true recombination events than others. For instance, recombinant edges which 
resulted in a recombinant tree with the same topology as the species tree appeared to 
be more prone to spurious sampling. 
Such edges were ones where the donor was the 
same branch, sister branch, or parent branch of the recipient branch. 
Therefore, in some cases we elected to remove such edges from consideration 
when measuring recombination intensity.
\texttt{IG: WE STILL NEED TO CHECK AND REFINE THIS IDEA. SC: WORKS ARE BEING DONE
FOR THIS ISSUE OF REMOVING UNINFORMATIVE RECOMBINANT EDGES}.

\subsection{Mowgli}
While ClonalOrigin could be used to understand recombining gene
transfers within the conserved core genome, we sought to understand
the history of the remainder of the genome where more complex
evolutionary processes occur, such as gene duplications, losses, and
horizontal transfers.  In recent years, several methods have been
developed for inferring these evolutionary events in gene families by
reconciling gene trees with a species tree
\citep{David2011,Doyon2011,Tofigh2011}.  In this analysis, we
complemented our study of recombination by using Mowgli
\citep{Doyon2011} to characterize events that can change the copy
number of a gene, such as duplications, losses, and non-recombining
horizontal gene transfers.  In contrast to ClonalOrigin, Mowgli
assumes a parsimony model for inferring events and it uses
pre-determined gene clusters where genes are treated as atomic units.
%The advantage of using Mowgli was that it did not assume the sequence
%data contains a single copy of each homologous segment, and it
%attempted to explain different copy numbers by a series of gain and
%loss events.  We here were interested in the approach of Mowgli to
%study non-recombining gene transfers.

We obtained gene family clusters defined in \citet{Suzuki2011} 
using the OrthoMCL method \citep{Li2003} over the full set of six genomes
including \emph{S.\ equi} subsp.\emph{\ equi} strain 4047 (SEE). 
For each family cluster,
the corresponding protein sequences were aligned using the MUSCLE
program \citep{Edgar2004a} and nucleotide alignments were constructed
by mapping nucleotide sequences onto each protein sequence alignment. We then
identified recombination within gene alignments using the Single
Breakpoint Recombination (SBP) method from the HyPhy package
\citep{KosakovskyPond2006}.  Using the Akaike Information Criterion (AIC), SBP
inferred that 47.5\% (1,094) of gene families contained at least one
topology-changing recombination. Each of these families was then
split into two gene fragments.
Gene trees were constructed from nucleotide alignments for each
gene fragment using the RAxML program \citep{Stamatakis2006}, and were
rooted by minimizing inferred duplications and losses.  Each rooted
gene tree was then reconciled using the Mowgli program
\citep{Doyon2011}, which inferred the most parsimonious gene
duplication, loss, and transfer events for each family.  In the case
of multiple maximum parsimonious reconciliations, one was chosen
uniformly at random.  Unless specified otherwise, default setings were used for
all programs.
Since the method of Mowgli did not directly model recombination events, 
recombination-based transfers were inferred in a post-processing step, by 
checking whether the transfer was coupled with a loss of a homologous gene in 
the recipient branch (indicating gene replacement rather than gain). 
A transfer event was classified as non-recombining if the 
destination species of that transfer had an extant descendant that contained
copies of that genes which were descendants of the transferred gene as well as 
copies which were non-descendants.
We noted that this parsimonious approach was imperfect and would tend to 
undercount events, especially recombining gene transfer events, which were
modeled using two events: a transfer and a loss. Nevertheless, this analysis 
allowed us to study non-recombining gene transfers, which were not modeled by 
ClonalOrigin, and to validate some of the ClonalOrigin results on recombining 
gene transfer.

\subsection{Functional categories association}
We performed a series of analyses to check for functional gene categories 
enriched for certain types of gene transfer events. Functional categories were 
assigned in the same manner as \citet{Suzuki2011}. In short, blastP was used to 
compare Streptococcus genes to bacterial proteins from the Uniref90 database.  
Matches with E-values less than $1.0\times10^{-5}$ were assumed to have the same Gene Ontology (GO) 
classification as the target gene, as determined from the uniProt GOA database. 
We also classified a category of  \emph{virulent genes} using the virulence 
profile described in \citet{Suzuki2011}. As virulent genes we used 207 loci, in 
which at least one of the genes in the gene family was identified by that 
profile.
Each gene covered by our analysis was associated with a recombination intensity 
score from the ClonalOrigin analysis and scores from the Mowgli 
analysis corresponding to number of duplications, losses, non-recombining 
gene transfers, and homologous recombining gene transfer events. 
\texttt{The recombination intensity scores were taken to be average site-wise intensity 
scores. SC: THIS NEEDS TO BE REANALYZED WITH RECOMBINATION INTENSITY WITH ONLY
TREE-TOPOLOGY CHANGING RECOMBINANT EDGES.} 
Mann-Whitney tests were used to identify GO categories 
whose genes showed significantly elevated values in each score. 
A 5\% False Discovery Rate (FDR) correction \citep{Benjamini1995} was used to 
determine an appropriate significance threshold for our p-values.

\subsection{Simulation Study}
We conducted experiments on simulation data to check the accuracy of 
estimates obtained by ClonalOrigin in a scenario similar to ours. We used a 
simulation feature of ClonalOrigin to simulate the process of mutation and 
recombination across the five species.
We started by testing the accuracy of the global estimates obtained from the 
first stage of ClonalOrigin for the three population parameters.
Ten recombinant trees were sampled given the tree parameters and the species
tree  set to the values inferred from the 
five bacterial genomes (see section \textbf{Results}).
Each of the recombinant trees was used to simulate
the Jukes-Cantor model for DNA sequence evolution to create 274 blocks of
alignments. 
The block lengths configuration was the same as that of the blocks 
used in the analysis of the five genomes.
We ran ClonalOrigin 
on the ten replicate data sets, and obtained global estimates by applying the 
weighted median scheme for the block-specific estimates. The means of the ten 
global estimates of mutation rate per site and recombination rate per site were 
0.081 with standard deviation (SD) of $0.0011$, and 0.011 with SD of $0.0003$, 
respectively, centering
around the true values: 0.081 and 0.012, respectively. The mean 
of the ten estimates of the recombinant tract length was 919 bp (SD 28), which 
was significantly higher than the true value of 744 bp. This was not entirely 
surprising as recombinant tract lengths were known to be difficult to 
estimate from relatively short alignment blocks \citep{Didelot2010}. When we 
performed similar simulations which much longer blocks (10000 bp), 
ClonalOrigin was able to estimate the average recombinant tract length 
accurately (data not shown).

% \texttt{IG: RESULTS FROM THIS POINT ON REQUIRE SOME REFINEMENT. SC: Let's see.}
Next we evaluated the accuracy of statistics estimated in 
the second stage of ClonalOrigin. We used 100 replicates of data sets obtained
by first
sampling ten recombinant trees given the mutation rate, recombination rate, and
tract length, and generating ten independent 
simulations of sequence evolution using each of the recombinant trees. For each measure 
of interest we thus obtained 100 separate estimates, which we could compare 
with either
the expected value from the ten recombinant trees, or 
true value from one of the recombinant trees. 
First, we tested the number of 
recombinant edges of each of the sixty different types. Figure S\ref{fig:h1} 
plots  the distribution of observed counts against the 
respective expected count given the model parameters and species tree 
for each recombinant edge type.
\texttt{FIX ``TRUE" TO BE EXPECTED VALUE RATHER THAN AVERAGE OVER THE 10 REC. 
TREES. SC: OKAY. LET'S FIND THE PRIOR EXPECTED NUMBER OF EVENTS. THIS MUST BE
GIVEN SOMEWHERE}. 
The estimated numbers appeared to be slightly over-estimated, but in 
very good correlation with the expected numbers.

\texttt{CONSIDER SHOWING DETAILED COMPARISON OF ESTIMATED VALUES TO ONES 
EXPECTED FROM THE ACTUAL REC. TREES. THESE DON'T LOOK AS NICE, BUT THEY MIGHT 
BE ABLE TO DEMONSTRATE OUR ABILITY TO DETECT DEPARTURE FROM EXPECTATION. SC:
THIS MAY BE ABOUT CHANGING AVERAGE TO PRIOR EXPECTED NUMBER OF EVENTS.}

\texttt{WE SHOULD PROBABLY CONSIDER ALTERNATIVE WAYS OF USING THESE RESULTS TO 
ARGUE THAT OUR DIRECTIONAL SIGNAL IS REAL. SC: I MIGHT CONSIDER SIMULATIONS AND
ARTIFICIALLY DOUBLE OR TRIPLE THE RECOMBINATION EVENTS OF ONE OF TYPES.}

With the slight over-estimation of recombination events we were concerned with 
any potential biases hidden in ClonalOrigin implementation. Because the 
simulated data were generated with no directionality between SPY and SDE,  if 
we estimate recombination events from the simulated data sets using 
ClonalOrigin and compare those estimates from those from the real data set, 
this could remove biases in ClonalOrigin implementation. We could use inferred 
estimates of recombination from the simulated data under the prior model of 
ClonalOrigin as the prior expected number of recombination. Figure 
S\ref{fig:h2} shows a plot of numbers of recombination events from the real 
data against those from the simulated data sets. Note that the numbers of 
recombination events from simulated data sets appear to deviate from the line 
of slope being 1 and intercept being 0. The numbers of recombination events for 
the recombinant edges from SPY to SDE increase by more than 2.5 fold whereas 
those for the recombinant edges from SDE to SPY increase by about 1.7 fold.  
This was not surprising because inference of number of recombinant edges was 
quite accurate. We considered that subtraction of effects of number of 
recombinant edges inferred from simulated data and subtraction of effects of 
the prior expected number of recombinant edges were effectively equivalent. 
However, this comparison between inferrence from simulated data and that of 
real data set suggested that the directional difference between SPY and SDE was 
due to neither lack of power of ClonalOrigin nor hidden bias thereof.

%\textbf{Simulation study for checking the surrogate measure of recombination intensity:} 
We tested if our measure of recombination intensity provided a reliable 
estimate of the true intensity. For this purpose, we artificially annotaded our 
synthetic blocks to genes according to the real annotations along the real 
bacterial alignments. For each gene in each of the 100 data sets, we plotted 
the average estimated recombination intensity along that gene against the true 
average intensity, i.e., the average number of different recombinant edge types 
implied by the recombinant tree used in simulation (Figure S\ref{fig:ri1}). 
Although the scatter plot did not strictly align along the $X=Y$ line, there 
was clear correlation of the estimated and true values (Pearson's correlation 
coefficient of 0.63). Estimated values appeared to be driven somewhat by the 
prior implied by the population parameter setting, explaining why lower values 
were over-estimated and higher values were under-estimated. Nonetheless, the 
correlation indicated that the measure of recombination intensity used in our 
functional gene category served as an adequate indicator of actual intensity of 
recombination events along genes.

\clearpage{}

\section{Results}

\subsection{ClonalOrigin analysis}
%\textbf{Genome sequence alignments and clonal frame:}
The species tree of the five genomes is shown in Figure S\ref{fig:tree5}, which
was estimated using 274 blocks of the core genome alignments with ClonalFrame. 
Figure S\ref{fig:blocksize} shows the distribution of block lengths. The average
length was 4215 bp, and total length of the blocks is 1155016 bp.  There was good 
agreement on the topology among independent chains, and reasonable agreement on 
branch lengths. The tree was also in good agreement with the one used by 
\citet{Suzuki2011}. ClonalFrame estimated the relative strength of recombination
compared to mutations to be $r/m = 4.66$, indicating 
that reconmination-driven substitutions affected more 
sites than did mutation-driven substitutions. The first stage of ClonalOrigin 
was then applied to obtain block-wise estimates of mutation rate, 
recombination rate, and average recombinant tract length 
(Figure \ref{fig:scatter3}).
In the first stage of ClonalOrigin MCMC a few alignment blocks were not
successfully finished even after a month of computation. 
Because those were not expected
to affect the global estimates of population parameters significantly, we used
only finished results of the first stage of ClonalOrigin to estimate mutation
rate, recombination rate, and recombinant tract length. 
The median of mutation
rate per site weighted over the lengths of blocks was estimated to
be 0.081 with (0.067,0.094) of interquartile range (IQR). Mutation rates
per site for many alignment blocks were almost centered around the
gobal median. The median of recombination rate per site was estimated
to be 0.012 with (0.006,0.019) of IQR.
The median value for the recombinant tract length was estimated
to be 744 base pairs with (346,2848) of IQR. 
The estimates of the three parameters were based on one of two independent
MCMC chains. 
The estimates from the second MCMC run were close to those of
the first run: mutation rate estimate of 0.081 with IQR of (0.067,0.099),
recombination rate of 0.012 with IQR of (0.006,0.020), and recombinant tract
length estimate of 723 with IQR of (348,2870).
We considered the similar 
results of the two independent chains as confirmation of 
the convergence of the first stage of ClonalOrigin MCMC.

% \texttt{THIS FIGURE OF RECOMBINATION BOUNDARIES NEEDS TO BE REPLACED OR REMOVED
% BECAUSE I DO NOT FIND REASONS OF ADDING THIS TO THE MANUSCRIPT.}
% For further validation of the 
% estimated recombnation rates, we recorded the number of recombinant edges 
% ClonalOrigin samples that are entirely contained within a block; 
% this number 
% should correlate well with the block-specific recombination rates. Figure 
% \ref{fig:recombination-boundaries} shows the numbers obtained in our MCMC runs, 
% and indeed these numbers appear to be well correlated with the block-specific 
% recombination rates.

%\textbf{Marginal evidence for unbalaned gene flow from SPY to SDE:}
We then proceeded with the second stage of ClonalOrigin, setting the population 
parameters to their global estimates. 
Using the posterior sample of recombinant trees, we counted 
rooted topologies of local genealogies along the blocks of alignments. 
The top eight 
topologies (out of 105 possible topologies) are listed in Table 
S\ref{tab:Gene-tree-topologies}. The species tree topology obtained a posterior 
probability of 66.7\%. The second most probable topology was one where SDD was 
the outgroup, and SPY and SDE were sister clades with probability of 9.3\%.
The other topologies had probability less than 4\%. We also recored the 
number of recombinant edges sampled by ClonalOrigin of each of the sixty 
possible types determined by donor/recepient species tree branches. The heat 
map in Figure \ref{fig:Heatmap-of-recombination} shows the log ratios of 
these numbers and their corresponding expected values given the species tree and 
the global parameter settings. 
% \texttt{IG: ADD FIGURE WITH NON-NORMALIZED VALUES? SC: Why?} 
The recombinant edge types that showed the highest enrichment in the 
sample were the ones from SPY1 and SPY2 branches to any branch within the SDE 
clade, with ratios in $[2.46,2.58]$. 
\texttt{IG: REPORT RATIOS TO SDE BRANCH? SC: List all of the values.} 
For comparison, the ratios for edges in the other directions were all smaller 
than 1.6. This appeared to indicate that gene flow driven by homologous 
recombination from SPY to SDE was stronger than gene flow in the reverse 
direction.

\subsection{Mowgli analysis}
%\textbf{Gene gain, loss, duplication, and non-recombining horizontal gene transfer:} 
For the Mowgli analysis we obtained 2314 gene families (Figure
S\ref{fig:famsize}), the largest of which consisted of 59 genes, while
nearly half of gene families (1066) had exactly one gene per species
(six genes total).  Figure \ref{fig:Gene-duplication-loss} shows the
inferred gene counts for each extant and ancestral species. Overall,
we inferred 304 duplications, 944 recombining gene transfers, and 294
non-recombining gene transfers, comprising 23.7\% of total transfer
events.  We found that duplications made up only a small fraction of
the events within the clade (10.2\%), while horizontal transfers and
losses were significantly more frequent (41.7\% and 48.0\%,
respectively). We found that duplications were most often
species-specific. Many gene losses likely represented horizontal
transfers that replaced the native gene sequence.  The species SDD and
SDE were the most expanded, both in terms of duplications and in-bound
transfers. Figure S\ref{fig:hgt-heatmap} shows the number of inferred
non-recombining gene transfer events between the different branches of
the species tree.
%\textbf{No evidence for unbalanced non-recombining gene transfer between SPY and SDE:}
SDE1 and SDE2 shared the most events (26 and 34) compared to other sample pairs. 
SDD appeared to acquire many genes from the ancestral SDE branch. Transfer 
events between the SPY and SDE clades were inferred in smaller numbers (<6) 
compared to those within each clade (23--34). Among the clade-crossing 
transfers, we did not see any significant enrichment for a specific direction, 
as opposed to the result for recombining transfers from the ClonalOrigin 
analysis (Figure S\ref{fig:mowgli-recomb-heatmap}).

\subsection{Functional categories associated with gene transfer}

\texttt{IG: THIS SECTION REQUIRES ADDITIONAL WORK, PENDING INTERPRETATION AND 
REFINEMENT OF RESULTS. }

\texttt{IG: THESE RESULTS MIGHT SLIGHTLY CHANGE BY REMOVING CERTAIN EDGE TYPES 
FROM CONSIDERATION. SC: ANALYSES WILL BE DONE BY REMOVING SOME EDGES.}

% \textbf{Functional categories associated with homologous recombination gene transfer:} 
% Having established relative correlation of the surrogate measure of recombination
% intensity in the simulation (Figure S\ref{fig:ri1}) we attempted to find
% functional categories that were associated with recombination.
% The nucleotide substitutions
% due to recombination events might not be selectively neutral because
% of large population of the bacteria, and accordingly natural selection
% could be more stronger than random genetic drift. Reflecting on the
% report of \citet{Lefebure2007} that genes that were likely to be
% under positive selection tended to be under higher recombination,
% we would expect that functional categories would be associated with higher
% recombination, and consequently homologous recombining gene transfer
% between species.

Table \ref{tab:functional} lists the functional categories that showed
significant associated with recombination intensity obtained from our 
ClonalOrigin analysis.
Several functional categories are related with ribosomes.
There was a functional category related with phosphotransferase systems
and ATP-binding cassette (ABC) transporters, i.e., carbohydrate transport
(GO:0008643). Translation (GO:0006412) included aminoacyl tRNA synthetase.
There were fatty acid biosynthetic process (GO:0006633) and lipid
biosynthetic process (GO:0008610). We did not find any significant association 
between recombination intensity and virulence genes. When measuring 
recombination only between the SDE and SPY clades, a smaller number of 
functional categories appeared to be significantly enriched, leaving the two 
biosynthetic processes out. For the direction from SPY to SDE only transport 
related
functional categories appeared to be significant, leaving out ribosome related 
functional categories and translation.

%\textbf{Functional categories associated with non-recombining gene transfer:}

Table \ref{tab:go-events} lists the functional gene categories that
showed significant associated with gene duplications, losses,
non-recombining and recombining gene transfer according to the Mowgli
analysis.  For the recombining transfers, Mowgli shows enrichment in
many of the same categories as ClonalOrigin, such as translation,
ribosome, and rRNA binding.  This agreement between the two different
methods, applied on somewhat different parts of the data, serves as
validation for the signal picked up by these separate association
studies.  However, for duplication, losses, and non-recombining
transfers, we find enrichment in very different categories related to
transposition, integration, and recombination.  Interestingly, gene
duplication and non-recombining transfers show enrichment in the same
categories, perhaps because both mechanisms increase gene copy number.
Gene losses also showed enrichment for transposition and integration,
but less significantly ($P<10^-6$) and also showed enrichment for
several additional categories such as transporter activity and
transcription initiation.  Unlike the associations with recombination
intensity, families with high numbers of duplications and
non-recombining gene transfer were significantly enriched for
virulence ($P<7.13 \times 10^{-3}$ and $P<1.33 \times 10^{-11}$,
respectively), whereas families with high losses and recombination
showed no enrichment ($P>0.54$ and $P>0.90$).  This might indicate
that non-recombining gene transfer served as a more effective means by
which SDE or SPY acquired virulence genes, compared to recombining
gene transfer.

\texttt{NEXT PARAGRAPH CONTAINS RESULTS ON INFERENCE OF SPECIFIC TRANSFER 
EVENTS. WE SHOULD DECIDE HOW TO RELATE THIS TO OUR OTHER RESULTS.}

%\textbf{Some detailed gene-level analyses:}
Having found these functional categories, we elected to search for specific 
genes that showed a clear and specific local signal of recombination.  We scanned 
all 274 alignment blocks to locate parts of genes for which a recombinant edge 
of a certain type was sampled with probability at least 0.9. 
% We focused on specific types of recombinant edges 
% \texttt{IG: ADD A BIT HERE. SC: I DO NOT SEE WHAT IG MEANT.}
We found a total of 58 genes showing high-confidence local 
recombination signal (Table S\ref{tab:genes-transfer}). The process of gene 
transfer does not necessarily move whole genes from one species to another 
\citep[e.g.,][]{Chan2009}, and we observed that units of 
recombining gene transfer do 
not correspond to single genes in their entirety. In some cases, only parts of 
a gene were inferred to have been transferred, for instance, transfer from SPY1 
to SDE1 was inferred for 29\% of the aspartyl-tRNA synthetase gene (see also 
Figure S\ref{fig:muts}). In other cases, the inferred transfer units cover sets 
of as much as eight adjacent genes.

%Not only parts of genes could be transferred via recombination, but a group of 
%genes could also be moved together. Of the genes listed in Table 
%S\ref{tab:genes-transfer} there were 12 groups of genes that were consecutive. 
%While transfer fragments of genes often covered two genes, there were transfer 
%fragments that covered eight genes together. Loci from SpyM3\_1512 to 
%SpyM3\_1525 were inferred to be transferred from SDE2 to SDD.

% \texttt{IG: NEXT PARAGRAPH CONTAINS SOME DISCUSSION ON THESE RESULTS. WE SHOULD 
% GET BACK TO THIS ONCE WE DECIDE WHAT TO DO WITH THE RESULTS}

\clearpage{}

\section{Discussions}
Closely related bacterial species were investigated here in our study to better
understand gene transfer processes. Two complementary methods were employed to
study recombining and non-recombining gene transfers. By discriminating the two
processes in analyzing the genomes, we provided evidence that the pathogenicity
of SDE and SPY was more attributed to non-recombining gene transfer process than
homologous recombining gene transfer.  Direction of gene flow between SDE and
SPY has been questioned and challenged
\citep{Kalia2001,Kalia2004,Towers2004,Bessen2005,Davies2007,Bessen2010} in order
to elucidate the origin of opportunistic human pathogenicity of SDE via transfer
from SPY into SDE.  We also showed in this study that the amount of recombining
gene flows from SPY to SDE was relatively greater than that of the reverse
direction using the core genome sequences.

\textbf{Virulence genes and gene transfer:} 
Because we found directional recombining gene transfer from SPY to SDE, and
did not found virulence genes association with recombining gene transfer, 
this indicated that 
pathogenicity of SDE might not be due to acquisition of virulence genes from
SPY via recombining transfer. This does not necessarily mean that virulence
genes were not exchanged between SPY and SDE via non-recombining transfer. 
Because the Mowgli analysis just did not reveal unbalanced non-recombining or
recombinding gene transfer, we suspected that
Mowgli's parsimony might not be powerful in detecting unbalanced gene flow.
We conjecture the following hypothesis of the evolution of SPY and SDE to
support the non-recombining virulence gene exchange between SPY and SDE.
Pathogenicity of SDE might have existed before the common ancestor of SPY
and SDE.  We speculated that SPY might have been split off from the common
ancestor that was SDE-like because the population size of SDE was larger than
that of SPY (CITE). After the split, SPY might have been specialized in
inhabting human hosts, and SDE in colonizing veterinary hosts. Recent surge in
cases of human diseases attributed to SDE might be because of secondary
contact of SPY and SDE since their split from their common ancestor.  When SPY
thrives in its environment of human hosts and another relatively close species,
SDE, starts to cohabit with the indigenous species SPY, genetic elements in one
of the two species could benefit another species. If the two species do not
compete with each other, then the increased genetic variation from multiple
species would eventually bring advantage to both of the species. 
Virulence genes could have been beneficial to both species in inhabting human hosts. 
Because the common ancestry of SPY and SDE, recombining gene transfer between
the two species might be relatively prevalent. But, these homologous
recombination might have rarely contributed to influx of virulence genes from
SPY to the opportunistic human pathogen of SDE. Because homologous parts
of SPY and SDE would be similar in their genetic compositions, the homologous
parts would be unlikely to contain virulence genes to human hosts. We suspected
that virulence genes were exchanged between SPY and SDE, but did not find any
directionality of the gene flow.

\textbf{Revisit to previous studies:} 
Reflection of previous studies in gene transfer in SPY and SDE would suggest
that using relatively small number of genomes we could detect homologous
recombining gene transfer that had been possible using many samples for
a few genes. We revisited the problem of gene transfer in other previous
studies.  We examined
the seven genes that \citet{Kalia2001} had studied for directionality
of gene transfer between SPY and SDE. Internal fragments of the seven
housekeeping genes, encoding putative glucose kinase (\emph{gki}),
glutamine transport protein (\emph{gtr}), glutamate racemase (\emph{murI}),
mismatch repair enzyme (\emph{mutS}), transketolase (\emph{recP}),
xanthine phosphoribosyltransferase (\emph{xpt}), and acetylcoenzyme
A acetyltransferase (\emph{yqiL}) were located in the SPY1 genome that we
investigated in this study.  Six of the seven genes except \emph{yqiL
}gene were in the core genomes ClonalOrigin analyzed. Genes \emph{mutS},
\emph{recP}, and \emph{xpt} had fragments of gene transfer with moderate
posterior probability of recombination. The other genes did not have
noticeable recombination. Except for \emph{mutS} where \citet{Kalia2001}
did not find recombination, the result was consistent with that of
\citet{Kalia2001}. 
We also revisited recombination in gene \emph{parC} in SPY1 genome
studied by \citet{Pinho2010}, where they studied fluoroquinolone
resistance in SDE and reported recombination in the gene. We also
observed moderate recombination in the gene.

\textbf{Mechanism of unbalanced gene flow between SPY and SDE:} 
If there were unbalaned gene transfer between the two
species, what were the barriers \citep{Thomas2005} to gene transfer
in the bacteria? Although barriers to bacterial recombination would
include geographic isolation, genetic isolation could prevent recombination
between coinhabiting bacteria. \citet{Kalia2001} discussed possible
mechanisms of the unbalanced gene flow by mentioning potential roles
of restriction-modification system (RMS) in the unbalanced gene flow.
Restriction-modification system is a biological self-defense mechanism
of bacteria, which is diverse in prokaryote world. Biological roles
of RMS might be to maintain and control species identity \citep{Jeltsch2003}.
Recently, \citet{Budroni2011a} studied that \emph{Neisseria meningitidis}
phylogenetic clades were associated with RMS, claiming that RMS modulated
homologous recombination in the bacteria of \emph{N.\ meningitidis}.
Most of \emph{S.\ pyogenes} genomes contain restriction-modification
type I system as well as type II system. The two genomes of SDE contained
only type II system. We conjecture that more tight self-defense system
of SPY may have made the species less capable of aquiring foreign
genetic material from SDE, and accordingly relatively unbalanced gene
flow between SPY and SDE may have been associated with different 
restriction-modification systems in the two species.

\textbf{Two functional categories and recombining gene transfer:}
Among the genes shown in Table S\ref{tab:functional} two classes of proteins
attracted our attentions: 
aminoacyl tRNA synthetases (aaRS) and phosphotransferase system (PTS).
Aminoacyl tRNA synthetases
charge tRNA with an amino acid so that amino acids are concatenated
from charged tRNA onto a growing peptides through translation. The
aaRS's are characterized by an amino acid that they charge tRNA with.
This gene family had been a target for horizontal gene transfer 
\citep[e.g.,][]{Lamour1994,Woese2000,Koonin2001},
which lent itself to an example of gene transfer in bacteria.
There were also recent reports of gene transfer in tRNA aminoacylation
\citep{Andam2010,Wang2011}. 
In our analysis, some of them appeared to be transferred
from SPY to SDE:
leucyl-tRNA synthetase, aspartyl-tRNA synthetase, and arginyl-tRNA
synthetase. Half of phenylalanyl-tRNA synthetase gene appeared to
be transferred in the reverse direction.

Bacterial phosphotransferase
system (PTS), also known as phosphoenolpyruvate (PEP) group translocation,
transports sugars such as glucose into the cell where phosphorylation
of the sugar happens. There have been studies of gene transfer of
PTS in bacteria \citep[e.g.,][]{Barabote2005,Zuniga2005}. As the
main source of energy for bacteria, sugars in the bacterial niches
must have been acting as important selective pressure. It is not surprising
that PTS sugars differ among bacterial groups because bacterial species
are correlated with their environment and energy sources theirof.
Table S\ref{tab:functional} shows that 
two PTS genes appeared to be transferred from SPY to SDE:
{}``PTS system, mannose-specific IIC component'' and {}``PTS system,
fructose-specific IIABC component.'' Whereas almost half of the former
gene was inferred to be affected by recombinant edges, only 3\% of the latter gene appeared
to be affected.

\textbf{Future models of bacterial evolution to handle recombining and
non-recombining gene transfer:}
The method of detecting homologous recombination gene transfer, or
the ClonalOrigin, is model-based approach as opposed to the other
method for non-recombining gene transfer, or the Mowgli.
The core genomes consisted of relatively long (i.e., 1500 base pairs
long) sequence alignments with almost no gaps. With this inherent
filtering of the input data preparation ClonalOrigin did not account
for gene gain and loss in its core genome analysis. On the contrary,
Mowgli used the pan genomes as well as the core genomes.  Considering
nucleotides as the level of gene duplication, loss, and transfer towards
the efforts of having more complete models for recombining and non-recombining
gene transfer
could be difficult because genome-scale application of those models
based on DNA or protein sequences might not be feasible.
The research into units
of gene transfer is underway \citep[e.g.,][]{Chan2009a}. Discriminating
the mode of gene transfer and using the units of each mode of gene
transfer towards the efforts of developing model-based approaches
to detecting gene transfer might be plausible. In doing so, more complicated
evolutionary models could be considered: genome rearrangement,
gene segment gain and loss, and gene duplications.

\section*{Supplementary Material}

Supplementary figures S1\textendash{}S8, and tables S1 and S2
are available at Genome Biology
and Evolution online (\url{http://www.gbe.oxfordjournals.org/}).

\section*{Acknowledgments}

We are indebted to Xavier Didelot for his help in setting up ClonalOrigin
analysis.  We are thankful to Haruo Suzuki for his help
in using the data of SDD genome. This work was supported by the National
Institute of Allergy and Infectious Disease, US National Institutes
of Health, under grant number AI073368-01A2 awarded to A.S. and M.J.S.

\bibliographystyle{gberefs}
\bibliography{siepel-strep}

\clearpage{}

\section*{Tables}

\clearpage{}

%
\begin{table}
\caption{\label{tab:functional}Functional categories asscoiated with higher
recombination.}
\begin{tabular}{|c|c|c|}
\hline 
p-value & count & description\tabularnewline
\hline
\hline 
\multicolumn{3}{|c|}{\# Total number of recombinant edge types}\tabularnewline
\hline 
2.994e-07 &  28 & rRNA binding\tabularnewline
\hline 
1.447e-06 &  42 & structural constituent of ribosome\tabularnewline
\hline 
1.694e-05 &  39 & ribonucleoprotein complex\tabularnewline
\hline 
2.048e-05 &  23 & fatty acid biosynthetic process\tabularnewline
\hline 
3.664e-05 &  72 & ribosome\tabularnewline
\hline 
1.018e-04 &  86 & translation\tabularnewline
\hline 
1.733e-04 &  22 & lipid biosynthetic process\tabularnewline
\hline 
2.792e-04 & 108 & RNA binding\tabularnewline
\hline 
4.164e-04 &  47 & carbohydrate transport\tabularnewline
\hline 
1.303e-03 &  38 & phosphoenolpyruvate-dependent sugar phosphotransferase 
system\tabularnewline
\hline 
1.345e-03 &  10 & pentose-phosphate shunt\tabularnewline
\hline 
1.819e-03 &  35 & protein-N(PI)-phosphohistidine-sugar phosphotransferase 
activity\tabularnewline
\hline 
\multicolumn{3}{|c|}{\# SDE->SPY recombinant edge types}\tabularnewline
\hline 
0.0001236 & 28 & rRNA binding\tabularnewline
\hline 
0.0001284 & 86 & translation\tabularnewline
\hline 
0.0001711 & 39 & ribonucleoprotein complex\tabularnewline
\hline 
0.0002230 & 10 & pentose-phosphate shunt\tabularnewline
\hline 
0.0003412 & 38 & phosphoenolpyruvate-dependent sugar phosphotransferase 
system\tabularnewline
\hline 
0.0004498 & 42 & structural constituent of ribosome\tabularnewline
\hline 
0.0007789 & 47 & carbohydrate transport\tabularnewline
\hline 
0.0008152 & 35 & protein-N(PI)-phosphohistidine-sugar phosphotransferase 
activity\tabularnewline
\hline 
\multicolumn{3}{|c|}{\# SPY->SDE recombinant edge types}\tabularnewline
\hline 
2.025e-06 & 47 & carbohydrate transport\tabularnewline
\hline 
1.875e-04 & 38 & phosphoenolpyruvate-dependent sugar phosphotransferase 
system\tabularnewline
\hline 
2.843e-04 & 18 & sodium ion transport\tabularnewline
\hline 
3.248e-04 & 35 & protein-N(PI)-phosphohistidine-sugar phosphotransferase 
activity\tabularnewline
\hline
\end{tabular}

\end{table}

\clearpage{}%

% Table 2. 
\begin{table}
\caption{\label{tab:go-events}
GO-term enrichment for different evolutionary events. }

\begin{tabular}{rrll}
P-value & Count & GO-term & Description \\
\hline
\multicolumn{4}{c}{Duplications} \\
\hline
4e-81 &  67 & GO:0006313 & transposition, DNA-mediated\\
5e-80 &  59 & GO:0004803 & transposase activity\\
6e-70 &  23 & GO:0032196 & transposition\\
2e-58 &  52 & GO:0015074 & DNA integration\\
8e-25 &  86 & GO:0006310 & DNA recombination\\
1e-20 &  36 & GO:0016987 & sigma factor activity\\
1e-20 &  36 & GO:0006352 & transcription initiation\\
2e-19 & 197 & GO:0003676 & nucleic acid binding\\
2e-10 & 527 & GO:0003677 & DNA binding\\
4e-06 & 142 & GO:0043565 & sequence-specific DNA binding\\
\hline
\hline
\multicolumn{4}{c}{Losses} \\
\hline
1e-06 &  67 & GO:0006313 & transposition, DNA-mediated\\
2e-05 &  59 & GO:0004803 & transposase activity\\
5e-05 &  52 & GO:0015074 & DNA integration\\
5e-04 &  23 & GO:0032196 & transposition\\
5e-04 &  10 & GO:0046873 & metal ion transmembrane transporter activity\\
7e-04 &  36 & GO:0016987 & sigma factor activity\\
7e-04 &  36 & GO:0006352 & transcription initiation\\
1e-03 &  31 & GO:0006814 & sodium ion transport\\
3e-03 &  86 & GO:0006310 & DNA recombination\\
6e-03 & 197 & GO:0003676 & nucleic acid binding\\
\hline
\hline
\multicolumn{4}{c}{Non-recombining transfers} \\
\hline
7e-61 &  67 & GO:0006313 & transposition, DNA-mediated\\
3e-58 &  59 & GO:0004803 & transposase activity\\
1e-30 &  52 & GO:0015074 & DNA integration\\
1e-28 &  23 & GO:0032196 & transposition\\
1e-13 & 527 & GO:0003677 & DNA binding\\
2e-12 &  86 & GO:0006310 & DNA recombination\\
1e-11 & 197 & GO:0003676 & nucleic acid binding\\
3e-09 &  36 & GO:0016987 & sigma factor activity\\
3e-09 &  36 & GO:0006352 & transcription initiation\\
3e-08 & 142 & GO:0043565 & sequence-specific DNA binding\\
\hline
\hline
\multicolumn{4}{c}{Recombining transfers} \\
\hline
7e-09 & 153 & GO:0006412 & translation\\
1e-08 &  67 & GO:0030529 & ribonucleoprotein complex\\
1e-07 &  45 & GO:0019843 & rRNA binding\\
2e-06 &  73 & GO:0003735 & structural constituent of ribosome\\
3e-05 & 157 & GO:0005840 & ribosome\\
7e-05 & 221 & GO:0003723 & RNA binding\\
2e-04 &  11 & GO:0004826 & phenylalanine-tRNA ligase activity\\
3e-04 &  14 & GO:0015413 & nickel-transporting ATPase activity\\
4e-04 &  93 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
5e-04 & 104 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
\end{tabular}

\end{table}

\clearpage{}

\section*{Figures\clearpage{}}

% Figure 1. A Recombinant Tree of ClonalOrigin Model
\begin{figure}
% \includegraphics[scale=0.42]{figures/mauve-analysis-002}
\includegraphics[scale=0.42]{figures/figure1}
\caption{\label{fig:clonalorigin}Recombinant tree in the ClonalOrigin model.
(A) A recombinant tree relating the five genomes with three recombinant edges.
(B) A genomic region of size 100 base pairs with subregions affected by the
three recombinant edges. Genomic positions are evenly marked by ticks every 
10-th base pair. Local gene trees induced by the recombinant tree are
illustrated for each region covered by some combinations of recombinant edges
along the genomic region (C-G). For example, the local gene tree (E) is created
by all of the three recombinant edges in the subregion between 51 base pairs
and 70 base pairs.}
\end{figure}
\clearpage{}

% Figure 2. Heat maps.
\begin{figure}
\includegraphics[scale=0.45]{figures/figure2}
\caption{\label{fig:Heatmap-of-recombination}Heatmap of recombination events.
Heat map representing the logarithm base 2 of the number of recombination
events inferred relative to its expectation under the prior, for each
donor/recipient pair of branches. The donor is given by the y-axis;
the recipient on the x-axis.  The
green cells were the ones for which the ratios were meaningless because
the recipient species tree branch of the cells was older than the
donor species branch.}
\end{figure}
\clearpage{}%

% Figure 3. The three main parameters.
\begin{figure}
\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-theta}}

\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-rho}}

\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-delta}}
\caption{\label{fig:scatter3}Scatter plots of mutation rate, recombination
rate, and average tract length for the 270 blocks. The red dashed
line is the global median of all the blocks.}
\end{figure}
\clearpage{}%

% Figure 4. Gene duplication, loss, and transfer
\begin{figure}
\includegraphics[width=6in]{figures/strep-events}
\caption{\label{fig:Gene-duplication-loss} Distribution of inferred
  gene duplication, loss, and transfer events across the phylogeny.
  Gene counts for each extant and ancestral species are given at each
  node of the phylogeny.  For each branch we denote the number of gains
  G*, duplications D*, losses L*, non-recombining transfers T*, and 
  recombinations R*.  The non-recombining transfer and recombination counts
  are indicated on the recipient branch (donors are not indicated).}
\end{figure}
\clearpage{}%

% Figure 5.
%=============================================================================
% FIGURE - Transfer heatmap
\begin{figure}
\begin{center}
\includegraphics[width=5in]{figures/trans-hgt-heatmap}
\end{center}
\vspace{-1.2in}
\caption{Distribution of horizontal gene transfers.}
\label{fig:hgt-heatmap}
\end{figure}
%=============================================================================

\clearpage{}\setcounter{figure}{0}
\setcounter{table}{0}
\renewcommand{\figurename}{Supplementary Figure}
\renewcommand{\tablename}{Supplementary Table}

\section*{Supplementary Tables}
\clearpage{}

% Table S1. Gene Tree topologies
\begin{table}
\caption{\label{tab:Gene-tree-topologies}Posterior probability of gene tree
topologies.}
\begin{tabular}{|c|c|}
\hline 
Gene tree topologies & Posterior probability\tabularnewline
\hline
\hline 
(((SDE1,SDE2),SDD),(SPY1,SPY2)) & 66.7\%\tabularnewline
\hline 
(((SDE1,SDE2),(SPY1,SPY2)),SDD) & 9.3\%\tabularnewline
\hline 
((SDE1,SDE2),(SDD,(SPY1,SPY2))) & 3.5\%\tabularnewline
\hline 
(((SDE1,(SDE2,SDD)),(SPY1,SPY2)) & 1.5\%\tabularnewline
\hline 
(((SDE1,SDD),SDE2),(SPY1,SPY2)) & 1.3\%\tabularnewline
\hline 
((SDE1,SDD),(SDE2,(SPY1,SPY2))) & 1.2\%\tabularnewline
\hline 
((((SDE1,SDE2),SPY1),SPY2),SDD) & 1.0\%\tabularnewline
\hline 
((SDE2,SDD),(SDE1,(SPY1,SPY2))) & 1.0\%\tabularnewline
\hline
\end{tabular}
\end{table}
\clearpage{}

% Table S2. 
\begin{table}
\caption{\label{tab:genes-transfer}Genes and their parts transferred via 
homologous
recombining gene transfer. }
{\footnotesize
\begin{tabular}{|lllllll|}
\hline 
Locus  & N$^a$  & S$^b$  & D$^c$ & F$^d$(\%)  & C$^e$(\%)  & gene 
product\tabularnewline
\hline 
SpyM3\_0074  &  & 4  & 0  & 20  & 100  & putative penicillin-binding protein 1b 
\tabularnewline
SpyM3\_0134  &  & 6  & 0  & 43, 0 & 79  & leucyl-tRNA synthetase \tabularnewline
SpyM3\_0156  &  & 0  & 4  & 77  & 100  & glucose-6-phosphate isomerase 
\tabularnewline
SpyM3\_0186  &  & 5  & 4  & 83  & 94  & putative glucose kinase \tabularnewline
SpyM3\_0192  & 1 & 6  & 5  & 14, 6 & 100  & ribulose-phosphate 3-epimerase 
\tabularnewline
SpyM3\_0193  & 1 & 6  & 5  & 99  & 100  & hypothetical protein\tabularnewline
SpyM3\_0197  &  & 6  & 5  & 7  & 100  & putative surface exclusion 
protein\tabularnewline
SpyM3\_0204  &  & 6  & 5  & 45  & 100  & hypothetical protein \tabularnewline
SpyM3\_0215  &  & 6  & 5  & 5, 26 & 100  & oligopeptide permease \tabularnewline
SpyM3\_0238  &  & 5  & 3  & 18  & 100  & hypothetical protein \tabularnewline
SpyM3\_0292  &  & 6  & 5  & 27  & 100  & putative tetrapyrrole methylase family 
protein \tabularnewline
SpyM3\_0429  &  & 2  & 1  & 0,22  & 100  & putative oligoendopeptidase 
F\tabularnewline
SpyM3\_0465  & 2 & 6  & 5  & 95  & 99  & putative dipeptidase \tabularnewline
SpyM3\_0466  & 2 & 6  & 5  & 16  & 100  & putative adhesion protein 
\tabularnewline
SpyM3\_0506  &  & 5  & 3  & 52  & 100  & phenylalanyl-tRNA synthetase subunit 
alpha \tabularnewline
SpyM3\_0580  & 3 & 3  & 0  & 3  & 100  & PTS system, fructose-specific IIABC 
component\tabularnewline
SpyM3\_0581  & 3 & 3  & 0  & 18  & 100  & putative peptidoglycan hydrolase 
\tabularnewline
SpyM3\_0581  &  & 6  & 5  & 11  & 100  & putative peptidoglycan hydrolase 
\tabularnewline
SpyM3\_0745  & 4 & 6  & 1  & 14  & 100  & putative two-component sensor 
histidine kinase \tabularnewline
SpyM3\_0746  & 4 & 6  & 1  & 55  & 100  & putative two-component response 
regulator \tabularnewline
SpyM3\_0785  & 5 & 4  & 1  & 78  & 92  & inorganic polyphosphate/ATP-NAD 
kinase\tabularnewline
SpyM3\_0786  & 5 & 4  & 1  & 100  & 100  & putative ribosomal large subunit 
pseudouridine synthase \tabularnewline
SpyM3\_0787  & 5 & 4  & 1  & 12  & 100  & phosphotransacetylase \tabularnewline
SpyM3\_1078  & 6 & 4  & 1  & 21  & 100  & putative cation/potassium uptake 
protein \tabularnewline
SpyM3\_1079  & 6 & 4  & 1  & 20  & 100  & putative ATP-dependent RNA helicase 
\tabularnewline
SpyM3\_1093  &  & 4  & 0  & 30  & 100  & putative heavy 
metal/cadmium-transporting ATPase \tabularnewline
SpyM3\_1151  &  & 3  & 1  & 34  & 100  & putative DNA repair and genetic 
recombination protein \tabularnewline
SpyM3\_1153  & 7 & 4  & 0  & 57  & 100  & putative hemolysin\tabularnewline
SpyM3\_1154  & 7 & 4  & 0  & 4  & 100  & putative geranyltranstransferase 
\tabularnewline
SpyM3\_1201  &  & 6  & 5  & 21  & 100  & putative two-component sensor 
histidine kinase \tabularnewline
SpyM3\_1267  & 8 & 6  & 5  & 7  & 100  & bifunctional methionine sulfoxide 
reductase A/B protein \tabularnewline
SpyM3\_1268  & 8 & 6  & 5  & 8  & 100  & hypothetical protein \tabularnewline
SpyM3\_1365  &  & 3  & 0  & 28  & 100  & putative Cof family 
protein/peptidyl-prolyl cis-trans isomerase, cyclophilin
type\tabularnewline
SpyM3\_1380  & 9 & 6  & 1  & 98  & 100  & putative Acetyl-CoA:acetoacetyl-CoA 
transferase b subunit \tabularnewline
SpyM3\_1381  & 9 & 6  & 1  & 100  & 100  & putative oxidoreductase 
\tabularnewline
SpyM3\_1382  & 9 & 6  & 1  & 23  & 38  & 3-hydroxybutyrate dehydrogenase 
\tabularnewline
SpyM3\_1512  &  & 4  & 0  & 59  & 100  & PTS system, mannose-specific IIC 
component \tabularnewline
SpyM3\_1518  & 10 & 1  & 2  & 63  & 100  & acetyl-CoA carboxylase subunit 
beta\tabularnewline
SpyM3\_1519  & 10 & 1  & 2  & 100  & 100  & acetyl-CoA carboxylase biotin 
carboxylase subunit \tabularnewline
SpyM3\_1520  & 10 & 1  & 2  & 100  & 100  & (3R)-hydroxymyristoyl-ACP 
dehydratase \tabularnewline
SpyM3\_1521  & 10 & 1  & 2  & 100  & 100  & acetyl-CoA carboxylase biotin 
carboxyl carrier protein subunit\tabularnewline
SpyM3\_1522  & 10 & 1  & 2  & 100  & 100  & 3-oxoacyl-(acyl carrier protein) 
synthase II \tabularnewline
SpyM3\_1523  & 10 & 1  & 2  & 100  & 100  & 3-ketoacyl-(acyl-carrier-protein) 
reductase \tabularnewline
SpyM3\_1524  & 10 & 1  & 2  & 100  & 100  & acyl-carrier-protein 
S-malonyltransferase\tabularnewline
SpyM3\_1525  & 10 & 1  & 2  & 100  & 100  & putative trans-2-enoyl-ACP 
reductase II \tabularnewline
SpyM3\_1561  &  & 6  & 5  & 11  & 97  & hypothetical protein \tabularnewline
SpyM3\_1577  &  & 4  & 1  & 11  & 100  & excinuclease ABC subunit A 
\tabularnewline
SpyM3\_1579  &  & 4  & 1  & 1  & 100  & hypothetical protein \tabularnewline
SpyM3\_1591  &  & 6  & 5  & 5  & 100  & ribonuclease HIII \tabularnewline
SpyM3\_1753  &  & 6  & 5  & 98  & 100  & putative transcriptional regulator 
\tabularnewline
SpyM3\_1753  & 11 & 3  & 1  & 53  & 100  & putative transcriptional regulator 
\tabularnewline
SpyM3\_1754  & 11 & 3  & 1  & 54  & 100  & putative transcriptional regulator 
\tabularnewline
SpyM3\_1754  &  & 6  & 5  & 95  & 100  & putative transcriptional regulator 
\tabularnewline
SpyM3\_1778  &  & 6  & 0  & 34  & 100  & putative cationic amino acid 
transporter protein \tabularnewline
SpyM3\_1809  &  & 4  & 0  & 5  & 100  & arginyl-tRNA synthetase \tabularnewline
SpyM3\_1813  & 12 & 3  & 0  & 19  & 100  & hypothetical protein \tabularnewline
SpyM3\_1814  & 12 & 3  & 0  & 29  & 100  & aspartyl-tRNA 
synthetase\tabularnewline
\hline
\end{tabular}

$^a$ Numbers represent a group of loci that are consecutive in gene order. Empty
cells denote stand alone loci,

$^b$ Donor branches of homologous recombinant edges, 0 (SDE1), 1 (SDE2), 2
(SDD), 3 (SPY1), 4 (SPY2), 5 (SDE), 6 (SPY), 7(SD), and 8(ROOT),

$^c$ Recipient branches of recombinant edges,

$^d$ Proportion of sites that the recombinant edge covers out of the sites in
the core alignment block, and

$^e$ Percentage of the sites in the core alignment block out of total sites of
the locus.
}

\end{table}
\clearpage{}

\section*{Supplementary Figures}

\clearpage{}

% Figure S1: Posterior probability of recobinant edges on aspartyl-tRNA.
\begin{figure}
\includegraphics[scale=0.6]{figures/yl-trna}
\caption{\label{fig:muts}Posterior probability of recombinant edges on 
aspartyl-tRNA
synthetase.  The
top line represents a genomic location of the current view in SPY1 genome. Locus
tags are represented below. The start of a locus tag is red colored, and the end
blue colored. Each of nine horizontal bars represents a branch of the five
species tree. The species tree branch is a recipient species.
The red area on SDE1 line represents gene fragment transfer from
SPY1 due to recombination in locus SpyM3\_1814 of aspartyl-tRNA synthetase.}
\end{figure}
\clearpage{}%

% Figure S2: 
\begin{figure}
\subfloat[]{\includegraphics[scale=0.5]{figures/h1-R}}
\subfloat[]{\includegraphics[scale=0.5]{figures/h1-R-inset}}
\caption{\label{fig:h1}Numer of recombination events. The diagonal dashed
lines are with slope of 1 and intercept of 0. For each order
pairs of branches of a reference species tree the true number of recombination
events is computed as the mean of the 10 replicates of recombinant
trees from the simulation. With ten estimates of each replicate total
number of 100 estimats are shown as intervals of 90\% of quantile
range, or the lower value of the interval is the 5\% quantile, and
the upper one is the 95\% quantile. }
\end{figure}
\clearpage{}%

% Figure S3:
\begin{figure}
\subfloat[]{\includegraphics[scale=0.5]{figures/h2-R}}
\subfloat[]{\includegraphics[scale=0.5]{figures/h2-R-inset}}

\subfloat[]{\includegraphics[scale=0.5]{figures/h2-R-inset2}}

\caption{\label{fig:h2}Comparison of number of recombinations between real
and simulated data sets. The diagonal dashed lines are with
slope of 1 and intercept of 0. For each order pairs of branches of
a reference species tree the number of recombination events is estimated
from the real data set. Total number of 100 estimats from a simulated
data set are shown as intervals of 90\% of quantile range, or the
lower value of the interval is the 5\% quantile, and the upper one
is the 95\% quantile. }
\end{figure}
\clearpage{}%

% Figure S4
\begin{figure}
\includegraphics{figures/ri-R}
\caption{\label{fig:ri1}The surrogate measure of recombination intensity
of genes. Each point represents a pair of a true value of average number of
recombinant edges and its estimated value for a gene. The dashed line is with
slope of 1 and intercept of 0.}
\end{figure}
\clearpage{}%

% Figure S5: The species tree
\begin{figure}
\includegraphics[scale=0.7]{figures/species}
\caption{\label{fig:tree5}Species tree of five taxa for ClonalOrigin analysis.
SPY is the branch of a clade\emph{ of Streptococcus pyogenes}, SDE
is that for \emph{S.\ dysgalactiae} ssp.\emph{\ equisimilis}, and SD is the
branch for the three individuals of \emph{S.\ dysgalactiae}. The scale is
proportional to the expected number of mutations given the Watterson's estimate 
of mutation
rate $\theta$ and the tree.}
\end{figure}
\clearpage{}%

% Figure S6: Histogram of block sizes
\begin{figure}
\begin{center}
\includegraphics[width=5in]{figures/blocksize}
\end{center}
\vspace{-.3in}
\caption{Distribution of core alignment block sizes.}
\label{fig:blocksize}
\end{figure}
\clearpage{}%

% Figure S7: gene family sizes
%=============================================================================
% FIGURE - gene family sizes
\begin{figure}
\begin{center}
\includegraphics[width=5in]{figures/famsizes}
\end{center}
\vspace{-.3in}
\caption{Distribution of gene family sizes.}
\label{fig:famsize}
\end{figure}
%=============================================================================

\clearpage

% Figure S8: recombination heatmap from mowgli
%=============================================================================
% FIGURE - Transfer heatmap
\begin{figure}
\begin{center}
\includegraphics[width=5in]{figures/trans-recomb-heatmap}
\end{center}
\vspace{-1.2in}
\caption{Distribution of recombining horizontal gene transfers.}
\label{fig:mowgli-recomb-heatmap}
\end{figure}
%=============================================================================

\end{document}

\begin{figure}
\includegraphics[scale=0.7]{figures/plot-number-recombination-within-blocks-1\lyxdot out\lyxdot recomb}
\caption{\label{fig:recombination-boundaries}Distribution of number of 
recombination
events boundaries. The number of recombination event boundaries per
site for each block along the \emph{S.\ dysgalactiae} ssp.\emph{\ equisimilis}
ATCC 12394.}
\end{figure}
\clearpage{}%

The mechanisms of horizontal gene transfer
include conjugation by plasmid, transduction by bacteriophages, or
transformation of naked DNA uptake \citep{Thomas2005}. Additionally,
marine bacteria was reported to acquire foreign genes via gene transfer
agents, i.e., viral-like particles produced by $\alpha$-Proteobacteria
\citep{McDaniel2010}.

Although
recombinant edges between two species branches do not explicitly denote
gene flow between the two species in their original model, \citet{Didelot2010}
exploited their method to infer flows of genetic segment between \emph{Bacillus
cereus} species. This model-based approach to detecting homologous
recombining gene transfer was used to the core genomes of closely
related \emph{Streptococcus} species.

{[}WHERE SHOULD I PUT THESE UNIT OF GENE TRANSFER?{]}

\textbf{Units of gene transfer:} \citet{Chan2009a} showed that protein
domains were not gene transfer units. In another paper \citet{Chan2009}
studied gene with and without breakpoints, and found that more often
entire genes transfer in pathogens than in non-pathogens. In pathogenic
bacteria they found more genes with breakpoints than genes without
breakpoints.

% Table and SDE and SPY
Consistent
with the number of recombinant edges were that gene transfer fragments
with high posterior probability shown in Table S\ref{tab:genes-transfer}
often appeared to be from SPY to SDE. While four genes were inferred
to be transferred to SPY1, there were more genes transferred to SDE1
or SDE2.

\textbf{Evolutionary hypotheses about the ancestor of SPY and SDE:}
\citet{Kalia2001} proposed two scenarios of the evolution of the
two species: The first one was that the most recent common ancestor
of SPY and SDE was a GAS-like pathogen, SDE was losing virulence factors
since the divergence of the ancestor, and SDE regained virulence factors
from SPY because of the same niche. The second scenario was that the
most recent common ancestor of SPY and SDE was a GGS/GCS-like commensal,
SPY evolved virulence factors since the divergence of the ancestor,
and SDE gained virulence factors from SPY. Commensal bacteria could
evolve by aquiring virulence genes from its closely related pathogenic
bacteria. \citet{Kalia2001} suggested the divergence time of SDD
to be around the time when human started to domesticate wildlife animals
around 10,000 years ago. A demographic model could be made with two
species to study the divergence time quantitatively. This would require
ancestral recombination graphs embedded in a species tree.

% From discussion
This distinction of the two processes of
gene transfer was envisioned a decade ago \citep[e.g.,][]{Ochman2001},
and the methods of gene transfer detection by distinguishing the processes
were not yet employed even if some of the methods were already available.
We by no means provided any solutions to the problem of detecting
gene transfer here. Yet, we hoped that we should be able to discriminate
the two processes because each of the processes of gene transfer could
affect biological functions that were different in nature.

% From discussion
For example, genes transferred via non-recombining
gene transfer could be parts of functional units such
as protein domains, and secondary structures of ribosomal RNAs. Genes
transferred via homologous recombining gene transfer could be small
fragments of genes as long as the flanking sequences of the fragments
were more or less homologous.

\textbf{Overlap of genes under positive selection and higher recombination:}
As the uncertainty in gene tree topologies becomes larger, false positive
rates of detecting sites under positive selection would be high (CITE).
Genes could be divided into parts where we hoped no multiple gene
tree topologies exist. Dividing genes into those fragments might be
a way to avoid potentially high false positive rates of positively
selected sites. Considering the fact that nucleotide substitutions
seemed to be affected more by recombinations than by mutations we
suspect that fragmenting genes might not be helpful.

% Virulence gene and non-recombining gene transfer
We investigated bacterial gene
transfer between SPY and SDE using the genome sequences by comparing
gene tree and species tree. We considered both core and pan genomes
for detecting homologous recombining gene transfer and non-recombining
gene transfer. 
%
The extent to which genes were transferred would vary on this
mode of gene transfer. 
While homologous
recombination gene transfer would happen in the genes whose donor
and recipient species were rather closely related, non-homologous
gene transfer could happen virtually in any genes of donor
and recipient species as long as the species shared their habitats.
%
We did not find an excess of virulence
genes associated homologous recombining gene transfer. Actually, many important
virulence genes were missing in the core genomes. We found association
of virulence genes with non-recombining gene transfer.

% Number of recombinant edges (Methods)
This would indicate the degree to which the clonal
species tree was disrupted by bacterial recombination. 
%
We used the method to infer which types of bacterial recombination
were more prevalent than which other types in the lineages of the
tree relating five streptococcus genomes.

% Watterson's estimate
in that the rate of mutation
on the branches of the genealogy and branch lengths of the species
tree $\mathcal{T}$ are reciprocally proportional.

% simulation for ClonalOrigin stage 1
We hoped that we could understand the power
and limitations of ClonalOrigin under the parameter setup that we
inferred from our data set.

% Heat map for recombination from ClonalOrigin
Because the
lengths of SPY1 and SPY2 were larger than those of SDE1 and SDE2,
recombinant edges could more often depart from one of SPY1 and SPY2
lineages and arrive at one of SDE1 and SDE2 lineages than those for
the reverse direction \emph{a priori} (Figure S\ref{fig:tree5}). 
We took into account of the prior expected number of recombinant edges.

\textbf{Gene transfer and functional categories:} We showed much clearer
examples of lateral transfer of gene fragments.
Modular structures of aaRS genes made them
easy targets of gene transfer \citep{Woese2000}. The two transporter
genes might have been under selective pressure from environment.

% rRNA binding
Of the genes in Table S\ref{tab:genes-transfer} we did not find ribosomal
proteins that were heavily enriched in the gene ontology term, {}``rRNA
binding.'' 
Ribosomal proteins tended to cluster together in the genome,
and they appeared to be affected by the recombinant edges whose departures
were on the root branch. 
Recombinant edges that connect the root branch
are hard to be considered as gene flow. We ruled out those recombinant
edges in Table S\ref{tab:genes-transfer} as we mentioned.

### 2011-07-13 pl/scatter-plot-parameter.pl

@@ -84,7 +84,7 @@ A prefix of ClonalOrigin output files in XML format is required.
 =item B<-out> <base name of output file>
 
 Three files are generated for the three parameters. The option string of out is
-used as a base name of them: i.e., out.theta, out.rho, and out.delta.
+used as a base name of them: i.e., out-theta, out-rho, and out-delta.
 
 =back
 
@@ -168,9 +168,9 @@ my $position;
 ##############################################################
 # Open the three output files.
 ##############################################################
-open OUTTHETA, ">$basenameOutFile.theta" or die $!;
-open OUTRHO, ">$basenameOutFile.rho" or die $!;
-open OUTDELTA, ">$basenameOutFile.delta" or die $!;
+open OUTTHETA, ">$basenameOutFile-theta" or die $!;
+open OUTRHO, ">$basenameOutFile-rho" or die $!;
+open OUTDELTA, ">$basenameOutFile-delta" or die $!;
 
 ##############################################################
 # Start to parse the XML file.

### 2011-07-13 sh/filter-blocks.sh

@@ -86,10 +86,10 @@ function run-lcb {
 }
 
 function mkdir-tmp {
-  mkdir -p $TMPINPUTDIR
+  mkdir -p $TEMPINPUTDIR
   read-species-genbank-files data/$SPECIES mkdir-tmp
 }
 
 function rmdir-tmp {
-  rm -rf $TMPDIR
+  rm -rf $TEMPDIR
 }

### 2011-07-13 sh/manuscript.sh

@@ -1,15 +1,56 @@
+# SPECIES, REPETITION, REPLICATE are three variables that need to be specified.
+# The following variables should be chosen by considering ClonalFrame runs.
+# CLONALFRAMEREPLICATE=1
+# RUNID=1
 function manuscript {
   SPECIES=cornellf
   REPETITION=3
+  REPLICATE=1
   set-more-global-variable $SPECIES $REPETITION
   CLONALFRAMEREPLICATE=1
   RUNID=1
   MANUSCRIPTOUTPUT=manuscript/output
-  manuscript-sequence-alignment
-  manuscript-watterson
-  manuscript-tree
+  # manuscript-core-alignment
+  # manuscript-sequence-alignment
+  # manuscript-watterson
+  # manuscript-clonalframe
+  # manuscript-tree
+  # manuscript-heatmap
+  # manuscript-summarize-clonalorigin-first
+  # manuscript-keynotes
+  # manuscript-simulation-clonalorigin1
+  # manuscript-simulation-clonalorigin2
+  # manuscript-probability-recedge
+  # manuscript-tree-topology
+  # manuscript-functional-category
+  cp $MANUSCRIPTOUTPUT/*.eps doc/figures/
 }
 
+# ClonalFrame estimated the relative strength of recombination
+# compared to mutations to be $r/m = 4.66$, indicating 
+# that reconmination-driven substitutions affected more 
+# sites than did mutation-driven substitutions. The first stage of ClonalOrigin 
+function manuscript-clonalframe {
+  CLONALFRAMEOUT=$RUNCLONALFRAME/output/${CLONALFRAMEREPLICATE}/core_clonalframe.out.${RUNID}
+  echo "Use ClonalFrame GUI to read $CLONALFRAMEOUT"
+  echo "Find values of r/m using ClonalFrame GUI"
+  echo -e "END of $FUNCNAME\n"
+}
+
+# The number of blocks is 274.
+function manuscript-core-alignment {
+  OUT=manuscript/$FUNCNAME.txt
+  rm -f $OUT
+  XMFA=$DATADIR/core_alignment.xmfa
[…]

### 2011-07-13 sh/run.sh

@@ -152,6 +152,7 @@ fi
 SIMULATIONS=$(ls species|grep ^s)
 SPECIESS=$(ls species|grep -v ^s)
 
+source sh/utility.sh
 source sh/hms.sh
 source sh/set-more-global-variable.sh
 source sh/mkdir-species.sh
@@ -1116,9 +1117,8 @@ CHOICES=( init-file-system \
           compute-global-median \
           create-ingene \
           extract-species-tree \
-          --- MANUSCRIPT-Figure1 ---\
-          manuscript \
-          xxx )
+          --- MANUSCRIPT ---\
+          manuscript )
 select CHOICE in ${CHOICES[@]}; do 
   if [ "$CHOICE" == "" ];  then
     echo -e "You need to enter something\n"

### 2011-07-13 sh/scatter-plot-parameter.sh

@@ -24,7 +24,7 @@ function scatter-plot-parameter {
       perl pl/$FUNCNAME.pl \
         -xmlbase $RUNCLONALORIGIN/output/$REPLICATE/core_co.phase2 \
         -xmfabase $DATADIR/core_alignment.xmfa \
-        -out $RUNANALYSIS/$FUNCNAME-$REPLICATE.out
+        -out $RUNANALYSIS/$FUNCNAME-$REPLICATE-out
       echo "$RUNANALYSIS/$FUNCNAME-$REPLICATE.out"
 
       NUMBER_BLOCK=$(echo `ls $DATADIR/core_alignment.xmfa.*|wc -l`)
@@ -39,7 +39,7 @@ function scatter-plot-parameter {
       MEDIAN_RHO=$(grep "Median rho" $RUNCLONALORIGIN/summary/${REPLICATE}/median.txt | cut -d ":" -f 2)
 
       analyze-run-clonalorigin-scatter-plot-parameter-rscript \
-        $RUNANALYSIS/$FUNCNAME-$REPLICATE.out 
+        $RUNANALYSIS/$FUNCNAME-$REPLICATE-out 
  
       #cat $RUNCLONALORIGIN/summary/${REPLICATE}/median.txt
       echo -e "Prepare 2nd run using prepare-run-2nd-clonalorigin menu!"
@@ -56,7 +56,7 @@ library("geneplotter")  ## from BioConductor
 require("RColorBrewer") ## from CRAN
 
 plotThreeParameter <- function (f, xlab, ylab, m, logscale) {
-  x.filename <- paste (f, ".ps", sep="")
+  x.filename <- paste (f, ".eps", sep="")
   x <- read.table (f);
 
   pos.median <- c() 
@@ -78,9 +78,9 @@ plotThreeParameter <- function (f, xlab, ylab, m, logscale) {
   }
   dev.off()
 }
-plotThreeParameter ("$S2OUT.theta", "Genomic position on S. dysgalactiae ssp. equisimilis ATCC 12394", "Mutation rate per site", $MEDIAN_THETA, FALSE)
-plotThreeParameter ("$S2OUT.rho", "Genomic position on S. dysgalactiae ssp. equisimilis ATCC 12394", "Recombination rate per site", $MEDIAN_RHO, FALSE)
[…]

### 2011-07-13 sh/set-more-global-variable.sh

@@ -99,8 +99,7 @@ function set-more-global-variable {
   # submitted in the cluster in CAC cluster, the job creates a temporary
   # directory where it can save the input and output files. JOBID is the CAC job
   # id for run-mauve. This job ID should be found in the Mauve alignment file.
-  TMPDIR=/tmp/$JOBID.scheduler.v4linux
-  TMPINPUTDIR=$TMPDIR/input
-
+  TEMPDIR=/tmp/$JOBID.scheduler.v4linux
+  TEMPINPUTDIR=$TEMPDIR/input
 }
 

### 2011-07-13 sh/summary-core-alignment.sh

@@ -17,6 +17,10 @@ function summary-core-alignment {
         -in $DATADIR/core_alignment.xmfa \
         -out $RUNANALYSIS/$FUNCNAME.txt
   
+      echo "Check $RUNANALYSIS/$FUNCNAME.txt"
+      echo "----"
+      cat $RUNANALYSIS/$FUNCNAME.txt
+      echo "----"
       break
     fi
   done

### 2011-09-26 doc/manuscript.tex
I am focusing on rewriting the manuscript as a resource paper.

@@
-%% LyX 1.6.7 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
+% We approach the project of streptococcus recombination and gene transfer from
+% the angle of a resource paper. 
 \documentclass[english]{article}
@@
 \usepackage{longtable}
-\usepackage{epstopdf}
 \doublespacing
@@
 
-Inferring two types of gene transfer using whole genomes of \emph{Streptococcus
-pyogenes} and\emph{ S.\ dysgalactiae} subspecies \emph{equisimilis}
+Inferring homologous recombination and horizontal gene transfer 
+using whole genomes of \emph{Streptococcus pyogenes} 
+and\emph{ S.\ dysgalactiae} subspecies \emph{equisimilis}
 
@@
 
-\subsubsection*{Version: 0.65 (July 8, 2011) \clearpage{}}
+\subsubsection*{Version: 0.66 (Sep 30, 2011) \clearpage{}}
 
 \section*{Abstract}
-
-Using two closely related species of \emph{Streptococcus pyogenes}
-(SPY) that is a pervasive human-specific pathogen, and \emph{S.\ dysgalactiae}
-ssp.\emph{\ equisimilis} (SDE) that is its closely related species
-and an opportunistic human pathogen we investigated two types of gene
-transfer: recombining and non-recombining
-gene transfer. The former lateral process was studied
-with the core genomes of \emph{Streptococcus} genomes, and the latter
-process with the pan genomes as well as the core genomes. Two
-complementary methods were employed to examine genes and their
-fragments that could have been exchanged between SPY and SDE as efforts
-of finding their roles in the formation of pathogenicity of the bacteria.
-We revisited the problem of directionality of gene transfer between
-SPY and SDE, and found evidence supporting that homologous
-recombining gene transfer was polarized to be relatively from SPY to SDE based
-on the core genomes.  Although we did not find evidence for directional
-non-recombining gene transfer, virulence genes appeared to undergo
-non-recombining gene transfer more than did non-virulence genes.
-Recombining gene transfer was not found to be associated with
-virulence genes.  This might indicate that additional genes obtained via
-non-recombining gene transfer were more effective than gene fragments
-transfer via bacterial recombination for SDE in the acquisition of virulence
-genes.  While horizontal gene transfer studies traditionally 
-focused on genes as a whole,
-we rediscovered smaller gene fragments and their transfer
-due to recombination between SPY and SDE in aminoacyl tRNA
-synthetase (aaRS), and phosphotransferase systems (PTS).
+Homologous recombination and horizontal gene transfer are investigated
+with complete genomes of closely related bacterial species of 
+\emph{Streptococcus pyogenes} (SPY) 
+and \emph{S.\ dysgalactiae} subspecies \emph{equisimilis} (SDE).
+The former is a pervasive human-specific pathogen, and the latter
+an opportunistic human pathogen.  Parts of the genomic configuration, including
+human pathogenicity, of SDE could have originated in those of SPY.  
+Our genome-wide survey of gene transfer indicated that gene transfer via homologous
+recombination occurred relatively from SPY to SDE.
+Although we did not find enrichment of virulence genes in the homologous recombining gene
+transfer, we found evidence that they transferred more often via
+non-recombining fashion than non-virulence genes. 
+A genome-wide resource of recombining and non-recombining gene transfer for the study of the
+genomes is provided so that research community in pathogenic bacteria can take
+advantage of it.
 
@@
 
-\textbf{Two lateral processes of gene transfer:} 
-Foreign bacterial genes can be laterally
-brought into a bacterial genome by two different processes 
-\citep{Ochman2001,Lawrence2009}:
-homologous recombining transfer and non-recombining transfer.
-In the former process donor genes that are homologous to those in
-a recipient species replace the indigenous genes. In the latter process
-donor genes are integrated into a recipient species without replacement
-of genes in the species. 
-Although the discrimination between the two
-lateral processes was noted a decade ago, methods of detecting horizontal
-gene transfer did not appear to properly take the differences into
-account. 
-Proper discrimination of the two processes of gene transfer
-could lead to better understanding of biological nature
-of transferred genes. For example, 
-bacterial lateral transfers of small fragments of ribosomal RNA genes 
-were known to be 
-due to homologous recombination \citep{Sneath1993,Yap1999} not non-recombining
-gene transfer. 
-Because traditionally non-recombining gene transfer in bacteria has been
-referred to as horizontal gene transfer, association of ribosomal RNA genes with
-just lateral transfer might be misleading.
+Bacteria evolves new biological features through horizontally transferred genes
+from other lineages as well as vertically transmitted mutations on genes within
+the same lineage.  Mutations on a gene modifiy relatively small portions of the
+gene, including substitutions, insertions, and deletions of deoxyribonucleic
+acids. Horizontal gene transfers can reshape the organization of a bacterial
+genome, which could revamp the biological function of the bacterial species.
+Recipient bacteria can substitute transferred genes for homologous parts of
+their genomes. They can also integrate transferred genes in their genomes
+without replacement. These two processes are discriminated by calling the former
+\textit{recombining gene transfer}, and the latter 
+\textit{horizontal gene transfer} \citep{Ochman2001,Lawrence2009}.  
+Despite the two distinct evolutionary processes in bacteria, reseraches on
+statistical analyses of complete genome sequences in efforts of understanding
+bacterial evolution have paid little attention to a combined approach to
+illuminating disparity of the two gene transfer processes.
+
+HERE
 
@@
 genomic segments flow between the two species and directions of thereof.
+Cite this: \citep{Ahmad2009}.
 
@@
 Figure S\ref{fig:blocksize} shows the distribution of block lengths. The average
-length was 4215 bp, and total length of the blocks is 1155016 bp.  There was good 
+length was 4188 bp.  There was good 
 agreement on the topology among independent chains, and reasonable agreement on 
@@
 \begin{figure}
-% \includegraphics[scale=0.42]{figures/mauve-analysis-002}
-\includegraphics[scale=0.42]{figures/figure1}
+\includegraphics[scale=0.42]{figures/mauve-analysis-002}
 \caption{\label{fig:clonalorigin}Recombinant tree in the ClonalOrigin model.
@@
 \begin{figure}
-\includegraphics[scale=0.45]{figures/figure2}
+\includegraphics[scale=0.45]{figures/heatmap-recedge}
 \caption{\label{fig:Heatmap-of-recombination}Heatmap of recombination events.
@@
 \begin{figure}
-\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-theta}}
-
-\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-rho}}
-
-\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-delta}}
+\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot theta}}
+
+\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot rho}}
+
+\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot delta}}
 \caption{\label{fig:scatter3}Scatter plots of mutation rate, recombination
@@
 \begin{center}
-\includegraphics[width=5in]{figures/trans-hgt-heatmap}
+\includegraphics[width=5in]{figures/trans-hgt-heatmap.pdf}
 \end{center}
@@
 \begin{figure}
-\includegraphics{figures/ri-R}
+\includegraphics{figures/ri1}
 \caption{\label{fig:ri1}The surrogate measure of recombination intensity
@@
 \begin{figure}
-\includegraphics[scale=0.7]{figures/species}
+\includegraphics[scale=0.7]{figures/cornellf-3-tree}
 \caption{\label{fig:tree5}Species tree of five taxa for ClonalOrigin analysis.
@@
 \begin{center}
-\includegraphics[width=5in]{figures/blocksize}
+\includegraphics[width=5in]{figures/blocksize.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=5in]{figures/famsizes}
+\includegraphics[width=5in]{figures/famsizes.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=5in]{figures/trans-recomb-heatmap}
+\includegraphics[width=5in]{figures/trans-recomb-heatmap.pdf}
 \end{center}

### 2011-09-28 doc/manuscript.tex
The result and UCSC genome browser resource is more refined to do something about virulence genes.

@@
 Inferring homologous recombination and horizontal gene transfer 
-using whole genomes of \emph{Streptococcus pyogenes} 
-and\emph{ S.\ dysgalactiae} subspecies \emph{equisimilis}
+using whole genomes of \textit{Streptococcus pyogenes} 
+and\textit{ S.\ dysgalactiae} subspecies \textit{equisimilis}
 
@@
 A streptococci (GAS), group C streptococci (GCS), group G streptococci
-(GGS), \emph{Streptococcus pyogenes} (SPY), \emph{S.\ dysgalactiae}
-subspecies \emph{dysgalactiae} (SDD), \emph{S.\ dysgalactiae} subspecies
-\emph{equisimilis} (SDE).
+(GGS), \textit{Streptococcus pyogenes} (SPY), \textit{S.\ dysgalactiae}
+subspecies \textit{dysgalactiae} (SDD), \textit{S.\ dysgalactiae} subspecies
+\textit{equisimilis} (SDE).
 
@@
 \section*{Abstract}
-Homologous recombination and horizontal gene transfer are investigated
-with complete genomes of closely related bacterial species of 
-\emph{Streptococcus pyogenes} (SPY) 
-and \emph{S.\ dysgalactiae} subspecies \emph{equisimilis} (SDE).
-The former is a pervasive human-specific pathogen, and the latter
-an opportunistic human pathogen.  Parts of the genomic configuration, including
-human pathogenicity, of SDE could have originated in those of SPY.  
-Our genome-wide survey of gene transfer indicated that gene transfer via homologous
-recombination occurred relatively from SPY to SDE.
-Although we did not find enrichment of virulence genes in the homologous recombining gene
-transfer, we found evidence that they transferred more often via
-non-recombining fashion than non-virulence genes. 
-A genome-wide resource of recombining and non-recombining gene transfer for the study of the
-genomes is provided so that research community in pathogenic bacteria can take
-advantage of it.
+Homologous recombination and horizontal gene transfer are investigated with
+complete genomes of closely related bacterial species of \textit{Streptococcus
+pyogenes} (SPY) and \textit{S.\ dysgalactiae} subspecies \textit{equisimilis}
+(SDE).  The former is a pervasive human-specific pathogen, and the latter an
+opportunistic human pathogen.  Parts of the genomic configuration, including
+human pathogenicity, of SDE could have originated in those of SPY.  Our
+genome-wide survey of gene transfer indicated that gene transfer via homologous
+recombination occurred relatively from SPY to SDE.  Although we did not find
+enrichment of virulence genes in the homologous recombining gene transfer, we
+found evidence that they transferred more often via non-recombining fashion than
+non-virulence genes.  A genome-wide resource of recombining and non-recombining
+gene transfer for the study of the genomes is provided so that research
+community in pathogenic bacteria can take advantage of it.
 
@@
 
+Bacteria infect humans with a variety of diseases: e.g., food poisoning
+(\textit{Eschericha Coli}), strep throat (\textit{Streptococcus pyrogenes}),
+tuberculosis (\textit{Mycobacterium tuberculosis}), cholera (\textit{Vibrio
+cholera}), anthrax (\textit{Bacillus anthracis}), pneumonia
+(\textit{Chlamydophila pneumoniae}), and influenza (\textit{Haemophilus
+influenzae}), to name a few.  \textit{Streptococcus pyogenes} (SPY) belonging to
+group A streptococci (GAS) primarily infect humans in the throat and skin
+\citep{Broyles2009}. It has high prevalence throughout the world causing
+diseases ranging from mild illnesses such as pharyngitis to severe invasive
+streptococcal diseases including necrotizing fasciitis and toxic shock syndrome
+\citep{Cunningham2000a}.  Although \textit{Streptococcus dysgalactiae}
+subspecies \textit{equisimilis} (SDE) belonging to group C or G streptococci
+(GCS/GGS) used to be generally regarded as a veterinary pathogen
+\citep{Vandamme1996}, it has been reported to cause serious and life-threatening
+streptococcal diseases to humans traditionally associated with GAS
+\citep[e.g.,][]{Brandt2009}.  Because of overlap in the disease profile and the
+ecological niche of tissue sites betwen SPY and SDE, these bacteria may as well
+interact with each other by exchanging parts of their genomes.
+
+The research topic on gene exchanges between SPY and SDE has been discussed in
+various ways. \citet{Sachse2002} showed gene transfer from SDE to SPY for the
+downstream of the genomic region of streptococcal pyrogenic exotoxins
+\textit{speG}. \citet{Davies2007} investigated phage 3396 from SDE as a vehicle
+of gene transfer from SDE to SPY. \citet{Davies2009} demonstrated that an
+integrative conjugative element of SDE allowed gene transfer from SDE to other
+streptococci.  \citet{Ahmad2009} did not conclusively deduce the net
+directionality of housekeeping gene transfer between SPY and SDE.  Notably,
+\citet{Kalia2001} attempted to show that the direction of gene transfer from SPY
+to SDE was more dominant than the other direction.  Because these researches on
+the controversial issue, directionality of gene transfer, have been limited to
+relatively small regions in the genomes, it remains to be elusive to determine
+the genome-wide degrees to which genomic segments flow between the two species,
+and directions of thereof.
+
+As sequencing costs plummet \citep{Mardis2011}, bacterial genome sequencing
+produces individual genomes of high quality even within a bacterial species
+\citep{Tettelin2009a}.  Studies on horizontal gene transfer in prokaryotes have
+used genes in very large sets of species \citep{Koonin2001} from a great deal
+of genomes available.  Although this type of study gave us insights into complex
+evolutionary histories of prokaryotes, with multiple individual genomes per
+species we now can go beyond the broad study on gene transfer towards
+fine-scaled history of gene transfer between very closely related prokaryotes.
+For instance, \citet{Liu2009} studied gene transfer events in
+\textit{Lactobacillus bulgaricus} and \textit{Streptococcus thermophilus} using
+the compositional approach developed by \citet{Karlin2001}. \citet{Luo2011}
+investigated gene transfer events in \textit{Eschericha Coli} using the embedded
+quartet decomposition analysis developed by \citet{Zhaxybayeva2006}.
+\citet{Hamady2006} combined phylogenetic and compositional approaches to
+detecting horizontal gene transfer.  \citet{Caro-Quintero2011} studied a number
+of \textit{Shewanella baltica} genomes to identify the genetic elements that
+enabled the species to adapt to redox gradients.
+
 Bacteria evolves new biological features through horizontally transferred genes
-from other lineages as well as vertically transmitted mutations on genes within
+from other lineages and vertically transmitted mutations on genes within
 the same lineage.  Mutations on a gene modifiy relatively small portions of the
@@
 acids. Horizontal gene transfers can reshape the organization of a bacterial
-genome, which could revamp the biological function of the bacterial species.
+genome, which could revamp biological functions of the bacterial species.
 Recipient bacteria can substitute transferred genes for homologous parts of
-their genomes. They can also integrate transferred genes in their genomes
-without replacement. These two processes are discriminated by calling the former
-\textit{recombining gene transfer}, and the latter 
-\textit{horizontal gene transfer} \citep{Ochman2001,Lawrence2009}.  
-Despite the two distinct evolutionary processes in bacteria, reseraches on
-statistical analyses of complete genome sequences in efforts of understanding
-bacterial evolution have paid little attention to a combined approach to
-illuminating disparity of the two gene transfer processes.
-
-HERE
-
-\textbf{Gene transfer between SPY and SDE:} Through virulence genes
-bacteria can infect humans
-with a variety of diseases: e.g., food poisoning by \emph{Eschericha
-Coli}, strep throat by \emph{Streptococcus pyrogenes}, tuberculosis
-by \emph{Mycobacterium tuberculosis}, cholera by \emph{Vibrio cholera},
-anthrax by \emph{Bacillus anthracis}, pneumonia by \emph{Streptococcus
-pneumoniae} and \emph{Chlamydophila pneumoniae}, and influenza by
-\emph{Haemophilus influenzae}, to name a few. \emph{S.\ pyogenes}
-(SPY) belonging to group A streptococci (GAS) primarily infect humans
-in the throat and skin \citep{Broyles2009}. It has high prevalence
-throughout the world causing diseases ranging from mild illnesses
-such as pharyngitis to severe invasive streptococcal diseases including
-necrotizing fasciitis and toxic shock syndrome \citep{Cunningham2000a}.
-A closely related species called\emph{ S.\ dysgalactiae} ssp.\emph{\ 
-equisimilis}
-(SDE) belonging to group C or G streptococci (GCS/GGS) used to be
-generally regarded as a veterinary pathogen \citep{Vandamme1996}.
-Researchers have been confirming that GCS/GGS can cause serious and
-life-threatening streptococcal diseases to humans traditionally associated
-with GAS \citep[e.g.,][]{Brandt2009}. Because SDE can cohabit with
-SPY at the same ecological niche of tissue sites, and both may exhibit
-overlap in the disease profile, the close proximity
-may allow the two species to interact with each other by exchanging
-parts of their genomes. It has been suspected that opportunistic pathogenicity
-of SDE may have been facilitated by the very genetic interaction between
-the two species. 
-Although virulence genes were suspected to be acquired via non-recombining gene
-transfer rather than recombining transfer \citep{Groisman1994}, systematic
-approaches to elucidating relative effects of the two types of gene transfer 
-rarely have been applied. This unfortunate situation motivated us to study the
-closely related bacteria as efforts of better understanding relative 
-association of gene transfer with virulence genes.
-We hoped that understanding their genetic exchanges
-would enhance our knowledge of mechanisms of their pathogenicity on
-humans.
-
-\textbf{Previous studies on gene transfer and its directionality between
-SPY and SDE:} 
-Because of the close interaction between SPY and SDE, 
-possible donor and recipient relationship of the two species were noted before.
-However, previous researches in gene transfer between
-the two pyogeneic species of SPY and SDE were limited to small portion
-of the bacterial genome, mostly focusing on several genes 
-\citep{Kalia2004,Towers2004,Bessen2005}.
-For example, \citet{Sachse2002} showed gene transfer from SDE to
-SPY for the downstream of the genomic region of streptococcal pyrogenic
-exotoxins \emph{speG}. \citet{Davies2007} investigated phage 3396
-from SDE as a vehicle of gene transfer from SDE to SPY. \citet{Davies2009}
-demonstrated that an integrative conjugative element of SDE allowed
-gene transfer from SDE to other streptococci. On the contrary, \citet{Kalia2001}
-attempted to show that the direction of gene transfer from SPY to
-SDE was more dominant than the other direction, albeit unfortunately
-appeared to be premature. Because their researches on direction of
-gene flow were limited to only small portion of regions in the genomes,
-it remains to be elusive to determine the genome-wide degrees to which
-genomic segments flow between the two species and directions of thereof.
-Cite this: \citep{Ahmad2009}.
-
-\textbf{Methods of studying gene transfer in bacteria:} 
-Studies on horizontal gene transfer in prokaryotes have focused on genes in very
-large sets of species \citep{Koonin2001} using a great deal of genomes
-available.  These studies gave us insights into complex evolutionary histories
-of bacteria and the Archaea.  As sequencing costs plummet \citep{Mardis2011},
-bacterial genome sequencing produces genomes of high quality even within
-bacterial species \citep{Tettelin2009a}.  With more individual genomes per
-species we now can go beyond the broad study on gene transfer towards
-fine-scaled history between very closely related prokaryotes.  For example,
-\citet{Liu2009} studied gene transfer events in \emph{Lactobacillus bulgaricus}
-and \emph{Streptococcus thermophilus} using the compositional approach developed
-by \citet{Karlin2001}. \citet{Luo2011} investigated gene transfer events in
-\emph{Eschericha Coli} using the embedded quartet decomposition analysis
-developed by \citet{Zhaxybayeva2006}. \citet{Hamady2006} combined phylogenetic
-and compositional approaches to detecting horizontal gene transfer.
-\citet{Caro-Quintero2011} studied a number of \emph{Shewanella baltica} genomes
-to identify the genetic elements that enabled the species to adapt to redox
-gradients. While compositional approaches were effective in analyzing
-genome-scale data, phylognetic approaches have allowed to understand
-evolutionary forces of gene exchange between species
-\citep[e.g.,][]{Bapteste2004,Simonson2005,Beiko2009} by comparing differences
-between gene trees and a reference species tree.  However, many of the methods
-did not have explicit models of homologous recombining gene transfer that was
-one of important evolutionary processes in horizontal gene transfer.  
-% Overall,
-% the development of methods for downstream analyses using bacterial genomes
-% appears to lag behind that of high-throughput sequencing technology
-% \citep[e.g.,][]{Green2011}.  
-Recently, a method of detecting recombination 
-with whole bacterial genomes \citep{Didelot2010} was developed. The explicit
-model of homologous recombination allowed them to study 
-flows of genetic segment between \emph{Bacillus cereus} species in the core
-genomes that shared by most of individuals. 
-In addition to recombination acting on bacterial genomes, genome rearragement,
-translocation, duplication, loss, and more complex processes would govern the
-evolution of bacteria. These processes, especially gene duplication, loss, and
-non-recombining gene transfer, would result in more complex relationship between
-even closely related bacteria. There have been progresses in 
-methods of inferring gene duplication, loss, and transfer
-events for gene families \citep{Doyon2011}. 
-We noted that
-the first approach could be used to infer transfer of genomic segments
-due to homologous recombination, and the second could do similarly
-due to non-recombining gene transfer.
-
-%\textbf{Last paragraph before methods:} 
-With these
-methods for detecting gene transfer we interrogated gene flows
-between SPY and SDE by measuring the relative impact of
-two different processes of gene transfer on the pathogenicity of SPY
-and SDE. This was possible thanks to recent release of bacterial
-genomes of SPY and SDE. 
-We here provided evidence that non-recombining
-gene transfer was more effective for acquisitions of virulence genes
-than homologous recombining gene transfer in SPY and SDE. 
-While recombination rates varied along
-the genomes, a few hot spots of recombination were observed. We found
-directional gene flows in homologous recombining gene transfer process from SPY
+their genomes. They can also integrate transferred genes into their genomes
+without replacement. These two processes are here discriminated by calling the
+former \textit{recombining gene transfer}, and the latter \textit{horizontal
+gene transfer} \citep{Ochman2001,Lawrence2009}.  Despite the two distinct
+evolutionary processes in bacteria, reseraches on statistical analyses of
+complete genome sequences in efforts of better understanding bacterial evolution
+have paid little attention to a combined approach to illuminating disparity of
+the two gene transfer processes.
+
+Homologous recombining gene transfer could be studied more readily using the
+core genomes from multiple individual bacterial genomes while horizontal gene
+transfer might be more easily detected using the pan genomes. Recently, a method
+of inferring homologous recombination with whole bacterial genomes was developed
+\citep{Didelot2010}. The model of homologous recombination allowed them to study
+flows of genetic segment between bacterial individuals of \textit{Bacillus
+cereus} species in the core genomes.  More complex evolutionary processes acting
+on bacterial genomes include genome rearragement, translocation, inversion,
+duplication, gain, loss, etc.  Because of the complexity, these processes do not
+readily lend themselves to statistically more sound approaches such as maximum
+likelihood. Instead, a parsimony-based approach to inferring gene gain, loss,
+and transfer events was developed by  \citep{Doyon2011}.  We noted that the
+first approach could be used to infer transfer of genomic segments due to
+homologous recombination, and the second could do similarly due to horizontal
+gene transfer. We expect that model-based approaches using whole bacterial
+genomes will be developed in the near future.
+
+The complete genomes of the two spcies of SPY and SDE presented a unique
+opportunity to study gene transfer over relatively recent time scales.  We used
+both of the model-based approach for inferring homologous recombination and the
+parsimony-based one for detecting gene gain, loss, and horizontal transfer. We
+also compared both of the results in recombining gene transfer events to show
+that the parimony-based method inferred gene transfer events in similar
+biological functional categories to those inferred with the model-based
+approach. This is the first attempt to analyaze SPY and SDE in genome-wide scale
+for detecting gene transfer events not only in the core genome but also in the
+pan genome, which make the analysis complete in terms of the amount of used
+genome data. Our analysis of core genomes also showed for the first time that
+the net directionality of, at least, recombining gene transfer was more from SPY
 to SDE.
-We showcased gene fragments of the inferred homologous gene transfers with high
-posterior probabilities, mostly appearing to move from SPY to SDE.
 
@@
 
-\section{Methods and Materials}
-
-\subsection{Data preparation}
-Two genomes of \emph{S.\ pyogenes} (SPY), and two of \emph{S.\ dysgalactiae}
-ssp.\emph{\ equisimilis} (SDE) were downloaded from NCBI with the following 
-accessions:
-NC\_004070 {Beres2002}, NC\_008024 \citep{Beres2006}, NC\_012891
-\citep{Shimomura2011}, and CP002215 \citep{Suzuki2011}. We refer to these 
-samples as SPY1, SPY2, SDE1, and SDE2, respectively. One genome
-of \emph{S.\ dysgalactiae} 
-ssp.\emph{\ dysgalactiae} was provided
-by \citet{Suzuki2011} with accession identifiers of SddyATCC27957, referred to 
-here as SDD. We also downloaded the genome (NC\_012471) of 
-\emph{S.\ equi} ssp.\emph{\ equi} strain 4047 referred to as SEE.
-
-\subsection{ClonalOrigin}
-We present a short summary of the statistical framework developed by
-\citet{Didelot2007} and \citet{Didelot2010} for studying bacterial 
-recombination. A detailed account was found in the two papers and in the user 
-guide available at \url{http://code.google.com/p/clonalorigin}. 
-The analysis required first identifying sufficiently long contiguous blocks of 
-high confident alignment. Multiple alignments with the six genomes
-led to a dramatic reduction in the core 
-genome length compared with those without SEE genome.
-Because blocks of long multiple sequence alignments
-were needed in order to obtain reliable 
-estimates of recombination rate, we omitted the SEE sample 
-from the ClonalOrigin analysis. We used progressiveMauve v2.3.1 
-\citep{Darling2004,Darling2010} with default options
-to align the five genomes.
-We then used stripSubsetLCBs, a program of the ClonalOrigin package
-to identify 276 blocks of sequence alignments with a length threshold of 
-1500 base pairs.  We excluded two of these
-blocks because they contained long consecutive gaps. Using these blocks, we ran 
-ClonalFrame v1.1 available at
-\url{http://www.xavierdidelot.xtreemhost.com} to obtain an 
-estimate of the species tree relating the five genomes. ClonalFrame also 
-outputs estimates of additional parameters, such as the average recombinant 
-tract length and the ratio between recombination and mutation rates, however, 
-these were not used in the subsequent analysis. We used a single Markov chain 
-Monte Carlo (MCMC) with 10000 burn-in generations and 10000 sampling 
-generations to obtain the tree estimate with branch lengths, 
-and seven additional independent runs to validate adequate 
-convergence of the first.
-
-The next step of the analysis involved running ClonalOrigin in order to infer 
-recombination histories across the alignment blocks. ClonalOrigin performs MCMC 
-sampling of population parameters and recombination histories, given the 
-species tree. The population parameters include mutation rate, 
-recombination rate, and average tract length of a recombinant edge. A 
-recombination history is represented by a \emph{recombinant tree}, which is a 
-species tree augmented with a set of \emph{recombinant edges}. 
-Although \citet{Didelot2010} described the concept of a recombinant tree
-in detail, we elected to use Figure \ref{fig:clonalorigin} 
-to illustrate it in order to clarify
-summarization procedures we employed of posterior samples of recombinant trees.
-Each recombinant 
-edge is defined by time and branch of origin, time and branch of destination, 
-and genomic interval in which it affects. The branch of origin is referred 
-to as the \emph{donor} and the destination branch is referred to as the 
-\emph{recipient}. A recombinant tree can be viewed as a restricted version of 
-an ancestral recombination graph (ARG), and as such it induces a local 
-genealogy for each site in the genomic block it describes. ClonalOrigin could
-jointly sample population parameters and 
-recombinant trees according to the posterior probabilities defined by a model 
-approximating the coalescence process with recombination.
-% \texttt{MAYBE ELABORATE MORE ON RESTRICTIONS OF CLONAL-ORIGIN'S MODEL? 
-% SC:I HAVE NOT THOUGHT ABOUT MORE RESTRICTIONS ON THE RECOMBINANT TREES.}
-
-% ClonalOrigin is applied in two stages. In both stages, sampling is performed 
-% separately for the different alignment blocks. In the first stage, full 
-% sampling is performed to obtain block-specific estimates of the model 
-% parameters integrating out the possible recombinant trees. The block-specific 
-% estimates of mutation rate, recombination rate, and average tract length are 
-% then combined using a weighted median scheme to obtain a global estimate for 
-% each parameter. In this scheme, each estimate is weighted according to the 
-% length of the block with which it is associated, down-weighing the less 
-% reliable estimates obtained from short blocks. In the second stage of 
-% ClonalOrigin, the model parameters are fixed to their global estimate across 
-% all blocks, and only the recombinant tree is sampled. Later this section we 
-% describe several summaries of this posterior sample that are used in our 
-% analysis. We used ClonalOrigin subversion r19 (downloaded from 
-% \url{http://clonalorigin.googlecode.com/svn/trunk}). For the first stage, we 
-% used two independent chains with 1,000,000 burn-in generations and  10,000,000 
-% sampling generations, sub-sampled every 100,000 generations. The first chain 
-% was used to obtain the block-specific estimates, and the second chain was used 
-% to validate adequate convergence of the first. For the second stage, we used a 
-% single chain with 10,000,000 burn-in generations and  100,000,000 sampling 
-% generations, sub-sampled every 100,000 generations.
-
-Two stages of MCMC with ClonalOrigin
-were employed as described in \citet{Didelot2010}. 
-Both stages of MCMC were performed independently for each of the 274 alignment blocks. 
-We used ClonalOrigin subversion r19 that had been downloaded from 
-\url{http://clonalorigin.googlecode.com/svn/trunk}. 
-In the first MCMC we obtained estimates of 
-mutation rate, recombination rate, and recombinant tract length. 
-Two independent chains were executed with $10^6$ burn-in generations and $10^7$ 
-sampling generations, sub-sampled every $10^5$-th generations. One of the two chains
-was used to obtain the block-wise mean estimates, and the other chain was used 
-to validate convergence of the first stage of MCMC. 
-The estimators of the three population parameters were medians weighted over
-block lengths, which down-weighed less reliable estimates obtained from shorter
-alignment blocks. 
-The second stage of MCMC with ClonalOrigin was executed with the three population
-parameters fixed to the estimated values in the first stage. 
-Two independent chains of the second stage
-of ClonalOrigin's MCMC were also executed with a total generation being
-$1.1\times10^8$. The first $10^7$ generations of each chain were discarded
-as burn-in. The remaining $10^8$ generations were sub-sampled
-every $10^5$-th generation.
-We now describe the method of summarizing
-the posterior sample to examine aspects of recombinant trees from the second
-stage of ClonalOrigin's MCMC.
-
-\citet{Didelot2010} presented two procedures of summarizing recombinant trees by
-viewing trees from two different angles: one from recombinant trees by
-aggregating all of the sites, and another from site-wise recombinant edge
-changes to the species tree. By following their approaches
-we quantified more aspects of recombinant trees: from the view of recombinant
-trees with aggregated sites we counted gene tree topologies, and from the view of
-site-wise recombinant edge changes we quantified recombination intensity along
-the genome. 
-We first describe procedures of summarizing recombinant trees from
-the first viewpoint.
-To summarize the posterior samples of ClonalOrigin we examined
-the distribution of topologies of local genealogies induced by the sampled 
-recombinant trees. Recall that a recombinant tree induces a local genealogy for 
-each site along the analyzed block. Figure \ref{fig:clonalorigin} shows six
-local gene trees including one that is the same as the species tree from the
-most left-hand side of the block of 100 base pairs to the right.
-We expected the dominant topology across 
-sampled genealogies to be concordant with that of the species tree. However, 
-other frequently sampled topologies were expected to be potentially informative about 
-common recombination events. There are 105 possible 
-rooted topologies relating five taxa, which allowed us to count topologies
-induced by a posterior sample of recombinant trees.
-Another useful summary suggested by 
-\citet{Didelot2010} was to classify recombinant edges by types defined according 
-to donor/recipient branch pairs, and record the number of edges sampled for
-each recombinant edge type. For instance, our species tree has nine branches
-including rooting branch and 
-81 branch pairs, out of which 60 are possible donor/recipient pairs 
-by prohibiting recombinant edges from going backward in time. The posterior 
-expected number of edges of each type can be compared to the prior expected 
-number determined by the three population parameters fixed in the second stage of 
-ClonalOrigin. This comparison was displayed in a heat map of 
-Figure \ref{fig:Heatmap-of-recombination}, which indicated branch pairs that 
-might experience an excess or depletion of recombination events across all
-blocks.
-
-For the second viewpoint of recombinant trees
-\citet{Didelot2010} suggested an approach used to trace
-recombination intensity along the genome. 
-For each edge type and each site, we recorded the sampling frequency of a 
-recombinant edge of the type at that site (e.g., Figure S\ref{fig:muts}).
-Figure \ref{fig:clonalorigin}B illustrates the posterior probability that a site
-experiences influx from a donor branch using a single recombinant tree.
-% This sampling frequency is an 
-% estimate of the posterior probability taking into consideration the 
-% sequence data and the setting chosen for the model parameters in the second 
-% stage of ClonalOrigin. 
-Summing these posteriors over a certain set of edge 
-types, we could obtain a measure of recombination intensity along the genome 
-(Figure \ref{fig:clonalorigin}B). 
-This allowed us to find genomic regions enriched for recombination events in general, 
-or regions enriched for recombination events of certain types (e.g.,
-between SPY and SDE).  We note that the recombination intensity measure is not 
-normalized to the interval $[0,1]$. Rather giving probabilistic meanings to the
-recombination intensity we interpreted it as an average number of recombinant
-edge types per site.
-The measure appeared to act as an 
-effective indicator of recombination intensity 
-(see section of \textbf{Simulation Study}). 
-Our simulation study indicated that edges of certain types were better indicators of 
-true recombination events than others. For instance, recombinant edges which 
-resulted in a recombinant tree with the same topology as the species tree appeared to 
-be more prone to spurious sampling. 
-Such edges were ones where the donor was the 
-same branch, sister branch, or parent branch of the recipient branch. 
-Therefore, in some cases we elected to remove such edges from consideration 
-when measuring recombination intensity.
-\texttt{IG: WE STILL NEED TO CHECK AND REFINE THIS IDEA. SC: WORKS ARE BEING DONE
-FOR THIS ISSUE OF REMOVING UNINFORMATIVE RECOMBINANT EDGES}.
-
-\subsection{Mowgli}
-While ClonalOrigin could be used to understand recombining gene
-transfers within the conserved core genome, we sought to understand
-the history of the remainder of the genome where more complex
-evolutionary processes occur, such as gene duplications, losses, and
-horizontal transfers.  In recent years, several methods have been
-developed for inferring these evolutionary events in gene families by
-reconciling gene trees with a species tree
-\citep{David2011,Doyon2011,Tofigh2011}.  In this analysis, we
-complemented our study of recombination by using Mowgli
-\citep{Doyon2011} to characterize events that can change the copy
-number of a gene, such as duplications, losses, and non-recombining
-horizontal gene transfers.  In contrast to ClonalOrigin, Mowgli
-assumes a parsimony model for inferring events and it uses
-pre-determined gene clusters where genes are treated as atomic units.
-%The advantage of using Mowgli was that it did not assume the sequence
-%data contains a single copy of each homologous segment, and it
-%attempted to explain different copy numbers by a series of gain and
-%loss events.  We here were interested in the approach of Mowgli to
-%study non-recombining gene transfers.
-
-We obtained gene family clusters defined in \citet{Suzuki2011} 
-using the OrthoMCL method \citep{Li2003} over the full set of six genomes
-including \emph{S.\ equi} subsp.\emph{\ equi} strain 4047 (SEE). 
-For each family cluster,
-the corresponding protein sequences were aligned using the MUSCLE
-program \citep{Edgar2004a} and nucleotide alignments were constructed
-by mapping nucleotide sequences onto each protein sequence alignment. We then
-identified recombination within gene alignments using the Single
-Breakpoint Recombination (SBP) method from the HyPhy package
-\citep{KosakovskyPond2006}.  Using the Akaike Information Criterion (AIC), SBP
-inferred that 47.5\% (1,094) of gene families contained at least one
-topology-changing recombination. Each of these families was then
-split into two gene fragments.
-Gene trees were constructed from nucleotide alignments for each
-gene fragment using the RAxML program \citep{Stamatakis2006}, and were
-rooted by minimizing inferred duplications and losses.  Each rooted
-gene tree was then reconciled using the Mowgli program
-\citep{Doyon2011}, which inferred the most parsimonious gene
-duplication, loss, and transfer events for each family.  In the case
-of multiple maximum parsimonious reconciliations, one was chosen
-uniformly at random.  Unless specified otherwise, default setings were used for
-all programs.
-Since the method of Mowgli did not directly model recombination events, 
-recombination-based transfers were inferred in a post-processing step, by 
-checking whether the transfer was coupled with a loss of a homologous gene in 
-the recipient branch (indicating gene replacement rather than gain). 
-A transfer event was classified as non-recombining if the 
-destination species of that transfer had an extant descendant that contained
-copies of that genes which were descendants of the transferred gene as well as 
-copies which were non-descendants.
-We noted that this parsimonious approach was imperfect and would tend to 
-undercount events, especially recombining gene transfer events, which were
-modeled using two events: a transfer and a loss. Nevertheless, this analysis 
-allowed us to study non-recombining gene transfers, which were not modeled by 
-ClonalOrigin, and to validate some of the ClonalOrigin results on recombining 
+\section{Results}
+
+\subsection{The five or six genomes}
+
+\subsection{The genome alignment of the five genomes}
+
+\subsection{The species tree inferred using the 274 alignment blocks}
+
+\subsection{The homologous recombination events}
+An alignment of the five genomes showed two distinct genome structures of SPY
+and SDE (Figure S\ref{fig:mauve}). The SDE genomes were larger in length than
+those of SPY. Although it was unclear that SDD was different from either of SDE1
+and SDE2 from the graphical view of the genome alignment, the spcies tree from
+the genome alignment showed that SDD branch was attached to the clade branch of
+SDE (Figure S\ref{fig:tree5}). The distribution of the alignment block lengths
+was skewed to the right with mean of 4188 base pairs long (Figure
+S\ref{fig:blocksize}), which implied that the model assumption of geometric
+length distribution might be plausible in \texttt{ClonalOrigin} method.  The
+relative strength of recombination compared to mutations indicated that
+reconmination-driven substitutions affected more sites than did mutation-driven
+substitutions (Table S\ref{tab:clonalframe}).
+
+\subsubsection{The three population parameters}
+Along the genome of SDE1 population mutation rate did not appear to change much
+(Figure \ref{fig:scatter3}).  We found a few spots of genomic regions where the
+mutation rates were larger than twice the global estimate of mutation rate.
+\texttt{WHAT ARE THESE BLOCKS? WHAT GENES ARE LOCATED IN THE BLOCK?}
+The recombination rate appeared to change more dynamically. Espeically, the
+genomic region of SDE1 near 20 kilobase pairs from the origin of replication
+contains alignment blocks with bell-shaped recombination rates.
+\texttt{USE UCSC GENOME BROWSER TO FIND WHAT GENES BELONG TO THOSE BLOCKS.}
+The recombinant tract length changes in a way that we could not note any
+particular positions along the genome.
+There appeared to be negative correlation between recombinant tract length and
+recombination rate.
+\texttt{PLOT PAIRWISE SCATTER PLOTS OF THE THREE PARAMETERS.}
+In the first stage of \texttt{ClonalOrigin} MCMC a few alignment blocks were not
+successfully finished even after a month of computation.  Because those were not
+expected to affect the global estimates of population parameters significantly,
+we used only finished results of the first stage of \texttt{ClonalOrigin} to
+estimate mutation rate, recombination rate, and recombinant tract length.  Two
+independent runs of the first stage MCMC resulted in similar estimates of the
+three population parameters (Table S\ref{tab:three}).
+
+\subsection{Tree topology changes along the genome}
+Most of local tree topologies, or 66.7\%, along the alignment blocks were the
+same as the species tree, which allowed us to be more confident of the species
+tree inferred using the alignment blocks (Table
+S\ref{tab:Gene-tree-topologies}).  The second most probable topology was one
+where SDD was the outgroup, and SPY and SDE were sister clades with probability
+of 9.3\%.  Therefore, SPY and SDE appeared to share about tenth of the core
+genomes that were more closely related with each other than each to SDD's
+genome. The other topologies had less than 4\% of proportions.
+
+\subsection{Evidence for unbalaned gene flow from SPY to SDE}
+The heat map in Figure \ref{fig:Heatmap-of-recombination} shows the log base 2
+ratios of these numbers and their corresponding expected values given the
+species tree and the global parameter settings.  Because the recombinant edge
+from any species tree branch to the root branch is not possible under
+\texttt{ClonalOrigin} model (see Figure \ref{fig:clonalorigin}), the column for
+the root as recipient was colored green. Likewise, all of the cells where a
+recipient branch was older than the donor, and two branches did not share any
+contemporary time were colored green. The diagonal cells were colored gray
+because the observed number of recombinant edges and the expected one were
+similar. Because recombinant edges starting from and ending in the same species
+tree branch would not change the species tree, which would not change the
+likelihood, the cells should be colored gray.  Because SDD is strictly
+veterinary hosts, we did not expect that SDD received much gene transfer from
+either SPY1 or SPY2.  The log base 2 ratio from SPY1 to SDD was -2.20, and that
+from SPY2 to SDD was -1.97. On the contrary, the recombing gene transfer from
+SDE1 to SDD was 0.756, and that from SDE2 to SDD was 0.698. The relatively
+higher recombining gene transfer from two SDE individuals to SDD compared with
+that from two SPY individuals to SDD reflected that the SDD was more closely
+related to SDE than to SPY.  The gene transfer from SDD to either SDE1 or SDE2
+was less frequent than the case of reverse direction.
+
+The recombining gene transfer log base 2 ratios between SDE1 and SDE2 were about
+twice as large as that from SDE1 and SDE2 to SPY1 and SPY2. Because the
+recombining gene transfer within a clade would be larger than that between
+clades, this was expected. Surprisingly, the recombining gene transfer log base
+2 ratios between SPY1 and SPY2 was comparable with that from SPY1 and SPY2 to
+SDE1 and SDE2.  The recombining gene transfer from a clade SPY1 or SPY2 to a
+different clade SDE1 or SDE2 was not different from that within the clade, or
+SPY1 and SPY2.  Consistent with this result, the gene transfer log base 2 ratio
+from SPY1 and SPY2 to SDE branch was larger than any type of gene transfer: 1.88
+and 1.94, respectively. This clearly showed that the recombining gene transfer
+was biased toward the net direction from SPY to SDE.
+
+% \texttt{IG: ADD FIGURE WITH NON-NORMALIZED VALUES? SC: Why?}
+
+% No evidence for unbalanced non-recombining gene transfer between SPY and SDE
+Figure S\ref{fig:hgt-heatmap} shows the number of inferred
+non-recombining gene transfer events between the different branches of
+the species tree.
+SDE1 and SDE2 shared the most events (26 and 34) compared to other sample pairs. 
+SDD appeared to acquire many genes from the ancestral SDE branch. Transfer 
+events between the SPY and SDE clades were inferred in smaller numbers (<6) 
+compared to those within each clade (23--34). Among the clade-crossing 
+transfers, we did not see any significant enrichment for a specific direction, 
+as opposed to the result for recombining transfers from the model-based 
+analysis (Figure S\ref{fig:mowgli-recomb-heatmap}).
+
+\subsection{Transport-related genes are associated with recombining gene
+transfer}
+
+\subsection{The virulence genes}
+\texttt{Find cases of any virulence genes that were transferred via recombining and
+horizontally.} 
+\texttt{I need to see how the virulence genes list was created.}
+\texttt{I want to know the virulence gene categories. Which kind of virulence
+genes are more under gene transfer?}
+\texttt{Make a list of virulence genes and their categories with more
+interpretations.}
+
+\subsection{Horizontal gene transfer is associated with virulence genes}
+
+\subsection{Comparison between model-based and parsimony-based approaches}
+\texttt{I might want to analyze using only core genomes and pan genomes
+separately. By separating data set of 6-gene families from the other families,
+we could have more complete comparison.}
+
+\subsection{UCSC genome browser for the dataset}
+We created a web-resource for gene transfer available at
+\url{http://strep-genome.bscb.cornell.edu}.  For each of the five individual
+genomes as a reference genome several annotation tracks were displayed:
+alignment blocks, recombination rate per site for each block, local gene tree
+topologies, recombination
+intensity, locations of virulence genes, posterior probability of recombining
 gene transfer.
 
-\subsection{Functional categories association}
-We performed a series of analyses to check for functional gene categories 
-enriched for certain types of gene transfer events. Functional categories were 
-assigned in the same manner as \citet{Suzuki2011}. In short, blastP was used to 
-compare Streptococcus genes to bacterial proteins from the Uniref90 database.  
-Matches with E-values less than $1.0\times10^{-5}$ were assumed to have the same Gene Ontology (GO) 
-classification as the target gene, as determined from the uniProt GOA database. 
-We also classified a category of  \emph{virulent genes} using the virulence 
-profile described in \citet{Suzuki2011}. As virulent genes we used 207 loci, in 
-which at least one of the genes in the gene family was identified by that 
-profile.
-Each gene covered by our analysis was associated with a recombination intensity 
-score from the ClonalOrigin analysis and scores from the Mowgli 
-analysis corresponding to number of duplications, losses, non-recombining 
-gene transfers, and homologous recombining gene transfer events. 
-\texttt{The recombination intensity scores were taken to be average site-wise intensity 
-scores. SC: THIS NEEDS TO BE REANALYZED WITH RECOMBINATION INTENSITY WITH ONLY
-TREE-TOPOLOGY CHANGING RECOMBINANT EDGES.} 
-Mann-Whitney tests were used to identify GO categories 
-whose genes showed significantly elevated values in each score. 
-A 5\% False Discovery Rate (FDR) correction \citep{Benjamini1995} was used to 
-determine an appropriate significance threshold for our p-values.
-
-\subsection{Simulation Study}
-We conducted experiments on simulation data to check the accuracy of 
-estimates obtained by ClonalOrigin in a scenario similar to ours. We used a 
-simulation feature of ClonalOrigin to simulate the process of mutation and 
-recombination across the five species.
-We started by testing the accuracy of the global estimates obtained from the 
-first stage of ClonalOrigin for the three population parameters.
-Ten recombinant trees were sampled given the tree parameters and the species
-tree  set to the values inferred from the 
-five bacterial genomes (see section \textbf{Results}).
-Each of the recombinant trees was used to simulate
-the Jukes-Cantor model for DNA sequence evolution to create 274 blocks of
-alignments. 
-The block lengths configuration was the same as that of the blocks 
-used in the analysis of the five genomes.
-We ran ClonalOrigin 
-on the ten replicate data sets, and obtained global estimates by applying the 
-weighted median scheme for the block-specific estimates. The means of the ten 
-global estimates of mutation rate per site and recombination rate per site were 
-0.081 with standard deviation (SD) of $0.0011$, and 0.011 with SD of $0.0003$, 
-respectively, centering
-around the true values: 0.081 and 0.012, respectively. The mean 
-of the ten estimates of the recombinant tract length was 919 bp (SD 28), which 
-was significantly higher than the true value of 744 bp. This was not entirely 
-surprising as recombinant tract lengths were known to be difficult to 
-estimate from relatively short alignment blocks \citep{Didelot2010}. When we 
-performed similar simulations which much longer blocks (10000 bp), 
-ClonalOrigin was able to estimate the average recombinant tract length 
-accurately (data not shown).
-
-% \texttt{IG: RESULTS FROM THIS POINT ON REQUIRE SOME REFINEMENT. SC: Let's see.}
-Next we evaluated the accuracy of statistics estimated in 
-the second stage of ClonalOrigin. We used 100 replicates of data sets obtained
-by first
-sampling ten recombinant trees given the mutation rate, recombination rate, and
-tract length, and generating ten independent 
-simulations of sequence evolution using each of the recombinant trees. For each measure 
-of interest we thus obtained 100 separate estimates, which we could compare 
-with either
-the expected value from the ten recombinant trees, or 
-true value from one of the recombinant trees. 
-First, we tested the number of 
-recombinant edges of each of the sixty different types. Figure S\ref{fig:h1} 
-plots  the distribution of observed counts against the 
-respective expected count given the model parameters and species tree 
-for each recombinant edge type.
-\texttt{FIX ``TRUE" TO BE EXPECTED VALUE RATHER THAN AVERAGE OVER THE 10 REC. 
-TREES. SC: OKAY. LET'S FIND THE PRIOR EXPECTED NUMBER OF EVENTS. THIS MUST BE
-GIVEN SOMEWHERE}. 
-The estimated numbers appeared to be slightly over-estimated, but in 
-very good correlation with the expected numbers.
-
-\texttt{CONSIDER SHOWING DETAILED COMPARISON OF ESTIMATED VALUES TO ONES 
-EXPECTED FROM THE ACTUAL REC. TREES. THESE DON'T LOOK AS NICE, BUT THEY MIGHT 
-BE ABLE TO DEMONSTRATE OUR ABILITY TO DETECT DEPARTURE FROM EXPECTATION. SC:
-THIS MAY BE ABOUT CHANGING AVERAGE TO PRIOR EXPECTED NUMBER OF EVENTS.}
-
-\texttt{WE SHOULD PROBABLY CONSIDER ALTERNATIVE WAYS OF USING THESE RESULTS TO 
-ARGUE THAT OUR DIRECTIONAL SIGNAL IS REAL. SC: I MIGHT CONSIDER SIMULATIONS AND
-ARTIFICIALLY DOUBLE OR TRIPLE THE RECOMBINATION EVENTS OF ONE OF TYPES.}
-
-With the slight over-estimation of recombination events we were concerned with 
-any potential biases hidden in ClonalOrigin implementation. Because the 
-simulated data were generated with no directionality between SPY and SDE,  if 
-we estimate recombination events from the simulated data sets using 
-ClonalOrigin and compare those estimates from those from the real data set, 
-this could remove biases in ClonalOrigin implementation. We could use inferred 
-estimates of recombination from the simulated data under the prior model of 
-ClonalOrigin as the prior expected number of recombination. Figure 
-S\ref{fig:h2} shows a plot of numbers of recombination events from the real 
-data against those from the simulated data sets. Note that the numbers of 
-recombination events from simulated data sets appear to deviate from the line 
-of slope being 1 and intercept being 0. The numbers of recombination events for 
-the recombinant edges from SPY to SDE increase by more than 2.5 fold whereas 
-those for the recombinant edges from SDE to SPY increase by about 1.7 fold.  
-This was not surprising because inference of number of recombinant edges was 
-quite accurate. We considered that subtraction of effects of number of 
-recombinant edges inferred from simulated data and subtraction of effects of 
-the prior expected number of recombinant edges were effectively equivalent. 
-However, this comparison between inferrence from simulated data and that of 
-real data set suggested that the directional difference between SPY and SDE was 
-due to neither lack of power of ClonalOrigin nor hidden bias thereof.
-
-%\textbf{Simulation study for checking the surrogate measure of recombination intensity:} 
-We tested if our measure of recombination intensity provided a reliable 
-estimate of the true intensity. For this purpose, we artificially annotaded our 
-synthetic blocks to genes according to the real annotations along the real 
-bacterial alignments. For each gene in each of the 100 data sets, we plotted 
-the average estimated recombination intensity along that gene against the true 
-average intensity, i.e., the average number of different recombinant edge types 
-implied by the recombinant tree used in simulation (Figure S\ref{fig:ri1}). 
-Although the scatter plot did not strictly align along the $X=Y$ line, there 
-was clear correlation of the estimated and true values (Pearson's correlation 
-coefficient of 0.63). Estimated values appeared to be driven somewhat by the 
-prior implied by the population parameter setting, explaining why lower values 
-were over-estimated and higher values were under-estimated. Nonetheless, the 
-correlation indicated that the measure of recombination intensity used in our 
-functional gene category served as an adequate indicator of actual intensity of 
-recombination events along genes.
-
-\clearpage{}
-
-\section{Results}
-
-\subsection{ClonalOrigin analysis}
-%\textbf{Genome sequence alignments and clonal frame:}
-The species tree of the five genomes is shown in Figure S\ref{fig:tree5}, which
-was estimated using 274 blocks of the core genome alignments with ClonalFrame. 
-Figure S\ref{fig:blocksize} shows the distribution of block lengths. The average
-length was 4188 bp.  There was good 
-agreement on the topology among independent chains, and reasonable agreement on 
-branch lengths. The tree was also in good agreement with the one used by 
-\citet{Suzuki2011}. ClonalFrame estimated the relative strength of recombination
-compared to mutations to be $r/m = 4.66$, indicating 
-that reconmination-driven substitutions affected more 
-sites than did mutation-driven substitutions. The first stage of ClonalOrigin 
-was then applied to obtain block-wise estimates of mutation rate, 
-recombination rate, and average recombinant tract length 
-(Figure \ref{fig:scatter3}).
-In the first stage of ClonalOrigin MCMC a few alignment blocks were not
-successfully finished even after a month of computation. 
-Because those were not expected
-to affect the global estimates of population parameters significantly, we used
-only finished results of the first stage of ClonalOrigin to estimate mutation
-rate, recombination rate, and recombinant tract length. 
-The median of mutation
-rate per site weighted over the lengths of blocks was estimated to
-be 0.081 with (0.067,0.094) of interquartile range (IQR). Mutation rates
-per site for many alignment blocks were almost centered around the
-gobal median. The median of recombination rate per site was estimated
-to be 0.012 with (0.006,0.019) of IQR.
-The median value for the recombinant tract length was estimated
-to be 744 base pairs with (346,2848) of IQR. 
-The estimates of the three parameters were based on one of two independent
-MCMC chains. 
-The estimates from the second MCMC run were close to those of
-the first run: mutation rate estimate of 0.081 with IQR of (0.067,0.099),
-recombination rate of 0.012 with IQR of (0.006,0.020), and recombinant tract
-length estimate of 723 with IQR of (348,2870).
-We considered the similar 
-results of the two independent chains as confirmation of 
-the convergence of the first stage of ClonalOrigin MCMC.
-
-% \texttt{THIS FIGURE OF RECOMBINATION BOUNDARIES NEEDS TO BE REPLACED OR REMOVED
-% BECAUSE I DO NOT FIND REASONS OF ADDING THIS TO THE MANUSCRIPT.}
-% For further validation of the 
-% estimated recombnation rates, we recorded the number of recombinant edges 
-% ClonalOrigin samples that are entirely contained within a block; 
-% this number 
-% should correlate well with the block-specific recombination rates. Figure 
-% \ref{fig:recombination-boundaries} shows the numbers obtained in our MCMC runs, 
-% and indeed these numbers appear to be well correlated with the block-specific 
-% recombination rates.
-
-%\textbf{Marginal evidence for unbalaned gene flow from SPY to SDE:}
-We then proceeded with the second stage of ClonalOrigin, setting the population 
-parameters to their global estimates. 
-Using the posterior sample of recombinant trees, we counted 
-rooted topologies of local genealogies along the blocks of alignments. 
-The top eight 
-topologies (out of 105 possible topologies) are listed in Table 
-S\ref{tab:Gene-tree-topologies}. The species tree topology obtained a posterior 
-probability of 66.7\%. The second most probable topology was one where SDD was 
-the outgroup, and SPY and SDE were sister clades with probability of 9.3\%.
-The other topologies had probability less than 4\%. We also recored the 
-number of recombinant edges sampled by ClonalOrigin of each of the sixty 
-possible types determined by donor/recepient species tree branches. The heat 
-map in Figure \ref{fig:Heatmap-of-recombination} shows the log ratios of 
-these numbers and their corresponding expected values given the species tree and 
-the global parameter settings. 
-% \texttt{IG: ADD FIGURE WITH NON-NORMALIZED VALUES? SC: Why?} 
-The recombinant edge types that showed the highest enrichment in the 
-sample were the ones from SPY1 and SPY2 branches to any branch within the SDE 
-clade, with ratios in $[2.46,2.58]$. 
-\texttt{IG: REPORT RATIOS TO SDE BRANCH? SC: List all of the values.} 
-For comparison, the ratios for edges in the other directions were all smaller 
-than 1.6. This appeared to indicate that gene flow driven by homologous 
-recombination from SPY to SDE was stronger than gene flow in the reverse 
-direction.
-
-\subsection{Mowgli analysis}
-%\textbf{Gene gain, loss, duplication, and non-recombining horizontal gene transfer:} 
-For the Mowgli analysis we obtained 2314 gene families (Figure
-S\ref{fig:famsize}), the largest of which consisted of 59 genes, while
+\subsection{Parsimony-based approach for gene gain, loss, and duplication}
+% Gene gain, loss, duplication, and non-recombining horizontal gene transfer
+Among the 2314 gene families
+the largest family consisted of 59 genes, while
 nearly half of gene families (1066) had exactly one gene per species
-(six genes total).  Figure \ref{fig:Gene-duplication-loss} shows the
+, or six genes total (Figure S\ref{fig:famsize}).
+Figure \ref{fig:Gene-duplication-loss} shows the
 inferred gene counts for each extant and ancestral species. Overall,
@@
 SDE were the most expanded, both in terms of duplications and in-bound
-transfers. Figure S\ref{fig:hgt-heatmap} shows the number of inferred
-non-recombining gene transfer events between the different branches of
-the species tree.
-%\textbf{No evidence for unbalanced non-recombining gene transfer between SPY and SDE:}
-SDE1 and SDE2 shared the most events (26 and 34) compared to other sample pairs. 
-SDD appeared to acquire many genes from the ancestral SDE branch. Transfer 
-events between the SPY and SDE clades were inferred in smaller numbers (<6) 
-compared to those within each clade (23--34). Among the clade-crossing 
-transfers, we did not see any significant enrichment for a specific direction, 
-as opposed to the result for recombining transfers from the ClonalOrigin 
-analysis (Figure S\ref{fig:mowgli-recomb-heatmap}).
+transfers.
 
 \subsection{Functional categories associated with gene transfer}
-
-\texttt{IG: THIS SECTION REQUIRES ADDITIONAL WORK, PENDING INTERPRETATION AND 
-REFINEMENT OF RESULTS. }
-
-\texttt{IG: THESE RESULTS MIGHT SLIGHTLY CHANGE BY REMOVING CERTAIN EDGE TYPES 
-FROM CONSIDERATION. SC: ANALYSES WILL BE DONE BY REMOVING SOME EDGES.}
 
@@
 Table \ref{tab:functional} lists the functional categories that showed
-significant associated with recombination intensity obtained from our 
-ClonalOrigin analysis.
+significantly associated with recombination intensity. 
 Several functional categories are related with ribosomes.
@@
 Table \ref{tab:go-events} lists the functional gene categories that
-showed significant associated with gene duplications, losses,
-non-recombining and recombining gene transfer according to the Mowgli
-analysis.  For the recombining transfers, Mowgli shows enrichment in
-many of the same categories as ClonalOrigin, such as translation,
+showed significantly associated with gene duplications, losses,
+non-recombining and recombining gene transfer using the pan genome of the six
+\textit{Streptococcus} individuals.
+
+For the recombining transfers, Mowgli shows enrichment in
+many of the same categories as \texttt{ClonalOrigin}, such as translation,
 ribosome, and rRNA binding.  This agreement between the two different
@@
 validation for the signal picked up by these separate association
-studies.  However, for duplication, losses, and non-recombining
+studies.
+
+However, for duplication, losses, and non-recombining
 transfers, we find enrichment in very different categories related to
@@
 Gene losses also showed enrichment for transposition and integration,
-but less significantly ($P<10^-6$) and also showed enrichment for
+but less significantly ($P<10^{-6}$) and also showed enrichment for
 several additional categories such as transporter activity and
@@
 \section{Discussions}
-Closely related bacterial species were investigated here in our study to better
+Closely related bacterial species were investigated here to better
 understand gene transfer processes. Two complementary methods were employed to
@@
 of gene transfer between SPY and SDE. Internal fragments of the seven
-housekeeping genes, encoding putative glucose kinase (\emph{gki}),
-glutamine transport protein (\emph{gtr}), glutamate racemase (\emph{murI}),
-mismatch repair enzyme (\emph{mutS}), transketolase (\emph{recP}),
-xanthine phosphoribosyltransferase (\emph{xpt}), and acetylcoenzyme
-A acetyltransferase (\emph{yqiL}) were located in the SPY1 genome that we
-investigated in this study.  Six of the seven genes except \emph{yqiL
-}gene were in the core genomes ClonalOrigin analyzed. Genes \emph{mutS},
-\emph{recP}, and \emph{xpt} had fragments of gene transfer with moderate
+housekeeping genes, encoding putative glucose kinase (\textit{gki}),
+glutamine transport protein (\textit{gtr}), glutamate racemase (\textit{murI}),
+mismatch repair enzyme (\textit{mutS}), transketolase (\textit{recP}),
+xanthine phosphoribosyltransferase (\textit{xpt}), and acetylcoenzyme
+A acetyltransferase (\textit{yqiL}) were located in the SPY1 genome that we
+investigated in this study.  Six of the seven genes except \textit{yqiL
+}gene were in the core genomes \texttt{ClonalOrigin} analyzed. Genes \textit{mutS},
+\textit{recP}, and \textit{xpt} had fragments of gene transfer with moderate
 posterior probability of recombination. The other genes did not have
-noticeable recombination. Except for \emph{mutS} where \citet{Kalia2001}
+noticeable recombination. Except for \textit{mutS} where \citet{Kalia2001}
 did not find recombination, the result was consistent with that of
 \citet{Kalia2001}. 
-We also revisited recombination in gene \emph{parC} in SPY1 genome
+We also revisited recombination in gene \textit{parC} in SPY1 genome
 studied by \citet{Pinho2010}, where they studied fluoroquinolone
@@
 \textbf{Mechanism of unbalanced gene flow between SPY and SDE:} 
-If there were unbalaned gene transfer between the two
+If there were unbalanced gene transfer between the two
 species, what were the barriers \citep{Thomas2005} to gene transfer
@@
 of RMS might be to maintain and control species identity \citep{Jeltsch2003}.
-Recently, \citet{Budroni2011a} studied that \emph{Neisseria meningitidis}
+Recently, \citet{Budroni2011a} studied that \textit{Neisseria meningitidis}
 phylogenetic clades were associated with RMS, claiming that RMS modulated
-homologous recombination in the bacteria of \emph{N.\ meningitidis}.
-Most of \emph{S.\ pyogenes} genomes contain restriction-modification
+homologous recombination in the bacteria of \textit{N.\ meningitidis}.
+Most of \textit{S.\ pyogenes} genomes contain restriction-modification
 type I system as well as type II system. The two genomes of SDE contained
@@
 The method of detecting homologous recombination gene transfer, or
-the ClonalOrigin, is model-based approach as opposed to the other
+the \texttt{ClonalOrigin}, is model-based approach as opposed to the other
 method for non-recombining gene transfer, or the Mowgli.
@@
 long) sequence alignments with almost no gaps. With this inherent
-filtering of the input data preparation ClonalOrigin did not account
+filtering of the input data preparation \texttt{ClonalOrigin} did not account
 for gene gain and loss in its core genome analysis. On the contrary,
@@
 
-\section*{Supplementary Material}
-
-Supplementary figures S1\textendash{}S8, and tables S1 and S2
-are available at Genome Biology
-and Evolution online (\url{http://www.gbe.oxfordjournals.org/}).
+\section{Materials and Methods}
+
+\subsection{Complete bacterial genomes}
+Two genomes of \textit{S.\ pyogenes} (SPY), and two of \textit{S.\ dysgalactiae}
+ssp.\textit{\ equisimilis} (SDE) were downloaded from NCBI with the following
+accessions: NC\_004070 {Beres2002}, NC\_008024 \citep{Beres2006}, NC\_012891
+\citep{Shimomura2011}, and CP002215 \citep{Suzuki2011}. We refer to these
+samples as SPY1, SPY2, SDE1, and SDE2, respectively. One genome of 
+\textit{S.\ dysgalactiae} ssp.\textit{\ dysgalactiae} was provided by 
+\citet{Suzuki2011}
+with accession identifiers of SddyATCC27957, referred to here as SDD. We also
+downloaded the genome (NC\_012471) of \textit{S.\ equi} ssp.\textit{\ equi}
+strain 4047 referred to as SEE.
+
+\subsection{Model-based method for recombining gene transfer}
+We present a short summary of the statistical method developed by
+\citet{Didelot2007} and \citet{Didelot2010} for studying bacterial
+recombination. A detailed account was found in the two papers and in the user
+guide available at \url{http://code.google.com/p/clonalorigin}.
+
+We used \texttt{progressiveMauve} v2.3.1 \citep{Darling2004,Darling2010} with
+default options to attempt to align the six genomes, which led to too small
+proportion of a core genome. We had to remove the SEE sample from the genome
+alignment. Note that we kept it in the parsimony-based approach.  We then used
+\texttt{stripSubsetLCBs}, a program of the \texttt{ClonalOrigin} package, and
+identified 276 blocks of sequence alignments with length threshold of 1500 base
+pairs.  We excluded two of these blocks because they contained long consecutive
+gaps.
+
+Using 274 blocks, we applied \texttt{ClonalFrame} v1.1 available at
+\url{http://www.xavierdidelot.xtreemhost.com}, and obtained an estimate of the
+species tree relating the five genomes. ClonalFrame also outputs estimates of
+parameters including the average recombinant tract length and the ratio between
+recombination and mutation rates, however, these were not used in the subsequent
+analysis. We used a single Markov chain Monte Carlo (MCMC) with 10000 burn-in
+generations and 10000 sampling generations to obtain the tree estimate with
+branch lengths, and additionally seven independent MCMC runs to validate
+adequate convergence of the first.
+
+Two stages of MCMC with \texttt{ClonalOrigin}, subversion r19 that had been
+downloaded from \url{http://clonalorigin.googlecode.com/svn/trunk}, were
+employed independently for each of the 274 alignment blocks as described in
+\citet{Didelot2010}.   In the first MCMC we obtained estimates of mutation rate,
+recombination rate, and recombinant tract length.  Two independent chains were
+executed with $10^6$ burn-in generations and $10^7$ sampling generations,
+sub-sampled every $10^5$-th generations. One of the two chains was used to
+obtain the block-wise mean estimates, and the other chain was used to validate
+convergence of the first stage of MCMC.  The estimators of the three population
+parameters were medians weighted over block lengths, which down-weighed less
+reliable estimates obtained from shorter alignment blocks.  The second stage of
+MCMC with \texttt{ClonalOrigin} was executed with the three population
+parameters fixed to the estimated values in the first stage.  Two independent
+chains of the second stage of \texttt{ClonalOrigin}'s MCMC were also executed
+with a total generation being $1.1\times10^8$. The first $10^7$ generations of
+each chain were discarded as burn-in. The remaining $10^8$ generations were
+sub-sampled every $10^5$-th generation.
+
+\subsection{Local gene trees and recombination intensity}
+Because the procedure of summarizing posterior sample of \textit{recombinant
+tree} was not detailed in the user guide at the time when we used
+\texttt{ClonalOrigin}, we elected to use Figure \ref{fig:clonalorigin} to
+illustrate a recombinant tree in order to clarify summarization procedures we
+employed of posterior samples of recombinant trees.  A recombination history is
+represented by a \textit{recombinant tree}, which is a species tree augmented
+with a set of \textit{recombinant edges}.  Each recombinant edge is defined by
+time and branch of origin, time and branch of destination, and genomic interval
+that it affects. The branch of origin is referred to as the \textit{donor} and
+the destination branch is referred to as the \textit{recipient}. A recombinant
+tree can be viewed as a restricted version of an ancestral recombination graph
+(ARG), and as such it induces a local genealogy for each site in the genomic
+block it describes.  \citet{Didelot2010} presented two procedures of summarizing
+recombinant trees by viewing trees from two different angles: one from
+recombinant trees by aggregating all of the sites, and another from site-wise
+recombinant edge changes to the species tree.  By following their approaches we
+quantified more aspects of recombinant trees: from the view of recombinant trees
+with aggregated sites we counted gene tree topologies, and from the view of
+site-wise recombinant edge changes we quantified recombination intensity along
+the genome.
+
+We first describe procedures of summarizing recombinant trees from the first
+viewpoint.  To summarize the posterior samples of \texttt{ClonalOrigin} we
+examined the distribution of topologies of local genealogies induced by the
+sampled recombinant trees. Recall that a recombinant tree induces a local
+genealogy for each site along the analyzed block.  Figure \ref{fig:clonalorigin}
+shows six local gene trees including one that is the same as the species tree
+from the most left-hand side of the block of 100 base pairs to the right.  We
+expected the dominant topology across sampled genealogies to be concordant with
+that of the species tree. However, other frequently sampled topologies were
+expected to be potentially informative about common recombination events. There
+are 105 possible rooted topologies relating five taxa, which allowed us to count
+topologies induced by a posterior sample of recombinant trees.
+
+For the second viewpoint of recombinant trees \citet{Didelot2010} suggested an
+approach used to trace recombination intensity along the genome.  For each edge
+type and each site, we recorded the sampling frequency of a recombinant edge of
+the type at that site (e.g., Figure S\ref{fig:muts}).  Figure
+\ref{fig:clonalorigin}B illustrates the posterior probability that a site
+experiences influx from a donor branch using a single recombinant tree.  Summing
+these posteriors over a certain set of edge types, we could obtain a measure of
+recombination intensity along the genome (Figure \ref{fig:clonalorigin}B).  This
+allowed us to find genomic regions enriched for recombination events in general,
+or regions enriched for recombination events of certain types (e.g., between SPY
+and SDE).
+
+We note that the recombination intensity is not normalized to
+the interval $[0,1]$. Rather giving probabilistic meanings to the recombination
+intensity we interpreted it as an average number of recombinant edge types per
+site.  The measure appeared to act as an effective indicator of recombination
+intensity (see subsection of \textbf{Simulation Study}).  Our simulation study
+indicated that edges of certain types were better indicators of true
+recombination events than others. For instance, recombinant edges which resulted
+in a recombinant tree with the same topology as the species tree appeared to be
+more prone to spurious sampling.  Such edges were ones where the donor was the
+same branch, sister branch, or parent branch of the recipient branch.
+Therefore, in some cases we elected to remove such edges from consideration when
+measuring recombination intensity.
+
+\subsection{Parsimony-based method for horizontal gene transfer}
+While \texttt{ClonalOrigin} could be used to understand recombining gene
+transfers within the conserved core genome, we sought to understand the history
+of the remainder of the genome where more complex evolutionary processes occur,
+such as gene duplications, losses, and horizontal transfers.  In recent years,
+several methods have been developed for inferring these evolutionary events in
+gene families by reconciling gene trees with a species tree
+\citep{David2011,Doyon2011,Tofigh2011}.  In this analysis, we complemented our
+study of recombination by using Mowgli \citep{Doyon2011} to characterize events
+that can change the copy number of a gene, such as duplications, losses, and
+non-recombining horizontal gene transfers.  In contrast to
+\texttt{ClonalOrigin}, Mowgli assumes a parsimony model for inferring events and
+it uses pre-determined gene clusters where genes are treated as atomic units.
+%The advantage of using Mowgli was that it did not assume the sequence
+%data contains a single copy of each homologous segment, and it
+%attempted to explain different copy numbers by a series of gain and
+%loss events.  We here were interested in the approach of Mowgli to
+%study non-recombining gene transfers.
+We obtained gene family clusters defined in \citet{Suzuki2011} using the
+OrthoMCL method \citep{Li2003} over the full set of six genomes including
+\textit{S.\ equi} subsp.\textit{\ equi} strain 4047 (SEE).  For each family
+cluster, the corresponding protein sequences were aligned using the MUSCLE
+program \citep{Edgar2004a} and nucleotide alignments were constructed by mapping
+nucleotide sequences onto each protein sequence alignment. We then identified
+recombination within gene alignments using the Single Breakpoint Recombination
+(SBP) method from the HyPhy package \citep{KosakovskyPond2006}.  Using the
+Akaike Information Criterion (AIC), SBP inferred that 47.5\% (1,094) of gene
+families contained at least one topology-changing recombination. Each of these
+families was then split into two gene fragments.  Gene trees were constructed
+from nucleotide alignments for each gene fragment using the RAxML program
+\citep{Stamatakis2006}, and were rooted by minimizing inferred duplications and
+losses.  Each rooted gene tree was then reconciled using the Mowgli program
+\citep{Doyon2011}, which inferred the most parsimonious gene duplication, loss,
+and transfer events for each family.  In the case of multiple maximum
+parsimonious reconciliations, one was chosen uniformly at random.  Unless
+specified otherwise, default setings were used for all programs.  Since the
+method of Mowgli did not directly model recombination events,
+recombination-based transfers were inferred in a post-processing step, by
+checking whether the transfer was coupled with a loss of a homologous gene in
+the recipient branch (indicating gene replacement rather than gain).  A transfer
+event was classified as non-recombining if the destination species of that
+transfer had an extant descendant that contained copies of that genes which were
+descendants of the transferred gene as well as copies which were
+non-descendants.  We noted that this parsimonious approach was imperfect and
+would tend to undercount events, especially recombining gene transfer events,
+which were modeled using two events: a transfer and a loss. Nevertheless, this
+analysis allowed us to study non-recombining gene transfers, which were not
+modeled by \texttt{ClonalOrigin}, and to validate some of the
+\texttt{ClonalOrigin} results on recombining gene transfer.
+
+\subsection{Functional categories association}
+We performed a series of analyses to check for functional gene categories
+enriched for certain types of gene transfer events. Functional categories were
+assigned in the same manner as \citet{Suzuki2011}. In short, blastP was used to
+compare Streptococcus genes to bacterial proteins from the Uniref90 database.
+Matches with E-values less than $1.0\times10^{-5}$ were assumed to have the same
+Gene Ontology (GO) classification as the target gene, as determined from the
+uniProt GOA database.  We also classified a category of  \textit{virulent genes}
+using the virulence profile described in \citet{Suzuki2011}. As virulent genes
+we used 207 loci, in which at least one of the genes in the gene family was
+identified by that profile.  Each gene covered by our analysis was associated
+with a recombination intensity score from the \texttt{ClonalOrigin} analysis and
+scores from the Mowgli analysis corresponding to number of duplications, losses,
+non-recombining gene transfers, and homologous recombining gene transfer events.
+\texttt{The recombination intensity scores were taken to be average site-wise
+intensity scores. SC: THIS NEEDS TO BE REANALYZED WITH RECOMBINATION INTENSITY
+WITH ONLY TREE-TOPOLOGY CHANGING RECOMBINANT EDGES.} Mann-Whitney tests were
+used to identify GO categories whose genes showed significantly elevated values
+in each score.  A 5\% False Discovery Rate (FDR) correction
+\citep{Benjamini1995} was used to determine an appropriate significance
+threshold for our p-values.
+
+\subsection{Simulation Study}
+We conducted experiments on simulation data to check the accuracy of estimates
+obtained by \texttt{ClonalOrigin} in a scenario similar to ours. We used a
+simulation feature of \texttt{ClonalOrigin} to simulate the process of mutation
+and recombination across the five species.  While populaiton mutation and
+recombination parameters were accurately estimated, the recombinant
+tract length was over-estimated  (Table
+S\ref{tab:sim-three-population-parameters}).  This was not entirely surprising
+as recombinant tract lengths were known to be difficult to estimate from
+relatively short alignment blocks \citep{Didelot2010}. When we performed similar
+simulations which much longer blocks (10000 bp), \texttt{ClonalOrigin} was able
+to estimate the average recombinant tract length accurately (data not shown).
+
+We also evaluated the accuracy of statistics estimated in the second stage of
+\texttt{ClonalOrigin}. The number of recombinant edges for each pair of recombinant
+tree branches was correlated with the true value in general 
+while there were slight over-estimation (Figure S\ref{fig:h1}). 
+However, this comparison between inferrence from simulated data and that of 
+real data set suggested that the directional difference between SPY and SDE was 
+due to neither lack of power of \texttt{ClonalOrigin} nor hidden bias thereof.
+
+Recombination intensities that were measured on genes also showed rough
+correlation (Pearson's correlation coefficient of 0.63) of the estimated and
+true values (Figure S\ref{fig:ri1}).  Estimated values appeared to be driven
+somewhat by the prior implied by the population parameter setting, explaining
+why lower values were over-estimated and higher values were under-estimated.
+Nonetheless, the correlation indicated that the measure of recombination
+intensity used in our functional gene category served as an adequate indicator
+of actual intensity of recombination events along genes.
+
+\clearpage{}
+
+\section*{Supplementary Information}
+
+\subsection*{Simulation for the first stage MCMC of ClonalOrigin.}
+We started by testing the accuracy
+of the global estimates obtained from the first stage of \texttt{ClonalOrigin}
+for the three population parameters.  Ten recombinant trees were sampled given
+the three parameters and the species tree  set to the values inferred from the
+five bacterial genomes (see section \textbf{Results}).  Each of the recombinant
+trees was used to simulate the Jukes-Cantor model for DNA sequence evolution to
+create 274 blocks of alignments.  The block lengths configuration was the same
+as that of the blocks used in the analysis of the five genomes.  We ran
+\texttt{ClonalOrigin} on the ten replicate data sets, and obtained global
+estimates by applying the weighted median scheme for the block-specific
+estimates. The means of the ten global estimates of mutation rate per site and
+recombination rate per site were 0.081 with standard deviation (SD) of $0.0011$,
+and 0.011 with SD of $0.0003$, respectively, centering around the true values:
+0.081 and 0.012, respectively. The mean of the ten estimates of the recombinant
+tract length was 919 bp (SD 28), which was significantly higher than the true
+value of 744 bp. This was not entirely surprising as recombinant tract lengths
+were known to be difficult to estimate from relatively short alignment blocks
+\citep{Didelot2010}. When we performed similar simulations which much longer
+blocks (10000 bp), \texttt{ClonalOrigin} was able to estimate the average
+recombinant tract length accurately (data not shown).
+
+Table S\ref{tab:sim-three-population-parameters}.~
+Recombinant tract length was over-estimated.
+
+\subsection*{Simulation study for the second stage MCMC of ClonalOrigin}
+% \texttt{IG: RESULTS FROM THIS POINT ON REQUIRE SOME REFINEMENT. SC: Let's see.}
+Next we evaluated the accuracy of statistics estimated in the second stage of
+\texttt{ClonalOrigin}. We used 100 replicates of data sets obtained by first
+sampling ten recombinant trees given the mutation rate, recombination rate, and
+tract length, and generating ten independent simulations of sequence evolution
+using each of the recombinant trees. For each measure of interest we thus
+obtained 100 separate estimates, which we could compare with either the expected
+value from the ten recombinant trees, or true value from one of the recombinant
+trees.  First, we tested the number of recombinant edges of each of the sixty
+different types. Figure S\ref{fig:h1} plots  the distribution of observed counts
+against the respective expected count given the model parameters and species
+tree for each recombinant edge type.  \texttt{FIX ``TRUE" TO BE EXPECTED VALUE
+RATHER THAN AVERAGE OVER THE 10 REC.  TREES. SC: OKAY. LET'S FIND THE PRIOR
+EXPECTED NUMBER OF EVENTS. THIS MUST BE GIVEN SOMEWHERE}.  The estimated numbers
+appeared to be slightly over-estimated, but in very good correlation with the
+expected numbers.
+
+\texttt{CONSIDER SHOWING DETAILED COMPARISON OF ESTIMATED VALUES TO ONES 
+EXPECTED FROM THE ACTUAL REC. TREES. THESE DON'T LOOK AS NICE, BUT THEY MIGHT 
+BE ABLE TO DEMONSTRATE OUR ABILITY TO DETECT DEPARTURE FROM EXPECTATION. SC:
+THIS MAY BE ABOUT CHANGING AVERAGE TO PRIOR EXPECTED NUMBER OF EVENTS.}
+
+\texttt{WE SHOULD PROBABLY CONSIDER ALTERNATIVE WAYS OF USING THESE RESULTS TO 
+ARGUE THAT OUR DIRECTIONAL SIGNAL IS REAL. SC: I MIGHT CONSIDER SIMULATIONS AND
+ARTIFICIALLY DOUBLE OR TRIPLE THE RECOMBINATION EVENTS OF ONE OF TYPES.}
+
+With the slight over-estimation of recombination events we were concerned with 
+any potential biases hidden in \texttt{ClonalOrigin} implementation. Because the 
+simulated data were generated with no directionality between SPY and SDE,  if 
+we estimate recombination events from the simulated data sets using 
+\texttt{ClonalOrigin} and compare those estimates from those from the real data set, 
+this could remove biases in \texttt{ClonalOrigin} implementation. We could use inferred 
+estimates of recombination from the simulated data under the prior model of 
+\texttt{ClonalOrigin} as the prior expected number of recombination. Figure 
+S\ref{fig:h2} shows a plot of numbers of recombination events from the real 
+data against those from the simulated data sets. Note that the numbers of 
+recombination events from simulated data sets appear to deviate from the line 
+of slope being 1 and intercept being 0. The numbers of recombination events for 
+the recombinant edges from SPY to SDE increase by more than 2.5 fold whereas 
+those for the recombinant edges from SDE to SPY increase by about 1.7 fold.  
+This was not surprising because inference of number of recombinant edges was 
+quite accurate. We considered that subtraction of effects of number of 
+recombinant edges inferred from simulated data and subtraction of effects of 
+the prior expected number of recombinant edges were effectively equivalent. 
+However, this comparison between inferrence from simulated data and that of 
+real data set suggested that the directional difference between SPY and SDE was 
+due to neither lack of power of \texttt{ClonalOrigin} nor hidden bias thereof.
+
+We tested if our measure of recombination intensity provided a reliable 
+estimate of the true intensity. For this purpose, we artificially annotaded our 
+synthetic blocks to genes according to the real annotations along the real 
+bacterial alignments. For each gene in each of the 100 data sets, we plotted 
+the average estimated recombination intensity along that gene against the true 
+average intensity, i.e., the average number of different recombinant edge types 
+implied by the recombinant tree used in simulation (Figure S\ref{fig:ri1}). 
+Although the scatter plot did not strictly align along the $X=Y$ line, there 
+was clear correlation of the estimated and true values (Pearson's correlation 
+coefficient of 0.63). Estimated values appeared to be driven somewhat by the 
+prior implied by the population parameter setting, explaining why lower values 
+were over-estimated and higher values were under-estimated. Nonetheless, the 
+correlation indicated that the measure of recombination intensity used in our 
+functional gene category served as an adequate indicator of actual intensity of 
+recombination events along genes.
 
@@
 
-We are indebted to Xavier Didelot for his help in setting up ClonalOrigin
+We are indebted to Xavier Didelot for his help in setting up \texttt{ClonalOrigin}
 analysis.  We are thankful to Haruo Suzuki for his help
@@
 of Health, under grant number AI073368-01A2 awarded to A.S. and M.J.S.
+
+\section*{Author Contributions}
+
+Conceived and designed the experiments: SCC MDR MJS AS.
+Performed the experiments: SCC MDR MJH.
+Analyzed the data: SCC MDR MJH IG.
+Contributed reagents/materials/analysis tools: MJS.
+Wrote the paper: SCC MDR MJH IG.
 
@@
 \includegraphics[scale=0.42]{figures/mauve-analysis-002}
-\caption{\label{fig:clonalorigin}Recombinant tree in the ClonalOrigin model.
+\caption{\label{fig:clonalorigin}Recombinant tree in the \texttt{ClonalOrigin} model.
 (A) A recombinant tree relating the five genomes with three recombinant edges.
@@
 \begin{center}
-\includegraphics[width=5in]{figures/trans-hgt-heatmap.pdf}
+\includegraphics[width=5in]{figures/trans-hgt-heatmap}
 \end{center}
@@
 
+% Table S3. Simulation study for the three population parameters.
+\begin{table}
+\caption{\label{tab:sim-three-population-parameters}
+Simulation study of the three population parameters.}
+\noindent \begin{centering}
+\begin{tabular}{ccc}
+ & True & Estimates (Standard Deviation)\tabularnewline
+\hline
+Mutation rate per site & 0.081 & 0.081 (0.0011)\tabularnewline
+Recombination rate per site & 0.012 & 0.011 (0.0003)\tabularnewline
+Recombinant tract length & 744 bp & 919 bp (28)\tabularnewline
+\end{tabular}
+\par\end{centering}
+\end{table}
+\clearpage{}
+
+% Table S4. ClonalFrame Result.
+\begin{table}
+\caption{\label{tab:clonalframe}Species tree inference using ClonalFrame.}
+\noindent \centering{}\begin{tabular}{cc}
+Parameter & Estimates\tabularnewline
+\hline
+$\theta$ & 92871.84 (N/A)\tabularnewline
+$\nu$ & 0.134 (0.132, 0.135)\tabularnewline
+$R$ & 9370 (8950, 9780)\tabularnewline
+$\delta$ & 362 (347, 377)\tabularnewline
+$r/m$ & 4.66 (4.45, 4.87)\tabularnewline
+$\rho/\theta$ & 0.101 (0.096, 0.105)\tabularnewline
+TMRCA & 0.331 (0.324, 0.337)\tabularnewline
+$T$ & 1.02 (1.01, 1.04)\tabularnewline
+\end{tabular}
+\end{table}
+\clearpage{}
+
+% Table S5. ClonalOrigin's 1st stage result.
+\begin{table}
+\caption{\label{tab:three}Three population parameter estimates from
+ClonalOrigin. 
+The estimates of the three parameters were based on one of two independent
+MCMC chains.  We considered the similar 
+results of the two independent chains as confirmation of 
+the convergence of the first stage of \texttt{ClonalOrigin} MCMC.
+}
+\noindent \centering{}\begin{tabular}{ccc}
+& Replicate 1 & Replicate 2\tabularnewline
+Parameter & Estimate (IQR$^a$) & Estimate (IQR)\tabularnewline
+\hline
+$\theta^b$ & 0.081 (0.067, 0.094) & 0.081 (0.067, 0.099)\tabularnewline
+$\rho^c$ & 0.012 (0.006, 0.019) & 0.012 (0.006, 0.020)\tabularnewline
+$\delta^d$ & 744 (346, 2848) & 723 (348, 2870)\tabularnewline
+\end{tabular}
+
+$^a$ IQR stands for interquartile range.
+
+$^b$ The median of mutation rate per site weighted over the lengths of blocks.
+
+$^c$ The median of recombination rate per site weighted over the lengths of blocks.
+
+$^d$ The median value for the recombinant tract length weighted over the lengths of blocks.
+
+\end{table}
+\clearpage{}
+
 \section*{Supplementary Figures}
@@
 \includegraphics[scale=0.7]{figures/cornellf-3-tree}
-\caption{\label{fig:tree5}Species tree of five taxa for ClonalOrigin analysis.
-SPY is the branch of a clade\emph{ of Streptococcus pyogenes}, SDE
-is that for \emph{S.\ dysgalactiae} ssp.\emph{\ equisimilis}, and SD is the
-branch for the three individuals of \emph{S.\ dysgalactiae}. The scale is
+\caption{\label{fig:tree5}Species tree of five taxa for \texttt{ClonalOrigin} analysis.
+SPY is the branch of a clade\textit{ of Streptococcus pyogenes}, SDE
+is that for \textit{S.\ dysgalactiae} ssp.\textit{\ equisimilis}, and SD is the
+branch for the three individuals of \textit{S.\ dysgalactiae}. The scale is
 proportional to the expected number of mutations given the Watterson's estimate 
@@
 \begin{center}
-\includegraphics[width=5in]{figures/famsizes.pdf}
+\includegraphics[width=5in]{figures/famsizes}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=5in]{figures/trans-recomb-heatmap.pdf}
+\includegraphics[width=5in]{figures/trans-recomb-heatmap}
 \end{center}
@@
 
+% Figure S9: Mauve alignment
+%=============================================================================
+% FIGURE - Mauve alignment
+\begin{figure}
+\begin{center}
+\includegraphics[width=5in]{figures/mauve.pdf}
+\end{center}
+\vspace{-1.2in}
+\caption{Alignment of the five genomes.}
+\label{fig:mauve}
+\end{figure}
+%=============================================================================
+
 \end{document}
@@
 events boundaries. The number of recombination event boundaries per
-site for each block along the \emph{S.\ dysgalactiae} ssp.\emph{\ equisimilis}
+site for each block along the \textit{S.\ dysgalactiae} ssp.\textit{\ equisimilis}
 ATCC 12394.}
@@
 gene flow between the two species in their original model, \citet{Didelot2010}
-exploited their method to infer flows of genetic segment between \emph{Bacillus
+exploited their method to infer flows of genetic segment between \textit{Bacillus
 cereus} species. This model-based approach to detecting homologous
 recombining gene transfer was used to the core genomes of closely
-related \emph{Streptococcus} species.
+related \textit{Streptococcus} species.
 
@@
 We hoped that we could understand the power
-and limitations of ClonalOrigin under the parameter setup that we
+and limitations of \texttt{ClonalOrigin} under the parameter setup that we
 inferred from our data set.
@@
 lineages and arrive at one of SDE1 and SDE2 lineages than those for
-the reverse direction \emph{a priori} (Figure S\ref{fig:tree5}). 
+the reverse direction \textit{a priori} (Figure S\ref{fig:tree5}). 
 We took into account of the prior expected number of recombinant edges.
@@
 edges in Table S\ref{tab:genes-transfer} as we mentioned.
+
+% From Introduction
+It has been suspected that opportunistic pathogenicity
+of SDE may have been facilitated by the very genetic interaction between
+the two species. 
+Although virulence genes were suspected to be acquired via non-recombining gene
+transfer rather than recombining transfer \citep{Groisman1994}, systematic
+approaches to elucidating relative effects of the two types of gene transfer 
+rarely have been applied.
+
+This unfortunate situation motivated us to study the
+closely related bacteria as efforts of better understanding relative 
+association of gene transfer with virulence genes.
+We hoped that understanding their genetic exchanges
+would enhance our knowledge of mechanisms of their pathogenicity on
+humans.
+
+While compositional approaches were effective in analyzing
+genome-scale data, phylognetic approaches have allowed to understand
+evolutionary forces of gene exchange between species
+\citep[e.g.,][]{Bapteste2004,Simonson2005,Beiko2009} by comparing differences
+between gene trees and a reference species tree.  However, many of the methods
+did not have explicit models of homologous recombining gene transfer that was
+one of important evolutionary processes in horizontal gene transfer.  
+% Overall,
+% the development of methods for downstream analyses using bacterial genomes
+% appears to lag behind that of high-throughput sequencing technology
+% \citep[e.g.,][]{Green2011}.
+
+% Heat map
+% Another useful
+% summary suggested by \citet{Didelot2010} was to classify recombinant edges by
+% types defined according to donor/recipient branch pairs, and record the number
+% of edges sampled for each recombinant edge type. For instance, our species tree
+% has nine branches including rooting branch and 81 branch pairs, out of which 60
+% are possible donor/recipient pairs by prohibiting recombinant edges from going
+% backward in time. The posterior expected number of edges of each type can be
+% compared to the prior expected number determined by the three population
+% parameters fixed in the second stage of \texttt{ClonalOrigin}. This comparison
+% was displayed in a heat map of Figure \ref{fig:Heatmap-of-recombination}, which
+% indicated branch pairs that might experience an excess or depletion of
+% recombination events across all blocks.

### 2011-09-29 pl/scatter-plot-parameter.pl
The posterior probability of recombination tracks were added for all of 5 genomes.

@@ -218,15 +218,15 @@ sub endElement {
   if($tag eq "theta"){
     my $thetaPerSite = $content / $blockLength;
     # print OUTTHETA "$blockID\t$thetaPerSite\n";
-    print OUTTHETA "$position\t$thetaPerSite\n";
+    print OUTTHETA "$position\t$thetaPerSite\t$blockID\n";
   }
   if($tag eq "rho"){
     $rhoPerBlock = $content;
   }
   if($tag eq "delta"){
-    print OUTDELTA "$position\t$content\n";
+    print OUTDELTA "$position\t$content\t$blockID\n";
     my $rhoPerSite = $rhoPerBlock / ($blockLength + $content - 1);
-    print OUTRHO "$position\t$rhoPerSite\n";
+    print OUTRHO "$position\t$rhoPerSite\t$blockID\n";
   }
   $tag = "";
   $content = "";
@@ -273,7 +273,7 @@ sub printError {
 __END__
 =head1 NAME
 
-scatter-plot-parameter.pl - Build a heat map of recombination.
+scatter-plot-parameter.pl - File for plotting three population parameter estimates
 
 =head1 VERSION
 
@@ -281,8 +281,7 @@ scatter-plot-parameter.pl 1.0
 
 =head1 SYNOPSIS
 
-perl scatter-plot-parameter.pl [-h] [-help] [-version] 
-  [-xmlbase filename] [-out file]
+perl pl/scatter-plot-parameter.pl three -xmlbase path/to/co1output -xmfabase filepath/to/corexmfa
 
 perl pl/scatter-plot-parameter.pl bed -in scatter-plot-parameter-1-out-rho
 

### 2011-09-29 pl/virulence.pl

@@ -23,6 +23,7 @@ use Getopt::Long;
 use Pod::Usage;
 require 'pl/sub-error.pl';
 require 'pl/sub-ptt.pl';
+require 'pl/sub-bed.pl';
 $| = 1; # Do not buffer output
 my $VERSION = 'virulence.pl 1.0';
 
@@ -40,6 +41,7 @@ GetOptions( \%params,
             'verbose',
             'version' => sub { print $VERSION."\n"; exit; },
             'ptt=s',
+            'bed=s',
             'virulence=s',
             'ortholog=s',
             'in=s',
@@ -87,7 +89,8 @@ if ($cmd eq "table")
 }
 elsif ($cmd eq "bed")
 {
-  unless (exists $params{virulence} and exists $params{ptt})
+  unless (exists $params{virulence} 
+          and (exists $params{ptt} or exists $params{bed}))
   {
     &printError("command bed needs options -virulence, and -ptt");
   }
@@ -226,19 +229,44 @@ elsif ($cmd eq "bed")
   }
   close VIR;
 
-  my @genesPtt = rnaseqPttParse ($params{ptt});
-  foreach my $g (@genesPtt)
+  if (exists $params{ptt})
   {
-    if (exists $virulence{$g->{Synonym}})
+    my @genesPtt = rnaseqPttParse ($params{ptt});
+    foreach my $g (@genesPtt)
     {
-      if ($virulence{$g->{Synonym}} eq "TRUE")
+      if (exists $virulence{$g->{Synonym}})
       {
-        $g->{Location} =~ /(\d+)\.\.(\d+)/;
-        my $start = $1;
-        my $end = $2;
-        print $outfile "chr1\t$start\t$end\n";
+        if ($virulence{$g->{Synonym}} eq "TRUE")
+        {
+          $g->{Location} =~ /(\d+)\.\.(\d+)/;
+          my $start = $1;
+          my $end = $2;
+          my $name = $g->{Synonym};
+          print $outfile "chr1\t$start\t$end\t$name\n";
+        }
       }
     }
+  }
+  elsif (exists $params{bed})
+  {
+    my @genesBed = rnaseqBedParse ($params{bed});
+    foreach my $g (@genesBed)
+    {
+      if (exists $virulence{$g->{name}})
+      {
+        if ($virulence{$g->{name}} eq "TRUE")
+        {
+          my $start = $g->{start};
+          my $end = $g->{end};
+          my $name = $g->{name};
[…]

### 2011-09-29 pl/xmfaToMaf.pl

@@ -46,6 +46,8 @@ GetOptions( \%params,
             'version' => sub { print $VERSION."\n"; exit; },
             'ref=i',
             'xmfa=s',
+            'scorefile=s',
+            'numberblock',
             'xmfa2maf=s',
             'rename=s',
             'reorder=s',
@@ -76,11 +78,38 @@ if ($cmd eq 'ucsc')
     &printError("Command $cmd requires option -xmfa2maf");
   }
 }
+elsif ($cmd eq 'block')
+{
+  unless (exists $params{xmfa2maf})
+  {
+    &printError("Command $cmd requires option -xmfa2maf");
+  }
+}
 
 ################################################################################
 ## DATA PROCESSING
 ################################################################################
-if ($cmd eq 'ucsc')
+if ($cmd eq 'number')
+{
+  my $count = 0;
+  my $l;
+  open MAF, $params{xmfa2maf} or die "cannot open < $params{xmfa2maf} $!";
+  while ($l = <MAF>)
+  {
+    chomp $l;
+    if ($l =~ /^a/)
+    {
+      $count++;
+      print $outfile "a $count\n";
+    }
+    else
+    {
+      print $outfile "$l\n";
+    }
+  }
+  close MAF; 
+}
+elsif ($cmd eq 'ucsc')
 {
   my @countLineAlignment = peachMafCountFromMauveXmfa2Maf ($params{xmfa2maf});
   my @strand = peachMafStrandFromMauveXmfa2Maf ($params{xmfa2maf});
@@ -126,7 +155,14 @@ if ($cmd eq 'ucsc')
       {
         die "The first character must be character a" unless $line =~ /^a/;
         # print $outfile "$line\n";
-        push @alignment, $line;
+        if (exists $params{numberblock})
+        {
+          push @alignment, "a $i";
+        }
+        else
+        {
+          push @alignment, $line;
+        }
       }
       elsif ($j == $countLineAlignment[$i] - 1)
       {
@@ -143,7 +179,7 @@ if ($cmd eq 'ucsc')
         my @e = split /\s+/, $line;
         if ($strand[$i-1][$reorder[1]-1] eq '-')
         {
-          $e[2] = $e[5] - $e[2];
+          # $e[2] = $e[5] - $e[2];
           if ($e[4] eq '+')
           {
             $e[4] = '-';
@@ -164,6 +200,53 @@ if ($cmd eq 'ucsc')
 }
[…]

### 2011-09-29 sh/recombination-intensity1-probability.sh

@@ -52,7 +52,8 @@ function recombination-intensity1-probability {
       echo -n "Do you wish to create wiggle files of recombination probability using blocks (y/n)? "
       read WISH
       if [ "$WISH" == "y" ]; then
-        mkdir $RUNANALYSIS/recombprob-$REPLICATE
+        OUTDIR=$RUNANALYSIS/recombprob-ref$REFGENOME-rep$REPLICATE
+        echo mkdir $OUTDIR > batch.sh
         echo perl pl/$FUNCNAME.pl wiggle \
           -xml $RUNCLONALORIGIN/output2/${REPLICATE}/core_co.phase3.xml \
           -xmfa2maf $DATADIR/core_alignment.maf \
@@ -61,7 +62,7 @@ function recombination-intensity1-probability {
           -gbk $GBKFILE \
           -ri1map $RUNANALYSIS/rimap.txt \
           -clonaloriginsamplesize $NUMBER_SAMPLE \
-          -out $RUNANALYSIS/recombprob-$REPLICATE > batch.sh
+          -out $OUTDIR >> batch.sh
       else
         echo -e "  Skipping drawing recombination probability ..." 
       fi

### 2011-09-29 doc/manuscript.tex
Just an update.

@@
 
+Furthermore, we found a total of 58 genes 
+for which a recombinant edge 
+of certain types were sampled with probability at least 0.9
+by focusing on specific types of recombinant edges.
+(Table S\ref{tab:genes-transfer}).  Many of the genes in the list were
+transferred from SPY to SDE rather than the reverse direction overall.
+
 % \texttt{IG: ADD FIGURE WITH NON-NORMALIZED VALUES? SC: Why?}
@@
 \texttt{I need to see how the virulence genes list was created.}
-\texttt{I want to know the virulence gene categories. Which kind of virulence
+\texttt{I want to know the virulence gene categories. What kind of virulence
 genes are more under gene transfer?}
@@
 
+A list of virulence genes that might be transferred via recombining gene
+transfer.
+block 76: SDEG\_0641
+block 97: SDEG\_0673
+block 101: SDEG\_0739
+block 151: SDEG\_1447
+block 244: SDEG\_2040
+block 242: SDD27957\_00955
+block 240: SDD27957\_00985
+block 67: SDD27957\_03085
+block 133: SDD27957\_04505
+block 167: SDD27957\_06790
+block 170: SDD27957\_07725
+block 186: MGAS10750\_Spy1444
+block 255: MGAS10750\_Spy1922
+
+Blocks with virulence genes.
+
+Check virulence genes (they were combined) at 
+SDD27957\_02890, SDD27957\_02895
+MGAS10750\_Spy0455, MGAS10750\_Spy0456 (block 75)
+
 \subsection{Horizontal gene transfer is associated with virulence genes}
@@
 \subsection{Comparison between model-based and parsimony-based approaches}
-\texttt{I might want to analyze using only core genomes and pan genomes
+\texttt{I might want to analyze only core genomes and pan genomes
 separately. By separating data set of 6-gene families from the other families,
@@
 
-\subsection{Functional categories associated with gene transfer}
-
-% \textbf{Functional categories associated with homologous recombination gene transfer:} 
-% Having established relative correlation of the surrogate measure of recombination
-% intensity in the simulation (Figure S\ref{fig:ri1}) we attempted to find
-% functional categories that were associated with recombination.
-% The nucleotide substitutions
-% due to recombination events might not be selectively neutral because
-% of large population of the bacteria, and accordingly natural selection
-% could be more stronger than random genetic drift. Reflecting on the
-% report of \citet{Lefebure2007} that genes that were likely to be
-% under positive selection tended to be under higher recombination,
-% we would expect that functional categories would be associated with higher
-% recombination, and consequently homologous recombining gene transfer
-% between species.
-
-Table \ref{tab:functional} lists the functional categories that showed
-significantly associated with recombination intensity. 
-Several functional categories are related with ribosomes.
-There was a functional category related with phosphotransferase systems
-and ATP-binding cassette (ABC) transporters, i.e., carbohydrate transport
-(GO:0008643). Translation (GO:0006412) included aminoacyl tRNA synthetase.
-There were fatty acid biosynthetic process (GO:0006633) and lipid
-biosynthetic process (GO:0008610). We did not find any significant association 
-between recombination intensity and virulence genes. When measuring 
-recombination only between the SDE and SPY clades, a smaller number of 
-functional categories appeared to be significantly enriched, leaving the two 
-biosynthetic processes out. For the direction from SPY to SDE only transport 
-related
-functional categories appeared to be significant, leaving out ribosome related 
-functional categories and translation.
-
-%\textbf{Functional categories associated with non-recombining gene transfer:}
-
-Table \ref{tab:go-events} lists the functional gene categories that
-showed significantly associated with gene duplications, losses,
-non-recombining and recombining gene transfer using the pan genome of the six
-\textit{Streptococcus} individuals.
-
-For the recombining transfers, Mowgli shows enrichment in
-many of the same categories as \texttt{ClonalOrigin}, such as translation,
-ribosome, and rRNA binding.  This agreement between the two different
-methods, applied on somewhat different parts of the data, serves as
-validation for the signal picked up by these separate association
+\subsection{Transporter and aminoacyl tRNA synthetases}
+Lateral transfer of transporter related genes are well known.  Functional
+categories associated with higher recombining gene transfer were found to
+include phosphotransferase systems and ATP-binding cassette (ABC) transporters,
+i.e., carbohydrate transport (GO:0008643) (Table \ref{tab:functional}).  We also
+found several aminoacyl tRNA synthetases in gene ontology of translation
+(GO:0006412) to be significantly associated with higher recombining gene
+transfer.  Note that functional categories concerning ribosomes were also
+associated with higher recombining gene transfer. These categories were not
+found significantly associated with higher recombining gene transfer when we
+considered only gene transfer between SPY and SDE genomes, which implied that
+many recombining gene transfer events should be attributed to those of internal
+species branches as donors.  Interestingly, transports functional categories
+were found to be associated with recombining gene transfer for the direction
+from SPY to SDE.  The parsimony-based approach also  shows enrichment of the
+recombining transfers in many of the same categories such as translation,
+ribosome, and rRNA binding (Table \ref{tab:go-events}).  This agreement between
+the two different methods, applied on somewhat different parts of the data,
+serves as validation for the signal picked up by these separate association
 studies.  
-
-However, for duplication, losses, and non-recombining
-transfers, we find enrichment in very different categories related to
-transposition, integration, and recombination.  Interestingly, gene
-duplication and non-recombining transfers show enrichment in the same
-categories, perhaps because both mechanisms increase gene copy number.
-Gene losses also showed enrichment for transposition and integration,
-but less significantly ($P<10^{-6}$) and also showed enrichment for
-several additional categories such as transporter activity and
-transcription initiation.  Unlike the associations with recombination
-intensity, families with high numbers of duplications and
-non-recombining gene transfer were significantly enriched for
-virulence ($P<7.13 \times 10^{-3}$ and $P<1.33 \times 10^{-11}$,
-respectively), whereas families with high losses and recombination
-showed no enrichment ($P>0.54$ and $P>0.90$).  This might indicate
+\texttt{Comparison of Mawgli and CO?}
+Recombining gene transfers were not enriched for virulence genes (P-value of $>0.90$)
+while horizontal gene transfers were found to be significantly associated with
+virulence genes (P-value of $<1.33 \times 10^{-11}$). This was also consistent
+with the fact that we could not find the association of recombining gene
+transfer with virulence genes in the model-based approach.
+This supported more evidence
 that non-recombining gene transfer served as a more effective means by
@@
 gene transfer.
-
-\texttt{NEXT PARAGRAPH CONTAINS RESULTS ON INFERENCE OF SPECIFIC TRANSFER 
-EVENTS. WE SHOULD DECIDE HOW TO RELATE THIS TO OUR OTHER RESULTS.}
-
-%\textbf{Some detailed gene-level analyses:}
-Having found these functional categories, we elected to search for specific 
-genes that showed a clear and specific local signal of recombination.  We scanned 
-all 274 alignment blocks to locate parts of genes for which a recombinant edge 
-of a certain type was sampled with probability at least 0.9. 
-% We focused on specific types of recombinant edges 
-% \texttt{IG: ADD A BIT HERE. SC: I DO NOT SEE WHAT IG MEANT.}
-We found a total of 58 genes showing high-confidence local 
-recombination signal (Table S\ref{tab:genes-transfer}). The process of gene 
+For horizontal gene transfer we found enrichment in very different 
+categories related to
+transposition, integration, and recombination.  
+\texttt{More interpretations about Mawgli analsis of function categories?}
+
+Gene duplication and loss events were enriched for categories similar to those
+of horizontal gene transfer's enrichment.  Interestingly, gene duplication and
+horizontal transfers showed enrichment in the same categories, perhaps because
+both mechanisms increase gene copy number.  Gene losses showed enrichment
+for transposition and integration, but less significantly (P-value of
+$<10^{-6}$), and showed enrichment for several additional categories such as
+transporter activity and transcription initiation.  Families with high numbers
+of duplications were significantly enriched for virulence genes (P-value of
+$<7.13 \times 10^{-3}$), whereas families with high losses showed no enrichment
+(P-value of $>0.54$).
+
+% \texttt{IG: NEXT PARAGRAPH CONTAINS SOME DISCUSSION ON THESE RESULTS. WE SHOULD 
+% GET BACK TO THIS ONCE WE DECIDE WHAT TO DO WITH THE RESULTS}
+
+\subsection{Units of recombining gene transfer}
+The process of gene 
 transfer does not necessarily move whole genes from one species to another 
@@
 recombining gene transfer do 
-not correspond to single genes in their entirety. In some cases, only parts of 
+not correspond to single genes in their entirety
+(Table S\ref{tab:genes-transfer}). In some cases, only parts of 
 a gene were inferred to have been transferred, for instance, transfer from SPY1 
@@
 of as much as eight adjacent genes.
-
-%Not only parts of genes could be transferred via recombination, but a group of 
-%genes could also be moved together. Of the genes listed in Table 
-%S\ref{tab:genes-transfer} there were 12 groups of genes that were consecutive. 
-%While transfer fragments of genes often covered two genes, there were transfer 
-%fragments that covered eight genes together. Loci from SpyM3\_1512 to 
-%SpyM3\_1525 were inferred to be transferred from SDE2 to SDD.
-
-% \texttt{IG: NEXT PARAGRAPH CONTAINS SOME DISCUSSION ON THESE RESULTS. WE SHOULD 
-% GET BACK TO THIS ONCE WE DECIDE WHAT TO DO WITH THE RESULTS}
+Not only parts of genes could be transferred via recombination, but a group of 
+genes could also be moved together. Of the genes listed in Table 
+S\ref{tab:genes-transfer} there were 12 groups of genes that were consecutive. 
+While transfer fragments of genes often covered two genes, there were transfer 
+fragments that covered eight genes together. Loci from SpyM3\_1512 to 
+SpyM3\_1525 were inferred to be transferred from SDE2 to SDD.
 
@@
 % recombination events across all blocks.
+
+% \textbf{Functional categories associated with homologous recombination gene transfer:} 
+% Having established relative correlation of the surrogate measure of recombination
+% intensity in the simulation (Figure S\ref{fig:ri1}) we attempted to find
+% functional categories that were associated with recombination.
+% The nucleotide substitutions
+% due to recombination events might not be selectively neutral because
+% of large population of the bacteria, and accordingly natural selection
+% could be more stronger than random genetic drift. Reflecting on the
+% report of \citet{Lefebure2007} that genes that were likely to be
+% under positive selection tended to be under higher recombination,
+% we would expect that functional categories would be associated with higher
+% recombination, and consequently homologous recombining gene transfer
+% between species.

### 2011-10-03 doc/manuscript.tex
The posterior predictive checking procedure is being done with s17 simulation.

@@
 
-phosphotransferase system (PTS), ATP-binding cassette (ABC), Aminoacyl
-tRNA synthetase (aaRS), ancestral recombination graph (ARG), group
-A streptococci (GAS), group C streptococci (GCS), group G streptococci
+group A streptococci (GAS), group C streptococci (GCS), group G streptococci
 (GGS), \textit{Streptococcus pyogenes} (SPY), \textit{S.\ dysgalactiae}
 subspecies \textit{dysgalactiae} (SDD), \textit{S.\ dysgalactiae} subspecies
-\textit{equisimilis} (SDE).
+\textit{equisimilis} (SDE)
 
@@
 (SDE).  The former is a pervasive human-specific pathogen, and the latter an
-opportunistic human pathogen.  Parts of the genomic configuration, including
-human pathogenicity, of SDE could have originated in those of SPY.  Our
-genome-wide survey of gene transfer indicated that gene transfer via homologous
-recombination occurred relatively from SPY to SDE.  Although we did not find
-enrichment of virulence genes in the homologous recombining gene transfer, we
-found evidence that they transferred more often via non-recombining fashion than
-non-virulence genes.  A genome-wide resource of recombining and non-recombining
-gene transfer for the study of the genomes is provided so that research
-community in pathogenic bacteria can take advantage of it.
+opportunistic human pathogen.  We aims at which parts of the genomic
+organization, including human pathogenic virulence factors, of SDE could have
+originated in those of SPY, or vice versa.  Our genome-wide survey of gene
+transfer indicated that gene transfer via homologous recombination occurred
+relatively from SPY to SDE.  Although we did not find enrichment of virulence
+genes in the homologous recombining gene transfer, we found evidence that they
+transferred more often via non-recombining fashion without replacing parts of
+destination genomes than did non-virulence genes.  A genome-wide resource of
+recombining and non-recombining gene transfer for the study of the genomes is
+provided so that research community in pathogenic bacteria could take advantage
+of it.
 
@@
 
-The research topic on gene exchanges between SPY and SDE has been discussed in
-various ways. \citet{Sachse2002} showed gene transfer from SDE to SPY for the
-downstream of the genomic region of streptococcal pyrogenic exotoxins
-\textit{speG}. \citet{Davies2007} investigated phage 3396 from SDE as a vehicle
-of gene transfer from SDE to SPY. \citet{Davies2009} demonstrated that an
-integrative conjugative element of SDE allowed gene transfer from SDE to other
-streptococci.  \citet{Ahmad2009} did not conclusively deduce the net
+Therefore, the research topic on gene exchanges between SPY and SDE has been
+discussed in various ways. \citet{Sachse2002} showed gene transfer from SDE to
+SPY for the downstream of the genomic region of streptococcal pyrogenic
+exotoxins \textit{speG}. \citet{Davies2007} investigated phage 3396 from SDE as
+a vehicle of gene transfer from SDE to SPY. \citet{Davies2009} demonstrated that
+an integrative conjugative element of SDE allowed gene transfer from SDE to
+other streptococci.  \citet{Ahmad2009} did not conclusively deduce the net
 directionality of housekeeping gene transfer between SPY and SDE.  Notably,
@@
 to SDE was more dominant than the other direction.  Because these researches on
-the controversial issue, directionality of gene transfer, have been limited to
-relatively small regions in the genomes, it remains to be elusive to determine
-the genome-wide degrees to which genomic segments flow between the two species,
-and directions of thereof.
+the controversial issue, directionality of gene transfer between SPY and SDE,
+have been limited to relatively small regions in the genomes primarily on
+representative loci from the Multi Locus Sequence Typing database, it remains to
+be elusive to determine the genome-wide degrees to which genomic segments flow
+between the two species, and net directions of thereof.
 
@@
 of \textit{Shewanella baltica} genomes to identify the genetic elements that
-enabled the species to adapt to redox gradients.
+enabled the species to adapt to redox gradients.  Although several complete
+genomes from SPY have been available, \citet{Suzuki2011} recently added one more
+complete genome of SDE to the previous set of a unique SDE genome. This set of
+SPY and SDE genomes prompted us to study genome-wide gene flows between the two
+species that afflicted human beings around the globe.
 
@@
 the same lineage.  Mutations on a gene modifiy relatively small portions of the
-gene, including substitutions, insertions, and deletions of deoxyribonucleic
+gene through substitutions, insertions, and deletions of deoxyribonucleic
 acids. Horizontal gene transfers can reshape the organization of a bacterial
 genome, which could revamp biological functions of the bacterial species.
-Recipient bacteria can substitute transferred genes for homologous parts of
-their genomes. They can also integrate transferred genes into their genomes
-without replacement. These two processes are here discriminated by calling the
+Recipient bacteria of transferred genes can substitute transferred genes for homologous parts of
+their genomes, or integrate transferred genes into their genomes
+without replacement. These two processes are here in this paper discriminated by calling the
 former \textit{recombining gene transfer}, and the latter \textit{horizontal
@@
 complete genome sequences in efforts of better understanding bacterial evolution
-have paid little attention to a combined approach to illuminating disparity of
-the two gene transfer processes.
+have paid less attention to a combined approach to illuminating disparity of
+the two gene transfer processes. We attempted here to use and compare two
+approaches for distinguishing recombining gene transfer process from horizontal, or
+non-recombining, gene transfer one, and more systematically tackled their roles
+in association with virulence genes.
 
@@
 duplication, gain, loss, etc.  Because of the complexity, these processes do not
-readily lend themselves to statistically more sound approaches such as maximum
-likelihood. Instead, a parsimony-based approach to inferring gene gain, loss,
-and transfer events was developed by  \citep{Doyon2011}.  We noted that the
-first approach could be used to infer transfer of genomic segments due to
-homologous recombination, and the second could do similarly due to horizontal
-gene transfer. We expect that model-based approaches using whole bacterial
-genomes will be developed in the near future.
-
-The complete genomes of the two spcies of SPY and SDE presented a unique
+readily lend themselves to statistically more sound model-based approaches such
+as maximum likelihood. Instead, a parsimony-based approach to inferring gene
+gain, loss, and transfer events was developed by  \citep{Doyon2011}.  We noted
+that the first approach could be used to infer transfer of genomic segments due
+to homologous recombination, and the second could do similarly attributed to
+horizontal gene transfer. Although this heterogeneously combined approach could
+be unsatisfactory, we hoped to take the unique opportunity of new genomes to
+study gene transfer in genome-wide scale among SPY and SDE individual bacteria.
+
+In short, the complete genomes of the two spcies of SPY and SDE presented a unique
 opportunity to study gene transfer over relatively recent time scales.  We used
@@
 biological functional categories to those inferred with the model-based
-approach. This is the first attempt to analyaze SPY and SDE in genome-wide scale
-for detecting gene transfer events not only in the core genome but also in the
-pan genome, which make the analysis complete in terms of the amount of used
-genome data. Our analysis of core genomes also showed for the first time that
-the net directionality of, at least, recombining gene transfer was more from SPY
-to SDE.
+approach. This is the first attempt as far as we know to analyaze SPY and SDE in
+genome-wide scale for detecting gene transfer events not only in the core genome
+but also in the pan genome, which make the analysis complete in terms of the
+amount of used genome data. Our analysis of core genomes also showed for the
+first time that the net directionality of, at least, recombining gene transfer
+from SPY to SDE was stronger than the reverse direction.
 
@@
 
-\subsection{The five or six genomes}
-
-\subsection{The genome alignment of the five genomes}
-
-\subsection{The species tree inferred using the 274 alignment blocks}
-
-\subsection{The homologous recombination events}
-An alignment of the five genomes showed two distinct genome structures of SPY
-and SDE (Figure S\ref{fig:mauve}). The SDE genomes were larger in length than
-those of SPY. Although it was unclear that SDD was different from either of SDE1
-and SDE2 from the graphical view of the genome alignment, the spcies tree from
-the genome alignment showed that SDD branch was attached to the clade branch of
-SDE (Figure S\ref{fig:tree5}). The distribution of the alignment block lengths
-was skewed to the right with mean of 4188 base pairs long (Figure
-S\ref{fig:blocksize}), which implied that the model assumption of geometric
-length distribution might be plausible in \texttt{ClonalOrigin} method.  The
-relative strength of recombination compared to mutations indicated that
-reconmination-driven substitutions affected more sites than did mutation-driven
-substitutions (Table S\ref{tab:clonalframe}).
-
-\subsubsection{The three population parameters}
+\subsection{Recombination-driven substitutions were stronger than mutation-driven ones}
+
+The genome alignment of the five genomes (Table S\ref{tab:genome}) showed two distinct
+genome structures of SPY and SDE (Figure S\ref{fig:mauve}). The SDE genomes were
+larger in length than those of SPY. Although it was unclear that SDD was
+different from either of SDE1 and SDE2 from the graphical view of the genome
+alignment, the spcies tree inferred from the genome alignment showed that SDD
+branch was attached to the clade branch of SDE1 and SDE2 (Figure
+S\ref{fig:tree5}). The distribution of the alignment block lengths was skewed to
+the right with mean of 4188 base pairs long (Figure S\ref{fig:blocksize}), which
+implied that the model assumption of geometric length distribution might be
+plausible in \texttt{ClonalOrigin} method (see section \textbf{Materials and
+Method} and \citealt{Didelot2010}).  The relative strength of recombination
+compared to mutations indicated that reconmination-driven substitutions affected
+more sites than did mutation-driven substitutions (Table
+S\ref{tab:clonalframe}). In the \textit{Streptococcus} bacteria homologous
+recombination appeared to be a stronger evolutionary force than point mutations 
+process.
+
+\subsubsection{Recombination hotspots were observed in genomes}
+
 Along the genome of SDE1 population mutation rate did not appear to change much
-(Figure \ref{fig:scatter3}).  We found a few spots of genomic regions where the
+(Figure S\ref{fig:scatter3}).  We found a few spots of genomic regions where the
 mutation rates were larger than twice the global estimate of mutation rate.
-\texttt{WHAT ARE THESE BLOCKS? WHAT GENES ARE LOCATED IN THE BLOCK?}
+% \texttt{WHAT ARE THESE BLOCKS? WHAT GENES ARE LOCATED IN THE BLOCK?}
 The recombination rate appeared to change more dynamically. Espeically, the
@@
 contains alignment blocks with bell-shaped recombination rates.
-\texttt{USE UCSC GENOME BROWSER TO FIND WHAT GENES BELONG TO THOSE BLOCKS.}
+% \texttt{USE UCSC GENOME BROWSER TO FIND WHAT GENES BELONG TO THOSE BLOCKS.}
 The recombinant tract length changes in a way that we could not note any
@@
 recombination rate.
-\texttt{PLOT PAIRWISE SCATTER PLOTS OF THE THREE PARAMETERS.}
-In the first stage of \texttt{ClonalOrigin} MCMC a few alignment blocks were not
-successfully finished even after a month of computation.  Because those were not
-expected to affect the global estimates of population parameters significantly,
-we used only finished results of the first stage of \texttt{ClonalOrigin} to
-estimate mutation rate, recombination rate, and recombinant tract length.  Two
-independent runs of the first stage MCMC resulted in similar estimates of the
-three population parameters (Table S\ref{tab:three}).
-
-\subsection{Tree topology changes along the genome}
-Most of local tree topologies, or 66.7\%, along the alignment blocks were the
+% \texttt{PLOT PAIRWISE SCATTER PLOTS OF THE THREE PARAMETERS.}
+
+\subsection{SPY and SDE were more closely related in about tenth of the genome}
+
+Most of local tree topologies (see section \textbf{Materials and Methods}), 
+or 66.7\%, along the alignment blocks were the
 same as the species tree, which allowed us to be more confident of the species
 tree inferred using the alignment blocks (Table
-S\ref{tab:Gene-tree-topologies}).  The second most probable topology was one
-where SDD was the outgroup, and SPY and SDE were sister clades with probability
-of 9.3\%.  Therefore, SPY and SDE appeared to share about tenth of the core
-genomes that were more closely related with each other than each to SDD's
+S\ref{tab:Gene-tree-topologies}).  The second most probable topology (9.3\%)  was one
+where SDD was the outgroup, and SPY and SDE were sister clades 
+SPY and SDE appeared to share about tenth of the core
+genomes that they were more closely related with each other than each to SDD's
 genome. The other topologies had less than 4\% of proportions.
 
-\subsection{Evidence for unbalaned gene flow from SPY to SDE}
+\subsection{Unbalaned gene flow from SPY to SDE in genome-wide scale}
+
 The heat map in Figure \ref{fig:Heatmap-of-recombination} shows the log base 2
-ratios of these numbers and their corresponding expected values given the
-species tree and the global parameter settings.  Because the recombinant edge
-from any species tree branch to the root branch is not possible under
-\texttt{ClonalOrigin} model (see Figure \ref{fig:clonalorigin}), the column for
-the root as recipient was colored green. Likewise, all of the cells where a
-recipient branch was older than the donor, and two branches did not share any
-contemporary time were colored green. The diagonal cells were colored gray
-because the observed number of recombinant edges and the expected one were
-similar. Because recombinant edges starting from and ending in the same species
-tree branch would not change the species tree, which would not change the
-likelihood, the cells should be colored gray.  Because SDD is strictly
-veterinary hosts, we did not expect that SDD received much gene transfer from
-either SPY1 or SPY2.  The log base 2 ratio from SPY1 to SDD was -2.20, and that
-from SPY2 to SDD was -1.97. On the contrary, the recombing gene transfer from
-SDE1 to SDD was 0.756, and that from SDE2 to SDD was 0.698. The relatively
-higher recombining gene transfer from two SDE individuals to SDD compared with
-that from two SPY individuals to SDD reflected that the SDD was more closely
-related to SDE than to SPY.  The gene transfer from SDD to either SDE1 or SDE2
-was less frequent than the case of reverse direction.
-
-The recombining gene transfer log base 2 ratios between SDE1 and SDE2 were about
-twice as large as that from SDE1 and SDE2 to SPY1 and SPY2. Because the
-recombining gene transfer within a clade would be larger than that between
-clades, this was expected. Surprisingly, the recombining gene transfer log base
-2 ratios between SPY1 and SPY2 was comparable with that from SPY1 and SPY2 to
-SDE1 and SDE2.  The recombining gene transfer from a clade SPY1 or SPY2 to a
-different clade SDE1 or SDE2 was not different from that within the clade, or
-SPY1 and SPY2.  Consistent with this result, the gene transfer log base 2 ratio
-from SPY1 and SPY2 to SDE branch was larger than any type of gene transfer: 1.88
-and 1.94, respectively. This clearly showed that the recombining gene transfer
-was biased toward the net direction from SPY to SDE.
-
-Furthermore, we found a total of 58 genes 
-for which a recombinant edge 
-of certain types were sampled with probability at least 0.9
-by focusing on specific types of recombinant edges.
-(Table S\ref{tab:genes-transfer}).  Many of the genes in the list were
-transferred from SPY to SDE rather than the reverse direction overall.
-
-% \texttt{IG: ADD FIGURE WITH NON-NORMALIZED VALUES? SC: Why?}
-
-% No evidence for unbalanced non-recombining gene transfer between SPY and SDE
-Figure S\ref{fig:hgt-heatmap} shows the number of inferred
-non-recombining gene transfer events between the different branches of
-the species tree.
-SDE1 and SDE2 shared the most events (26 and 34) compared to other sample pairs. 
-SDD appeared to acquire many genes from the ancestral SDE branch. Transfer 
-events between the SPY and SDE clades were inferred in smaller numbers (<6) 
-compared to those within each clade (23--34). Among the clade-crossing 
-transfers, we did not see any significant enrichment for a specific direction, 
-as opposed to the result for recombining transfers from the model-based 
-analysis (Figure S\ref{fig:mowgli-recomb-heatmap}).
-
-\subsection{Transport-related genes are associated with recombining gene
-transfer}
-
-\subsection{The virulence genes}
-\texttt{Find cases of any virulence genes that were transferred via recombining and
-horizontally.} 
-\texttt{I need to see how the virulence genes list was created.}
-\texttt{I want to know the virulence gene categories. What kind of virulence
-genes are more under gene transfer?}
-\texttt{Make a list of virulence genes and their categories with more
-interpretations.}
-
-A list of virulence genes that might be transferred via recombining gene
-transfer.
-block 76: SDEG\_0641
-block 97: SDEG\_0673
-block 101: SDEG\_0739
-block 151: SDEG\_1447
-block 244: SDEG\_2040
-block 242: SDD27957\_00955
-block 240: SDD27957\_00985
-block 67: SDD27957\_03085
-block 133: SDD27957\_04505
-block 167: SDD27957\_06790
-block 170: SDD27957\_07725
-block 186: MGAS10750\_Spy1444
-block 255: MGAS10750\_Spy1922
-
-Blocks with virulence genes.
-
-Check virulence genes (they were combined) at 
-SDD27957\_02890, SDD27957\_02895
-MGAS10750\_Spy0455, MGAS10750\_Spy0456 (block 75)
-
-\subsection{Horizontal gene transfer is associated with virulence genes}
-
-\subsection{Comparison between model-based and parsimony-based approaches}
-\texttt{I might want to analyze only core genomes and pan genomes
-separately. By separating data set of 6-gene families from the other families,
-we could have more complete comparison.}
-
-\subsection{UCSC genome browser for the dataset}
-We created a web-resource for gene transfer available at
-\url{http://strep-genome.bscb.cornell.edu}.  For each of the five individual
-genomes as a reference genome several annotation tracks were displayed:
-alignment blocks, recombination rate per site for each block, local gene tree
-topologies, recombination
-intensity, locations of virulence genes, posterior probability of recombining
-gene transfer.
-
-\subsection{Parsimony-based approach for gene gain, loss, and duplication}
-% Gene gain, loss, duplication, and non-recombining horizontal gene transfer
+ratios of the \textit{a posteriori} number of recombinant edges for a pair of
+donor and recipient species tree branches and their corresponding expected
+values given the species tree and the global parameter settings.  Because
+recombinant edges from any species tree branch to the root branch are not
+possible under \texttt{ClonalOrigin} model (see Figure S\ref{fig:clonalorigin}),
+the column for the root as recipient was colored green. Similarly, all of the
+cells where a recipient branch was older than the donor, and two branches did
+not share any contemporary time were colored green. The diagonal cells were
+colored gray, indicating no significant number of recombinant edges, because the
+observed number of recombinant edges and the expected one were similar. As
+recombinant edges starting from and ending in the same species tree branch would
+not change the species tree, which would not change the likelihood much, the
+cells should be colored gray.  Because SDD is a strictly veterinary pathogenic
+bacteria, we did not expect that SDD received much gene transfer from either
+SPY1 or SPY2: the log base 2 ratio from SPY1 to SDD was -2.20, and that from
+SPY2 to SDD was -1.97. On the contrary, the recombing gene transfer from SDE1 to
+SDD was 0.756, and that from SDE2 to SDD was 0.698. This relatively higher
+recombining gene transfer from two SDE individuals to SDD compared with that
+from two SPY individuals to SDD reflected that the SDD was more closely related
+to SDE than to SPY in most of the genomes.
+
+The genomic organization between SDD and SDE1/SDE2 were much similar (Figures
+S\ref{fig:mauve}), and they formed a clade in the species tree
+(S\ref{fig:tree5}). However, the gene transfer from SDD to either SDE1 or SDE2
+was less frequent than the case of reverse direction; SDD served as a recipient
+rather than a donor. It is not unclear why SDD received more genes from SDE1 or
+SDE2 than did SDE1 or SDE2 from SDD.
+
+The recombining gene transfer log base 2 ratios between SDE1 and SDE2 (1.17 and
+1.20) were about twice as large as that from SDE1 and SDE2 to SPY1 and SPY2
+(0.67, 0.66, 0.60, and 0.39).  The intensities of recombining gene transfer from
+external branches of SDE to external branches to SPY was similar to those from
+SDE1 or SDE2 to SDD (0.76 and 0.70).  Either SPY1 or SPY2, and SDD appeared to
+be no different from the view point of SDE1 or SDE2 because of the similar
+values of recombining gene transfer intensity from SDE1 or SDE2 to the other
+three external branches.  It was expected that the recombining gene transfer
+within a clade would be larger than that between clades.  Surprisingly, the
+recombining gene transfer log base 2 ratios between SPY1 and SPY2 (1.06 and
+1.07) was comparable with those from SPY1 and SPY2 to SDE1 and SDE2 (1.25, 1.25,
+1.17, and 1.29).  Effectively, the recombining gene transfer from a clade SPY1
+or SPY2 to a different clade SDE1 or SDE2 was no different from that within the
+clade of SPY1 and SPY2. The two external branches of SDE received genes from
+either SPY1 or SPY2 via recombining gene transfer as if SDE1 and SDE2 were in
+the same clade as SPY1 and SPY2.  The bacterial recombination between SPY and
+SDE might happen more often than that between SDE and SDD whereas the genome
+configurations of SDE and SDD are more similar than each to SPY. This might be
+consistent with the fact that SPY and SDE shared the same ecological niche.  The
+gene transfer log base 2 ratio from SPY1 and SPY2 to SDE branch was larger than
+any type of gene transfer: 1.88 and 1.94, respectively.  Furthermore, we found a
+total of 58 genes for which a recombinant edge of certain types were sampled
+with probability at least 0.9 by focusing on specific types of recombinant
+edges.  (Table S\ref{tab:genes-transfer}).  Many of the genes in the list were
+transferred from SPY to SDE rather than the reverse direction overall.  All
+things considered, this clearly showed that the recombining gene transfer was
+biased toward the net direction from SPY to SDE.
+
+Figure \ref{fig:hgt-heatmap} shows the number of inferred non-recombining gene
+transfer events between the different branches of the species tree.  SDE1 and
+SDE2 shared the most events (26 and 34) compared to other sample pairs.  SDD
+appeared to acquire many genes from the ancestral SDE branch. Transfer events
+between the SPY and SDE clades were inferred in smaller numbers (<6) compared to
+those within each clade (23--34). Among the clade-crossing transfers, we did not
+see any significant enrichment for a specific direction, as opposed to the
+result for recombining transfers from the model-based analysis (Figure
+S\ref{fig:mowgli-recomb-heatmap}).
+
+\subsection{More frequent recombining gene transfer than horizontal gene transfer}
 Among the 2314 gene families
@@
 transfers.
+
+\subsection{Virulence genes via recombining gene transfer}
+
+\texttt{Find cases of any virulence genes that were transferred via horizontally.} 
+\texttt{Make a list of virulence genes and their categories with more interpretations.}
+
+About 10\% of the genes comprised of virulence facters, about half of which were
+found in the core genome of the five genome alignment. Using the UCSC genome
+browser tracks we located virulence genes with notable posterior recombination
+probability.  Stressful host conditions include heat shock, starvation, and iron
+depletion.  Iron uptake of bacteria was known to be essential for survival of
+the bacteria under hostile host condition. We found that parts of the gene,
+putative calcium transporter (MgtA), in SDE1 had been transferred from SPY1
+(SpyM3\_0440).  An extracellular protein, putative dipeptidase (SpyM3\_0465),
+was found to be transferred from SPY to SDE. The gene as a whole was
+transferred, which was a case for the finding that virulence genes tend to be
+transferred as a whole unit \citep{Chan2009}. Dipeptidase may be important for
+the survival of bacteria to scavenge dipeptides from the host.  An adhesins
+gene, putative internalin A precursor (SpyM3\_1035), was found to be affected by
+recombining gene transfer from SPY to SDE. Some of the middle parts of the gene
+was affected.  A middle part of a chaperonin GroEL, putative heat shock protein
+(SpyM3\_1765), was inferred to transfer from SPY to SDE1.  Increase of GroEL
+expression was known to allow for the bacteria to survive the hostile host
+environment of acid shock.  Superoxide dismutase was reported to be associated
+virulence of \textit{Streptococcus agalactiae} \citep{Poyart2001}. This whole
+gene including its 5' UTR, (SpyM3\_1071), was affected by recombining gene
+transfer from SPY to SDE.  Pyruvate formate lyase was recently reported to have
+virulence \citep{Yesilkaya2009}. A few parts of the gene (SpyM3\_1749) seemed to
+be affected by recombining gene transfer.  One MLST locus, called DNA mismatch
+repair protein MutS, appeared to be affected by recombining gene transfer.  A
+putative spermidine/putrescine ABC transporter ATP-binding protein,
+(SpyM3\_0764), was found to be transferred from SPY to SDE.  Under starvation
+condition bacteria must use more effective energy metabolism.  A putative
+glucose kinase, (SpyM3\_1180), might be transferred from SPY to SDE to adapt
+SDE's energy metabolism to survive such harsh condition.
+
+\subsection{Horizontal gene transfer is associated with virulence genes}
+\texttt{Select genes with high horizontal gene transfers. Relate these to
+categories of virulence genes. I should find many more virulence genes in Matt's
+analysis.}
+
+\subsection{Comparison between model-based and parsimony-based approaches}
+\texttt{I might want to analyze only core genomes and pan genomes
+separately. By separating data set of 6-gene families from the other families,
+we could have more complete comparison.}
+
+\subsection{UCSC genome browser for the dataset}
+We created a web-resource for gene transfer available at
+\url{http://strep-genome.bscb.cornell.edu}.  For each of the five individual
+genomes as a reference genome several annotation tracks were displayed:
+alignment blocks, recombination rate per site for each block, local gene tree
+topologies, recombination
+intensity, locations of virulence genes, posterior probability of recombining
+gene transfer.
+
+\subsection{Transport-related genes are associated with recombining gene transfer}
 
@@
 
-\clearpage{}
-
 \section{Discussions}
@@
 downloaded the genome (NC\_012471) of \textit{S.\ equi} ssp.\textit{\ equi}
-strain 4047 referred to as SEE.
+strain 4047 referred to as SEE \citep{Holden2009}.
 
@@
 parameters were medians weighted over block lengths, which down-weighed less
-reliable estimates obtained from shorter alignment blocks.  The second stage of
+reliable estimates obtained from shorter alignment blocks.
+
+In the first stage of \texttt{ClonalOrigin} MCMC a few alignment blocks were not
+successfully finished even after a month of computation.  Because those were not
+expected to affect the global estimates of population parameters significantly,
+we used only finished results of the first stage of \texttt{ClonalOrigin} to
+estimate mutation rate, recombination rate, and recombinant tract length.  Two
+independent runs of the first stage MCMC resulted in similar estimates of the
+three population parameters (Table S\ref{tab:three}).
+
+The second stage of
 MCMC with \texttt{ClonalOrigin} was executed with the three population
@@
 
-% Figure 1. A Recombinant Tree of ClonalOrigin Model
-\begin{figure}
-\includegraphics[scale=0.42]{figures/mauve-analysis-002}
-\caption{\label{fig:clonalorigin}Recombinant tree in the \texttt{ClonalOrigin} model.
-(A) A recombinant tree relating the five genomes with three recombinant edges.
-(B) A genomic region of size 100 base pairs with subregions affected by the
-three recombinant edges. Genomic positions are evenly marked by ticks every 
-10-th base pair. Local gene trees induced by the recombinant tree are
-illustrated for each region covered by some combinations of recombinant edges
-along the genomic region (C-G). For example, the local gene tree (E) is created
-by all of the three recombinant edges in the subregion between 51 base pairs
-and 70 base pairs.}
-\end{figure}
-\clearpage{}
-
-% Figure 2. Heat maps.
+%=============================================================================
+% Figure 1. Heat maps of recombining gene transfer (Add horizontal gene transfer)
 \begin{figure}
@@
 \clearpage{}%
-
-% Figure 3. The three main parameters.
+%=============================================================================
+
+%=============================================================================
+% FIGURE 2. - Transfer heatmap
 \begin{figure}
-\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot theta}}
-
-\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot rho}}
-
-\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot delta}}
-\caption{\label{fig:scatter3}Scatter plots of mutation rate, recombination
-rate, and average tract length for the 270 blocks. The red dashed
-line is the global median of all the blocks.}
+\begin{center}
+\includegraphics[width=5in]{figures/trans-hgt-heatmap}
+\end{center}
+\vspace{-1.2in}
+\caption{Distribution of horizontal gene transfers.}
+\label{fig:hgt-heatmap}
 \end{figure}
-\clearpage{}%
+%=============================================================================
 
@@
 
-% Figure 5.
-%=============================================================================
-% FIGURE - Transfer heatmap
-\begin{figure}
-\begin{center}
-\includegraphics[width=5in]{figures/trans-hgt-heatmap}
-\end{center}
-\vspace{-1.2in}
-\caption{Distribution of horizontal gene transfers.}
-\label{fig:hgt-heatmap}
-\end{figure}
-%=============================================================================
-
 \clearpage{}\setcounter{figure}{0}
@@
 \clearpage{}
+
+%=============================================================================
+% Table S1. List of genomes
+\begin{table}
+\caption{\label{tab:genome}The set of complete bacterial genomes}
+\noindent \begin{centering}
+\begin{tabular}{cccc}
+\hline 
+Name & NCBI accession & Size (base pairs) & Reference\tabularnewline
+\hline
+SDE1 & CP002215 & 2159491 & Suzuki et al. (2011)\tabularnewline
+SDE2 & NC\_012891 & 2106340 & Shimomura et al. (2011)\tabularnewline
+SDD & N/A & 2141837 & Suzuki et al. (2011)\tabularnewline
+SPY1 & NC\_004070 & 1900521 & Beres et al. (2002)\tabularnewline
+SPY2 & NC\_008024 & 1937111 & Beres et al. (2006)\tabularnewline
+SEE & NC\_012471 & 2253793 & Holden et al. (2009)\tabularnewline
+\hline
+\end{tabular}
+\par\end{centering}
+\end{table}
+\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% Table S2. ClonalFrame Result.
+\begin{table}
+\caption{\label{tab:clonalframe}Species tree inference using ClonalFrame.}
+\noindent \centering{}\begin{tabular}{cc}
+Parameter & Estimates\tabularnewline
+\hline
+$\theta$ & 92871.84 (N/A)\tabularnewline
+$\nu$ & 0.134 (0.132, 0.135)\tabularnewline
+$R$ & 9370 (8950, 9780)\tabularnewline
+$\delta$ & 362 (347, 377)\tabularnewline
+$r/m$ & 4.66 (4.45, 4.87)\tabularnewline
+$\rho/\theta$ & 0.101 (0.096, 0.105)\tabularnewline
+TMRCA & 0.331 (0.324, 0.337)\tabularnewline
+$T$ & 1.02 (1.01, 1.04)\tabularnewline
+\hline
+\end{tabular}
+\end{table}
+\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% Table S3. ClonalOrigin's 1st stage result.
+\begin{table}
+\caption{\label{tab:three}Three population parameter estimates from
+ClonalOrigin. 
+The estimates of the three parameters were based on one of two independent
+MCMC chains.  We considered the similar 
+results of the two independent chains as confirmation of 
+the convergence of the first stage of \texttt{ClonalOrigin} MCMC.
+}
+\noindent \centering{}\begin{tabular}{ccc}
+& Replicate 1 & Replicate 2\tabularnewline
+Parameter & Estimate (IQR$^a$) & Estimate (IQR)\tabularnewline
+\hline
+$\theta^b$ & 0.081 (0.067, 0.094) & 0.081 (0.067, 0.099)\tabularnewline
+$\rho^c$ & 0.012 (0.006, 0.019) & 0.012 (0.006, 0.020)\tabularnewline
+$\delta^d$ & 744 (346, 2848) & 723 (348, 2870)\tabularnewline
+\end{tabular}
+
+$^a$ IQR stands for interquartile range.
+
+$^b$ The median of mutation rate per site weighted over the lengths of blocks.
+
+$^c$ The median of recombination rate per site weighted over the lengths of blocks.
+
+$^d$ The median value for the recombinant tract length weighted over the lengths of blocks.
+
+\end{table}
+\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% Table S4. The values of the heat map.
+\begin{table}
+\caption{\label{tab:heatmap}}
+\noindent \centering{}\begin{tabular}{ccccccccc}
+0.176  &   NA  &1.170  &    NA & 0.7564  & NA & 0.67 &    NA & 0.663\tabularnewline
+-0.118 & 0.081 &-0.202 &     NA&  0.3395 &  NA& -0.15&  0.045& -0.323\tabularnewline
+1.201  &   NA  &0.092  &    NA & 0.6978  & NA & 0.60 &    NA & 0.387\tabularnewline
+-0.629 & 0.022 &-0.661 &0.00670&  0.0078 &  NA& -0.94& -0.299& -1.030\tabularnewline
+-0.317 & 0.371 &-0.197 &     NA&  0.0752 &  NA& -1.06& -0.525& -0.915\tabularnewline
+-1.745 &-0.698 &-1.786 &0.23039& -0.4854 &  NA& -1.37&  0.497& -1.306\tabularnewline
+1.245  &1.883  &1.251  &    NA &-2.1997  & NA & 0.06 &    NA & 1.062\tabularnewline
+0.089  &0.200  &0.298 &0.00098 &-0.7903  & NA &-0.19 &-0.058 &-0.159\tabularnewline
+1.174  &1.938  &1.287  &    NA &-1.9740  & NA & 1.07 &    NA & 0.063\tabularnewline
+\end{tabular}
+\end{table}
+\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% Table S5. The values of the heat map of actual counts.
+\begin{table}
+\caption{\label{tab:obsheatmap}}
+\noindent \centering{}\begin{tabular}{ccccccccc}
+8.896 &  0.00 &17.712  & 0.00 & 12.872   & 0 &12.59  &  0.00  &12.53\tabularnewline
+44.626& 136.49& 42.088 &  0.00& 227.061  &  0& 73.39 &  99.95 & 65.25\tabularnewline
+18.099&   0.00&  8.393 &  0.00&  12.360  &  0& 11.97 &   0.00 & 10.36\tabularnewline
+9.671 &101.74 & 9.464  &31.26 &115.796   & 0 &13.68  & 98.38  &12.86\tabularnewline
+45.539& 168.58& 49.513 &  0.00& 194.706  &  0& 43.09 &  67.14 & 47.65\tabularnewline
+43.553& 602.67& 42.337 &765.53& 802.813  &  0& 98.86 &2146.01 &103.60\tabularnewline
+40.800&  11.46& 40.973 &  0.00&   4.371  &  0& 21.15 &   0.00 & 42.35\tabularnewline
+57.618& 262.08& 66.594 & 30.92& 163.518  &  0& 83.56 & 207.00 & 85.10\tabularnewline
+38.837&  11.90& 42.000 &  0.00&   5.111  &  0& 42.45 &   0.00 & 21.19\tabularnewline
+\end{tabular}
+\end{table}
+\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% Table SXXX. Simulation study for the three population parameters.
+\begin{table}
+\caption{\label{tab:sim-three-population-parameters}
+Simulation study of the three population parameters.}
+\noindent \begin{centering}
+\begin{tabular}{ccc}
+ & True & Estimates (Standard Deviation)\tabularnewline
+\hline
+Mutation rate per site & 0.081 & 0.081 (0.0011)\tabularnewline
+Recombination rate per site & 0.012 & 0.011 (0.0003)\tabularnewline
+Recombinant tract length & 744 bp & 919 bp (28)\tabularnewline
+\end{tabular}
+\par\end{centering}
+\end{table}
+\clearpage{}
+%=============================================================================
 
@@
 
-% Table S3. Simulation study for the three population parameters.
-\begin{table}
-\caption{\label{tab:sim-three-population-parameters}
-Simulation study of the three population parameters.}
-\noindent \begin{centering}
-\begin{tabular}{ccc}
- & True & Estimates (Standard Deviation)\tabularnewline
-\hline
-Mutation rate per site & 0.081 & 0.081 (0.0011)\tabularnewline
-Recombination rate per site & 0.012 & 0.011 (0.0003)\tabularnewline
-Recombinant tract length & 744 bp & 919 bp (28)\tabularnewline
-\end{tabular}
-\par\end{centering}
-\end{table}
+\section*{Supplementary Figures}
+
 \clearpage{}
 
-% Table S4. ClonalFrame Result.
-\begin{table}
-\caption{\label{tab:clonalframe}Species tree inference using ClonalFrame.}
-\noindent \centering{}\begin{tabular}{cc}
-Parameter & Estimates\tabularnewline
-\hline
-$\theta$ & 92871.84 (N/A)\tabularnewline
-$\nu$ & 0.134 (0.132, 0.135)\tabularnewline
-$R$ & 9370 (8950, 9780)\tabularnewline
-$\delta$ & 362 (347, 377)\tabularnewline
-$r/m$ & 4.66 (4.45, 4.87)\tabularnewline
-$\rho/\theta$ & 0.101 (0.096, 0.105)\tabularnewline
-TMRCA & 0.331 (0.324, 0.337)\tabularnewline
-$T$ & 1.02 (1.01, 1.04)\tabularnewline
-\end{tabular}
-\end{table}
+%=============================================================================
+% Figure S1: Mauve alignment
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{figures/mauve.pdf}
+\end{center}
+\vspace{-1.2in}
+\caption{Alignment of the five genomes.}
+\label{fig:mauve}
+\end{figure}
 \clearpage{}
-
-% Table S5. ClonalOrigin's 1st stage result.
-\begin{table}
-\caption{\label{tab:three}Three population parameter estimates from
-ClonalOrigin. 
-The estimates of the three parameters were based on one of two independent
-MCMC chains.  We considered the similar 
-results of the two independent chains as confirmation of 
-the convergence of the first stage of \texttt{ClonalOrigin} MCMC.
-}
-\noindent \centering{}\begin{tabular}{ccc}
-& Replicate 1 & Replicate 2\tabularnewline
-Parameter & Estimate (IQR$^a$) & Estimate (IQR)\tabularnewline
-\hline
-$\theta^b$ & 0.081 (0.067, 0.094) & 0.081 (0.067, 0.099)\tabularnewline
-$\rho^c$ & 0.012 (0.006, 0.019) & 0.012 (0.006, 0.020)\tabularnewline
-$\delta^d$ & 744 (346, 2848) & 723 (348, 2870)\tabularnewline
-\end{tabular}
-
-$^a$ IQR stands for interquartile range.
-
-$^b$ The median of mutation rate per site weighted over the lengths of blocks.
-
-$^c$ The median of recombination rate per site weighted over the lengths of blocks.
-
-$^d$ The median value for the recombinant tract length weighted over the lengths of blocks.
-
-\end{table}
-\clearpage{}
-
-\section*{Supplementary Figures}
-
-\clearpage{}
-
-% Figure S1: Posterior probability of recobinant edges on aspartyl-tRNA.
+%=============================================================================
+
+%=============================================================================
+% Figure S2: Species tree
 \begin{figure}
-\includegraphics[scale=0.6]{figures/yl-trna}
+\includegraphics[scale=0.7]{figures/cornellf-3-tree}
+\caption{\label{fig:tree5}Species tree of five taxa for \texttt{ClonalOrigin} analysis.
+SPY is the branch of a clade\textit{ of Streptococcus pyogenes}, SDE
+is that for \textit{S.\ dysgalactiae} ssp.\textit{\ equisimilis}, and SD is the
+branch for the three individuals of \textit{S.\ dysgalactiae}. The scale is
+proportional to the expected number of mutations given the Watterson's estimate 
+of mutation
+rate $\theta$ and the tree.}
+\end{figure}
+\clearpage{}%
+%=============================================================================
+
+%=============================================================================
+% Figure S3: Histogram of block sizes
+\begin{figure}
+\begin{center}
+\includegraphics[width=5in]{figures/blocksize.pdf}
+\end{center}
+\vspace{-.3in}
+\caption{Distribution of core alignment block sizes.}
+\label{fig:blocksize}
+\end{figure}
+\clearpage{}%
+%=============================================================================
+
+%=============================================================================
+% Figure S4. Three population parameters.
+\begin{figure}
+\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot theta}}
+
+\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot rho}}
+
+\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot delta}}
+\caption{\label{fig:scatter3}Scatter plots of mutation rate, recombination
+rate, and average tract length for the 270 blocks. The red dashed
+line is the global median of all the blocks.}
+\end{figure}
+\clearpage{}%
+%=============================================================================
+
+%=============================================================================
+% Figure S5. A Recombinant Tree of ClonalOrigin Model
+\begin{figure}
+\includegraphics[scale=0.42]{figures/mauve-analysis-002}
+\caption{\label{fig:clonalorigin}Recombinant tree in the \texttt{ClonalOrigin} model.
+(A) A recombinant tree relating the five genomes with three recombinant edges.
+(B) A genomic region of size 100 base pairs with subregions affected by the
+three recombinant edges. Genomic positions are evenly marked by ticks every 
+10-th base pair. Local gene trees induced by the recombinant tree are
+illustrated for each region covered by some combinations of recombinant edges
+along the genomic region (C-G). For example, the local gene tree (E) is created
+by all of the three recombinant edges in the subregion between 51 base pairs
+and 70 base pairs.}
+\end{figure}
+%=============================================================================
+
+%=============================================================================
+% Figure XXX: View of UCSC genome browser for bacterial gene transfer.
+\begin{figure}
+\includegraphics[scale=0.6]{figures/ucsc}
 \caption{\label{fig:muts}Posterior probability of recombinant edges on 
@@
 
-% Figure S5: The species tree
-\begin{figure}
-\includegraphics[scale=0.7]{figures/cornellf-3-tree}
-\caption{\label{fig:tree5}Species tree of five taxa for \texttt{ClonalOrigin} analysis.
-SPY is the branch of a clade\textit{ of Streptococcus pyogenes}, SDE
-is that for \textit{S.\ dysgalactiae} ssp.\textit{\ equisimilis}, and SD is the
-branch for the three individuals of \textit{S.\ dysgalactiae}. The scale is
-proportional to the expected number of mutations given the Watterson's estimate 
-of mutation
-rate $\theta$ and the tree.}
-\end{figure}
-\clearpage{}%
-
-% Figure S6: Histogram of block sizes
-\begin{figure}
-\begin{center}
-\includegraphics[width=5in]{figures/blocksize.pdf}
-\end{center}
-\vspace{-.3in}
-\caption{Distribution of core alignment block sizes.}
-\label{fig:blocksize}
-\end{figure}
-\clearpage{}%
-
 % Figure S7: gene family sizes
@@
 \end{figure}
-%=============================================================================
-
-% Figure S9: Mauve alignment
-%=============================================================================
-% FIGURE - Mauve alignment
-\begin{figure}
-\begin{center}
-\includegraphics[width=5in]{figures/mauve.pdf}
-\end{center}
-\vspace{-1.2in}
-\caption{Alignment of the five genomes.}
-\label{fig:mauve}
-\end{figure}
+\clearpage{}
 %=============================================================================
@@
 % between species.
+
+% block 76: SDEG\_0641
+% ORTHOMCL43-1    SEQ\_1497        SpyM3\_0440      MGAS10750\_Spy0535 SDD27957\_03210  SDE12394\_03445  SDEG\_0641
+% Not found in Davies2007
+% putative calcium transporter
+% MgtA
+
+% block 97: SDEG\_0673
+% ORTHOMCL1244-1  SEQ\_0860        SDD27957\_03385  MGAS10750\_Spy0623 SpyM3\_0465      SDE12394\_03650  SDEG\_0673
+% Putative (put.) extracellular proteins: put. dipeptidase (SpyM3\_0465)
+% Dipeptiase for the same human host might be important for SDE to survive human
+% host environment. Scarvenge dipeptides.
+
+% block 101: SDEG\_0739
+% ORTHOMCL1207-1  SDD27957\_04880  SEQ\_0932        SDD27957\_03660  SDEG\_0739 SDE12394\_03975  MGAS10750\_Spy0677       SpyM3\_0511
+
+% block 151: SDEG\_1447
+% ORTHOMCL99-1	SEQ\_0680	SDD27957\_06980	SDEG\_1447	SDE12394\_07540	SpyM3\_1071	MGAS10750\_Spy1252
+% Not found in Davies2007
+% superoxide dismutase
+% sodM
+% Poyart2001
+% Contribution of Mn-Cofactored Superoxide Dismutase (SodA) to the Virulence of Streptococcus agalactiae.
+
+% block 244: SDEG\_2040
+% ORTHOMCL37-1	SEQ\_0209	SpyM3\_1765	MGAS10750\_Spy1855	SDD27957\_10625	SDE12394\_10325	SDEG\_2040
+% Not found in Davies2007
+% chaperonin GroEL, 
+% putative heat shock protein
+% What is the role of this GroEL?
+% Increase of GroEL in acid shock, which might be important for survival of the
+% bacteria in human hosts.
+
+% block 242: SDD27957\_00955
+% ORTHOMCL730-1	SDE12394\_10000	SEQ\_2112	SDD27957\_00955	SDE12394\_00680	SDEG\_0169	SpyM3\_1755	MGAS10750\_Spy1843
+% Not found in Davies2007
+% putative pyruvate formate-lyase activating enzyme
+% There are some upstream parts were replaced.
+
+% block 240: SDD27957\_00985
+% ORTHOMCL731-1	SDE12394\_10005	SEQ\_2106	SDD27957\_00985	SpyM3\_1749	MGAS10750\_Spy1837	SDE12394\_00710	SDEG\_0175
+% Not found in Davies2007
+% putative pyruvate formate-lyase
+% pflD
+% Yesilkaya2009
+% Pyruvate Formate Lyase Is Required for Pneumococcal Fermentative Metabolism and Virulence
+% SDE2 received from SPY1 via recombining gene transfer.
+
+% block 167: SDD27957\_06790
+% ORTHOMCL1139-1	SEQ\_0705	SpyM3\_1035	MGAS10750\_Spy1208	SDD27957\_06790	SDE12394\_07350	SDEG\_1372
+% Not found in Davies2007
+% putative internalin A precursor
+% Adhesins
+
+% block 255: MGAS10750\_Spy1922
+% ORTHOMCL158-1	SEQ\_2161	SDD27957\_10925	SDE12394\_10560	SDEG\_2091	SpyM3\_1806	MGAS10750\_Spy1922
+% Not found in Davies2007
+% DNA mismatch repair protein MutS
+% What makes MutS as a virulence genes?
+
+% block 67: SDD27957\_03085
+% ORTHOMCL121-1	SEQ\_1525	SDD27957\_03085	SpyM3\_0420	SDE12394\_03265	SDEG\_0610	MGAS10750\_Spy0514
+% Not found in Davies2007
+% lysyl-tRNA synthetase
+
+% block 133: SDD27957\_04505
+% ORTHOMCL58-1	SEQ\_1129	SDD27957\_04505	SDE12394\_04895	SDEG\_0866	SpyM3\_0764	MGAS10750\_Spy0977
+% Not found in Davies2007
+% putative spermidine/putrescine ABC transporter ATP-binding protein
+
+% List of virulence genes.
+% Genes for the Majority Of Group A Streptococcal Virulence Factors and
+% Extracellular Surface Proteins Do Not Confer an Increased Propensity to Cause
+% Invasive Disease
+
+% block 170: SDD27957\_07725
+% ORTHOMCL96-1	SEQ\_0613	SDD27957\_07725	SDEG\_1515	SDE12394\_07910	SpyM3\_1180	MGAS10750\_Spy1365
+% Not found in Davies2007
+% putative glucose kinase
+
+% Evidence for Reductive Genome Evolution and Lateral Acquisition of Virulence
+% Functions in Two Corynebacterium pseudotuberculosis Strains
+
+% block 186: MGAS10750\_Spy1444
+% ORTHOMCL93-1	SEQ\_1774	SDD27957\_08550	SDE12394\_08460	SDEG\_1691	SpyM3\_1369	MGAS10750\_Spy1444
+% Not found in Davies2007
+% putative serine/threonine kinase
+
+% Blocks with virulence genes.
+% Check virulence genes (they were combined) at 
+% SDD27957\_02890, SDD27957\_02895
+% MGAS10750\_Spy0455, MGAS10750\_Spy0456 (block 75)

### 2011-10-03 sh/choose-simulation.sh

@@ -1,6 +1,22 @@
+###############################################################################
+# Copyright (C) 2011 Sang Chul Choi
+#
+# This file is part of Mauve Analysis.
+# 
+# Mauve Analysis is free software: you can redistribute it and/or modify
+# it under the terms of the GNU General Public License as published by
+# the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# Mauve Analysis is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with Mauve Analysis.  If not, see <http://www.gnu.org/licenses/>.
+###############################################################################
 
-# 0. For simulation I make directories in CAC and copy genomes files to the data
-# directory. 
 function choose-simulation {
   PS3="Choose the simulation for clonalorigin: "
   select SPECIES in ${SIMULATIONS[@]}; do 
@@ -37,6 +53,8 @@ function mkdir-simulation {
   echo -e " done"
   mkdir $OUTPUTDIR/$1
   mkdir $OUTPUTDIR/$1/run-analysis
+  return 0
+
   echo -n "  Creating a simulation $1 at $CAC_OUTPUTDIR in $CAC_USERHOST ..."
   ssh -x $CAC_USERHOST mkdir $CAC_OUTPUTDIR/$1
   echo -e " done"
@@ -65,6 +83,8 @@ function mkdir-simulation-repeat {
         $RUNCLONALORIGIN \
         $RUNANALYSIS
   echo -e " done"
+  return 0
+
   CAC_BASEDIR=$CAC_OUTPUTDIR/$1/$2
   CAC_DATADIR=$CAC_BASEDIR/data
   CAC_RUNMAUVE=$CAC_BASEDIR/run-mauve

### 2011-10-03 sh/map-tree-topology.sh

@@ -36,6 +36,7 @@ function map-tree-topology {
           -numberblock $NUMBER_BLOCK \
           -endblockid
         echo " done."
+
       else
         echo "  Skipping splitting ClonalOrigin xml output files..."
       fi

### 2011-10-03 sh/run.sh

@@ -81,6 +81,7 @@ source sh/analyze-run-clonalorigin2-simulation2.sh
 source sh/sim3-prepare.sh 
 source sh/sim3-receive.sh 
 source sh/sim3-analyze.sh 
+source sh/sim5-prepare.sh 
 source sh/create-ingene.sh
 source sh/convert-gff-ingene.sh 
 source sh/locate-gene-in-block.sh
@@ -131,6 +132,8 @@ CHOICES=( init-file-system \
           sim3-prepare \
           sim3-receive \
           sim3-analyze \
+          ---SIMULATION5---\
+          sim5-prepare \
           # ---SIMULATION4---\
           # clonalorigin2-simulation3 \
           # sim4-prepare \
@@ -229,6 +232,7 @@ select CHOICE in ${CHOICES[@]}; do
   elif [ "$CHOICE" == "sim3-prepare" ]; then $CHOICE; break
   elif [ "$CHOICE" == "sim3-receive" ]; then $CHOICE; break
   elif [ "$CHOICE" == "sim3-analyze" ]; then $CHOICE; break
+  elif [ "$CHOICE" == "sim5-prepare" ]; then $CHOICE; break
   elif [ "$CHOICE" == "create-ingene" ]; then $CHOICE; break
   elif [ "$CHOICE" == "convert-gff-ingene" ]; then $CHOICE; break
   elif [ "$CHOICE" == "locate-gene-in-block" ]; then $CHOICE; break

### 2011-10-03 sh/sim5-prepare.sh

@@ -1,0 +1,99 @@
+###############################################################################
+# Copyright (C) 2011 Sang Chul Choi
+#
+# This file is part of Mauve Analysis.
+# 
+# Mauve Analysis is free software: you can redistribute it and/or modify
+# it under the terms of the GNU General Public License as published by
+# the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# Mauve Analysis is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with Mauve Analysis.  If not, see <http://www.gnu.org/licenses/>.
+###############################################################################
+
+function sim5-prepare {
+  PS3="Choose the species to do $FUNCNAME: "
+  select SPECIES in ${SIMULATIONS[@]}; do 
+    if [ "$SPECIES" == "" ];  then
+      echo -e "You need to enter something\n"
+      continue
+    elif [ "$SPECIES" == "s17" ] \
+         || [ "$SPECIES" == "sxx" ]; then
+      read-species
+
+      BASEDIR=$OUTPUTDIR/$SPECIES
+      BASERUNANALYSIS=$BASEDIR/run-analysis
+      NUMBER_BLOCK=$(grep NUMBER_BLOCK $SPECIESFILE | cut -d":" -f2)
+      NUMBER_REPLICATE=$(grep NUMBER_REPLICATE $SPECIESFILE | cut -d":" -f2)
+      THETA_PER_SITE=$(grep ThetaPerSite $SPECIESFILE | cut -d":" -f2)
+      XMLBASEDIR=$(grep XMLBASEDIR $SPECIESFILE | cut -d":" -f2)
+      HOW_MANY_REPETITION=$(grep Repetition $SPECIESFILE | cut -d":" -f2)
+      HOW_MANY_REPLICATE=$(grep Replicate $SPECIESFILE | cut -d":" -f2)
+      TREE=$(grep SpeicesTree $SPECIESFILE | cut -d":" -f2)
+
+
+      echo -n "Do you wish to simulate data using recombinant trees (y/n)? "
+      read WISH
+      if [ "$WISH" == "y" ]; then
+        echo -e "  Simulating data using recombinant trees..."
+
[…]

### 2011-10-03 src/clonalorigin/rectree.cpp

@@ -34,6 +34,9 @@ RecTree::RecTree(unsigned int numsites,string newick,bool isFilename,bool forcea
     edge=vector<RecEdge*>(0);
 }
 
+/**
+ * Note that this constructor already exists.
+ */
 RecTree::RecTree(unsigned int numsites,WargXml *infile,bool addedges,bool forceages,bool isfinal,std::streampos itstart):Tree(infile->getTree(itstart,isfinal),false,forceages)
 {
     tabNode=vector<int>(n*2-1);

### 2011-10-03 src/clonalorigin/weakargsim.cpp

@@ -2,6 +2,27 @@
  * The main source file for simulating under ClonalOrigin model.
  * This is the main source file for the project of simulating data under
  * ClonalOrigin model.
+ *
+ * 1. Simulate data for a given set of species tree, rho, delta, and theta:
+      --cmd-sim-given-tree
+      --tree-file Newick_Tree_File
+      --block-file Block_Length_File
+      --out-file OUTFILE_PREFIX
+      --number-data NUMBER_REPLICATES
+      -T s$THETA_PER_SITE -D $DELTA -R s$RHO_PER_SITE
+ *
+ * 2. Extract gene tree topologies:
+      --cmd-extract-tree (or --gene-tree : this may be obsolete)
+      --xml-file ri/core_co.phase3.xml.blockID.replicateID
+      --out-file ri-out/core_co.phase3.xml.blockID.replicateID
+      --block-length BLOCK_LENGTH
+ *
+ * 3. Simulate data for a given recombinant tree and theta:
+      --cmd-sim-given-rectree
+      --xml-file ri/core_co.phase3.xml.blockID.replicateID
+      -T s$THETA_PER_SITE
+      --number-data NUMBER_REPLICATES
+      --out-file core_alignment
  */
 #ifdef HAVE_CONFIG_H
 #include "config.h"
@@ -98,6 +119,10 @@ static const char * help=
     --gene-tree\n\
                   Gene trees for all of iteration and site are exported.\n\
                   Not random!\n\
+    --cmd-sim-given-rectree\n\
+                  Simulate data given a recombinant tree and mutation rate.\n\
+    --cmd-sim-given-tree\n\
+                  Simulate data given a species tree, recombination, delta, mutation rate.\n\
     --block-length\n\
                   Length of a block (used with --gene-tree)\n\
     -o FILE       Export the data to the given file in XMFA format\n\
@@ -143,6 +168,8 @@ enum { OPT_HELP,
        OPT_XML_FILE,
        OPT_BLOCK_LENGTH,
        OPT_GENE_TREE,
+       OPT_CMD_SIM_GIVEN_TREE,
+       OPT_CMD_SIM_GIVEN_RECTREE,
        OPT_NUMBER_DATA,
        OPT_OUTPUT_FILE,
        OPT_CLONALTREE_FILE,
@@ -205,8 +232,12 @@ CSimpleOpt::SOption g_rgOptions[] = {
     {   OPT_GENE_TREE,
[…]

### 2011-10-04 doc/manuscript.tex
It is not clear how the manuscript is made.

@@
 opportunistic human pathogen.  We aims at which parts of the genomic
-organization, including human pathogenic virulence factors, of SDE could have
+organization and human pathogenic virulence factors of SDE could have
 originated in those of SPY, or vice versa.  Our genome-wide survey of gene
@@
 relatively from SPY to SDE.  Although we did not find enrichment of virulence
-genes in the homologous recombining gene transfer, we found evidence that they
-transferred more often via non-recombining fashion without replacing parts of
-destination genomes than did non-virulence genes.  A genome-wide resource of
-recombining and non-recombining gene transfer for the study of the genomes is
-provided so that research community in pathogenic bacteria could take advantage
-of it.
+genes in the homologous recombining gene transfer, we found evidence that
+virulence genes transferred more often via non-recombining fashion without
+replacing parts of destination genomes than did non-virulence genes.  A
+genome-wide resource of recombining and non-recombining gene transfer for the
+study of the genomes is provided so that research community in pathogenic
+bacteria could take advantage of it.
 
@@
 SPY for the downstream of the genomic region of streptococcal pyrogenic
-exotoxins \textit{speG}. \citet{Davies2007} investigated phage 3396 from SDE as
+exotoxins. \citet{Davies2007} investigated phage 3396 from SDE as
 a vehicle of gene transfer from SDE to SPY. \citet{Davies2009} demonstrated that
@@
 to SDE was more dominant than the other direction.  Because these researches on
-the controversial issue, directionality of gene transfer between SPY and SDE,
-have been limited to relatively small regions in the genomes primarily on
-representative loci from the Multi Locus Sequence Typing database, it remains to
-be elusive to determine the genome-wide degrees to which genomic segments flow
-between the two species, and net directions of thereof.
+the controversial issue, net directionality of gene transfer between SPY and
+SDE, have been limited to relatively small regions in the genomes primarily on
+representative loci from the Multi Locus Sequence Typing (MLST) database, it
+remains to be elusive to determine the genome-wide degrees to which genomic
+segments flowed between the two species, and net directions of thereof.
 
@@
 evolutionary histories of prokaryotes, with multiple individual genomes per
-species we now can go beyond the broad study on gene transfer towards
+species we now can go beyond the rough broad study on gene transfer towards
 fine-scaled history of gene transfer between very closely related prokaryotes.
@@
 genomes from SPY have been available, \citet{Suzuki2011} recently added one more
-complete genome of SDE to the previous set of a unique SDE genome. This set of
-SPY and SDE genomes prompted us to study genome-wide gene flows between the two
-species that afflicted human beings around the globe.
+complete genome of SDE to the previous set of a unique SDE genome
+\citep{Shimomura2011}. The multiple SDE genomes prompted us to study
+genome-wide gene flows between the two species that afflicted human beings
+around the globe.
 
 Bacteria evolves new biological features through horizontally transferred genes
-from other lineages and vertically transmitted mutations on genes within
-the same lineage.  Mutations on a gene modifiy relatively small portions of the
-gene through substitutions, insertions, and deletions of deoxyribonucleic
-acids. Horizontal gene transfers can reshape the organization of a bacterial
-genome, which could revamp biological functions of the bacterial species.
-Recipient bacteria of transferred genes can substitute transferred genes for homologous parts of
-their genomes, or integrate transferred genes into their genomes
-without replacement. These two processes are here in this paper discriminated by calling the
-former \textit{recombining gene transfer}, and the latter \textit{horizontal
-gene transfer} \citep{Ochman2001,Lawrence2009}.  Despite the two distinct
-evolutionary processes in bacteria, reseraches on statistical analyses of
-complete genome sequences in efforts of better understanding bacterial evolution
-have paid less attention to a combined approach to illuminating disparity of
-the two gene transfer processes. We attempted here to use and compare two
-approaches for distinguishing recombining gene transfer process from horizontal, or
-non-recombining, gene transfer one, and more systematically tackled their roles
+from other lineages and vertically transmitted mutations on genes within the
+same lineage.  Mutations on a gene modifiy relatively small portions of the gene
+through substitutions, insertions, and deletions of deoxyribonucleic acids.
+Horizontal gene transfers can reshape the organization of a bacterial genome,
+which could revamp biological functions of the bacterial species.  Recipient
+bacteria of transferred genes can substitute transferred genes for homologous
+parts of their genomes, or integrate transferred genes into their genomes
+without replacement. These two processes are here in this paper discriminated by
+calling the former \textit{recombining gene transfer}, and the latter
+\textit{horizontal gene transfer} \citep{Ochman2001,Lawrence2009}.  Despite the
+two distinct evolutionary processes in bacteria, analyses of complete genome
+sequences in efforts of better understanding bacterial evolution have paid less
+attention to a combined approach to illuminating disparity of the two gene
+transfer processes. We attempted here to use and compare two approaches for
+distinguishing recombining gene transfer process from horizontal, or
+non-recombining, gene transfer, and more systematically tackled their roles
 in association with virulence genes.
@@
 to homologous recombination, and the second could do similarly attributed to
-horizontal gene transfer. Although this heterogeneously combined approach could
-be unsatisfactory, we hoped to take the unique opportunity of new genomes to
-study gene transfer in genome-wide scale among SPY and SDE individual bacteria.
-
-In short, the complete genomes of the two spcies of SPY and SDE presented a unique
-opportunity to study gene transfer over relatively recent time scales.  We used
-both of the model-based approach for inferring homologous recombination and the
-parsimony-based one for detecting gene gain, loss, and horizontal transfer. We
-also compared both of the results in recombining gene transfer events to show
-that the parimony-based method inferred gene transfer events in similar
-biological functional categories to those inferred with the model-based
-approach. This is the first attempt as far as we know to analyaze SPY and SDE in
-genome-wide scale for detecting gene transfer events not only in the core genome
-but also in the pan genome, which make the analysis complete in terms of the
-amount of used genome data. Our analysis of core genomes also showed for the
-first time that the net directionality of, at least, recombining gene transfer
-from SPY to SDE was stronger than the reverse direction.
+horizontal gene transfer.
+
+In short, the complete genomes of the two spcies of SPY and SDE presented a
+unique opportunity to study gene transfer over relatively recent time scales.
+We used both of the model-based approach for inferring homologous recombination
+and the parsimony-based one for detecting gene gain, loss, and horizontal
+transfer. We also compared both of the results in recombining gene transfer
+events to show that the parimony-based method inferred recombining gene transfer
+events in similar biological functional categories to those inferred with the
+model-based approach. This was the first attempt as far as we know to analyaze
+SPY and SDE in genome-wide scale for detecting gene transfer events not only in
+the core genome but also in the pan genome, which make the analysis complete in
+terms of the amount of used genome data. Our analysis of core genomes also
+showed for the first time that the net directionality of, at least, recombining
+gene transfer from SPY to SDE was stronger than the reverse direction.
 
@@
 
-\subsection{Recombination-driven substitutions were stronger than mutation-driven ones}
-
-The genome alignment of the five genomes (Table S\ref{tab:genome}) showed two distinct
-genome structures of SPY and SDE (Figure S\ref{fig:mauve}). The SDE genomes were
-larger in length than those of SPY. Although it was unclear that SDD was
-different from either of SDE1 and SDE2 from the graphical view of the genome
-alignment, the spcies tree inferred from the genome alignment showed that SDD
-branch was attached to the clade branch of SDE1 and SDE2 (Figure
-S\ref{fig:tree5}). The distribution of the alignment block lengths was skewed to
-the right with mean of 4188 base pairs long (Figure S\ref{fig:blocksize}), which
-implied that the model assumption of geometric length distribution might be
-plausible in \texttt{ClonalOrigin} method (see section \textbf{Materials and
-Method} and \citealt{Didelot2010}).  The relative strength of recombination
-compared to mutations indicated that reconmination-driven substitutions affected
-more sites than did mutation-driven substitutions (Table
-S\ref{tab:clonalframe}). In the \textit{Streptococcus} bacteria homologous
-recombination appeared to be a stronger evolutionary force than point mutations 
-process.
+\subsection{Recombination dirved more substitutions than did point mutation}
+
+Five genomes except \textit{S.\ equi} ssp.\textit{\ equi} strain 4047 were
+chosen (Table S\ref{tab:genome}) to study recombining gene transfer.  The genome
+alignment of the five genomes showed two distinct genome structures of SPY and
+SDE (Figure S\ref{fig:mauve}). The SDE genomes were larger in length than those
+of SPY. Although it was unclear whether SDD was different from either of SDE1
+and SDE2 from the graphical view of the genome alignment, the spcies tree
+inferred from the genome alignment showed that SDD branch was attached to the
+clade branch of SDE1 and SDE2 (Figure S\ref{fig:tree5}). The distribution of the
+alignment block lengths was skewed to the right with mean of 4188 base pairs
+long (Figure S\ref{fig:blocksize}), which implied that the model assumption of
+geometric block length distribution might be plausible in \texttt{ClonalOrigin}
+method (see section \textbf{Materials and Method} and \citealt{Didelot2010}).
+The relative strength of recombination compared to mutations indicated that
+reconmination-driven substitutions affected more sites than did mutation-driven
+substitutions (Table S\ref{tab:clonalframe}). In the \textit{Streptococcus}
+bacteria homologous recombination appeared to affect more sites than did point
+mutations process.
 
@@
 (Figure S\ref{fig:scatter3}).  We found a few spots of genomic regions where the
-mutation rates were larger than twice the global estimate of mutation rate.
+mutation rates were larger than twice the global estimate of mutation rate.  The
+recombination rate appeared to change more dynamically. Espeically, the genomic
+region of SDE1 near 20 kilobase pairs from the origin of replication contained
+alignment blocks with bell-shaped recombination rates.  The recombinant tract
+length changes in a way that we could not note any particular positions along
+the genome.  There appeared to be negative correlation between recombinant tract
+length and recombination rate (Figure S\ref{fig:rhologdelta}).
 % \texttt{WHAT ARE THESE BLOCKS? WHAT GENES ARE LOCATED IN THE BLOCK?}
-The recombination rate appeared to change more dynamically. Espeically, the
-genomic region of SDE1 near 20 kilobase pairs from the origin of replication
-contains alignment blocks with bell-shaped recombination rates.
 % \texttt{USE UCSC GENOME BROWSER TO FIND WHAT GENES BELONG TO THOSE BLOCKS.}
-The recombinant tract length changes in a way that we could not note any
-particular positions along the genome.
-There appeared to be negative correlation between recombinant tract length and
-recombination rate.
-% \texttt{PLOT PAIRWISE SCATTER PLOTS OF THE THREE PARAMETERS.}
 
@@
 
-Most of local tree topologies (see section \textbf{Materials and Methods}), 
-or 66.7\%, along the alignment blocks were the
-same as the species tree, which allowed us to be more confident of the species
-tree inferred using the alignment blocks (Table
-S\ref{tab:Gene-tree-topologies}).  The second most probable topology (9.3\%)  was one
-where SDD was the outgroup, and SPY and SDE were sister clades 
-SPY and SDE appeared to share about tenth of the core
-genomes that they were more closely related with each other than each to SDD's
-genome. The other topologies had less than 4\% of proportions.
+Most of local gene tree topologies (see section \textbf{Materials and Methods}),
+or 66.7\%, along the alignment blocks were the same as the species tree, which
+allowed us to be more confident of the species tree inferred using the alignment
+blocks (Table S\ref{tab:Gene-tree-topologies}).  The second most probable
+topology (9.3\%) was one where SDD was the outgroup, and SPY and SDE were
+sister clades. SPY and SDE appeared to share about tenth of the core genomes that
+they were more closely related with each other than each to SDD's genome. The
+other topologies had less than 4\% of proportions.
 
@@
 
-The heat map in Figure \ref{fig:Heatmap-of-recombination} shows the log base 2
-ratios of the \textit{a posteriori} number of recombinant edges for a pair of
-donor and recipient species tree branches and their corresponding expected
-values given the species tree and the global parameter settings.  Because
-recombinant edges from any species tree branch to the root branch are not
-possible under \texttt{ClonalOrigin} model (see Figure S\ref{fig:clonalorigin}),
-the column for the root as recipient was colored green. Similarly, all of the
-cells where a recipient branch was older than the donor, and two branches did
-not share any contemporary time were colored green. The diagonal cells were
-colored gray, indicating no significant number of recombinant edges, because the
-observed number of recombinant edges and the expected one were similar. As
-recombinant edges starting from and ending in the same species tree branch would
-not change the species tree, which would not change the likelihood much, the
-cells should be colored gray.  Because SDD is a strictly veterinary pathogenic
-bacteria, we did not expect that SDD received much gene transfer from either
-SPY1 or SPY2: the log base 2 ratio from SPY1 to SDD was -2.20, and that from
-SPY2 to SDD was -1.97. On the contrary, the recombing gene transfer from SDE1 to
-SDD was 0.756, and that from SDE2 to SDD was 0.698. This relatively higher
-recombining gene transfer from two SDE individuals to SDD compared with that
-from two SPY individuals to SDD reflected that the SDD was more closely related
-to SDE than to SPY in most of the genomes.
-
-The genomic organization between SDD and SDE1/SDE2 were much similar (Figures
+The heat map of Figure \ref{fig:Heatmap-of-recombination} (see Table
+S\ref{tab:heatmap} for the values) shows the log base 2 ratios of the \textit{a
+posteriori} number of recombinant edges for a pair of donor and recipient
+species tree branches and their corresponding expected values given the species
+tree and the three population parameters.  Because recombinant edges from any
+species tree branch to the root branch are not possible under
+\texttt{ClonalOrigin} model (see Figure S\ref{fig:clonalorigin}), the column for
+the root as recipient was colored green, indicating invalid recombinant edges.
+Similarly, all of the cells where a recipient branch was older than the donor,
+and two branches did not share any contemporary time were colored green. The
+diagonal cells were colored gray, indicating no significant number of
+recombinant edges, because the observed and expected numbers of recombinant
+edges were similar. As recombinant edges starting from and ending in the same
+species tree branch would not change the species tree, which would not change
+the likelihood much, the cells would be colored gray.  Because SDD is a strictly
+veterinary pathogenic bacteria, we did not expect that SDD received much gene
+transfer from either SPY1 or SPY2: the log base 2 ratio from SPY1 to SDD was
+-2.20, and that from SPY2 to SDD was -1.97.  On the contrary, the recombing gene
+transfer from SDE1 to SDD was 0.756, and that from SDE2 to SDD was 0.698. This
+relatively higher recombining gene transfer from two SDE individuals to SDD
+compared with that from two SPY individuals to SDD reflected that the SDD was
+more closely related to SDE than to SPY in most of the genomes.
+
+The genomic organization between SDD and SDE1/SDE2 were much similar (Figure
 S\ref{fig:mauve}), and they formed a clade in the species tree
@@
 
-The recombining gene transfer log base 2 ratios between SDE1 and SDE2 (1.17 and
-1.20) were about twice as large as that from SDE1 and SDE2 to SPY1 and SPY2
-(0.67, 0.66, 0.60, and 0.39).  The intensities of recombining gene transfer from
-external branches of SDE to external branches to SPY was similar to those from
-SDE1 or SDE2 to SDD (0.76 and 0.70).  Either SPY1 or SPY2, and SDD appeared to
-be no different from the view point of SDE1 or SDE2 because of the similar
-values of recombining gene transfer intensity from SDE1 or SDE2 to the other
-three external branches.  It was expected that the recombining gene transfer
-within a clade would be larger than that between clades.  Surprisingly, the
-recombining gene transfer log base 2 ratios between SPY1 and SPY2 (1.06 and
+Because it was expected that the recombining gene transfer within a clade would
+be larger than that between clades, the recombining gene transfer log base 2
+ratios between SDE1 and SDE2 (1.17 and 1.20) were about twice as large as that
+from SDE1 and SDE2 to SPY1 and SPY2 (0.67, 0.66, 0.60, and 0.39).  The
+intensities of recombining gene transfer from external branches of SDE to
+external branches to SPY was similar to those from SDE1 or SDE2 to SDD (0.76 and
+0.70).  Either SPY1 or SPY2, and SDD appeared to be no different from the view
+point of SDE1 or SDE2 since the similar values of recombining gene transfer
+intensity from SDE1 or SDE2 to the other three external branches.  Surprisingly,
+the recombining gene transfer log base 2 ratios between SPY1 and SPY2 (1.06 and
 1.07) was comparable with those from SPY1 and SPY2 to SDE1 and SDE2 (1.25, 1.25,
 1.17, and 1.29).  Effectively, the recombining gene transfer from a clade SPY1
-or SPY2 to a different clade SDE1 or SDE2 was no different from that within the
-clade of SPY1 and SPY2. The two external branches of SDE received genes from
-either SPY1 or SPY2 via recombining gene transfer as if SDE1 and SDE2 were in
-the same clade as SPY1 and SPY2.  The bacterial recombination between SPY and
-SDE might happen more often than that between SDE and SDD whereas the genome
-configurations of SDE and SDD are more similar than each to SPY. This might be
-consistent with the fact that SPY and SDE shared the same ecological niche.  The
-gene transfer log base 2 ratio from SPY1 and SPY2 to SDE branch was larger than
-any type of gene transfer: 1.88 and 1.94, respectively.  Furthermore, we found a
-total of 58 genes for which a recombinant edge of certain types were sampled
-with probability at least 0.9 by focusing on specific types of recombinant
-edges.  (Table S\ref{tab:genes-transfer}).  Many of the genes in the list were
-transferred from SPY to SDE rather than the reverse direction overall.  All
-things considered, this clearly showed that the recombining gene transfer was
-biased toward the net direction from SPY to SDE.
-
-Figure \ref{fig:hgt-heatmap} shows the number of inferred non-recombining gene
-transfer events between the different branches of the species tree.  SDE1 and
-SDE2 shared the most events (26 and 34) compared to other sample pairs.  SDD
-appeared to acquire many genes from the ancestral SDE branch. Transfer events
-between the SPY and SDE clades were inferred in smaller numbers (<6) compared to
-those within each clade (23--34). Among the clade-crossing transfers, we did not
-see any significant enrichment for a specific direction, as opposed to the
-result for recombining transfers from the model-based analysis (Figure
-S\ref{fig:mowgli-recomb-heatmap}).
+or SPY2 to a different clade SDE1 or SDE2 seemed to be no different from that
+within the clade of SPY1 and SPY2. The two external branches of SDE received
+genes from either SPY1 or SPY2 via recombining gene transfer as if SDE1 and SDE2
+were in the same clade as SPY1 and SPY2.  The bacterial recombination between
+SPY and SDE might happen more often than that between SDE and SDD even though
+the genome configurations of SDE and SDD are more similar than each to SPY. This
+might be consistent with the fact that SPY and SDE shared the same ecological
+niche.  The gene transfer log base 2 ratio from SPY1 and SPY2 to SDE branch was
+larger than any type of gene transfer: 1.88 and 1.94, respectively.  
+Furthermore, we found a total of 58 genes for which a recombinant edge of
+certain types were sampled with probability at least 0.9 by focusing on specific
+types of recombinant edges.  (Table S\ref{tab:genes-transfer}).  Many of the
+genes in the list were transferred from SPY to SDE rather than the reverse
+direction overall.  All things considered, this clearly showed that the
+recombining gene transfer was biased toward the net direction from SPY to SDE.
+
+Figure \ref{fig:hgt-heatmap} shows the number of inferred non-recombining
+horizontal gene transfer events between the different branches of the species
+tree.  SDE1 and SDE2 shared the most events (26 and 34) compared to other sample
+pairs.  SDD appeared to acquire many genes from the ancestral SDE branch.
+Transfer events between the SPY and SDE clades were inferred in smaller numbers
+(<6) compared to those within each clade (23--34). Among the clade-crossing
+transfers, we did not see any significant enrichment for a specific direction,
+as opposed to the result for recombining transfers from the model-based analysis
+(Figure S\ref{fig:mowgli-recomb-heatmap}).
 
 \subsection{More frequent recombining gene transfer than horizontal gene transfer}
-Among the 2314 gene families
-the largest family consisted of 59 genes, while
-nearly half of gene families (1066) had exactly one gene per species
-, or six genes total (Figure S\ref{fig:famsize}).
-Figure \ref{fig:Gene-duplication-loss} shows the
-inferred gene counts for each extant and ancestral species. Overall,
+
+Among the 2314 gene families the largest family consisted of 59 genes, while
+nearly half of gene families (1066) had exactly one gene per species, or six
+genes total (Figure S\ref{fig:famsize}).  Figure \ref{fig:Gene-duplication-loss}
+shows the inferred gene counts for each extant and ancestral species. Overall,
 we inferred 304 duplications, 944 recombining gene transfers, and 294
-non-recombining gene transfers, comprising 23.7\% of total transfer
-events.  We found that duplications made up only a small fraction of
-the events within the clade (10.2\%), while horizontal transfers and
-losses were significantly more frequent (41.7\% and 48.0\%,
-respectively). We found that duplications were most often
-species-specific. Many gene losses likely represented horizontal
-transfers that replaced the native gene sequence.  The species SDD and
-SDE were the most expanded, both in terms of duplications and in-bound
-transfers.
+non-recombining gene transfers, comprising 23.7\% of total transfer events.  We
+found that duplications made up only a small fraction of the events within the
+clade (10.2\%), while horizontal transfers and losses were significantly more
+frequent (41.7\% and 48.0\%, respectively). We found that duplications were most
+often species-specific. Many gene losses likely represented horizontal transfers
+that replaced the native gene sequence.  The species SDD and SDE were the most
+expanded, both in terms of duplications and in-bound transfers.
 
@@
 \subsection{Horizontal gene transfer is associated with virulence genes}
+
 \texttt{Select genes with high horizontal gene transfers. Relate these to
 categories of virulence genes. I should find many more virulence genes in Matt's
-analysis.}
+analysis. Use the virulence gene types from SDE2's paper. Which virulence gene
+types have more gene transfer?}
+
+We found that some virulence genes tended to have more horizontally transfers
+than others. 
+Streptococcal C5a peptidase (\textit{scpB}) (SDEG\_0933) was inferred to have six
+horizontal gene transfer. 
+Putative transposase were also found to have relatively large number of
+horizontal gene transfer events.
+(SpyM3\_1016, SpyM3\_1053, SpyM3\_1054)
+
+Putative response regulator (SpyM3\_0174).
+Putative maltose operon transcriptional repressor (SpyM3\_0982).
+Putative histidine kinase (SpyM3\_0171).
+Putative cell wall hydrolase (SpyM3\_0731, SpyM3\_0922, SpyM3\_1208).
+
+GGS\_124 possesses 5 genes that encode putative nucleases with a secretion
+signal peptide: genome-encoded streptodornase (SDEG\_0541), others,
+prophage-associated deoxyribonuclease (SDEG\_1103), all of which are predicted to code for a secretion signal peptide.
+Streptodornase (Sdn) - phage associated  (SpyM3\_1409).
+
+Protein F2 like fibronectin-binding protein (SpyM3\_0104).
 
 \subsection{Comparison between model-based and parsimony-based approaches}
+
 \texttt{I might want to analyze only core genomes and pan genomes
@@
 we could have more complete comparison.}
+
+\texttt{Genes and recombining gene transfer events count and recombination
+intensity might be use to compare two methods.}
 
@@
 alignment blocks, recombination rate per site for each block, local gene tree
-topologies, recombination
-intensity, locations of virulence genes, posterior probability of recombining
-gene transfer.
+topologies, recombination intensity, locations of virulence genes, posterior
+probability of recombining gene transfer.
 
@@
 \texttt{Comparison of Mawgli and CO?}
-Recombining gene transfers were not enriched for virulence genes (P-value of $>0.90$)
-while horizontal gene transfers were found to be significantly associated with
-virulence genes (P-value of $<1.33 \times 10^{-11}$). This was also consistent
-with the fact that we could not find the association of recombining gene
-transfer with virulence genes in the model-based approach.
-This supported more evidence
-that non-recombining gene transfer served as a more effective means by
-which SDE or SPY acquired virulence genes, compared to recombining
-gene transfer.
-For horizontal gene transfer we found enrichment in very different 
-categories related to
-transposition, integration, and recombination.  
+Recombining gene transfers were not enriched for virulence genes (P-value of
+$>0.90$) while horizontal gene transfers were found to be significantly
+associated with virulence genes (P-value of $<1.33 \times 10^{-11}$). This was
+also consistent with the fact that we could not find the association of
+recombining gene transfer with virulence genes in the model-based approach.
+This supported more evidence that non-recombining gene transfer served as a more
+effective means by which SDE or SPY acquired virulence genes, compared to
+recombining gene transfer.  For horizontal gene transfer we found enrichment in
+very different categories related to transposition, integration, and
+recombination.  
 \texttt{More interpretations about Mawgli analsis of function categories?}
@@
 \subsection{Units of recombining gene transfer}
-The process of gene 
-transfer does not necessarily move whole genes from one species to another 
-\citep[e.g.,][]{Chan2009}, and we observed that units of 
-recombining gene transfer do 
-not correspond to single genes in their entirety
-(Table S\ref{tab:genes-transfer}). In some cases, only parts of 
-a gene were inferred to have been transferred, for instance, transfer from SPY1 
-to SDE1 was inferred for 29\% of the aspartyl-tRNA synthetase gene (see also 
-Figure S\ref{fig:muts}). In other cases, the inferred transfer units cover sets 
-of as much as eight adjacent genes.
-Not only parts of genes could be transferred via recombination, but a group of 
-genes could also be moved together. Of the genes listed in Table 
-S\ref{tab:genes-transfer} there were 12 groups of genes that were consecutive. 
-While transfer fragments of genes often covered two genes, there were transfer 
-fragments that covered eight genes together. Loci from SpyM3\_1512 to 
-SpyM3\_1525 were inferred to be transferred from SDE2 to SDD.
+The process of gene transfer does not necessarily move whole genes from one
+species to another \citep[e.g.,][]{Chan2009}, and we observed that units of
+recombining gene transfer do not correspond to single genes in their entirety
+(Table S\ref{tab:genes-transfer}). In some cases, only parts of a gene were
+inferred to have been transferred, for instance, transfer from SPY1 to SDE1 was
+inferred for 29\% of the aspartyl-tRNA synthetase gene (see also Figure
+S\ref{fig:muts}). In other cases, the inferred transfer units cover sets of as
+much as eight adjacent genes.  Not only parts of genes could be transferred via
+recombination, but a group of genes could also be moved together. Of the genes
+listed in Table S\ref{tab:genes-transfer} there were 12 groups of genes that
+were consecutive.  While transfer fragments of genes often covered two genes,
+there were transfer fragments that covered eight genes together. Loci from
+SpyM3\_1512 to SpyM3\_1525 were inferred to be transferred from SDE2 to SDD.
 
 \section{Discussions}
-Closely related bacterial species were investigated here to better
-understand gene transfer processes. Two complementary methods were employed to
-study recombining and non-recombining gene transfers. By discriminating the two
+
+Closely related bacterial species were investigated here to better understand
+gene transfer processes. Two complementary methods were employed to study
+recombining and non-recombining gene transfers. By discriminating the two
 processes in analyzing the genomes, we provided evidence that the pathogenicity
@@
 \textbf{Virulence genes and gene transfer:} 
-Because we found directional recombining gene transfer from SPY to SDE, and
-did not found virulence genes association with recombining gene transfer, 
-this indicated that 
-pathogenicity of SDE might not be due to acquisition of virulence genes from
-SPY via recombining transfer. This does not necessarily mean that virulence
-genes were not exchanged between SPY and SDE via non-recombining transfer. 
-Because the Mowgli analysis just did not reveal unbalanced non-recombining or
-recombinding gene transfer, we suspected that
-Mowgli's parsimony might not be powerful in detecting unbalanced gene flow.
-We conjecture the following hypothesis of the evolution of SPY and SDE to
-support the non-recombining virulence gene exchange between SPY and SDE.
-Pathogenicity of SDE might have existed before the common ancestor of SPY
-and SDE.  We speculated that SPY might have been split off from the common
-ancestor that was SDE-like because the population size of SDE was larger than
-that of SPY (CITE). After the split, SPY might have been specialized in
-inhabting human hosts, and SDE in colonizing veterinary hosts. Recent surge in
-cases of human diseases attributed to SDE might be because of secondary
-contact of SPY and SDE since their split from their common ancestor.  When SPY
-thrives in its environment of human hosts and another relatively close species,
-SDE, starts to cohabit with the indigenous species SPY, genetic elements in one
-of the two species could benefit another species. If the two species do not
-compete with each other, then the increased genetic variation from multiple
-species would eventually bring advantage to both of the species. 
-Virulence genes could have been beneficial to both species in inhabting human hosts. 
-Because the common ancestry of SPY and SDE, recombining gene transfer between
-the two species might be relatively prevalent. But, these homologous
-recombination might have rarely contributed to influx of virulence genes from
-SPY to the opportunistic human pathogen of SDE. Because homologous parts
-of SPY and SDE would be similar in their genetic compositions, the homologous
-parts would be unlikely to contain virulence genes to human hosts. We suspected
-that virulence genes were exchanged between SPY and SDE, but did not find any
-directionality of the gene flow.
+Because we found directional recombining gene transfer from SPY to SDE, and did
+not found virulence genes association with recombining gene transfer, this
+indicated that pathogenicity of SDE might not be due to acquisition of virulence
+genes from SPY via recombining transfer. This does not necessarily mean that
+virulence genes were not exchanged between SPY and SDE via non-recombining
+transfer.  Because the \texttt{Mowgli} analysis just did not reveal unbalanced
+non-recombining or recombinding gene transfer, we suspected that \texttt{Mowgli}'s
+parsimony might not be powerful in detecting unbalanced gene flow.  We
+conjecture the following hypothesis of the evolution of SPY and SDE to support
+the non-recombining virulence gene exchange between SPY and SDE.  Pathogenicity
+of SDE might have existed before the common ancestor of SPY and SDE.  We
+speculated that SPY might have been split off from the common ancestor that was
+SDE-like because the population size of SDE was larger than that of SPY (CITE).
+After the split, SPY might have been specialized in inhabting human hosts, and
+SDE in colonizing veterinary hosts. Recent surge in cases of human diseases
+attributed to SDE might be because of secondary contact of SPY and SDE since
+their split from their common ancestor.  When SPY thrives in its environment of
+human hosts and another relatively close species, SDE, starts to cohabit with
+the indigenous species SPY, genetic elements in one of the two species could
+benefit another species. If the two species do not compete with each other, then
+the increased genetic variation from multiple species would eventually bring
+advantage to both of the species.  Virulence genes could have been beneficial to
+both species in inhabting human hosts.  Because the common ancestry of SPY and
+SDE, recombining gene transfer between the two species might be relatively
+prevalent. But, these homologous recombination might have rarely contributed to
+influx of virulence genes from SPY to the opportunistic human pathogen of SDE.
+Because homologous parts of SPY and SDE would be similar in their genetic
+compositions, the homologous parts would be unlikely to contain virulence genes
+to human hosts. We suspected that virulence genes were exchanged between SPY and
+SDE, but did not find any directionality of the gene flow.
 
@@
 that using relatively small number of genomes we could detect homologous
-recombining gene transfer that had been possible using many samples for
-a few genes. We revisited the problem of gene transfer in other previous
-studies.  We examined
-the seven genes that \citet{Kalia2001} had studied for directionality
+recombining gene transfer that had been possible using many samples for a few
+genes. We revisited the problem of gene transfer in other previous studies.  We
+examined the seven genes that \citet{Kalia2001} had studied for directionality
 of gene transfer between SPY and SDE. Internal fragments of the seven
-housekeeping genes, encoding putative glucose kinase (\textit{gki}),
-glutamine transport protein (\textit{gtr}), glutamate racemase (\textit{murI}),
-mismatch repair enzyme (\textit{mutS}), transketolase (\textit{recP}),
-xanthine phosphoribosyltransferase (\textit{xpt}), and acetylcoenzyme
-A acetyltransferase (\textit{yqiL}) were located in the SPY1 genome that we
-investigated in this study.  Six of the seven genes except \textit{yqiL
-}gene were in the core genomes \texttt{ClonalOrigin} analyzed. Genes \textit{mutS},
-\textit{recP}, and \textit{xpt} had fragments of gene transfer with moderate
-posterior probability of recombination. The other genes did not have
-noticeable recombination. Except for \textit{mutS} where \citet{Kalia2001}
-did not find recombination, the result was consistent with that of
-\citet{Kalia2001}. 
-We also revisited recombination in gene \textit{parC} in SPY1 genome
-studied by \citet{Pinho2010}, where they studied fluoroquinolone
-resistance in SDE and reported recombination in the gene. We also
-observed moderate recombination in the gene.
+housekeeping genes, encoding putative glucose kinase (\textit{gki}), glutamine
+transport protein (\textit{gtr}), glutamate racemase (\textit{murI}), mismatch
+repair enzyme (\textit{mutS}), transketolase (\textit{recP}), xanthine
+phosphoribosyltransferase (\textit{xpt}), and acetylcoenzyme A acetyltransferase
+(\textit{yqiL}) were located in the SPY1 genome that we investigated in this
+study.  Six of the seven genes except \textit{yqiL }gene were in the core
+genomes \texttt{ClonalOrigin} analyzed. Genes \textit{mutS}, \textit{recP}, and
+\textit{xpt} had fragments of gene transfer with moderate posterior probability
+of recombination. The other genes did not have noticeable recombination. Except
+for \textit{mutS} where \citet{Kalia2001} did not find recombination, the result
+was consistent with that of \citet{Kalia2001}.  We also revisited recombination
+in gene \textit{parC} in SPY1 genome studied by \citet{Pinho2010}, where they
+studied fluoroquinolone resistance in SDE and reported recombination in the
+gene. We also observed moderate recombination in the gene.
 
 \textbf{Mechanism of unbalanced gene flow between SPY and SDE:} 
-If there were unbalanced gene transfer between the two
-species, what were the barriers \citep{Thomas2005} to gene transfer
-in the bacteria? Although barriers to bacterial recombination would
-include geographic isolation, genetic isolation could prevent recombination
-between coinhabiting bacteria. \citet{Kalia2001} discussed possible
-mechanisms of the unbalanced gene flow by mentioning potential roles
-of restriction-modification system (RMS) in the unbalanced gene flow.
-Restriction-modification system is a biological self-defense mechanism
-of bacteria, which is diverse in prokaryote world. Biological roles
-of RMS might be to maintain and control species identity \citep{Jeltsch2003}.
-Recently, \citet{Budroni2011a} studied that \textit{Neisseria meningitidis}
-phylogenetic clades were associated with RMS, claiming that RMS modulated
-homologous recombination in the bacteria of \textit{N.\ meningitidis}.
-Most of \textit{S.\ pyogenes} genomes contain restriction-modification
-type I system as well as type II system. The two genomes of SDE contained
-only type II system. We conjecture that more tight self-defense system
-of SPY may have made the species less capable of aquiring foreign
-genetic material from SDE, and accordingly relatively unbalanced gene
-flow between SPY and SDE may have been associated with different 
-restriction-modification systems in the two species.
+If there were unbalanced gene transfer between the two species, what were the
+barriers \citep{Thomas2005} to gene transfer in the bacteria? Although barriers
+to bacterial recombination would include geographic isolation, genetic isolation
+could prevent recombination between coinhabiting bacteria. \citet{Kalia2001}
+discussed possible mechanisms of the unbalanced gene flow by mentioning
+potential roles of restriction-modification system (RMS) in the unbalanced gene
+flow.  Restriction-modification system is a biological self-defense mechanism of
+bacteria, which is diverse in prokaryote world. Biological roles of RMS might be
+to maintain and control species identity \citep{Jeltsch2003}.  Recently,
+\citet{Budroni2011a} studied that \textit{Neisseria meningitidis} phylogenetic
+clades were associated with RMS, claiming that RMS modulated homologous
+recombination in the bacteria of \textit{N.\ meningitidis}.  Most of \textit{S.\
+pyogenes} genomes contain restriction-modification type I system as well as type
+II system. The two genomes of SDE contained only type II system. We conjecture
+that more tight self-defense system of SPY may have made the species less
+capable of aquiring foreign genetic material from SDE, and accordingly
+relatively unbalanced gene flow between SPY and SDE may have been associated
+with different restriction-modification systems in the two species.
 
@@
 Among the genes shown in Table S\ref{tab:functional} two classes of proteins
-attracted our attentions: 
-aminoacyl tRNA synthetases (aaRS) and phosphotransferase system (PTS).
-Aminoacyl tRNA synthetases
-charge tRNA with an amino acid so that amino acids are concatenated
-from charged tRNA onto a growing peptides through translation. The
-aaRS's are characterized by an amino acid that they charge tRNA with.
-This gene family had been a target for horizontal gene transfer 
-\citep[e.g.,][]{Lamour1994,Woese2000,Koonin2001},
-which lent itself to an example of gene transfer in bacteria.
-There were also recent reports of gene transfer in tRNA aminoacylation
-\citep{Andam2010,Wang2011}. 
-In our analysis, some of them appeared to be transferred
-from SPY to SDE:
-leucyl-tRNA synthetase, aspartyl-tRNA synthetase, and arginyl-tRNA
-synthetase. Half of phenylalanyl-tRNA synthetase gene appeared to
-be transferred in the reverse direction.
-
-Bacterial phosphotransferase
-system (PTS), also known as phosphoenolpyruvate (PEP) group translocation,
-transports sugars such as glucose into the cell where phosphorylation
-of the sugar happens. There have been studies of gene transfer of
-PTS in bacteria \citep[e.g.,][]{Barabote2005,Zuniga2005}. As the
-main source of energy for bacteria, sugars in the bacterial niches
-must have been acting as important selective pressure. It is not surprising
-that PTS sugars differ among bacterial groups because bacterial species
-are correlated with their environment and energy sources theirof.
-Table S\ref{tab:functional} shows that 
-two PTS genes appeared to be transferred from SPY to SDE:
-{}``PTS system, mannose-specific IIC component'' and {}``PTS system,
-fructose-specific IIABC component.'' Whereas almost half of the former
-gene was inferred to be affected by recombinant edges, only 3\% of the latter gene appeared
-to be affected.
+attracted our attentions: aminoacyl tRNA synthetases (aaRS) and
+phosphotransferase system (PTS).  Aminoacyl tRNA synthetases charge tRNA with an
+amino acid so that amino acids are concatenated from charged tRNA onto a growing
+peptides through translation. The aaRS's are characterized by an amino acid that
+they charge tRNA with.  This gene family had been a target for horizontal gene
+transfer \citep[e.g.,][]{Lamour1994,Woese2000,Koonin2001}, which lent itself to
+an example of gene transfer in bacteria.  There were also recent reports of gene
+transfer in tRNA aminoacylation \citep{Andam2010,Wang2011}.  In our analysis,
+some of them appeared to be transferred from SPY to SDE: leucyl-tRNA synthetase,
+aspartyl-tRNA synthetase, and arginyl-tRNA synthetase. Half of phenylalanyl-tRNA
+synthetase gene appeared to be transferred in the reverse direction.
+
+Bacterial phosphotransferase system (PTS), also known as phosphoenolpyruvate
+(PEP) group translocation, transports sugars such as glucose into the cell where
+phosphorylation of the sugar happens. There have been studies of gene transfer
+of PTS in bacteria \citep[e.g.,][]{Barabote2005,Zuniga2005}. As the main source
+of energy for bacteria, sugars in the bacterial niches must have been acting as
+important selective pressure. It is not surprising that PTS sugars differ among
+bacterial groups because bacterial species are correlated with their environment
+and energy sources theirof.  Table S\ref{tab:functional} shows that two PTS
+genes appeared to be transferred from SPY to SDE: {}``PTS system,
+mannose-specific IIC component'' and {}``PTS system, fructose-specific IIABC
+component.'' Whereas almost half of the former gene was inferred to be affected
+by recombinant edges, only 3\% of the latter gene appeared to be affected.
 
@@
 non-recombining gene transfer:}
-The method of detecting homologous recombination gene transfer, or
-the \texttt{ClonalOrigin}, is model-based approach as opposed to the other
-method for non-recombining gene transfer, or the Mowgli.
-The core genomes consisted of relatively long (i.e., 1500 base pairs
-long) sequence alignments with almost no gaps. With this inherent
-filtering of the input data preparation \texttt{ClonalOrigin} did not account
-for gene gain and loss in its core genome analysis. On the contrary,
-Mowgli used the pan genomes as well as the core genomes.  Considering
-nucleotides as the level of gene duplication, loss, and transfer towards
-the efforts of having more complete models for recombining and non-recombining
-gene transfer
-could be difficult because genome-scale application of those models
-based on DNA or protein sequences might not be feasible.
-The research into units
-of gene transfer is underway \citep[e.g.,][]{Chan2009a}. Discriminating
-the mode of gene transfer and using the units of each mode of gene
-transfer towards the efforts of developing model-based approaches
-to detecting gene transfer might be plausible. In doing so, more complicated
-evolutionary models could be considered: genome rearrangement,
-gene segment gain and loss, and gene duplications.
+The method of detecting homologous recombination gene transfer, or the
+\texttt{ClonalOrigin}, is model-based approach as opposed to the other method
+for non-recombining gene transfer, or the \texttt{Mowgli}.  The core genomes consisted of
+relatively long (i.e., 1500 base pairs long) sequence alignments with almost no
+gaps. With this inherent filtering of the input data preparation
+\texttt{ClonalOrigin} did not account for gene gain and loss in its core genome
+analysis. On the contrary, \texttt{Mowgli} used the pan genomes as well as the core
+genomes.  Considering nucleotides as the level of gene duplication, loss, and
+transfer towards the efforts of having more complete models for recombining and
+non-recombining gene transfer could be difficult because genome-scale
+application of those models based on DNA or protein sequences might not be
+feasible.  The research into units of gene transfer is underway
+\citep[e.g.,][]{Chan2009a}. Discriminating the mode of gene transfer and using
+the units of each mode of gene transfer towards the efforts of developing
+model-based approaches to detecting gene transfer might be plausible. In doing
+so, more complicated evolutionary models could be considered: genome
+rearrangement, gene segment gain and loss, and gene duplications.
 
@@
 \subsection{Complete bacterial genomes}
+
 Two genomes of \textit{S.\ pyogenes} (SPY), and two of \textit{S.\ dysgalactiae}
 ssp.\textit{\ equisimilis} (SDE) were downloaded from NCBI with the following
-accessions: NC\_004070 {Beres2002}, NC\_008024 \citep{Beres2006}, NC\_012891
-\citep{Shimomura2011}, and CP002215 \citep{Suzuki2011}. We refer to these
-samples as SPY1, SPY2, SDE1, and SDE2, respectively. One genome of 
+accessions: NC\_004070 \citep{Beres2002}, NC\_008024 \citep{Beres2006},
+NC\_012891 \citep{Shimomura2011}, and CP002215 \citep{Suzuki2011}. We refer to
+these samples as SPY1, SPY2, SDE1, and SDE2, respectively. The genome of
 \textit{S.\ dysgalactiae} ssp.\textit{\ dysgalactiae} was provided by
-\citet{Suzuki2011}
-with accession identifiers of SddyATCC27957, referred to here as SDD. We also
-downloaded the genome (NC\_012471) of \textit{S.\ equi} ssp.\textit{\ equi}
-strain 4047 referred to as SEE \citep{Holden2009}.
+\citet{Suzuki2011} with accession identifiers of SddyATCC27957, referred to here
+as SDD. We also downloaded the genome (NC\_012471) of \textit{S.\ equi}
+ssp.\textit{\ equi} strain 4047 referred to as SEE \citep{Holden2009}.
 
 \subsection{Model-based method for recombining gene transfer}
+
 We present a short summary of the statistical method developed by
@@
 recombination. A detailed account was found in the two papers and in the user
-guide available at \url{http://code.google.com/p/clonalorigin}.
-
-We used \texttt{progressiveMauve} v2.3.1 \citep{Darling2004,Darling2010} with
-default options to attempt to align the six genomes, which led to too small
-proportion of a core genome. We had to remove the SEE sample from the genome
-alignment. Note that we kept it in the parsimony-based approach.  We then used
+guide available at \url{http://code.google.com/p/clonalorigin}.  We used
+\texttt{progressiveMauve} v2.3.1 \citep{Darling2004,Darling2010} with default
+options to attempt to align the six genomes, which led to too small proportion
+of a core genome. We had to remove the SEE sample from the genome alignment
+although we kept it in the parsimony-based approach.  We then used
 \texttt{stripSubsetLCBs}, a program of the \texttt{ClonalOrigin} package, and
 identified 276 blocks of sequence alignments with length threshold of 1500 base
-pairs.  We excluded two of these blocks because they contained long consecutive
-gaps.
-
-Using 274 blocks, we applied \texttt{ClonalFrame} v1.1 available at
-\url{http://www.xavierdidelot.xtreemhost.com}, and obtained an estimate of the
-species tree relating the five genomes. ClonalFrame also outputs estimates of
-parameters including the average recombinant tract length and the ratio between
-recombination and mutation rates, however, these were not used in the subsequent
-analysis. We used a single Markov chain Monte Carlo (MCMC) with 10000 burn-in
-generations and 10000 sampling generations to obtain the tree estimate with
-branch lengths, and additionally seven independent MCMC runs to validate
-adequate convergence of the first.
+pairs.  We excluded two of these blocks in a subsequent analysis because they
+contained long consecutive gaps.  We applied \texttt{ClonalFrame} v1.1 available
+at \url{http://www.xavierdidelot.xtreemhost.com} to the remaining 274 blocks,
+and estimated the species tree relating the five genomes. \texttt{ClonalFrame}
+estimates the average recombinant tract length and the ratio between
+recombination and mutation rates, which were not used in the following analysis.
+We used a single Markov chain Monte Carlo (MCMC) with 10000 burn-in generations
+and 10000 sampling generations to obtain the tree estimate with branch lengths,
+and additionally seven independent MCMC runs to validate adequate convergence of
+the first run.
 
@@
 recombination rate, and recombinant tract length.  Two independent chains were
-executed with $10^6$ burn-in generations and $10^7$ sampling generations,
-sub-sampled every $10^5$-th generations. One of the two chains was used to
-obtain the block-wise mean estimates, and the other chain was used to validate
-convergence of the first stage of MCMC.  The estimators of the three population
-parameters were medians weighted over block lengths, which down-weighed less
-reliable estimates obtained from shorter alignment blocks.
-
-In the first stage of \texttt{ClonalOrigin} MCMC a few alignment blocks were not
-successfully finished even after a month of computation.  Because those were not
-expected to affect the global estimates of population parameters significantly,
-we used only finished results of the first stage of \texttt{ClonalOrigin} to
-estimate mutation rate, recombination rate, and recombinant tract length.  Two
-independent runs of the first stage MCMC resulted in similar estimates of the
-three population parameters (Table S\ref{tab:three}).
-
-The second stage of
-MCMC with \texttt{ClonalOrigin} was executed with the three population
-parameters fixed to the estimated values in the first stage.  Two independent
-chains of the second stage of \texttt{ClonalOrigin}'s MCMC were also executed
-with a total generation being $1.1\times10^8$. The first $10^7$ generations of
-each chain were discarded as burn-in. The remaining $10^8$ generations were
-sub-sampled every $10^5$-th generation.
+executed with $10^6$ burn-in generations and $10^7$ sampling generations. We
+sub-sampled the result every $10^5$-th generations.  A few alignment blocks were
+not successfully finished even after a month of computation.  Because those were
+not expected to affect the global estimates of population parameters
+significantly, we used only finished blocks to estimate mutation rate,
+recombination rate, and recombinant tract length.  The global estimators of the
+three population parameters were medians weighted over block lengths, which
+down-weighed less reliable estimates obtained from shorter alignment blocks.
+One of the two independent chains was used to obtain the block-wise mean
+estimates, and the other chain was used to validate convergence of the first
+stage of MCMC.  Two independent MCMC runs resulted in similar estimates of the
+three population parameters (Table S\ref{tab:three}).  The second stage of MCMC
+with \texttt{ClonalOrigin} was executed with the three population parameters
+fixed to the estimated values in the first stage.  Two independent MCMC chains
+were also executed with a total generation being $1.1\times10^8$. The first
+$10^7$ generations of each chain were discarded as burn-in. The remaining $10^8$
+generations were sub-sampled every $10^5$-th generation.
 
@@
 tree} was not detailed in the user guide at the time when we used
-\texttt{ClonalOrigin}, we elected to use Figure \ref{fig:clonalorigin} to
-illustrate a recombinant tree in order to clarify summarization procedures we
-employed of posterior samples of recombinant trees.  A recombination history is
+\texttt{ClonalOrigin}, we elected to use Figure S\ref{fig:clonalorigin} to
+illustrate a recombinant tree in order to clarify summarization of
+\textit{recombinant history} posterior samples.  A recombination history is
 represented by a \textit{recombinant tree}, which is a species tree augmented
-with a set of \textit{recombinant edges}.  Each recombinant edge is defined by
-time and branch of origin, time and branch of destination, and genomic interval
-that it affects. The branch of origin is referred to as the \textit{donor} and
-the destination branch is referred to as the \textit{recipient}. A recombinant
+with \textit{recombinant edges}.  Each recombinant edge is defined by the time
+and branch of origin and destination, and the genomic interval that it affects.
+The branch of origin is referred to as the \textit{donor} and the destination
+branch is referred to as the \textit{recipient}. The type of a recombinant edge
+is determined by the donor and recipient species tree branches.  A recombinant
 tree can be viewed as a restricted version of an ancestral recombination graph
@@
 block it describes.  \citet{Didelot2010} presented two procedures of summarizing
-recombinant trees by viewing trees from two different angles: one from
-recombinant trees by aggregating all of the sites, and another from site-wise
-recombinant edge changes to the species tree.  By following their approaches we
-quantified more aspects of recombinant trees: from the view of recombinant trees
-with aggregated sites we counted gene tree topologies, and from the view of
-site-wise recombinant edge changes we quantified recombination intensity along
-the genome.
-
-We first describe procedures of summarizing recombinant trees from the first
-viewpoint.  To summarize the posterior samples of \texttt{ClonalOrigin} we
-examined the distribution of topologies of local genealogies induced by the
-sampled recombinant trees. Recall that a recombinant tree induces a local
-genealogy for each site along the analyzed block.  Figure \ref{fig:clonalorigin}
-shows six local gene trees including one that is the same as the species tree
-from the most left-hand side of the block of 100 base pairs to the right.  We
-expected the dominant topology across sampled genealogies to be concordant with
-that of the species tree. However, other frequently sampled topologies were
-expected to be potentially informative about common recombination events. There
-are 105 possible rooted topologies relating five taxa, which allowed us to count
-topologies induced by a posterior sample of recombinant trees.
-
-For the second viewpoint of recombinant trees \citet{Didelot2010} suggested an
-approach used to trace recombination intensity along the genome.  For each edge
-type and each site, we recorded the sampling frequency of a recombinant edge of
-the type at that site (e.g., Figure S\ref{fig:muts}).  Figure
-\ref{fig:clonalorigin}B illustrates the posterior probability that a site
-experiences influx from a donor branch using a single recombinant tree.  Summing
-these posteriors over a certain set of edge types, we could obtain a measure of
-recombination intensity along the genome (Figure \ref{fig:clonalorigin}B).  This
-allowed us to find genomic regions enriched for recombination events in general,
-or regions enriched for recombination events of certain types (e.g., between SPY
-and SDE).
-
-We note that the recombination intensity is not normalized to
-the interval $[0,1]$. Rather giving probabilistic meanings to the recombination
-intensity we interpreted it as an average number of recombinant edge types per
-site.  The measure appeared to act as an effective indicator of recombination
-intensity (see subsection of \textbf{Simulation Study}).  Our simulation study
-indicated that edges of certain types were better indicators of true
-recombination events than others. For instance, recombinant edges which resulted
-in a recombinant tree with the same topology as the species tree appeared to be
-more prone to spurious sampling.  Such edges were ones where the donor was the
-same branch, sister branch, or parent branch of the recipient branch.
-Therefore, in some cases we elected to remove such edges from consideration when
-measuring recombination intensity.
+recombinant trees by viewing trees from two different angles: by counting types
+of recombinant edges regardless of genomic positions affected by the edges, and
+by measuring recombinant edge changes to the species tree in site-wise manner.
+By similarly following their approaches we quantified additional aspects of
+recombinant trees; we counted gene tree topologies for each site, and also
+quantified recombination intensity along the genome.
+
+The recombination intensity allowed us to find genomic regions enriched for
+recombination events in general, or regions enriched for recombination events of
+certain types (e.g., between SPY and SDE).  We note that the recombination
+intensity is not normalized to the interval $[0,1]$. Rather than giving
+probabilistic meanings to the recombination intensity we interpreted it as an
+average number of recombinant edge types per site.  The measure appeared to act
+as an effective indicator of recombination intensity (see subsection of
+\textbf{Simulation Study}).  Our simulation study indicated that edges of
+certain types were better indicators of true recombination events than others.
+For instance, recombinant edges which resulted in a recombinant tree with the
+same topology as the species tree appeared to be more prone to spurious
+sampling.  Such edges were ones where the donor was the same branch, sister
+branch, or parent branch of the recipient branch.  Therefore, in some cases we
+elected to remove such edges from consideration when measuring recombination
+intensity.  The tree topologies changes would be correlated with recombination
+intensity.  The tree topologies would show more effectively regions where
+recombination occurs along the genome (see subsection of \textbf{UCSC Genome
+Browser}).
 
 \subsection{Parsimony-based method for horizontal gene transfer}
+
 While \texttt{ClonalOrigin} could be used to understand recombining gene
@@
 \citep{David2011,Doyon2011,Tofigh2011}.  In this analysis, we complemented our
-study of recombination by using Mowgli \citep{Doyon2011} to characterize events
-that can change the copy number of a gene, such as duplications, losses, and
-non-recombining horizontal gene transfers.  In contrast to
-\texttt{ClonalOrigin}, Mowgli assumes a parsimony model for inferring events and
-it uses pre-determined gene clusters where genes are treated as atomic units.
-%The advantage of using Mowgli was that it did not assume the sequence
-%data contains a single copy of each homologous segment, and it
-%attempted to explain different copy numbers by a series of gain and
-%loss events.  We here were interested in the approach of Mowgli to
-%study non-recombining gene transfers.
-We obtained gene family clusters defined in \citet{Suzuki2011} using the
-OrthoMCL method \citep{Li2003} over the full set of six genomes including
-\textit{S.\ equi} subsp.\textit{\ equi} strain 4047 (SEE).  For each family
-cluster, the corresponding protein sequences were aligned using the MUSCLE
-program \citep{Edgar2004a} and nucleotide alignments were constructed by mapping
-nucleotide sequences onto each protein sequence alignment. We then identified
-recombination within gene alignments using the Single Breakpoint Recombination
-(SBP) method from the HyPhy package \citep{KosakovskyPond2006}.  Using the
-Akaike Information Criterion (AIC), SBP inferred that 47.5\% (1,094) of gene
-families contained at least one topology-changing recombination. Each of these
-families was then split into two gene fragments.  Gene trees were constructed
-from nucleotide alignments for each gene fragment using the RAxML program
-\citep{Stamatakis2006}, and were rooted by minimizing inferred duplications and
-losses.  Each rooted gene tree was then reconciled using the Mowgli program
-\citep{Doyon2011}, which inferred the most parsimonious gene duplication, loss,
-and transfer events for each family.  In the case of multiple maximum
-parsimonious reconciliations, one was chosen uniformly at random.  Unless
-specified otherwise, default setings were used for all programs.  Since the
-method of Mowgli did not directly model recombination events,
-recombination-based transfers were inferred in a post-processing step, by
-checking whether the transfer was coupled with a loss of a homologous gene in
-the recipient branch (indicating gene replacement rather than gain).  A transfer
-event was classified as non-recombining if the destination species of that
-transfer had an extant descendant that contained copies of that genes which were
-descendants of the transferred gene as well as copies which were
-non-descendants.  We noted that this parsimonious approach was imperfect and
-would tend to undercount events, especially recombining gene transfer events,
-which were modeled using two events: a transfer and a loss. Nevertheless, this
-analysis allowed us to study non-recombining gene transfers, which were not
-modeled by \texttt{ClonalOrigin}, and to validate some of the
+study of recombination by using \texttt{Mowgli} \citep{Doyon2011} to
+characterize events that can change the copy number of a gene, such as
+duplications, losses, and non-recombining horizontal gene transfers.  In
+contrast to \texttt{ClonalOrigin}, \texttt{Mowgli} assumes a parsimony model for
+inferring events and uses pre-determined gene clusters where genes are treated
+as atomic units.  The advantage of using \texttt{Mowgli} was that it did not
+assume the sequence data contains a single copy of each homologous segment, and
+it attempted to explain different copy numbers by a series of gain and loss
+events.  We here were interested in the approach of \texttt{Mowgli} to study
+non-recombining horizontal gene transfers.  We obtained gene family clusters
+defined in \citet{Suzuki2011} using the OrthoMCL method \citep{Li2003} over the
+full set of six genomes including \textit{S.\ equi} subsp.\textit{\ equi} strain
+4047 (SEE).  For each family cluster, the corresponding protein sequences were
+aligned using the MUSCLE program \citep{Edgar2004a} and nucleotide alignments
+were constructed by mapping nucleotide sequences onto each protein sequence
+alignment. We then identified recombination within gene alignments using the
+Single Breakpoint Recombination (SBP) method from the HyPhy package
+\citep{KosakovskyPond2006}.  Using the Akaike Information Criterion (AIC), SBP
+inferred that 47.5\% (1,094) of gene families contained at least one
+topology-changing recombination. Each of these families was then split into two
+gene fragments.  Gene trees were constructed from nucleotide alignments for each
+gene fragment using the RAxML program \citep{Stamatakis2006}, and were rooted by
+minimizing inferred duplications and losses.  Each rooted gene tree was then
+reconciled using the \texttt{Mowgli} program \citep{Doyon2011}, which inferred
+the most parsimonious gene duplication, loss, and transfer events for each
+family.  In the case of multiple maximum parsimonious reconciliations, one was
+chosen uniformly at random.  Unless specified otherwise, default setings were
+used for all programs.  Since the method of \texttt{Mowgli} did not directly
+model recombination events, recombination-based transfers were inferred in a
+post-processing step, by checking whether the transfer was coupled with a loss
+of a homologous gene in the recipient branch (indicating gene replacement rather
+than gain).  A transfer event was classified as non-recombining if the
+destination species of that transfer had an extant descendant that contained
+copies of that genes which were descendants of the transferred gene as well as
+copies which were non-descendants.  We noted that this parsimonious approach was
+imperfect and would tend to undercount events, especially recombining gene
+transfer events, which were modeled using two events: a transfer and a loss.
+Nevertheless, this analysis allowed us to study non-recombining gene transfers,
+which were not modeled by \texttt{ClonalOrigin}, and to validate some of the
 \texttt{ClonalOrigin} results on recombining gene transfer.
@@
 \subsection{Functional categories association}
+
+% FIXME
+% \texttt{The recombination intensity scores were taken to be average site-wise
+% intensity scores. SC: THIS NEEDS TO BE REANALYZED WITH RECOMBINATION INTENSITY
+% WITH ONLY TREE-TOPOLOGY CHANGING RECOMBINANT EDGES.} 
 We performed a series of analyses to check for functional gene categories
@@
 uniProt GOA database.  We also classified a category of  \textit{virulent genes}
-using the virulence profile described in \citet{Suzuki2011}. As virulent genes
-we used 207 loci, in which at least one of the genes in the gene family was
-identified by that profile.  Each gene covered by our analysis was associated
-with a recombination intensity score from the \texttt{ClonalOrigin} analysis and
-scores from the Mowgli analysis corresponding to number of duplications, losses,
-non-recombining gene transfers, and homologous recombining gene transfer events.
-\texttt{The recombination intensity scores were taken to be average site-wise
-intensity scores. SC: THIS NEEDS TO BE REANALYZED WITH RECOMBINATION INTENSITY
-WITH ONLY TREE-TOPOLOGY CHANGING RECOMBINANT EDGES.} Mann-Whitney tests were
-used to identify GO categories whose genes showed significantly elevated values
-in each score.  A 5\% False Discovery Rate (FDR) correction
+using the virulence profile described in \citet{Suzuki2011}. Among 207 virulent
+genes at least one of the genes in the gene family was identified by that
+profile.  Each gene covered by our analysis was associated with a recombination
+intensity score from the \texttt{ClonalOrigin} analysis and scores from the
+\texttt{Mowgli} analysis corresponding to number of duplications, losses,
+horizontal gene transfers, and recombining gene transfer events.  Mann-Whitney
+tests were used to identify GO categories whose genes showed significantly
+elevated values in each score.  A 5\% False Discovery Rate (FDR) correction
 \citep{Benjamini1995} was used to determine an appropriate significance
@@
 \subsection{Simulation Study}
+
 We conducted experiments on simulation data to check the accuracy of estimates
@@
 and recombination across the five species.  While populaiton mutation and
-recombination parameters were accurately estimated, the recombinant
-tract length was over-estimated  (Table
-S\ref{tab:sim-three-population-parameters}).  This was not entirely surprising
-as recombinant tract lengths were known to be difficult to estimate from
-relatively short alignment blocks \citep{Didelot2010}. When we performed similar
-simulations which much longer blocks (10000 bp), \texttt{ClonalOrigin} was able
-to estimate the average recombinant tract length accurately (data not shown).
-
-We also evaluated the accuracy of statistics estimated in the second stage of
-\texttt{ClonalOrigin}. The number of recombinant edges for each pair of recombinant
-tree branches was correlated with the true value in general 
-while there were slight over-estimation (Figure S\ref{fig:h1}). 
-However, this comparison between inferrence from simulated data and that of 
-real data set suggested that the directional difference between SPY and SDE was 
-due to neither lack of power of \texttt{ClonalOrigin} nor hidden bias thereof.
-
-Recombination intensities that were measured on genes also showed rough
-correlation (Pearson's correlation coefficient of 0.63) of the estimated and
-true values (Figure S\ref{fig:ri1}).  Estimated values appeared to be driven
-somewhat by the prior implied by the population parameter setting, explaining
-why lower values were over-estimated and higher values were under-estimated.
-Nonetheless, the correlation indicated that the measure of recombination
-intensity used in our functional gene category served as an adequate indicator
-of actual intensity of recombination events along genes.
+recombination parameters were accurately estimated, the recombinant tract length
+was over-estimated  (Table S\ref{tab:sim-three-population-parameters}).  This
+was not surprising as recombinant tract lengths were known to be difficult to
+estimate from relatively short alignment blocks \citep{Didelot2010}. 
+% When we
+% performed simulations which much longer blocks (10000 bp), \texttt{ClonalOrigin}
+% was able to estimate the average recombinant tract length accurately (data not
+% shown).  
+% FIXME: Show this result of 10,000 blocks.
+We also evaluated the accuracy of the statistics estimated in the second stage
+of \texttt{ClonalOrigin}. The number of recombinant edges for each pair of
+recombinant tree branches was correlated with the true value in general while
+there were slight over-estimation (Figure S\ref{fig:h1}).  However, this
+comparison between inferrence from simulated data and that of real data set
+suggested that the directional difference between SPY and SDE was due to neither
+lack of power of \texttt{ClonalOrigin} nor hidden bias thereof.  Recombination
+intensities that were measured on genes also showed rough correlation (Pearson's
+correlation coefficient of 0.63) of the estimated and true values (Figure
+S\ref{fig:ri1}).  Estimated values appeared to be driven somewhat by the prior
+implied by the population parameter setting, explaining why lower values were
+over-estimated and higher values were under-estimated.  Nonetheless, the
+correlation indicated that the measure of recombination intensity used in our
+functional gene category served as an adequate indicator of actual intensity of
+recombination events along genes.
 
@@
 
-\section*{Supplementary Information}
-
-\subsection*{Simulation for the first stage MCMC of ClonalOrigin.}
-We started by testing the accuracy
-of the global estimates obtained from the first stage of \texttt{ClonalOrigin}
-for the three population parameters.  Ten recombinant trees were sampled given
-the three parameters and the species tree  set to the values inferred from the
-five bacterial genomes (see section \textbf{Results}).  Each of the recombinant
-trees was used to simulate the Jukes-Cantor model for DNA sequence evolution to
-create 274 blocks of alignments.  The block lengths configuration was the same
-as that of the blocks used in the analysis of the five genomes.  We ran
-\texttt{ClonalOrigin} on the ten replicate data sets, and obtained global
-estimates by applying the weighted median scheme for the block-specific
-estimates. The means of the ten global estimates of mutation rate per site and
-recombination rate per site were 0.081 with standard deviation (SD) of $0.0011$,
-and 0.011 with SD of $0.0003$, respectively, centering around the true values:
-0.081 and 0.012, respectively. The mean of the ten estimates of the recombinant
-tract length was 919 bp (SD 28), which was significantly higher than the true
-value of 744 bp. This was not entirely surprising as recombinant tract lengths
-were known to be difficult to estimate from relatively short alignment blocks
-\citep{Didelot2010}. When we performed similar simulations which much longer
-blocks (10000 bp), \texttt{ClonalOrigin} was able to estimate the average
-recombinant tract length accurately (data not shown).
-
-Table S\ref{tab:sim-three-population-parameters}.~
-Recombinant tract length was over-estimated.
-
-\subsection*{Simulation study for the second stage MCMC of ClonalOrigin}
-% \texttt{IG: RESULTS FROM THIS POINT ON REQUIRE SOME REFINEMENT. SC: Let's see.}
-Next we evaluated the accuracy of statistics estimated in the second stage of
-\texttt{ClonalOrigin}. We used 100 replicates of data sets obtained by first
-sampling ten recombinant trees given the mutation rate, recombination rate, and
-tract length, and generating ten independent simulations of sequence evolution
-using each of the recombinant trees. For each measure of interest we thus
-obtained 100 separate estimates, which we could compare with either the expected
-value from the ten recombinant trees, or true value from one of the recombinant
-trees.  First, we tested the number of recombinant edges of each of the sixty
-different types. Figure S\ref{fig:h1} plots  the distribution of observed counts
-against the respective expected count given the model parameters and species
-tree for each recombinant edge type.  \texttt{FIX ``TRUE" TO BE EXPECTED VALUE
-RATHER THAN AVERAGE OVER THE 10 REC.  TREES. SC: OKAY. LET'S FIND THE PRIOR
-EXPECTED NUMBER OF EVENTS. THIS MUST BE GIVEN SOMEWHERE}.  The estimated numbers
-appeared to be slightly over-estimated, but in very good correlation with the
-expected numbers.
-
-\texttt{CONSIDER SHOWING DETAILED COMPARISON OF ESTIMATED VALUES TO ONES 
-EXPECTED FROM THE ACTUAL REC. TREES. THESE DON'T LOOK AS NICE, BUT THEY MIGHT 
-BE ABLE TO DEMONSTRATE OUR ABILITY TO DETECT DEPARTURE FROM EXPECTATION. SC:
-THIS MAY BE ABOUT CHANGING AVERAGE TO PRIOR EXPECTED NUMBER OF EVENTS.}
-
-\texttt{WE SHOULD PROBABLY CONSIDER ALTERNATIVE WAYS OF USING THESE RESULTS TO 
-ARGUE THAT OUR DIRECTIONAL SIGNAL IS REAL. SC: I MIGHT CONSIDER SIMULATIONS AND
-ARTIFICIALLY DOUBLE OR TRIPLE THE RECOMBINATION EVENTS OF ONE OF TYPES.}
+\section{Supplementary Information}
+
+\subsection{Population mutation and recombination rates were accurately
+estimated}
+
+% FIXME: Newick tree in text in a file.
+We describe the detailed procedure of simulations for checking if the three
+population parameters were accurately recovered from the first stage of
+\texttt{ClonalOrigin} MCMC.  Ten recombinant trees were sampled given the three
+parameters and the species tree  set to the values inferred from the five
+bacterial genomes (see section \textbf{Results}).  Each of the recombinant trees
+was used to simulate the Jukes-Cantor model for DNA sequence evolution to create
+274 blocks of alignments.  The block lengths configuration was the same as that
+of the blocks used in the analysis of the five genomes.  The mean of the ten
+estimates of the recombinant tract length was significantly higher than the true
+value of (919 bp versus 744 bp) whereas the values of the mutation and
+recombination rates were accurately estimated (Table
+S\ref{tab:sim-three-population-parameters}).  Recombinant tract lengths were
+known to be difficult to estimate from relatively short alignment blocks
+\citep{Didelot2010}. Performing similar simulations with much longer blocks
+(i.e., 10000 bp) we estimated the average recombinant tract length accurately
+\texttt{(Table S data not shown)}.
+
+\subsection{Estimating numbers of recombinant edges were accurate}
+
+% FIXME 1: IG: RESULTS FROM THIS POINT ON REQUIRE SOME REFINEMENT. SC: Let's see.
+% FIXME 2: ``TRUE" TO BE EXPECTED VALUE
+% RATHER THAN AVERAGE OVER THE 10 REC.  TREES. SC: OKAY. LET'S FIND THE PRIOR
+% EXPECTED NUMBER OF EVENTS. THIS MUST BE GIVEN SOMEWHERE}.  
+% FIXME 3: CONSIDER SHOWING DETAILED COMPARISON OF ESTIMATED VALUES TO ONES 
+% EXPECTED FROM THE ACTUAL REC. TREES. THESE DON'T LOOK AS NICE, BUT THEY MIGHT 
+% BE ABLE TO DEMONSTRATE OUR ABILITY TO DETECT DEPARTURE FROM EXPECTATION. SC:
+% THIS MAY BE ABOUT CHANGING AVERAGE TO PRIOR EXPECTED NUMBER OF EVENTS.
+% FIXME 4: WE SHOULD PROBABLY CONSIDER ALTERNATIVE WAYS OF USING THESE RESULTS TO 
+% ARGUE THAT OUR DIRECTIONAL SIGNAL IS REAL. SC: I MIGHT CONSIDER SIMULATIONS AND
+% ARTIFICIALLY DOUBLE OR TRIPLE THE RECOMBINATION EVENTS OF ONE OF TYPES.
+We checked if numbers of recombinant edge types were inferred correctly. Because
+of the two-stage Bayesian method we first sampled ten recombinant trees given
+the mutation rate, recombination rate, and tract length (see section
+\textbf{Results}), and then created ten independent simulated data sets with the
+Jukes-Cantor model for DNA sequence evolution using each of the recombinant
+trees. This consisted of 100 replicates of data sets.  Figure S\ref{fig:h1}
+shows the observed count of each recombinant edge type against the expected
+count; the estimated numbers appeared to be slightly over-estimated, but in
+correlation with the true numbers.
+
+\subsection{Unbalanced gene flow model was adequate}
+
+We checked if the model with unbalanced gene flow was adequate using posterior
+predictive check. We used 100 recombinant trees from the posterior distribution
+of the recombiant tree given the real data of the 274 alignment blocks in order
+to simulate the ClonalOrigin model to create a new data set of 100 274-block
+alignments.  Using the simulated data sets of 100 274-block alignments we
+infer recombinant trees to draw the distribution of log base 2 ratio of
+observed recombinant edge counts to the prior expected number of recombinant
+edges. If the model with unbalanced gene flow was adequate, the log base 2
+ratio from the real data set should center around the distribution. 
+\texttt{What is the result of this simulation?}
+
+\subsection{}
 
@@
 any potential biases hidden in \texttt{ClonalOrigin} implementation. Because the
-simulated data were generated with no directionality between SPY and SDE,  if 
-we estimate recombination events from the simulated data sets using 
-\texttt{ClonalOrigin} and compare those estimates from those from the real data set, 
-this could remove biases in \texttt{ClonalOrigin} implementation. We could use inferred 
-estimates of recombination from the simulated data under the prior model of 
-\texttt{ClonalOrigin} as the prior expected number of recombination. Figure 
-S\ref{fig:h2} shows a plot of numbers of recombination events from the real 
-data against those from the simulated data sets. Note that the numbers of 
-recombination events from simulated data sets appear to deviate from the line 
-of slope being 1 and intercept being 0. The numbers of recombination events for 
+simulated data were generated with no directionality between SPY and SDE,  if we
+estimate recombination events from the simulated data sets using
+\texttt{ClonalOrigin} and compare those estimates from those from the real data
+set, this could remove biases in \texttt{ClonalOrigin} implementation. We could
+use inferred estimates of recombination from the simulated data under the prior
+model of \texttt{ClonalOrigin} as the prior expected number of recombination.
+Figure S\ref{fig:h2} shows a plot of numbers of recombination events from the
+real data against those from the simulated data sets. Note that the numbers of
+recombination events from simulated data sets appear to deviate from the line of
+slope being 1 and intercept being 0.  The numbers of recombination events for
 the recombinant edges from SPY to SDE increase by more than 2.5 fold whereas
@@
 quite accurate. We considered that subtraction of effects of number of
-recombinant edges inferred from simulated data and subtraction of effects of 
-the prior expected number of recombinant edges were effectively equivalent. 
-However, this comparison between inferrence from simulated data and that of 
-real data set suggested that the directional difference between SPY and SDE was 
-due to neither lack of power of \texttt{ClonalOrigin} nor hidden bias thereof.
-
-We tested if our measure of recombination intensity provided a reliable 
-estimate of the true intensity. For this purpose, we artificially annotaded our 
-synthetic blocks to genes according to the real annotations along the real 
-bacterial alignments. For each gene in each of the 100 data sets, we plotted 
-the average estimated recombination intensity along that gene against the true 
+recombinant edges inferred from simulated data and subtraction of effects of the
+prior expected number of recombinant edges were effectively equivalent.
+However, this comparison between inferrence from simulated data and that of real
+data set suggested that the directional difference between SPY and SDE was due
+to neither lack of power of \texttt{ClonalOrigin} nor hidden bias thereof.
+
+\subsection{Simulation for recombination intensity}
+
+We tested if our measure of recombination intensity provided a reliable estimate
+of the true intensity. For this purpose, we artificially annotaded our synthetic
+blocks to genes according to the real annotations along the real bacterial
+alignments. For each gene in each of the simulated 100 data sets, we plotted the
+average estimated recombination intensity along that gene against the true
 average intensity, i.e., the average number of different recombinant edge types
 implied by the recombinant tree used in simulation (Figure S\ref{fig:ri1}).
-Although the scatter plot did not strictly align along the $X=Y$ line, there 
-was clear correlation of the estimated and true values (Pearson's correlation 
+Although the scatter plot did not strictly follow along the $X=Y$ line, there
+was correlation of the estimated and true values (Pearson's correlation
 coefficient of 0.63). Estimated values appeared to be driven somewhat by the
@@
 \begin{center}
-\includegraphics[width=5in]{figures/trans-hgt-heatmap}
+% \includegraphics[width=5in]{figures/trans-hgt-heatmap}
+\includegraphics[width=5in]{figures/heatmap-hgtedge}
 \end{center}
@@
 $\delta^d$ & 744 (346, 2848) & 723 (348, 2870)\tabularnewline
+\hline
 \end{tabular}
@@
 \noindent \centering{}\begin{tabular}{ccccccccc}
-0.176  &   NA  &1.170  &    NA & 0.7564  & NA & 0.67 &    NA & 0.663\tabularnewline
--0.118 & 0.081 &-0.202 &     NA&  0.3395 &  NA& -0.15&  0.045& -0.323\tabularnewline
-1.201  &   NA  &0.092  &    NA & 0.6978  & NA & 0.60 &    NA & 0.387\tabularnewline
--0.629 & 0.022 &-0.661 &0.00670&  0.0078 &  NA& -0.94& -0.299& -1.030\tabularnewline
--0.317 & 0.371 &-0.197 &     NA&  0.0752 &  NA& -1.06& -0.525& -0.915\tabularnewline
--1.745 &-0.698 &-1.786 &0.23039& -0.4854 &  NA& -1.37&  0.497& -1.306\tabularnewline
-1.245  &1.883  &1.251  &    NA &-2.1997  & NA & 0.06 &    NA & 1.062\tabularnewline
-0.089  &0.200  &0.298 &0.00098 &-0.7903  & NA &-0.19 &-0.058 &-0.159\tabularnewline
-1.174  &1.938  &1.287  &    NA &-1.9740  & NA & 1.07 &    NA & 0.063\tabularnewline
+\hline
+& SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2\tabularnewline
+\hline
+SDE1& 0.176  &   NA  &1.170  &    NA & 0.7564  & NA & 0.67 &    NA & 0.663\tabularnewline
+SDE & -0.118 & 0.081 &-0.202 &     NA&  0.3395 &  NA& -0.15&  0.045& -0.323\tabularnewline
+SDE2& 1.201  &   NA  &0.092  &    NA & 0.6978  & NA & 0.60 &    NA & 0.387\tabularnewline
+SD  & -0.629 & 0.022 &-0.661 &0.00670&  0.0078 &  NA& -0.94& -0.299& -1.030\tabularnewline
+SDD & -0.317 & 0.371 &-0.197 &     NA&  0.0752 &  NA& -1.06& -0.525& -0.915\tabularnewline
+ROOT& -1.745 &-0.698 &-1.786 &0.23039& -0.4854 &  NA& -1.37&  0.497& -1.306\tabularnewline
+SPY1& 1.245  &1.883  &1.251  &    NA &-2.1997  & NA & 0.06 &    NA & 1.062\tabularnewline
+SPY & 0.089  &0.200  &0.298 &0.00098 &-0.7903  & NA &-0.19 &-0.058 &-0.159\tabularnewline
+SPY2& 1.174  &1.938  &1.287  &    NA &-1.9740  & NA & 1.07 &    NA & 0.063\tabularnewline
+\hline
 \end{tabular}
@@
 \noindent \centering{}\begin{tabular}{ccccccccc}
-8.896 &  0.00 &17.712  & 0.00 & 12.872   & 0 &12.59  &  0.00  &12.53\tabularnewline
-44.626& 136.49& 42.088 &  0.00& 227.061  &  0& 73.39 &  99.95 & 65.25\tabularnewline
-18.099&   0.00&  8.393 &  0.00&  12.360  &  0& 11.97 &   0.00 & 10.36\tabularnewline
-9.671 &101.74 & 9.464  &31.26 &115.796   & 0 &13.68  & 98.38  &12.86\tabularnewline
-45.539& 168.58& 49.513 &  0.00& 194.706  &  0& 43.09 &  67.14 & 47.65\tabularnewline
-43.553& 602.67& 42.337 &765.53& 802.813  &  0& 98.86 &2146.01 &103.60\tabularnewline
-40.800&  11.46& 40.973 &  0.00&   4.371  &  0& 21.15 &   0.00 & 42.35\tabularnewline
-57.618& 262.08& 66.594 & 30.92& 163.518  &  0& 83.56 & 207.00 & 85.10\tabularnewline
-38.837&  11.90& 42.000 &  0.00&   5.111  &  0& 42.45 &   0.00 & 21.19\tabularnewline
+\hline
+& SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2\tabularnewline
+\hline
+SDE1&8.896 &  0.00 &17.712  & 0.00 & 12.872   & 0 &12.59  &  0.00  &12.53\tabularnewline
+SDE &44.626& 136.49& 42.088 &  0.00& 227.061  &  0& 73.39 &  99.95 & 65.25\tabularnewline
+SDE2&18.099&   0.00&  8.393 &  0.00&  12.360  &  0& 11.97 &   0.00 & 10.36\tabularnewline
+SD  &9.671 &101.74 & 9.464  &31.26 &115.796   & 0 &13.68  & 98.38  &12.86\tabularnewline
+SDD &45.539& 168.58& 49.513 &  0.00& 194.706  &  0& 43.09 &  67.14 & 47.65\tabularnewline
+ROOT&43.553& 602.67& 42.337 &765.53& 802.813  &  0& 98.86 &2146.01 &103.60\tabularnewline
+SPY1&40.800&  11.46& 40.973 &  0.00&   4.371  &  0& 21.15 &   0.00 & 42.35\tabularnewline
+SPY &57.618& 262.08& 66.594 & 30.92& 163.518  &  0& 83.56 & 207.00 & 85.10\tabularnewline
+SPY2&38.837&  11.90& 42.000 &  0.00&   5.111  &  0& 42.45 &   0.00 & 21.19\tabularnewline
+\hline
 \end{tabular}
@@
 \begin{center}
-\includegraphics[width=5in]{figures/blocksize.pdf}
+\includegraphics[width=5in]{figures/blocksize}
 \end{center}
@@
 %=============================================================================
+% Figure S: Scatter plot of rho and log(delta)
+\begin{figure}
+\begin{center}
+\includegraphics[width=5in]{figures/rho-vs-delta}
+\end{center}
+\vspace{-.3in}
+\caption{Scatter plot of recombination rate and average tract length.}
+\label{fig:rhologdelta}
+\end{figure}
+\clearpage{}%
+%=============================================================================
+
+%=============================================================================
 % Figure S5. A Recombinant Tree of ClonalOrigin Model
 \begin{figure}
-\includegraphics[scale=0.42]{figures/mauve-analysis-002}
+\includegraphics[scale=0.42]{figures/mauve-analysis-001}
 \caption{\label{fig:clonalorigin}Recombinant tree in the \texttt{ClonalOrigin} model.
@@
 by all of the three recombinant edges in the subregion between 51 base pairs
-and 70 base pairs.}
+and 70 base pairs.
+
+We first describe procedures of summarizing recombinant trees from the first
+viewpoint.  To summarize the posterior samples of \texttt{ClonalOrigin} we
+examined the distribution of topologies of local genealogies induced by the
+sampled recombinant trees. Recall that a recombinant tree induces a local
+genealogy for each site along the analyzed block.  Figure \ref{fig:clonalorigin}
+shows six local gene trees including one that is the same as the species tree
+from the most left-hand side of the block of 100 base pairs to the right.  We
+expected the dominant topology across sampled genealogies to be concordant with
+that of the species tree. However, other frequently sampled topologies were
+expected to be potentially informative about common recombination events. There
+are 105 possible rooted topologies relating five taxa, which allowed us to count
+topologies induced by a posterior sample of recombinant trees.
+
+For the second viewpoint of recombinant trees \citet{Didelot2010} suggested an
+approach used to trace recombination intensity along the genome.  For each edge
+type and each site, we recorded the sampling frequency of a recombinant edge of
+the type at that site (e.g., Figure S\ref{fig:muts}).  Figure
+\ref{fig:clonalorigin}B illustrates the posterior probability that a site
+experiences influx from a donor branch using a single recombinant tree.  Summing
+these posteriors over a certain set of edge types, we could obtain a measure of
+recombination intensity along the genome (Figure \ref{fig:clonalorigin}B).
+
+}
 \end{figure}
@@
 %=============================================================================
-% Figure XXX: View of UCSC genome browser for bacterial gene transfer.
+% Figure S6: View of UCSC genome browser for bacterial gene transfer.
 \begin{figure}
@@
 \clearpage{}%
-
-% Figure S2: 
+%=============================================================================
+
+%=============================================================================
+% Figure S7: 
 \begin{figure}
@@
 \clearpage{}%
-
-% Figure S3:
+%=============================================================================
+
+%=============================================================================
+% Figure S8:
+\begin{figure}
+\includegraphics{figures/ri1}
+\caption{\label{fig:ri1}The surrogate measure of recombination intensity
+of genes. Each point represents a pair of a true value of average number of
+recombinant edges and its estimated value for a gene. The dashed line is with
+slope of 1 and intercept of 0.}
+\end{figure}
+\clearpage{}%
+%=============================================================================
+
+%=============================================================================
+% Figure S9:
 \begin{figure}
@@
 \subfloat[]{\includegraphics[scale=0.5]{figures/h2-R-inset}}
-
 \subfloat[]{\includegraphics[scale=0.5]{figures/h2-R-inset2}}
@@
 \clearpage{}%
-
-% Figure S4
-\begin{figure}
-\includegraphics{figures/ri1}
-\caption{\label{fig:ri1}The surrogate measure of recombination intensity
-of genes. Each point represents a pair of a true value of average number of
-recombinant edges and its estimated value for a gene. The dashed line is with
-slope of 1 and intercept of 0.}
-\end{figure}
-\clearpage{}%
+%=============================================================================
 
@@
 \begin{center}
-\includegraphics[width=5in]{figures/trans-recomb-heatmap}
+% \includegraphics[width=5in]{figures/trans-recomb-heatmap}
+\includegraphics[width=5in]{figures/heatmap-recedge-mowgli}
 \end{center}
@@
 % MGAS10750\_Spy0455, MGAS10750\_Spy0456 (block 75)
+
+% HGT and virulene genes in detail
+% ORTHOMCL4-1	0	1	6	0	TRUE
+% ORTHOMCL4-1	SDD27957_02925	SEQ_0563	MGAS10750_Spy0339	SpyM3_0298	MGAS10750_Spy0340	SDEG_0469	SDD27957_02245	SDE12394_02260	SDD27957_05495	SDEG_0933	SpyM3_1726	MGAS10750_Spy1807
+% streptococcal C5a peptidase (scpB) (SDEG_0933) 
+% 
+% ORTHOMCL1826-1	0	2	6	0	TRUE
+% ORTHOMCL1826-1	SDEG_1261	SDE12394_05305	SDE12394_03405	SDEG_0634	SDD27957_00125	SDD27957_02625	SDD27957_04790	SDD27957_05475	SDE12394_02555
+% Nothing in Shimomura.
+% 
+% ORTHOMCL912-1	3	2	5	0	TRUE
+% ORTHOMCL912-1	SDD27957_04815	SDD27957_05095	SDD27957_02300	SDD27957_11185	SDD27957_05765	SDD27957_06935	SpyM3_1053	MGAS10750_Spy1189	MGAS10750_Spy1234	SpyM3_1016	SDD27957_02260	MGAS10750_Spy0401	SpyM3_1054	MGAS10750_Spy1235
+% Nothing in Shimomura.
+% SpyM3_1053 /product="putative transposase"
+% SpyM3_1054 /product="putative transposase"
+% SpyM3_1016 /product="IS1076 transposase"
+% 
+% # ORTHOMCL4-2	1	1	4	0	TRUE
+% 
+% # ORTHOMCL29-2	0	0	4	0	TRUE
+% ORTHOMCL29-1	1	1	3	1	TRUE
+% ORTHOMCL29-1	SEQ_1712	SEQ_0303	SpyM3_0174	MGAS10750_Spy0201	SDD27957_10325	SDE12394_09770	SDEG_1964	SEQ_1095	SDEG_0515	MGAS10750_Spy0403
+% Nothing in Shimomura.
+% SpyM3_0174 /product="putative response regulator"
+% 
+% # ORTHOMCL26-2	0	0	3	0	TRUE
+% ORTHOMCL26-1	0	0	3	0	TRUE
+% ORTHOMCL26-1	SpyM3_0982	MGAS10750_Spy1149	SEQ_1437	SDD27957_06480	SDEG_1294	SDE12394_06795	SEQ_2022	SDE12394_00945	SDEG_0221
+% Nothing in Shimomura.
+% SpyM3_0982 /product="putative maltose operon transcriptional repressor"
+% 
+% # ORTHOMCL1564-2	0	2	3	0	TRUE
+% 
+% #ORTHOMCL126-2	0	1	3	0	TRUE
+% ORTHOMCL126-1	0	1	3	0	TRUE
+% ORTHOMCL126-1	SEQ_0301	SDD27957_10335	SDEG_1966	SpyM3_0171	MGAS10750_Spy0199	SEQ_1711	MGAS10750_Spy0404	SDE12394_02570	SDEG_0516
+% Nothing in Shimomura.
+% SpyM3_0171 /product="putative histidine kinase"
+% 
+% ORTHOMCL1071-2	1	2	3	2	TRUE
+% % ORTHOMCL1071-2	SEQ_0845	SEQ_0190	SEQ_1729	MGAS10750_Spy0620	SpyM3_0731	SpyM3_0922	SpyM3_1208	SDEG_1106	SDD27957_07165
+% Nothing in Shimomura.
+% SpyM3_0731 /product="putative holin - phage associated"
+% SpyM3_0922 /product="putative cell wall hydrolase - phage associated"
+% SpyM3_1208 /product="putative cell wall hydrolase, lysin - phage associated"
+% 
+% ORTHOMCL1037-1	0	2	3	0	TRUE
+% ORTHOMCL1037-1	MGAS10750_Spy1697	SDE12394_04580	MGAS10750_Spy0948	SDD27957_05395	SDD27957_04295	SDE12394_07035
+% Nothing in Shimomura.
+% 
+% ORTHOMCL2721	0	4	2	0	TRUE
+% ORTHOMCL2721	SDE12394_10035	SDE12394_09945	SDE12394_09940
+% Nothing in Shimomura.
+% 
+% # ORTHOMCL1826-2	4	0	2	0	TRUE
+% # ORTHOMCL1826-2	SDEG_1261	SDE12394_05305	SDE12394_03405	SDEG_0634	SDD27957_00125	SDD27957_02625	SDD27957_04790	SDD27957_05475	SDE12394_02555
+% 
+% ORTHOMCL1564-1	0	2	2	0	TRUE
+% ORTHOMCL1564-1	SDEG_1103	SpyM3_1409	SDD27957_09010	SEQ_0696	SDD27957_02745	SDEG_0541
+% GGS_124 possesses 5 genes that encode putative nucleases with a secretion signal peptide: genome-encoded streptodornase (SDEG_0541), others,  prophage-associated deoxyribonuclease (SDEG_1103), all of which are predicted to code for a secretion signal peptide.
+% SpyM3_1409 /product="streptodornase (Sdn) - phage associated"
+% 
+% 
+% ORTHOMCL1439-1	0	1	2	0	TRUE
+% ORTHOMCL1439-1	SEQ_1606	SEQ_1999	SpyM3_0104	SDE12394_09890	SDD27957_10420	SDE12394_09885	SDEG_1984
+% Nothing in Shimomura.
+% SpyM3_0104 /product="protein F2 like fibronectin-binding protein"
+% 
+% ORTHOMCL1369-2	2	1	2	0	TRUE
+% ORTHOMCL1369-2	MGAS10750_Spy1683	SDEG_1338	SDEG_1346	SDE12394_07225	SDE12394_04610	SDD27957_04325	SDD27957_05365	SDD27957_08360
+% Nothing in Shimomura.
+% 
+% ORTHOMCL1283-1	3	0	2	0	TRUE
+% ORTHOMCL1283-1	SDD27957_00840	SDEG_1543	SDEG_1531	SDE12394_03300	SDEG_0983	SDE12394_05495	SDE12394_08020	SDE12394_07960
+% Nothing in Shimomura.
+% 
+% ORTHOMCL1232-2	0	1	2	0	TRUE
+% ORTHOMCL1232-2	SDE12394_10675	SDE12394_10850	SDEG_2142	MGAS10750_Spy1897	SDD27957_00535	SDD27957_02435
+% Nothing in Shimomura.
+% 
+% ORTHOMCL1229-2	0	2	2	0	TRUE
+% ORTHOMCL1229-2	SDD27957_03510	SDD27957_05295	SDD27957_08295	SEQ_0753	MGAS10750_Spy1709
+% Nothing in Shimomura.
+% 
+% --------------------------------
+% 
+% 
+% ORTHOMCL9-2	1	1	1	0	TRUE
+% ORTHOMCL9-2	SpyM3_0160	MGAS10750_Spy0188	SEQ_0271a	SEQ_0289a	SDD27957_10400	SDE12394_09855	SDEG_1980	SpyM3_1853
+% 
+% ORTHOMCL912-2	5	1	1	0	TRUE
+% ORTHOMCL9-1	1	0	1	0	TRUE
+% ORTHOMCL848-2	0	1	1	0	TRUE
+%

### 2011-10-04 pl/sub-bed.pl

@@ -1,0 +1,47 @@
+###############################################################################
+# Copyright (C) 2011 Sang Chul Choi
+#
+# This file is part of Mauve Analysis.
+# 
+# Mauve Analysis is free software: you can redistribute it and/or modify
+# it under the terms of the GNU General Public License as published by
+# the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# Mauve Analysis is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with Mauve Analysis.  If not, see <http://www.gnu.org/licenses/>.
+###############################################################################
+use strict;
+use warnings;
+
+sub rnaseqBedParse ($) {
+  my ($f) = @_;
+  my @v;
+  open BED, $f or die "cannot open $f $!";
+  while (<BED>)
+  {
+    # print $outfile "chr1\t$t->{start}\t$t->{end}\t$t->{name}\t$t->{score}\t$t->{strand}\n";
+    chomp;
+    my @e = split /\t/;
+    my $rec = {};
+    $rec->{chr}    = $e[0];
+    $rec->{start}  = $e[1];
+    $rec->{end}    = $e[2];
+    $rec->{name}   = $e[3];
+    $rec->{score}  = $e[4];
+    if (scalar @e >= 6)
+    {
+      $rec->{strand} = $e[5];
+    }
+    push @v, $rec;
+  }
+  close BED;
+  return @v;
+}
+
+1;

### 2011-10-04 doc/manuscript.tex
The manuscript needs to be revised.

@@
 
-\subsubsection*{Version: 0.66 (Sep 30, 2011) \clearpage{}}
+\subsubsection*{Version: 0.67 (Oct 7, 2011) \clearpage{}}
 
@@
 \caption{\label{tab:functional}Functional categories asscoiated with higher
-recombination.}
-\begin{tabular}{|c|c|c|}
+recombination - Total number of recombinant edge types.}
+\begin{tabular}{ccc}
 \hline 
@@
 \hline 
-\hline 
-\multicolumn{3}{|c|}{\# Total number of recombinant edge types}\tabularnewline
-\hline 
 2.994e-07 &  28 & rRNA binding\tabularnewline
-\hline 
 1.447e-06 &  42 & structural constituent of ribosome\tabularnewline
-\hline 
 1.694e-05 &  39 & ribonucleoprotein complex\tabularnewline
-\hline 
 2.048e-05 &  23 & fatty acid biosynthetic process\tabularnewline
-\hline 
 3.664e-05 &  72 & ribosome\tabularnewline
-\hline 
 1.018e-04 &  86 & translation\tabularnewline
-\hline 
 1.733e-04 &  22 & lipid biosynthetic process\tabularnewline
-\hline 
 2.792e-04 & 108 & RNA binding\tabularnewline
-\hline 
 4.164e-04 &  47 & carbohydrate transport\tabularnewline
-\hline 
-1.303e-03 &  38 & phosphoenolpyruvate-dependent sugar phosphotransferase 
-system\tabularnewline
-\hline 
+1.303e-03 &  38 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
 1.345e-03 &  10 & pentose-phosphate shunt\tabularnewline
-\hline 
-1.819e-03 &  35 & protein-N(PI)-phosphohistidine-sugar phosphotransferase 
-activity\tabularnewline
-\hline 
-\multicolumn{3}{|c|}{\# SDE->SPY recombinant edge types}\tabularnewline
-\hline 
-0.0001236 & 28 & rRNA binding\tabularnewline
-\hline 
-0.0001284 & 86 & translation\tabularnewline
-\hline 
-0.0001711 & 39 & ribonucleoprotein complex\tabularnewline
-\hline 
-0.0002230 & 10 & pentose-phosphate shunt\tabularnewline
-\hline 
-0.0003412 & 38 & phosphoenolpyruvate-dependent sugar phosphotransferase 
-system\tabularnewline
-\hline 
-0.0004498 & 42 & structural constituent of ribosome\tabularnewline
-\hline 
-0.0007789 & 47 & carbohydrate transport\tabularnewline
-\hline 
-0.0008152 & 35 & protein-N(PI)-phosphohistidine-sugar phosphotransferase 
-activity\tabularnewline
-\hline 
-\multicolumn{3}{|c|}{\# SPY->SDE recombinant edge types}\tabularnewline
-\hline 
-2.025e-06 & 47 & carbohydrate transport\tabularnewline
-\hline 
-1.875e-04 & 38 & phosphoenolpyruvate-dependent sugar phosphotransferase 
-system\tabularnewline
-\hline 
-2.843e-04 & 18 & sodium ion transport\tabularnewline
-\hline 
-3.248e-04 & 35 & protein-N(PI)-phosphohistidine-sugar phosphotransferase 
-activity\tabularnewline
+1.819e-03 &  35 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
 \hline 
 \end{tabular}
-
 \end{table}
-
-\clearpage{}%
-
-% Table 2. 
+\clearpage{}
+
+%
 \begin{table}
-\caption{\label{tab:go-events}
-GO-term enrichment for different evolutionary events. }
-
-\begin{tabular}{rrll}
+\caption{\label{tab:go-events}GO-term enrichment for different
+evolutionary events - Non-recombining transfers.}
+\begin{tabular}{cccc}
+\hline 
 P-value & Count & GO-term & Description \\
-\hline
-\multicolumn{4}{c}{Duplications} \\
-\hline
-4e-81 &  67 & GO:0006313 & transposition, DNA-mediated\\
-5e-80 &  59 & GO:0004803 & transposase activity\\
-6e-70 &  23 & GO:0032196 & transposition\\
-2e-58 &  52 & GO:0015074 & DNA integration\\
-8e-25 &  86 & GO:0006310 & DNA recombination\\
-1e-20 &  36 & GO:0016987 & sigma factor activity\\
-1e-20 &  36 & GO:0006352 & transcription initiation\\
-2e-19 & 197 & GO:0003676 & nucleic acid binding\\
-2e-10 & 527 & GO:0003677 & DNA binding\\
-4e-06 & 142 & GO:0043565 & sequence-specific DNA binding\\
-\hline
-\hline
-\multicolumn{4}{c}{Losses} \\
-\hline
-1e-06 &  67 & GO:0006313 & transposition, DNA-mediated\\
-2e-05 &  59 & GO:0004803 & transposase activity\\
-5e-05 &  52 & GO:0015074 & DNA integration\\
-5e-04 &  23 & GO:0032196 & transposition\\
-5e-04 &  10 & GO:0046873 & metal ion transmembrane transporter activity\\
-7e-04 &  36 & GO:0016987 & sigma factor activity\\
-7e-04 &  36 & GO:0006352 & transcription initiation\\
-1e-03 &  31 & GO:0006814 & sodium ion transport\\
-3e-03 &  86 & GO:0006310 & DNA recombination\\
-6e-03 & 197 & GO:0003676 & nucleic acid binding\\
-\hline
-\hline
-\multicolumn{4}{c}{Non-recombining transfers} \\
 \hline 
@@
 \hline 
-\hline
-\multicolumn{4}{c}{Recombining transfers} \\
-\hline
-7e-09 & 153 & GO:0006412 & translation\\
-1e-08 &  67 & GO:0030529 & ribonucleoprotein complex\\
-1e-07 &  45 & GO:0019843 & rRNA binding\\
-2e-06 &  73 & GO:0003735 & structural constituent of ribosome\\
-3e-05 & 157 & GO:0005840 & ribosome\\
-7e-05 & 221 & GO:0003723 & RNA binding\\
-2e-04 &  11 & GO:0004826 & phenylalanine-tRNA ligase activity\\
-3e-04 &  14 & GO:0015413 & nickel-transporting ATPase activity\\
-4e-04 &  93 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
-5e-04 & 104 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
 \end{tabular}
-
 \end{table}
-
 \clearpage{}
@@
 %=============================================================================
-% FIGURE 2. - Transfer heatmap
+% FIGURE 2. - HGT Transfer heatmap
 \begin{figure}
@@
 % \includegraphics[width=5in]{figures/trans-hgt-heatmap}
-\includegraphics[width=5in]{figures/heatmap-hgtedge}
+\includegraphics[scale=0.45]{figures/heatmap-hgtedge}
 \end{center}
-\vspace{-1.2in}
 \caption{Distribution of horizontal gene transfers.}
@@
 \end{figure}
+\clearpage{}%
 %=============================================================================
@@
 \caption{\label{tab:heatmap}}
-\noindent \centering{}\begin{tabular}{ccccccccc}
-\hline
-& SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2\tabularnewline
+\noindent \centering{}\begin{tabular}{cccccccccc}
+\hline
+& SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2 \tabularnewline
 \hline
@@
 \caption{\label{tab:obsheatmap}}
-\noindent \centering{}\begin{tabular}{ccccccccc}
-\hline
-& SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2\tabularnewline
+\noindent \centering{}\begin{tabular}{cccccccccc}
+\hline
+& SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2 \tabularnewline
 \hline
@@
 Recombinant tract length & 744 bp & 919 bp (28)\tabularnewline
+\hline
 \end{tabular}
@@
 topologies.}
-\begin{tabular}{|c|c|}
+\noindent \begin{centering}
+\begin{tabular}{cc}
 \hline 
@@
 \hline
-\hline 
 (((SDE1,SDE2),SDD),(SPY1,SPY2)) & 66.7\%\tabularnewline
-\hline 
 (((SDE1,SDE2),(SPY1,SPY2)),SDD) & 9.3\%\tabularnewline
-\hline 
 ((SDE1,SDE2),(SDD,(SPY1,SPY2))) & 3.5\%\tabularnewline
-\hline 
 (((SDE1,(SDE2,SDD)),(SPY1,SPY2)) & 1.5\%\tabularnewline
-\hline 
 (((SDE1,SDD),SDE2),(SPY1,SPY2)) & 1.3\%\tabularnewline
-\hline 
 ((SDE1,SDD),(SDE2,(SPY1,SPY2))) & 1.2\%\tabularnewline
-\hline 
 ((((SDE1,SDE2),SPY1),SPY2),SDD) & 1.0\%\tabularnewline
-\hline 
 ((SDE2,SDD),(SDE1,(SPY1,SPY2))) & 1.0\%\tabularnewline
@@
 \end{tabular}
+\par\end{centering}
 \end{table}
@@
 {\footnotesize
-\begin{tabular}{|lllllll|}
+\begin{tabular}{lllllll}
 \hline 
@@
 
+%
+\begin{table}
+\caption{\label{tab:functional-sde-spy}Functional categories asscoiated with higher
+recombination - \# SDE->SPY recombinant edge types.}
+\begin{tabular}{ccc}
+\hline 
+p-value & count & description\tabularnewline
+\hline 
+0.0001236 & 28 & rRNA binding\tabularnewline
+0.0001284 & 86 & translation\tabularnewline
+0.0001711 & 39 & ribonucleoprotein complex\tabularnewline
+0.0002230 & 10 & pentose-phosphate shunt\tabularnewline
+0.0003412 & 38 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
+0.0004498 & 42 & structural constituent of ribosome\tabularnewline
+0.0007789 & 47 & carbohydrate transport\tabularnewline
+0.0008152 & 35 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
+\hline 
+\end{tabular}
+\end{table}
+\clearpage{}
+
+%
+\begin{table}
+\caption{\label{tab:functional-spy-sde}Functional categories asscoiated with higher
+recombination - \# SPY->SDE recombinant edge types.}
+\begin{tabular}{ccc}
+\hline 
+p-value & count & description\tabularnewline
+\hline 
+2.025e-06 & 47 & carbohydrate transport\tabularnewline
+1.875e-04 & 38 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
+2.843e-04 & 18 & sodium ion transport\tabularnewline
+3.248e-04 & 35 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
+\hline 
+\end{tabular}
+\end{table}
+\clearpage{}
+
+%
+\begin{table}
+\caption{\label{tab:go-events}GO-term enrichment for different
+evolutionary events - Duplication.}
+\begin{tabular}{cccc}
+\hline 
+P-value & Count & GO-term & Description \\
+\hline 
+4e-81 &  67 & GO:0006313 & transposition, DNA-mediated\\
+5e-80 &  59 & GO:0004803 & transposase activity\\
+6e-70 &  23 & GO:0032196 & transposition\\
+2e-58 &  52 & GO:0015074 & DNA integration\\
+8e-25 &  86 & GO:0006310 & DNA recombination\\
+1e-20 &  36 & GO:0016987 & sigma factor activity\\
+1e-20 &  36 & GO:0006352 & transcription initiation\\
+2e-19 & 197 & GO:0003676 & nucleic acid binding\\
+2e-10 & 527 & GO:0003677 & DNA binding\\
+4e-06 & 142 & GO:0043565 & sequence-specific DNA binding\\
+\hline 
+\end{tabular}
+\end{table}
+\clearpage{}
+
+%
+\begin{table}
+\caption{\label{tab:go-events}GO-term enrichment for different
+evolutionary events - Loss.}
+\begin{tabular}{cccc}
+\hline 
+P-value & Count & GO-term & Description \\
+\hline 
+1e-06 &  67 & GO:0006313 & transposition, DNA-mediated\\
+2e-05 &  59 & GO:0004803 & transposase activity\\
+5e-05 &  52 & GO:0015074 & DNA integration\\
+5e-04 &  23 & GO:0032196 & transposition\\
+5e-04 &  10 & GO:0046873 & metal ion transmembrane transporter activity\\
+7e-04 &  36 & GO:0016987 & sigma factor activity\\
+7e-04 &  36 & GO:0006352 & transcription initiation\\
+1e-03 &  31 & GO:0006814 & sodium ion transport\\
+3e-03 &  86 & GO:0006310 & DNA recombination\\
+6e-03 & 197 & GO:0003676 & nucleic acid binding\\
+\hline 
+\end{tabular}
+\end{table}
+\clearpage{}
+
+%
+\begin{table}
+\caption{\label{tab:go-events}GO-term enrichment for different
+evolutionary events - Recombining transfers.}
+\begin{tabular}{cccc}
+\hline 
+P-value & Count & GO-term & Description \\
+\hline 
+7e-09 & 153 & GO:0006412 & translation\\
+1e-08 &  67 & GO:0030529 & ribonucleoprotein complex\\
+1e-07 &  45 & GO:0019843 & rRNA binding\\
+2e-06 &  73 & GO:0003735 & structural constituent of ribosome\\
+3e-05 & 157 & GO:0005840 & ribosome\\
+7e-05 & 221 & GO:0003723 & RNA binding\\
+2e-04 &  11 & GO:0004826 & phenylalanine-tRNA ligase activity\\
+3e-04 &  14 & GO:0015413 & nickel-transporting ATPase activity\\
+4e-04 &  93 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
+5e-04 & 104 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
+\hline 
+\end{tabular}
+\end{table}
+\clearpage{}
+
 \section*{Supplementary Figures}
@@
 \begin{center}
-\includegraphics[scale=0.5]{figures/mauve.pdf}
+\includegraphics[scale=0.2]{figures/mauve}
 \end{center}
-\vspace{-1.2in}
 \caption{Alignment of the five genomes.}
@@
 and 70 base pairs.
-
 We first describe procedures of summarizing recombinant trees from the first
@@
 topologies induced by a posterior sample of recombinant trees.  
-
 For the second viewpoint of recombinant trees \citet{Didelot2010} suggested an
@@
 recombination intensity along the genome (Figure \ref{fig:clonalorigin}B).  
-
 }
 \end{figure}
+\clearpage{}%
 %=============================================================================
@@
 \begin{figure}
-\includegraphics[scale=0.6]{figures/ucsc}
+\includegraphics[scale=0.8]{figures/ucsc}
 \caption{\label{fig:muts}Posterior probability of recombinant edges on 
@@
 \subfloat[]{\includegraphics[scale=0.5]{figures/h2-R-inset}}
+
 \subfloat[]{\includegraphics[scale=0.5]{figures/h2-R-inset2}}
@@
 % \includegraphics[width=5in]{figures/trans-recomb-heatmap}
-\includegraphics[width=5in]{figures/heatmap-recedge-mowgli}
+\includegraphics[scale=0.45]{figures/heatmap-recedge-mowgli}
 \end{center}
-\vspace{-1.2in}
 \caption{Distribution of recombining horizontal gene transfers.}

### 2011-10-06 doc/manuscript.tex
The manuscript is being rewritten.

@@
 Bacteria infect humans with a variety of diseases: e.g., food poisoning
-(\textit{Eschericha Coli}), strep throat (\textit{Streptococcus pyrogenes}),
+(\textit{Eschericha Coli}), strep throat (\textit{Streptococcus pyogenes}),
 tuberculosis (\textit{Mycobacterium tuberculosis}), cholera (\textit{Vibrio
@@
 streptococcal diseases including necrotizing fasciitis and toxic shock syndrome
-\citep{Cunningham2000a}.  Although \textit{Streptococcus dysgalactiae}
-subspecies \textit{equisimilis} (SDE) belonging to group C or G streptococci
-(GCS/GGS) used to be generally regarded as a veterinary pathogen
-\citep{Vandamme1996}, it has been reported to cause serious and life-threatening
-streptococcal diseases to humans traditionally associated with GAS
-\citep[e.g.,][]{Brandt2009}.  Because of overlap in the disease profile and the
-ecological niche of tissue sites betwen SPY and SDE, these bacteria may as well
-interact with each other by exchanging parts of their genomes.
-
-Therefore, the research topic on gene exchanges between SPY and SDE has been
-discussed in various ways. \citet{Sachse2002} showed gene transfer from SDE to
-SPY for the downstream of the genomic region of streptococcal pyrogenic
-exotoxins. \citet{Davies2007} investigated phage 3396 from SDE as
-a vehicle of gene transfer from SDE to SPY. \citet{Davies2009} demonstrated that
-an integrative conjugative element of SDE allowed gene transfer from SDE to
-other streptococci.  \citet{Ahmad2009} did not conclusively deduce the net
+\citep{Cunningham2000a}.  Its close relative, \textit{Streptococcus
+dysgalactiae} subspecies \textit{equisimilis} (SDE) belonging to group C or G
+streptococci (GCS/GGS), used to be generally regarded as a veterinary pathogen
+\citep{Vandamme1996}. However, it has been reported to cause serious and
+life-threatening streptococcal diseases to humans traditionally associated with
+GAS \citep[e.g.,][]{Brandt2009}.  Because of the overlap in disease profile and
+ecological niche of tissue sites betwen SPY and SDE, these bacteria may as
+well interact with each other by exchanging parts of their genomes.
+
+Therefore, the research topic on gene exchange between SPY and SDE has been
+performed in various ways. \citet{Sachse2002} showed gene transfer from SDE to
+SPY for the downstream of the genomic region of streptococcal pyogenic
+exotoxins. \citet{Davies2007} investigated phage 3396 from SDE as a vehicle of
+gene transfer from SDE to SPY. \citet{Davies2009} demonstrated that an
+integrative conjugative element of SDE allowed gene transfer from SDE to other
+streptococci.  \citet{Ahmad2009} did not conclusively deduce the net
 directionality of housekeeping gene transfer between SPY and SDE.  Notably,
-\citet{Kalia2001} attempted to show that the direction of gene transfer from SPY
-to SDE was more dominant than the other direction.  Because these researches on
-the controversial issue, net directionality of gene transfer between SPY and
+\citet{Kalia2001} attempted to show that the net direction of gene transfer from
+SPY to SDE was more dominant than the other direction.  Because these researches
+on the controversial issue, net directionality of gene transfer between SPY and
 SDE, have been limited to relatively small regions in the genomes primarily on
@@
 \citep{Tettelin2009a}.  Studies on horizontal gene transfer in prokaryotes have
-used genes in very large sets of species \citep{Koonin2001} from a great deal
-of genomes available.  Although this type of study gave us insights into complex
+used genes in very large sets of species \citep{Koonin2001} from a great deal of
+genomes available.  Although this type of study gave us insights into deep
 evolutionary histories of prokaryotes, with multiple individual genomes per
@@
 \textit{Lactobacillus bulgaricus} and \textit{Streptococcus thermophilus} using
-the compositional approach developed by \citet{Karlin2001}. \citet{Luo2011}
-investigated gene transfer events in \textit{Eschericha Coli} using the embedded
-quartet decomposition analysis developed by \citet{Zhaxybayeva2006}.
-\citet{Hamady2006} combined phylogenetic and compositional approaches to
-detecting horizontal gene transfer.  \citet{Caro-Quintero2011} studied a number
-of \textit{Shewanella baltica} genomes to identify the genetic elements that
-enabled the species to adapt to redox gradients.  Although several complete
-genomes from SPY have been available, \citet{Suzuki2011} recently added one more
-complete genome of SDE to the previous set of a unique SDE genome
-\citep{Shimomura2011}. The multiple SDE genomes prompted us to study
-genome-wide gene flows between the two species that afflicted human beings
-around the globe.
+the compositional approach of \citet{Karlin2001}. \citet{Luo2011} investigated
+gene transfer events in \textit{Eschericha Coli} using the embedded quartet
+decomposition due to \citet{Zhaxybayeva2006}.  \citet{Hamady2006} combined
+phylogenetic and compositional approaches to detecting horizontal gene transfer.
+\citet{Caro-Quintero2011} studied a number of \textit{Shewanella baltica}
+genomes to identify the genetic elements that enabled the species to adapt to
+redox gradients.  Although several complete genomes from SPY have been
+available, \citet{Suzuki2011} just recently added one more complete genome of
+SDE to the set of a previously unique SDE genome \citep{Shimomura2011}. These
+multiple SDE genomes prompted us to study genome-wide gene flows between SPY and
+SDE that afflicted human beings around the globe.
 
@@
 through substitutions, insertions, and deletions of deoxyribonucleic acids.
-Horizontal gene transfers can reshape the organization of a bacterial genome,
-which could revamp biological functions of the bacterial species.  Recipient
-bacteria of transferred genes can substitute transferred genes for homologous
-parts of their genomes, or integrate transferred genes into their genomes
-without replacement. These two processes are here in this paper discriminated by
-calling the former \textit{recombining gene transfer}, and the latter
-\textit{horizontal gene transfer} \citep{Ochman2001,Lawrence2009}.  Despite the
-two distinct evolutionary processes in bacteria, analyses of complete genome
-sequences in efforts of better understanding bacterial evolution have paid less
-attention to a combined approach to illuminating disparity of the two gene
-transfer processes. We attempted here to use and compare two approaches for
+Gene transfers can reshape the organization of a bacterial genome,
+which could revamp biological functions of the species.  Recipient bacteria of
+transferred genes can substitute transferred genes for homologous parts of their
+genomes, or integrate transferred genes into their genomes without replacement.
+These two processes are here in this manuscript discriminated by calling the
+former \textit{recombining gene transfer}, and the latter \textit{horizontal
+gene transfer} \citep{Ochman2001,Lawrence2009}.  Despite the two distinct
+evolutionary processes in bacteria, analyses of complete genome sequences in
+efforts of better understanding bacterial evolution have paid less attention to
+a combined approach to illuminating disparity of the two gene transfer
+processes. We attempted here to use and compare two approaches for
 distinguishing recombining gene transfer process from horizontal, or
-non-recombining, gene transfer, and more systematically tackled their roles
-in association with virulence genes.
+non-recombining, gene transfer, and tackled their roles in association with
+virulence genes.
 
 Homologous recombining gene transfer could be studied more readily using the
-core genomes from multiple individual bacterial genomes while horizontal gene
-transfer might be more easily detected using the pan genomes. Recently, a method
-of inferring homologous recombination with whole bacterial genomes was developed
+core genomes of individual bacteria while horizontal gene transfer might be more
+easily detected using the pan genomes. Recently, a method of inferring
+homologous recombination with whole bacterial genomes was developed
 \citep{Didelot2010}. The model of homologous recombination allowed them to study
-flows of genetic segment between bacterial individuals of \textit{Bacillus
-cereus} species in the core genomes.  More complex evolutionary processes acting
-on bacterial genomes include genome rearragement, translocation, inversion,
-duplication, gain, loss, etc.  Because of the complexity, these processes do not
-readily lend themselves to statistically more sound model-based approaches such
-as maximum likelihood. Instead, a parsimony-based approach to inferring gene
-gain, loss, and transfer events was developed by  \citep{Doyon2011}.  We noted
-that the first approach could be used to infer transfer of genomic segments due
-to homologous recombination, and the second could do similarly attributed to
-horizontal gene transfer.
-
-In short, the complete genomes of the two spcies of SPY and SDE presented a
+flows of genetic segment between bacteria individuals of \textit{Bacillus
+cereus} species in the core genomes.  Other evolutionary processes acting on
+bacterial genomes include genome rearragement, translocation, inversion,
+duplication, gain, loss, etc.  These processes did not readily lend themselves
+to statistically more sound model-based approaches such as maximum likelihood.
+Instead, a parsimony-based approach to inferring gene gain, loss, and transfer
+events was developed by \citep{Doyon2011}.  We noted that the first approach
+could be used to infer transfer of genomic segments due to homologous
+recombination, and the second could do similarly attributed to horizontal gene
+transfer.
+
+In short, the complete genomes of the two spcies of SPY and SDE presented us a
 unique opportunity to study gene transfer over relatively recent time scales.
-We used both of the model-based approach for inferring homologous recombination
-and the parsimony-based one for detecting gene gain, loss, and horizontal
-transfer. We also compared both of the results in recombining gene transfer
-events to show that the parimony-based method inferred recombining gene transfer
-events in similar biological functional categories to those inferred with the
-model-based approach. This was the first attempt as far as we know to analyaze
-SPY and SDE in genome-wide scale for detecting gene transfer events not only in
-the core genome but also in the pan genome, which make the analysis complete in
-terms of the amount of used genome data. Our analysis of core genomes also
-showed for the first time that the net directionality of, at least, recombining
-gene transfer from SPY to SDE was stronger than the reverse direction.
+We took advantage of the model-based approach for inferring homologous
+recombination, and the parsimony-based one for detecting gene duplication, loss,
+and horizontal transfer. We compared both of the results in recombining gene
+transfer events to show that the parimony-based method inferred recombining gene
+transfer events in biological functional categories similar to those inferred
+with the model-based approach. This was the first attempt as far as we know to
+analyaze SPY and SDE in genome-wide scale for detecting gene transfer events not
+only in the core genome but also in the pan genome, which makes the analysis
+complete in terms of the amount of used genome data. Our analysis of core
+genomes also showed for the first time that the net directionality of
+recombining gene transfer from SPY to SDE was stronger than the reverse
+direction.
 
@@
 
-\subsection{Recombination dirved more substitutions than did point mutation}
+\subsection{Recombination drived more substitutions than did point mutation}
 
 Five genomes except \textit{S.\ equi} ssp.\textit{\ equi} strain 4047 were
-chosen (Table S\ref{tab:genome}) to study recombining gene transfer.  The genome
-alignment of the five genomes showed two distinct genome structures of SPY and
-SDE (Figure S\ref{fig:mauve}). The SDE genomes were larger in length than those
-of SPY. Although it was unclear whether SDD was different from either of SDE1
-and SDE2 from the graphical view of the genome alignment, the spcies tree
-inferred from the genome alignment showed that SDD branch was attached to the
-clade branch of SDE1 and SDE2 (Figure S\ref{fig:tree5}). The distribution of the
-alignment block lengths was skewed to the right with mean of 4188 base pairs
-long (Figure S\ref{fig:blocksize}), which implied that the model assumption of
-geometric block length distribution might be plausible in \texttt{ClonalOrigin}
-method (see section \textbf{Materials and Method} and \citealt{Didelot2010}).
-The relative strength of recombination compared to mutations indicated that
-reconmination-driven substitutions affected more sites than did mutation-driven
-substitutions (Table S\ref{tab:clonalframe}). In the \textit{Streptococcus}
-bacteria homologous recombination appeared to affect more sites than did point
-mutations process.
+chosen (Table S\ref{tab:genome}) to study recombining gene transfer using the
+model-based approach.  The genome alignment of the five genomes showed two
+distinct genome structures of SPY and SDE (Figure S\ref{fig:mauve}). The SDE and
+SDD genomes were larger in length than those of SPY. Although it was unclear
+whether SDD was different from either of SDE1 and SDE2 from the graphical view
+of the genome alignment, the spcies tree inferred from the genome alignment
+showed that SDD branch was attached to the clade branch of SDE1 and SDE2 (Figure
+S\ref{fig:tree5}). The distribution of the alignment block lengths was skewed to
+the right with mean of 4188 base pairs long (Figure S\ref{fig:blocksize}), which
+supported that the model assumption of geometric block length distribution might
+be plausible in \texttt{ClonalOrigin} method (see section \textbf{Materials and
+Methods} and \citealt{Didelot2010}).  The relative strength ($r/m$) of
+recombination compared to mutations indicated that recombination-driven
+substitutions affected more sites than did mutation-driven substitutions (Table
+S\ref{tab:clonalframe}). This could imply that in the \textit{Streptococcus}
+species bacterial recombination process appeared to accelerate molecular
+evolution of more sites than did point mutations process.
 
@@
 mutation rates were larger than twice the global estimate of mutation rate.  The
-recombination rate appeared to change more dynamically. Espeically, the genomic
-region of SDE1 near 20 kilobase pairs from the origin of replication contained
-alignment blocks with bell-shaped recombination rates.  The recombinant tract
-length changes in a way that we could not note any particular positions along
-the genome.  There appeared to be negative correlation between recombinant tract
-length and recombination rate (Figure S\ref{fig:rhologdelta}).
-% \texttt{WHAT ARE THESE BLOCKS? WHAT GENES ARE LOCATED IN THE BLOCK?}
-% \texttt{USE UCSC GENOME BROWSER TO FIND WHAT GENES BELONG TO THOSE BLOCKS.}
+recombinant tract length changed in a way that we could not note any particular
+positions with notable tract length values compared with other positions along
+the genome.  As recombinant tract lengths were smaller, recombination rates per
+site seemed larger (Figure S\ref{fig:rhologdelta}).  The recombination rate
+appeared to change more dynamically. Especially, a genomic region of SDE1 near
+20 kilobase pairs from the origin of replication contained alignment blocks with
+bell-shaped recombination rates.
 
@@
 
-Most of local gene tree topologies (see section \textbf{Materials and Methods}),
-or 66.7\%, along the alignment blocks were the same as the species tree, which
-allowed us to be more confident of the species tree inferred using the alignment
-blocks (Table S\ref{tab:Gene-tree-topologies}).  The second most probable
-topology (9.3\%) was one where SDD was the outgroup, and SPY and SDE were
-sister clades. SPY and SDE appeared to share about tenth of the core genomes that
-they were more closely related with each other than each to SDD's genome. The
-other topologies had less than 4\% of proportions.
-
-\subsection{Unbalaned gene flow from SPY to SDE in genome-wide scale}
-
-The heat map of Figure \ref{fig:Heatmap-of-recombination} (see Table
-S\ref{tab:heatmap} for the values) shows the log base 2 ratios of the \textit{a
-posteriori} number of recombinant edges for a pair of donor and recipient
-species tree branches and their corresponding expected values given the species
-tree and the three population parameters.  Because recombinant edges from any
-species tree branch to the root branch are not possible under
-\texttt{ClonalOrigin} model (see Figure S\ref{fig:clonalorigin}), the column for
-the root as recipient was colored green, indicating invalid recombinant edges.
-Similarly, all of the cells where a recipient branch was older than the donor,
-and two branches did not share any contemporary time were colored green. The
-diagonal cells were colored gray, indicating no significant number of
-recombinant edges, because the observed and expected numbers of recombinant
-edges were similar. As recombinant edges starting from and ending in the same
-species tree branch would not change the species tree, which would not change
-the likelihood much, the cells would be colored gray.  Because SDD is a strictly
-veterinary pathogenic bacteria, we did not expect that SDD received much gene
-transfer from either SPY1 or SPY2: the log base 2 ratio from SPY1 to SDD was
--2.20, and that from SPY2 to SDD was -1.97.  On the contrary, the recombing gene
-transfer from SDE1 to SDD was 0.756, and that from SDE2 to SDD was 0.698. This
-relatively higher recombining gene transfer from two SDE individuals to SDD
-compared with that from two SPY individuals to SDD reflected that the SDD was
-more closely related to SDE than to SPY in most of the genomes.
-
-The genomic organization between SDD and SDE1/SDE2 were much similar (Figure
-S\ref{fig:mauve}), and they formed a clade in the species tree
-(S\ref{fig:tree5}). However, the gene transfer from SDD to either SDE1 or SDE2
-was less frequent than the case of reverse direction; SDD served as a recipient
-rather than a donor. It is not unclear why SDD received more genes from SDE1 or
-SDE2 than did SDE1 or SDE2 from SDD.
+There are 105 possible rooted topologies relating five taxa.  We expected the
+dominant topology across sampled genealogies to be concordant with that of the
+species tree.  Indeed, 66.7\% of the local gene tree topologies (see section
+\textbf{Materials and Methods}) along the alignment blocks were the same as the
+species tree, which allowed us to be more confident of the species tree inferred
+using the alignment blocks (Table S\ref{tab:Gene-tree-topologies}).  Other
+frequently sampled topologies were expected to be potentially informative about
+common recombination events.  The second most frequent topology (9.3\%) was one
+where SDD was the outgroup, and SPY and SDE were sister clades. Therefore, SPY
+and SDE appeared be more closely related with each other than each to SDD's
+genome in about tenth of the core genome. The other topologies had less than 4\%
+of proportions.
+
+\subsection{Unbalaned recombining gene flow from SPY to SDE in genome-wide scale}
+
+The heat map of Figure \ref{fig:Heatmap-of-recombination} (see Tables
+S\ref{tab:heatmap} and S\ref{tab:obsheatmap} for the heat map values and the
+observed number of recombinant edges) shows the log base 2 ratios of the
+\textit{a posteriori} number of recombinant edges for a pair of donor and
+recipient species tree branches and their corresponding expected values given
+the species tree (Figure S\ref{fig:tree5}) and the three population parameters
+(Table S\ref{tab:three}).  Because recombinant edges from any species tree
+branch to the root branch are impossible (see Figure S\ref{fig:clonalorigin}),
+the column for the root as recipient was colored green, indicating invalid
+recombinant edges.  Similarly, all of the cells for which a recipient branch was
+older than the donor, and two branches did not share any contemporary time were
+colored green. The diagonal cells were colored gray, indicating no significant
+number of recombinant edges, because the observed and expected numbers of
+recombinant edges were similar. As recombinant edges starting from and ending in
+the same species tree branch would not change the species tree, which would not
+change the likelihood much, the cells would be colored gray.  Because SDD is a
+strictly veterinary pathogenic bacteria, we did not expect that SDD received
+much gene transfer from either SPY1 or SPY2: the log base 2 ratio from SPY1 to
+SDD was -2.20, and that from SPY2 to SDD was -1.97.  On the contrary, the
+recombing gene transfer from SDE1 to SDD was 0.756, and that from SDE2 to SDD
+was 0.698. This relatively higher recombining gene transfer from two SDE
+individuals to SDD compared with that from two SPY individuals to SDD reflected
+that the SDD was more closely related to SDE than to SPY in most of the genomes,
+and the genomic organization between SDD and SDE1/SDE2 were much similar
+(Figure S\ref{fig:mauve}).  However, the gene transfer from SDD to either SDE1
+or SDE2 was less frequent than the case of reverse direction; SDD effectively
+served as a recipient rather than a donor. It was not unclear why SDD received
+more genes from SDE1 or SDE2 than did SDE1 or SDE2 from SDD.
 
 Because it was expected that the recombining gene transfer within a clade would
-be larger than that between clades, the recombining gene transfer log base 2
-ratios between SDE1 and SDE2 (1.17 and 1.20) were about twice as large as that
-from SDE1 and SDE2 to SPY1 and SPY2 (0.67, 0.66, 0.60, and 0.39).  The
-intensities of recombining gene transfer from external branches of SDE to
-external branches to SPY was similar to those from SDE1 or SDE2 to SDD (0.76 and
-0.70).  Either SPY1 or SPY2, and SDD appeared to be no different from the view
-point of SDE1 or SDE2 since the similar values of recombining gene transfer
-intensity from SDE1 or SDE2 to the other three external branches.  Surprisingly,
-the recombining gene transfer log base 2 ratios between SPY1 and SPY2 (1.06 and
-1.07) was comparable with those from SPY1 and SPY2 to SDE1 and SDE2 (1.25, 1.25,
-1.17, and 1.29).  Effectively, the recombining gene transfer from a clade SPY1
-or SPY2 to a different clade SDE1 or SDE2 seemed to be no different from that
-within the clade of SPY1 and SPY2. The two external branches of SDE received
-genes from either SPY1 or SPY2 via recombining gene transfer as if SDE1 and SDE2
-were in the same clade as SPY1 and SPY2.  The bacterial recombination between
-SPY and SDE might happen more often than that between SDE and SDD even though
-the genome configurations of SDE and SDD are more similar than each to SPY. This
-might be consistent with the fact that SPY and SDE shared the same ecological
-niche.  The gene transfer log base 2 ratio from SPY1 and SPY2 to SDE branch was
-larger than any type of gene transfer: 1.88 and 1.94, respectively.  
-Furthermore, we found a total of 58 genes for which a recombinant edge of
-certain types were sampled with probability at least 0.9 by focusing on specific
-types of recombinant edges.  (Table S\ref{tab:genes-transfer}).  Many of the
-genes in the list were transferred from SPY to SDE rather than the reverse
-direction overall.  All things considered, this clearly showed that the
-recombining gene transfer was biased toward the net direction from SPY to SDE.
-
-Figure \ref{fig:hgt-heatmap} shows the number of inferred non-recombining
-horizontal gene transfer events between the different branches of the species
-tree.  SDE1 and SDE2 shared the most events (26 and 34) compared to other sample
-pairs.  SDD appeared to acquire many genes from the ancestral SDE branch.
-Transfer events between the SPY and SDE clades were inferred in smaller numbers
-(<6) compared to those within each clade (23--34). Among the clade-crossing
-transfers, we did not see any significant enrichment for a specific direction,
-as opposed to the result for recombining transfers from the model-based analysis
-(Figure S\ref{fig:mowgli-recomb-heatmap}).
+be more frequent than that between clades, not surprisingly the recombining gene
+transfer log base 2 ratios between SDE1 and SDE2 (1.17 and 1.20) were about
+twice as large as that from SDE1 and SDE2 to SPY1 and SPY2 (0.67, 0.66, 0.60,
+and 0.39).  The intensities of recombining gene transfer from external branches
+of SDE1 and SDE2 to external branches of SPY1 and SPY2 was similar to those from
+SDE1 or SDE2 to SDD (0.76 and 0.70).  After all, either SPY1 or SPY2, and SDD as
+recipient appeared to be no different from the view point of SDE1 or SDE2 as
+donor since the similar values of recombining gene transfer intensity from SDE1
+or SDE2 to the other three external branches.  Surprisingly, the recombining
+gene transfer log base 2 ratios between SPY1 and SPY2 (1.06 and 1.07) was
+comparable with, but even less than those from SPY1 and SPY2 to SDE1 and SDE2
+(1.25, 1.25, 1.17, and 1.29).  Effectively, the recombining gene transfer from a
+clade SPY1 or SPY2 to a different clade SDE1 or SDE2 seemed to be no different
+from that within the clade of SPY1 and SPY2.  The two external branches of SDE
+received genes from either SPY1 or SPY2 via recombining gene transfer as if SDE1
+and SDE2 were in the same clade as SPY1 and SPY2.  This was more surprising
+because the bacterial recombination between SPY and SDE seemed to happen more
+often than that between SDE and SDD even though the genome configurations of SDE
+and SDD were more similar than each to SPY.  However, this result could be
+concordant with the fact that SPY and SDE shared the same ecological niche
+despite their different genomic organization.  The gene transfer log base 2
+ratio from SPY1 and SPY2 to SDE branch was also larger than any type of
+recombining gene transfer: 1.88 and 1.94, respectively.  Additionally we found
+58 genes for which a recombinant edge of certain types were sampled with
+probability at least 0.9 by focusing on specific types of recombinant edges
+(Table S\ref{tab:genes-transfer}).  Many of the genes in the list were
+transferred from SPY to SDE rather than the reverse direction overall.  All
+things considered, this epitomized that the recombining gene transfer was biased
+toward the net direction from SPY to SDE.
+
+We tried in vain to quantify differential horizontal gene transfer between SPY
+and SDE with the parsimony-based approach (Figure \ref{fig:hgt-heatmap}).  Among
+the clade-crossing transfers, we did not see any significant enrichment for a
+specific direction, as opposed to the result for recombining transfers from the
+model-based analysis (Figure S\ref{fig:mowgli-recomb-heatmap}).  SDE1 and SDE2
+shared the most events (26 and 34) compared to other sample pairs.  SDD appeared
+to acquire many genes from the ancestral SDE branch.  Transfer events between
+the SPY and SDE clades were inferred in smaller numbers (<6) compared to those
+within each clade (23--34). We were puzzled by this finding because virulence
+genes were found to be associated with horizontal gene transfer, and we could
+not show net directionality of horizontal gene transfer between SPY and SDE.
 
@@
 
-Among the 2314 gene families the largest family consisted of 59 genes, while
-nearly half of gene families (1066) had exactly one gene per species, or six
-genes total (Figure S\ref{fig:famsize}).  Figure \ref{fig:Gene-duplication-loss}
-shows the inferred gene counts for each extant and ancestral species. Overall,
-we inferred 304 duplications, 944 recombining gene transfers, and 294
-non-recombining gene transfers, comprising 23.7\% of total transfer events.  We
-found that duplications made up only a small fraction of the events within the
-clade (10.2\%), while horizontal transfers and losses were significantly more
-frequent (41.7\% and 48.0\%, respectively). We found that duplications were most
-often species-specific. Many gene losses likely represented horizontal transfers
-that replaced the native gene sequence.  The species SDD and SDE were the most
-expanded, both in terms of duplications and in-bound transfers.
+Overall, we inferred 304 duplications, 944 recombining gene transfers, and 294
+horizontal gene transfers, comprising 23.7\% of total transfer events (Figure
+\ref{fig:Gene-duplication-loss}).  Among the 2314 gene families the largest
+family consisted of 59 genes, while nearly half of gene families (1066) had
+exactly one gene per species, or six genes total (Figure S\ref{fig:famsizes}).
+Within clades we found that duplications made up only a small fraction of the
+events  (10.2\%), while horizontal transfers and losses were significantly more
+frequent (41.7\% and 48.0\%, respectively). Duplications were most often
+species-specific.  The species SDD and SDE were the most expanded, both in terms
+of duplications and in-bound transfers.
 
 \subsection{Virulence genes via recombining gene transfer}
-
-\texttt{Find cases of any virulence genes that were transferred via horizontally.} 
-\texttt{Make a list of virulence genes and their categories with more interpretations.}
 
@@
 browser tracks we located virulence genes with notable posterior recombination
-probability.  Stressful host conditions include heat shock, starvation, and iron
-depletion.  Iron uptake of bacteria was known to be essential for survival of
-the bacteria under hostile host condition. We found that parts of the gene,
-putative calcium transporter (MgtA), in SDE1 had been transferred from SPY1
-(SpyM3\_0440).  An extracellular protein, putative dipeptidase (SpyM3\_0465),
-was found to be transferred from SPY to SDE. The gene as a whole was
-transferred, which was a case for the finding that virulence genes tend to be
-transferred as a whole unit \citep{Chan2009}. Dipeptidase may be important for
-the survival of bacteria to scavenge dipeptides from the host.  An adhesins
-gene, putative internalin A precursor (SpyM3\_1035), was found to be affected by
-recombining gene transfer from SPY to SDE. Some of the middle parts of the gene
-was affected.  A middle part of a chaperonin GroEL, putative heat shock protein
-(SpyM3\_1765), was inferred to transfer from SPY to SDE1.  Increase of GroEL
-expression was known to allow for the bacteria to survive the hostile host
-environment of acid shock.  Superoxide dismutase was reported to be associated
-virulence of \textit{Streptococcus agalactiae} \citep{Poyart2001}. This whole
-gene including its 5' UTR, (SpyM3\_1071), was affected by recombining gene
-transfer from SPY to SDE.  Pyruvate formate lyase was recently reported to have
-virulence \citep{Yesilkaya2009}. A few parts of the gene (SpyM3\_1749) seemed to
-be affected by recombining gene transfer.  One MLST locus, called DNA mismatch
-repair protein MutS, appeared to be affected by recombining gene transfer.  A
-putative spermidine/putrescine ABC transporter ATP-binding protein,
-(SpyM3\_0764), was found to be transferred from SPY to SDE.  Under starvation
-condition bacteria must use more effective energy metabolism.  A putative
-glucose kinase, (SpyM3\_1180), might be transferred from SPY to SDE to adapt
-SDE's energy metabolism to survive such harsh condition.
-
-\subsection{Horizontal gene transfer is associated with virulence genes}
-
-\texttt{Select genes with high horizontal gene transfers. Relate these to
-categories of virulence genes. I should find many more virulence genes in Matt's
-analysis. Use the virulence gene types from SDE2's paper. Which virulence gene
-types have more gene transfer?}
-
-We found that some virulence genes tended to have more horizontally transfers
-than others. 
-Streptococcal C5a peptidase (\textit{scpB}) (SDEG\_0933) was inferred to have six
-horizontal gene transfer. 
-Putative transposase were also found to have relatively large number of
-horizontal gene transfer events.
-(SpyM3\_1016, SpyM3\_1053, SpyM3\_1054)
-
-Putative response regulator (SpyM3\_0174).
-Putative maltose operon transcriptional repressor (SpyM3\_0982).
-Putative histidine kinase (SpyM3\_0171).
-Putative cell wall hydrolase (SpyM3\_0731, SpyM3\_0922, SpyM3\_1208).
-
-GGS\_124 possesses 5 genes that encode putative nucleases with a secretion
-signal peptide: genome-encoded streptodornase (SDEG\_0541), others,
-prophage-associated deoxyribonuclease (SDEG\_1103), all of which are predicted to code for a secretion signal peptide.
-Streptodornase (Sdn) - phage associated  (SpyM3\_1409).
-
-Protein F2 like fibronectin-binding protein (SpyM3\_0104).
+probability.  These genes were by no means comprehensive list of virulence
+factors that might have been transferred via homologous recombination.
+Stressful host conditions for pathogeneic bacteria can include heat shock,
+starvation, and iron depletion.  Iron uptake of bacteria was known to be
+essential for survival of the bacteria under hostile host condition. We found
+that parts of putative calcium transporter (MgtA) in SDE1 could have been
+transferred from SPY1 (SpyM3\_0440).  An extracellular protein, putative
+dipeptidase (SpyM3\_0465) that could be important for the survival of bacteria
+to scavenge dipeptides from the host, was inferred to be transferred from SPY to
+SDE: the whole gene was transferred, which was a case for the finding that
+virulence genes tended to be transferred as a whole unit \citep{Chan2009}.
+Another example of whole gene transfer was superoxide dismutase that was
+reported to be associated virulence of \textit{Streptococcus agalactiae}
+\citep{Poyart2001}: the whole gene including its 5' UTR, (SpyM3\_1071), was
+affected by recombining gene transfer from SPY to SDE.  Some of the genes were
+only affected in their parts.  The middle parts of an adhesins gene, putative
+internalin A precursor (SpyM3\_1035), was found to be affected by recombining
+gene transfer from SPY to SDE.  The middle part of a chaperonin GroEL, putative
+heat shock protein (SpyM3\_1765), was inferred to transfer from SPY to SDE1:
+increase of GroEL expression was known to allow for the bacteria to survive the
+hostile host environment of acid shock.  Parts of the gene (SpyM3\_1749),
+pyruvate formate lyase that was recently reported to have virulence
+\citep{Yesilkaya2009}, seemed to be affected by recombining gene transfer.
+Under starvation condition bacteria must use more effective energy metabolism: a
+putative glucose kinase, (SpyM3\_1180), were inferred to be transferred from SPY
+to SDE to adapt SDE's energy metabolism to survive such harsh condition.  We
+also found that an MLST locus, called DNA mismatch repair protein (MutS),
+appeared to be affected by recombining gene transfer.
+
+\subsection{Virulence genes via horizontal gene transfer}
+
+Horizontal gene transfers were found to be significantly associated with
+virulence genes (P-value of $<1.33 \times 10^{-11}$) whereas recombining gene
+transfers were not enriched for virulence genes (P-value of $>0.90$) using the
+parsimony-based approach.  This was also consistent with the fact that we could
+not find the association of recombining gene transfer with virulence genes in
+the model-based approach.  This supported more evidence that non-recombining
+gene transfer served as a more effective means by which SDE or SPY acquired
+virulence genes, compared to recombining gene transfer.  
+Families with high numbers of duplications were also significantly enriched for
+virulence genes (P-value of $<7.13 \times 10^{-3}$) whereas families with high
+losses showed no enrichment (P-value of $>0.54$).
+
+Roughly, half of the virulence genes were in the pan genome.  We did not attempt
+to comprehensively list virulence genes that were inferred to be transferred
+horizontally. We rather sort virulence genes by the number of horizontal gene
+transfer events to list the first several genes with high ranks.  Streptococcal
+C5a peptidase (\textit{scpB}, SDEG\_0933) was inferred to have six horizontal
+gene transfer.  Putative transposase were found to have a relatively large
+number of horizontal gene transfer events (SpyM3\_1016, SpyM3\_1053,
+SpyM3\_1054). Other genes with relatively large number of horizontal gene
+transfer events included putative response regulator (SpyM3\_0174), putative
+maltose operon transcriptional repressor (SpyM3\_0982), putative histidine
+kinase (SpyM3\_0171), and putative cell wall hydrolase (SpyM3\_0731,
+SpyM3\_0922, SpyM3\_1208).  Virulence genes encoding putative nucleases with a
+secretion signal peptide were also found to have had experienced horizontal gene
+transfer: genome-encoded streptodornase (SDEG\_0541), and prophage-associated
+deoxyribonuclease (SDEG\_1103).  Protein F2 like fibronectin-binding protein
+(SpyM3\_0104), an adhesin and invasin, was inferred to horizontally
+transferred.
 
@@
 
-\texttt{I might want to analyze only core genomes and pan genomes
-separately. By separating data set of 6-gene families from the other families,
-we could have more complete comparison.}
-
-\texttt{Genes and recombining gene transfer events count and recombination
-intensity might be use to compare two methods.}
-
-\subsection{UCSC genome browser for the dataset}
+Both of the model-based and parimony-based methods used discrepancy between a
+reference species tree and gene trees to detect recombining gene transfer. In
+essence the two methods worked in relatively similar manner. Firstly, functional
+categories that were associated with recombining gene transfer were similar in
+both methods (Tables S\ref{tab:functional-all} and
+S\label{tab:go-events-recombining}).  Secondly, both of them did not show
+association of virulence genes with recombining gene transfer. Lastly, we found
+a positive correlation of 0.39 with P-value of $2.2\times10^{-16}$ between the
+recombination intensity and the number of recombining gene transfer for the
+genes (Figure S\ref{fig:cmpcomowgli}).  The correlation (0.03 with P-value of
+0.18) disappeared in the comparison between recombination intensity and number
+of horizontal gene transfer events (Figure S\ref{fig:cmpcomowglihgt}).
+
+\subsection{UCSC genome browser for bacterial gene transfer study}
+
 We created a web-resource for gene transfer available at
@@
 probability of recombining gene transfer.  
-
-\subsection{Transport-related genes are associated with recombining gene transfer}
-
-\subsection{Transporter and aminoacyl tRNA synthetases}
-Lateral transfer of transporter related genes are well known.  Functional
-categories associated with higher recombining gene transfer were found to
-include phosphotransferase systems and ATP-binding cassette (ABC) transporters,
-i.e., carbohydrate transport (GO:0008643) (Table \ref{tab:functional}).  We also
-found several aminoacyl tRNA synthetases in gene ontology of translation
-(GO:0006412) to be significantly associated with higher recombining gene
-transfer.  Note that functional categories concerning ribosomes were also
-associated with higher recombining gene transfer. These categories were not
-found significantly associated with higher recombining gene transfer when we
-considered only gene transfer between SPY and SDE genomes, which implied that
-many recombining gene transfer events should be attributed to those of internal
-species branches as donors.  Interestingly, transports functional categories
-were found to be associated with recombining gene transfer for the direction
-from SPY to SDE.  The parsimony-based approach also  shows enrichment of the
-recombining transfers in many of the same categories such as translation,
-ribosome, and rRNA binding (Table \ref{tab:go-events}).  This agreement between
-the two different methods, applied on somewhat different parts of the data,
-serves as validation for the signal picked up by these separate association
-studies.  
-\texttt{Comparison of Mawgli and CO?}
-Recombining gene transfers were not enriched for virulence genes (P-value of
-$>0.90$) while horizontal gene transfers were found to be significantly
-associated with virulence genes (P-value of $<1.33 \times 10^{-11}$). This was
-also consistent with the fact that we could not find the association of
-recombining gene transfer with virulence genes in the model-based approach.
-This supported more evidence that non-recombining gene transfer served as a more
-effective means by which SDE or SPY acquired virulence genes, compared to
-recombining gene transfer.  For horizontal gene transfer we found enrichment in
-very different categories related to transposition, integration, and
-recombination.  
-\texttt{More interpretations about Mawgli analsis of function categories?}
-
-Gene duplication and loss events were enriched for categories similar to those
-of horizontal gene transfer's enrichment.  Interestingly, gene duplication and
-horizontal transfers showed enrichment in the same categories, perhaps because
-both mechanisms increase gene copy number.  Gene losses showed enrichment
-for transposition and integration, but less significantly (P-value of
-$<10^{-6}$), and showed enrichment for several additional categories such as
-transporter activity and transcription initiation.  Families with high numbers
-of duplications were significantly enriched for virulence genes (P-value of
-$<7.13 \times 10^{-3}$), whereas families with high losses showed no enrichment
-(P-value of $>0.54$).
-
+% To Matt: what could we show in the pan-genome? or from Mowgli's analysis?
+
+\subsection{Units of recombining gene transfer}
+
+The process of gene transfer may or may not move whole genes from one species to
+another \citep[e.g.,][]{Chan2009}. We observed that units of recombining gene
+transfer could happen with genes (Table S\ref{tab:genes-transfer}). In some
+cases, parts of a gene were inferred to have been transferred, for instance,
+transfer from SPY1 to SDE1 was inferred for 29\% of the aspartyl-tRNA synthetase
+gene.  In other cases, the inferred transfer units covered a set of as much as
+eight adjacent genes.  Of the genes listed in Table S\ref{tab:genes-transfer}
+there were 12 groups of genes that recombinant tracts covered their gene
+boudaries.  Most of the recombinant tracts often covered two genes. There were
+transfer fragments that covered eight genes together: loci from SpyM3\_1518 to
+SpyM3\_1525.
+
+\subsection{Transporter, transferase, and aminoacyl tRNA synthetases}
+
+Two classes of functional categories notably were associated with recombining
+gene transfer.  Firstly, genes involved in sensing queues of environmental
+changes would be important for bacteria to adapt to changing environment.  Genes
+encoding membrane-bound proteins would be involved in the function.  Several of
+functional categories that were associated with gene transfer might be related
+with membrane-bound protein coding genes (Table \ref{tab:functional}).  Sugar
+phosphotransferase classes were ranked top in the list: phosphrylation must be
+important in signal transduction to recognize environmental changes.  The
+transport functional categories of carbohydrate transport and sodium ion
+transport also belonged to genes encoding membrane-bound proteins.  Also highly
+ranked were fatty acid or lipid biosyntheses that must be involved in producing
+membrane-bound molecules. These functional categories related with
+membrane-bound proteins might allow for bacteria to adapt to changing
+environment via recombining gene transfer.
+
+Secondly, aminoacyl tRNA synthetases had been a focus of gene transfer
+\citep{Woese2000}, and these aminoacyl tRNA synthetases belonged to the
+functional category of translation (GO:0006412).  They were inferred to be 
+associated with recombining gene transfer not horizontal gene transfer because
+translation category was siginificant only in recombining gene transfer (see
+Tables \ref{tab:functional}, \label{tab:go-events} and S\label{tab:go-events-recombining}).
+
+In addition to the two kinds of functional categories, functional categories
+concerning ribosomes were also associated with higher recombining gene transfer.
+Because genes related with ribosomes were considered to be housekeeping genes,
+gene transfer in ribosome genes would be rare, it was unexpected that ribosome
+related genes were found to be associated with recombining gene transfer. This
+was more significant when we considered all of the recombinant edges (Table
+S\ref{tab:functional-all}) not just topology-changing recombinant edges (Table
+\ref{tab:functional}) in computing recombination intensity.  Therefore, high
+recombination intensity in ribosome related genes would likely to be attributed
+to topology-not-changing recombinant edges (Table
+S\ref{tab:functional-notopology}).  The parsimony-based approach also showed
+enrichment of the recombining transfers in many of the same categories such as
+translation, ribosome, and rRNA binding when using all types of recombinant
+edges in computing recombination intensity (see Tables
+S\ref{tab:go-events-recombining} and S\ref{tab:functional-all}).  Because the
+parsimony-based method did not take species tree branch lengths into account, we
+were likely to the agreement between the two different methods.
+
+For horizontal gene transfer we found enrichment in very different categories
+related to transposition, integration, and recombination (Table
+\ref{tab:go-events}).  Gene duplication and loss events were enriched for
+categories similar to those of horizontal gene transfer's enrichment (Tables
+S\ref{tab:go-events-duplication} and S\ref{tab:go-events-loss}).
+Gene duplication and horizontal transfers showed enrichment in the same
+categories perhaps because both mechanisms increase gene copy number.  
+Gene losses showed enrichment for transposition and integration, but less
+significantly (P-value of $<10^{-6}$), and showed enrichment for several
+additional categories such as transporter activity and transcription initiation.
+
+Comparing two types of recombinant edges, one from SDE to SPY and another from
+SPY to SDE transport related categories were more associated for the direction
+from SPY to SDE
+(Table S\ref{tab:functional-sde-spy} and S\ref{tab:functional-spy-sde}).
+
+% More interpretations about Mawgli analysis of function categories?
+% Why transposases?
 % \texttt{IG: NEXT PARAGRAPH CONTAINS SOME DISCUSSION ON THESE RESULTS. WE SHOULD 
@@
 
-\subsection{Units of recombining gene transfer}
-The process of gene transfer does not necessarily move whole genes from one
-species to another \citep[e.g.,][]{Chan2009}, and we observed that units of
-recombining gene transfer do not correspond to single genes in their entirety
-(Table S\ref{tab:genes-transfer}). In some cases, only parts of a gene were
-inferred to have been transferred, for instance, transfer from SPY1 to SDE1 was
-inferred for 29\% of the aspartyl-tRNA synthetase gene (see also Figure
-S\ref{fig:muts}). In other cases, the inferred transfer units cover sets of as
-much as eight adjacent genes.  Not only parts of genes could be transferred via
-recombination, but a group of genes could also be moved together. Of the genes
-listed in Table S\ref{tab:genes-transfer} there were 12 groups of genes that
-were consecutive.  While transfer fragments of genes often covered two genes,
-there were transfer fragments that covered eight genes together. Loci from
-SpyM3\_1512 to SpyM3\_1525 were inferred to be transferred from SDE2 to SDD.
-
 \section{Discussions}
 
-Closely related bacterial species were investigated here to better understand
-gene transfer processes. Two complementary methods were employed to study
-recombining and non-recombining gene transfers. By discriminating the two
-processes in analyzing the genomes, we provided evidence that the pathogenicity
-of SDE and SPY was more attributed to non-recombining gene transfer process than
-homologous recombining gene transfer.  Direction of gene flow between SDE and
-SPY has been questioned and challenged
-\citep{Kalia2001,Kalia2004,Towers2004,Bessen2005,Davies2007,Bessen2010} in order
-to elucidate the origin of opportunistic human pathogenicity of SDE via transfer
-from SPY into SDE.  We also showed in this study that the amount of recombining
-gene flows from SPY to SDE was relatively greater than that of the reverse
-direction using the core genome sequences.
-
-\textbf{Virulence genes and gene transfer:} 
-Because we found directional recombining gene transfer from SPY to SDE, and did
-not found virulence genes association with recombining gene transfer, this
-indicated that pathogenicity of SDE might not be due to acquisition of virulence
-genes from SPY via recombining transfer. This does not necessarily mean that
-virulence genes were not exchanged between SPY and SDE via non-recombining
-transfer.  Because the \texttt{Mowgli} analysis just did not reveal unbalanced
-non-recombining or recombinding gene transfer, we suspected that \texttt{Mowgli}'s
-parsimony might not be powerful in detecting unbalanced gene flow.  We
-conjecture the following hypothesis of the evolution of SPY and SDE to support
-the non-recombining virulence gene exchange between SPY and SDE.  Pathogenicity
-of SDE might have existed before the common ancestor of SPY and SDE.  We
-speculated that SPY might have been split off from the common ancestor that was
-SDE-like because the population size of SDE was larger than that of SPY (CITE).
-After the split, SPY might have been specialized in inhabting human hosts, and
-SDE in colonizing veterinary hosts. Recent surge in cases of human diseases
-attributed to SDE might be because of secondary contact of SPY and SDE since
-their split from their common ancestor.  When SPY thrives in its environment of
-human hosts and another relatively close species, SDE, starts to cohabit with
-the indigenous species SPY, genetic elements in one of the two species could
-benefit another species. If the two species do not compete with each other, then
-the increased genetic variation from multiple species would eventually bring
-advantage to both of the species.  Virulence genes could have been beneficial to
-both species in inhabting human hosts.  Because the common ancestry of SPY and
-SDE, recombining gene transfer between the two species might be relatively
-prevalent. But, these homologous recombination might have rarely contributed to
-influx of virulence genes from SPY to the opportunistic human pathogen of SDE.
-Because homologous parts of SPY and SDE would be similar in their genetic
-compositions, the homologous parts would be unlikely to contain virulence genes
-to human hosts. We suspected that virulence genes were exchanged between SPY and
-SDE, but did not find any directionality of the gene flow.
-
-\textbf{Revisit to previous studies:} 
-Reflection of previous studies in gene transfer in SPY and SDE would suggest
-that using relatively small number of genomes we could detect homologous
-recombining gene transfer that had been possible using many samples for a few
-genes. We revisited the problem of gene transfer in other previous studies.  We
-examined the seven genes that \citet{Kalia2001} had studied for directionality
-of gene transfer between SPY and SDE. Internal fragments of the seven
-housekeeping genes, encoding putative glucose kinase (\textit{gki}), glutamine
-transport protein (\textit{gtr}), glutamate racemase (\textit{murI}), mismatch
-repair enzyme (\textit{mutS}), transketolase (\textit{recP}), xanthine
-phosphoribosyltransferase (\textit{xpt}), and acetylcoenzyme A acetyltransferase
-(\textit{yqiL}) were located in the SPY1 genome that we investigated in this
-study.  Six of the seven genes except \textit{yqiL }gene were in the core
-genomes \texttt{ClonalOrigin} analyzed. Genes \textit{mutS}, \textit{recP}, and
+We investigated closely related bacterial species to better understand gene
+transfer processes between SPY and SDE. Two complementary methods were employed
+to study recombining and horizontal gene transfers. By discriminating the two
+processes in analyzing the genomes, we showed that the pathogenicity of SDE and
+SPY was more attributed to horizontal gene transfer process than recombining
+gene transfer.  Because recombination in bacteria seemed to occur at somewhat
+random positions along the genome, the homologous parts transferred might be
+unlikely to be enriched with virulence factors to human hosts.  Recombining gene
+transfer might have rarely contributed to influx of virulence genes from SPY to
+the opportunistic human pathogen of SDE as we showed a few cases of virulence
+gene that might have been transferred via recombining gene transfer.
+
+We also showed that the net amount of recombining gene flows from SPY to SDE was
+relatively greater than that of the reverse direction in the core genome
+sequences.  If there were unbalanced recombining gene transfer between the two
+species, what were the barriers \citep{Thomas2005} to the gene transfer in the
+bacteria? Although barriers to bacterial recombination would include geographic
+isolation, genetic isolation could prevent recombination between coinhabiting
+bacteria.  \citet{Kalia2001} discussed restriction-modification system (RMS) as
+a possible mechanism of the unbalanced gene flow.  Restriction-modification
+system is a biological self-defense mechanism of bacteria, which is diverse in
+prokaryote world.  Biological roles of RMS might be to maintain and control
+species identity \citep{Jeltsch2003}.  Recently, \citet{Budroni2011a} studied
+that \textit{Neisseria meningitidis} phylogenetic clades were associated with
+RMS, claiming that RMS modulated homologous recombination in the bacteria of
+\textit{N.\ meningitidis}.  Most of \textit{S.\ pyogenes} genomes contain
+restriction-modification type I system as well as type II system. The two
+genomes of SDE contained only type II system. More tight self-defense system of
+SPY may have made the species less capable of aquiring foreign genetic material
+from SDE, and accordingly relatively unbalanced gene flow between SPY and SDE
+may have been associated with different restriction-modification systems in the
+two species \citep{Kalia2001}.
+
+We used relatively small number of genomes to detect recombining gene transfer
+that had only been possible using many samples for a few genes. We revisited the
+problem of gene transfer in other previous studies.  We reexamined the seven
+genes that \citet{Kalia2001} had studied for net directionality of gene transfer
+between SPY and SDE.  Internal fragments of the six housekeeping genes, encoding
+putative glucose kinase (\textit{gki}), glutamine transport protein
+(\textit{gtr}), glutamate racemase (\textit{murI}), mismatch repair enzyme
+(\textit{mutS}), transketolase (\textit{recP}), and xanthine
+phosphoribosyltransferase (\textit{xpt}) were located in the core genome that we
+investigated in this study.  Genes \textit{mutS}, \textit{recP}, and
 \textit{xpt} had fragments of gene transfer with moderate posterior probability
@@
 studied fluoroquinolone resistance in SDE and reported recombination in the
-gene. We also observed moderate recombination in the gene.
-
-\textbf{Mechanism of unbalanced gene flow between SPY and SDE:} 
-If there were unbalanced gene transfer between the two species, what were the
-barriers \citep{Thomas2005} to gene transfer in the bacteria? Although barriers
-to bacterial recombination would include geographic isolation, genetic isolation
-could prevent recombination between coinhabiting bacteria. \citet{Kalia2001}
-discussed possible mechanisms of the unbalanced gene flow by mentioning
-potential roles of restriction-modification system (RMS) in the unbalanced gene
-flow.  Restriction-modification system is a biological self-defense mechanism of
-bacteria, which is diverse in prokaryote world. Biological roles of RMS might be
-to maintain and control species identity \citep{Jeltsch2003}.  Recently,
-\citet{Budroni2011a} studied that \textit{Neisseria meningitidis} phylogenetic
-clades were associated with RMS, claiming that RMS modulated homologous
-recombination in the bacteria of \textit{N.\ meningitidis}.  Most of \textit{S.\
-pyogenes} genomes contain restriction-modification type I system as well as type
-II system. The two genomes of SDE contained only type II system. We conjecture
-that more tight self-defense system of SPY may have made the species less
-capable of aquiring foreign genetic material from SDE, and accordingly
-relatively unbalanced gene flow between SPY and SDE may have been associated
-with different restriction-modification systems in the two species.
-
-\textbf{Two functional categories and recombining gene transfer:}
-Among the genes shown in Table S\ref{tab:functional} two classes of proteins
-attracted our attentions: aminoacyl tRNA synthetases (aaRS) and
-phosphotransferase system (PTS).  Aminoacyl tRNA synthetases charge tRNA with an
-amino acid so that amino acids are concatenated from charged tRNA onto a growing
-peptides through translation. The aaRS's are characterized by an amino acid that
-they charge tRNA with.  This gene family had been a target for horizontal gene
-transfer \citep[e.g.,][]{Lamour1994,Woese2000,Koonin2001}, which lent itself to
-an example of gene transfer in bacteria.  There were also recent reports of gene
-transfer in tRNA aminoacylation \citep{Andam2010,Wang2011}.  In our analysis,
-some of them appeared to be transferred from SPY to SDE: leucyl-tRNA synthetase,
-aspartyl-tRNA synthetase, and arginyl-tRNA synthetase. Half of phenylalanyl-tRNA
-synthetase gene appeared to be transferred in the reverse direction.
-
-Bacterial phosphotransferase system (PTS), also known as phosphoenolpyruvate
-(PEP) group translocation, transports sugars such as glucose into the cell where
-phosphorylation of the sugar happens. There have been studies of gene transfer
-of PTS in bacteria \citep[e.g.,][]{Barabote2005,Zuniga2005}. As the main source
-of energy for bacteria, sugars in the bacterial niches must have been acting as
-important selective pressure. It is not surprising that PTS sugars differ among
-bacterial groups because bacterial species are correlated with their environment
-and energy sources theirof.  Table S\ref{tab:functional} shows that two PTS
-genes appeared to be transferred from SPY to SDE: {}``PTS system,
-mannose-specific IIC component'' and {}``PTS system, fructose-specific IIABC
-component.'' Whereas almost half of the former gene was inferred to be affected
-by recombinant edges, only 3\% of the latter gene appeared to be affected.
-
-\textbf{Future models of bacterial evolution to handle recombining and
-non-recombining gene transfer:}
-The method of detecting homologous recombination gene transfer, or the
-\texttt{ClonalOrigin}, is model-based approach as opposed to the other method
-for non-recombining gene transfer, or the \texttt{Mowgli}.  The core genomes consisted of
-relatively long (i.e., 1500 base pairs long) sequence alignments with almost no
-gaps. With this inherent filtering of the input data preparation
-\texttt{ClonalOrigin} did not account for gene gain and loss in its core genome
-analysis. On the contrary, \texttt{Mowgli} used the pan genomes as well as the core
-genomes.  Considering nucleotides as the level of gene duplication, loss, and
-transfer towards the efforts of having more complete models for recombining and
-non-recombining gene transfer could be difficult because genome-scale
-application of those models based on DNA or protein sequences might not be
-feasible.  The research into units of gene transfer is underway
-\citep[e.g.,][]{Chan2009a}. Discriminating the mode of gene transfer and using
-the units of each mode of gene transfer towards the efforts of developing
-model-based approaches to detecting gene transfer might be plausible. In doing
-so, more complicated evolutionary models could be considered: genome
-rearrangement, gene segment gain and loss, and gene duplications.
+gene. We observed moderate recombination in the gene.
+
+We used two heterogeneous methods: one model-based and another parsimony-based
+for studying recombining and horizontal gene transfer.  The core genomes
+consisted of sequence alignments with rare gaps. With this inherent filtering of
+the input data preparation the model-based method did not accommodate gene
+dupliation and loss in its core genome analysis. On the contrary, the
+parsimony-based method used all of the genomes not just the core genomes.
+Considering nucleotides as the level of gene duplication, loss, and transfer
+towards a complete model for recombining and horizontal gene transfer could be
+difficult because genome-scale application of those models based on DNA or
+protein sequences might not be feasible.  Combining the two approaches for
+studying the mode of gene transfer would be desirable to study more complicated
+evolutionary processes including recombining gene transfer, gene duplication,
+loss, and horizontal gene transfer.
 
@@
 
-%
 \begin{table}
-\caption{\label{tab:functional}Functional categories asscoiated with higher
-recombination - Total number of recombinant edge types.}
-\begin{tabular}{ccc}
-\hline 
-p-value & count & description\tabularnewline
-\hline 
-2.994e-07 &  28 & rRNA binding\tabularnewline
-1.447e-06 &  42 & structural constituent of ribosome\tabularnewline
-1.694e-05 &  39 & ribonucleoprotein complex\tabularnewline
-2.048e-05 &  23 & fatty acid biosynthetic process\tabularnewline
-3.664e-05 &  72 & ribosome\tabularnewline
-1.018e-04 &  86 & translation\tabularnewline
-1.733e-04 &  22 & lipid biosynthetic process\tabularnewline
-2.792e-04 & 108 & RNA binding\tabularnewline
-4.164e-04 &  47 & carbohydrate transport\tabularnewline
-1.303e-03 &  38 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
-1.345e-03 &  10 & pentose-phosphate shunt\tabularnewline
-1.819e-03 &  35 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
+\caption{\label{tab:functional}Functional categories asscoiated with 
+recombination intensity based on topology changing recombinant edge types.}
+\begin{tabular}{cccc}
+\hline 
+P-Value & Count & GO term & Description\tabularnewline
+\hline 
+1.731e-06 & 38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
+2.775e-06 & 33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
+4.581e-05 & 22 & GO:0006633 & fatty acid biosynthetic process\tabularnewline
+0.0001248 & 20 & GO:0008610 & lipid biosynthetic process\tabularnewline
+0.0002675 & 45 & GO:0008643 & carbohydrate transport\tabularnewline
+0.000513 & 81 & GO:0006412 & translation\tabularnewline
+0.0005617 & 13 & GO:0006814 & sodium ion transport\tabularnewline
+0.001019 & 40 & GO:0030529 & ribonucleoprotein complex\tabularnewline
+0.001203 & 12 & GO:0031402 & sodium ion binding\tabularnewline
+0.00206 & 39 & GO:0003735 & structural constituent of ribosome\tabularnewline
+0.002115 & 237 & GO:0008152 & metabolic process\tabularnewline
+0.00215 & 26 & GO:0019843 & rRNA binding\tabularnewline
+0.00287 & 371 & GO:0016740 & transferase activity\tabularnewline
+0.003552 & 66 & GO:0005840 & ribosome\tabularnewline
 \hline 
@@
 \begin{table}
-\caption{\label{tab:go-events}GO-term enrichment for different
-evolutionary events - Non-recombining transfers.}
+\caption{\label{tab:go-events}Functional categories asscoiated with 
+non-recombining horizontal gene transfer.}
 \begin{tabular}{cccc}
 \hline 
-P-value & Count & GO-term & Description \\
+P-Value & Count & GO term & Description \\
 \hline 
@@
 %=============================================================================
-% Figure 1. Heat maps of recombining gene transfer (Add horizontal gene transfer)
+% Figure 1. Heat map of recombining gene transfer from ClonalOrigin
 \begin{figure}
 \includegraphics[scale=0.45]{figures/heatmap-recedge}
-\caption{\label{fig:Heatmap-of-recombination}Heatmap of recombination events.
-Heat map representing the logarithm base 2 of the number of recombination
-events inferred relative to its expectation under the prior, for each
-donor/recipient pair of branches. The donor is given by the y-axis;
-the recipient on the x-axis.  The
-green cells were the ones for which the ratios were meaningless because
-the recipient species tree branch of the cells was older than the
-donor species branch.}
+\caption{\label{fig:Heatmap-of-recombination}Heatmap of recombination events
+using the five genomes.  Each cell of the heat map represents the logarithm base
+2 of the number of recombination events inferred relative to its expectation
+under the prior, for a donor/recipient pair of branches. The donor is given by
+the y-axis; the recipient on the x-axis.  The green cells are meaningless
+because the recipient species tree branch of the cells are older than the donor
+species branch.}
 \end{figure}
-\clearpage{}%
+\clearpage{}
 %=============================================================================
@@
 \end{center}
-\caption{Distribution of horizontal gene transfers.}
-\label{fig:hgt-heatmap}
+\caption{\label{fig:hgt-heatmap}Distribution of horizontal gene transfers.
+Each cell of the color matrix shows the number of inferred non-recombining
+horizontal gene transfer events between the different branches of the species
+tree.}
 \end{figure}
@@
 
-% Figure 4. Gene duplication, loss, and transfer
+%=============================================================================
+% Figure 3. Gene duplication, loss, and transfer
 \begin{figure}
 \includegraphics[width=6in]{figures/strep-events}
-\caption{\label{fig:Gene-duplication-loss} Distribution of inferred
-  gene duplication, loss, and transfer events across the phylogeny.
-  Gene counts for each extant and ancestral species are given at each
-  node of the phylogeny.  For each branch we denote the number of gains
-  G*, duplications D*, losses L*, non-recombining transfers T*, and 
-  recombinations R*.  The non-recombining transfer and recombination counts
-  are indicated on the recipient branch (donors are not indicated).}
+\caption{\label{fig:Gene-duplication-loss} Distribution of inferred gene
+duplication, loss, and transfer events across the phylogeny.  Gene counts for
+each extant and ancestral species are given at each node of the phylogeny.  For
+each branch we denote the number of gains G*, duplications D*, losses L*,
+non-recombining transfers T*, and recombinations R*.  The non-recombining
+transfer and recombination counts are indicated on the recipient branch (donors
+are not indicated).}
 \end{figure}
 \clearpage{}%
+%=============================================================================
 
@@
 %=============================================================================
-% Table S1. List of genomes
+% Table S1. List of six genomes.
 \begin{table}
-\caption{\label{tab:genome}The set of complete bacterial genomes}
+\caption{\label{tab:genome}The six complete bacterial genomes. SDD genome was
+not yet available from NCBI.}
 \noindent \begin{centering}
@@
 \begin{table}
-\caption{\label{tab:clonalframe}Species tree inference using ClonalFrame.}
+\caption{\label{tab:clonalframe}Parameter estimates from ClonalFrame. 
+$\theta$ is population mutation rate,
+$\delta$ is the average recombinant tract length,
+$r/m$ is the relative strength of recombination compared to mutations,
+TMRCA is the time to most recent common ancestor, and
+$T$ is the total branch length of the inferred species tree.}
 \noindent \centering{}\begin{tabular}{cc}
-Parameter & Estimates\tabularnewline
+\hline
+Parameter & Estimates (95\% credibility interval) \tabularnewline
 \hline
@@
 %=============================================================================
-% Table S3. ClonalOrigin's 1st stage result.
+% Table S3. Gene Tree topologies
 \begin{table}
-\caption{\label{tab:three}Three population parameter estimates from
-ClonalOrigin. 
-The estimates of the three parameters were based on one of two independent
-MCMC chains.  We considered the similar 
-results of the two independent chains as confirmation of 
-the convergence of the first stage of \texttt{ClonalOrigin} MCMC.
-}
-\noindent \centering{}\begin{tabular}{ccc}
-& Replicate 1 & Replicate 2\tabularnewline
-Parameter & Estimate (IQR$^a$) & Estimate (IQR)\tabularnewline
-\hline
-$\theta^b$ & 0.081 (0.067, 0.094) & 0.081 (0.067, 0.099)\tabularnewline
-$\rho^c$ & 0.012 (0.006, 0.019) & 0.012 (0.006, 0.020)\tabularnewline
-$\delta^d$ & 744 (346, 2848) & 723 (348, 2870)\tabularnewline
+\caption{\label{tab:Gene-tree-topologies}Posterior probability of gene tree
+topologies. See Table S\ref{tab:genome} for the reference of the bacteria
+individual names.}
+\noindent \begin{centering}
+\begin{tabular}{cc}
+\hline 
+Gene tree topologies & Posterior probability\tabularnewline
+\hline
+(((SDE1,SDE2),SDD),(SPY1,SPY2)) & 66.7\%\tabularnewline
+(((SDE1,SDE2),(SPY1,SPY2)),SDD) & 9.3\%\tabularnewline
+((SDE1,SDE2),(SDD,(SPY1,SPY2))) & 3.5\%\tabularnewline
+(((SDE1,(SDE2,SDD)),(SPY1,SPY2)) & 1.5\%\tabularnewline
+(((SDE1,SDD),SDE2),(SPY1,SPY2)) & 1.3\%\tabularnewline
+((SDE1,SDD),(SDE2,(SPY1,SPY2))) & 1.2\%\tabularnewline
+((((SDE1,SDE2),SPY1),SPY2),SDD) & 1.0\%\tabularnewline
+((SDE2,SDD),(SDE1,(SPY1,SPY2))) & 1.0\%\tabularnewline
 \hline
 \end{tabular}
-
-$^a$ IQR stands for interquartile range.
-
-$^b$ The median of mutation rate per site weighted over the lengths of blocks.
-
-$^c$ The median of recombination rate per site weighted over the lengths of blocks.
-
-$^d$ The median value for the recombinant tract length weighted over the lengths of blocks.
-
+\par\end{centering}
 \end{table}
@@
 \begin{table}
-\caption{\label{tab:heatmap}}
+\caption{\label{tab:heatmap}Ratios of logarithm base 2 of the observed
+recombinant edges to the expected number of recombinant edges.  Figure
+\ref{fig:Heatmap-of-recombination} shows the values graphically using heat map.
+NA's are for the branch pairs that are impossible. Rows are for donor, and
+columns for recipient.}
 \noindent \centering{}\begin{tabular}{cccccccccc}
@@
 \begin{table}
-\caption{\label{tab:obsheatmap}}
+\caption{\label{tab:obsheatmap}The observed number of recombinant edges.
+Rows are for donor, and columns for recipient.}
 \noindent \centering{}\begin{tabular}{cccccccccc}
@@
 %=============================================================================
-% Table SXXX. Simulation study for the three population parameters.
-\begin{table}
-\caption{\label{tab:sim-three-population-parameters}
-Simulation study of the three population parameters.}
-\noindent \begin{centering}
-\begin{tabular}{ccc}
- & True & Estimates (Standard Deviation)\tabularnewline
-\hline
-Mutation rate per site & 0.081 & 0.081 (0.0011)\tabularnewline
-Recombination rate per site & 0.012 & 0.011 (0.0003)\tabularnewline
-Recombinant tract length & 744 bp & 919 bp (28)\tabularnewline
-\hline
-\end{tabular}
-\par\end{centering}
-\end{table}
-\clearpage{}
-%=============================================================================
-
-% Table S1. Gene Tree topologies
-\begin{table}
-\caption{\label{tab:Gene-tree-topologies}Posterior probability of gene tree
-topologies.}
-\noindent \begin{centering}
-\begin{tabular}{cc}
-\hline 
-Gene tree topologies & Posterior probability\tabularnewline
-\hline
-(((SDE1,SDE2),SDD),(SPY1,SPY2)) & 66.7\%\tabularnewline
-(((SDE1,SDE2),(SPY1,SPY2)),SDD) & 9.3\%\tabularnewline
-((SDE1,SDE2),(SDD,(SPY1,SPY2))) & 3.5\%\tabularnewline
-(((SDE1,(SDE2,SDD)),(SPY1,SPY2)) & 1.5\%\tabularnewline
-(((SDE1,SDD),SDE2),(SPY1,SPY2)) & 1.3\%\tabularnewline
-((SDE1,SDD),(SDE2,(SPY1,SPY2))) & 1.2\%\tabularnewline
-((((SDE1,SDE2),SPY1),SPY2),SDD) & 1.0\%\tabularnewline
-((SDE2,SDD),(SDE1,(SPY1,SPY2))) & 1.0\%\tabularnewline
-\hline
-\end{tabular}
-\par\end{centering}
-\end{table}
-\clearpage{}
-
-% Table S2. 
+% Table S6. Genes with evident recombining gene transfer.
 \begin{table}
@@
 \clearpage{}
-
-%
+%=============================================================================
+
+%=============================================================================
+% Table S. ClonalOrigin's 1st stage result.
 \begin{table}
-\caption{\label{tab:functional-sde-spy}Functional categories asscoiated with higher
-recombination - \# SDE->SPY recombinant edge types.}
+\caption{\label{tab:three}Three population parameter estimates from
+ClonalOrigin. 
+The estimates of the three parameters were based on one of two independent
+MCMC chains.  We considered the similar 
+results of the two independent chains as confirmation of 
+the convergence of the first stage of \texttt{ClonalOrigin} MCMC.
+}
+\noindent \begin{centering}
 \begin{tabular}{ccc}
-\hline 
-p-value & count & description\tabularnewline
-\hline 
-0.0001236 & 28 & rRNA binding\tabularnewline
-0.0001284 & 86 & translation\tabularnewline
-0.0001711 & 39 & ribonucleoprotein complex\tabularnewline
-0.0002230 & 10 & pentose-phosphate shunt\tabularnewline
-0.0003412 & 38 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
-0.0004498 & 42 & structural constituent of ribosome\tabularnewline
-0.0007789 & 47 & carbohydrate transport\tabularnewline
-0.0008152 & 35 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
+& Replicate 1 & Replicate 2\tabularnewline
+Parameter & Estimate (IQR$^a$) & Estimate (IQR)\tabularnewline
+\hline
+$\theta^b$ & 0.081 (0.067, 0.094) & 0.081 (0.067, 0.099)\tabularnewline
+$\rho^c$ & 0.012 (0.006, 0.019) & 0.012 (0.006, 0.020)\tabularnewline
+$\delta^d$ & 744 (346, 2848) & 723 (348, 2870)\tabularnewline
+\hline
+\end{tabular}
+\par\end{centering}
+
+$^a$ IQR stands for interquartile range.
+
+$^b$ The median of mutation rate per site weighted over the lengths of blocks.
+
+$^c$ The median of recombination rate per site weighted over the lengths of blocks.
+
+$^d$ The median value for the recombinant tract length weighted over the lengths of blocks.
+
+\end{table}
+\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% Table SXXX. Simulation study for the three population parameters.
+\begin{table}
+\caption{\label{tab:sim-three-population-parameters}
+Simulation study of the three population parameters.}
+\noindent \begin{centering}
+\begin{tabular}{ccc}
+ & True & Estimates (Standard Deviation)\tabularnewline
+\hline
+Mutation rate per site & 0.081 & 0.081 (0.0011)\tabularnewline
+Recombination rate per site & 0.012 & 0.011 (0.0003)\tabularnewline
+Recombinant tract length & 744 bp & 919 bp (28)\tabularnewline
+\hline
+\end{tabular}
+\par\end{centering}
+\end{table}
+\clearpage{}
+%=============================================================================
+
+\begin{table}
+\caption{\label{tab:functional-all}Functional categories asscoiated with 
+recombination intensity based on all of the recombinant edge types.}
+\begin{tabular}{cccc}
+\hline 
+P-Value & Count & GO term & Description\tabularnewline
+\hline 
+4.596e-06 & 26 & GO:0019843 & rRNA binding\tabularnewline
+5.395e-06 & 40 & GO:0030529 & ribonucleoprotein complex\tabularnewline
+7.935e-06 & 39 & GO:0003735 & structural constituent of ribosome\tabularnewline
+1.546e-05 & 81 & GO:0006412 & translation\tabularnewline
+3.518e-05 & 66 & GO:0005840 & ribosome\tabularnewline
+4.288e-05 & 22 & GO:0006633 & fatty acid biosynthetic process\tabularnewline
+9.098e-05 & 20 & GO:0008610 & lipid biosynthetic process\tabularnewline
+0.0003186 & 33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
+0.000357 & 45 & GO:0008643 & carbohydrate transport\tabularnewline
+0.0006318 & 38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
+0.0009886 & 105 & GO:0003723 & RNA binding\tabularnewline
 \hline 
@@
 \begin{table}
-\caption{\label{tab:functional-spy-sde}Functional categories asscoiated with higher
-recombination - \# SPY->SDE recombinant edge types.}
-\begin{tabular}{ccc}
-\hline 
-p-value & count & description\tabularnewline
-\hline 
-2.025e-06 & 47 & carbohydrate transport\tabularnewline
-1.875e-04 & 38 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
-2.843e-04 & 18 & sodium ion transport\tabularnewline
-3.248e-04 & 35 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
+\caption{\label{tab:functional-notopology}Functional categories asscoiated with 
+recombination intensity based on the topology-not-changing recombinant edge types.}
+\begin{tabular}{cccc}
+\hline 
+P-Value & Count & GO term & Description\tabularnewline
+\hline 
+6.515e-05 & 39 & GO:0003735 & structural constituent of ribosome\tabularnewline
+0.00020986 & 40 & GO:0030529 & ribonucleoprotein complex\tabularnewline
+0.00045049 & 66 & GO:0005840 & ribosome\tabularnewline
+0.00052925 & 26 & GO:0019843 & rRNA binding\tabularnewline
 \hline 
@@
 
-%
 \begin{table}
-\caption{\label{tab:go-events}GO-term enrichment for different
+\caption{\label{tab:functional-sde-spy}Functional categories asscoiated with 
+recombination intensity based on SDE-to-SPY recombinant edge types.}
+\begin{tabular}{cccc}
+\hline 
+P-Value & Count & GO term & Description\tabularnewline
+\hline 
+1.854e-05 & 38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
+2.703e-05 & 81 & GO:0006412 & translation\tabularnewline
+6.482e-05 & 33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
+9.777e-05 & 105 & GO:0003723 & RNA binding\tabularnewline
+0.00012793 & 40 & GO:0030529 & ribonucleoprotein complex\tabularnewline
+0.00025176 & 26 & GO:0019843 & rRNA binding\tabularnewline
+0.00031236 & 72 & GO:0005975 & carbohydrate metabolic process\tabularnewline
+0.00044756 & 317 & GO:0005737 & cytoplasm\tabularnewline
+0.00059675 & 45 & GO:0008643 & carbohydrate transport\tabularnewline
+0.00066874 & 39 & GO:0003735 & structural constituent of ribosome\tabularnewline
+0.00135304 & 10 & GO:0006098 & pentose-phosphate shunt\tabularnewline
+\hline 
+\end{tabular}
+\end{table}
+\clearpage{}
+
+\begin{table}
+\caption{\label{tab:functional-spy-sde}Functional categories asscoiated with 
+recombination intensity based on SPY-to-SDE recombinant edge types.}
+\begin{tabular}{cccc}
+\hline 
+P-Value & Count & GO term & Description\tabularnewline
+\hline 
+1.106e-06 & 45 & GO:0008643 & carbohydrate transport\tabularnewline
+1.511e-06 & 33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
+1.595e-06 & 38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
+1.443e-05 & 187 & GO:0006810 & transport\tabularnewline
+0.0001106 & 371 & GO:0016740 & transferase activity\tabularnewline
+0.0003596 & 72 & GO:0005975 & carbohydrate metabolic process\tabularnewline
+0.0004023 & 13 & GO:0006814 & sodium ion transport\tabularnewline
+0.000823 & 317 & GO:0005737 & cytoplasm\tabularnewline
+0.001142 & 12 & GO:0031402 & sodium ion binding\tabularnewline
+\hline 
+\end{tabular}
+\end{table}
+\clearpage{}
+
+%
+\begin{table}
+\caption{\label{tab:go-events-duplication}GO-term enrichment for different
 evolutionary events - Duplication.}
@@
 \begin{table}
-\caption{\label{tab:go-events}GO-term enrichment for different
+\caption{\label{tab:go-events-loss}GO-term enrichment for different
 evolutionary events - Loss.}
@@
 \begin{table}
-\caption{\label{tab:go-events}GO-term enrichment for different
+\caption{\label{tab:go-events-recombining}GO-term enrichment for different
 evolutionary events - Recombining transfers.}
@@
 %=============================================================================
-% Figure S1: Mauve alignment
+% Figure S1: Mauve alignment of the five genomes.
 \begin{figure}
 \begin{center}
-\includegraphics[scale=0.2]{figures/mauve}
+\includegraphics[scale=0.25]{figures/mauve}
 \end{center}
-\caption{Alignment of the five genomes.}
-\label{fig:mauve}
+\caption{\label{fig:mauve}The genome alignment of the five genomes. The top
+first and second rows are SDE1 and SDE2 genomes, respectively. The third is SDD
+genome. The fourth and fifth rows are SPY1 and SPY2 genomes, respectively.}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[scale=0.7]{figures/cornellf-3-tree}
-\caption{\label{fig:tree5}Species tree of five taxa for \texttt{ClonalOrigin} analysis.
-SPY is the branch of a clade\textit{ of Streptococcus pyogenes}, SDE
-is that for \textit{S.\ dysgalactiae} ssp.\textit{\ equisimilis}, and SD is the
-branch for the three individuals of \textit{S.\ dysgalactiae}. The scale is
-proportional to the expected number of mutations given the Watterson's estimate 
-of mutation
-rate $\theta$ and the tree.}
+\includegraphics[scale=0.5]{figures/cornellf-3-tree}
+\caption{\label{fig:tree5}The species tree of the five genomes inferred using
+\texttt{ClonalFrame}.  SPY is the branch of a clade of \textit{Streptococcus
+pyogenes}, SDE is that for \textit{S.\ dysgalactiae} ssp.\textit{\ equisimilis},
+and SD is the branch for the three individual genomes of \textit{S.\
+dysgalactiae}. The scale is proportional to the expected number of mutations
+given the Watterson's estimate of mutation rate $\theta$ and the tree.}
 \end{figure}
@@
 \end{center}
-\vspace{-.3in}
-\caption{Distribution of core alignment block sizes.}
-\label{fig:blocksize}
+\caption{\label{fig:blocksize}The distribution of core alignment block sizes.}
 \end{figure}
@@
 \begin{figure}
+\begin{center}
 \subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot theta}}
@@
 \subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot delta}}
-\caption{\label{fig:scatter3}Scatter plots of mutation rate, recombination
-rate, and average tract length for the 270 blocks. The red dashed
-line is the global median of all the blocks.}
+\end{center}
+\caption{\label{fig:scatter3}Three scatter plots of mutation rate, recombination
+rate, and average tract length for the blocks. The plus signs are mean values
+for each block. The density of each value for each block is depicted by black
+coulds.  The red dashed lines are the global block-length weighted median of the
+blocks.}
 \end{figure}
@@
 %=============================================================================
-% Figure S: Scatter plot of rho and log(delta)
+% Figure S5: Scatter plot of rho and log(delta)
 \begin{figure}
@@
 \end{center}
-\vspace{-.3in}
-\caption{Scatter plot of recombination rate and average tract length.}
+\caption{Plot of recombination rate and logarithm of average tract length.}
 \label{fig:rhologdelta}
@@
 %=============================================================================
-% Figure S5. A Recombinant Tree of ClonalOrigin Model
+% Figure S6. A Recombinant Tree of ClonalOrigin Model
 \begin{figure}
@@
 \caption{\label{fig:clonalorigin}Recombinant tree in the \texttt{ClonalOrigin} model.
-(A) A recombinant tree relating the five genomes with three recombinant edges.
+(A) A recombinant tree relating the five genomes with four recombinant edges.
 (B) A genomic region of size 100 base pairs with subregions affected by the
-three recombinant edges. Genomic positions are evenly marked by ticks every 
-10-th base pair. Local gene trees induced by the recombinant tree are
-illustrated for each region covered by some combinations of recombinant edges
-along the genomic region (C-G). For example, the local gene tree (E) is created
-by all of the three recombinant edges in the subregion between 51 base pairs
-and 70 base pairs.
-We first describe procedures of summarizing recombinant trees from the first
-viewpoint.  To summarize the posterior samples of \texttt{ClonalOrigin} we
-examined the distribution of topologies of local genealogies induced by the
-sampled recombinant trees. Recall that a recombinant tree induces a local
-genealogy for each site along the analyzed block.  Figure \ref{fig:clonalorigin}
-shows six local gene trees including one that is the same as the species tree
-from the most left-hand side of the block of 100 base pairs to the right.  We
-expected the dominant topology across sampled genealogies to be concordant with
-that of the species tree. However, other frequently sampled topologies were
-expected to be potentially informative about common recombination events. There
-are 105 possible rooted topologies relating five taxa, which allowed us to count
-topologies induced by a posterior sample of recombinant trees.  
-For the second viewpoint of recombinant trees \citet{Didelot2010} suggested an
-approach used to trace recombination intensity along the genome.  For each edge
-type and each site, we recorded the sampling frequency of a recombinant edge of
-the type at that site (e.g., Figure S\ref{fig:muts}).  Figure
-\ref{fig:clonalorigin}B illustrates the posterior probability that a site
-experiences influx from a donor branch using a single recombinant tree.  Summing
-these posteriors over a certain set of edge types, we could obtain a measure of
-recombination intensity along the genome (Figure \ref{fig:clonalorigin}B).  
-}
+four recombinant edges. Genomic positions are evenly marked by ticks every 10-th
+base pair.  For each edge type and each site, we recorded the sampling frequency
+of a recombinant edge of the type at that site.  The posterior probability that
+a site experiences influx from a donor branch using a single recombinant tree.
+Summing these posteriors over a certain set of edge types, we obtain a measure
+of recombination intensity along the genome.  Each segment from b to g are
+scored by the number of recombinant edges per site.  (C-G) Local gene trees
+induced by the recombinant tree are illustrated for each region covered by
+combinations of recombinant edges along the genomic region. For example, the
+local gene tree (E) is created by three recombinant edges in the subregion
+between 51 base pairs and 70 base pairs. The local gene tree (C) is the same as
+the species tree (A) in their topology.  Because there are 105 possible rooted
+topologies relating five taxa, we could count topologies.}
 \end{figure}
@@
 %=============================================================================
-% Figure S6: View of UCSC genome browser for bacterial gene transfer.
+% Figure S7: recombination heatmap from mowgli
+\begin{figure}
+\begin{center}
+% \includegraphics[width=5in]{figures/trans-recomb-heatmap}
+\includegraphics[scale=0.45]{figures/heatmap-recedge-mowgli}
+\end{center}
+\caption{\label{fig:mowgli-recomb-heatmap}Distribution of recombining horizontal
+gene transfers.}
+\end{figure}
+\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% Figure S8: gene family sizes
+% FIGURE - gene family sizes
+\begin{figure}
+\begin{center}
+\includegraphics[width=5in]{figures/famsizes}
+\end{center}
+\caption{\label{fig:famsizes}Distribution of gene family sizes.}
+\end{figure}
+\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% Figure S9: Comparison of ClonalOrigin and Mowgli analysis
+\begin{figure}
+\begin{center}
+\includegraphics[width=5in]{figures/compareMowgliCoRecombining}
+\end{center}
+\caption{\label{fig:cmpcomowgli}Plot of recombination intensity against number
+of recombining gene transfer.}
+\end{figure}
+\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% Figure S10: HGT Comparison of ClonalOrigin and Mowgli analysis
+\begin{figure}
+\begin{center}
+\includegraphics[width=5in]{figures/compareMowgliCoHgt}
+\end{center}
+\caption{\label{fig:cmpcomowglihgt}Plot of recombination intensity against number
+of horizontal gene transfer.}
+\end{figure}
+\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% Figure S: View of UCSC genome browser for bacterial gene transfer.
 \begin{figure}
@@
 %=============================================================================
-% Figure S7: 
+% Figure S: 
 \begin{figure}
@@
 %=============================================================================
-% Figure S8:
+% Figure S:
 \begin{figure}
@@
 %=============================================================================
-% Figure S9:
+% Figure S:
 \begin{figure}
@@
 
-% Figure S7: gene family sizes
-%=============================================================================
-% FIGURE - gene family sizes
-\begin{figure}
-\begin{center}
-\includegraphics[width=5in]{figures/famsizes}
-\end{center}
-\vspace{-.3in}
-\caption{Distribution of gene family sizes.}
-\label{fig:famsize}
-\end{figure}
-%=============================================================================
-
 \clearpage
-
-% Figure S8: recombination heatmap from mowgli
-%=============================================================================
-% FIGURE - Transfer heatmap
-\begin{figure}
-\begin{center}
-% \includegraphics[width=5in]{figures/trans-recomb-heatmap}
-\includegraphics[scale=0.45]{figures/heatmap-recedge-mowgli}
-\end{center}
-\caption{Distribution of recombining horizontal gene transfers.}
-\label{fig:mowgli-recomb-heatmap}
-\end{figure}
-\clearpage{}
-%=============================================================================
 
@@
 %
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Functional category association
+Lateral transfer of transporter related genes are well known.  Functional
+categories associated with higher recombining gene transfer were found to
+include phosphotransferase systems and ATP-binding cassette (ABC) transporters,
+i.e., carbohydrate transport (GO:0008643) (Table \ref{tab:functional}).
+
+We also
+found several aminoacyl tRNA synthetases in gene ontology of translation
+(GO:0006412) to be significantly associated with higher recombining gene
+transfer.
+
+Functional categories concerning ribosomes were also
+associated with higher recombining gene transfer. These categories were not
+found significantly associated with higher recombining gene transfer when we
+considered only gene transfer between SPY and SDE genomes, which implied that
+many recombining gene transfer events should be attributed to those of internal
+species branches as donors.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% From the discussion about virulence genes and two gene transfer
+We suspected that virulence genes were exchanged between SPY and
+SDE, but did not find any directionality of the gene flow.
+This does not necessarily mean that
+virulence genes were not exchanged between SPY and SDE via non-recombining
+transfer.  Because the \texttt{Mowgli} analysis just did not reveal unbalanced
+non-recombining or recombinding gene transfer, we suspected that \texttt{Mowgli}'s
+parsimony might not be powerful in detecting unbalanced gene flow.  We
+conjecture the following hypothesis of the evolution of SPY and SDE to support
+the non-recombining virulence gene exchange between SPY and SDE.  Pathogenicity
+of SDE might have existed before the common ancestor of SPY and SDE.  We
+speculated that SPY might have been split off from the common ancestor that was
+SDE-like because the population size of SDE was larger than that of SPY (CITE).
+After the split, SPY might have been specialized in inhabting human hosts, and
+SDE in colonizing veterinary hosts. Recent surge in cases of human diseases
+attributed to SDE might be because of secondary contact of SPY and SDE since
+their split from their common ancestor.  When SPY thrives in its environment of
+human hosts and another relatively close species, SDE, starts to cohabit with
+the indigenous species SPY, genetic elements in one of the two species could
+benefit another species. If the two species do not compete with each other, then
+the increased genetic variation from multiple species would eventually bring
+advantage to both of the species.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Functional category association 
+\textbf{Two functional categories and recombining gene transfer:}
+Among the genes shown in Table S\ref{tab:functional} two classes of proteins
+attracted our attentions: aminoacyl tRNA synthetases (aaRS) and
+phosphotransferase system (PTS).  Aminoacyl tRNA synthetases charge tRNA with an
+amino acid so that amino acids are concatenated from charged tRNA onto a growing
+peptides through translation. The aaRS's are characterized by an amino acid that
+they charge tRNA with.  This gene family had been a target for horizontal gene
+transfer \citep[e.g.,][]{Lamour1994,Woese2000,Koonin2001}, which lent itself to
+an example of gene transfer in bacteria.  There were also recent reports of gene
+transfer in tRNA aminoacylation \citep{Andam2010,Wang2011}.  In our analysis,
+some of them appeared to be transferred from SPY to SDE: leucyl-tRNA synthetase,
+aspartyl-tRNA synthetase, and arginyl-tRNA synthetase. Half of phenylalanyl-tRNA
+synthetase gene appeared to be transferred in the reverse direction.
+
+Bacterial phosphotransferase system (PTS), also known as phosphoenolpyruvate
+(PEP) group translocation, transports sugars such as glucose into the cell where
+phosphorylation of the sugar happens. There have been studies of gene transfer
+of PTS in bacteria \citep[e.g.,][]{Barabote2005,Zuniga2005}. As the main source
+of energy for bacteria, sugars in the bacterial niches must have been acting as
+important selective pressure. It is not surprising that PTS sugars differ among
+bacterial groups because bacterial species are correlated with their environment
+and energy sources theirof.  Table S\ref{tab:functional} shows that two PTS
+genes appeared to be transferred from SPY to SDE: {}``PTS system,
+mannose-specific IIC component'' and {}``PTS system, fructose-specific IIABC
+component.'' Whereas almost half of the former gene was inferred to be affected
+by recombinant edges, only 3\% of the latter gene appeared to be affected.
+
+% This agreement
+% between the two different methods, applied on somewhat different parts of the
+% data, serves as validation for the signal picked up by these separate
+% association studies.
+
+% These topology-not-changing recombinant edges tend change only branch lengths.

### 2011-10-06 sh/sim5-prepare.sh

@@ -66,7 +66,7 @@ function sim5-prepare {
         echo -e "  Skipping generating local trees..." 
       fi
 
-      echo -n "Do you wish to creat jobidfile for submission (y/n)? "
+      echo -n "Do you wish to create jobidfile for submission (y/n)? "
       read WISH
       if [ "$WISH" == "y" ]; then
         echo -e "  Simulating data using recombinant trees..."
@@ -93,6 +93,27 @@ function sim5-prepare {
       else
         echo -e "  Skipping creating jobidfile ..." 
       fi
+
+      echo -n "Do you wish to get the output result files (y/n)? "
+      read WISH
+      if [ "$WISH" == "y" ]; then
+        for b in $(eval echo {1..$NUMBER_BLOCK}); do
+          for REPETITION in $(eval echo {1..$HOW_MANY_REPETITION}); do
+            DATADIR=output/s17/$REPETITION/data
+            g=$(($REPETITION * 10 + 1))
+            for REPLICATE in $(eval echo {1..$HOW_MANY_REPLICATE}); do
+              mkdir -p output/s17/$REPETITION/run-clonalorigin/output2/$REPLICATE
+              XML=output/s17/$REPETITION/run-clonalorigin/output2/$REPLICATE/core_co.phase3.xml.$b
+              scp -q cac:run/mauve/100311/$XML $XML
+            done
+            echo -ne "block $b - $g\r"
+          done
+          echo -ne "                                           \r"
+        done
+      else
+        echo -e "  Skipping creating jobidfile ..." 
+      fi
+
       break
     fi
   done

### 2011-10-07 doc/manuscript.tex
The whole procedure must be rerun.

@@
 opportunistic human pathogen.  We aims at which parts of the genomic
-organization and human pathogenic virulence factors of SDE could have
-originated in those of SPY, or vice versa.  Our genome-wide survey of gene
-transfer indicated that gene transfer via homologous recombination occurred
-relatively from SPY to SDE.  Although we did not find enrichment of virulence
-genes in the homologous recombining gene transfer, we found evidence that
+organization and human pathogenic virulence factors of SDE could have originated
+in those of SPY, or vice versa.  Our genome-wide survey of gene transfer
+indicated that gene transfer via homologous recombination occurred more for the
+net direction from SPY to SDE.  Although we did not find enrichment of virulence
+genes in homologous recombining gene transfer, we found evidence that
 virulence genes transferred more often via non-recombining fashion without
@@
 genome-wide resource of recombining and non-recombining gene transfer for the
-study of the genomes is provided so that research community in pathogenic
-bacteria could take advantage of it.
+study of the SPY and SDE genomes is provided so that research community in
+pathogenic bacteria could take advantage of it.
 
@@
 \citep{Didelot2010}. The model of homologous recombination allowed them to study
-flows of genetic segment between bacteria individuals of \textit{Bacillus
+flows of genetic segments between bacteria individuals of \textit{Bacillus
 cereus} species in the core genomes.  Other evolutionary processes acting on
@@
 to statistically more sound model-based approaches such as maximum likelihood.
-Instead, a parsimony-based approach to inferring gene gain, loss, and transfer
-events was developed by \citep{Doyon2011}.  We noted that the first approach
-could be used to infer transfer of genomic segments due to homologous
-recombination, and the second could do similarly attributed to horizontal gene
+Instead, a parsimony-based approach to inferring gene duplication, loss, and
+transfer events was developed by \citet{Doyon2011}.  We noted that the first
+approach could be used to infer transfer of genomic segments due to homologous
+recombination, and the second could help similarly attributed to horizontal gene
 transfer.
@@
 with the model-based approach. This was the first attempt as far as we know to
-analyaze SPY and SDE in genome-wide scale for detecting gene transfer events not
+analyze SPY and SDE in genome-wide scale for detecting gene transfer events not
 only in the core genome but also in the pan genome, which makes the analysis
@@
 chosen (Table S\ref{tab:genome}) to study recombining gene transfer using the
-model-based approach.  The genome alignment of the five genomes showed two
+model-based approach.  A genome alignment of the five genomes showed two
 distinct genome structures of SPY and SDE (Figure S\ref{fig:mauve}). The SDE and
@@
 whether SDD was different from either of SDE1 and SDE2 from the graphical view
-of the genome alignment, the spcies tree inferred from the genome alignment
+of the genome alignment, a spcies tree inferred from the genome alignment
 showed that SDD branch was attached to the clade branch of SDE1 and SDE2 (Figure
@@
 appeared to change more dynamically. Especially, a genomic region of SDE1 near
-20 kilobase pairs from the origin of replication contained alignment blocks with
+200 kilobase pairs from the origin of replication contained alignment blocks with
 bell-shaped recombination rates.
@@
 dominant topology across sampled genealogies to be concordant with that of the
-species tree.  Indeed, 66.7\% of the local gene tree topologies (see section
+species tree.  Indeed, two thirds of the local gene tree topologies (see section
 \textbf{Materials and Methods}) along the alignment blocks were the same as the
@@
 recombinant edges.  Similarly, all of the cells for which a recipient branch was
-older than the donor, and two branches did not share any contemporary time were
+older than the donor, and the donor and recipient branches did not share any contemporary time were
 colored green. The diagonal cells were colored gray, indicating no significant
@@
 recombinant edges were similar. As recombinant edges starting from and ending in
-the same species tree branch would not change the species tree, which would not
+the same species tree branch would not change the species tree topology, which
+might not
 change the likelihood much, the cells would be colored gray.  Because SDD is a
@@
 donor since the similar values of recombining gene transfer intensity from SDE1
-or SDE2 to the other three external branches.  Surprisingly, the recombining
+or SDE2 to one of the other three external branches. Surprisingly, the recombining
 gene transfer log base 2 ratios between SPY1 and SPY2 (1.06 and 1.07) was
-comparable with, but even less than those from SPY1 and SPY2 to SDE1 and SDE2
+comparable with, yet even less than those from SPY1 and SPY2 to SDE1 and SDE2
 (1.25, 1.25, 1.17, and 1.29).  Effectively, the recombining gene transfer from a
@@
 concordant with the fact that SPY and SDE shared the same ecological niche
-despite their different genomic organization.  The gene transfer log base 2
+despite their dissimilar genomic organization.  The gene transfer log base 2
 ratio from SPY1 and SPY2 to SDE branch was also larger than any type of
@@
 the SPY and SDE clades were inferred in smaller numbers (<6) compared to those
-within each clade (23--34). We were puzzled by this finding because virulence
-genes were found to be associated with horizontal gene transfer, and we could
+within each clade. We were puzzled by this finding because virulence
+genes were found to be associated with horizontal gene transfer, and yet we could
 not show net directionality of horizontal gene transfer between SPY and SDE.
@@
 both methods (Tables S\ref{tab:functional-all} and
-S\label{tab:go-events-recombining}).  Secondly, both of them did not show
+S\ref{tab:go-events-recombining}).  Secondly, both of them did not show
 association of virulence genes with recombining gene transfer. Lastly, we found
@@
 
-We created a web-resource for gene transfer available at
+We created a web-resource (Figure S\ref{fig:ucsc}) for gene transfer available at
 \url{http://strep-genome.bscb.cornell.edu}.  For each of the five individual
@@
 
-Two classes of functional categories notably were associated with recombining
-gene transfer.  Firstly, genes involved in sensing queues of environmental
-changes would be important for bacteria to adapt to changing environment.  Genes
-encoding membrane-bound proteins would be involved in the function.  Several of
-functional categories that were associated with gene transfer might be related
-with membrane-bound protein coding genes (Table \ref{tab:functional}).  Sugar
-phosphotransferase classes were ranked top in the list: phosphrylation must be
-important in signal transduction to recognize environmental changes.  The
-transport functional categories of carbohydrate transport and sodium ion
-transport also belonged to genes encoding membrane-bound proteins.  Also highly
-ranked were fatty acid or lipid biosyntheses that must be involved in producing
-membrane-bound molecules. These functional categories related with
-membrane-bound proteins might allow for bacteria to adapt to changing
-environment via recombining gene transfer.
-
-Secondly, aminoacyl tRNA synthetases had been a focus of gene transfer
-\citep{Woese2000}, and these aminoacyl tRNA synthetases belonged to the
+Two classes of functional categories were notably to be inferred to associated
+with recombining gene transfer.  Firstly, genes involved in sensing queues of
+environmental changes would be important for bacteria to adapt to changing
+environment.  Genes encoding membrane-bound proteins would be involved in the
+function.  Several of functional categories that were associated with gene
+transfer might be related with membrane-bound protein coding genes (Table
+\ref{tab:functional}).  Sugar phosphotransferase classes were ranked top in the
+list: phosphrylation could be important in signal transduction to recognize
+environmental changes.  The transport functional categories of carbohydrate
+transport and sodium ion transport also belonged to genes encoding
+membrane-bound proteins.  Also highly ranked were fatty acid and lipid
+biosyntheses that must be involved in producing membrane-bound protein
+molecules. These functional categories related with membrane-bound proteins
+might allow for bacteria to adapt to changing environment via recombining gene
+transfer.  Secondly, aminoacyl tRNA synthetases had been a focus of gene
+transfer \citep{Woese2000}, and these aminoacyl tRNA synthetases belonged to the
 functional category of translation (GO:0006412).  They were inferred to be
@@
 translation category was siginificant only in recombining gene transfer (see
-Tables \ref{tab:functional}, \label{tab:go-events} and S\label{tab:go-events-recombining}).
+Tables \ref{tab:functional}, \ref{tab:go-events} and
+S\ref{tab:go-events-recombining}).
 
@@
 \ref{tab:functional}) in computing recombination intensity.  Therefore, high
-recombination intensity in ribosome related genes would likely to be attributed
+recombination intensity in ribosome related genes would be likely attributed
 to topology-not-changing recombinant edges (Table
@@
 edges in computing recombination intensity (see Tables
-S\ref{tab:go-events-recombining} and S\ref{tab:functional-all}).  Because the
-parsimony-based method did not take species tree branch lengths into account, we
-were likely to the agreement between the two different methods.
+S\ref{tab:functional-all} and S\ref{tab:go-events-recombining}).  Because the
+parsimony-based method did not take species tree branch lengths into account, 
+the agreement between the two different methods might be likely.
 
@@
 recombination intensity based on topology changing recombinant edge types.}
-\begin{tabular}{cccc}
+\begin{tabular}{cccl}
 \hline 
@@
 non-recombining horizontal gene transfer.}
-\begin{tabular}{cccc}
+\begin{tabular}{cccl}
 \hline 
@@
 2 of the number of recombination events inferred relative to its expectation
-under the prior, for a donor/recipient pair of branches. The donor is given by
-the y-axis; the recipient on the x-axis.  The green cells are meaningless
+under the prior, for a donor and recipient pair of branches. The donor is given by
+the y-axis; the recipient on the x-axis.  The green colored cells are meaningless
 because the recipient species tree branch of the cells are older than the donor
@@
 \end{center}
-\caption{\label{fig:hgt-heatmap}Distribution of horizontal gene transfers.
+\caption{\label{fig:hgt-heatmap}Heat map of horizontal gene transfers using the
+six genomes.
 Each cell of the color matrix shows the number of inferred non-recombining
 horizontal gene transfer events between the different branches of the species
-tree.}
+tree. Black colored cells represent no inferred events.}
 \end{figure}
@@
 \begin{table}
-\caption{\label{tab:clonalframe}Parameter estimates from ClonalFrame. 
+\caption{\label{tab:clonalframe}Parameter estimates from \texttt{ClonalFrame}. 
 $\theta$ is population mutation rate,
@@
 \end{tabular}
+\end{table}
+\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% Table S6. ClonalOrigin's 1st stage result.
+\begin{table}
+\caption{\label{tab:three}Three population parameter estimates from
+ClonalOrigin. 
+The estimates of the three parameters were based on one of two independent
+MCMC chains.  We considered the similar 
+results of the two independent chains as confirmation of 
+the convergence of the first stage of \texttt{ClonalOrigin} MCMC.
+}
+\noindent \begin{centering}
+\begin{tabular}{ccc}
+\hline
+& Replicate 1 & Replicate 2\tabularnewline
+Parameter & Estimate (IQR$^a$) & Estimate (IQR)\tabularnewline
+\hline
+$\theta^b$ & 0.081 (0.067, 0.094) & 0.081 (0.067, 0.099)\tabularnewline
+$\rho^c$ & 0.012 (0.006, 0.019) & 0.012 (0.006, 0.020)\tabularnewline
+$\delta^d$ & 744 (346, 2848) & 723 (348, 2870)\tabularnewline
+\hline
+\end{tabular}
+\par\end{centering}
+
+$^a$ IQR stands for interquartile range.
+
+$^b$ The median of mutation rate per site weighted over the lengths of blocks.
+
+$^c$ The median of recombination rate per site weighted over the lengths of blocks.
+
+$^d$ The median value for the recombinant tract length weighted over the lengths of blocks.
+
 \end{table}
@@
 %=============================================================================
-% Table S. ClonalOrigin's 1st stage result.
-\begin{table}
-\caption{\label{tab:three}Three population parameter estimates from
-ClonalOrigin. 
-The estimates of the three parameters were based on one of two independent
-MCMC chains.  We considered the similar 
-results of the two independent chains as confirmation of 
-the convergence of the first stage of \texttt{ClonalOrigin} MCMC.
-}
-\noindent \begin{centering}
-\begin{tabular}{ccc}
-& Replicate 1 & Replicate 2\tabularnewline
-Parameter & Estimate (IQR$^a$) & Estimate (IQR)\tabularnewline
-\hline
-$\theta^b$ & 0.081 (0.067, 0.094) & 0.081 (0.067, 0.099)\tabularnewline
-$\rho^c$ & 0.012 (0.006, 0.019) & 0.012 (0.006, 0.020)\tabularnewline
-$\delta^d$ & 744 (346, 2848) & 723 (348, 2870)\tabularnewline
-\hline
-\end{tabular}
-\par\end{centering}
-
-$^a$ IQR stands for interquartile range.
-
-$^b$ The median of mutation rate per site weighted over the lengths of blocks.
-
-$^c$ The median of recombination rate per site weighted over the lengths of blocks.
-
-$^d$ The median value for the recombinant tract length weighted over the lengths of blocks.
-
-\end{table}
-\clearpage{}
-%=============================================================================
-
-%=============================================================================
 % Table SXXX. Simulation study for the three population parameters.
@@
 recombination intensity based on all of the recombinant edge types.}
-\begin{tabular}{cccc}
+\begin{tabular}{cccl}
 \hline 
@@
 recombination intensity based on the topology-not-changing recombinant edge types.}
-\begin{tabular}{cccc}
+\begin{tabular}{cccl}
 \hline 
@@
 recombination intensity based on SDE-to-SPY recombinant edge types.}
-\begin{tabular}{cccc}
+\begin{tabular}{cccl}
 \hline 
@@
 evolutionary events - Duplication.}
-\begin{tabular}{cccc}
+\begin{tabular}{cccl}
 \hline 
@@
 evolutionary events - Loss.}
-\begin{tabular}{cccc}
+\begin{tabular}{cccl}
 \hline 
@@
 evolutionary events - Recombining transfers.}
-\begin{tabular}{cccc}
+\begin{tabular}{cccl}
 \hline 
@@
 dysgalactiae}. The scale is proportional to the expected number of mutations
-given the Watterson's estimate of mutation rate $\theta$ and the tree.}
+given the Watterson's estimate of mutation rate and the tree.}
 \end{figure}
@@
 \end{center}
-\caption{\label{fig:mowgli-recomb-heatmap}Distribution of recombining horizontal
-gene transfers.}
+\caption{\label{fig:mowgli-recomb-heatmap}Heat map of recombining gene transfers using the
+six genomes.
+Each cell of the color matrix shows the number of inferred recombining
+gene transfer events between the different branches of the species
+tree. Black colored cells represent no inferred events.}
 \end{figure}
@@
 %=============================================================================
-% Figure S: View of UCSC genome browser for bacterial gene transfer.
+% Figure S11: View of UCSC genome browser for bacterial gene transfer.
 \begin{figure}
 \includegraphics[scale=0.8]{figures/ucsc}
-\caption{\label{fig:muts}Posterior probability of recombinant edges on 
-aspartyl-tRNA
-synthetase.  The
-top line represents a genomic location of the current view in SPY1 genome. Locus
-tags are represented below. The start of a locus tag is red colored, and the end
-blue colored. Each of nine horizontal bars represents a branch of the five
-species tree. The species tree branch is a recipient species.
-The red area on SDE1 line represents gene fragment transfer from
-SPY1 due to recombination in locus SpyM3\_1814 of aspartyl-tRNA synthetase.}
+\caption{\label{fig:ucsc}UCSC genome browser tracks for gene transfer study of
+SPY and SDE.}
 \end{figure}

### 2011-10-07 pl/count-observed-recedge.pl

@@ -27,7 +27,6 @@
 
 use strict;
 use warnings;
-use IO::Uncompress::Bunzip2 qw(bunzip2 $Bunzip2Error);
 use XML::Parser;
 use Getopt::Long;
 use Pod::Usage;
@@ -36,14 +35,18 @@ require "pl/sub-error.pl";
 require "pl/sub-simple-parser.pl";
 require "pl/sub-array.pl";
 require "pl/sub-heatmap.pl";
+require "pl/sub-newick-parser.pl";
 sub makeXMLFilename($$$$);
 sub get_exp_map($$);
 sub get_obs_map_iteration($$);
 
+my $cmd = ""; 
+sub process {
+  my ($a) = @_; 
+  $cmd = $a; 
+}
 $| = 1; # Do not buffer output
-
 my $VERSION = 'count-observed-recedge.pl 1.0';
-
 my $man = 0;
 my $help = 0;
 my %params = ('help' => \$help, 'h' => \$help, 'man' => \$man);        
@@ -56,13 +59,15 @@ GetOptions( \%params,
             'e=s',
             'n=i',
             's=i',
+            'out=s',
             'meanonly',
             'meanfile=s',
             'xmlbasename=s',
             'endblockid',
             'append',
             'obsonly',
-            'check'
+            'check',
+            '<>' => \&process
             ) or pod2usage(2);
 pod2usage(1) if $help;
 pod2usage(-exitstatus => 0, -verbose => 2) if $man;
@@ -71,7 +76,6 @@ my $xmlDir;
 my $heatDir = "";
 my $numBlocks;
 my $numSpecies;
-
 my $obsonly = 0;
 my $meanonly = 0;
 my $meanfile = "";
@@ -79,6 +83,8 @@ my $append = 0;
 my $check = 0;
 my $xmlBasename = "core_co.phase3";
 my $endblockid = 0;
+my $outfile;
+
 if (exists $params{d})
 {
   $xmlDir = $params{d};
@@ -93,31 +99,15 @@ if (exists $params{e})
   $heatDir = $params{e};
 }
 
-if (exists $params{obsonly})
-{
-  $obsonly = 1;
-  unless ($heatDir eq "")
-  {
-    &printError("obsonly is used without -e option");
-  }
-}
-
 if (exists $params{n})
 {
   $numBlocks = $params{n};
 }
-else
-{
-  &printError("you did not specify a number of blocks");
-}
+
 
 if (exists $params{s})
 {
[…]

### 2011-10-07 pl/sub-newick-parser.pl

@@ -324,6 +324,4 @@ sub maNewicktFindRedEdgePair ($$)
   return $m;
 }
 
-
-
 1;

### 2011-10-07 sh/sim5-prepare.sh

@@ -66,6 +66,27 @@ function sim5-prepare {
         echo -e "  Skipping generating local trees..." 
       fi
 
+      echo -n "Do you wish to check the simulate data for gap characters (y/n)? "
+      read WISH
+      if [ "$WISH" == "y" ]; then
+        echo -e "  Checking data using recombinant trees..."
+
+        for b in $(eval echo {1..$NUMBER_BLOCK}); do
+          for REPETITION in $(eval echo {1..$HOW_MANY_REPETITION}); do
+            DATADIR=$BASEDIR/$REPETITION/data
+            g=$(($REPETITION * 10 + 1))
+            BLOCKSIZE=$(echo `perl pl/get-block-length.pl $XMLBASEDIR/core_co.phase3.xml.$b.$g`) 
+            for REPLICATE in $(eval echo {1..$HOW_MANY_REPLICATE}); do
+              XMFA=$DATADIR/core_alignment.$b.$g.$REPLICATE.xmfa
+              perl pl/checkxmfa.pl $XMFA
+            done
+          done
+        done
+      else
+        echo -e "  Skipping generating local trees..." 
+      fi
+
+
       echo -n "Do you wish to create jobidfile for submission (y/n)? "
       read WISH
       if [ "$WISH" == "y" ]; then
@@ -110,6 +131,48 @@ function sim5-prepare {
           done
           echo -ne "                                           \r"
         done
+      else
+        echo -e "  Skipping creating jobidfile ..." 
+      fi
+
+      echo -n "Do you wish to analyze the output result files (y/n)? "
+      read WISH
+      if [ "$WISH" == "y" ]; then
+        for REPETITION in $(eval echo {1..$HOW_MANY_REPETITION}); do
+        # for REPETITION in $(eval echo {1..1}); do
+          RUNCLONALORIGIN=$BASEDIR/$REPETITION/run-clonalorigin
+          RUNANALYSIS=$BASEDIR/$REPETITION/run-analysis
+          mkdir $RUNANALYSIS
+          for REPLICATE in $(eval echo {1..$HOW_MANY_REPLICATE}); do
+            perl pl/count-observed-recedge.pl obsonly \
+              -d $RUNCLONALORIGIN/output2/${REPLICATE} \
+              -n $NUMBER_BLOCK \
+              -endblockid \
+              -obsonly \
+              -out $RUNANALYSIS/obsonly-recedge-$REPLICATE.txt
[…]

### 2011-10-13 doc/manuscript.tex
Just an update.

@@
 \section*{Abstract}
+
 Homologous recombination and horizontal gene transfer are investigated with
@@
 in those of SPY, or vice versa.  Our genome-wide survey of gene transfer
-indicated that gene transfer via homologous recombination occurred more for the
-net direction from SPY to SDE.  Although we did not find enrichment of virulence
-genes in homologous recombining gene transfer, we found evidence that
-virulence genes transferred more often via non-recombining fashion without
-replacing parts of destination genomes than did non-virulence genes.  A
-genome-wide resource of recombining and non-recombining gene transfer for the
-study of the SPY and SDE genomes is provided so that research community in
-pathogenic bacteria could take advantage of it.
+indicated that gene transfer via homologous recombination occurred more often
+for the net direction from SPY to SDE.  We found evidence that virulence genes
+transferred more often via non-recombining fashion without replacing parts of
+destination genomes than did non-virulence genes.  A genome-wide resource of
+recombining and non-recombining gene transfer for the study of the SPY and SDE
+genomes is provided so that research community in the pathogenic SPY and SDE
+bacteria could take advantage of it.
 
@@
 from other lineages and vertically transmitted mutations on genes within the
-same lineage.  Mutations on a gene modifiy relatively small portions of the gene
-through substitutions, insertions, and deletions of deoxyribonucleic acids.
-Gene transfers can reshape the organization of a bacterial genome,
+same lineage.  While mutations on a gene modifiy relatively small portions of the 
+gene through substitutions, insertions, and deletions of deoxyribonucleic acids,
+gene transfers can reshape the organization of a bacterial genome,
 which could revamp biological functions of the species.  Recipient bacteria of
@@
 distinguishing recombining gene transfer process from horizontal, or
-non-recombining, gene transfer, and tackled their roles in association with
+non-recombining, gene transfer, and tackled the two processes in association with
 virulence genes.
@@
 analyze SPY and SDE in genome-wide scale for detecting gene transfer events not
-only in the core genome but also in the pan genome, which makes the analysis
+only in the core genome but also in the pan genome, which made the analysis
 complete in terms of the amount of used genome data. Our analysis of core
@@
 direction.
-
-\clearpage{}
 
@@
 
+\texttt{Limitation of ClonalOrigin: inability of inferring recombinant edges}
+
+\texttt{Need of a new model with asymmetry of gene transfer}
+
+\texttt{Need of more SDE genomes}
+
 \section{Materials and Methods}
@@
 options to attempt to align the six genomes, which led to too small proportion
-of a core genome. We had to remove the SEE sample from the genome alignment
+of core genomes. We had to remove the SEE sample from the genome alignment
 although we kept it in the parsimony-based approach.  We then used
@@
 and estimated the species tree relating the five genomes. \texttt{ClonalFrame}
-estimates the average recombinant tract length and the ratio between
-recombination and mutation rates, which were not used in the following analysis.
+also estimates the ratio between recombination and mutation rates.
 We used a single Markov chain Monte Carlo (MCMC) with 10000 burn-in generations
@@
 
-\subsection{Local gene trees and recombination intensity}
-Because the procedure of summarizing posterior sample of \textit{recombinant
-tree} was not detailed in the user guide at the time when we used
-\texttt{ClonalOrigin}, we elected to use Figure S\ref{fig:clonalorigin} to
-illustrate a recombinant tree in order to clarify summarization of
-\textit{recombinant history} posterior samples.  A recombination history is
-represented by a \textit{recombinant tree}, which is a species tree augmented
-with \textit{recombinant edges}.  Each recombinant edge is defined by the time
-and branch of origin and destination, and the genomic interval that it affects.
-The branch of origin is referred to as the \textit{donor} and the destination
-branch is referred to as the \textit{recipient}. The type of a recombinant edge
-is determined by the donor and recipient species tree branches.  A recombinant
-tree can be viewed as a restricted version of an ancestral recombination graph
-(ARG), and as such it induces a local genealogy for each site in the genomic
-block it describes.  \citet{Didelot2010} presented two procedures of summarizing
-recombinant trees by viewing trees from two different angles: by counting types
-of recombinant edges regardless of genomic positions affected by the edges, and
-by measuring recombinant edge changes to the species tree in site-wise manner.
-By similarly following their approaches we quantified additional aspects of
-recombinant trees; we counted gene tree topologies for each site, and also
-quantified recombination intensity along the genome.
-
-The recombination intensity allowed us to find genomic regions enriched for
-recombination events in general, or regions enriched for recombination events of
-certain types (e.g., between SPY and SDE).  We note that the recombination
-intensity is not normalized to the interval $[0,1]$. Rather than giving
-probabilistic meanings to the recombination intensity we interpreted it as an
-average number of recombinant edge types per site.  The measure appeared to act
-as an effective indicator of recombination intensity (see subsection of
-\textbf{Simulation Study}).  Our simulation study indicated that edges of
-certain types were better indicators of true recombination events than others.
-For instance, recombinant edges which resulted in a recombinant tree with the
-same topology as the species tree appeared to be more prone to spurious
-sampling.  Such edges were ones where the donor was the same branch, sister
-branch, or parent branch of the recipient branch.  Therefore, in some cases we
-elected to remove such edges from consideration when measuring recombination
-intensity.  The tree topologies changes would be correlated with recombination
-intensity.  The tree topologies would show more effectively regions where
-recombination occurs along the genome (see subsection of \textbf{UCSC Genome
-Browser}).
-
 \subsection{Parsimony-based method for horizontal gene transfer}
@@
 duplications, losses, and non-recombining horizontal gene transfers.  In
-contrast to \texttt{ClonalOrigin}, \texttt{Mowgli} assumes a parsimony model for
+contrast to \texttt{ClonalOrigin}, \texttt{Mowgli} is parsimony-based for
 inferring events and uses pre-determined gene clusters where genes are treated
@@
 it attempted to explain different copy numbers by a series of gain and loss
-events.  We here were interested in the approach of \texttt{Mowgli} to study
-non-recombining horizontal gene transfers.  We obtained gene family clusters
-defined in \citet{Suzuki2011} using the OrthoMCL method \citep{Li2003} over the
-full set of six genomes including \textit{S.\ equi} subsp.\textit{\ equi} strain
-4047 (SEE).  For each family cluster, the corresponding protein sequences were
-aligned using the MUSCLE program \citep{Edgar2004a} and nucleotide alignments
-were constructed by mapping nucleotide sequences onto each protein sequence
+events.  We obtained gene family clusters defined in \citet{Suzuki2011} using
+the \texttt{OrthoMCL} method \citep{Li2003} over the full set of the six genomes
+including \textit{S.\ equi} subsp.\textit{\ equi} strain 4047 (SEE).  For each
+family cluster, the corresponding protein sequences were aligned using the
+\texttt{MUSCLE} program \citep{Edgar2004a} and nucleotide alignments were
+constructed by mapping nucleotide sequences onto each protein sequence
 alignment. We then identified recombination within gene alignments using the
-Single Breakpoint Recombination (SBP) method from the HyPhy package
-\citep{KosakovskyPond2006}.  Using the Akaike Information Criterion (AIC), SBP
-inferred that 47.5\% (1,094) of gene families contained at least one
-topology-changing recombination. Each of these families was then split into two
-gene fragments.  Gene trees were constructed from nucleotide alignments for each
-gene fragment using the RAxML program \citep{Stamatakis2006}, and were rooted by
-minimizing inferred duplications and losses.  Each rooted gene tree was then
-reconciled using the \texttt{Mowgli} program \citep{Doyon2011}, which inferred
-the most parsimonious gene duplication, loss, and transfer events for each
-family.  In the case of multiple maximum parsimonious reconciliations, one was
-chosen uniformly at random.  Unless specified otherwise, default setings were
-used for all programs.  Since the method of \texttt{Mowgli} did not directly
-model recombination events, recombination-based transfers were inferred in a
-post-processing step, by checking whether the transfer was coupled with a loss
-of a homologous gene in the recipient branch (indicating gene replacement rather
-than gain).  A transfer event was classified as non-recombining if the
-destination species of that transfer had an extant descendant that contained
-copies of that genes which were descendants of the transferred gene as well as
-copies which were non-descendants.  We noted that this parsimonious approach was
-imperfect and would tend to undercount events, especially recombining gene
-transfer events, which were modeled using two events: a transfer and a loss.
-Nevertheless, this analysis allowed us to study non-recombining gene transfers,
-which were not modeled by \texttt{ClonalOrigin}, and to validate some of the
+Single Breakpoint Recombination (\texttt{SBP}) method from the \texttt{HyPhy}
+package \citep{KosakovskyPond2006}.  Using the Akaike Information Criterion
+(AIC), \texttt{SBP} inferred that 47.5\% (1,094) of gene families contained at
+least one topology-changing recombination. Each of these families was then split
+into two gene fragments.  Gene trees were constructed from nucleotide alignments
+for each gene fragment using the \texttt{RAxML} program \citep{Stamatakis2006},
+and were rooted by minimizing inferred duplications and losses.  Each rooted
+gene tree was then reconciled using the \texttt{Mowgli} program
+\citep{Doyon2011}, which inferred the most parsimonious gene duplication, loss,
+and transfer events for each family.  In the case of multiple maximum
+parsimonious reconciliations, one was chosen uniformly at random.  Unless
+specified otherwise, default setings were used for all the programs.  Since the
+method of \texttt{Mowgli} did not directly model recombination events,
+recombination-based transfers were inferred in a post-processing step, by
+checking whether the transfer was coupled with a loss of a homologous gene in
+the recipient branch (indicating gene replacement rather than gain).  A transfer
+event was classified as non-recombining if the destination species of that
+transfer had an extant descendant that contained copies of that genes which were
+descendants of the transferred gene as well as copies which were
+non-descendants.  We noted that this parsimonious approach was imperfect and
+would tend to undercount events, especially recombining gene transfer events,
+which were modeled using two events: a transfer and a loss.  Nevertheless, this
+analysis allowed us to study non-recombining gene transfers, which were not
+modeled by \texttt{ClonalOrigin}, and to validate some of the
 \texttt{ClonalOrigin} results on recombining gene transfer.
@@
 
-% FIXME
-% \texttt{The recombination intensity scores were taken to be average site-wise
-% intensity scores. SC: THIS NEEDS TO BE REANALYZED WITH RECOMBINATION INTENSITY
-% WITH ONLY TREE-TOPOLOGY CHANGING RECOMBINANT EDGES.} 
 We performed a series of analyses to check for functional gene categories
 enriched for certain types of gene transfer events. Functional categories were
-assigned in the same manner as \citet{Suzuki2011}. In short, blastP was used to
+assigned in the same manner as \citet{Suzuki2011}. In short, \texttt{blastP} 
+was used to
 compare Streptococcus genes to bacterial proteins from the Uniref90 database.
@@
 We conducted experiments on simulation data to check the accuracy of estimates
-obtained by \texttt{ClonalOrigin} in a scenario similar to ours. We used a
+obtained by \texttt{ClonalOrigin} in a scenario similar to one from the real 
+data set.  We used a
 simulation feature of \texttt{ClonalOrigin} to simulate the process of mutation
-and recombination across the five species.  While populaiton mutation and
+and recombination across the five species 
+(see section \textbf{Supplementary Information} for detail).
+While populaiton mutation and
 recombination parameters were accurately estimated, the recombinant tract length
@@
 estimate from relatively short alignment blocks \citep{Didelot2010}. 
-% When we
-% performed simulations which much longer blocks (10000 bp), \texttt{ClonalOrigin}
-% was able to estimate the average recombinant tract length accurately (data not
-% shown).  
-% FIXME: Show this result of 10,000 blocks.
 We also evaluated the accuracy of the statistics estimated in the second stage
@@
 
-\clearpage{}
-
 \section{Supplementary Information}
+
+\subsection{Local gene trees and recombination intensity}
+
+Because the procedure of summarizing posterior samples of \textit{recombinant
+tree}, which will be defined in a moment, 
+was not detailed in the user guide at the time when we used
+\texttt{ClonalOrigin}, we elected to use Figure S\ref{fig:clonalorigin} to
+illustrate a recombinant tree in order to clarify summarization of
+\textit{recombinant history} posterior samples.  A recombination history is
+represented by a \textit{recombinant tree}, which is a species tree augmented
+with \textit{recombinant edges}.  A recombinant edge connect two species tree
+branches.  Each recombinant edge is defined by the time
+and branch of origin and destination, and the genomic interval that it affects.
+The branch of origin is referred to as the \textit{donor} and the destination
+branch is referred to as the \textit{recipient}. The type of a recombinant edge
+is determined by the donor and recipient species tree branches.  A recombinant
+tree can be viewed as a restricted version of an ancestral recombination graph
+(ARG), and as such it induces a local genealogy for each site in the genomic
+block it describes.  \citet{Didelot2010} presented two procedures of summarizing
+recombinant trees by viewing trees from two different angles: by counting types
+of recombinant edges regardless of genomic positions affected by the edges, and
+by measuring recombinant edge changes to the species tree in site-wise manner.
+By similarly following their approaches we quantified additional aspects of
+recombinant trees; we counted gene tree topologies for each site, and also
+quantified recombination intensity along the genome.
+
+The recombination intensity allowed us to find genomic regions enriched for
+recombination events in general, or regions enriched for recombination events of
+certain types (e.g., between SPY and SDE).  We noted that the recombination
+intensity was not normalized to the interval of $[0,1]$. Rather than giving
+probabilistic meanings to the recombination intensity we interpreted it as an
+average number of recombinant edge types per site.  The measure appeared to act
+as an effective indicator of recombination intensity (see subsection of
+\textbf{Simulation Study}).  Our simulation study indicated that edges of
+certain types were better indicators of recombination events than others.
+For instance, recombinant edges which resulted in a recombinant tree with the
+same topology as the species tree appeared to be more prone to spurious
+sampling.  Such edges were ones where the donor was the same branch, sister
+branch, or parent branch of the recipient branch.  Therefore, in some cases we
+elected to remove such edges from consideration when measuring recombination
+intensity.  
+% The tree topologies changes would be correlated with recombination
+% intensity.  The tree topologies would show more effectively regions where
+% recombination occurs along the genome (see subsection of \textbf{UCSC Genome
+% Browser}).
 
@@
 % FIXME: Newick tree in text in a file.
+% FIXME: Publish the result and the data set with the code.
 We describe the detailed procedure of simulations for checking if the three
@@
 parameters and the species tree  set to the values inferred from the five
-bacterial genomes (see section \textbf{Results}).  Each of the recombinant trees
-was used to simulate the Jukes-Cantor model for DNA sequence evolution to create
-274 blocks of alignments.  The block lengths configuration was the same as that
-of the blocks used in the analysis of the five genomes.  The mean of the ten
-estimates of the recombinant tract length was significantly higher than the true
-value of (919 bp versus 744 bp) whereas the values of the mutation and
-recombination rates were accurately estimated (Table
-S\ref{tab:sim-three-population-parameters}).  Recombinant tract lengths were
-known to be difficult to estimate from relatively short alignment blocks
-\citep{Didelot2010}. Performing similar simulations with much longer blocks
-(i.e., 10000 bp) we estimated the average recombinant tract length accurately
-\texttt{(Table S data not shown)}.
+bacterial genomes.  Each of the recombinant trees was used to simulate the
+Jukes-Cantor model for DNA sequence evolution to create 274 blocks of
+alignments.  The block lengths configuration was the same as that of the blocks
+used in the analysis of the five genomes.  The mean of the ten estimates of the
+recombinant tract length was significantly higher than the true value of (919 bp
+versus 744 bp) whereas the values of the mutation and recombination rates were
+accurately estimated (Table S\ref{tab:sim-three-population-parameters}).
+Recombinant tract lengths were known to be difficult to estimate from relatively
+short alignment blocks \citep{Didelot2010}.
 
@@
 % ARTIFICIALLY DOUBLE OR TRIPLE THE RECOMBINATION EVENTS OF ONE OF TYPES.
-We checked if numbers of recombinant edge types were inferred correctly. Because
-of the two-stage Bayesian method we first sampled ten recombinant trees given
-the mutation rate, recombination rate, and tract length (see section
-\textbf{Results}), and then created ten independent simulated data sets with the
-Jukes-Cantor model for DNA sequence evolution using each of the recombinant
-trees. This consisted of 100 replicates of data sets.  Figure S\ref{fig:h1}
-shows the observed count of each recombinant edge type against the expected
-count; the estimated numbers appeared to be slightly over-estimated, but in
-correlation with the true numbers.
+Because of the two-stage Bayesian method we first sampled ten recombinant trees
+given the mutation rate, recombination rate, and tract length, and then created
+ten independent simulated data sets with the Jukes-Cantor model for DNA sequence
+evolution using each of the recombinant trees. This consisted of 100 replicates
+of data sets.  Figure S\ref{fig:h1} shows the observed count of each recombinant
+edge type against the expected count; the estimated numbers appeared to be
+slightly over-estimated, but in correlation with the true numbers.
 
@@
 
-\subsection{}
-
-With the slight over-estimation of recombination events we were concerned with
-any potential biases hidden in \texttt{ClonalOrigin} implementation. Because the
-simulated data were generated with no directionality between SPY and SDE,  if we
-estimate recombination events from the simulated data sets using
-\texttt{ClonalOrigin} and compare those estimates from those from the real data
-set, this could remove biases in \texttt{ClonalOrigin} implementation. We could
-use inferred estimates of recombination from the simulated data under the prior
-model of \texttt{ClonalOrigin} as the prior expected number of recombination.
-Figure S\ref{fig:h2} shows a plot of numbers of recombination events from the
-real data against those from the simulated data sets. Note that the numbers of
-recombination events from simulated data sets appear to deviate from the line of
-slope being 1 and intercept being 0.  The numbers of recombination events for
-the recombinant edges from SPY to SDE increase by more than 2.5 fold whereas
-those for the recombinant edges from SDE to SPY increase by about 1.7 fold.
-This was not surprising because inference of number of recombinant edges was
-quite accurate. We considered that subtraction of effects of number of
-recombinant edges inferred from simulated data and subtraction of effects of the
-prior expected number of recombinant edges were effectively equivalent.
-However, this comparison between inferrence from simulated data and that of real
-data set suggested that the directional difference between SPY and SDE was due
-to neither lack of power of \texttt{ClonalOrigin} nor hidden bias thereof.
-
 \subsection{Simulation for recombination intensity}
@@
 We tested if our measure of recombination intensity provided a reliable estimate
-of the true intensity. For this purpose, we artificially annotaded our synthetic
+of the true intensity. For this purpose, we artificially annotated our synthetic
 blocks to genes according to the real annotations along the real bacterial
@@
 average intensity, i.e., the average number of different recombinant edge types
-implied by the recombinant tree used in simulation (Figure S\ref{fig:ri1}).
-Although the scatter plot did not strictly follow along the $X=Y$ line, there
-was correlation of the estimated and true values (Pearson's correlation
-coefficient of 0.63). Estimated values appeared to be driven somewhat by the
-prior implied by the population parameter setting, explaining why lower values
-were over-estimated and higher values were under-estimated. Nonetheless, the
-correlation indicated that the measure of recombination intensity used in our
-functional gene category served as an adequate indicator of actual intensity of
-recombination events along genes.
+implied by the recombinant tree used in simulation 
+(Pearson's correlation coefficient of 0.63 and Figure S\ref{fig:ri1}).
 
@@
 %=============================================================================
-% Figure S: 
+% Figure S12: 
+\begin{figure}
+\subfloat[]{\includegraphics[scale=0.5]{figures/h3-R-ratio-topology-changing}}
+\subfloat[]{\includegraphics[scale=0.5]{figures/h3-R-ratio-not-topology-changing}}
+\caption{\label{fig:h3}Numer of recombination events. The diagonal dashed
+lines are with slope of 1 and intercept of 0. For each order
+pairs of branches of a reference species tree the true number of recombination
+events is computed as the mean of the 10 replicates of recombinant
+trees from the simulation. With ten estimates of each replicate total
+number of 100 estimats are shown as intervals of 90\% of quantile
+range, or the lower value of the interval is the 5\% quantile, and
+the upper one is the 95\% quantile. }
+\end{figure}
+\clearpage{}%
+%=============================================================================
+
+%=============================================================================
+% Figure S12: 
 \begin{figure}
@@
 %=============================================================================
-% Figure S:
+% Figure S13:
 \begin{figure}
@@
 % Figure S:
-\begin{figure}
-\subfloat[]{\includegraphics[scale=0.5]{figures/h2-R}}
-\subfloat[]{\includegraphics[scale=0.5]{figures/h2-R-inset}}
-
-\subfloat[]{\includegraphics[scale=0.5]{figures/h2-R-inset2}}
-
-\caption{\label{fig:h2}Comparison of number of recombinations between real
-and simulated data sets. The diagonal dashed lines are with
-slope of 1 and intercept of 0. For each order pairs of branches of
-a reference species tree the number of recombination events is estimated
-from the real data set. Total number of 100 estimats from a simulated
-data set are shown as intervals of 90\% of quantile range, or the
-lower value of the interval is the 5\% quantile, and the upper one
-is the 95\% quantile. }
-\end{figure}
-\clearpage{}%
-%=============================================================================
-
-\clearpage
+%\begin{figure}
+%\subfloat[]{\includegraphics[scale=0.5]{figures/h2-R}}
+%\subfloat[]{\includegraphics[scale=0.5]{figures/h2-R-inset}}
+
+%\subfloat[]{\includegraphics[scale=0.5]{figures/h2-R-inset2}}
+
+%\caption{\label{fig:h2}Comparison of number of recombinations between real
+%and simulated data sets. The diagonal dashed lines are with
+%slope of 1 and intercept of 0. For each order pairs of branches of
+%a reference species tree the number of recombination events is estimated
+%from the real data set. Total number of 100 estimats from a simulated
+%data set are shown as intervals of 90\% of quantile range, or the
+%lower value of the interval is the 5\% quantile, and the upper one
+%is the 95\% quantile. }
+%\end{figure}
+%\clearpage{}%
+%=============================================================================
 
@@
 % These topology-not-changing recombinant edges tend change only branch lengths.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Should we keep this section?}
+
+With the slight over-estimation of recombination events we were concerned with
+any potential biases hidden in \texttt{ClonalOrigin} implementation. Because the
+simulated data were generated with no directionality between SPY and SDE,  if we
+estimate recombination events from the simulated data sets using
+\texttt{ClonalOrigin} and compare those estimates from those from the real data
+set, this could remove biases in \texttt{ClonalOrigin} implementation. We could
+use inferred estimates of recombination from the simulated data under the prior
+model of \texttt{ClonalOrigin} as the prior expected number of recombination.
+Figure S\ref{fig:h2} shows a plot of numbers of recombination events from the
+real data against those from the simulated data sets. Note that the numbers of
+recombination events from simulated data sets appear to deviate from the line of
+slope being 1 and intercept being 0.  The numbers of recombination events for
+the recombinant edges from SPY to SDE increase by more than 2.5 fold whereas
+those for the recombinant edges from SDE to SPY increase by about 1.7 fold.
+This was not surprising because inference of number of recombinant edges was
+quite accurate. We considered that subtraction of effects of number of
+recombinant edges inferred from simulated data and subtraction of effects of the
+prior expected number of recombinant edges were effectively equivalent.
+However, this comparison between inferrence from simulated data and that of real
+data set suggested that the directional difference between SPY and SDE was due
+to neither lack of power of \texttt{ClonalOrigin} nor hidden bias thereof.

### 2011-10-13 doc/manuscript.tex
Functional categories are being discussed in the manuscript.

@@
 processes. We attempted here to use and compare two approaches for
-distinguishing recombining gene transfer process from horizontal, or
+distinguishing the recombining gene transfer process from the horizontal, or
 non-recombining, gene transfer, and tackled the two processes in association with
@@
 approach could be used to infer transfer of genomic segments due to homologous
-recombination, and the second could help similarly attributed to horizontal gene
+recombination, and the second could be useful in studying horizontal gene
 transfer.
@@
 recombination, and the parsimony-based one for detecting gene duplication, loss,
-and horizontal transfer. We compared both of the results in recombining gene
-transfer events to show that the parimony-based method inferred recombining gene
-transfer events in biological functional categories similar to those inferred
-with the model-based approach. This was the first attempt as far as we know to
-analyze SPY and SDE in genome-wide scale for detecting gene transfer events not
-only in the core genome but also in the pan genome, which made the analysis
-complete in terms of the amount of used genome data. Our analysis of core
-genomes also showed for the first time that the net directionality of
-recombining gene transfer from SPY to SDE was stronger than the reverse
-direction.
+and horizontal transfer. Virulence genes were found to be more associated
+non-recombining horizontal gene transfer than non-virulence genes.  We compared
+both of the results in recombining gene transfer events to show that the
+parimony-based method inferred recombining gene transfer events in biological
+functional categories similar to those inferred with the model-based approach.
+This was the first attempt as far as we know of systematically analyzing SPY and
+SDE in genome-wide scale to detect gene transfer events not only in the core
+genome but also in the pan genome, which made the analysis complete in terms of
+the amount of used genome data. Our analysis of core genomes also showed for the
+first time that the net directionality of recombining gene transfer from SPY to
+SDE was stronger than the reverse direction.
 
@@
 distinct genome structures of SPY and SDE (Figure S\ref{fig:mauve}). The SDE and
-SDD genomes were larger in length than those of SPY. Although it was unclear
-whether SDD was different from either of SDE1 and SDE2 from the graphical view
-of the genome alignment, a spcies tree inferred from the genome alignment
-showed that SDD branch was attached to the clade branch of SDE1 and SDE2 (Figure
-S\ref{fig:tree5}). The distribution of the alignment block lengths was skewed to
-the right with mean of 4188 base pairs long (Figure S\ref{fig:blocksize}), which
-supported that the model assumption of geometric block length distribution might
-be plausible in \texttt{ClonalOrigin} method (see section \textbf{Materials and
-Methods} and \citealt{Didelot2010}).  The relative strength ($r/m$) of
-recombination compared to mutations indicated that recombination-driven
-substitutions affected more sites than did mutation-driven substitutions (Table
+SDD genomes were larger in length than those of SPY. Although the graphical view
+of the genome alignment did not clearly show whether SDD was different from
+either of SDE1 and SDE2 in the genome organization, a spcies tree inferred from
+the genome alignment showed that the external SDD branch was attached to the
+internal clade branch of SDE1 and SDE2 (Figure S\ref{fig:tree5}). The
+distribution of the alignment block lengths was skewed to the right with mean of
+4188 base pairs long (Figure S\ref{fig:blocksize}), which supported that the
+model assumption of geometric block length distribution might be plausible in
+\texttt{ClonalOrigin} method (see section \textbf{Materials and Methods} and
+\citealt{Didelot2010}).  The relative strength ($r/m$) of recombination compared
+to mutations indicated that recombination-driven substitutions affected more
+sites of the core genomes than did mutation-driven substitutions (Table
 S\ref{tab:clonalframe}). This could imply that in the \textit{Streptococcus}
@@
 
-\subsubsection{Recombination hotspots were observed in genomes}
-
-Along the genome of SDE1 population mutation rate did not appear to change much
+\subsection{Recombination hotspots were observed in genomes}
+
+Population mutation rate did not appear to change much across the genome of SDE1 
 (Figure S\ref{fig:scatter3}).  We found a few spots of genomic regions where the
@@
 
-\subsection{SPY and SDE were more closely related in about tenth of the genome}
-
-There are 105 possible rooted topologies relating five taxa.  We expected the
+\subsection{SPY and SDE were more closely related in about tenth of the core genome}
+
+There are 105 possible rooted labeled topologies relating five taxa.  We expected the
 dominant topology across sampled genealogies to be concordant with that of the
@@
 S\ref{tab:heatmap} and S\ref{tab:obsheatmap} for the heat map values and the
-observed number of recombinant edges) shows the log base 2 ratios of the
+observed number of recombinant edges) shows the logarithm base 2 ratios of the
 \textit{a posteriori} number of recombinant edges for a pair of donor and
@@
 the column for the root as recipient was colored green, indicating invalid
-recombinant edges.  Similarly, all of the cells for which a recipient branch was
+recombinant edges in Figure \ref{fig:Heatmap-of-recombination}.  
+Similarly, all of the cells for which a recipient branch was
 older than the donor, and the donor and recipient branches did not share any contemporary time were
@@
 number of recombinant edges, because the observed and expected numbers of
-recombinant edges were similar. As recombinant edges starting from and ending in
+recombinant edges were similar. As recombinant edges that start from and end in
 the same species tree branch would not change the species tree topology, which
@@
 strictly veterinary pathogenic bacteria, we did not expect that SDD received
-much gene transfer from either SPY1 or SPY2: the log base 2 ratio from SPY1 to
+much gene transfer from either SPY1 or SPY2: the logarithm base 2 ratio from SPY1 to
 SDD was -2.20, and that from SPY2 to SDD was -1.97.  On the contrary, the
@@
 individuals to SDD compared with that from two SPY individuals to SDD reflected
+the fact
 that the SDD was more closely related to SDE than to SPY in most of the genomes,
@@
 be more frequent than that between clades, not surprisingly the recombining gene
-transfer log base 2 ratios between SDE1 and SDE2 (1.17 and 1.20) were about
-twice as large as that from SDE1 and SDE2 to SPY1 and SPY2 (0.67, 0.66, 0.60,
+transfer logarithm base 2 ratios between SDE1 and SDE2 (1.17 and 1.20) were about
+twice as large as those from SDE1 and SDE2 to SPY1 and SPY2 (0.67, 0.66, 0.60,
 and 0.39).  The intensities of recombining gene transfer from external branches
 of SDE1 and SDE2 to external branches of SPY1 and SPY2 was similar to those from
-SDE1 or SDE2 to SDD (0.76 and 0.70).  After all, either SPY1 or SPY2, and SDD as
+SDE1 or SDE2 to SDD (0.76 and 0.70).  Therefore, either SPY1 or SPY2, and SDD as
 recipient appeared to be no different from the view point of SDE1 or SDE2 as
@@
 or SDE2 to one of the other three external branches. Surprisingly, the recombining
-gene transfer log base 2 ratios between SPY1 and SPY2 (1.06 and 1.07) was
+gene transfer logarithm base 2 ratios between SPY1 and SPY2 (1.06 and 1.07) was
 comparable with, yet even less than those from SPY1 and SPY2 to SDE1 and SDE2
-(1.25, 1.25, 1.17, and 1.29).  Effectively, the recombining gene transfer from a
-clade SPY1 or SPY2 to a different clade SDE1 or SDE2 seemed to be no different
+(1.25, 1.25, 1.17, and 1.29).  The recombining gene transfer from a clade SPY1 
+or SPY2 to a different clade SDE1 or SDE2 seemed to be effectively no different
 from that within the clade of SPY1 and SPY2.  The two external branches of SDE
@@
 concordant with the fact that SPY and SDE shared the same ecological niche
-despite their dissimilar genomic organization.  The gene transfer log base 2
+despite their dissimilar genomic organization.  The gene transfer logarithm base 2
 ratio from SPY1 and SPY2 to SDE branch was also larger than any type of
@@
 specific direction, as opposed to the result for recombining transfers from the
-model-based analysis (Figure S\ref{fig:mowgli-recomb-heatmap}).  SDE1 and SDE2
-shared the most events (26 and 34) compared to other sample pairs.  SDD appeared
-to acquire many genes from the ancestral SDE branch.  Transfer events between
-the SPY and SDE clades were inferred in smaller numbers (<6) compared to those
-within each clade. We were puzzled by this finding because virulence
-genes were found to be associated with horizontal gene transfer, and yet we could
-not show net directionality of horizontal gene transfer between SPY and SDE.
-
-\subsection{More frequent recombining gene transfer than horizontal gene transfer}
-
-Overall, we inferred 304 duplications, 944 recombining gene transfers, and 294
-horizontal gene transfers, comprising 23.7\% of total transfer events (Figure
-\ref{fig:Gene-duplication-loss}).  Among the 2314 gene families the largest
-family consisted of 59 genes, while nearly half of gene families (1066) had
-exactly one gene per species, or six genes total (Figure S\ref{fig:famsizes}).
-Within clades we found that duplications made up only a small fraction of the
-events  (10.2\%), while horizontal transfers and losses were significantly more
-frequent (41.7\% and 48.0\%, respectively). Duplications were most often
-species-specific.  The species SDD and SDE were the most expanded, both in terms
-of duplications and in-bound transfers.
+model-based analysis.  SDE1 and SDE2 shared the most events (26 and 34) compared
+to other sample pairs.  SDD appeared to acquire many genes from the ancestral
+SDE branch.  Transfer events between the SPY and SDE clades were inferred in
+smaller numbers (<6) compared to those within each clade. We were puzzled by
+this finding because virulence genes were found to be associated with horizontal
+gene transfer, and yet we could not show net directionality of horizontal gene
+transfer between SPY and SDE.  However, we inferred relatively more recombining
+gene transfer events from the SPY branch to SDE branch even in the
+parsimony-based approach (Figure S\ref{fig:mowgli-recomb-heatmap}).
+
+\subsection{Reduction of genes and divergence of SPY and SDE species}
+
+The parsimony-based approach estimated 304 duplications, 944 recombining gene
+transfers, and 294 horizontal gene transfers, comprising 23.7\% of total
+transfer events (Figure \ref{fig:Gene-duplication-loss}).  Among the 2314 gene
+families the largest family consisted of 59 genes, while nearly half of gene
+families (1066) had exactly one gene per species, or six genes total (Figure
+S\ref{fig:famsizes}).  Numbers of genes at the five ancestors were inferred to
+be smaller than those at the extant species.  SDE lineages from the root
+appeared to have increased in the number of genes while the common ancestor of
+SPY1 and SPY2 decreased in the number of genes.
+
+The four external lineages of SDE1, SDE2, SDD, and SPY1 had almost equal numbers
+of gene gain and loss events while the four lineages of SP, SD, SPY2, and SEE
+had larger numbers of gene gain events than those of gene loss events. In the
+two lineages of SP and SEE we might have inferred gene gain and loss events less
+confidently than events in other lineages because of lack of outgroup lineages
+to compare with the two lineages. The number of gene gain events in SPY2 lineage
+seemed to be large compared with those of gene gain events in the other
+lineages. On the contrary, two internal branches of SPY and SDE had smaller
+numbers of gene gain events than those of gene loss events.  This might indicate
+that SPY and SDE lineages had reduced number of genes to specialize in their
+specific niches after each split from their corresponding direct ancestors.
+
+Duplications in extant lineages were more frequent than those in the other
+lineages.  Although the number of duplication events along SPY2 branch seemed
+smaller than those of the other extant lineages, duplication events appeared to
+be most often species-specific.  Within clades we found that duplications made
+up only a small fraction of the events  (10.2\%), while horizontal transfers and
+losses were significantly more frequent (41.7\% and 48.0\%, respectively).  The
+extant species of SDD and two SDE's were the most expanded, both in terms of
+duplications and in-bound transfers. This was apparently related with the larger
+genome sizes of the two species than that of SPY.  
+The numbers of recombining gene transfer events tended to be larger than those
+of horizontal gene transfer events except SD lineage.  Numbers of recombining
+gene transfer in the extant lineages might have been larger than those in the
+internal branches because the number of recombining gene transfer events must
+had been affected by the availability of more closely related individual
+genomes.
 
@@
 About 10\% of the genes comprised of virulence facters, about half of which were
-found in the core genome of the five genome alignment. Using the UCSC genome
-browser tracks we located virulence genes with notable posterior recombination
-probability.  These genes were by no means comprehensive list of virulence
-factors that might have been transferred via homologous recombination.
-Stressful host conditions for pathogeneic bacteria can include heat shock,
-starvation, and iron depletion.  Iron uptake of bacteria was known to be
-essential for survival of the bacteria under hostile host condition. We found
+found in the core genome of the five genome alignment \citep{Suzuki2011}.  Using
+customized UCSC genome browser tracks available at
+\url{http://strep-genome.bscb.cornell.edu} we located virulence genes with
+notable posterior recombination probability.  These genes were by no means a
+comprehensive list of virulence factors that might have been transferred via
+homologous recombination.  We wished to showcase virulence genes using the
+result of our strudy of gene transfer.
+
+Stressful host conditions for pathogenic bacteria can include heat shock,
+starvation, and iron depletion, etc.  Iron uptake of bacteria was known to be
+essential for survival of bacteria under the hostile host condition. We found
 that parts of putative calcium transporter (MgtA) in SDE1 could have been
 transferred from SPY1 (SpyM3\_0440).  An extracellular protein, putative
-dipeptidase (SpyM3\_0465) that could be important for the survival of bacteria
-to scavenge dipeptides from the host, was inferred to be transferred from SPY to
-SDE: the whole gene was transferred, which was a case for the finding that
-virulence genes tended to be transferred as a whole unit \citep{Chan2009}.
-Another example of whole gene transfer was superoxide dismutase that was
-reported to be associated virulence of \textit{Streptococcus agalactiae}
+dipeptidase (SpyM3\_0465) that could be important for bacteria to scavenge
+dipeptides from their host, was inferred to be transferred from SPY to SDE: the
+whole gene was transferred, which was a case for the finding that virulence
+genes tended to be transferred as a whole unit \citep{Chan2009}.  Another
+example of whole gene transfer was superoxide dismutase that was reported to be
+associated with virulence of \textit{Streptococcus agalactiae}
 \citep{Poyart2001}: the whole gene including its 5' UTR, (SpyM3\_1071), was
@@
 
-\subsection{Virulence genes via horizontal gene transfer}
+\subsection{Association of virulence genes and horizontal gene transfer}
 
@@
 gene transfer served as a more effective means by which SDE or SPY acquired
-virulence genes, compared to recombining gene transfer.  
-Families with high numbers of duplications were also significantly enriched for
-virulence genes (P-value of $<7.13 \times 10^{-3}$) whereas families with high
-losses showed no enrichment (P-value of $>0.54$).
-
-Roughly, half of the virulence genes were in the pan genome.  We did not attempt
-to comprehensively list virulence genes that were inferred to be transferred
-horizontally. We rather sort virulence genes by the number of horizontal gene
-transfer events to list the first several genes with high ranks.  Streptococcal
-C5a peptidase (\textit{scpB}, SDEG\_0933) was inferred to have six horizontal
-gene transfer.  Putative transposase were found to have a relatively large
-number of horizontal gene transfer events (SpyM3\_1016, SpyM3\_1053,
-SpyM3\_1054). Other genes with relatively large number of horizontal gene
-transfer events included putative response regulator (SpyM3\_0174), putative
-maltose operon transcriptional repressor (SpyM3\_0982), putative histidine
-kinase (SpyM3\_0171), and putative cell wall hydrolase (SpyM3\_0731,
-SpyM3\_0922, SpyM3\_1208).  Virulence genes encoding putative nucleases with a
-secretion signal peptide were also found to have had experienced horizontal gene
-transfer: genome-encoded streptodornase (SDEG\_0541), and prophage-associated
-deoxyribonuclease (SDEG\_1103).  Protein F2 like fibronectin-binding protein
-(SpyM3\_0104), an adhesin and invasin, was inferred to horizontally
-transferred.
-
-\subsection{Comparison between model-based and parsimony-based approaches}
-
-Both of the model-based and parimony-based methods used discrepancy between a
-reference species tree and gene trees to detect recombining gene transfer. In
-essence the two methods worked in relatively similar manner. Firstly, functional
-categories that were associated with recombining gene transfer were similar in
-both methods (Tables S\ref{tab:functional-all} and
-S\ref{tab:go-events-recombining}).  Secondly, both of them did not show
-association of virulence genes with recombining gene transfer. Lastly, we found
-a positive correlation of 0.39 with P-value of $2.2\times10^{-16}$ between the
-recombination intensity and the number of recombining gene transfer for the
-genes (Figure S\ref{fig:cmpcomowgli}).  The correlation (0.03 with P-value of
-0.18) disappeared in the comparison between recombination intensity and number
-of horizontal gene transfer events (Figure S\ref{fig:cmpcomowglihgt}).
-
-\subsection{UCSC genome browser for bacterial gene transfer study}
-
-We created a web-resource (Figure S\ref{fig:ucsc}) for gene transfer available at
+virulence genes, compared to recombining gene transfer.  Families with high
+numbers of duplications were also significantly enriched for virulence genes
+(P-value of $<7.13 \times 10^{-3}$) whereas families with high losses showed no
+enrichment (P-value of $>0.54$).
+
+Roughly, half of the virulence genes were in the pan genome as mentioned above.
+We did not attempt to comprehensively list virulence genes that were inferred to
+be transferred horizontally or non-recombining fashion. We rather sort virulence
+genes by the number of horizontal gene transfer events to list the first several
+genes with high ranks.  Streptococcal C5a peptidase (\textit{scpB}, SDEG\_0933)
+was inferred to have six horizontal gene transfer.  Putative transposase were
+found to have a relatively large number of horizontal gene transfer events
+(SpyM3\_1016, SpyM3\_1053, SpyM3\_1054). Other genes with relatively large
+number of horizontal gene transfer events included putative response regulator
+(SpyM3\_0174), putative maltose operon transcriptional repressor (SpyM3\_0982),
+putative histidine kinase (SpyM3\_0171), and putative cell wall hydrolase
+(SpyM3\_0731, SpyM3\_0922, SpyM3\_1208).  Virulence genes encoding putative
+nucleases with a secretion signal peptide were also found to have had
+experienced horizontal gene transfer: genome-encoded streptodornase
+(SDEG\_0541), and prophage-associated deoxyribonuclease (SDEG\_1103).  Protein
+F2 like fibronectin-binding protein (SpyM3\_0104), an adhesin and invasin, was
+inferred to horizontally transferred.
+
+\subsection{Model-based approach or parsimony-based one for recombining gene
+transfer study}
+
+A full model-based approach would be desirable in analyzing data for testing
+relevant hypotheses more systematically. Model complexity might not allow easy
+access to the implementation of a model. Our study involved a model-based
+approach and a parsimony-based one for understanding evolution in the core
+genomes of several bacteria individuals. Because of inability of the model-based
+method for studying the pan genomes we had to resort to the parsimony-based
+method. However, the parsimony-based method was also applied to the core
+genomes. We wished to compare the two methods in their results of the core
+genome analyses.  Because of two independent analyses on the same data set, this
+would allow us to be more confident of the results if the two analyses performed
+similarly Generally, phylogenetic methods of detecting gene transfers often take
+advantage of discrepancy between a reference species tree and gene trees. The
+two approaches that we employed in this study were no exception. We found a few
+similar results in both methods; firstly, functional categories that were
+associated with recombining gene transfer were similar (Tables
+S\ref{tab:functional-all} and S\ref{tab:go-events-recombining}).  Secondly,
+either of them did not show association of virulence genes with recombining gene
+transfer.  To more directly compare the methods we used the recombination
+intensity and the number of recombining gene transfer events for each gene.  We
+found a weak positive correlation of 0.39 with a very small significant P-value
+of $2.2\times10^{-16}$ for testing a null hypothesis of the correlation being
+zero.  (Figure S\ref{fig:cmpcomowgli}).  Interestingly, the correlation (0.03
+with P-value of 0.18) disappeared in a comparison between recombination
+intensity and number of horizontal gene transfer events (Figure
+S\ref{fig:cmpcomowglihgt}).
+
+\subsection{UCSC genome browser tracks for bacterial gene transfer study}
+
+Often computational and statistical analyses on genomes could benefit other
+researchers when those resources were provided: e.g., UCSC Genome Browser, and
+Ensembl Genome Browser. We wished to provide research community especially in
+SPY and SDE bacteria with some of the results from our analysis of gene
+transfer. We used the UCSC genome browser for the purpose.  The resource shown
+in Figure S\ref{fig:ucsc} for gene transfer is available at
 \url{http://strep-genome.bscb.cornell.edu}.  For each of the five individual
@@
 topologies, recombination intensity, locations of virulence genes, posterior
-probability of recombining gene transfer.  
+probability of recombining gene transfer. The genome track of probability of
+recombining gene transfer allowed us to easily browse the bacterial genome to
+locate regions that we might be interested in. 
 % To Matt: what could we show in the pan-genome? or from Mowgli's analysis?
@@
 another \citep[e.g.,][]{Chan2009}. We observed that units of recombining gene
-transfer could happen with genes (Table S\ref{tab:genes-transfer}). In some
+transfer could be multiple genes (Table S\ref{tab:genes-transfer}). In some
 cases, parts of a gene were inferred to have been transferred, for instance,
 transfer from SPY1 to SDE1 was inferred for 29\% of the aspartyl-tRNA synthetase
-gene.  In other cases, the inferred transfer units covered a set of as much as
-eight adjacent genes.  Of the genes listed in Table S\ref{tab:genes-transfer}
-there were 12 groups of genes that recombinant tracts covered their gene
-boudaries.  Most of the recombinant tracts often covered two genes. There were
-transfer fragments that covered eight genes together: loci from SpyM3\_1518 to
-SpyM3\_1525.
-
-\subsection{Transporter, transferase, and aminoacyl tRNA synthetases}
-
-Two classes of functional categories were notably to be inferred to associated
-with recombining gene transfer.  Firstly, genes involved in sensing queues of
-environmental changes would be important for bacteria to adapt to changing
-environment.  Genes encoding membrane-bound proteins would be involved in the
-function.  Several of functional categories that were associated with gene
-transfer might be related with membrane-bound protein coding genes (Table
-\ref{tab:functional}).  Sugar phosphotransferase classes were ranked top in the
-list: phosphrylation could be important in signal transduction to recognize
-environmental changes.  The transport functional categories of carbohydrate
-transport and sodium ion transport also belonged to genes encoding
-membrane-bound proteins.  Also highly ranked were fatty acid and lipid
-biosyntheses that must be involved in producing membrane-bound protein
-molecules. These functional categories related with membrane-bound proteins
-might allow for bacteria to adapt to changing environment via recombining gene
-transfer.  Secondly, aminoacyl tRNA synthetases had been a focus of gene
-transfer \citep{Woese2000}, and these aminoacyl tRNA synthetases belonged to the
-functional category of translation (GO:0006412).  They were inferred to be
-associated with recombining gene transfer not horizontal gene transfer because
-translation category was siginificant only in recombining gene transfer (see
-Tables \ref{tab:functional}, \ref{tab:go-events} and
-S\ref{tab:go-events-recombining}).
-
-In addition to the two kinds of functional categories, functional categories
-concerning ribosomes were also associated with higher recombining gene transfer.
-Because genes related with ribosomes were considered to be housekeeping genes,
-gene transfer in ribosome genes would be rare, it was unexpected that ribosome
-related genes were found to be associated with recombining gene transfer. This
-was more significant when we considered all of the recombinant edges (Table
-S\ref{tab:functional-all}) not just topology-changing recombinant edges (Table
-\ref{tab:functional}) in computing recombination intensity.  Therefore, high
-recombination intensity in ribosome related genes would be likely attributed
-to topology-not-changing recombinant edges (Table
-S\ref{tab:functional-notopology}).  The parsimony-based approach also showed
+gene. Recombinant tracts often were shown to span two genes.  In other cases,
+the inferred transfer units spaned a set of as much as eight adjacent genes:
+loci from SpyM3\_1518 to SpyM3\_1525.
+
+\subsection{Transporters, transferases, aminoacyl tRNA synthetases, and
+transposases}
+
+Molecular evolutionary forces could act on genetic materials of organisms, and
+changes to the genotypes might lead to some phenotypes that could change
+biological functions of the species. So, the bacterial gene transfer process as
+an evolutionary force could be associated with some biological functions. We
+attempted to find which biological functions were associated with recombining or
+horizontal gene transfers.  Several functional categories that were found to be
+associated with recombining gene transfer were related with either
+membrane-bound molecules or aminoacyl tRNA synthetases (Table
+\ref{tab:functional}).  However, functional categories associated with
+horizontal gene transfer were different from those from recombining gene
+transfer analysis (Table \ref{tab:go-events}).  The top ranked functional
+category associated with horizontal gene transfer was related with transposases.
+These differences in the profile of functional categories between the two
+processes were interesting to us. An approach of dicriminating the two gene
+transfer processes would result in the differences.
+
+Because bacteria as a unicellular organism should be able to respond to changes
+to environment, some evolutionary pressures must have been acting on genes
+encoding membrane-bound molecules. Recombining gene transfer process seemed to
+be more associated with biological functions that had been critical of sensing
+queues of bacterial environmental changes.  Sugar phosphotransferase classes
+were ranked top in the list; phosphrylation could be important in signal
+transduction to recognize environmental changes.  The transport functional
+categories of carbohydrate transport and sodium ion transport also belonged to
+genes encoding membrane-bound molecules.  Also highly ranked were fatty acid and
+lipid biosyntheses that must be involved in producing membrane-bound molecules.
+Horizontal gene transfer process would allow bacteria to acquire foreign genes
+without the requirement of homologous regions in the bacteria.  Transposases
+would help foreign genes be excised from a donor genome and integrated into a
+recipient genome.
+
+In addition to these biological categories, we also found several aminoacyl tRNA
+synthetases (GO Term translation GO:0006412) in association with recombining
+gene transfer.  In the parsimony-based approach the term translation was top
+ranked (Table S\ref{tab:go-events-recombining}).  They had been a focus of gene
+transfer in bacterial species \citep{Woese2000}.  Some of the functional
+categories were unexpected; functional categories concerning ribosomes were
+found to be associated with recombining gene transfer.  Because genes related
+with ribosomes were considered to be housekeeping genes, gene transfer in which
+would be rare, it was unexpected that ribosome related genes were found to be
+associated with recombining gene transfer.  Because it was a bizarre result, we
+investigated the association by considering different recombinant edge types. We
+noted that two different classes of recombinant edge types: one that changes
+species tree topologies, and another that does not. Because recombinant edges
+that did not change species tree topology might be difficult to infer reliably,
+considering those recombinant edges in computing recombinantion intensity might
+lead to dubious results.  Ribosome related functional categories were found to
+be top ranked if we took those recombinant edges that did not change species
+tree topology (Tables S\ref{tab:functional-notopology} and
+S\ref{tab:functional-all}).  In Table \ref{tab:functional} ribosome related
+functional categories were not top ranked.  Therefore, high recombination
+intensity in ribosome related genes were likely to be attributed to
+topology-not-changing recombinant edges. We should be cautious of recombining
+transfer of genes related with ribosome related functions.
+
+The parsimony-based approach also showed
 enrichment of the recombining transfers in many of the same categories such as
@@
 edges in computing recombination intensity (see Tables
-S\ref{tab:functional-all} and S\ref{tab:go-events-recombining}).  Because the
+S\ref{tab:functional-all} and S\ref{tab:go-events-recombining}).  
+Because the
 parsimony-based method did not take species tree branch lengths into account, 
@@
 
-% More interpretations about Mawgli analysis of function categories?
-% Why transposases?
+\texttt{More interpretations about Mawgli analysis of function categories?
+Why transposases?}
+
 % \texttt{IG: NEXT PARAGRAPH CONTAINS SOME DISCUSSION ON THESE RESULTS. WE SHOULD 
@@
 alignments.  Using the simulated data sets of 100 274-block alignments we
-infer recombinant trees to draw the distribution of log base 2 ratio of
+infer recombinant trees to draw the distribution of logarithm base 2 ratio of
 observed recombinant edge counts to the prior expected number of recombinant
-edges. If the model with unbalanced gene flow was adequate, the log base 2
+edges. If the model with unbalanced gene flow was adequate, the logarithm base 2
 ratio from the real data set should center around the distribution. 
@@
 recombination intensity based on topology changing recombinant edge types.}
+\noindent \begin{centering}
 \begin{tabular}{cccl}
@@
 non-recombining horizontal gene transfer.}
+\noindent \begin{centering}
 \begin{tabular}{cccl}
@@
 \begin{table}
-\caption{\label{tab:genome}The six complete bacterial genomes. SDD genome was
-not yet available from NCBI.}
+\caption{\label{tab:genome}The six complete bacterial genomes. 
+Each bacteria and its genome is referred to as the given name.
+SDD genome was not yet available from NCBI.}
 \noindent \begin{centering}
@@
 \ref{fig:Heatmap-of-recombination} shows the values graphically using heat map.
-NA's are for the branch pairs that are impossible. Rows are for donor, and
-columns for recipient.}
+NA's are for the branch pairs that are impossible under the model-based
+approached. Rows are for donor, and columns for recipient.}
 \noindent \centering{}\begin{tabular}{cccccccccc}
@@
 \caption{\label{tab:obsheatmap}The observed number of recombinant edges.
-Rows are for donor, and columns for recipient.}
+The zeros are for the branch pairs that are impossible under the model-based
+approached. Rows are for donor, and columns for recipient.}
 \noindent \centering{}\begin{tabular}{cccccccccc}
@@
 \caption{\label{tab:three}Three population parameter estimates from
-ClonalOrigin. 
+the model-based approach.
 The estimates of the three parameters were based on one of two independent
@@
 results of the two independent chains as confirmation of 
-the convergence of the first stage of \texttt{ClonalOrigin} MCMC.
-}
+the convergence of the first stage of MCMC.}
 \noindent \begin{centering}
@@
 %=============================================================================
-% Table SXXX. Simulation study for the three population parameters.
-\begin{table}
-\caption{\label{tab:sim-three-population-parameters}
-Simulation study of the three population parameters.}
-\noindent \begin{centering}
-\begin{tabular}{ccc}
- & True & Estimates (Standard Deviation)\tabularnewline
-\hline
-Mutation rate per site & 0.081 & 0.081 (0.0011)\tabularnewline
-Recombination rate per site & 0.012 & 0.011 (0.0003)\tabularnewline
-Recombinant tract length & 744 bp & 919 bp (28)\tabularnewline
-\hline
-\end{tabular}
-\par\end{centering}
-\end{table}
-\clearpage{}
-%=============================================================================
-
+% Table S8. ClonalOrigin - All Topologies
 \begin{table}
@@
 recombination intensity based on all of the recombinant edge types.}
+\noindent \begin{centering}
 \begin{tabular}{cccl}
@@
 \end{tabular}
+\par\end{centering}
 \end{table}
 \clearpage{}
-
-%
+%=============================================================================
+
+%=============================================================================
+% Table S. Mowgli - Recombining Transfer
+\begin{table}
+\caption{\label{tab:go-events-recombining}Functional categories enriched in 
+recombining transfer events that are inferred using the parsimony-based approach.}
+\noindent \begin{centering}
+\begin{tabular}{cccl}
+\hline 
+P-value & Count & GO-term & Description \\
+\hline 
+7e-09 & 153 & GO:0006412 & translation\\
+1e-08 &  67 & GO:0030529 & ribonucleoprotein complex\\
+1e-07 &  45 & GO:0019843 & rRNA binding\\
+2e-06 &  73 & GO:0003735 & structural constituent of ribosome\\
+3e-05 & 157 & GO:0005840 & ribosome\\
+7e-05 & 221 & GO:0003723 & RNA binding\\
+2e-04 &  11 & GO:0004826 & phenylalanine-tRNA ligase activity\\
+3e-04 &  14 & GO:0015413 & nickel-transporting ATPase activity\\
+4e-04 &  93 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
+5e-04 & 104 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
+\hline 
+\end{tabular}
+\par\end{centering}
+\end{table}
+\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% Table S. Simulation study for the three population parameters.
 \begin{table}
@@
 recombination intensity based on the topology-not-changing recombinant edge types.}
+\noindent \begin{centering}
 \begin{tabular}{cccl}
@@
 \end{tabular}
+\par\end{centering}
 \end{table}
 \clearpage{}
-
+%=============================================================================
+
+%=============================================================================
+% Table S. ClonalOrigin - SDE-to-SPY
 \begin{table}
@@
 recombination intensity based on SDE-to-SPY recombinant edge types.}
+\noindent \begin{centering}
 \begin{tabular}{cccl}
@@
 \end{tabular}
+\par\end{centering}
 \end{table}
 \clearpage{}
-
+%=============================================================================
+
+%=============================================================================
+% Table S. ClonalOrigin - SPY-to-SDE
 \begin{table}
@@
 recombination intensity based on SPY-to-SDE recombinant edge types.}
-\begin{tabular}{cccc}
+\noindent \begin{centering}
+\begin{tabular}{cccl}
 \hline 
@@
 \end{tabular}
+\par\end{centering}
 \end{table}
 \clearpage{}
-
-%
+%=============================================================================
+
+%=============================================================================
+% Table S. Mowgli - Duplication 
 \begin{table}
-\caption{\label{tab:go-events-duplication}GO-term enrichment for different
-evolutionary events - Duplication.}
+\caption{\label{tab:go-events-duplication}Functional categories enriched in 
+gene duplication events that are inferred using the parsimony-based approach.}
+\noindent \begin{centering}
 \begin{tabular}{cccl}
@@
 \end{tabular}
+\par\end{centering}
 \end{table}
 \clearpage{}
-
-%
+%=============================================================================
+
+%=============================================================================
+% Table S. Mowgli - Loss
 \begin{table}
-\caption{\label{tab:go-events-loss}GO-term enrichment for different
-evolutionary events - Loss.}
+\caption{\label{tab:go-events-loss}Functional categories enriched in 
+gene loss events that are inferred using the parsimony-based approach.}
+\noindent \begin{centering}
 \begin{tabular}{cccl}
@@
 \end{tabular}
+\par\end{centering}
 \end{table}
 \clearpage{}
-
-%
+%=============================================================================
+
+%=============================================================================
+% Table S15. Simulation study for the three population parameters.
 \begin{table}
-\caption{\label{tab:go-events-recombining}GO-term enrichment for different
-evolutionary events - Recombining transfers.}
-\begin{tabular}{cccl}
-\hline 
-P-value & Count & GO-term & Description \\
-\hline 
-7e-09 & 153 & GO:0006412 & translation\\
-1e-08 &  67 & GO:0030529 & ribonucleoprotein complex\\
-1e-07 &  45 & GO:0019843 & rRNA binding\\
-2e-06 &  73 & GO:0003735 & structural constituent of ribosome\\
-3e-05 & 157 & GO:0005840 & ribosome\\
-7e-05 & 221 & GO:0003723 & RNA binding\\
-2e-04 &  11 & GO:0004826 & phenylalanine-tRNA ligase activity\\
-3e-04 &  14 & GO:0015413 & nickel-transporting ATPase activity\\
-4e-04 &  93 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
-5e-04 & 104 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
+\caption{\label{tab:sim-three-population-parameters}
+Simulation study of the three population parameters.}
+\noindent \begin{centering}
+\begin{tabular}{ccc}
+\hline
+ & True & Estimates (Standard Deviation)\tabularnewline
+\hline
+Mutation rate per site & 0.081 & 0.081 (0.0011)\tabularnewline
+Recombination rate per site & 0.012 & 0.011 (0.0003)\tabularnewline
+Recombinant tract length & 744 bp & 919 bp (28)\tabularnewline
 \hline
 \end{tabular}
+\par\end{centering}
 \end{table}
 \clearpage{}
+%=============================================================================
 
@@
 \begin{figure}
-\includegraphics[scale=0.5]{figures/cornellf-3-tree}
+\includegraphics[scale=0.7]{figures/cornellf-3-tree}
 \caption{\label{fig:tree5}The species tree of the five genomes inferred using
@@
 \end{center}
-\caption{\label{fig:blocksize}The distribution of core alignment block sizes.}
+\caption{\label{fig:blocksize}Length distribution of the 274 core alignment blocks.}
 \end{figure}
@@
 for each block. The density of each value for each block is depicted by black
-coulds.  The red dashed lines are the global block-length weighted median of the
-blocks.}
+clouds.  The red dashed lines are the global block-length weighted median of the
+three parameter estimates.}
 \end{figure}
@@
 \end{center}
-\caption{Plot of recombination rate and logarithm of average tract length.}
+\caption{Plot of recombination rate and logarithm of average tract length
+of the posterior samples.}
 \label{fig:rhologdelta}
@@
 the species tree (A) in their topology.  Because there are 105 possible rooted
-topologies relating five taxa, we could count topologies.}
+labeled topologies relating five taxa, we could count topologies.}
 \end{figure}
@@
 \begin{figure}
-\subfloat[]{\includegraphics[scale=0.5]{figures/h3-R-ratio-topology-changing}}
-\subfloat[]{\includegraphics[scale=0.5]{figures/h3-R-ratio-not-topology-changing}}
+
+\subfloat[]{\includegraphics[scale=0.4]{figures/h3-R-ratio-topology-changing}}
+
+\subfloat[]{\includegraphics[scale=0.4]{figures/h3-R-ratio-not-topology-changing}}
+
 \caption{\label{fig:h3}Numer of recombination events. The diagonal dashed
@@
 %=============================================================================
-% Figure S12: 
-\begin{figure}
-\subfloat[]{\includegraphics[scale=0.5]{figures/h1-R}}
-\subfloat[]{\includegraphics[scale=0.5]{figures/h1-R-inset}}
-\caption{\label{fig:h1}Numer of recombination events. The diagonal dashed
-lines are with slope of 1 and intercept of 0. For each order
-pairs of branches of a reference species tree the true number of recombination
-events is computed as the mean of the 10 replicates of recombinant
-trees from the simulation. With ten estimates of each replicate total
-number of 100 estimats are shown as intervals of 90\% of quantile
-range, or the lower value of the interval is the 5\% quantile, and
-the upper one is the 95\% quantile. }
-\end{figure}
-\clearpage{}%
-%=============================================================================
-
-%=============================================================================
 % Figure S13:
@@
 
+%=============================================================================
+% Figure S13: 
+%\begin{figure}
+%\subfloat[]{\includegraphics[scale=0.5]{figures/h1-R}}
+%\subfloat[]{\includegraphics[scale=0.5]{figures/h1-R-inset}}
+%\caption{\label{fig:h1}Numer of recombination events. The diagonal dashed
+%lines are with slope of 1 and intercept of 0. For each order
+%pairs of branches of a reference species tree the true number of recombination
+%events is computed as the mean of the 10 replicates of recombinant
+%trees from the simulation. With ten estimates of each replicate total
+%number of 100 estimats are shown as intervals of 90\% of quantile
+%range, or the lower value of the interval is the 5\% quantile, and
+%the upper one is the 95\% quantile. }
+%\end{figure}
+%\clearpage{}%
+
+%=============================================================================
 %=============================================================================

### 2011-10-14 doc/manuscript.tex
An interesting vim extension for saving sessions.

@@
 intensity in ribosome related genes were likely to be attributed to
-topology-not-changing recombinant edges. We should be cautious of recombining
+topology-not-changing recombinant edges. We would be cautious of recombining
 transfer of genes related with ribosome related functions.
 
-The parsimony-based approach also showed
-enrichment of the recombining transfers in many of the same categories such as
-translation, ribosome, and rRNA binding when using all types of recombinant
-edges in computing recombination intensity (see Tables
-S\ref{tab:functional-all} and S\ref{tab:go-events-recombining}).  
-Because the
-parsimony-based method did not take species tree branch lengths into account, 
-the agreement between the two different methods might be likely.
-
-For horizontal gene transfer we found enrichment in very different categories
-related to transposition, integration, and recombination (Table
-\ref{tab:go-events}).  Gene duplication and loss events were enriched for
-categories similar to those of horizontal gene transfer's enrichment (Tables
-S\ref{tab:go-events-duplication} and S\ref{tab:go-events-loss}).
-Gene duplication and horizontal transfers showed enrichment in the same
-categories perhaps because both mechanisms increase gene copy number.  
-Gene losses showed enrichment for transposition and integration, but less
-significantly (P-value of $<10^{-6}$), and showed enrichment for several
-additional categories such as transporter activity and transcription initiation.
-
-Comparing two types of recombinant edges, one from SDE to SPY and another from
-SPY to SDE transport related categories were more associated for the direction
-from SPY to SDE
-(Table S\ref{tab:functional-sde-spy} and S\ref{tab:functional-spy-sde}).
-
-\texttt{More interpretations about Mawgli analysis of function categories?
-Why transposases?}
-
-% \texttt{IG: NEXT PARAGRAPH CONTAINS SOME DISCUSSION ON THESE RESULTS. WE SHOULD 
-% GET BACK TO THIS ONCE WE DECIDE WHAT TO DO WITH THE RESULTS}
+\subsection{Simulations}
+
+1. Topology-not-changing recombinant edge was unreliably estimated compared with
+topology-changing recombinant edge.
+2. Balanced case
+3. Unbalanced case
+
+We conducted experiments on simulation data to check the accuracy of estimates
+obtained by the model-based approach in a scenario similar to one from the real 
+data set.
+
+While populaiton mutation and
+recombination parameters were accurately estimated, the recombinant tract length
+was over-estimated  (Table S\ref{tab:sim-three-population-parameters}).  
+Recombinant tract lengths were reported to be difficult to estimate from relatively
+short alignment blocks \citep{Didelot2010}.
+
+Figure S\ref{fig:h1} shows the observed count of each recombinant
+edge type against the expected count; the estimated numbers appeared to be
+slightly over-estimated, but in correlation with the true numbers.
+
+We also evaluated the accuracy of the statistics estimated in the second stage
+of \texttt{ClonalOrigin}.
+
+The number of recombinant edges for each pair of
+recombinant tree branches was correlated with the true value in general while
+there were slight over-estimation (Figure \ref{fig:h3}).
+
+However, this
+comparison between inferrence from simulated data and that of real data set
+suggested that the directional difference between SPY and SDE was due to neither
+lack of power of \texttt{ClonalOrigin} nor hidden bias thereof.
+
+Recombination
+intensities that were measured on genes also showed rough correlation (Pearson's
+correlation coefficient of 0.63) of the estimated and true values (Figure
+S\ref{fig:ri1}).
+
+Estimated values appeared to be driven somewhat by the prior
+implied by the population parameter setting, explaining why lower values were
+over-estimated and higher values were under-estimated.
+
+Nonetheless, the
+correlation indicated that the measure of recombination intensity used in our
+functional gene category served as an adequate indicator of actual intensity of
+recombination events along genes.
 
@@
 
-We investigated closely related bacterial species to better understand gene
-transfer processes between SPY and SDE. Two complementary methods were employed
-to study recombining and horizontal gene transfers. By discriminating the two
-processes in analyzing the genomes, we showed that the pathogenicity of SDE and
-SPY was more attributed to horizontal gene transfer process than recombining
-gene transfer.  Because recombination in bacteria seemed to occur at somewhat
-random positions along the genome, the homologous parts transferred might be
-unlikely to be enriched with virulence factors to human hosts.  Recombining gene
-transfer might have rarely contributed to influx of virulence genes from SPY to
-the opportunistic human pathogen of SDE as we showed a few cases of virulence
-gene that might have been transferred via recombining gene transfer.
-
-We also showed that the net amount of recombining gene flows from SPY to SDE was
-relatively greater than that of the reverse direction in the core genome
-sequences.  If there were unbalanced recombining gene transfer between the two
-species, what were the barriers \citep{Thomas2005} to the gene transfer in the
-bacteria? Although barriers to bacterial recombination would include geographic
-isolation, genetic isolation could prevent recombination between coinhabiting
-bacteria.  \citet{Kalia2001} discussed restriction-modification system (RMS) as
-a possible mechanism of the unbalanced gene flow.  Restriction-modification
+Molecular evolution of bacterial genomes is multifaceted. Studies of bacterial
+evolution often focused on one side of many evolutionary processes.  While
+understanding one aspect of bacterial evolutionary processes we could be
+ignorant of the other features.  Horizontal gene transfer process was often
+considered in macro-evolutionary scale using a great deal of bacterial species.
+As closely related species or groups of bacterial individuals are becoming a
+focus of evolutionary study, recombining gene transfer process would surface as
+another facet of evolutionary forces.  Because the two gene transfer processes
+would act on bacterial genomes simultaneously, ignoring one of the two processes
+in studying evolution of bacteria would lead to a narrowed point of view to the
+underlying evolutionary forces. Our study of SPY and SDE genomes undertook the
+work of discriminating recombining and non-recombining horizontal gene transfer
+processes because the two species were very closely related.  We showed that the
+two gene transfer processes acted on different biological functions. Recombining
+gene transfer process in the analysis of SPY and SDE genomes was related with
+genes encoding membrane-bound molecules and aminoacyl tRNA synthetases. On the
+contrary, horizontal gene transfer process was found to be associated with
+transposases, which would allow horizontal gene transfer of their neighboring
+genes.  Because SPY was a human-specific pathogen, and SDE was an opportunistic
+human pathogen, we were interested in association of pathogenicity with the two
+gene transfer processes. We showed that virulence genes were more attributed to
+horizontal gene transfer process than to recombining gene transfer process.
+With the current data of only SPY and SDE genomes we could not conclude that the
+association of virulence genes with horizontal gene transfer not recombining
+gene transfer was applicable to other species.  However, without the
+discrimination of the two processes it would have been difficult to find the
+disparate association of biological functions or virulence genes with the
+different gene transfer processes.
+
+Another contribution of our study to bacterial gene transfer research is to
+elucidate the net directionality of gene flow from SPY to SDE in the core
+genomes. Although the issue of net directionality has been tackled, our study
+was the first attempt of resolving the issue in genome-wide scale. Yet, it still
+remains to be seen whether this net directionality of gene flow from SPY to SDE
+happened in the pan genome. A model-based method with more refined models would
+be necessary for the reserach direction because the parsimony-based method we
+employed simply could not show any net directionality between SPY and SDE in the
+pan genome. Although we are unsure whether much larger number of genomes could
+be useful in finding net directionality, we consider that more SDE genomes would
+be beneficial to future studies.
+
+Were there unbalanced recombining gene transfer between the two species, what
+were the barriers \citep{Thomas2005} to cause the unbalanced gene transfer in
+the species?  \citet{Kalia2001} discussed restriction-modification system (RMS)
+as a possible mechanism of the unbalanced gene flow.  Restriction-modification
 system is a biological self-defense mechanism of bacteria, which is diverse in
@@
 RMS, claiming that RMS modulated homologous recombination in the bacteria of
-\textit{N.\ meningitidis}.  Most of \textit{S.\ pyogenes} genomes contain
+\textit{N.\ meningitidis}.  Most of \textit{S.\ pyogenes} genomes contained
 restriction-modification type I system as well as type II system. The two
@@
 
-We used relatively small number of genomes to detect recombining gene transfer
-that had only been possible using many samples for a few genes. We revisited the
-problem of gene transfer in other previous studies.  We reexamined the seven
-genes that \citet{Kalia2001} had studied for net directionality of gene transfer
-between SPY and SDE.  Internal fragments of the six housekeeping genes, encoding
-putative glucose kinase (\textit{gki}), glutamine transport protein
-(\textit{gtr}), glutamate racemase (\textit{murI}), mismatch repair enzyme
-(\textit{mutS}), transketolase (\textit{recP}), and xanthine
-phosphoribosyltransferase (\textit{xpt}) were located in the core genome that we
-investigated in this study.  Genes \textit{mutS}, \textit{recP}, and
-\textit{xpt} had fragments of gene transfer with moderate posterior probability
-of recombination. The other genes did not have noticeable recombination. Except
-for \textit{mutS} where \citet{Kalia2001} did not find recombination, the result
-was consistent with that of \citet{Kalia2001}.  We also revisited recombination
-in gene \textit{parC} in SPY1 genome studied by \citet{Pinho2010}, where they
-studied fluoroquinolone resistance in SDE and reported recombination in the
-gene. We observed moderate recombination in the gene.
-
-We used two heterogeneous methods: one model-based and another parsimony-based
-for studying recombining and horizontal gene transfer.  The core genomes
-consisted of sequence alignments with rare gaps. With this inherent filtering of
-the input data preparation the model-based method did not accommodate gene
-dupliation and loss in its core genome analysis. On the contrary, the
-parsimony-based method used all of the genomes not just the core genomes.
-Considering nucleotides as the level of gene duplication, loss, and transfer
-towards a complete model for recombining and horizontal gene transfer could be
-difficult because genome-scale application of those models based on DNA or
-protein sequences might not be feasible.  Combining the two approaches for
-studying the mode of gene transfer would be desirable to study more complicated
-evolutionary processes including recombining gene transfer, gene duplication,
-loss, and horizontal gene transfer.
-
-\texttt{Limitation of ClonalOrigin: inability of inferring recombinant edges}
-
-\texttt{Need of a new model with asymmetry of gene transfer}
-
-\texttt{Need of more SDE genomes}
+We reexamined the seven genes that \citet{Kalia2001} had studied for net
+directionality of gene transfer between SPY and SDE.  The internal fragments of
+the six housekeeping genes (\textit{gki}, \textit{gtr}, \textit{murI},
+\textit{mutS}, \textit{recP}, \textit{xpt}) were located in the core genome that
+we investigated in this study.  Except for \textit{mutS} where \citet{Kalia2001}
+did not find recombination, our results were consistent with that of
+\citet{Kalia2001}.  \citet{Pinho2010} reported recombination in the gene
+\textit{parC}, in which we also observed moderate recombination.
+
+It remains to be seen whether the inability of detecting net directionality
+between SPY and SDE in the pan genomes was because of lack of power of the
+parsimony-based approach that we employed. Considering gene duplication, loss,
+and horizontal gene transfer in a model-based approach would be desirable in the
+coming ages of abundant bacterial genomes. The model might consider genome
+alignments where each taxon does not necessarily have a unique gene; some taxa
+might have no gene due to gene loss, and other taxa multiple genes due to gene
+gain, duplication, or horizontal gene transfer. With these types of models will
+we be able to address more interesting evolutionary questions in bacteria. This
+need of more refined models would be also true even for the study of net
+directionality between species in the core genome.  In other words, one could
+envision a model where a population two-taxa (i.e., SPY and SD groups) tree is
+superimposed on the species tree shown in Figure S\ref{fig:clonalorigin}A. A
+splitting event of the population tree could happen somewhere between the root
+of the species tree and the internal node of SPY. Recombinant edges that connect
+branches within SPY or SD clades would be more frequent than those that connect
+branches between SPY and SDE clades.  With the reference to the recombination
+within a clade one parameter can dictate the degree of gene from from SPY to
+SDE, and another for the reverse direction.  If the two parameters could be
+incorporated in the model, then they might be used to compare the degrees of
+gene flow between the two species in a more statistically sound way.  We hope
+that we will see thie refinement of the model to better illustrate bacterial
+evolution in forthcoming bacterial genomic era.
 
@@
 \subsection{Simulation Study}
-
-We conducted experiments on simulation data to check the accuracy of estimates
-obtained by \texttt{ClonalOrigin} in a scenario similar to one from the real 
-data set.  We used a
-simulation feature of \texttt{ClonalOrigin} to simulate the process of mutation
-and recombination across the five species 
+We performed three sets of simulations.  In the first simulation, we wished to
+show that \texttt{ClonalOrigin} could infer population mutation rate, tract
+length, and recombination rate with reasonable accuracy. In the second
+simulation, we wished to check whether recombinant edges were recovered from
+simulated data with recombinant trees generated from the prior. The third
+simulation was conducted using simulated data with recombinant trees sampled the
+posterior distribution.  We used a simulation feature of \texttt{ClonalOrigin}
+to simulate the process of mutation and recombination across the five species
 (see section \textbf{Supplementary Information} for detail).
-While populaiton mutation and
-recombination parameters were accurately estimated, the recombinant tract length
-was over-estimated  (Table S\ref{tab:sim-three-population-parameters}).  This
-was not surprising as recombinant tract lengths were known to be difficult to
-estimate from relatively short alignment blocks \citep{Didelot2010}. 
-We also evaluated the accuracy of the statistics estimated in the second stage
-of \texttt{ClonalOrigin}. The number of recombinant edges for each pair of
-recombinant tree branches was correlated with the true value in general while
-there were slight over-estimation (Figure S\ref{fig:h1}).  However, this
-comparison between inferrence from simulated data and that of real data set
-suggested that the directional difference between SPY and SDE was due to neither
-lack of power of \texttt{ClonalOrigin} nor hidden bias thereof.  Recombination
-intensities that were measured on genes also showed rough correlation (Pearson's
-correlation coefficient of 0.63) of the estimated and true values (Figure
-S\ref{fig:ri1}).  Estimated values appeared to be driven somewhat by the prior
-implied by the population parameter setting, explaining why lower values were
-over-estimated and higher values were under-estimated.  Nonetheless, the
-correlation indicated that the measure of recombination intensity used in our
-functional gene category served as an adequate indicator of actual intensity of
-recombination events along genes.
 
@@
 
-\subsection{Local gene trees and recombination intensity}
+\subsection{Summarization of posterior samples for local gene trees and recombination intensity}
 
 Because the procedure of summarizing posterior samples of \textit{recombinant
-tree}, which will be defined in a moment, 
-was not detailed in the user guide at the time when we used
-\texttt{ClonalOrigin}, we elected to use Figure S\ref{fig:clonalorigin} to
-illustrate a recombinant tree in order to clarify summarization of
-\textit{recombinant history} posterior samples.  A recombination history is
-represented by a \textit{recombinant tree}, which is a species tree augmented
-with \textit{recombinant edges}.  A recombinant edge connect two species tree
-branches.  Each recombinant edge is defined by the time
+tree}, which will be defined in a moment, was not detailed in the user guide at
+the time when we used \texttt{ClonalOrigin}, we elected to use Figure
+S\ref{fig:clonalorigin} to illustrate a recombinant tree in order to clarify
+summarization of \textit{recombinant history} posterior samples.  A
+recombination history is represented by a \textit{recombinant tree}, which is a
+species tree augmented with \textit{recombinant edges}.  A recombinant edge
+connect two species tree branches.  Each recombinant edge is defined by the time
 and branch of origin and destination, and the genomic interval that it affects.
@@
 \textbf{Simulation Study}).  Our simulation study indicated that edges of
-certain types were better indicators of recombination events than others.
-For instance, recombinant edges which resulted in a recombinant tree with the
-same topology as the species tree appeared to be more prone to spurious
-sampling.  Such edges were ones where the donor was the same branch, sister
-branch, or parent branch of the recipient branch.  Therefore, in some cases we
-elected to remove such edges from consideration when measuring recombination
-intensity.  
-% The tree topologies changes would be correlated with recombination
-% intensity.  The tree topologies would show more effectively regions where
+certain types were better indicators of recombination events than others.  For
+instance, recombinant edges which resulted in a recombinant tree with the same
+topology as the species tree appeared to be more prone to spurious sampling.
+Such edges were ones where the donor was the same branch, sister branch, or
+parent branch of the recipient branch.  Therefore, in some cases we elected to
+remove such edges from consideration when measuring recombination intensity.  
+% The tree topologies would show more effectively regions where
 % recombination occurs along the genome (see subsection of \textbf{UCSC Genome
@@
 
-\subsection{Population mutation and recombination rates were accurately
-estimated}
-
+\subsection{Simulation 1 - three population parameters}
+
+We describe a procedure of simulations for checking if the three
+population parameters were accurately recovered using the model-based approach.
+Ten recombinant trees were sampled given the three parameters (Table
+S\ref{tab:three}) and the species tree (Figure S\ref{fig:tree5}) set to the
+values inferred from the five bacterial genomes.  Each of the recombinant trees
+was used to simulate the Jukes-Cantor model for DNA sequence evolution to create
+274 blocks of alignments.  The block lengths configuration was the same as that
+of the blocks used in the analysis of the five genomes.
+
+% s15
 % FIXME: Newick tree in text in a file.
 % FIXME: Publish the result and the data set with the code.
-We describe the detailed procedure of simulations for checking if the three
-population parameters were accurately recovered from the first stage of
-\texttt{ClonalOrigin} MCMC.  Ten recombinant trees were sampled given the three
-parameters and the species tree  set to the values inferred from the five
-bacterial genomes.  Each of the recombinant trees was used to simulate the
-Jukes-Cantor model for DNA sequence evolution to create 274 blocks of
-alignments.  The block lengths configuration was the same as that of the blocks
-used in the analysis of the five genomes.  The mean of the ten estimates of the
-recombinant tract length was significantly higher than the true value of (919 bp
-versus 744 bp) whereas the values of the mutation and recombination rates were
-accurately estimated (Table S\ref{tab:sim-three-population-parameters}).
-Recombinant tract lengths were known to be difficult to estimate from relatively
-short alignment blocks \citep{Didelot2010}.
-
-\subsection{Estimating numbers of recombinant edges were accurate}
-
+
+\subsection{Simulation 2 - numbers of recombinant edges from prior}
+
+We first sampled ten recombinant trees given the mutation rate, recombination
+rate, and tract length (Table S\ref{tab:three}) and the species tree (Figure
+S\ref{fig:tree5}), and then created ten independent simulated data sets with
+the Jukes-Cantor model for DNA sequence evolution using each of the recombinant
+trees. This consisted of 100 replicates of data sets.  The recombinant trees
+were created from the prior distribution and recombinant edges were added to the
+species tree with no particular directions of recombining gene transfers between
+species tree branches.  We computed the ratio of the mean number of recombinant
+edges for a pair of donor and recipient species tree branches over the simulated
+10 recombinant trees against the median of recombinant edges estimated from the
+10 simulated data sets, and then took the logarithm of the ratio to base 2. We
+also computed 5\% and 95\% quantiles to find the 90\% quantile range for the log
+base 2 ratio.  We expect that the log base 2 ratio would near zero. We also
+wished to see if how the recombinant edge count estimates from the real data set
+deviate from the prior. Instead of using the number of recombinant edges from
+the ten recombinant trees, we used the numbers of recombinant edges that were
+estimated from the real data.
+
+We tested if our measure of recombination intensity provided a reliable estimate
+of the true intensity. We artificially annotated our synthetic blocks to genes
+according to the real annotations from SPY1 genome.  For each gene in each of
+the simulated 100 data sets, we plotted the average estimated recombination
+intensity along that gene against the average number of different recombinant
+edge types from the ten simulated recombinant tree.
+
+% s16
+
+% s16
 % FIXME 1: IG: RESULTS FROM THIS POINT ON REQUIRE SOME REFINEMENT. SC: Let's see.
@@
 % ARTIFICIALLY DOUBLE OR TRIPLE THE RECOMBINATION EVENTS OF ONE OF TYPES.
-Because of the two-stage Bayesian method we first sampled ten recombinant trees
-given the mutation rate, recombination rate, and tract length, and then created
-ten independent simulated data sets with the Jukes-Cantor model for DNA sequence
-evolution using each of the recombinant trees. This consisted of 100 replicates
-of data sets.  Figure S\ref{fig:h1} shows the observed count of each recombinant
-edge type against the expected count; the estimated numbers appeared to be
-slightly over-estimated, but in correlation with the true numbers.
-
-\subsection{Unbalanced gene flow model was adequate}
-
-We checked if the model with unbalanced gene flow was adequate using posterior
-predictive check. We used 100 recombinant trees from the posterior distribution
-of the recombiant tree given the real data of the 274 alignment blocks in order
-to simulate the ClonalOrigin model to create a new data set of 100 274-block
-alignments.  Using the simulated data sets of 100 274-block alignments we
-infer recombinant trees to draw the distribution of logarithm base 2 ratio of
-observed recombinant edge counts to the prior expected number of recombinant
-edges. If the model with unbalanced gene flow was adequate, the logarithm base 2
-ratio from the real data set should center around the distribution. 
-\texttt{What is the result of this simulation?}
-
-\subsection{Simulation for recombination intensity}
-
-We tested if our measure of recombination intensity provided a reliable estimate
-of the true intensity. For this purpose, we artificially annotated our synthetic
-blocks to genes according to the real annotations along the real bacterial
-alignments. For each gene in each of the simulated 100 data sets, we plotted the
-average estimated recombination intensity along that gene against the true
-average intensity, i.e., the average number of different recombinant edge types
-implied by the recombinant tree used in simulation 
-(Pearson's correlation coefficient of 0.63 and Figure S\ref{fig:ri1}).
+
+\subsection{Simulation 3 - number of recombinant edges from posterior}
+
+We used 100 recombinant trees sampled from the posterior distribution
+of the recombiant tree instead of using simulated recombinant trees.
+We created one simulated data set for each of the 100 recombinant trees with
+the Jukes-Cantor model for DNA sequence evolution 
+using the mutation rate estimate (Table S\ref{tab:three}).
+This consisted of 100 replicates of data sets.  
+We follow the same computation of the ratio as in the simulation 2 with 
+the numbers of recombinant edges estimated from the real data
+as numerators.
+
+% s17
 
@@
 are not indicated).}
+\end{figure}
+\clearpage{}%
+%=============================================================================
+
+%=============================================================================
+% Figure S12: 
+\begin{figure}
+
+\subfloat[]{\includegraphics[scale=0.4]{figures/h3-R-ratio-topology-changing}}
+
+\subfloat[]{\includegraphics[scale=0.4]{figures/h3-R-ratio-not-topology-changing}}
+
+\caption{\label{fig:h3}Numer of recombination events. The diagonal dashed
+lines are with slope of 1 and intercept of 0. For each order
+pairs of branches of a reference species tree the true number of recombination
+events is computed as the mean of the 10 replicates of recombinant
+trees from the simulation. With ten estimates of each replicate total
+number of 100 estimats are shown as intervals of 90\% of quantile
+range, or the lower value of the interval is the 5\% quantile, and
+the upper one is the 95\% quantile. }
 \end{figure}
@@
 % Table S. ClonalOrigin - SDE-to-SPY
-\begin{table}
-\caption{\label{tab:functional-sde-spy}Functional categories asscoiated with 
-recombination intensity based on SDE-to-SPY recombinant edge types.}
-\noindent \begin{centering}
-\begin{tabular}{cccl}
-\hline 
-P-Value & Count & GO term & Description\tabularnewline
-\hline 
-1.854e-05 & 38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
-2.703e-05 & 81 & GO:0006412 & translation\tabularnewline
-6.482e-05 & 33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
-9.777e-05 & 105 & GO:0003723 & RNA binding\tabularnewline
-0.00012793 & 40 & GO:0030529 & ribonucleoprotein complex\tabularnewline
-0.00025176 & 26 & GO:0019843 & rRNA binding\tabularnewline
-0.00031236 & 72 & GO:0005975 & carbohydrate metabolic process\tabularnewline
-0.00044756 & 317 & GO:0005737 & cytoplasm\tabularnewline
-0.00059675 & 45 & GO:0008643 & carbohydrate transport\tabularnewline
-0.00066874 & 39 & GO:0003735 & structural constituent of ribosome\tabularnewline
-0.00135304 & 10 & GO:0006098 & pentose-phosphate shunt\tabularnewline
-\hline 
-\end{tabular}
-\par\end{centering}
-\end{table}
-\clearpage{}
+%\begin{table}
+%\caption{\label{tab:functional-sde-spy}Functional categories asscoiated with 
+%recombination intensity based on SDE-to-SPY recombinant edge types.}
+%\noindent \begin{centering}
+%\begin{tabular}{cccl}
+%\hline 
+%P-Value & Count & GO term & Description\tabularnewline
+%\hline 
+%1.854e-05 & 38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
+%2.703e-05 & 81 & GO:0006412 & translation\tabularnewline
+%6.482e-05 & 33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
+%9.777e-05 & 105 & GO:0003723 & RNA binding\tabularnewline
+%0.00012793 & 40 & GO:0030529 & ribonucleoprotein complex\tabularnewline
+%0.00025176 & 26 & GO:0019843 & rRNA binding\tabularnewline
+%0.00031236 & 72 & GO:0005975 & carbohydrate metabolic process\tabularnewline
+%0.00044756 & 317 & GO:0005737 & cytoplasm\tabularnewline
+%0.00059675 & 45 & GO:0008643 & carbohydrate transport\tabularnewline
+%0.00066874 & 39 & GO:0003735 & structural constituent of ribosome\tabularnewline
+%0.00135304 & 10 & GO:0006098 & pentose-phosphate shunt\tabularnewline
+%\hline 
+%\end{tabular}
+%\par\end{centering}
+%\end{table}
+%\clearpage{}
 %=============================================================================
@@
 % Table S. ClonalOrigin - SPY-to-SDE
-\begin{table}
-\caption{\label{tab:functional-spy-sde}Functional categories asscoiated with 
-recombination intensity based on SPY-to-SDE recombinant edge types.}
-\noindent \begin{centering}
-\begin{tabular}{cccl}
-\hline 
-P-Value & Count & GO term & Description\tabularnewline
-\hline 
-1.106e-06 & 45 & GO:0008643 & carbohydrate transport\tabularnewline
-1.511e-06 & 33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
-1.595e-06 & 38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
-1.443e-05 & 187 & GO:0006810 & transport\tabularnewline
-0.0001106 & 371 & GO:0016740 & transferase activity\tabularnewline
-0.0003596 & 72 & GO:0005975 & carbohydrate metabolic process\tabularnewline
-0.0004023 & 13 & GO:0006814 & sodium ion transport\tabularnewline
-0.000823 & 317 & GO:0005737 & cytoplasm\tabularnewline
-0.001142 & 12 & GO:0031402 & sodium ion binding\tabularnewline
-\hline 
-\end{tabular}
-\par\end{centering}
-\end{table}
-\clearpage{}
+%\begin{table}
+%\caption{\label{tab:functional-spy-sde}Functional categories asscoiated with 
+%recombination intensity based on SPY-to-SDE recombinant edge types.}
+%\noindent \begin{centering}
+%\begin{tabular}{cccl}
+%\hline 
+%P-Value & Count & GO term & Description\tabularnewline
+%\hline 
+%1.106e-06 & 45 & GO:0008643 & carbohydrate transport\tabularnewline
+%1.511e-06 & 33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
+%1.595e-06 & 38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
+%1.443e-05 & 187 & GO:0006810 & transport\tabularnewline
+%0.0001106 & 371 & GO:0016740 & transferase activity\tabularnewline
+%0.0003596 & 72 & GO:0005975 & carbohydrate metabolic process\tabularnewline
+%0.0004023 & 13 & GO:0006814 & sodium ion transport\tabularnewline
+%0.000823 & 317 & GO:0005737 & cytoplasm\tabularnewline
+%0.001142 & 12 & GO:0031402 & sodium ion binding\tabularnewline
+%\hline 
+%\end{tabular}
+%\par\end{centering}
+%\end{table}
+%\clearpage{}
 %=============================================================================
@@
 % Table S. Mowgli - Duplication 
-\begin{table}
-\caption{\label{tab:go-events-duplication}Functional categories enriched in 
-gene duplication events that are inferred using the parsimony-based approach.}
-\noindent \begin{centering}
-\begin{tabular}{cccl}
-\hline 
-P-value & Count & GO-term & Description \\
-\hline 
-4e-81 &  67 & GO:0006313 & transposition, DNA-mediated\\
-5e-80 &  59 & GO:0004803 & transposase activity\\
-6e-70 &  23 & GO:0032196 & transposition\\
-2e-58 &  52 & GO:0015074 & DNA integration\\
-8e-25 &  86 & GO:0006310 & DNA recombination\\
-1e-20 &  36 & GO:0016987 & sigma factor activity\\
-1e-20 &  36 & GO:0006352 & transcription initiation\\
-2e-19 & 197 & GO:0003676 & nucleic acid binding\\
-2e-10 & 527 & GO:0003677 & DNA binding\\
-4e-06 & 142 & GO:0043565 & sequence-specific DNA binding\\
-\hline 
-\end{tabular}
-\par\end{centering}
-\end{table}
-\clearpage{}
+%\begin{table}
+%\caption{\label{tab:go-events-duplication}Functional categories enriched in 
+%gene duplication events that are inferred using the parsimony-based approach.}
+%\noindent \begin{centering}
+%\begin{tabular}{cccl}
+%\hline 
+%P-value & Count & GO-term & Description \\
+%\hline 
+%4e-81 &  67 & GO:0006313 & transposition, DNA-mediated\\
+%5e-80 &  59 & GO:0004803 & transposase activity\\
+%6e-70 &  23 & GO:0032196 & transposition\\
+%2e-58 &  52 & GO:0015074 & DNA integration\\
+%8e-25 &  86 & GO:0006310 & DNA recombination\\
+%1e-20 &  36 & GO:0016987 & sigma factor activity\\
+%1e-20 &  36 & GO:0006352 & transcription initiation\\
+%2e-19 & 197 & GO:0003676 & nucleic acid binding\\
+%2e-10 & 527 & GO:0003677 & DNA binding\\
+%4e-06 & 142 & GO:0043565 & sequence-specific DNA binding\\
+%\hline 
+%\end{tabular}
+%\par\end{centering}
+%\end{table}
+%\clearpage{}
 %=============================================================================
@@
 % Table S. Mowgli - Loss
-\begin{table}
-\caption{\label{tab:go-events-loss}Functional categories enriched in 
-gene loss events that are inferred using the parsimony-based approach.}
-\noindent \begin{centering}
-\begin{tabular}{cccl}
-\hline 
-P-value & Count & GO-term & Description \\
-\hline 
-1e-06 &  67 & GO:0006313 & transposition, DNA-mediated\\
-2e-05 &  59 & GO:0004803 & transposase activity\\
-5e-05 &  52 & GO:0015074 & DNA integration\\
-5e-04 &  23 & GO:0032196 & transposition\\
-5e-04 &  10 & GO:0046873 & metal ion transmembrane transporter activity\\
-7e-04 &  36 & GO:0016987 & sigma factor activity\\
-7e-04 &  36 & GO:0006352 & transcription initiation\\
-1e-03 &  31 & GO:0006814 & sodium ion transport\\
-3e-03 &  86 & GO:0006310 & DNA recombination\\
-6e-03 & 197 & GO:0003676 & nucleic acid binding\\
-\hline 
-\end{tabular}
-\par\end{centering}
-\end{table}
-\clearpage{}
+%\begin{table}
+%\caption{\label{tab:go-events-loss}Functional categories enriched in 
+%gene loss events that are inferred using the parsimony-based approach.}
+%\noindent \begin{centering}
+%\begin{tabular}{cccl}
+%\hline 
+%P-value & Count & GO-term & Description \\
+%\hline 
+%1e-06 &  67 & GO:0006313 & transposition, DNA-mediated\\
+%2e-05 &  59 & GO:0004803 & transposase activity\\
+%5e-05 &  52 & GO:0015074 & DNA integration\\
+%5e-04 &  23 & GO:0032196 & transposition\\
+%5e-04 &  10 & GO:0046873 & metal ion transmembrane transporter activity\\
+%7e-04 &  36 & GO:0016987 & sigma factor activity\\
+%7e-04 &  36 & GO:0006352 & transcription initiation\\
+%1e-03 &  31 & GO:0006814 & sodium ion transport\\
+%3e-03 &  86 & GO:0006310 & DNA recombination\\
+%6e-03 & 197 & GO:0003676 & nucleic acid binding\\
+%\hline 
+%\end{tabular}
+%\par\end{centering}
+%\end{table}
+%\clearpage{}
 %=============================================================================
@@
 SPY and SDE.}
-\end{figure}
-\clearpage{}%
-%=============================================================================
-
-%=============================================================================
-% Figure S12: 
-\begin{figure}
-
-\subfloat[]{\includegraphics[scale=0.4]{figures/h3-R-ratio-topology-changing}}
-
-\subfloat[]{\includegraphics[scale=0.4]{figures/h3-R-ratio-not-topology-changing}}
-
-\caption{\label{fig:h3}Numer of recombination events. The diagonal dashed
-lines are with slope of 1 and intercept of 0. For each order
-pairs of branches of a reference species tree the true number of recombination
-events is computed as the mean of the 10 replicates of recombinant
-trees from the simulation. With ten estimates of each replicate total
-number of 100 estimats are shown as intervals of 90\% of quantile
-range, or the lower value of the interval is the 5\% quantile, and
-the upper one is the 95\% quantile. }
 \end{figure}
@@
 to neither lack of power of \texttt{ClonalOrigin} nor hidden bias thereof.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% 
+%(Table S\ref{tab:functional-sde-spy})
+%(Table S\ref{tab:functional-spy-sde})
+%(Table S\ref{tab:go-events-duplication})
+%(Table S\ref{tab:go-events-loss})
+%
+%
+%For horizontal gene transfer we found enrichment in very different categories
+%related to transposition, integration, and recombination (Table
+%\ref{tab:go-events}).  
+%
+%Gene duplication and loss events were enriched for
+%categories similar to those of horizontal gene transfer's enrichment (Tables
+%S\ref{tab:go-events-duplication} and S\ref{tab:go-events-loss}).
+%
+%Gene duplication and horizontal transfers showed enrichment in the same
+%categories perhaps because both mechanisms increase gene copy number.  
+%
+%Gene losses showed enrichment for transposition and integration, but less
+%significantly (P-value of $<10^{-6}$), and showed enrichment for several
+%additional categories such as transporter activity and transcription initiation.
+%
+%Comparing two types of recombinant edges, one from SDE to SPY and another from
+%SPY to SDE transport related categories were more associated for the direction
+%from SPY to SDE
+%(Table S\ref{tab:functional-sde-spy} and S\ref{tab:functional-spy-sde}).
+%
+%\texttt{More interpretations about Mawgli analysis of function categories?
+%Why transposases?}
+
+% \texttt{IG: NEXT PARAGRAPH CONTAINS SOME DISCUSSION ON THESE RESULTS. WE SHOULD 
+% GET BACK TO THIS ONCE WE DECIDE WHAT TO DO WITH THE RESULTS}
+
+%The parsimony-based approach also showed
+%enrichment of the recombining transfers in many of the same categories such as
+%translation, ribosome, and rRNA binding when using all types of recombinant
+%edges in computing recombination intensity (see Tables
+%S\ref{tab:functional-all} and S\ref{tab:go-events-recombining}).  
+%Because the
+%parsimony-based method did not take species tree branch lengths into account, 
+%the agreement between the two different methods might be likely.

### 2011-10-14 sh/batch.sh

@@ -41,8 +41,13 @@ function batch {
   batch-summary-coii
   batch-summary-coii-r
   batch-rmessage
+  batch-sim-three
+  batch-sim-prior
+  batch-sim-posterior
   scp -q output/cornell5/* cac:run/mauve/101011
-  scp species/cornell5 cac:run/mauve/101011
+  scp -q species/cornell5 cac:run/mauve/101011
+  scp -q data/$SIM1INBLOCK cac:run/mauve/101011
+  scp -q data/$SIM1TREE cac:run/mauve/101011
 }
 
 function batch-variable {
@@ -92,6 +97,10 @@ function batch-speciesfile {
   COIIWALLTIME=$(grep ^COIIWALLTIME\: $SPECIESFILE | cut -d":" -f2)
   AOIWALLTIME=$(grep ^AOIWALLTIME\: $SPECIESFILE | cut -d":" -f2)
   AOIIWALLTIME=$(grep ^AOIIWALLTIME\: $SPECIESFILE | cut -d":" -f2)
+  SCO1WALLTIME=$(grep ^SCO1WALLTIME\: $SPECIESFILE | cut -d":" -f2)
+  SIM1WALLTIME=$(grep ^SIM1WALLTIME\: $SPECIESFILE | cut -d":" -f2)
+  SIM2WALLTIME=$(grep ^SIM2WALLTIME\: $SPECIESFILE | cut -d":" -f2)
+  SIM3WALLTIME=$(grep ^SIM3WALLTIME\: $SPECIESFILE | cut -d":" -f2)
   LCBMINLEN=$(grep ^LCBMINLEN\: $SPECIESFILE | cut -d":" -f2)
   NUMBERSPECIES=$(grep ^NUMBERSPECIES\: $SPECIESFILE | cut -d":" -f2)
   COIHOWMANYNODE=$(grep ^COIHOWMANYNODE\: $SPECIESFILE | cut -d":" -f2)
@@ -114,6 +123,18 @@ function batch-speciesfile {
   GBK=$(grep ^GBK\: $SPECIESFILE | cut -d":" -f2)
   TREETOPOLOGY=$(grep ^TREETOPOLOGY\: $SPECIESFILE | cut -d":" -f2)
   REFGENOMELENGTH=$(grep ^REFGENOMELENGTH\: $SPECIESFILE | cut -d":" -f2)
+
+  # For simulation 1
[…]

### 2011-10-14 sh/simulate-data-clonalorigin1-prepare.sh

@@ -1,3 +1,21 @@
+###############################################################################
+# Copyright (C) 2011 Sang Chul Choi
+#
+# This file is part of Mauve Analysis.
+# 
+# Mauve Analysis is free software: you can redistribute it and/or modify
+# it under the terms of the GNU General Public License as published by
+# the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# Mauve Analysis is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with Mauve Analysis.  If not, see <http://www.gnu.org/licenses/>.
+###############################################################################
 
 # 11. Prepare the first stage of clonalorigin for simulation.
 # -----------------------------------------------------------
@@ -64,6 +82,7 @@
 #
 # Finishing the above I make a list of jobs over all the repeats. I also create
 # the main batch script for the jobs of all the repeats.
+
 function simulate-data-clonalorigin1-prepare {
   PS3="Choose a menu of simulation with clonalorigin: "
   select SPECIES in ${SIMULATIONS[@]}; do 

### 2011-10-15 doc/manuscript.tex
Just an update.

@@
-% We approach the project of streptococcus recombination and gene transfer from
-% the angle of a resource paper. 
 \documentclass[english]{article}
@@
 
-\subsubsection*{Version: 0.67 (Oct 7, 2011) \clearpage{}}
+\subsubsection*{Version: 0.68 (Oct 15, 2011) \clearpage{}}
 
@@
 evolutionary histories of prokaryotes, with multiple individual genomes per
-species we now can go beyond the rough broad study on gene transfer towards
+species we now can go beyond such rough broad study on gene transfer towards
 fine-scaled history of gene transfer between very closely related prokaryotes.
@@
 
-In short, the complete genomes of the two spcies of SPY and SDE presented us a
+In short, the complete genomes of the two species of SPY and SDE presented us a
 unique opportunity to study gene transfer over relatively recent time scales.
@@
 both of the results in recombining gene transfer events to show that the
-parimony-based method inferred recombining gene transfer events in biological
+parsimony-based method inferred recombining gene transfer events in biological
 functional categories similar to those inferred with the model-based approach.
@@
 where SDD was the outgroup, and SPY and SDE were sister clades. Therefore, SPY
-and SDE appeared be more closely related with each other than each to SDD's
+and SDE appeared to be more closely related with each other than each to SDD's
 genome in about tenth of the core genome. The other topologies had less than 4\%
@@
 S\ref{tab:heatmap} and S\ref{tab:obsheatmap} for the heat map values and the
-observed number of recombinant edges) shows the logarithm base 2 ratios of the
+observed number of recombinant edges) shows the logarithm base 2 fractions of the
 \textit{a posteriori} number of recombinant edges for a pair of donor and
@@
 strictly veterinary pathogenic bacteria, we did not expect that SDD received
-much gene transfer from either SPY1 or SPY2: the logarithm base 2 ratio from SPY1 to
+much gene transfer from either SPY1 or SPY2: the logarithm base 2 fraction from SPY1 to
 SDD was -2.20, and that from SPY2 to SDD was -1.97.  On the contrary, the
@@
 or SDE2 was less frequent than the case of reverse direction; SDD effectively
-served as a recipient rather than a donor. It was not unclear why SDD received
+served as a recipient rather than a donor. It was unclear why SDD received
 more genes from SDE1 or SDE2 than did SDE1 or SDE2 from SDD.
@@
 be more frequent than that between clades, not surprisingly the recombining gene
-transfer logarithm base 2 ratios between SDE1 and SDE2 (1.17 and 1.20) were about
+transfer logarithm base 2 fractions between SDE1 and SDE2 (1.17 and 1.20) were about
 twice as large as those from SDE1 and SDE2 to SPY1 and SPY2 (0.67, 0.66, 0.60,
@@
 or SDE2 to one of the other three external branches. Surprisingly, the recombining
-gene transfer logarithm base 2 ratios between SPY1 and SPY2 (1.06 and 1.07) was
+gene transfer logarithm base 2 fractions between SPY1 and SPY2 (1.06 and 1.07) was
 comparable with, yet even less than those from SPY1 and SPY2 to SDE1 and SDE2
@@
 despite their dissimilar genomic organization.  The gene transfer logarithm base 2
-ratio from SPY1 and SPY2 to SDE branch was also larger than any type of
+fraction from SPY1 and SPY2 to SDE branch was also larger than any type of
 recombining gene transfer: 1.88 and 1.94, respectively.  Additionally we found
@@
 that did not change species tree topology might be difficult to infer reliably,
-considering those recombinant edges in computing recombinantion intensity might
+considering those recombinant edges in computing recombination intensity might
 lead to dubious results.  Ribosome related functional categories were found to
@@
 
-1. Topology-not-changing recombinant edge was unreliably estimated compared with
-topology-changing recombinant edge.
-2. Balanced case
-3. Unbalanced case
-
 We conducted experiments on simulation data to check the accuracy of estimates
-obtained by the model-based approach in a scenario similar to one from the real 
-data set.
-
-While populaiton mutation and
-recombination parameters were accurately estimated, the recombinant tract length
-was over-estimated  (Table S\ref{tab:sim-three-population-parameters}).  
-Recombinant tract lengths were reported to be difficult to estimate from relatively
-short alignment blocks \citep{Didelot2010}.
-
-Figure S\ref{fig:h1} shows the observed count of each recombinant
-edge type against the expected count; the estimated numbers appeared to be
-slightly over-estimated, but in correlation with the true numbers.
-
-We also evaluated the accuracy of the statistics estimated in the second stage
-of \texttt{ClonalOrigin}.
-
-The number of recombinant edges for each pair of
-recombinant tree branches was correlated with the true value in general while
-there were slight over-estimation (Figure \ref{fig:h3}).
-
-However, this
-comparison between inferrence from simulated data and that of real data set
-suggested that the directional difference between SPY and SDE was due to neither
-lack of power of \texttt{ClonalOrigin} nor hidden bias thereof.
-
-Recombination
-intensities that were measured on genes also showed rough correlation (Pearson's
-correlation coefficient of 0.63) of the estimated and true values (Figure
-S\ref{fig:ri1}).
-
-Estimated values appeared to be driven somewhat by the prior
-implied by the population parameter setting, explaining why lower values were
-over-estimated and higher values were under-estimated.
-
-Nonetheless, the
-correlation indicated that the measure of recombination intensity used in our
-functional gene category served as an adequate indicator of actual intensity of
-recombination events along genes.
+obtained by the model-based approach in a scenario similar to the one from the
+real data set.  Population mutation and recombination parameters were accurately
+estimated (Table S\ref{tab:sim-three}).  Yet, the recombinant tract length was
+over-estimated; recombinant tract lengths were reported to be difficult to
+estimate from relatively short alignment blocks \citep{Didelot2010}.  Numbers of
+recombinant edges from the simulated recombinant trees of Simulation 2 (see
+\textbf{Supplementary Information} for the description of Simulation 2) were
+compared with the corresponding numbers of recombinant edges that were estimated
+from the simulated data set from Simulation 2 (see the solid-line intervals with
+open circles in Figure \ref{fig:h3}).  We noted that tree-topology-changing
+recombinant edges were more reliably recovered compared with
+tree-topology-not-changing ones (Figure \ref{fig:h3}). All of the
+tree-topology-not-changing recombinant edge types were over-estimated (Figure
+\ref{fig:h3}B). The heat map of Figure \ref{fig:Heatmap-of-recombination} did
+not show how confident of the values we could be. Comparison of numbers of
+recombinant edges estimated from the real data set and the corresponding numbers
+estimated from Simulation 2 showed the interval estimates for the heat map
+values.  Dotted-line intervals from the red and blue colored cells in the heat
+map were above and below the reference of the horizontal line at the zero on the
+Y-axis, respectively (Figure \ref{fig:h3}).
+
+for each recombinant edge type to assess the effect of posterior recombinant
+trees samples with respect to that of prior recombinant tree samples. The ten
+\textit{a priori} recombinant trees from Simulation 2 would have balanced
+recombining gene transfer events between SPY and SDE with respect to the prior
+expected number of recombinant edges. The posterior recombinant trees used in
+Simulation 3 would have unbalanced recombining gene transfer between the two
+species branches as we observed in the real data analysis. For significantly
+strong gene flows such as ones from SPY to SDE a solid-line interval with a
+closed circle would be different from its count part of the dotted-line
+interval.  Figure \ref{fig:h3} showed that the recombining gene flow from SPY to
+SDE was stronger enough to be nearer to the horizontal line at the zero on the
+Y-axis. However, the recombining gene transfer from SDE to SPY appeared to be
+too weak.  This comparison supported the biased directionality of recombining
+gene transfer from SPY to SDE.
+
+Recombination intensities that were measured on genes also showed rough
+correlation (Pearson's correlation coefficient of 0.63) of the estimated and
+true values (Figure S\ref{fig:ri1}).  Estimated values appeared to be driven
+somewhat by the prior implied by the population parameter setting, explaining
+why lower values were over-estimated and higher values were under-estimated.
+Nonetheless, the correlation indicated that the measure of recombination
+intensity used in our functional gene category served as an adequate indicator
+of actual intensity of recombination events along genes.
 
@@
 and estimated the species tree relating the five genomes. \texttt{ClonalFrame}
-also estimates the ratio between recombination and mutation rates.
+also estimates the fraction between recombination and mutation rates.
 We used a single Markov chain Monte Carlo (MCMC) with 10000 burn-in generations
@@
 enriched for certain types of gene transfer events. Functional categories were
-assigned in the same manner as \citet{Suzuki2011}. In short, \texttt{blastP} 
-was used to
-compare Streptococcus genes to bacterial proteins from the Uniref90 database.
-Matches with E-values less than $1.0\times10^{-5}$ were assumed to have the same
-Gene Ontology (GO) classification as the target gene, as determined from the
-uniProt GOA database.  We also classified a category of  \textit{virulent genes}
-using the virulence profile described in \citet{Suzuki2011}. Among 207 virulent
-genes at least one of the genes in the gene family was identified by that
-profile.  Each gene covered by our analysis was associated with a recombination
-intensity score from the \texttt{ClonalOrigin} analysis and scores from the
-\texttt{Mowgli} analysis corresponding to number of duplications, losses,
-horizontal gene transfers, and recombining gene transfer events.  Mann-Whitney
-tests were used to identify GO categories whose genes showed significantly
-elevated values in each score.  A 5\% False Discovery Rate (FDR) correction
-\citep{Benjamini1995} was used to determine an appropriate significance
-threshold for our p-values.
+assigned in the same manner as \citet{Suzuki2011}. In short, \texttt{blastP} was
+used to compare Streptococcus genes to bacterial proteins from the Uniref90
+database.  Matches with E-values less than $1.0\times10^{-5}$ were assumed to
+have the same Gene Ontology (GO) classification as the target gene, as
+determined from the uniProt GOA database.  We also classified a category of
+\textit{virulent genes} using the virulence profile described in
+\citet{Suzuki2011}. Among 207 virulent genes at least one of the genes in the
+gene family was identified by that profile.  Each gene covered by our analysis
+was associated with a recombination intensity score from the
+\texttt{ClonalOrigin} analysis and scores from the \texttt{Mowgli} analysis
+corresponding to number of duplications, losses, horizontal gene transfers, and
+recombining gene transfer events.  Mann-Whitney tests were used to identify GO
+categories whose genes showed significantly elevated values in each score.  A
+5\% False Discovery Rate (FDR) correction \citep{Benjamini1995} was used to
+determine an appropriate significance threshold for our p-values.
 
 \subsection{Simulation Study}
+
 We performed three sets of simulations.  In the first simulation, we wished to
@@
 species tree with no particular directions of recombining gene transfers between
-species tree branches.  We computed the ratio of the mean number of recombinant
+species tree branches.  We computed the fraction of the mean number of recombinant
 edges for a pair of donor and recipient species tree branches over the simulated
 10 recombinant trees against the median of recombinant edges estimated from the
-10 simulated data sets, and then took the logarithm of the ratio to base 2. We
+10 simulated data sets, and then took the logarithm of the fraction to base 2. We
 also computed 5\% and 95\% quantiles to find the 90\% quantile range for the log
-base 2 ratio.  We expect that the log base 2 ratio would near zero. We also
+base 2 fraction.  We expect that the log base 2 fraction would near zero. We also
 wished to see if how the recombinant edge count estimates from the real data set
@@
 This consisted of 100 replicates of data sets.  
-We follow the same computation of the ratio as in the simulation 2 with 
+We follow the same computation of the fraction as in the simulation 2 with 
 the numbers of recombinant edges estimated from the real data
@@
 
+% \clearpage{}
+
+% \section*{Tables}
+
 \clearpage{}
 
-\section*{Tables}
-
-\clearpage{}
-
+%=============================================================================
+% Table 1. 
 \begin{table}
@@
 \end{table}
-\clearpage{}
-
-%
+%\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% Table 2. 
 \begin{table}
@@
 \clearpage{}
-
-\section*{Figures\clearpage{}}
+%=============================================================================
+
+%\section*{Figures\clearpage{}}
 
@@
 %=============================================================================
-% Figure S12: 
+% Figure 4. 
 \begin{figure}
@@
 
-\caption{\label{fig:h3}Numer of recombination events. The diagonal dashed
-lines are with slope of 1 and intercept of 0. For each order
-pairs of branches of a reference species tree the true number of recombination
-events is computed as the mean of the 10 replicates of recombinant
-trees from the simulation. With ten estimates of each replicate total
-number of 100 estimats are shown as intervals of 90\% of quantile
-range, or the lower value of the interval is the 5\% quantile, and
-the upper one is the 95\% quantile. }
+\caption{\label{fig:h3}
+Simulations of the model-based approach for assessing the power of recovering
+recombinant edges. Pairs of species tree branches are shown at x-axis labels.
+Three intervals are displayed for each pair. Intervals with solid line with open
+circle in the middle are for the logarithm of the fraction to base 2 of the average
+number of recombinant edges from the ten simulated recombinant trees 
+to the numbers of recombinant edges estimated from Simulation 2. 
+Interverals with dotted line are for the logarithm of the fraction to base 2 of
+the number of recombinant edges estimated from the real data
+to the numbers of recombinant edges estimated from Simulation 2.
+Intervals with solid line with closed circle in the middle are for the logarithm
+of the fraction to base 2 of the number of recombinant edges estimated from the real data
+to the numbers of recombinant edges estimated from Simulation 3. 
+The horizontal lines are at the zero.
+The intervals range from 5\% quantile to 95\% of quantile.}
 \end{figure}
@@
 
-\section*{Supplementary Tables}
-\clearpage{}
+% \section*{Supplementary Tables}
+% \clearpage{}
 
@@
 \end{table}
-\clearpage{}
+%\clearpage{}
 %=============================================================================
@@
 \begin{table}
-\caption{\label{tab:sim-three-population-parameters}
+\caption{\label{tab:sim-three}
 Simulation study of the three population parameters.}
@@
 \end{table}
-\clearpage{}
-%=============================================================================
-
-\section*{Supplementary Figures}
+%\clearpage{}
+%=============================================================================
+
+%\section*{Supplementary Figures}
 

### 2011-10-17 doc/manuscript_plos.tex
A PLoS template is added with the same content as one in manuscript.tex.

--- 2011-10-15 doc/manuscript.tex
@@
-\documentclass[english]{article}
-\usepackage[T1]{fontenc}
-\usepackage[latin9]{inputenc}
-\usepackage[letterpaper]{geometry}
-\geometry{verbose,tmargin=1in,bmargin=1in}
-\setlength{\parskip}{\bigskipamount}
-\setlength{\parindent}{0pt}
+% Four potential reviewers whose e-mail addresses you can provide
+% Claus O. Wilke
+% Debra Bessen
+% Bruce R. Levin
+% Michael Wessels
+
+% Template for PLoS
+% Version 1.0 January 2009
+%
+% To compile to pdf, run:
+% latex plos.template
+% bibtex plos.template
+% latex plos.template
+% latex plos.template
+% dvipdf plos.template
+
+\documentclass[10pt]{article}
+
+% amsmath package, useful for mathematical formulas
+\usepackage{amsmath}
+% amssymb package, useful for mathematical symbols
+\usepackage{amssymb}
+
+% graphicx package, useful for including eps and pdf graphics
+% include graphics with the command \includegraphics
+\usepackage{graphicx}
+
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
+
+\usepackage{color}
+
+% Use doublespacing - comment out for single spacing
+\usepackage{setspace} 
+\doublespacing
+
+% FIXME: Remove this package when submitting the manuscript.
 \usepackage{url}
-\usepackage{graphicx}
-\usepackage{setspace}
-\usepackage[authoryear]{natbib}
-\usepackage{longtable}
-\doublespacing
-
-\newcommand{\chck}{\textbf{[CHECK]}}
-\newcommand{\addfig}{\textbf{[ADD FIGURE]}}
-\newcommand{\tba}{\textbf{[TO BE ADDED]}}
-
+
+% Text layout
+\topmargin 0.0cm
+\oddsidemargin 0.5cm
+\evensidemargin 0.5cm
+\textwidth 16cm 
+\textheight 21cm
+
+% Bold the 'Figure #' in the caption and separate it with a period
+% Captions will be left justified
+\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
+
+% Use the PLoS provided bibtex style
+\bibliographystyle{plos}
+
+% Remove brackets from numbering in List of References
 \makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-%% Because html converters don't know tabularnewline
+\renewcommand{\@biblabel}[1]{\quad#1.}
+
+% FIXME: Remove the followings if you have to when submitting the manuscript.
+\let\citep\cite
+\let\citet\cite
 \providecommand{\tabularnewline}{\\}
-%% A simple dot to overcome graphicx limitations
 \newcommand{\lyxdot}{.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
-\newcommand{\lyxaddress}[1]{
-\par {\raggedright #1
-\vspace{1.4em}
-\noindent\par}
-}
-
 \@ifundefined{showcaptionsetup}{}{%
@@
 \usepackage{subfig}
+
 \makeatother
 
-\usepackage{babel}
+% Leave date blank
+\date{}
+
+\pagestyle{myheadings}
+%% ** EDIT HERE **
+
+%% ** EDIT HERE **
+%% PLEASE INCLUDE ALL MACROS BELOW
+%% END MACROS SECTION
 
 \begin{document}
-\renewcommand\refname{Literature Cited}
-
-\raggedright
-
-\subsubsection*{Title: }
-
-Inferring homologous recombination and horizontal gene transfer 
+
+% Title must be 150 characters or less
+\begin{flushleft}
+{\Large
+\textbf{Inferring homologous recombination and horizontal gene transfer 
 using whole genomes of \textit{Streptococcus pyogenes} 
-and\textit{ S.\ dysgalactiae} subspecies \textit{equisimilis}
-
-\subsubsection*{Authors:}
-
-Sang Chul Choi$^{1}$, Matthew D. Rasmussen$^{1}$, Melissa Jane Hubisz$^{1}$, Ilan 
-Gronau$^{1}$,
-Michael J. Stanhope$^{2}$, and Adam Siepel$^{1}$
-
-\subsubsection*{Affiliations:}
-
-\lyxaddress{$^{1}$Department of Biological Statistics and Computational Biology,
-Cornell University, Ithaca, NY 14853, USA.}
-
-\lyxaddress{$^{2}$Department of Population Medicine and Diagnostic Sciences,
-College of Veterinary Medicine, Cornell University, Ithaca, NY 14853,
-USA.}
-
-\subsubsection*{Author for Correspondence:}
-
-Adam Siepel, Department of Biological Statistics and Computational
-Biology, Cornell University, Ithaca, NY 14853, USA, Phone: 607-254-1157,
-Fax: 607-255-4698, Email: acs4@cornell.edu\clearpage{}
-
-\subsubsection*{Abbreviations:}
-
-group A streptococci (GAS), group C streptococci (GCS), group G streptococci
-(GGS), \textit{Streptococcus pyogenes} (SPY), \textit{S.\ dysgalactiae}
-subspecies \textit{dysgalactiae} (SDD), \textit{S.\ dysgalactiae} subspecies
-\textit{equisimilis} (SDE)
-
-\subsubsection*{Keywords:}
-
-bacterial recombination, gene duplication and transfer, ClonalOrigin,
-ClonalFrame, Mowgli
-
-\subsubsection*{Version: 0.68 (Oct 15, 2011) \clearpage{}}
-
+and\textit{ S.\ dysgalactiae} ssp.\textit{\ equisimilis}}
+}
+% Short Title (running head): Gene transfer using Streptococcus genomes
+% Insert Author names, affiliations and corresponding author email.
+\\
+Sang Chul Choi$^{1}$, Matthew D. Rasmussen$^{1}$, 
+Melissa Jane Hubisz$^{1}$, Ilan Gronau$^{1}$,
+Michael J. Stanhope$^{2}$, and Adam Siepel$^{1,\ast}$
+\\
+\bf{1} 
+Department of Biological Statistics and Computational Biology,
+Cornell University, Ithaca, NY 14853, USA
+\\
+\bf{2}
+Department of Population Medicine and Diagnostic Sciences,
+College of Veterinary Medicine, Cornell University, Ithaca, NY 14853, USA
+\\
+$\ast$ E-mail: acs4@cornell.edu, Phone: 607-254-1157, Fax: 607-255-4698
+\end{flushleft}
+
+% Please keep the abstract between 250 and 300 words
 \section*{Abstract}
@@
 
-\clearpage{}
-
-\section{Introduction}
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLoS ONE authors please skip this step. 
+% Author Summary not valid for PLoS ONE submissions.   
+\section*{Author Summary}
+
+\textit{Streptococcus pyogenes} (SPY) has afflicted people around the globe
+causing many diseases. Its closely related species, \textit{S.\ dysgalactiae}
+subspecies \textit{equisimilis} (SDE), used to be known as veterinary pathogens
+have been reported to share disease profiles with SPY.  While gene transfers
+between the two species were shown in small portions of genomic loci, no study
+on the gene transfer at genome-scale has been performed because of lack of
+methodologies implemented or genomic data available. Recent advances in both
+method development and genomic data allowed to study gene transfer between the
+two species in genome scale.  Recipient bacteria of transferred genes can
+substitute transferred genes for homologous parts of their genomes, or integrate
+transferred genes into their genomes without replacement.  The former is called
+recombining gene transfer, and the latter horizontal gene transfer. By
+discriminating the two processes, We found that virulence genes were more
+associated with horizontal gene transfer than recombining gene transfer.  We
+also found a signficant signal of recombining gene transfer for the net
+direction from SPY to SDE in the core genome.  Our study provides new insights
+into gene transfer between SPY snd SDE in genome scale.
+
+\section*{Introduction}
 
@@
 life-threatening streptococcal diseases to humans traditionally associated with
-GAS \citep[e.g.,][]{Brandt2009}.  Because of the overlap in disease profile and
+GAS \citep{Brandt2009}.  Because of the overlap in disease profile and
 ecological niche of tissue sites betwen SPY and SDE, these bacteria may as
@@
 Therefore, the research topic on gene exchange between SPY and SDE has been
-performed in various ways. \citet{Sachse2002} showed gene transfer from SDE to
+performed in various ways. Sachse et al.\ \citet{Sachse2002} showed gene transfer from SDE to
 SPY for the downstream of the genomic region of streptococcal pyogenic
-exotoxins. \citet{Davies2007} investigated phage 3396 from SDE as a vehicle of
-gene transfer from SDE to SPY. \citet{Davies2009} demonstrated that an
+exotoxins. Davies et al.\ \citet{Davies2007} investigated phage 3396 from SDE as a vehicle of
+gene transfer from SDE to SPY. They also \citet{Davies2009} demonstrated that an
 integrative conjugative element of SDE allowed gene transfer from SDE to other
-streptococci.  \citet{Ahmad2009} did not conclusively deduce the net
+streptococci. Ahmad et al.\ \citet{Ahmad2009} did not conclusively deduce the net
 directionality of housekeeping gene transfer between SPY and SDE.  Notably,
-\citet{Kalia2001} attempted to show that the net direction of gene transfer from
+Kalia et al.\ \citet{Kalia2001} attempted to show that the net direction of gene transfer from
 SPY to SDE was more dominant than the other direction.  Because these researches
@@
 fine-scaled history of gene transfer between very closely related prokaryotes.
-For instance, \citet{Liu2009} studied gene transfer events in
+For instance, Liu et al.\ \citet{Liu2009} studied gene transfer events in
 \textit{Lactobacillus bulgaricus} and \textit{Streptococcus thermophilus} using
-the compositional approach of \citet{Karlin2001}. \citet{Luo2011} investigated
+the compositional approach of Karlin \citet{Karlin2001}. Luo et al.\ \citet{Luo2011} investigated
 gene transfer events in \textit{Eschericha Coli} using the embedded quartet
-decomposition due to \citet{Zhaxybayeva2006}.  \citet{Hamady2006} combined
+decomposition due to Zhaxybayeva et al.\ \citet{Zhaxybayeva2006}. Hamady et al.\ \citet{Hamady2006} combined
 phylogenetic and compositional approaches to detecting horizontal gene transfer.
-\citet{Caro-Quintero2011} studied a number of \textit{Shewanella baltica}
+Caro-Quintero et al.\ \citet{Caro-Quintero2011} studied a number of \textit{Shewanella baltica}
 genomes to identify the genetic elements that enabled the species to adapt to
 redox gradients.  Although several complete genomes from SPY have been
-available, \citet{Suzuki2011} just recently added one more complete genome of
+available, Suzuki et al.\ \citet{Suzuki2011} just recently added one more complete genome of
 SDE to the set of a previously unique SDE genome \citep{Shimomura2011}. These
@@
 Instead, a parsimony-based approach to inferring gene duplication, loss, and
-transfer events was developed by \citet{Doyon2011}.  We noted that the first
+transfer events was developed by Doyon et al.\ \citet{Doyon2011}.  We noted that the first
 approach could be used to infer transfer of genomic segments due to homologous
@@
 
-\section{Results}
-
-\subsection{Recombination drived more substitutions than did point mutation}
+% Results and Discussion can be combined.
+\section*{Results}
+
+\subsection*{Recombination drived more substitutions than did point mutation}
 
@@
 \texttt{ClonalOrigin} method (see section \textbf{Materials and Methods} and
-\citealt{Didelot2010}).  The relative strength ($r/m$) of recombination compared
+\citet{Didelot2010}).  The relative strength ($r/m$) of recombination compared
 to mutations indicated that recombination-driven substitutions affected more
@@
 
-\subsection{Recombination hotspots were observed in genomes}
+\subsection*{Recombination hotspots were observed in genomes}
 
@@
 
-\subsection{SPY and SDE were more closely related in about tenth of the core genome}
+\subsection*{SPY and SDE were more closely related in about tenth of the core genome}
 
@@
 
-\subsection{Unbalaned recombining gene flow from SPY to SDE in genome-wide scale}
+\subsection*{Unbalaned recombining gene flow from SPY to SDE in genome-wide scale}
 
@@
 
-\subsection{Reduction of genes and divergence of SPY and SDE species}
+\subsection*{Reduction of genes and divergence of SPY and SDE species}
 
@@
 
-\subsection{Virulence genes via recombining gene transfer}
+\subsection*{Virulence genes via recombining gene transfer}
 
@@
 
-\subsection{Association of virulence genes and horizontal gene transfer}
+\subsection*{Association of virulence genes and horizontal gene transfer}
 
@@
 
-\subsection{Model-based approach or parsimony-based one for recombining gene
+\subsection*{Model-based approach or parsimony-based one for recombining gene
 transfer study}
@@
 
-\subsection{UCSC genome browser tracks for bacterial gene transfer study}
+\subsection*{UCSC genome browser tracks for bacterial gene transfer study}
 
@@
 
-\subsection{Units of recombining gene transfer}
+\subsection*{Units of recombining gene transfer}
 
@@
 
-\subsection{Transporters, transferases, aminoacyl tRNA synthetases, and
+\subsection*{Transporters, transferases, aminoacyl tRNA synthetases, and
 transposases}
@@
 
-\subsection{Simulations}
+\subsection*{Simulations}
 
@@
 
-\section{Discussions}
+\section*{Discussion}
 
@@
 were the barriers \citep{Thomas2005} to cause the unbalanced gene transfer in
-the species?  \citet{Kalia2001} discussed restriction-modification system (RMS)
+the species?  Kalia et al.\ \citet{Kalia2001} discussed restriction-modification system (RMS)
 as a possible mechanism of the unbalanced gene flow.  Restriction-modification
@@
 prokaryote world.  Biological roles of RMS might be to maintain and control
-species identity \citep{Jeltsch2003}.  Recently, \citet{Budroni2011a} studied
+species identity \citep{Jeltsch2003}.  Recently, Budroni et al.\ \citet{Budroni2011a} studied
 that \textit{Neisseria meningitidis} phylogenetic clades were associated with
@@
 
-We reexamined the seven genes that \citet{Kalia2001} had studied for net
+We reexamined the seven genes that Kalia et al.\ \citet{Kalia2001} had studied for net
 directionality of gene transfer between SPY and SDE.  The internal fragments of
@@
 \textit{mutS}, \textit{recP}, \textit{xpt}) were located in the core genome that
-we investigated in this study.  Except for \textit{mutS} where \citet{Kalia2001}
+we investigated in this study.  Except for \textit{mutS} where Kalia et al.\ \citet{Kalia2001}
 did not find recombination, our results were consistent with that of
-\citet{Kalia2001}.  \citet{Pinho2010} reported recombination in the gene
+Kalia et al.\ \citet{Kalia2001}.  Pinho et al.\ \citet{Pinho2010} reported recombination in the gene
 \textit{parC}, in which we also observed moderate recombination.
@@
 
-\section{Materials and Methods}
-
-\subsection{Complete bacterial genomes}
+% You may title this section "Methods" or "Models". 
+% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
+% authors may use "Analysis" 
+\section*{Materials and Methods}
+
+\subsection*{Complete bacterial genomes}
 
@@
 \textit{S.\ dysgalactiae} ssp.\textit{\ dysgalactiae} was provided by
-\citet{Suzuki2011} with accession identifiers of SddyATCC27957, referred to here
+Suzuki et al.\ \citet{Suzuki2011} with accession identifiers of SddyATCC27957, referred to here
 as SDD. We also downloaded the genome (NC\_012471) of \textit{S.\ equi}
@@
 
-\subsection{Model-based method for recombining gene transfer}
+\subsection*{Model-based method for recombining gene transfer}
 
 We present a short summary of the statistical method developed by
-\citet{Didelot2007} and \citet{Didelot2010} for studying bacterial
+Didelot et al.\ \citet{Didelot2007} and \citet{Didelot2010} for studying bacterial
 recombination. A detailed account was found in the two papers and in the user
@@
 employed independently for each of the 274 alignment blocks as described in
-\citet{Didelot2010}.   In the first MCMC we obtained estimates of mutation rate,
+Didelot et al.\ \citet{Didelot2010}.   In the first MCMC we obtained estimates of mutation rate,
 recombination rate, and recombinant tract length.  Two independent chains were
@@
 
-\subsection{Parsimony-based method for horizontal gene transfer}
+\subsection*{Parsimony-based method for horizontal gene transfer}
 
@@
 it attempted to explain different copy numbers by a series of gain and loss
-events.  We obtained gene family clusters defined in \citet{Suzuki2011} using
+events.  We obtained gene family clusters defined in Suzuki et al.\ \citet{Suzuki2011} using
 the \texttt{OrthoMCL} method \citep{Li2003} over the full set of the six genomes
@@
 
-\subsection{Functional categories association}
+\subsection*{Functional categories association}
 
@@
 enriched for certain types of gene transfer events. Functional categories were
-assigned in the same manner as \citet{Suzuki2011}. In short, \texttt{blastP} was
+assigned in the same manner as Suzuki et al.\ \citet{Suzuki2011}. In short, \texttt{blastP} was
 used to compare Streptococcus genes to bacterial proteins from the Uniref90
@@
 \textit{virulent genes} using the virulence profile described in
-\citet{Suzuki2011}. Among 207 virulent genes at least one of the genes in the
+Suzuki et al.\ \citet{Suzuki2011}. Among 207 virulent genes at least one of the genes in the
 gene family was identified by that profile.  Each gene covered by our analysis
@@
 
-\subsection{Simulation Study}
+\subsection*{Simulation Study}
 
@@
 
-\section{Supplementary Information}
-
-\subsection{Summarization of posterior samples for local gene trees and recombination intensity}
-
-Because the procedure of summarizing posterior samples of \textit{recombinant
-tree}, which will be defined in a moment, was not detailed in the user guide at
-the time when we used \texttt{ClonalOrigin}, we elected to use Figure
-S\ref{fig:clonalorigin} to illustrate a recombinant tree in order to clarify
-summarization of \textit{recombinant history} posterior samples.  A
-recombination history is represented by a \textit{recombinant tree}, which is a
-species tree augmented with \textit{recombinant edges}.  A recombinant edge
-connect two species tree branches.  Each recombinant edge is defined by the time
-and branch of origin and destination, and the genomic interval that it affects.
-The branch of origin is referred to as the \textit{donor} and the destination
-branch is referred to as the \textit{recipient}. The type of a recombinant edge
-is determined by the donor and recipient species tree branches.  A recombinant
-tree can be viewed as a restricted version of an ancestral recombination graph
-(ARG), and as such it induces a local genealogy for each site in the genomic
-block it describes.  \citet{Didelot2010} presented two procedures of summarizing
-recombinant trees by viewing trees from two different angles: by counting types
-of recombinant edges regardless of genomic positions affected by the edges, and
-by measuring recombinant edge changes to the species tree in site-wise manner.
-By similarly following their approaches we quantified additional aspects of
-recombinant trees; we counted gene tree topologies for each site, and also
-quantified recombination intensity along the genome.
-
-The recombination intensity allowed us to find genomic regions enriched for
-recombination events in general, or regions enriched for recombination events of
-certain types (e.g., between SPY and SDE).  We noted that the recombination
-intensity was not normalized to the interval of $[0,1]$. Rather than giving
-probabilistic meanings to the recombination intensity we interpreted it as an
-average number of recombinant edge types per site.  The measure appeared to act
-as an effective indicator of recombination intensity (see subsection of
-\textbf{Simulation Study}).  Our simulation study indicated that edges of
-certain types were better indicators of recombination events than others.  For
-instance, recombinant edges which resulted in a recombinant tree with the same
-topology as the species tree appeared to be more prone to spurious sampling.
-Such edges were ones where the donor was the same branch, sister branch, or
-parent branch of the recipient branch.  Therefore, in some cases we elected to
-remove such edges from consideration when measuring recombination intensity.  
-% The tree topologies would show more effectively regions where
-% recombination occurs along the genome (see subsection of \textbf{UCSC Genome
-% Browser}).
-
-\subsection{Simulation 1 - three population parameters}
-
-We describe a procedure of simulations for checking if the three
-population parameters were accurately recovered using the model-based approach.
-Ten recombinant trees were sampled given the three parameters (Table
-S\ref{tab:three}) and the species tree (Figure S\ref{fig:tree5}) set to the
-values inferred from the five bacterial genomes.  Each of the recombinant trees
-was used to simulate the Jukes-Cantor model for DNA sequence evolution to create
-274 blocks of alignments.  The block lengths configuration was the same as that
-of the blocks used in the analysis of the five genomes.
-
-% s15
-% FIXME: Newick tree in text in a file.
-% FIXME: Publish the result and the data set with the code.
-
-\subsection{Simulation 2 - numbers of recombinant edges from prior}
-
-We first sampled ten recombinant trees given the mutation rate, recombination
-rate, and tract length (Table S\ref{tab:three}) and the species tree (Figure
-S\ref{fig:tree5}), and then created ten independent simulated data sets with
-the Jukes-Cantor model for DNA sequence evolution using each of the recombinant
-trees. This consisted of 100 replicates of data sets.  The recombinant trees
-were created from the prior distribution and recombinant edges were added to the
-species tree with no particular directions of recombining gene transfers between
-species tree branches.  We computed the fraction of the mean number of recombinant
-edges for a pair of donor and recipient species tree branches over the simulated
-10 recombinant trees against the median of recombinant edges estimated from the
-10 simulated data sets, and then took the logarithm of the fraction to base 2. We
-also computed 5\% and 95\% quantiles to find the 90\% quantile range for the log
-base 2 fraction.  We expect that the log base 2 fraction would near zero. We also
-wished to see if how the recombinant edge count estimates from the real data set
-deviate from the prior. Instead of using the number of recombinant edges from
-the ten recombinant trees, we used the numbers of recombinant edges that were
-estimated from the real data.
-
-We tested if our measure of recombination intensity provided a reliable estimate
-of the true intensity. We artificially annotated our synthetic blocks to genes
-according to the real annotations from SPY1 genome.  For each gene in each of
-the simulated 100 data sets, we plotted the average estimated recombination
-intensity along that gene against the average number of different recombinant
-edge types from the ten simulated recombinant tree.
-
-% s16
-
-% s16
-% FIXME 1: IG: RESULTS FROM THIS POINT ON REQUIRE SOME REFINEMENT. SC: Let's see.
-% FIXME 2: ``TRUE" TO BE EXPECTED VALUE
-% RATHER THAN AVERAGE OVER THE 10 REC.  TREES. SC: OKAY. LET'S FIND THE PRIOR
-% EXPECTED NUMBER OF EVENTS. THIS MUST BE GIVEN SOMEWHERE}.  
-% FIXME 3: CONSIDER SHOWING DETAILED COMPARISON OF ESTIMATED VALUES TO ONES 
-% EXPECTED FROM THE ACTUAL REC. TREES. THESE DON'T LOOK AS NICE, BUT THEY MIGHT 
-% BE ABLE TO DEMONSTRATE OUR ABILITY TO DETECT DEPARTURE FROM EXPECTATION. SC:
-% THIS MAY BE ABOUT CHANGING AVERAGE TO PRIOR EXPECTED NUMBER OF EVENTS.
-% FIXME 4: WE SHOULD PROBABLY CONSIDER ALTERNATIVE WAYS OF USING THESE RESULTS TO 
-% ARGUE THAT OUR DIRECTIONAL SIGNAL IS REAL. SC: I MIGHT CONSIDER SIMULATIONS AND
-% ARTIFICIALLY DOUBLE OR TRIPLE THE RECOMBINATION EVENTS OF ONE OF TYPES.
-
-\subsection{Simulation 3 - number of recombinant edges from posterior}
-
-We used 100 recombinant trees sampled from the posterior distribution
-of the recombiant tree instead of using simulated recombinant trees.
-We created one simulated data set for each of the 100 recombinant trees with
-the Jukes-Cantor model for DNA sequence evolution 
-using the mutation rate estimate (Table S\ref{tab:three}).
-This consisted of 100 replicates of data sets.  
-We follow the same computation of the fraction as in the simulation 2 with 
-the numbers of recombinant edges estimated from the real data
-as numerators.
-
-% s17
-
+% Do NOT remove this, even if you are not including acknowledgments
 \section*{Acknowledgments}
@@
 Contributed reagents/materials/analysis tools: MJS.
-Wrote the paper: SCC MDR MJH IG.
-
-\bibliographystyle{gberefs}
+Wrote the paper: SCC MDR MJH IG AS.
+
+%\section*{References}
+% The bibtex filename
+\bibliographystyle{plos}
 \bibliography{siepel-strep}
 
-% \clearpage{}
-
-% \section*{Tables}
-
-\clearpage{}
+\section*{Figure Legends}
+%\begin{figure}[!ht]
+%\begin{center}
+%%\includegraphics[width=4in]{figure_name.2.eps}
+%\end{center}
+%\caption{
+%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
+%should be left justified, as specified by the options to the caption 
+%package.
+%}
+%\label{Figure_label}
+%\end{figure}
+
+\section*{Tables}
+%\begin{table}[!ht]
+%\caption{
+%\bf{Table title}}
+%\begin{tabular}{|c|c|c|}
+%table information
+%\end{tabular}
+%\begin{flushleft}Table caption
+%\end{flushleft}
+%\label{tab:label}
+% \end{table}
 
@@
 
+\section{Supplementary Information}
+
+\subsection{Summarization of posterior samples for local gene trees and recombination intensity}
+
+Because the procedure of summarizing posterior samples of \textit{recombinant
+tree}, which will be defined in a moment, was not detailed in the user guide at
+the time when we used \texttt{ClonalOrigin}, we elected to use Figure
+S\ref{fig:clonalorigin} to illustrate a recombinant tree in order to clarify
+summarization of \textit{recombinant history} posterior samples.  A
+recombination history is represented by a \textit{recombinant tree}, which is a
+species tree augmented with \textit{recombinant edges}.  A recombinant edge
+connect two species tree branches.  Each recombinant edge is defined by the time
+and branch of origin and destination, and the genomic interval that it affects.
+The branch of origin is referred to as the \textit{donor} and the destination
+branch is referred to as the \textit{recipient}. The type of a recombinant edge
+is determined by the donor and recipient species tree branches.  A recombinant
+tree can be viewed as a restricted version of an ancestral recombination graph
+(ARG), and as such it induces a local genealogy for each site in the genomic
+block it describes.  Didelot et al.\ \citet{Didelot2010} presented two procedures of summarizing
+recombinant trees by viewing trees from two different angles: by counting types
+of recombinant edges regardless of genomic positions affected by the edges, and
+by measuring recombinant edge changes to the species tree in site-wise manner.
+By similarly following their approaches we quantified additional aspects of
+recombinant trees; we counted gene tree topologies for each site, and also
+quantified recombination intensity along the genome.
+
+The recombination intensity allowed us to find genomic regions enriched for
+recombination events in general, or regions enriched for recombination events of
+certain types (e.g., between SPY and SDE).  We noted that the recombination
+intensity was not normalized to the interval of $[0,1]$. Rather than giving
+probabilistic meanings to the recombination intensity we interpreted it as an
+average number of recombinant edge types per site.  The measure appeared to act
+as an effective indicator of recombination intensity (see subsection of
+\textbf{Simulation Study}).  Our simulation study indicated that edges of
+certain types were better indicators of recombination events than others.  For
+instance, recombinant edges which resulted in a recombinant tree with the same
+topology as the species tree appeared to be more prone to spurious sampling.
+Such edges were ones where the donor was the same branch, sister branch, or
+parent branch of the recipient branch.  Therefore, in some cases we elected to
+remove such edges from consideration when measuring recombination intensity.  
+% The tree topologies would show more effectively regions where
+% recombination occurs along the genome (see subsection of \textbf{UCSC Genome
+% Browser}).
+
+\subsection{Simulation 1 - three population parameters}
+
+We describe a procedure of simulations for checking if the three
+population parameters were accurately recovered using the model-based approach.
+Ten recombinant trees were sampled given the three parameters (Table
+S\ref{tab:three}) and the species tree (Figure S\ref{fig:tree5}) set to the
+values inferred from the five bacterial genomes.  Each of the recombinant trees
+was used to simulate the Jukes-Cantor model for DNA sequence evolution to create
+274 blocks of alignments.  The block lengths configuration was the same as that
+of the blocks used in the analysis of the five genomes.
+
+% s15
+% FIXME: Newick tree in text in a file.
+% FIXME: Publish the result and the data set with the code.
+
+\subsection{Simulation 2 - numbers of recombinant edges from prior}
+
+We first sampled ten recombinant trees given the mutation rate, recombination
+rate, and tract length (Table S\ref{tab:three}) and the species tree (Figure
+S\ref{fig:tree5}), and then created ten independent simulated data sets with
+the Jukes-Cantor model for DNA sequence evolution using each of the recombinant
+trees. This consisted of 100 replicates of data sets.  The recombinant trees
+were created from the prior distribution and recombinant edges were added to the
+species tree with no particular directions of recombining gene transfers between
+species tree branches.  We computed the fraction of the mean number of recombinant
+edges for a pair of donor and recipient species tree branches over the simulated
+10 recombinant trees against the median of recombinant edges estimated from the
+10 simulated data sets, and then took the logarithm of the fraction to base 2. We
+also computed 5\% and 95\% quantiles to find the 90\% quantile range for the log
+base 2 fraction.  We expect that the log base 2 fraction would near zero. We also
+wished to see if how the recombinant edge count estimates from the real data set
+deviate from the prior. Instead of using the number of recombinant edges from
+the ten recombinant trees, we used the numbers of recombinant edges that were
+estimated from the real data.
+
+We tested if our measure of recombination intensity provided a reliable estimate
+of the true intensity. We artificially annotated our synthetic blocks to genes
+according to the real annotations from SPY1 genome.  For each gene in each of
+the simulated 100 data sets, we plotted the average estimated recombination
+intensity along that gene against the average number of different recombinant
+edge types from the ten simulated recombinant tree.
+
+% s16
+
+% s16
+% FIXME 1: IG: RESULTS FROM THIS POINT ON REQUIRE SOME REFINEMENT. SC: Let's see.
+% FIXME 2: ``TRUE" TO BE EXPECTED VALUE
+% RATHER THAN AVERAGE OVER THE 10 REC.  TREES. SC: OKAY. LET'S FIND THE PRIOR
+% EXPECTED NUMBER OF EVENTS. THIS MUST BE GIVEN SOMEWHERE}.  
+% FIXME 3: CONSIDER SHOWING DETAILED COMPARISON OF ESTIMATED VALUES TO ONES 
+% EXPECTED FROM THE ACTUAL REC. TREES. THESE DON'T LOOK AS NICE, BUT THEY MIGHT 
+% BE ABLE TO DEMONSTRATE OUR ABILITY TO DETECT DEPARTURE FROM EXPECTATION. SC:
+% THIS MAY BE ABOUT CHANGING AVERAGE TO PRIOR EXPECTED NUMBER OF EVENTS.
+% FIXME 4: WE SHOULD PROBABLY CONSIDER ALTERNATIVE WAYS OF USING THESE RESULTS TO 
+% ARGUE THAT OUR DIRECTIONAL SIGNAL IS REAL. SC: I MIGHT CONSIDER SIMULATIONS AND
+% ARTIFICIALLY DOUBLE OR TRIPLE THE RECOMBINATION EVENTS OF ONE OF TYPES.
+
+\subsection{Simulation 3 - number of recombinant edges from posterior}
+
+We used 100 recombinant trees sampled from the posterior distribution
+of the recombiant tree instead of using simulated recombinant trees.
+We created one simulated data set for each of the 100 recombinant trees with
+the Jukes-Cantor model for DNA sequence evolution 
+using the mutation rate estimate (Table S\ref{tab:three}).
+This consisted of 100 replicates of data sets.  
+We follow the same computation of the fraction as in the simulation 2 with 
+the numbers of recombinant edges estimated from the real data
+as numerators.
+
+% s17
+
 \end{document}
-
-\begin{figure}
-\includegraphics[scale=0.7]{figures/plot-number-recombination-within-blocks-1\lyxdot out\lyxdot recomb}
-\caption{\label{fig:recombination-boundaries}Distribution of number of 
-recombination
-events boundaries. The number of recombination event boundaries per
-site for each block along the \textit{S.\ dysgalactiae} ssp.\textit{\ equisimilis}
-ATCC 12394.}
-\end{figure}
-\clearpage{}%
-
-The mechanisms of horizontal gene transfer
-include conjugation by plasmid, transduction by bacteriophages, or
-transformation of naked DNA uptake \citep{Thomas2005}. Additionally,
-marine bacteria was reported to acquire foreign genes via gene transfer
-agents, i.e., viral-like particles produced by $\alpha$-Proteobacteria
-\citep{McDaniel2010}.
-
-Although
-recombinant edges between two species branches do not explicitly denote
-gene flow between the two species in their original model, \citet{Didelot2010}
-exploited their method to infer flows of genetic segment between \textit{Bacillus
-cereus} species. This model-based approach to detecting homologous
-recombining gene transfer was used to the core genomes of closely
-related \textit{Streptococcus} species.
-
-{[}WHERE SHOULD I PUT THESE UNIT OF GENE TRANSFER?{]}
-
-\textbf{Units of gene transfer:} \citet{Chan2009a} showed that protein
-domains were not gene transfer units. In another paper \citet{Chan2009}
-studied gene with and without breakpoints, and found that more often
-entire genes transfer in pathogens than in non-pathogens. In pathogenic
-bacteria they found more genes with breakpoints than genes without
-breakpoints.
-
-% Table and SDE and SPY
-Consistent
-with the number of recombinant edges were that gene transfer fragments
-with high posterior probability shown in Table S\ref{tab:genes-transfer}
-often appeared to be from SPY to SDE. While four genes were inferred
-to be transferred to SPY1, there were more genes transferred to SDE1
-or SDE2.
-
-\textbf{Evolutionary hypotheses about the ancestor of SPY and SDE:}
-\citet{Kalia2001} proposed two scenarios of the evolution of the
-two species: The first one was that the most recent common ancestor
-of SPY and SDE was a GAS-like pathogen, SDE was losing virulence factors
-since the divergence of the ancestor, and SDE regained virulence factors
-from SPY because of the same niche. The second scenario was that the
-most recent common ancestor of SPY and SDE was a GGS/GCS-like commensal,
-SPY evolved virulence factors since the divergence of the ancestor,
-and SDE gained virulence factors from SPY. Commensal bacteria could
-evolve by aquiring virulence genes from its closely related pathogenic
-bacteria. \citet{Kalia2001} suggested the divergence time of SDD
-to be around the time when human started to domesticate wildlife animals
-around 10,000 years ago. A demographic model could be made with two
-species to study the divergence time quantitatively. This would require
-ancestral recombination graphs embedded in a species tree.
-
-% From discussion
-This distinction of the two processes of
-gene transfer was envisioned a decade ago \citep[e.g.,][]{Ochman2001},
-and the methods of gene transfer detection by distinguishing the processes
-were not yet employed even if some of the methods were already available.
-We by no means provided any solutions to the problem of detecting
-gene transfer here. Yet, we hoped that we should be able to discriminate
-the two processes because each of the processes of gene transfer could
-affect biological functions that were different in nature.
-
-% From discussion
-For example, genes transferred via non-recombining
-gene transfer could be parts of functional units such
-as protein domains, and secondary structures of ribosomal RNAs. Genes
-transferred via homologous recombining gene transfer could be small
-fragments of genes as long as the flanking sequences of the fragments
-were more or less homologous.
-
-\textbf{Overlap of genes under positive selection and higher recombination:}
-As the uncertainty in gene tree topologies becomes larger, false positive
-rates of detecting sites under positive selection would be high (CITE).
-Genes could be divided into parts where we hoped no multiple gene
-tree topologies exist. Dividing genes into those fragments might be
-a way to avoid potentially high false positive rates of positively
-selected sites. Considering the fact that nucleotide substitutions
-seemed to be affected more by recombinations than by mutations we
-suspect that fragmenting genes might not be helpful.
-
-% Virulence gene and non-recombining gene transfer
-We investigated bacterial gene
-transfer between SPY and SDE using the genome sequences by comparing
-gene tree and species tree. We considered both core and pan genomes
-for detecting homologous recombining gene transfer and non-recombining
-gene transfer. 
-%
-The extent to which genes were transferred would vary on this
-mode of gene transfer. 
-While homologous
-recombination gene transfer would happen in the genes whose donor
-and recipient species were rather closely related, non-homologous
-gene transfer could happen virtually in any genes of donor
-and recipient species as long as the species shared their habitats.
-%
-We did not find an excess of virulence
-genes associated homologous recombining gene transfer. Actually, many important
-virulence genes were missing in the core genomes. We found association
-of virulence genes with non-recombining gene transfer.
-
-% Number of recombinant edges (Methods)
-This would indicate the degree to which the clonal
-species tree was disrupted by bacterial recombination. 
-%
-We used the method to infer which types of bacterial recombination
-were more prevalent than which other types in the lineages of the
-tree relating five streptococcus genomes.
-
-% Watterson's estimate
-in that the rate of mutation
-on the branches of the genealogy and branch lengths of the species
-tree $\mathcal{T}$ are reciprocally proportional.
-
-% simulation for ClonalOrigin stage 1
-We hoped that we could understand the power
-and limitations of \texttt{ClonalOrigin} under the parameter setup that we
-inferred from our data set.
-
-% Heat map for recombination from ClonalOrigin
-Because the
-lengths of SPY1 and SPY2 were larger than those of SDE1 and SDE2,
-recombinant edges could more often depart from one of SPY1 and SPY2
-lineages and arrive at one of SDE1 and SDE2 lineages than those for
-the reverse direction \textit{a priori} (Figure S\ref{fig:tree5}). 
-We took into account of the prior expected number of recombinant edges.
-
-\textbf{Gene transfer and functional categories:} We showed much clearer
-examples of lateral transfer of gene fragments.
-Modular structures of aaRS genes made them
-easy targets of gene transfer \citep{Woese2000}. The two transporter
-genes might have been under selective pressure from environment.
-
-% rRNA binding
-Of the genes in Table S\ref{tab:genes-transfer} we did not find ribosomal
-proteins that were heavily enriched in the gene ontology term, {}``rRNA
-binding.'' 
-Ribosomal proteins tended to cluster together in the genome,
-and they appeared to be affected by the recombinant edges whose departures
-were on the root branch. 
-Recombinant edges that connect the root branch
-are hard to be considered as gene flow. We ruled out those recombinant
-edges in Table S\ref{tab:genes-transfer} as we mentioned.
-
-% From Introduction
-It has been suspected that opportunistic pathogenicity
-of SDE may have been facilitated by the very genetic interaction between
-the two species. 
-Although virulence genes were suspected to be acquired via non-recombining gene
-transfer rather than recombining transfer \citep{Groisman1994}, systematic
-approaches to elucidating relative effects of the two types of gene transfer 
-rarely have been applied.
-
-This unfortunate situation motivated us to study the
-closely related bacteria as efforts of better understanding relative 
-association of gene transfer with virulence genes.
-We hoped that understanding their genetic exchanges
-would enhance our knowledge of mechanisms of their pathogenicity on
-humans.
-
-While compositional approaches were effective in analyzing
-genome-scale data, phylognetic approaches have allowed to understand
-evolutionary forces of gene exchange between species
-\citep[e.g.,][]{Bapteste2004,Simonson2005,Beiko2009} by comparing differences
-between gene trees and a reference species tree.  However, many of the methods
-did not have explicit models of homologous recombining gene transfer that was
-one of important evolutionary processes in horizontal gene transfer.  
-% Overall,
-% the development of methods for downstream analyses using bacterial genomes
-% appears to lag behind that of high-throughput sequencing technology
-% \citep[e.g.,][]{Green2011}.
-
-% Heat map
-% Another useful
-% summary suggested by \citet{Didelot2010} was to classify recombinant edges by
-% types defined according to donor/recipient branch pairs, and record the number
-% of edges sampled for each recombinant edge type. For instance, our species tree
-% has nine branches including rooting branch and 81 branch pairs, out of which 60
-% are possible donor/recipient pairs by prohibiting recombinant edges from going
-% backward in time. The posterior expected number of edges of each type can be
-% compared to the prior expected number determined by the three population
-% parameters fixed in the second stage of \texttt{ClonalOrigin}. This comparison
-% was displayed in a heat map of Figure \ref{fig:Heatmap-of-recombination}, which
-% indicated branch pairs that might experience an excess or depletion of
-% recombination events across all blocks.
-
-% \textbf{Functional categories associated with homologous recombination gene transfer:} 
-% Having established relative correlation of the surrogate measure of recombination
-% intensity in the simulation (Figure S\ref{fig:ri1}) we attempted to find
-% functional categories that were associated with recombination.
-% The nucleotide substitutions
-% due to recombination events might not be selectively neutral because
-% of large population of the bacteria, and accordingly natural selection
-% could be more stronger than random genetic drift. Reflecting on the
-% report of \citet{Lefebure2007} that genes that were likely to be
-% under positive selection tended to be under higher recombination,
-% we would expect that functional categories would be associated with higher
-% recombination, and consequently homologous recombining gene transfer
-% between species.
-
-% block 76: SDEG\_0641
-% ORTHOMCL43-1    SEQ\_1497        SpyM3\_0440      MGAS10750\_Spy0535 SDD27957\_03210  SDE12394\_03445  SDEG\_0641
-% Not found in Davies2007
-% putative calcium transporter
-% MgtA
-
-% block 97: SDEG\_0673
-% ORTHOMCL1244-1  SEQ\_0860        SDD27957\_03385  MGAS10750\_Spy0623 SpyM3\_0465      SDE12394\_03650  SDEG\_0673
-% Putative (put.) extracellular proteins: put. dipeptidase (SpyM3\_0465)
-% Dipeptiase for the same human host might be important for SDE to survive human
-% host environment. Scarvenge dipeptides.
-
-% block 101: SDEG\_0739
-% ORTHOMCL1207-1  SDD27957\_04880  SEQ\_0932        SDD27957\_03660  SDEG\_0739 SDE12394\_03975  MGAS10750\_Spy0677       SpyM3\_0511
-
-% block 151: SDEG\_1447
-% ORTHOMCL99-1	SEQ\_0680	SDD27957\_06980	SDEG\_1447	SDE12394\_07540	SpyM3\_1071	MGAS10750\_Spy1252
-% Not found in Davies2007
-% superoxide dismutase
-% sodM
-% Poyart2001
-% Contribution of Mn-Cofactored Superoxide Dismutase (SodA) to the Virulence of Streptococcus agalactiae.
-
-% block 244: SDEG\_2040
-% ORTHOMCL37-1	SEQ\_0209	SpyM3\_1765	MGAS10750\_Spy1855	SDD27957\_10625	SDE12394\_10325	SDEG\_2040
-% Not found in Davies2007
-% chaperonin GroEL, 
-% putative heat shock protein
-% What is the role of this GroEL?
-% Increase of GroEL in acid shock, which might be important for survival of the
-% bacteria in human hosts.
-
-% block 242: SDD27957\_00955
-% ORTHOMCL730-1	SDE12394\_10000	SEQ\_2112	SDD27957\_00955	SDE12394\_00680	SDEG\_0169	SpyM3\_1755	MGAS10750\_Spy1843
-% Not found in Davies2007
-% putative pyruvate formate-lyase activating enzyme
-% There are some upstream parts were replaced.
-
-% block 240: SDD27957\_00985
-% ORTHOMCL731-1	SDE12394\_10005	SEQ\_2106	SDD27957\_00985	SpyM3\_1749	MGAS10750\_Spy1837	SDE12394\_00710	SDEG\_0175
-% Not found in Davies2007
-% putative pyruvate formate-lyase
-% pflD
-% Yesilkaya2009
-% Pyruvate Formate Lyase Is Required for Pneumococcal Fermentative Metabolism and Virulence
-% SDE2 received from SPY1 via recombining gene transfer.
-
-% block 167: SDD27957\_06790
-% ORTHOMCL1139-1	SEQ\_0705	SpyM3\_1035	MGAS10750\_Spy1208	SDD27957\_06790	SDE12394\_07350	SDEG\_1372
-% Not found in Davies2007
-% putative internalin A precursor
-% Adhesins
-
-% block 255: MGAS10750\_Spy1922
-% ORTHOMCL158-1	SEQ\_2161	SDD27957\_10925	SDE12394\_10560	SDEG\_2091	SpyM3\_1806	MGAS10750\_Spy1922
-% Not found in Davies2007
-% DNA mismatch repair protein MutS
-% What makes MutS as a virulence genes?
-
-% block 67: SDD27957\_03085
-% ORTHOMCL121-1	SEQ\_1525	SDD27957\_03085	SpyM3\_0420	SDE12394\_03265	SDEG\_0610	MGAS10750\_Spy0514
-% Not found in Davies2007
-% lysyl-tRNA synthetase
-
-% block 133: SDD27957\_04505
-% ORTHOMCL58-1	SEQ\_1129	SDD27957\_04505	SDE12394\_04895	SDEG\_0866	SpyM3\_0764	MGAS10750\_Spy0977
-% Not found in Davies2007
-% putative spermidine/putrescine ABC transporter ATP-binding protein
-
-% List of virulence genes.
-% Genes for the Majority Of Group A Streptococcal Virulence Factors and
-% Extracellular Surface Proteins Do Not Confer an Increased Propensity to Cause
-% Invasive Disease
-
-% block 170: SDD27957\_07725
-% ORTHOMCL96-1	SEQ\_0613	SDD27957\_07725	SDEG\_1515	SDE12394\_07910	SpyM3\_1180	MGAS10750\_Spy1365
-% Not found in Davies2007
-% putative glucose kinase
-
-% Evidence for Reductive Genome Evolution and Lateral Acquisition of Virulence
-% Functions in Two Corynebacterium pseudotuberculosis Strains
-
-% block 186: MGAS10750\_Spy1444
-% ORTHOMCL93-1	SEQ\_1774	SDD27957\_08550	SDE12394\_08460	SDEG\_1691	SpyM3\_1369	MGAS10750\_Spy1444
-% Not found in Davies2007
-% putative serine/threonine kinase
-
-% Blocks with virulence genes.
-% Check virulence genes (they were combined) at 
-% SDD27957\_02890, SDD27957\_02895
-% MGAS10750\_Spy0455, MGAS10750\_Spy0456 (block 75)
-
-% HGT and virulene genes in detail
-% ORTHOMCL4-1	0	1	6	0	TRUE
-% ORTHOMCL4-1	SDD27957_02925	SEQ_0563	MGAS10750_Spy0339	SpyM3_0298	MGAS10750_Spy0340	SDEG_0469	SDD27957_02245	SDE12394_02260	SDD27957_05495	SDEG_0933	SpyM3_1726	MGAS10750_Spy1807
-% streptococcal C5a peptidase (scpB) (SDEG_0933) 
-% 
-% ORTHOMCL1826-1	0	2	6	0	TRUE
-% ORTHOMCL1826-1	SDEG_1261	SDE12394_05305	SDE12394_03405	SDEG_0634	SDD27957_00125	SDD27957_02625	SDD27957_04790	SDD27957_05475	SDE12394_02555
-% Nothing in Shimomura.
-% 
-% ORTHOMCL912-1	3	2	5	0	TRUE
-% ORTHOMCL912-1	SDD27957_04815	SDD27957_05095	SDD27957_02300	SDD27957_11185	SDD27957_05765	SDD27957_06935	SpyM3_1053	MGAS10750_Spy1189	MGAS10750_Spy1234	SpyM3_1016	SDD27957_02260	MGAS10750_Spy0401	SpyM3_1054	MGAS10750_Spy1235
-% Nothing in Shimomura.
-% SpyM3_1053 /product="putative transposase"
-% SpyM3_1054 /product="putative transposase"
-% SpyM3_1016 /product="IS1076 transposase"
-% 
-% # ORTHOMCL4-2	1	1	4	0	TRUE
-% 
-% # ORTHOMCL29-2	0	0	4	0	TRUE
-% ORTHOMCL29-1	1	1	3	1	TRUE
-% ORTHOMCL29-1	SEQ_1712	SEQ_0303	SpyM3_0174	MGAS10750_Spy0201	SDD27957_10325	SDE12394_09770	SDEG_1964	SEQ_1095	SDEG_0515	MGAS10750_Spy0403
-% Nothing in Shimomura.
-% SpyM3_0174 /product="putative response regulator"
-% 
-% # ORTHOMCL26-2	0	0	3	0	TRUE
-% ORTHOMCL26-1	0	0	3	0	TRUE
-% ORTHOMCL26-1	SpyM3_0982	MGAS10750_Spy1149	SEQ_1437	SDD27957_06480	SDEG_1294	SDE12394_06795	SEQ_2022	SDE12394_00945	SDEG_0221
-% Nothing in Shimomura.
-% SpyM3_0982 /product="putative maltose operon transcriptional repressor"
-% 
-% # ORTHOMCL1564-2	0	2	3	0	TRUE
-% 
-% #ORTHOMCL126-2	0	1	3	0	TRUE
-% ORTHOMCL126-1	0	1	3	0	TRUE
-% ORTHOMCL126-1	SEQ_0301	SDD27957_10335	SDEG_1966	SpyM3_0171	MGAS10750_Spy0199	SEQ_1711	MGAS10750_Spy0404	SDE12394_02570	SDEG_0516
-% Nothing in Shimomura.
-% SpyM3_0171 /product="putative histidine kinase"
-% 
-% ORTHOMCL1071-2	1	2	3	2	TRUE
-% % ORTHOMCL1071-2	SEQ_0845	SEQ_0190	SEQ_1729	MGAS10750_Spy0620	SpyM3_0731	SpyM3_0922	SpyM3_1208	SDEG_1106	SDD27957_07165
-% Nothing in Shimomura.
-% SpyM3_0731 /product="putative holin - phage associated"
-% SpyM3_0922 /product="putative cell wall hydrolase - phage associated"
-% SpyM3_1208 /product="putative cell wall hydrolase, lysin - phage associated"
-% 
-% ORTHOMCL1037-1	0	2	3	0	TRUE
-% ORTHOMCL1037-1	MGAS10750_Spy1697	SDE12394_04580	MGAS10750_Spy0948	SDD27957_05395	SDD27957_04295	SDE12394_07035
-% Nothing in Shimomura.
-% 
-% ORTHOMCL2721	0	4	2	0	TRUE
-% ORTHOMCL2721	SDE12394_10035	SDE12394_09945	SDE12394_09940
-% Nothing in Shimomura.
-% 
-% # ORTHOMCL1826-2	4	0	2	0	TRUE
-% # ORTHOMCL1826-2	SDEG_1261	SDE12394_05305	SDE12394_03405	SDEG_0634	SDD27957_00125	SDD27957_02625	SDD27957_04790	SDD27957_05475	SDE12394_02555
-% 
-% ORTHOMCL1564-1	0	2	2	0	TRUE
-% ORTHOMCL1564-1	SDEG_1103	SpyM3_1409	SDD27957_09010	SEQ_0696	SDD27957_02745	SDEG_0541
-% GGS_124 possesses 5 genes that encode putative nucleases with a secretion signal peptide: genome-encoded streptodornase (SDEG_0541), others,  prophage-associated deoxyribonuclease (SDEG_1103), all of which are predicted to code for a secretion signal peptide.
-% SpyM3_1409 /product="streptodornase (Sdn) - phage associated"
-% 
-% 
-% ORTHOMCL1439-1	0	1	2	0	TRUE
-% ORTHOMCL1439-1	SEQ_1606	SEQ_1999	SpyM3_0104	SDE12394_09890	SDD27957_10420	SDE12394_09885	SDEG_1984
-% Nothing in Shimomura.
-% SpyM3_0104 /product="protein F2 like fibronectin-binding protein"
-% 
-% ORTHOMCL1369-2	2	1	2	0	TRUE
-% ORTHOMCL1369-2	MGAS10750_Spy1683	SDEG_1338	SDEG_1346	SDE12394_07225	SDE12394_04610	SDD27957_04325	SDD27957_05365	SDD27957_08360
-% Nothing in Shimomura.
-% 
-% ORTHOMCL1283-1	3	0	2	0	TRUE
-% ORTHOMCL1283-1	SDD27957_00840	SDEG_1543	SDEG_1531	SDE12394_03300	SDEG_0983	SDE12394_05495	SDE12394_08020	SDE12394_07960
-% Nothing in Shimomura.
-% 
-% ORTHOMCL1232-2	0	1	2	0	TRUE
-% ORTHOMCL1232-2	SDE12394_10675	SDE12394_10850	SDEG_2142	MGAS10750_Spy1897	SDD27957_00535	SDD27957_02435
-% Nothing in Shimomura.
-% 
-% ORTHOMCL1229-2	0	2	2	0	TRUE
-% ORTHOMCL1229-2	SDD27957_03510	SDD27957_05295	SDD27957_08295	SEQ_0753	MGAS10750_Spy1709
-% Nothing in Shimomura.
-% 
-% --------------------------------
-% 
-% 
-% ORTHOMCL9-2	1	1	1	0	TRUE
-% ORTHOMCL9-2	SpyM3_0160	MGAS10750_Spy0188	SEQ_0271a	SEQ_0289a	SDD27957_10400	SDE12394_09855	SDEG_1980	SpyM3_1853
-% 
-% ORTHOMCL912-2	5	1	1	0	TRUE
-% ORTHOMCL9-1	1	0	1	0	TRUE
-% ORTHOMCL848-2	0	1	1	0	TRUE
-%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Functional category association
-Lateral transfer of transporter related genes are well known.  Functional
-categories associated with higher recombining gene transfer were found to
-include phosphotransferase systems and ATP-binding cassette (ABC) transporters,
-i.e., carbohydrate transport (GO:0008643) (Table \ref{tab:functional}).
-
-We also
-found several aminoacyl tRNA synthetases in gene ontology of translation
-(GO:0006412) to be significantly associated with higher recombining gene
-transfer.
-
-Functional categories concerning ribosomes were also
-associated with higher recombining gene transfer. These categories were not
-found significantly associated with higher recombining gene transfer when we
-considered only gene transfer between SPY and SDE genomes, which implied that
-many recombining gene transfer events should be attributed to those of internal
-species branches as donors.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% From the discussion about virulence genes and two gene transfer
-We suspected that virulence genes were exchanged between SPY and
-SDE, but did not find any directionality of the gene flow.
-This does not necessarily mean that
-virulence genes were not exchanged between SPY and SDE via non-recombining
-transfer.  Because the \texttt{Mowgli} analysis just did not reveal unbalanced
-non-recombining or recombinding gene transfer, we suspected that \texttt{Mowgli}'s
-parsimony might not be powerful in detecting unbalanced gene flow.  We
-conjecture the following hypothesis of the evolution of SPY and SDE to support
-the non-recombining virulence gene exchange between SPY and SDE.  Pathogenicity
-of SDE might have existed before the common ancestor of SPY and SDE.  We
-speculated that SPY might have been split off from the common ancestor that was
-SDE-like because the population size of SDE was larger than that of SPY (CITE).
-After the split, SPY might have been specialized in inhabting human hosts, and
-SDE in colonizing veterinary hosts. Recent surge in cases of human diseases
-attributed to SDE might be because of secondary contact of SPY and SDE since
-their split from their common ancestor.  When SPY thrives in its environment of
-human hosts and another relatively close species, SDE, starts to cohabit with
-the indigenous species SPY, genetic elements in one of the two species could
-benefit another species. If the two species do not compete with each other, then
-the increased genetic variation from multiple species would eventually bring
-advantage to both of the species.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Functional category association 
-\textbf{Two functional categories and recombining gene transfer:}
-Among the genes shown in Table S\ref{tab:functional} two classes of proteins
-attracted our attentions: aminoacyl tRNA synthetases (aaRS) and
-phosphotransferase system (PTS).  Aminoacyl tRNA synthetases charge tRNA with an
-amino acid so that amino acids are concatenated from charged tRNA onto a growing
-peptides through translation. The aaRS's are characterized by an amino acid that
-they charge tRNA with.  This gene family had been a target for horizontal gene
-transfer \citep[e.g.,][]{Lamour1994,Woese2000,Koonin2001}, which lent itself to
-an example of gene transfer in bacteria.  There were also recent reports of gene
-transfer in tRNA aminoacylation \citep{Andam2010,Wang2011}.  In our analysis,
-some of them appeared to be transferred from SPY to SDE: leucyl-tRNA synthetase,
-aspartyl-tRNA synthetase, and arginyl-tRNA synthetase. Half of phenylalanyl-tRNA
-synthetase gene appeared to be transferred in the reverse direction.
-
-Bacterial phosphotransferase system (PTS), also known as phosphoenolpyruvate
-(PEP) group translocation, transports sugars such as glucose into the cell where
-phosphorylation of the sugar happens. There have been studies of gene transfer
-of PTS in bacteria \citep[e.g.,][]{Barabote2005,Zuniga2005}. As the main source
-of energy for bacteria, sugars in the bacterial niches must have been acting as
-important selective pressure. It is not surprising that PTS sugars differ among
-bacterial groups because bacterial species are correlated with their environment
-and energy sources theirof.  Table S\ref{tab:functional} shows that two PTS
-genes appeared to be transferred from SPY to SDE: {}``PTS system,
-mannose-specific IIC component'' and {}``PTS system, fructose-specific IIABC
-component.'' Whereas almost half of the former gene was inferred to be affected
-by recombinant edges, only 3\% of the latter gene appeared to be affected.
-
-% This agreement
-% between the two different methods, applied on somewhat different parts of the
-% data, serves as validation for the signal picked up by these separate
-% association studies.
-
-% These topology-not-changing recombinant edges tend change only branch lengths.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Should we keep this section?}
-
-With the slight over-estimation of recombination events we were concerned with
-any potential biases hidden in \texttt{ClonalOrigin} implementation. Because the
-simulated data were generated with no directionality between SPY and SDE,  if we
-estimate recombination events from the simulated data sets using
-\texttt{ClonalOrigin} and compare those estimates from those from the real data
-set, this could remove biases in \texttt{ClonalOrigin} implementation. We could
-use inferred estimates of recombination from the simulated data under the prior
-model of \texttt{ClonalOrigin} as the prior expected number of recombination.
-Figure S\ref{fig:h2} shows a plot of numbers of recombination events from the
-real data against those from the simulated data sets. Note that the numbers of
-recombination events from simulated data sets appear to deviate from the line of
-slope being 1 and intercept being 0.  The numbers of recombination events for
-the recombinant edges from SPY to SDE increase by more than 2.5 fold whereas
-those for the recombinant edges from SDE to SPY increase by about 1.7 fold.
-This was not surprising because inference of number of recombinant edges was
-quite accurate. We considered that subtraction of effects of number of
-recombinant edges inferred from simulated data and subtraction of effects of the
-prior expected number of recombinant edges were effectively equivalent.
-However, this comparison between inferrence from simulated data and that of real
-data set suggested that the directional difference between SPY and SDE was due
-to neither lack of power of \texttt{ClonalOrigin} nor hidden bias thereof.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% 
-%(Table S\ref{tab:functional-sde-spy})
-%(Table S\ref{tab:functional-spy-sde})
-%(Table S\ref{tab:go-events-duplication})
-%(Table S\ref{tab:go-events-loss})
-%
-%
-%For horizontal gene transfer we found enrichment in very different categories
-%related to transposition, integration, and recombination (Table
-%\ref{tab:go-events}).  
-%
-%Gene duplication and loss events were enriched for
-%categories similar to those of horizontal gene transfer's enrichment (Tables
-%S\ref{tab:go-events-duplication} and S\ref{tab:go-events-loss}).
-%
-%Gene duplication and horizontal transfers showed enrichment in the same
-%categories perhaps because both mechanisms increase gene copy number.  
-%
-%Gene losses showed enrichment for transposition and integration, but less
-%significantly (P-value of $<10^{-6}$), and showed enrichment for several
-%additional categories such as transporter activity and transcription initiation.
-%
-%Comparing two types of recombinant edges, one from SDE to SPY and another from
-%SPY to SDE transport related categories were more associated for the direction
-%from SPY to SDE
-%(Table S\ref{tab:functional-sde-spy} and S\ref{tab:functional-spy-sde}).
-%
-%\texttt{More interpretations about Mawgli analysis of function categories?
-%Why transposases?}
-
-% \texttt{IG: NEXT PARAGRAPH CONTAINS SOME DISCUSSION ON THESE RESULTS. WE SHOULD 
-% GET BACK TO THIS ONCE WE DECIDE WHAT TO DO WITH THE RESULTS}
-
-%The parsimony-based approach also showed
-%enrichment of the recombining transfers in many of the same categories such as
-%translation, ribosome, and rRNA binding when using all types of recombinant
-%edges in computing recombination intensity (see Tables
-%S\ref{tab:functional-all} and S\ref{tab:go-events-recombining}).  
-%Because the
-%parsimony-based method did not take species tree branch lengths into account, 
-%the agreement between the two different methods might be likely.

### 2011-10-18 src/warg2/SimpleGlob.h
warg2's version is implemented.

@@ -130,24 +130,24 @@
  */
 enum SG_Flags
 {
-    SG_GLOB_ERR         = 1 << 0,
-    SG_GLOB_MARK        = 1 << 1,
-    SG_GLOB_NOSORT      = 1 << 2,
-    SG_GLOB_NOCHECK     = 1 << 3,
-    SG_GLOB_TILDE       = 1 << 4,
-    SG_GLOB_ONLYDIR     = 1 << 5,
-    SG_GLOB_ONLYFILE    = 1 << 6,
-    SG_GLOB_NODOT       = 1 << 7,
-    SG_GLOB_FULLSORT    = 1 << 8
+  SG_GLOB_ERR         = 1 << 0,
+  SG_GLOB_MARK        = 1 << 1,
+  SG_GLOB_NOSORT      = 1 << 2,
+  SG_GLOB_NOCHECK     = 1 << 3,
+  SG_GLOB_TILDE       = 1 << 4,
+  SG_GLOB_ONLYDIR     = 1 << 5,
+  SG_GLOB_ONLYFILE    = 1 << 6,
+  SG_GLOB_NODOT       = 1 << 7,
+  SG_GLOB_FULLSORT    = 1 << 8
 };
 
 /*! @brief Error return codes */
 enum SG_Error
 {
-    SG_SUCCESS          =  0,
-    SG_ERR_NOMATCH      =  1,
-    SG_ERR_MEMORY       = -1,
-    SG_ERR_FAILURE      = -2
+  SG_SUCCESS          =  0,
+  SG_ERR_NOMATCH      =  1,
+  SG_ERR_MEMORY       = -1,
+  SG_ERR_FAILURE      = -2
 };
 
 // ---------------------------------------------------------------------------
@@ -217,111 +217,111 @@ enum SG_Error
 class SimpleGlobUtil
 {
 public:
-    static const char * strchr(const char *s, char c)
-    {
-        return (char *) sg_strchr((const SOCHAR_T *)s, c);
-    }
-    static const wchar_t * strchr(const wchar_t *s, wchar_t c)
-    {
-        return ::wcschr(s, c);
-    }
+  static const char * strchr(const char *s, char c)
+  {
+    return (char *) sg_strchr((const SOCHAR_T *)s, c);
+  }
+  static const wchar_t * strchr(const wchar_t *s, wchar_t c)
+  {
+    return ::wcschr(s, c);
+  }
 #if SG_HAVE_ICU
[…]

### 2011-10-18 src/warg2/exponential.h

@@ -8,27 +8,27 @@ namespace weakarg
 #define EXP_C 60801
 inline double exponential(double y)
 {
-    union
-    {
-        double d;
+  union
+  {
+    double d;
 #ifdef LITTLE_ENDIAN
-        struct
-        {
-            int j, i;
-        }
-        n;
-#else
-        struct
-        {
-            int i, j;
-        }
-        n;
-#endif
+    struct
+    {
+      int j, i;
     }
-    _eco;
-    _eco.n.i = (int)(EXP_A*(y)) + (1072693248 - EXP_C);
-    _eco.n.j = 0;
-    return _eco.d;
+    n;
+#else
+    struct
+    {
+      int i, j;
+    }
+    n;
+#endif
+  }
+  _eco;
+  _eco.n.i = (int)(EXP_A*(y)) + (1072693248 - EXP_C);
+  _eco.n.j = 0;
+  return _eco.d;
 }
 
 } // end namespace weakarg

### 2011-10-18 src/warg2/move.h

@@ -15,65 +15,65 @@ extern gsl_rng * rng;
 class Move
 {
 protected:
-    std::string name;///<The moves name
-    std::string description;///<The moves description
-    Param*param;///<Parameter set on which the move is performed
-    double alpha;///<Weighting for this move (probability is alpha/sum_i(alpha_i) )
-    int numcalls;///<Number of attempts at the move
-    int numaccept;///<Number of acceptances of the move
+  std::string name;///<The moves name
+  std::string description;///<The moves description
+  Param*param;///<Parameter set on which the move is performed
+  double alpha;///<Weighting for this move (probability is alpha/sum_i(alpha_i) )
+  int numcalls;///<Number of attempts at the move
+  int numaccept;///<Number of acceptances of the move
 public:
-    Move(Param*p,double a=1.0);///<Constructor specifying on which Parameter the move is to be performed
-    virtual Move * clone()=0;
-    virtual ~Move()=0;
-    virtual int move()=0;///<Performs the move once
-    virtual inline int getTopoCounts()
-    {
-        return(-1);
-    };
-    virtual inline int getTopoAcc()
-    {
-        return(-1);
-    };
-    virtual inline int move(vector<int> *samplespace)
-    {
-        cout<<"Problem in Move!"<<endl;
-        return(move());
-    };///<Performs the move once
-    inline void setParam(Param * param)
-    {
-        this->param=param;
-    }///<Change the Parameter on which the move is to be performed
-    inline std::string desc()
-    {
-        return description;
-    }///<Returns a description of the move
-    inline std::string getName()
-    {
-        return name;
-    }///<Returns a description of the move
-    inline void setName(std::string newname)
-    {
-        name=newname;
-    }
-    inline void setAlpha(double a)
-    {
-        alpha=a;
-    }
-    ;///<Sets alpha for the move
-    inline double getAlpha()
-    {
-        return(alpha);
-    }
-    inline int getCounts()
-    {
-        return(numcalls);
-    }
-    inline int getAcc()
-    {
-        return(numaccept);
-    }
+  Move(Param*p,double a=1.0);///<Constructor specifying on which Parameter the move is to be performed
+  virtual Move * clone()=0;
+  virtual ~Move()=0;
+  virtual int move()=0;///<Performs the move once
+  virtual inline int getTopoCounts()
+  {
+    return(-1);
+  };
[…]

### 2011-10-18 src/warg2/moveaddedge.cpp

@@ -5,65 +5,65 @@ namespace weakarg
 {
 
 MoveAddEdge::MoveAddEdge(Param *p,double a)
-    : Move(p,a)
+  : Move(p,a)
 {
-    description= "Update adding a recombinant edge";
-    name= "AddEdge";
+  description= "Update adding a recombinant edge";
+  name= "AddEdge";
 }
 
 
 int MoveAddEdge::move(vector<int> * samplespace)
 {
-    RecTree * rectree=param->getRecTree();
-    double tfrom,tto;
-    unsigned int start,end;
-    unsigned int efrom,eto;
-    //Draw start and end
-    rectree->setBlock(&start,&end,param->getDelta(),param->getData()->getBlocks());
-    //Draw eto and tto
-    bool insamplespace=false;
-    while(!insamplespace)
+  RecTree * rectree=param->getRecTree();
+  double tfrom,tto;
+  unsigned int start,end;
+  unsigned int efrom,eto;
+  //Draw start and end
+  rectree->setBlock(&start,&end,param->getDelta(),param->getData()->getBlocks());
+  //Draw eto and tto
+  bool insamplespace=false;
+  while(!insamplespace)
+    {
+      eto=rectree->getPoint(&tto);
+      //Draw efrom and tfrom
+      tfrom=tto+rectree->getNode(eto)->getAge();
+      efrom=rectree->getEdgeCoal(&tfrom);
+      tfrom-=rectree->getNode(efrom)->getAge();
+      if(samplespace==NULL) insamplespace=true;
+      else if (samplespace->size()==0) insamplespace=true;
+      else
         {
-            eto=rectree->getPoint(&tto);
-            //Draw efrom and tfrom
-            tfrom=tto+rectree->getNode(eto)->getAge();
-            efrom=rectree->getEdgeCoal(&tfrom);
-            tfrom-=rectree->getNode(efrom)->getAge();
-            if(samplespace==NULL) insamplespace=true;
-            else if (samplespace->size()==0) insamplespace=true;
-            else
-                {
-                    for(unsigned int i=0; i<samplespace->size(); i++) if((int)eto==samplespace->at(i)||(int)efrom==samplespace->at(i))
-                            {
-                                insamplespace=true;
-                                break;
-                            }
-                }
[…]

### 2011-10-18 src/warg2/moveaddedge.h

@@ -12,17 +12,17 @@ namespace weakarg
 class MoveAddEdge : public Move
 {
 public:
-    MoveAddEdge(Param * p,double a);
-    Move * clone()
-    {
-        return new MoveAddEdge(*this);
-    }
-    int move(vector<int> * samplespace=NULL);
-    inline int move()
-    {
-        return(move(NULL));
-    }
-    ~MoveAddEdge();
+  MoveAddEdge(Param * p,double a);
+  Move * clone()
+  {
+    return new MoveAddEdge(*this);
+  }
+  int move(vector<int> * samplespace=NULL);
+  inline int move()
+  {
+    return(move(NULL));
+  }
+  ~MoveAddEdge();
 
 };
 

### 2011-10-18 src/warg2/moveageclonal.cpp

@@ -7,216 +7,216 @@ namespace weakarg
 {
 
 MoveAgeClonal::MoveAgeClonal(Param * p, double a,bool allowclonal,int r,double t)
-    : Move(p,a)
+  : Move(p,a)
 {
-    description= "Update of the age of a clonal node";
-    name= "AgeClonal";
-    numtopo=0;
-    numtopoaccept=0;
-    reps=r;
-    tempering=t;
-    s1=param->getClonalStepSize();
-    s0=param->getClonalStepSize();
-    usenormalprop=true;
-    backuptree = NULL;
-    changeTopology=allowclonal;// we can forbid topology changes with this (no tempering is then needed)
+  description= "Update of the age of a clonal node";
+  name= "AgeClonal";
+  numtopo=0;
+  numtopoaccept=0;
+  reps=r;
+  tempering=t;
+  s1=param->getClonalStepSize();
+  s0=param->getClonalStepSize();
+  usenormalprop=true;
+  backuptree = NULL;
+  changeTopology=allowclonal;// we can forbid topology changes with this (no tempering is then needed)
 }
 
 inline double fround(double n, double d)
 {
-    return floor(n * pow(10., d) + .5) / pow(10., d);
+  return floor(n * pow(10., d) + .5) / pow(10., d);
 }
 
 int MoveAgeClonal::move()
 {
-    // Propose to change a nodes age
-    // reorder the nodes correctly
-    // change the recedge times
-    // account for factors of 2 in passing other clonal nodes
-    // accept/reject step
-    // if reject, put things back correctly
-    bool usetemper=false;// an indicator as to whether tempering is needed
-    if(!changeTopology)s1=0;// probabilities of the return move won't work unless we do this (because 2 reflecting boundaries)
+  // Propose to change a nodes age
+  // reorder the nodes correctly
+  // change the recedge times
+  // account for factors of 2 in passing other clonal nodes
+  // accept/reject step
+  // if reject, put things back correctly
+  bool usetemper=false;// an indicator as to whether tempering is needed
+  if(!changeTopology)s1=0;// probabilities of the return move won't work unless we do this (because 2 reflecting boundaries)
 
 
-    double temperllorig=0,temperllto=0,temperllbase=0,temperllfinal=0;
-    double temperpriororig=0,temperpriorto=0,temperpriorbase=0,temperpriorfinal=0;
[…]

### 2011-10-18 src/warg2/movedelta.cpp

@@ -5,47 +5,47 @@ namespace weakarg
 {
 
 MoveDelta::MoveDelta(Param * p,double a)
-    : Move(p,a)
+  : Move(p,a)
 {
-    description= "Update of the mean recombination length delta";
-    name= "Delta";
+  description= "Update of the mean recombination length delta";
+  name= "Delta";
 }
 
 int MoveDelta::move()
 {
-    double olddelta=param->getDelta();
-    numcalls++;
-    int X=0,Y=0;
-    for (int i=0; i<param->getRecTree()->numRecEdge(); i++)
-        {
-            Y+=param->getRecTree()->getRecEdge(i)->getEnd()-param->getRecTree()->getRecEdge(i)->getStart();
-            if (param->getData()->isBegEnd(param->getRecTree()->getRecEdge(i)->getStart()))
-                X++;
-            if (param->getData()->isBegEnd(param->getRecTree()->getRecEdge(i)->getEnd()))
-                Y--;
-            else X--;/*{
+  double olddelta=param->getDelta();
+  numcalls++;
+  int X=0,Y=0;
+  for (int i=0; i<param->getRecTree()->numRecEdge(); i++)
+    {
+      Y+=param->getRecTree()->getRecEdge(i)->getEnd()-param->getRecTree()->getRecEdge(i)->getStart();
+      if (param->getData()->isBegEnd(param->getRecTree()->getRecEdge(i)->getStart()))
+        X++;
+      if (param->getData()->isBegEnd(param->getRecTree()->getRecEdge(i)->getEnd()))
+        Y--;
+      else X--;/*{
             X--;
             Y--;
         }*/
-        }
-    for (int rep=0; rep<100; rep++)
-        {
-            double t=param->getDelta();
-            double t2=t+(gsl_rng_uniform(rng)-0.5)*10.0;
-            if (t2<1.0)
-                t2=2.0-t2;
-            int b=param->getData()->getB();
-            int L=param->getData()->getL();
-            int R=param->getRecTree()->numRecEdge();
-            double ratio=X*(log(t2)-log(t));
-            ratio+=Y*(log(1.0-1.0/t2)-log(1.0-1.0/t));
-            ratio-=R*(log(b*t2+L-b)-log(b*t+L-b));
-            if (log(gsl_rng_uniform(rng))<ratio)
[…]

### 2011-10-18 src/warg2/moveedgechange.cpp

@@ -7,80 +7,80 @@ namespace weakarg
 // the step size (sigma in a normal distribution) of
 
 MoveEdgeChange::MoveEdgeChange(Param * p,double a)
-    : Move(p,a)
+  : Move(p,a)
 {
-    description="Update moving the edge of a recedge";
-    name="EdgeChange";
+  description="Update moving the edge of a recedge";
+  name="EdgeChange";
 }
 
 int MoveEdgeChange::move(vector<int> * samplespace)
 {
-    RecTree * rectree=param->getRecTree();
-    dlog(1)<<"Proposing to change edge sites... ";
-    // first choose the move type
-    int which=rectree->sampleEdge(samplespace);//floor(gsl_rng_uniform(rng)*rectree->numRecEdge());
-    if (which<0)
-        {
-            dlog(1)<<"No valid edges to change!"<<endl;
-            return(-1);
-        }
-    numcalls++;
-    int movetto=gsl_rng_uniform_int(rng,2); //0 for no or 1 for yes
-    // collect the details of the edge
-    double l=param->getLL();
-    int oldedge;
-    double time;
-    if (movetto)
-        {
-            oldedge=rectree->getRecEdge(which)->getEdgeTo  ();
-            time=rectree->getEdgeTimeAbsTo  (which);
-        }
-    else
-        {
-            oldedge=rectree->getRecEdge(which)->getEdgeFrom();
-            time=rectree->getEdgeTimeAbsFrom(which);
-        }
-    unsigned int start=rectree->getRecEdge(which)->getStart();
-    unsigned int end  =rectree->getRecEdge(which)->getEnd();
-    dlog(1)<<"edge "<<which<<" ("<<start<<":"<<end<<")"<<flush;
-    vector<double> store(end-start);
-    for (unsigned int i=start; i<end; i++)
-        store[i-start]=param->getLLsite(i);
-    // change the edge
-    int tmpedge=oldedge;
-    vector<int> v=rectree->alive(time);
-    if (v.size()==0)
-        throw;
-    if (v.size()==1)
-        {
-            dlog(1)<<"Cannot change chosen edge!"<<endl;
-            return(-1);
-        }
-    while (tmpedge==oldedge)
-        tmpedge=v[gsl_rng_uniform_int(rng,v.size())];
-    if (movetto)
[…]

### 2011-10-18 src/warg2/movegreedytree.h

@@ -14,53 +14,53 @@ class MoveGreedyTree : public Move
 {
 
 public:
-    MoveGreedyTree(Param * p,double a);
-    Move * clone()
-    {
-        return new MoveGreedyTree(*this);
-    };
-    int move();
-    ~MoveGreedyTree();
+  MoveGreedyTree(Param * p,double a);
+  Move * clone()
+  {
+    return new MoveGreedyTree(*this);
+  };
+  int move();
+  ~MoveGreedyTree();
 
-    inline vector<int> getAllTips(int s,int N,RecTree * t)
-    {
-        vector<int> tlist=t->getAllChildren(s);
-        vector<int> ret;
-        for(unsigned int c1=0; c1<tlist.size(); c1++)
-            {
-                if(tlist[c1]<N) ret.push_back(tlist[c1]);
-            }
-        return(ret);
-    }///< gets all tips beneath a node on the tree
-    inline int getPairIndex(int a, int b, int N)
-    {
-        if(a==b) throw("getPairIndex:Impossible to pair individual with itself!");
-        if(a>b)
-            {
-                int c=a;    // now b>a
-                a=b;
-                b=c;
-            }
-        int ret=0;
-        for(int c1=0; c1<N; c1++)
-            {
-                for(int c2=c1+1; c2<N; c2++)
-                    {
-                        if((c1==a) &&(c2==b)) return(ret);
-                        ret ++;
-                    }
-            }
-        cerr<<a<<" or "<<b<<" is not less than N="<<N<<endl;
-        throw("getPairIndex:Not found a possible list index!");
-        return(-1);
-    }///< gets the pair index used internally to store the set of all tip pairings as a flat list
-    int recCount(int s1, int s2);///< Counts the number of recombinations between two sequences
-    vector<int> edgesBetween(int s1, int s2);///< returns a list of the edges that are between two sequences
-    vector<double> calcDists(vector<double> nmuts,vector<double> nsites);///calculates distances based on provided details
-    vector<double> calcDists(double * esttheta=NULL);///< Calculates the distances between each pair of sequences
-    vector<double> calcAges(vector<double> dists);///< Calculate the correct ages based on the distances
-    void applyChanges(vector<double> newage);///< Updates the ages of a tree to the newages
-    vector<vector<bool> >* localClonalFrame();///< returns the matrix of sites in the clonal frame between each pair of sequences
[…]

### 2011-10-18 src/warg2/moveregraftclonal.cpp

@@ -7,146 +7,146 @@ namespace weakarg
 //#define DEBUG
 
 MoveRegraftClonal::MoveRegraftClonal(Param * p,double a,int r,double t)
-    : Move(p,a)
+  : Move(p,a)
 {
-    description="Update to regraft a clonal node to a new parent";
-    name="regraft";
-    reps=r;
-    tempering=t;
-    backuptree=NULL;
+  description="Update to regraft a clonal node to a new parent";
+  name="regraft";
+  reps=r;
+  tempering=t;
+  backuptree=NULL;
 }
 
 inline double fround(double n, double d)
 {
-    return floor(n * pow(10., d) + .5) / pow(10., d);
+  return floor(n * pow(10., d) + .5) / pow(10., d);
 }
 
 
 int MoveRegraftClonal::move()
 {
-    numcalls++;
-    dlog(1)<<"Proposing to regraft tree...";
+  numcalls++;
+  dlog(1)<<"Proposing to regraft tree...";
 
-    if(backuptree == NULL)
-        backuptree = new RecTree(*(param->getRecTree()));
-    else
-        backuptree->assign(*(param->getRecTree()));
-    RecTree* rectree = backuptree;
+  if(backuptree == NULL)
+    backuptree = new RecTree(*(param->getRecTree()));
+  else
+    backuptree->assign(*(param->getRecTree()));
+  RecTree* rectree = backuptree;
 
-    RecTree * oldrectree=param->getRecTree();
-    param->setRecTree(rectree,true);
+  RecTree * oldrectree=param->getRecTree();
+  param->setRecTree(rectree,true);
 
-    int which=2*rectree->getN()-2;
-    if(which<=4)
-        {
-            dlog(1)<<" No regrafting of 3 or fewer nodes!"<<endl;
-            return(-1);
-        }
-    while(!validNode(which))
-        {
-            which=gsl_rng_uniform_int(rng,2*rectree->getN()-2);
-        }
-    int origfather=rectree->getNode(which)->getFather()->getId();
-    int notmovednode=rectree->otherChild(origfather,which);
-    int whichto=getAlive(which);
-    if(whichto<0)
-        {
-            cerr<<"Error in moveregraftclonal: can't find new place to regraft!"<<endl;
-            throw("Regraft error");
-        };
+  int which=2*rectree->getN()-2;
+  if(which<=4)
+    {
[…]

### 2011-10-18 src/warg2/moveremedge.cpp

@@ -6,50 +6,50 @@ namespace weakarg
 {
 
 MoveRemEdge::MoveRemEdge(Param * p,double a)
-    : Move(p,a)
+  : Move(p,a)
 {
-    description= "Update removing a recombinant edge";
-    name= "RemEdge";
+  description= "Update removing a recombinant edge";
+  name= "RemEdge";
 
 }
 
 
 int MoveRemEdge::move(vector<int> * samplespace)
 {
-    RecTree * rectree=param->getRecTree();
-    int which=rectree->sampleEdge(samplespace);//floor(gsl_rng_uniform(rng)*rectree->numRecEdge());
-    if (which<0)
-        {
-            dlog(1)<<"No valid edges to change!"<<endl;
-            return(-1);
-        }
-    numcalls++;
-    double l=param->getLL();
-    double tfrom=param->getRecTree()->getRecEdge(which)->getTimeFrom();
-    double tto  =param->getRecTree()->getRecEdge(which)->getTimeTo  ();
-    int start=param->getRecTree()->getRecEdge(which)->getStart();
-    int end  =param->getRecTree()->getRecEdge(which)->getEnd();
-    int efrom=param->getRecTree()->getRecEdge(which)->getEdgeFrom();
-    int eto  =param->getRecTree()->getRecEdge(which)->getEdgeTo  ();
-    rectree->remRecEdge(which);
-    vector<double> store(end-start);
-    for (int i=start; i<end; i++)
-        store[i-start]=param->getLLsite(i);
-    param->computeLikelihood(start,end);
-    double l2=param->getLL();
-    dlog(1)<<"Proposing to remove edge "<<which<<"...";
-    if (log(gsl_rng_uniform(rng))>l2-l+log((1.0+rectree->numRecEdge())*2.0/param->getRho()/rectree->getTTotal()))
-        {
-            dlog(1)<<"Rejected!"<<endl;
-            if(param->getRecTree()->addRecEdge(tfrom,tto,start,end,efrom,eto)<0) throw("MoveRemEdge: Can't restore edge!");
-            for (int i=start; i<end; i++)
-                param->setlocLL(i,store[i-start]);
-            param->setLL(l);
-            return(0);
-        }
-    else dlog(1)<<"Accepted!"<<endl;
-    numaccept++;
-    return(1);
+  RecTree * rectree=param->getRecTree();
[…]

### 2011-10-18 src/warg2/moverho.cpp

@@ -5,23 +5,23 @@ namespace weakarg
 {
 
 MoveRho::MoveRho(Param * p,double a)
-    : Move(p,a)
+  : Move(p,a)
 {
-    description= "Update of the recombination rate rho";
-    name= "Rho";
+  description= "Update of the recombination rate rho";
+  name= "Rho";
 }
 
 int MoveRho::move()
 {
-    int r=param->getRecTree()->numRecEdge();
-    double T=param->getRecTree()->getTTotal();
-    double rho=param->getRho();
-    double rho2=gsl_ran_gamma(rng,1.0+r,1.0/(param->hyperPriorOfRho()+T*0.5));
-    param->setRho(rho2);
-    dlog(1)<<"Gibbs update of rho from "<<rho<<" to "<<rho2<<"..."<<endl;
-    numcalls++;
-    numaccept++;
-    return(1);
+  int r=param->getRecTree()->numRecEdge();
+  double T=param->getRecTree()->getTTotal();
+  double rho=param->getRho();
+  double rho2=gsl_ran_gamma(rng,1.0+r,1.0/(param->hyperPriorOfRho()+T*0.5));
+  param->setRho(rho2);
+  dlog(1)<<"Gibbs update of rho from "<<rho<<" to "<<rho2<<"..."<<endl;
+  numcalls++;
+  numaccept++;
+  return(1);
 }
 
 MoveRho::~MoveRho()

### 2011-10-18 src/warg2/moverho.h

@@ -12,13 +12,13 @@ namespace weakarg
 class MoveRho : public Move
 {
 public:
-    MoveRho(Param * p,double a);
-    Move * clone()
-    {
-        return new MoveRho(*this);
-    };
-    int move();
-    ~MoveRho();
+  MoveRho(Param * p,double a);
+  Move * clone()
+  {
+    return new MoveRho(*this);
+  };
+  int move();
+  ~MoveRho();
 };
 
 } // end namespace weakarg

### 2011-10-18 src/warg2/movescaletree.cpp

@@ -7,91 +7,91 @@ namespace weakarg
 {
 
 MoveScaleTree::MoveScaleTree(Param * p,double a)
-    : Move(p,a)
+  : Move(p,a)
 {
-    description="Update all nodes and edges in the tree to a new rescaled time";
-    name="ScaleTree";
+  description="Update all nodes and edges in the tree to a new rescaled time";
+  name="ScaleTree";
 }
 
 inline double fround(double n, double d)
 {
-    return floor(n * pow(10., d) + .5) / pow(10., d);
+  return floor(n * pow(10., d) + .5) / pow(10., d);
 }
 
 int MoveScaleTree::move()
 {
-    numcalls++;
-    // Propose to change all nodes and recedges by a scaled amount
-    RecTree * rectree=param->getRecTree();
-    double lprior=rectree->prior(param)+param->logPriorOfRho()+param->logPriorOfTheta();
-    // update the tree
-    double logscale=param->getScaleTreeSize() *(gsl_rng_uniform(rng) * 2.0 -1.0);
-    double scale =exp(logscale);
-    dlog(1)<<"Proposing to scale tree TMRCA by "<<scale<<"...";
-    scaleTree(scale);
-    param->setRho(param->getRho()/scale);
-    param->setTheta(param->getTheta()/scale);
-    double newlprior=rectree->prior(param) + param->logPriorOfRho() + param->logPriorOfTheta();
-    if(log(gsl_rng_uniform(rng))>newlprior-lprior+(2.0*rectree->numRecEdge()+rectree->getN()-3.0)*logscale)
-        {
-            dlog(1)<<" Rejected!"<<endl;
-            scaleTree(1.0/scale);
-            param->setRho(param->getRho()*scale);
-            param->setTheta(param->getTheta()*scale);
+  numcalls++;
+  // Propose to change all nodes and recedges by a scaled amount
+  RecTree * rectree=param->getRecTree();
+  double lprior=rectree->prior(param)+param->logPriorOfRho()+param->logPriorOfTheta();
+  // update the tree
+  double logscale=param->getScaleTreeSize() *(gsl_rng_uniform(rng) * 2.0 -1.0);
+  double scale =exp(logscale);
+  dlog(1)<<"Proposing to scale tree TMRCA by "<<scale<<"...";
+  scaleTree(scale);
+  param->setRho(param->getRho()/scale);
+  param->setTheta(param->getTheta()/scale);
+  double newlprior=rectree->prior(param) + param->logPriorOfRho() + param->logPriorOfTheta();
[…]

### 2011-10-18 src/warg2/movescaletree.h

@@ -14,15 +14,15 @@ class MoveScaleTree : public Move
 {
 
 public:
-    MoveScaleTree(Param * p,double a);
-    Move * clone()
-    {
-        return new MoveScaleTree(*this);
-    };
-    int move();
-    ~MoveScaleTree();
+  MoveScaleTree(Param * p,double a);
+  Move * clone()
+  {
+    return new MoveScaleTree(*this);
+  };
+  int move();
+  ~MoveScaleTree();
 
-    void scaleTree(double scale);///<Scales the tree by a factor scale
+  void scaleTree(double scale);///<Scales the tree by a factor scale
 };
 
 } // end namespace weakarg

### 2011-10-18 src/warg2/movesitechange.h

@@ -12,20 +12,20 @@ namespace weakarg
 class MoveSiteChange : public Move
 {
 public:
-    MoveSiteChange(Param * p,double a);
-    Move * clone()
-    {
-        return new MoveSiteChange(*this);
-    }
-    int move(vector<int> * samplespace=NULL);
-    inline int move()
-    {
-        return(move(NULL));
-    }
-    ~MoveSiteChange();
-    double sumVec(std::vector<double> vec,int vmax=-1);///<Returns the sum of a vector up to vmax or the end of the vector
-    void RestoreLik(std::vector<double> *store,int dsign,int start,int end,int movestart,int imin=0,int imax=-1);///< Restores the likelihood from a stored vector between a given range
-    std::vector<double> calcPost(std::vector<double> *store1,std::vector<double> *store2,int which, int dsign,int movestart,int maxdist, bool keeplik);///<Calculates the posterior distribution of a move
+  MoveSiteChange(Param * p,double a);
+  Move * clone()
+  {
+    return new MoveSiteChange(*this);
+  }
+  int move(vector<int> * samplespace=NULL);
+  inline int move()
+  {
+    return(move(NULL));
+  }
+  ~MoveSiteChange();
+  double sumVec(std::vector<double> vec,int vmax=-1);///<Returns the sum of a vector up to vmax or the end of the vector
+  void RestoreLik(std::vector<double> *store,int dsign,int start,int end,int movestart,int imin=0,int imax=-1);///< Restores the likelihood from a stored vector between a given range
+  std::vector<double> calcPost(std::vector<double> *store1,std::vector<double> *store2,int which, int dsign,int movestart,int maxdist, bool keeplik);///<Calculates the posterior distribution of a move
 
 };
 

### 2011-10-18 src/warg2/movetheta.h

@@ -12,13 +12,13 @@ namespace weakarg
 class MoveTheta : public Move
 {
 public:
-    MoveTheta(Param * p,double a);
-    Move * clone()
-    {
-        return new MoveTheta(*this);
-    };
-    int move();
-    ~MoveTheta();
+  MoveTheta(Param * p,double a);
+  Move * clone()
+  {
+    return new MoveTheta(*this);
+  };
+  int move();
+  ~MoveTheta();
 
 
 };

### 2011-10-18 src/warg2/movetimechange.cpp

@@ -8,128 +8,128 @@ namespace weakarg
 {
 
 MoveTimeChange::MoveTimeChange(Param * p,double a)
-    : Move(p,a)
+  : Move(p,a)
 {
-    description="Update moving the time of a recedge";
-    name= "TimeChange";
+  description="Update moving the time of a recedge";
+  name= "TimeChange";
 }
 
 int MoveTimeChange::move(vector<int> * samplespace)
 {
-    RecTree * rectree=param->getRecTree();
-    int which=rectree->sampleEdge(samplespace);//floor(gsl_rng_uniform(rng)*rectree->numRecEdge());
-    if (which<0)
-        {
-            dlog(1)<<"No valid edges to change!"<<endl;
-            return(-1);
-        }
-    numcalls++;
-    int movetto=gsl_rng_uniform_int(rng,2);//0 for no or 1 for yes
-    // collect the details of the edge
-    int efrom=rectree->getRecEdge(which)->getEdgeFrom();
-    int eto  =rectree->getRecEdge(which)->getEdgeTo  ();
-    double tfrom=rectree->getEdgeTimeAbsFrom(which);
-    double tto  =rectree->getEdgeTimeAbsTo  (which);
-    double tmptime,tfromnew=tfrom,ttonew=tto;
-    int start=rectree->getRecEdge(which)->getStart();
-    int end  =rectree->getRecEdge(which)->getEnd();
-    int tmpedge,etonew=eto,efromnew=efrom;
-    double l=param->getLL();
-    double lpriorrat=0;// log of (prior ratio * transition ratio)
-    double rootage=rectree->getNode(rectree->getN()*2-2)->getAge();
-    vector<double> store(end-start);
-    for (int i=start; i<end; i++)
-        store[i-start]=param->getLLsite(i);
-    dlog(1)<<"Proposing to change edge time "<<efrom<<":"<<tfrom<<"->"<<eto<<":"<<tto<<" to"<<flush;
+  RecTree * rectree=param->getRecTree();
+  int which=rectree->sampleEdge(samplespace);//floor(gsl_rng_uniform(rng)*rectree->numRecEdge());
+  if (which<0)
+    {
+      dlog(1)<<"No valid edges to change!"<<endl;
+      return(-1);
+    }
+  numcalls++;
+  int movetto=gsl_rng_uniform_int(rng,2);//0 for no or 1 for yes
+  // collect the details of the edge
[…]

### 2011-10-18 src/warg2/mydata.cpp

@@ -8,190 +8,190 @@ namespace weakarg
 
 Data::Data(int n,vector<int> blocks)
 {
-    this->n=n;
-    this->L=blocks.back();
-    this->blocks=blocks;
-    data=vector<string>(n,string(L,'N'));
-    poly=vector<bool>(L,true);
-    begEnd=vector<bool>(L+1,false);
-    for (unsigned int i=0; i<blocks.size(); i++)
-        begEnd[blocks[i]]=true;
+  this->n=n;
+  this->L=blocks.back();
+  this->blocks=blocks;
+  data=vector<string>(n,string(L,'N'));
+  poly=vector<bool>(L,true);
+  begEnd=vector<bool>(L+1,false);
+  for (unsigned int i=0; i<blocks.size(); i++)
+    begEnd[blocks[i]]=true;
 }
 
 Data::Data(string filename)
 {
-    string line;
-    ifstream file;
-    file.open(filename.data());//Open file
-    <<<<<<< .mine
-    if(!file.is_open())
+  string line;
+  ifstream file;
+  file.open(filename.data());//Open file
+  <<<<<<< .mine
+  if(!file.is_open())
+    {
+      cerr << "Unable to read file \""<<filename<<"\".\n";
+      throw "File unreadable\n";
+    }
+  unsigned int which=0;
+  =======
+    unsigned int which=-1;
+  >>>>>>> .r336
+  blocks.push_back(0);
+  while (1)
+    {
+      getline(file,line);//Read next line from file
+      if (file.eof())
+        break;//Stop if end of file
+      <<<<<<< .mine
+      if(line[0]=='=')
         {
-            cerr << "Unable to read file \""<<filename<<"\".\n";
-            throw "File unreadable\n";
-        }
-    unsigned int which=0;
-    =======
-        unsigned int which=-1;
-    >>>>>>> .r336
-    blocks.push_back(0);
-    while (1)
-        {
-            getline(file,line);//Read next line from file
-            if (file.eof())
-                break;//Stop if end of file
-            <<<<<<< .mine
-            if(line[0]=='=')
-                {
-                    // pad all sequences with - to match current size
-                    // this enables the subset blocks in Mauve XMFA format to be read
-                    unsigned int maxblock = 0;
-                    for(unsigned int i=0; i<data.size(); i++)
-                        {
-                            if(maxblock < data[i].size()) maxblock = data[i].size();
-                        }
-                    data.resize(29);
-                    for(unsigned int i=0; i<29; i++)
[…]

### 2011-10-18 src/warg2/node.h

@@ -14,68 +14,68 @@ namespace weakarg
 class Node
 {
 protected:
-    Node * left;///<Left daughter node
-    Node * right;///<Right daughter node
-    Node * father;///<Father node
-    double age;///<Age of the node
-    int id;///<Index of the node
+  Node * left;///<Left daughter node
+  Node * right;///<Right daughter node
+  Node * father;///<Father node
+  double age;///<Age of the node
+  int id;///<Index of the node
 public:
-    Node();
-    Node(std::string * newick,Node * father,std::vector<Node*> * nodes);
-    ~Node();
-    Node* cloneSubtree() const;
-    std::string newick(int p=6) const;///<Convert the node to a Newick string (precision options)
-    std::string newickNoInternalLabels(int p=6) const;///<Convert the node to a Newick string (precision options)
-    inline int getId() const
-    {
-        return id;
-    }///<Returns the index of the node
-    inline void setId(int i)
-    {
-        id=i;
-    }///<Sets the index of the node
-    inline double getDist() const
-    {
-        if (father==NULL)
-            return 0;
-        return father->age-age;
-    }///<Returns the distance of the node to its father
-    inline double getAge() const
-    {
-        return age;
-    }///<Returns the age of the node
-    inline void setAge(double a)
-    {
-        age=a;
-    }///<Sets the age of the node
-    inline Node * getLeft() const
-    {
-        return left;
-    }///<Returns the left daughter node
-    inline Node * getRight() const
-    {
-        return right;
-    }///<Returns the right daughter node
-    inline Node * getFather() const
-    {
-        return father;
-    }///<Returns the father node
-    inline void setFather(Node * f)
-    {
-        father=f;
-    }///<Sets the father of the Node
-    inline void setLeft(Node * l)
-    {
-        left=l;
-    }///<Sets the left child of the Node
-    inline void setRight(Node * r)
-    {
-        right=r;
-    }///<Sets the right child of the Node
-    inline void changeAge(double a)
-    {
-        age=age+a;
-    }///<Changes the age of the node by adding an amount a
+  Node();
+  Node(std::string * newick,Node * father,std::vector<Node*> * nodes);
+  ~Node();
+  Node* cloneSubtree() const;
+  std::string newick(int p=6) const;///<Convert the node to a Newick string (precision options)
[…]

### 2011-10-18 src/warg2/param.h

@@ -22,202 +22,202 @@ class Param
 {
 
 public:
-    Param();
-    Param(RecTree* rectree,Data* data);
-    virtual ~Param();
-    void readParamsFromFile(WargXml * infile,std::streampos itstart=-1);///<Reads the parameter values from an output file
-    void readProgramOptions();///<Sets MCMC parameters based on command-line options
-    void metropolis(string comment=string(""));///<Performs the Metropolis-Hastings algorithm
-    void simulateData(std::vector<int> blocks);///<Simulates data given the weak ARG, using the data structure provided
-    inline RecTree * getTree()
-    {
-        return rectree;
-    }///<Returns the tree
-    inline void setTree(RecTree * t)
-    {
-        if(rectree!=NULL) delete(rectree);
-        rectree=t;
-    }///<Sets the tree
-    inline Data * getData()
-    {
-        return data;
-    }///<Returns the Data
-    inline void setData(Data*bd)
-    {
-        if(data!=NULL) delete(data);
-        data=bd;
-    }///<Sets the data
-    void exportXMLbegin(std::ostream& out,string comment=string(""));///<Creates the beginning of an output file
-    void exportXMLend(std::ostream& out);///<Creates the end of an output file
-    void exportXMLiter(std::ostream& out, long i=-1);///<Creates an iteration in the output file
-    void computeLikelihood();///<Calculates the current likelihood
-    void computeLikelihood(unsigned int start, unsigned int end);///<Calculates the current likelihood for a range of sites
-    double getRlenPrior(int edge);///<Returns the log of the prior due to genetic position for a recedge
-    inline double getTheta() const
-    {
-        return theta;
-    }///<Returns the value of theta
-    inline void setTheta(double t)
-    {
-        theta=t;
-    }///<Sets the value of theta
-    inline double getRho() const
-    {
-        return rho;
-    }///<Returns the value of rho
-    inline void setRho(double r)
-    {
-        rho=r;
-    }///<Sets the value of rho
-    inline double getDelta() const
-    {
-        return delta;
-    }///<Returns the value of delta
-    inline void setDelta(double d)
-    {
-        delta=d;
-    }///<Sets the value of delta
-    inline double getLL() const
-    {
-        return ll/tempering;
-    }///<Returns the value of the likelihood
[…]

### 2011-10-18 src/warg2/recedge.cpp

@@ -6,17 +6,17 @@ namespace weakarg
 
 RecEdge::RecEdge(double tfrom,double tto,unsigned int gstart,unsigned int gend,unsigned int edgefrom,unsigned int edgeto)
 {
-    MoveEdge(tfrom,tto,gstart,gend,edgefrom,edgeto);
+  MoveEdge(tfrom,tto,gstart,gend,edgefrom,edgeto);
 }
 
 void RecEdge::MoveEdge(double tfrom,double tto,unsigned int gstart,unsigned int gend,unsigned int edgefrom,unsigned int edgeto)
 {
-    timeFrom=tfrom;
-    timeTo=tto;
-    this->gstart=gstart;
-    this->gend=gend;
-    edgeFrom=edgefrom;
-    edgeTo=edgeto;
+  timeFrom=tfrom;
+  timeTo=tto;
+  this->gstart=gstart;
+  this->gend=gend;
+  edgeFrom=edgefrom;
+  edgeTo=edgeto;
 }
 
 RecEdge::~RecEdge()

### 2011-10-18 src/warg2/rectree.h

@@ -24,156 +24,156 @@ class RecTree:public Tree
 {
 protected:
 // Variables
-    std::vector<std::vector<int> > tabSons;///<Contains the branching order for the current local tree
-    std::vector<std::vector<double> > tabSonsDist;///<Contains the distances for the current local tree
-    std::vector<int> tabNode;///<Used only in makeLocalTree (made global for speed)
-    std::vector<int> tabRec;///<Used only in makeLocalTree (made global for speed)
-    std::vector<int> tabFather;///<Used only in makeLocalTree (made global for speed)
-    std::vector<double> age;///<Used only in makeLocalTree (made global for speed)
-    std::vector<int> tabEdge;///<Used only in makeLocalTree (made global for speed)
-    std::vector<bool> sameLTasPrev;///<Indicates for each site whether the local tree is the same as that of the previous site
-    unsigned int L;///< Number of sites in the sequences
-    std::vector<RecEdge *> edge;///< The recombinant edges
+  std::vector<std::vector<int> > tabSons;///<Contains the branching order for the current local tree
+  std::vector<std::vector<double> > tabSonsDist;///<Contains the distances for the current local tree
+  std::vector<int> tabNode;///<Used only in makeLocalTree (made global for speed)
+  std::vector<int> tabRec;///<Used only in makeLocalTree (made global for speed)
+  std::vector<int> tabFather;///<Used only in makeLocalTree (made global for speed)
+  std::vector<double> age;///<Used only in makeLocalTree (made global for speed)
+  std::vector<int> tabEdge;///<Used only in makeLocalTree (made global for speed)
+  std::vector<bool> sameLTasPrev;///<Indicates for each site whether the local tree is the same as that of the previous site
+  unsigned int L;///< Number of sites in the sequences
+  std::vector<RecEdge *> edge;///< The recombinant edges
 // Functions
-    int randomActiveNode(std::vector<int> nodelist);///<Returns a random node id from a vector of indicators for alive (-1 for not alive at current time)
-    int getEdgeCoal(std::vector<int> nl,int k,double* time);///<Returns the node the recombinant edge is from and sets the time
+  int randomActiveNode(std::vector<int> nodelist);///<Returns a random node id from a vector of indicators for alive (-1 for not alive at current time)
+  int getEdgeCoal(std::vector<int> nl,int k,double* time);///<Returns the node the recombinant edge is from and sets the time
 
[…]

### 2011-10-18 src/warg2/slotallocator.h

@@ -39,48 +39,48 @@ template< class T >
 class SlotAllocator
 {
 public:
-    static SlotAllocator<T>& GetSlotAllocator();
-    T* Allocate();
-    void Free( T* t );
-    void Free( std::vector<T*>& chunk );
-    ~SlotAllocator()
-    {
-        Purge();
-    };
-    void Purge()
-    {
-        std::vector<T*>& data = this->data;
-        unsigned& tail_free = this->tail_free;
-        unsigned& n_elems = this->n_elems;
-        std::vector< T* >& free_list = this->free_list;
-        for( unsigned dataI = 0; dataI < data.size(); dataI++ )
-            free(data[dataI]);
-        data.clear();
-        free_list.clear();
-        tail_free = 0;
-        n_elems = 0;
-    }
+  static SlotAllocator<T>& GetSlotAllocator();
+  T* Allocate();
+  void Free( T* t );
+  void Free( std::vector<T*>& chunk );
+  ~SlotAllocator()
+  {
+    Purge();
+  };
+  void Purge()
+  {
+    std::vector<T*>& data = this->data;
+    unsigned& tail_free = this->tail_free;
+    unsigned& n_elems = this->n_elems;
+    std::vector< T* >& free_list = this->free_list;
+    for( unsigned dataI = 0; dataI < data.size(); dataI++ )
+      free(data[dataI]);
+    data.clear();
+    free_list.clear();
+    tail_free = 0;
+    n_elems = 0;
+  }
 
 protected:
-    std::vector<T*> data;
-    unsigned tail_free;
-    unsigned n_elems;	/**< number of T in the most recently allocated block */
+  std::vector<T*> data;
+  unsigned tail_free;
+  unsigned n_elems;	/**< number of T in the most recently allocated block */
 
-    std::vector< T* > free_list;
+  std::vector< T* > free_list;
 
 private:
-    SlotAllocator() : tail_free(0), n_elems(0) {};
-    SlotAllocator& operator=( SlotAllocator& sa )
-    {
-        n_elems = sa.n_elems;
-        data = sa.data;
-        tail_free = sa.tail_free;
-        return *this;
-    };
-    SlotAllocator( SlotAllocator& sa )
-    {
-        *this = sa;
-    };
+  SlotAllocator() : tail_free(0), n_elems(0) {};
+  SlotAllocator& operator=( SlotAllocator& sa )
+  {
[…]

### 2011-10-18 src/warg2/tree.cpp

@@ -13,579 +13,579 @@ namespace weakarg
 
 Tree::Tree(int n)
 {
-    this->n=n;
-    nodes=vector<Node*>(n+n-1);
-    Node ** toCoal=(Node **)calloc(n,sizeof(Node*));
-    int k=n;
-    double time=0.0;
-    for (int i=0; i<n; i++)
-        {
-            nodes[i]=SlotAllocator<Node>::GetSlotAllocator().Allocate();
-            new (nodes[i]) Node();
-            nodes[i]->setId(i);
-            toCoal[i]=nodes[i];
-        }
-    while (k>1)
-        {
-            time-=2.0/(k*(k-1.0))*log(gsl_rng_uniform(rng));
-            int a=(int)floor(gsl_rng_uniform(rng)*k);
-            int b=a;
-            while (b==a)
-                b=(int)floor(gsl_rng_uniform(rng)*k);
-            nodes[n+n-k]=SlotAllocator<Node>::GetSlotAllocator().Allocate();
-            new (nodes[n+n-k]) Node();
-            nodes[n+n-k]->setId(n+n-k);
-            nodes[n+n-k]->setLeft(toCoal[a]);
-            nodes[n+n-k]->setRight(toCoal[b]);
-            nodes[n+n-k]->setAge(time);
-            toCoal[a]->setFather(nodes[n+n-k]);
-            toCoal[b]->setFather(nodes[n+n-k]);
-            toCoal[a]=nodes[n+n-k];
-            toCoal[b]=toCoal[k-1];
-            k--;
-        }
-    root=nodes[n+n-2];
-    free(toCoal);
-    computeTTotal();
+  this->n=n;
+  nodes=vector<Node*>(n+n-1);
+  Node ** toCoal=(Node **)calloc(n,sizeof(Node*));
+  int k=n;
+  double time=0.0;
+  for (int i=0; i<n; i++)
+    {
+      nodes[i]=SlotAllocator<Node>::GetSlotAllocator().Allocate();
+      new (nodes[i]) Node();
+      nodes[i]->setId(i);
+      toCoal[i]=nodes[i];
+    }
+  while (k>1)
+    {
+      time-=2.0/(k*(k-1.0))*log(gsl_rng_uniform(rng));
+      int a=(int)floor(gsl_rng_uniform(rng)*k);
+      int b=a;
+      while (b==a)
[…]

### 2011-10-18 src/warg2/wargxml.cpp

@@ -9,92 +9,92 @@ namespace weakarg
 
 WargXml::WargXml(std::string f)
 {
-    fname=f;
-    iterfile.open(fname.data());
-    if (!iterfile)
-        {
-            cerr << "Can't open input file " << f << endl;
-            exit(1);
-        }
+  fname=f;
+  iterfile.open(fname.data());
+  if (!iterfile)
+    {
+      cerr << "Can't open input file " << f << endl;
+      exit(1);
+    }
 }
 
 WargXml::~WargXml()
 {
-    if(iterfile.is_open()) iterfile.close();
+  if(iterfile.is_open()) iterfile.close();
 }
 
 streampos WargXml::gotoLineContaining(std::string str,bool getlast)
 {
-    streampos pos =iterfile.tellg (),it=-1;
-    string res;
-    size_t found;
-    int len=str.length();
-    while(1)
-        {
-            pos =iterfile.tellg ();
-            getline(iterfile,res);
-            found=res.find_first_of('<');
-            if(found!=string::npos) if(res.length()>=len+found) if(res.substr(found,len).compare(str)==0)
-                        {
-                            it=pos;
-                        }
-            if (iterfile.eof()||!iterfile.good()||(!getlast && it>=0)) break;
-        }
-    iterfile.clear();
-    iterfile.seekg(it);
-    return(it);
+  streampos pos =iterfile.tellg (),it=-1;
+  string res;
+  size_t found;
+  int len=str.length();
+  while(1)
+    {
+      pos =iterfile.tellg ();
+      getline(iterfile,res);
+      found=res.find_first_of('<');
+      if(found!=string::npos) if(res.length()>=len+found) if(res.substr(found,len).compare(str)==0)
+            {
+              it=pos;
+            }
+      if (iterfile.eof()||!iterfile.good()||(!getlast && it>=0)) break;
+    }
+  iterfile.clear();
+  iterfile.seekg(it);
+  return(it);
 }
 
 std::streampos WargXml::getLineContaining(std::string str,bool getlast)
 {
-    std::streampos p0=iterfile.tellg();
-    std::streampos r=gotoLineContaining(str,getlast);
-    iterfile.seekg(p0);
-    return(r);
+  std::streampos p0=iterfile.tellg();
+  std::streampos r=gotoLineContaining(str,getlast);
+  iterfile.seekg(p0);
+  return(r);
 }
 
[…]

### 2011-10-18 src/warg2/weakarg.cpp

@@ -22,8 +22,8 @@ vector<double> readInputFiles(Data* &data, RecTree* &rectree,vector<double> &sum
 
 ProgramOptions& opt()
 {
-    static ProgramOptions po;	// define a single instance of ProgramOptions per process.
-    return po;
+  static ProgramOptions po;	// define a single instance of ProgramOptions per process.
+  return po;
 }
 
 
@@ -31,22 +31,22 @@ ProgramOptions& opt()
 
 string getVersion()
 {
-    string ret;
+  string ret;
 #ifdef PACKAGE_STRING
-    ret.append(PACKAGE_STRING);
+  ret.append(PACKAGE_STRING);
 #else
-    ret.append("warg");
+  ret.append("warg");
 #endif
-    ret.append(" build date ");
-    ret.append(__DATE__);
-    ret.append(" at ");
-    ret.append(__TIME__);
-    return(ret);
+  ret.append(" build date ");
+  ret.append(__DATE__);
+  ret.append(" at ");
+  ret.append(__TIME__);
+  return(ret);
 }
 
 void printVersion()
 {
-    cout<<getVersion()<<endl;
+  cout<<getVersion()<<endl;
 }
 
 using namespace weakarg;
@@ -54,7 +54,7 @@ using namespace weakarg;
 // the linker can find it
 
 static const char * help=
-    "\
+  "\
     Usage: weakarg [OPTIONS] treefile datafile outputfile\n\
     \n\
     Options:\n\
@@ -106,307 +106,307 @@ static const char * help=
 
 int main(int argc, char *argv[])
 {
-    string comment="Command line: ";
-    for(int c1=0; c1< argc; c1++)
-        {
-            comment.append(argv[c1]);
-            comment.append(" ");
-        }
-    comment.append("\nVersion: ");
-    comment.append(getVersion());
-    vector<string> inputfiles;
-    initmpi(argc,argv);
-    makerng(true);
-    //optind=0;
-    bool upgma=false;
-    int c;
-    char * pch;
-    double simparrho=50.0;
-    double simpartheta=100.0;
-    double simpardelta=500.0;
-    int simparN=10;
-    int simparnumblocks=8;
-    int simparblocksize=500;
-    std::stringstream ss;
-    unsigned long seed=0;
-    bool readparams=false;
-    bool setregions=false;
-    while ((c = getopt (argc, argv, "w:x:y:z:s:va:T:R:D:L:C:r:t:i:S:G:fUhV")) != -1)
-        switch (c)
-            {
-            case('w'):
[…]

### 2011-10-18 src/warg2/weakarg.h

@@ -19,47 +19,47 @@ static const int NUMMOVES=11;
 class ProgramOptions
 {
 public:
-    ProgramOptions() :
-        theta(-1.0),
-        rho(-1.0),
-        thetaPerSite(false),
-        rhoPerSite(false),
-        delta(-1.0),
-        preburnin(100000),
-        burnin(100000),
-        additional(100000),
-        thinin(100),
-        allowclonal(true),
-        temperreps(0),
-        temperT(1.0),
-        subset(std::vector<int>(0)),
-        subsetSeed(-1),
-        greedyWeight(0),
-        verbose(0)
-    {
-        movep.resize(NUMMOVES,1.0);
-    }
-    std::vector<double> movep; //length (NUMMOVES) vector with weights set to default 1
+  ProgramOptions() :
+    theta(-1.0),
+    rho(-1.0),
+    thetaPerSite(false),
+    rhoPerSite(false),
+    delta(-1.0),
+    preburnin(100000),
+    burnin(100000),
+    additional(100000),
+    thinin(100),
+    allowclonal(true),
+    temperreps(0),
+    temperT(1.0),
+    subset(std::vector<int>(0)),
+    subsetSeed(-1),
+    greedyWeight(0),
+    verbose(0)
+  {
+    movep.resize(NUMMOVES,1.0);
+  }
+  std::vector<double> movep; //length (NUMMOVES) vector with weights set to default 1
 
-    double theta;
-    double rho;
-    bool thetaPerSite;
-    bool rhoPerSite;
-    double delta;
-    int preburnin;
-    int burnin;
-    int additional;
-    int thinin;
-    bool allowclonal;
-    int temperreps;
-    double temperT;
-    std::vector<int> subset;
-    int subsetSeed;
-    double greedyWeight;
-    int verbose;
-    std::string outfile;
-    std::string csvoutfile;
-    std::string logfile;
+  double theta;
+  double rho;
+  bool thetaPerSite;
+  bool rhoPerSite;
+  double delta;
+  int preburnin;
+  int burnin;
+  int additional;
+  int thinin;
+  bool allowclonal;
+  int temperreps;
+  double temperT;
+  std::vector<int> subset;
+  int subsetSeed;
+  double greedyWeight;
+  int verbose;
+  std::string outfile;
+  std::string csvoutfile;
+  std::string logfile;
 };
 ProgramOptions& opt();
 

### 2011-10-18 src/warg2/weakargsim.cpp

@@ -61,8 +61,8 @@ vector<int> readBlock (string& filename, unsigned b = 0);
 
 ProgramOptions& opt()
 {
-    static ProgramOptions po;	// define a single instance of ProgramOptions per process.
-    return po;
+  static ProgramOptions po;	// define a single instance of ProgramOptions per process.
+  return po;
 }
 
 
@@ -70,22 +70,22 @@ ProgramOptions& opt()
 
 string getVersion()
 {
-    string ret;
+  string ret;
 #ifdef PACKAGE_STRING
-    ret.append(PACKAGE_STRING);
+  ret.append(PACKAGE_STRING);
 #else
-    ret.append("warg");
+  ret.append("warg");
 #endif
-    ret.append(" build date ");
-    ret.append(__DATE__);
-    ret.append(" at ");
-    ret.append(__TIME__);
-    return(ret);
+  ret.append(" build date ");
+  ret.append(__DATE__);
+  ret.append(" at ");
+  ret.append(__TIME__);
+  return(ret);
 }
 
 void printVersion()
 {
-    cout<<getVersion()<<endl;
+  cout<<getVersion()<<endl;
 }
 
 using namespace weakarg;
@@ -93,7 +93,7 @@ using namespace weakarg;
 // the linker can find it
 
 static const char * help=
-    "\
+  "\
     Usage: weakargsim [OPTIONS] treefile datafile outputfile\n\
     \n\
     Options:\n\
@@ -151,7 +151,7 @@ static const char * help=
  */
 void ShowUsage()
 {
-    cout << help << endl;
+  cout << help << endl;
 }
 
 /**
@@ -188,100 +188,100 @@ enum { OPT_HELP,
  */
 CSimpleOpt::SOption g_rgOptions[] =
 {
-    // ID       TEXT          TYPE
-    {
-        OPT_NUM_ISOLATES,
-        "-N", SO_REQ_SEP
-    }, // "-N ARG"
-    {
-        OPT_THETA,
-        "-T", SO_REQ_SEP
-    }, // "-T ARG"
-    {
-        OPT_RHO,
-        "-R", SO_REQ_SEP
-    }, // "-R ARG"
-    {
-        OPT_DELTA,
-        "-D", SO_REQ_SEP
-    }, // "-D ARG"
-    {
-        OPT_LENGTH_FRAGMENT,
-        "-B", SO_REQ_SEP
-    }, // "-B ARG"
[…]

### 2011-10-18 doc/manuscript_plos.tex
I have to keep revisioning of the manuscript.

@@
 
+\section*{Things To Do for submitting this to PLoS Pathogens}
+\begin{enumerate}
+\item A paragraph for briefing the methods of ClonalOrigin and Mowgli before
+section \textbf{Results}.
+\item We can add at least 50 words in section \textbf{Abstract}.
+\item Figures, tables, and supporting information need to be ready.
+\item Publish source codes at public domain.
+\end{enumerate}
+
+\section*{A list our major and minor findings or observations}
+\begin{enumerate}
+\item We showed that recombining gene transfer from SPY to SDE was strong
+enough.
+\item A few evidences for the net directionality from SPY to SDE:
+time-constrained gene transfer between SPY to SDE also showed the directionality
+from SPY to SDE.
+\item We showed that virulence genes were associated with horizontal gene
+transfer. 
+\item We found that recombining gene tranfer processes are involved in
+membrane-bound protein encoding genes and aminoacyl tRNA synthetases.
+\item We found that horizontal gene transfer processes are involved in
+transposases.
+\item We provided evidence that recombination process affected more sites along
+the core genome of SPY and SDE than mutation process.
+\item We found a recombination hotspot at 200 kb from the origin of replication
+along the genome of SDE1.
+\item We showed that species-tree-topology-changing recombinant edges were more
+reliably recovered than species-tree-topology-not-changing edges.
+\item We showed that recombination intensity might be useful in finding genes
+under high recombining gene transfer.
+\item We found that SPY and SDE are more closely related than SDD and SDE in
+about 10\% of the core genome.
+\item Recombination intensity and number of recombining gene transfer events
+were positively correlated.
+\item We found reduction of numbers of genes after divergence of SDE and SPY.
+\item We confirmed the species tree that relates the five genomes.
+\item We listed several virulence genes associated with recombining and
+horizontal gene transfer.
+\end{enumerate}
+
+\section*{Suggestions about the manuscript}
+\begin{enumerate}
+\item Ilan thought that a species tree must be shown at the outset of the
+section \textbf{Results}.
+\item Ilan suggested a different order of four parts in \textbf{Results}:
+preliminary, funtional category and virulence gene analyses, gene gain and loss,
+and directionality.
+\end{enumerate}
+
 % Please keep the abstract between 250 and 300 words
@@
 organization and human pathogenic virulence factors of SDE could have originated
-in those of SPY, or vice versa.  Our genome-wide survey of gene transfer
+in those of SPY, or vice versa.  Two mechanisms of gene transfer in bacteria
+were known; recipient bacteria of transferred genes can substitute transferred
+genes for homologous parts of their genomes (Recombining Gene Transfer), or
+integrate transferred genes into their genomes without replacement (Horizontal
+Gene Transfer).  We discriminated the two processes by using a model-based
+approach for studying recombining gene transfer, and using a parsimony-based
+approach for both transfer processes.  Our genome-wide survey of gene transfer
 indicated that gene transfer via homologous recombination occurred more often
-for the net direction from SPY to SDE.  We found evidence that virulence genes
-transferred more often via non-recombining fashion without replacing parts of
-destination genomes than did non-virulence genes.  A genome-wide resource of
+for the net direction from SPY to SDE.  We also found evidence that virulence
+genes transferred more often via non-recombining fashion without replacing parts
+of destination genomes than did non-virulence genes.  
+A genome-wide resource of
 recombining and non-recombining gene transfer for the study of the SPY and SDE
@@
 
-Homologous recombining gene transfer could be studied more readily using the
-core genomes of individual bacteria while horizontal gene transfer might be more
-easily detected using the pan genomes. Recently, a method of inferring
+Recently, a method of inferring
 homologous recombination with whole bacterial genomes was developed
@@
 flows of genetic segments between bacteria individuals of \textit{Bacillus
-cereus} species in the core genomes.  Other evolutionary processes acting on
+cereus} species in the core genomes.   
+% Descriptions of ClonalOrigin model.
+Other evolutionary processes acting on
 bacterial genomes include genome rearragement, translocation, inversion,
@@
 Instead, a parsimony-based approach to inferring gene duplication, loss, and
-transfer events was developed by Doyon et al.\ \citet{Doyon2011}.  We noted that the first
+transfer events was developed by Doyon et al.\ \citet{Doyon2011}.  
+% Descriptions of Mowgli analsysis.
+We noted that the first
 approach could be used to infer transfer of genomic segments due to homologous
@@
 transfer.
+% Homologous recombining gene transfer could be studied more readily using the
+% core genomes of individual bacteria while horizontal gene transfer might be more
+% easily detected using the pan genomes.
 
@@
 genome but also in the pan genome, which made the analysis complete in terms of
-the amount of used genome data. Our analysis of core genomes also showed for the
+the amount of used genome data. Our analysis of the core genomes also showed for the
 first time that the net directionality of recombining gene transfer from SPY to
-SDE was stronger than the reverse direction.
+SDE was stronger at the genome scale.
 
@@
 \section*{Results}
+
+\subsection*{Distinguishing two gene transfer processes shows different functional
+categories associations}
+
+Bacterial gene transfer processes as an evolutionary force would affect more
+frequently in some biological functions than in others.  In our attempt of
+finding which biological functions were associated with recombining or
+horizontal gene transfers, we devised a measure of recombination intensity for
+genes by using frequency of recombinant edges that affect those genes, and used
+frequency of horizontal gene transfer events (see section \textbf{Materials and
+Methods}).  Table \ref{tab:functional} shows a list of function categories
+that were found to be associated with recombining gene transfers. We noted three
+biological functions in the list: membrane-bound molecules, aminoacyl tRNA
+synthetases, and ribosomes.
+
+Firstly, we have sugar phosphotransferases ranked
+top in the list; phosphrylation could be important in signal transduction to
+recognize environmental changes.  Transport functional categories of
+carbohydrate transport and sodium ion transport would belong to genes encoding
+membrane-bound molecules.  Also highly ranked were fatty acid and lipid
+biosyntheses that must be involved in producing membrane-bound molecules. 
+Because bacteria as a unicellular organism should be able to respond to changes
+to environment, some evolutionary pressures must have been acting on genes
+encoding membrane-bound molecules. Therefoure, recombining gene transfer process
+appeared to be more associated with biological functions that had been critical
+of sensing queues of bacterial environmental changes.
+
+Secondly, we found several aminoacyl tRNA synthetases (GO Term translation
+GO:0006412) in association with recombining gene transfer.  In the
+parsimony-based approach the term translation was also top ranked (Table
+S\ref{tab:go-events-recombining}).  Because they had been a focus of gene
+transfer in bacterial species \citep{Woese2000}, we could confirmed that the
+recombination intensity and requency of recombining gene transfer events would
+be useful in finding genes associated with recombining gene transfer.
+
+Thirdly, functional categories concerning ribosomes were unexpectedly
+found to be associated with recombining gene transfer.  Because genes related
+with ribosomes were considered to be housekeeping genes, gene transfer in which
+would be rare, it was unexpected that ribosome related genes were found to be
+associated with recombining gene transfer.  Because it was a bizarre result, we
+investigated the association by considering different recombinant edge types. We
+noted that two different classes of recombinant edge types: one that changes
+species tree topologies, and another that does not
+(see section \textbf{Materials and Methods}).
+Because recombinant edges
+that did not change species tree topology might be difficult to infer reliably,
+considering those recombinant edges in computing recombination intensity might
+lead to dubious results.  Ribosome related functional categories were found to
+be top ranked if we took those recombinant edges that did not change species
+tree topology (Tables S\ref{tab:functional-notopology} and
+S\ref{tab:functional-all}).  In Table \ref{tab:functional} ribosome related
+functional categories were not top ranked.  Therefore, high recombination
+intensity in ribosome related genes were likely to be attributed to
+topology-not-changing recombinant edges. We would be cautious of recombining
+transfer of genes related with ribosome related functions.
+
+Table \ref{tab:go-events} shows a list of functional categories associated with
+horizontal gene transfer, which were different from those from recombining gene
+transfer analysis.  Ranked top in the list was transposases, which were known to
+help foreign genes be excised from a donor genome and integrated into a
+recipient genome. The biological role of transposases was concordant with
+horizontal gene transfer process because the process involved in transposases
+would allow bacteria to acquire foreign genes without the requirement of
+homologous regions in the bacteria.  
+These differences in the profile of functional categories association of the two
+gene transfer processes were because of the approach of dicriminating the two
+gene transfer processes.
+
+\subsection*{Recombining gene transfer from SPY to SDE at genome-wide scale}
+
+We show a net directionality of recombining gene transfers from SPY to SDE.
+
+Describe our arguments of the net directionality. 
+Number of recombinant edges.
+Confidence about the numbers of recombinant edges.
+Not all of genes experience recombining gene transfers from SPY to SDE.
+Gene transfer counts from the parsimony-based approach; the power of the
+parsimony-based approach may be limited.
+Is this number of recombinant edges good enough for the argument?
+
+Reflection of the previous analyses, 
+heat maps of real data analysis, 
+and supporting simulations.
 
@@
 SDE branch.  Transfer events between the SPY and SDE clades were inferred in
-smaller numbers (<6) compared to those within each clade. We were puzzled by
+smaller numbers ($<6$) compared to those within each clade. We were puzzled by
 this finding because virulence genes were found to be associated with horizontal
@@
 homologous recombination.  We wished to showcase virulence genes using the
-result of our strudy of gene transfer.
+result of our study of gene transfer.
 
@@
 would allow us to be more confident of the results if the two analyses performed
-similarly Generally, phylogenetic methods of detecting gene transfers often take
+similarly. Generally, phylogenetic methods of detecting gene transfers often take
 advantage of discrepancy between a reference species tree and gene trees. The
@@
 intensity and the number of recombining gene transfer events for each gene.  We
-found a weak positive correlation of 0.39 with a very small significant P-value
+found a weak positive correlation of 0.39 with a significant P-value
 of $2.2\times10^{-16}$ for testing a null hypothesis of the correlation being
@@
 The process of gene transfer may or may not move whole genes from one species to
-another \citep[e.g.,][]{Chan2009}. We observed that units of recombining gene
+another \citep{Chan2009}. We observed that units of recombining gene
 transfer could be multiple genes (Table S\ref{tab:genes-transfer}). In some
@@
 transfer from SPY1 to SDE1 was inferred for 29\% of the aspartyl-tRNA synthetase
-gene. Recombinant tracts often were shown to span two genes.  In other cases,
+gene. Recombinant tracts were often shown to span two genes.  In other cases,
 the inferred transfer units spaned a set of as much as eight adjacent genes:
 loci from SpyM3\_1518 to SpyM3\_1525.
-
-\subsection*{Transporters, transferases, aminoacyl tRNA synthetases, and
-transposases}
-
-Molecular evolutionary forces could act on genetic materials of organisms, and
-changes to the genotypes might lead to some phenotypes that could change
-biological functions of the species. So, the bacterial gene transfer process as
-an evolutionary force could be associated with some biological functions. We
-attempted to find which biological functions were associated with recombining or
-horizontal gene transfers.  Several functional categories that were found to be
-associated with recombining gene transfer were related with either
-membrane-bound molecules or aminoacyl tRNA synthetases (Table
-\ref{tab:functional}).  However, functional categories associated with
-horizontal gene transfer were different from those from recombining gene
-transfer analysis (Table \ref{tab:go-events}).  The top ranked functional
-category associated with horizontal gene transfer was related with transposases.
-These differences in the profile of functional categories between the two
-processes were interesting to us. An approach of dicriminating the two gene
-transfer processes would result in the differences.
-
-Because bacteria as a unicellular organism should be able to respond to changes
-to environment, some evolutionary pressures must have been acting on genes
-encoding membrane-bound molecules. Recombining gene transfer process seemed to
-be more associated with biological functions that had been critical of sensing
-queues of bacterial environmental changes.  Sugar phosphotransferase classes
-were ranked top in the list; phosphrylation could be important in signal
-transduction to recognize environmental changes.  The transport functional
-categories of carbohydrate transport and sodium ion transport also belonged to
-genes encoding membrane-bound molecules.  Also highly ranked were fatty acid and
-lipid biosyntheses that must be involved in producing membrane-bound molecules.
-Horizontal gene transfer process would allow bacteria to acquire foreign genes
-without the requirement of homologous regions in the bacteria.  Transposases
-would help foreign genes be excised from a donor genome and integrated into a
-recipient genome.
-
-In addition to these biological categories, we also found several aminoacyl tRNA
-synthetases (GO Term translation GO:0006412) in association with recombining
-gene transfer.  In the parsimony-based approach the term translation was top
-ranked (Table S\ref{tab:go-events-recombining}).  They had been a focus of gene
-transfer in bacterial species \citep{Woese2000}.  Some of the functional
-categories were unexpected; functional categories concerning ribosomes were
-found to be associated with recombining gene transfer.  Because genes related
-with ribosomes were considered to be housekeeping genes, gene transfer in which
-would be rare, it was unexpected that ribosome related genes were found to be
-associated with recombining gene transfer.  Because it was a bizarre result, we
-investigated the association by considering different recombinant edge types. We
-noted that two different classes of recombinant edge types: one that changes
-species tree topologies, and another that does not. Because recombinant edges
-that did not change species tree topology might be difficult to infer reliably,
-considering those recombinant edges in computing recombination intensity might
-lead to dubious results.  Ribosome related functional categories were found to
-be top ranked if we took those recombinant edges that did not change species
-tree topology (Tables S\ref{tab:functional-notopology} and
-S\ref{tab:functional-all}).  In Table \ref{tab:functional} ribosome related
-functional categories were not top ranked.  Therefore, high recombination
-intensity in ribosome related genes were likely to be attributed to
-topology-not-changing recombinant edges. We would be cautious of recombining
-transfer of genes related with ribosome related functions.
 
@@
 
-for each recombinant edge type to assess the effect of posterior recombinant
+For each recombinant edge type to assess the effect of posterior recombinant
 trees samples with respect to that of prior recombinant tree samples. The ten

### 2011-10-18 src/warg2/weakargsim.cpp

@@ -172,13 +172,13 @@ enum { OPT_HELP,
        OPT_XML_FILE,
        OPT_BLOCK_LENGTH,
        OPT_GENE_TREE,
+       OPT_CMD_MCMC,
        OPT_CMD_SIM_GIVEN_TREE,
        OPT_CMD_SIM_GIVEN_RECTREE,
        OPT_NUMBER_DATA,
        OPT_OUTPUT_FILE,
        OPT_CLONALTREE_FILE,
        OPT_LOCALTREE_FILE,
-       //OPT_DOT_FILE,
        OPT_INCLUDE_ANCESTRAL_MATERIAL
      };
 

### 2011-10-19 doc/manuscript.tex
Ilan's Results section edits replaced mine. Mine is after the document at the end.

@@
 \usepackage{subfig}
+
 \makeatother
 
-\usepackage{babel}
+% Leave date blank
+\date{}
+
+\pagestyle{myheadings}
+%% ** EDIT HERE **
+
+%% ** EDIT HERE **
+%% PLEASE INCLUDE ALL MACROS BELOW
+%% END MACROS SECTION
 
 \begin{document}
-\renewcommand\refname{Literature Cited}
-
-\raggedright
-
-\subsubsection*{Title: }
-
-Inferring homologous recombination and horizontal gene transfer 
+
+% Title must be 150 characters or less
+\begin{flushleft}
+{\Large
+\textbf{Inferring homologous recombination and horizontal gene transfer 
 using whole genomes of \textit{Streptococcus pyogenes} 
-and\textit{ S.\ dysgalactiae} subspecies \textit{equisimilis}
-
-\subsubsection*{Authors:}
-
-Sang Chul Choi$^{1}$, Matthew D. Rasmussen$^{1}$, Melissa Jane Hubisz$^{1}$, Ilan 
-Gronau$^{1}$,
-Michael J. Stanhope$^{2}$, and Adam Siepel$^{1}$
-
-\subsubsection*{Affiliations:}
-
-\lyxaddress{$^{1}$Department of Biological Statistics and Computational Biology,
-Cornell University, Ithaca, NY 14853, USA.}
-
-\lyxaddress{$^{2}$Department of Population Medicine and Diagnostic Sciences,
-College of Veterinary Medicine, Cornell University, Ithaca, NY 14853,
-USA.}
-
-\subsubsection*{Author for Correspondence:}
-
-Adam Siepel, Department of Biological Statistics and Computational
-Biology, Cornell University, Ithaca, NY 14853, USA, Phone: 607-254-1157,
-Fax: 607-255-4698, Email: acs4@cornell.edu\clearpage{}
-
-\subsubsection*{Abbreviations:}
-
-group A streptococci (GAS), group C streptococci (GCS), group G streptococci
-(GGS), \textit{Streptococcus pyogenes} (SPY), \textit{S.\ dysgalactiae}
-subspecies \textit{dysgalactiae} (SDD), \textit{S.\ dysgalactiae} subspecies
-\textit{equisimilis} (SDE)
-
-\subsubsection*{Keywords:}
-
-bacterial recombination, gene duplication and transfer, ClonalOrigin,
-ClonalFrame, Mowgli
-
-\subsubsection*{Version: 0.68 (Oct 15, 2011) \clearpage{}}
-
+and\textit{ S.\ dysgalactiae} ssp.\textit{\ equisimilis}}
+}
+% Short Title (running head): Gene transfer using Streptococcus genomes
+% Insert Author names, affiliations and corresponding author email.
+\\
+Sang Chul Choi$^{1}$, Matthew D. Rasmussen$^{1}$, 
+Melissa Jane Hubisz$^{1}$, Ilan Gronau$^{1}$,
+Michael J. Stanhope$^{2}$, and Adam Siepel$^{1,\ast}$
+\\
+\bf{1} 
+Department of Biological Statistics and Computational Biology,
+Cornell University, Ithaca, NY 14853, USA
+\\
+\bf{2}
+Department of Population Medicine and Diagnostic Sciences,
+College of Veterinary Medicine, Cornell University, Ithaca, NY 14853, USA
+\\
+$\ast$ E-mail: acs4@cornell.edu, Phone: 607-254-1157, Fax: 607-255-4698
+\end{flushleft}
+
+\section*{Things To Do for submitting this to PLoS Pathogens}
+\begin{enumerate}
+\item A paragraph for briefing the methods of ClonalOrigin and Mowgli before
+section \textbf{Results}.
+\item We can add at least 50 words in section \textbf{Abstract}.
+\item Figures, tables, and supporting information need to be ready.
+\item Publish source codes at public domain.
+\end{enumerate}
+
+\section*{A list our major and minor findings or observations}
+\begin{enumerate}
+\item We showed that recombining gene transfer from SPY to SDE was strong
+enough.
+\item A few evidences for the net directionality from SPY to SDE:
+time-constrained gene transfer between SPY to SDE also showed the directionality
+from SPY to SDE.
+\item We showed that virulence genes were associated with horizontal gene
+transfer. 
+\item We found that recombining gene tranfer processes are involved in
+membrane-bound protein encoding genes and aminoacyl tRNA synthetases.
+\item We found that horizontal gene transfer processes are involved in
+transposases.
+\item We provided evidence that recombination process affected more sites along
+the core genome of SPY and SDE than mutation process.
+\item We found a recombination hotspot at 200 kb from the origin of replication
+along the genome of SDE1.
+\item We showed that species-tree-topology-changing recombinant edges were more
+reliably recovered than species-tree-topology-not-changing edges.
+\item We showed that recombination intensity might be useful in finding genes
+under high recombining gene transfer.
+\item We found that SPY and SDE are more closely related than SDD and SDE in
+about 10\% of the core genome.
+\item Recombination intensity and number of recombining gene transfer events
+were positively correlated.
+\item We found reduction of numbers of genes after divergence of SDE and SPY.
+\item We confirmed the species tree that relates the five genomes.
+\item We listed several virulence genes associated with recombining and
+horizontal gene transfer.
+\end{enumerate}
+
+\section*{Suggestions about the manuscript}
+\begin{enumerate}
+\item Ilan thought that a species tree must be shown at the outset of the
+section \textbf{Results}.
+\item Ilan suggested a different order of four parts in \textbf{Results}:
+preliminary, funtional category and virulence gene analyses, gene gain and loss,
+and directionality.
+\end{enumerate}
+
+% Please keep the abstract between 250 and 300 words
 \section*{Abstract}
@@
 Instead, a parsimony-based approach to inferring gene duplication, loss, and
-transfer events was developed by \citet{Doyon2011}.  We noted that the first
+transfer events was developed by Doyon et al.\ \citet{Doyon2011}.  
+% Descriptions of Mowgli analsysis.
+We noted that the first
 approach could be used to infer transfer of genomic segments due to homologous
@@
 genome but also in the pan genome, which made the analysis complete in terms of
-the amount of used genome data. Our analysis of core genomes also showed for the
+the amount of used genome data. Our analysis of the core genomes also showed for the
 first time that the net directionality of recombining gene transfer from SPY to
-SDE was stronger than the reverse direction.
-
-\section{Results}
-
-\subsection{Recombination drived more substitutions than did point mutation}
-
-Five genomes except \textit{S.\ equi} ssp.\textit{\ equi} strain 4047 were
-chosen (Table S\ref{tab:genome}) to study recombining gene transfer using the
-model-based approach.  A genome alignment of the five genomes showed two
-distinct genome structures of SPY and SDE (Figure S\ref{fig:mauve}). The SDE and
-SDD genomes were larger in length than those of SPY. Although the graphical view
-of the genome alignment did not clearly show whether SDD was different from
-either of SDE1 and SDE2 in the genome organization, a spcies tree inferred from
-the genome alignment showed that the external SDD branch was attached to the
-internal clade branch of SDE1 and SDE2 (Figure S\ref{fig:tree5}). The
-distribution of the alignment block lengths was skewed to the right with mean of
-4188 base pairs long (Figure S\ref{fig:blocksize}), which supported that the
-model assumption of geometric block length distribution might be plausible in
-\texttt{ClonalOrigin} method (see section \textbf{Materials and Methods} and
-\citealt{Didelot2010}).  The relative strength ($r/m$) of recombination compared
-to mutations indicated that recombination-driven substitutions affected more
-sites of the core genomes than did mutation-driven substitutions (Table
-S\ref{tab:clonalframe}). This could imply that in the \textit{Streptococcus}
-species bacterial recombination process appeared to accelerate molecular
-evolution of more sites than did point mutations process.
-
-\subsection{Recombination hotspots were observed in genomes}
-
-Population mutation rate did not appear to change much across the genome of SDE1 
-(Figure S\ref{fig:scatter3}).  We found a few spots of genomic regions where the
-mutation rates were larger than twice the global estimate of mutation rate.  The
-recombinant tract length changed in a way that we could not note any particular
-positions with notable tract length values compared with other positions along
-the genome.  As recombinant tract lengths were smaller, recombination rates per
-site seemed larger (Figure S\ref{fig:rhologdelta}).  The recombination rate
-appeared to change more dynamically. Especially, a genomic region of SDE1 near
-200 kilobase pairs from the origin of replication contained alignment blocks with
-bell-shaped recombination rates.
-
-\subsection{SPY and SDE were more closely related in about tenth of the core genome}
-
-There are 105 possible rooted labeled topologies relating five taxa.  We expected the
-dominant topology across sampled genealogies to be concordant with that of the
-species tree.  Indeed, two thirds of the local gene tree topologies (see section
-\textbf{Materials and Methods}) along the alignment blocks were the same as the
-species tree, which allowed us to be more confident of the species tree inferred
-using the alignment blocks (Table S\ref{tab:Gene-tree-topologies}).  Other
-frequently sampled topologies were expected to be potentially informative about
-common recombination events.  The second most frequent topology (9.3\%) was one
-where SDD was the outgroup, and SPY and SDE were sister clades. Therefore, SPY
-and SDE appeared to be more closely related with each other than each to SDD's
-genome in about tenth of the core genome. The other topologies had less than 4\%
-of proportions.
-
-\subsection{Unbalaned recombining gene flow from SPY to SDE in genome-wide scale}
-
-The heat map of Figure \ref{fig:Heatmap-of-recombination} (see Tables
-S\ref{tab:heatmap} and S\ref{tab:obsheatmap} for the heat map values and the
-observed number of recombinant edges) shows the logarithm base 2 fractions of the
-\textit{a posteriori} number of recombinant edges for a pair of donor and
-recipient species tree branches and their corresponding expected values given
-the species tree (Figure S\ref{fig:tree5}) and the three population parameters
-(Table S\ref{tab:three}).  Because recombinant edges from any species tree
-branch to the root branch are impossible (see Figure S\ref{fig:clonalorigin}),
-the column for the root as recipient was colored green, indicating invalid
-recombinant edges in Figure \ref{fig:Heatmap-of-recombination}.  
-Similarly, all of the cells for which a recipient branch was
-older than the donor, and the donor and recipient branches did not share any contemporary time were
-colored green. The diagonal cells were colored gray, indicating no significant
-number of recombinant edges, because the observed and expected numbers of
-recombinant edges were similar. As recombinant edges that start from and end in
-the same species tree branch would not change the species tree topology, which
-might not
-change the likelihood much, the cells would be colored gray.  Because SDD is a
-strictly veterinary pathogenic bacteria, we did not expect that SDD received
-much gene transfer from either SPY1 or SPY2: the logarithm base 2 fraction from SPY1 to
-SDD was -2.20, and that from SPY2 to SDD was -1.97.  On the contrary, the
-recombing gene transfer from SDE1 to SDD was 0.756, and that from SDE2 to SDD
-was 0.698. This relatively higher recombining gene transfer from two SDE
-individuals to SDD compared with that from two SPY individuals to SDD reflected
-the fact
-that the SDD was more closely related to SDE than to SPY in most of the genomes,
-and the genomic organization between SDD and SDE1/SDE2 were much similar
-(Figure S\ref{fig:mauve}).  However, the gene transfer from SDD to either SDE1
-or SDE2 was less frequent than the case of reverse direction; SDD effectively
-served as a recipient rather than a donor. It was unclear why SDD received
-more genes from SDE1 or SDE2 than did SDE1 or SDE2 from SDD.
-
-Because it was expected that the recombining gene transfer within a clade would
-be more frequent than that between clades, not surprisingly the recombining gene
-transfer logarithm base 2 fractions between SDE1 and SDE2 (1.17 and 1.20) were about
-twice as large as those from SDE1 and SDE2 to SPY1 and SPY2 (0.67, 0.66, 0.60,
-and 0.39).  The intensities of recombining gene transfer from external branches
-of SDE1 and SDE2 to external branches of SPY1 and SPY2 was similar to those from
-SDE1 or SDE2 to SDD (0.76 and 0.70).  Therefore, either SPY1 or SPY2, and SDD as
-recipient appeared to be no different from the view point of SDE1 or SDE2 as
-donor since the similar values of recombining gene transfer intensity from SDE1
-or SDE2 to one of the other three external branches. Surprisingly, the recombining
-gene transfer logarithm base 2 fractions between SPY1 and SPY2 (1.06 and 1.07) was
-comparable with, yet even less than those from SPY1 and SPY2 to SDE1 and SDE2
-(1.25, 1.25, 1.17, and 1.29).  The recombining gene transfer from a clade SPY1 
-or SPY2 to a different clade SDE1 or SDE2 seemed to be effectively no different
-from that within the clade of SPY1 and SPY2.  The two external branches of SDE
-received genes from either SPY1 or SPY2 via recombining gene transfer as if SDE1
-and SDE2 were in the same clade as SPY1 and SPY2.  This was more surprising
-because the bacterial recombination between SPY and SDE seemed to happen more
-often than that between SDE and SDD even though the genome configurations of SDE
-and SDD were more similar than each to SPY.  However, this result could be
-concordant with the fact that SPY and SDE shared the same ecological niche
-despite their dissimilar genomic organization.  The gene transfer logarithm base 2
-fraction from SPY1 and SPY2 to SDE branch was also larger than any type of
-recombining gene transfer: 1.88 and 1.94, respectively.  Additionally we found
-58 genes for which a recombinant edge of certain types were sampled with
-probability at least 0.9 by focusing on specific types of recombinant edges
-(Table S\ref{tab:genes-transfer}).  Many of the genes in the list were
-transferred from SPY to SDE rather than the reverse direction overall.  All
-things considered, this epitomized that the recombining gene transfer was biased
-toward the net direction from SPY to SDE.
-
-We tried in vain to quantify differential horizontal gene transfer between SPY
-and SDE with the parsimony-based approach (Figure \ref{fig:hgt-heatmap}).  Among
-the clade-crossing transfers, we did not see any significant enrichment for a
-specific direction, as opposed to the result for recombining transfers from the
-model-based analysis.  SDE1 and SDE2 shared the most events (26 and 34) compared
-to other sample pairs.  SDD appeared to acquire many genes from the ancestral
-SDE branch.  Transfer events between the SPY and SDE clades were inferred in
-smaller numbers (<6) compared to those within each clade. We were puzzled by
-this finding because virulence genes were found to be associated with horizontal
-gene transfer, and yet we could not show net directionality of horizontal gene
-transfer between SPY and SDE.  However, we inferred relatively more recombining
-gene transfer events from the SPY branch to SDE branch even in the
-parsimony-based approach (Figure S\ref{fig:mowgli-recomb-heatmap}).
-
-\subsection{Reduction of genes and divergence of SPY and SDE species}
+SDE was stronger at the genome scale.
+
+% Results and Discussion can be combined.
+\section*{Results}
+
+\subsection*{Distinguishing two gene transfer processes shows different functional
+categories associations}
+
+Bacterial gene transfer processes as an evolutionary force would affect more
+frequently in some biological functions than in others.  In our attempt of
+finding which biological functions were associated with recombining or
+horizontal gene transfers, we devised a measure of recombination intensity for
+genes by using frequency of recombinant edges that affect those genes, and used
+frequency of horizontal gene transfer events (see section \textbf{Materials and
+Methods}).  Table \ref{tab:functional} shows a list of function categories
+that were found to be associated with recombining gene transfers. We noted three
+biological functions in the list: membrane-bound molecules, aminoacyl tRNA
+synthetases, and ribosomes.
+
+Firstly, we have sugar phosphotransferases ranked
+top in the list; phosphrylation could be important in signal transduction to
+recognize environmental changes.  Transport functional categories of
+carbohydrate transport and sodium ion transport would belong to genes encoding
+membrane-bound molecules.  Also highly ranked were fatty acid and lipid
+biosyntheses that must be involved in producing membrane-bound molecules. 
+Because bacteria as a unicellular organism should be able to respond to changes
+to environment, some evolutionary pressures must have been acting on genes
+encoding membrane-bound molecules. Therefoure, recombining gene transfer process
+appeared to be more associated with biological functions that had been critical
+of sensing queues of bacterial environmental changes.
+
+Secondly, we found several aminoacyl tRNA synthetases (GO Term translation
+GO:0006412) in association with recombining gene transfer.  In the
+parsimony-based approach the term translation was also top ranked (Table
+S\ref{tab:go-events-recombining}).  Because they had been a focus of gene
+transfer in bacterial species \citep{Woese2000}, we could confirmed that the
+recombination intensity and requency of recombining gene transfer events would
+be useful in finding genes associated with recombining gene transfer.
+
+Thirdly, functional categories concerning ribosomes were unexpectedly
+found to be associated with recombining gene transfer.  Because genes related
+with ribosomes were considered to be housekeeping genes, gene transfer in which
+would be rare, it was unexpected that ribosome related genes were found to be
+associated with recombining gene transfer.  Because it was a bizarre result, we
+investigated the association by considering different recombinant edge types. We
+noted that two different classes of recombinant edge types: one that changes
+species tree topologies, and another that does not
+(see section \textbf{Materials and Methods}).
+Because recombinant edges
+that did not change species tree topology might be difficult to infer reliably,
+considering those recombinant edges in computing recombination intensity might
+lead to dubious results.  Ribosome related functional categories were found to
+be top ranked if we took those recombinant edges that did not change species
+tree topology (Tables S\ref{tab:functional-notopology} and
+S\ref{tab:functional-all}).  In Table \ref{tab:functional} ribosome related
+functional categories were not top ranked.  Therefore, high recombination
+intensity in ribosome related genes were likely to be attributed to
+topology-not-changing recombinant edges. We would be cautious of recombining
+transfer of genes related with ribosome related functions.
+
+Table \ref{tab:go-events} shows a list of functional categories associated with
+horizontal gene transfer, which were different from those from recombining gene
+transfer analysis.  Ranked top in the list was transposases, which were known to
+help foreign genes be excised from a donor genome and integrated into a
+recipient genome. The biological role of transposases was concordant with
+horizontal gene transfer process because the process involved in transposases
+would allow bacteria to acquire foreign genes without the requirement of
+homologous regions in the bacteria.  
+These differences in the profile of functional categories association of the two
+gene transfer processes were because of the approach of dicriminating the two
+gene transfer processes.
+
+\subsection*{Recombining gene transfer from SPY to SDE at genome-wide scale}
+
+We show a net directionality of recombining gene transfers from SPY to SDE.
+
+Describe our arguments of the net directionality. 
+Number of recombinant edges.
+Confidence about the numbers of recombinant edges.
+Not all of genes experience recombining gene transfers from SPY to SDE.
+Gene transfer counts from the parsimony-based approach; the power of the
+parsimony-based approach may be limited.
+Is this number of recombinant edges good enough for the argument?
+
+Reflection of the previous analyses, 
+heat maps of real data analysis, 
+and supporting simulations.
+
+\subsection{Bayesian analysis of recombining gene transfer}
+
+{\bf IG: REMOVED A LOT OF DETAIL (MAINLY IN FIRST PART). PLEASE CHECK.}
+
+We followed the pipeline and methods suggested in \cite{Didelot2010} to pursue a model based study of recombining gene transfer.
+A preliminary step in this analysis involved obtaining an estimate of the species tree with branch lengths (via ClonalFrame), and estimates of the three population parameters: mutation rate, recombination rate, and average recombinant tract length (via a first phase of ClonalOrigin).
+Due to divergence of the SEE sample from the other five, we could not obtain suficiently long reliable multiple sequence alginments with it, and had to exclude it from the bayesian analysis.
+The spcies tree inferred by ClonalFrame over the five species has the expected topology (Fig.
+{\bf [ADD FIGURE 1]} (see Figure S\ref{fig:tree5} for branch length estimates).
+ClonalFrame also produces estimates for the relative strength of recombination compared
+to mutations ($r/m$; Table S\ref{tab:clonalframe}).
+The estimate of $r/m=4.66$ provides an initial indication to the important part recombination took in the evolutionary process of shaping these genmes: four times more sites have been affected by recombination-driven substitutions than have been affected by mutation-driven substitutions.
+Estimates of population mutation rate ($\theta$), recombination rate ($\rho$), and average tract length ($\delta$) were obtained for the different alignment blocks analyzed (Figure S\ref{fig:scatter3}).
+Global estimates of these three population parameters were obtained by a length-weighted average (Materials and Methods).
+The local estimates typically seem to be moderately fluctuating around the global average, with the exception of what apears to be a recombination hotspot at a region 200 kilobase pairs from the origin of replication of SDE1.
+
+The species tree and estimated branch lengths, together with the global estimates of the three population parameters are used in the second phase of ClonalOrigin, where recombinant trees are sampled from the posterior distriution.
+One way to summarize the posterior sample of recombinant trees provided by ClonalOrigin is to observe the topologies of the local genealogies implied by it.
+Out of the 105 possible rooted labeled topologies spanning five taxa, the dominant topology (sampled 66.7\% of the time) was that of the species tree (Materials and Methods; Table S\ref{tab:Gene-tree-topologies}).
+The second most frequently sampled topology (sampled 9.3\% of the time) was the one in which SPY and SDE are sister clades (with SDD as an outgroup).
+Other topologies were sampled with frequency less than 4\%.
+This gives initial indication to the extent of recombining gene transfer  between SPY and SDE.
+
+To gain a more fine-scale view of the main types of recombination scenarios sampled by ClonalOrigin, we looked at the recombinant edges present in the sampled recombinant trees.
+Following the approach of \cite{Didelot2010}, we classified these edges by types defined by distinct pairs of donor and recepient branches of the species tree (Fig.
+[ADD FIG 1]).
+For each edge type, we recorded the average number of edges of that type present in the sampled recombinant tree, and computed the ratio of this observed count and the prior expected count as implied by the species tree (considering branch lengths) and the three population parameters estimated in the preliminary round of the analysis.
+The base 2 logarithms of these ratios are depicted in the heatmap of Figure \ref{fig:Heatmap-of-recombination} (see Tables
+S\ref{tab:heatmap} and S\ref{tab:obsheatmap}).
+
+Among the possible recombinant edge types, we distinguished between ones which preserve the species tree topology, and ones which alter it.
+Our main focus was on the topology-altering types, both because our main objective was to study cross-clade gene transfer, and because our simulation study indicates that the observed counts of topology-preserving recombinant edges are somewhat less reliable {\bf[IG: POINT TO RELEVANT SECTION]}.
+The heatmap allows us to find recombination scenarios that deviate from the expectation dictated by the probabilistic model of ClonalOrigin.
+{\bf [IG: CHECK THIS FOLLOWING LINE OF ARGUMENT]} We identify two main potential causes for such a deviation: (1) reduced or increased opportunities for species to interact and exchange genetic material; (2) selection acting on newly transferred genes.
+
+We observe a reduced count of recombinant edges sampled between SDD and the three branches of the SPY clade ($\sim$25\% of the expected count).
+This is consistent with the fact that these species have reduced opportunities to interact, SPY being a human pathogen and SDD being strictly veterinary.
+On the contrary, the number of sampled recombinant edges between SDD and branches in the SDE clade seems to be consistent with the prior, with a slight excess for edges from SDE1 and SDE2 to SDD {\bf [IG: IS THIS WORTH DWELLNG ON?]}.
+The greatest deviation of observed counts from expected ones was observed for recombinant edges between the SPY and SDE clades.
+This is consistent with the fact that these two species share an ecological niche (SDE being an opportunistic human pathogen).
+Interestingly, there appears to be much higher enrichment (up to four times more than expected) for recombinant edges originating in the SPY clade and ending up in the SDE clade.
+In a follow-up simulation study we show that this deviation from the expected count is indeed highly significant, and so is the observed asymmetry in pattern of recombination {\bf[IG: POINT TO RELEVANT SECTION]}.
+This bias in directionality of recombination pattern between SPY and SDE is especially intriguing in light of the hypothesis that SDE acquired human pathogenicity via gene transfer from SPY.
+
+\subsection{Parsimony-based reconciliation of gene trees with the species tree}
+
+{\bf[IG: DID NOT TOUCH THE FIRST 3 PARAGRAPHS. REQUIRES ``CONDENSING".]}
+
+{\bf[IG: START WITH SOME GENERAL STATISTICS - NUMBER OF GENES, ETC?]}
 
@@
 
-\subsection{Virulence genes via recombining gene transfer}
+{\bf [IG: MODIFIED THIS PARAGRAPH]}
+
+As in the Bayesian analysis of ClonalOrigin, we attempted to see what types of transfer events (horizontal and recombining) were most common in our inference.
+We classified events into types defined by donor-recepient branches in the species tree.
+We recorded the number of inferred events of each type in Figures \ref{fig:hgt-heatmap} \&S\ref{fig:mowgli-recomb-heatmap} {\bf [IG: CONSIDER MOVING BOTH INTO SUPPLEMENT]}.
+Unlike in the Bayesian analysis, we cannot normalize these counts by their expectation, since we do not have any model-based expectation for them.
+This restrict our ability to interprest these counts.
+We did notice, however, a very high number of inferred recombining gene transfer events from the ancestral SPY branch to the ancestral SDE branch, consistent with our observations from the Bayesian analysis.
+With horizontal gene transfer, we do not see a similar pattern, but we do observe a high number of horizontal transfers from the ancesral SDE branch to SDD.
+{\bf [IG: IS THERE ANYTHING ELSE WE CAN SAY ABOUT THIS?
+MAYBE SHOW A SCATTER PLOT OF THE TWO TYPES OF EVENTS?
+]}
+
+\subsection{Gene categories associated with recombining and horizontal gene transfer}
+
+%\subsection{Association of virulence genes and horizontal gene transfer}
+
+We set out to seek cateories of genes that are significantly enriched for transfer events of either type (recombining and  horizontal).
+For horizontal transfer, we used a score based on the number of events inferred by the parsinomy-based method, and for recombining gene transfer, we used both the parsimony-based inferred counts and a recombination intensity score based on the Bayesian analysis (Materials and Methods).
+First, we checked association of transfers with virulence genes {\bf [IG: A SENTENCE ON WHERE WE TAKE THESE GENES FROM]}.
+This set of virulence genes was found to be significantly associated with horizontal gene transfers (P-value $<1.33 \times 10^{-11}$), but not with recombining gene transfers (P-value $>0.90$ for parsimony-based score and P-value $>??$ for recombnation intensity).
+We also found a marginally significant association between virulence genes and gene duplications (P-value $<7.13 \times 10^{-3}$).
+
+%\subsection{Transporters, transferases, aminoacyl tRNA synthetases, and transposases}
+
+We conducted similar tests for other functional categories of genes.
+Several functional categories that were found to be associated with recombination intensity were related with either membrane-bound molecules or aminoacyl tRNA synthetases (Table
+\ref{tab:functional}).
+Consistent results were obtained for the parsimony-based counts of recombining gene transfers (Table S\ref{tab:go-events-recombining}).
+Functional categories associated with horizontal gene transfer were clearly different from those associated with recombining gene transfer (Table \ref{tab:go-events}).
+The top ranked functional categories in that list were related with transposases.
+These differences add to our previous findings about the different roles these two types of transfer play in the evolution of these species.
+
+{\bf [IG: I THINK THE FOLLOWING 3 PARAGRAPHS SHOULD BE MOVED TO DISCUSSION]}
+
+ Molecular evolutionary forces could act on genetic materials of organisms, and changes to the genotypes might lead to some phenotypes that could change biological functions of the species.
+So, the bacterial gene transfer process as an evolutionary force could be associated with some biological functions.
+We
+attempted to find which biological functions were associated with recombining or
+horizontal gene transfers.
+
+Because bacteria as a unicellular organism should be able to respond to changes to environment, some evolutionary pressures must have been acting on genes encoding membrane-bound molecules.
+Recombining gene transfer process seemed to be more associated with biological functions that had been critical of sensing queues of bacterial environmental changes.
+Sugar phosphotransferase classes
+were ranked top in the list; phosphrylation could be important in signal
+transduction to recognize environmental changes.  The transport functional
+categories of carbohydrate transport and sodium ion transport also belonged to
+genes encoding membrane-bound molecules.  Also highly ranked were fatty acid and
+lipid biosyntheses that must be involved in producing membrane-bound molecules.
+Horizontal gene transfer process would allow bacteria to acquire foreign genes
+without the requirement of homologous regions in the bacteria.  Transposases
+would help foreign genes be excised from a donor genome and integrated into a
+recipient genome.
+
+In addition to these biological categories, we also found several aminoacyl tRNA synthetases (GO Term translation GO:0006412) in association with recombining gene transfer.
+In the parsimony-based approach the term translation was top
+ranked (Table S\ref{tab:go-events-recombining}).  They had been a focus of gene
+transfer in bacterial species \citep{Woese2000}.
+
+{\bf [IG: DO WE NEED/WANT THE FOLLOWING DISCUSSION? MAYBE IN MATERIALS AND METHODS - JUSTIFYING WHY WE MEASURE INTENSITY USING TOPOLOGY-CHANGING REC EDGES.]}
+
+ Some of the functional categories were unexpected; functional categories concerning ribosomes were found to be associated with recombining gene transfer.
+Because genes related with ribosomes were considered to be housekeeping genes, gene transfer in which
+would be rare, it was unexpected that ribosome related genes were found to be
+associated with recombining gene transfer.  Because it was a bizarre result, we
+investigated the association by considering different recombinant edge types. We
+noted that two different classes of recombinant edge types: one that changes
+species tree topologies, and another that does not. Because recombinant edges
+that did not change species tree topology might be difficult to infer reliably,
+considering those recombinant edges in computing recombination intensity might
+lead to dubious results.  Ribosome related functional categories were found to
+be top ranked if we took those recombinant edges that did not change species
+tree topology (Tables S\ref{tab:functional-notopology} and
+S\ref{tab:functional-all}).  In Table \ref{tab:functional} ribosome related
+functional categories were not top ranked.  Therefore, high recombination
+intensity in ribosome related genes were likely to be attributed to
+topology-not-changing recombinant edges. We would be cautious of recombining
+transfer of genes related with ribosome related functions.
+
+\subsection{Comparison of Bayesian and parsimony-based inferences of recombining gene transfer}
+
+The similarity between the two lists of gene categories found to be associated with recombining gene transfers in the separate analyses -- Bayesian and parsimony-based --  indirectly insdicate correlation between the two approaches.
+In order to obtain a more direct measure of correlation, we used the set of genes present in both analyses to compute the correlation between recombination intensity scores and the number of recombinng gene transfers inferred by the parsimony-based approach.
+We observed a positive correlation of $0.39$ (P-value $2.2\times10^{-16}$; Figure S\ref{fig:cmpcomowgli}) between the two measures.
+For comparison, the correlation between recombination intensity and the number of horizontal gene transfers was 0.03 (P-value 0.18;  \&S\ref{fig:cmpcomowglihgt}).
+{\bf [IG: MAYBE THERE IS A BETTER WAY TO DEMONSTRATE THE CORRELATION BETWEEN METHODS]}.
+
+{\bf [IG: MOVE TO DISCUSSION:]}
+A full model-based approach would be desirable in analyzing data for testing
+relevant hypotheses more systematically. Model complexity might not allow easy
+access to the implementation of a model. Our study involved a model-based
+approach and a parsimony-based one for understanding evolution in the core
+genomes of several bacteria individuals. Because of inability of the model-based
+method for studying the pan genomes we had to resort to the parsimony-based
+method. However, the parsimony-based method was also applied to the core
+genomes. We wished to compare the two methods in their results of the core
+genome analyses.  Because of two independent analyses on the same data set, this
+would allow us to be more confident of the results if the two analyses performed
+similarly Generally, phylogenetic methods of detecting gene transfers often take
+advantage of discrepancy between a reference species tree and gene trees. The
+two approaches that we employed in this study were no exception.
+
+\subsection{A closer study of virulence genes}
+
+{\bf [IG: CONSIDER WHAT WE WANT TO SAY HERE]}
+
+Roughly, half of the virulence genes were in the pan genome as mentioned above.
+We did not attempt to comprehensively list virulence genes that were inferred to
+be transferred horizontally or non-recombining fashion. We rather sort virulence
+genes by the number of horizontal gene transfer events to list the first several
+genes with high ranks.  Streptococcal C5a peptidase (\textit{scpB}, SDEG\_0933)
+was inferred to have six horizontal gene transfer.  Putative transposase were
+found to have a relatively large number of horizontal gene transfer events
+(SpyM3\_1016, SpyM3\_1053, SpyM3\_1054). Other genes with relatively large
+number of horizontal gene transfer events included putative response regulator
+(SpyM3\_0174), putative maltose operon transcriptional repressor (SpyM3\_0982),
+putative histidine kinase (SpyM3\_0171), and putative cell wall hydrolase
+(SpyM3\_0731, SpyM3\_0922, SpyM3\_1208).  Virulence genes encoding putative
+nucleases with a secretion signal peptide were also found to have had
+experienced horizontal gene transfer: genome-encoded streptodornase
+(SDEG\_0541), and prophage-associated deoxyribonuclease (SDEG\_1103).  Protein
+F2 like fibronectin-binding protein (SpyM3\_0104), an adhesin and invasin, was
+inferred to horizontally transferred.
+
+%\subsection{Virulence genes via recombining gene transfer}
 
@@
 
-\subsection{Association of virulence genes and horizontal gene transfer}
-
-Horizontal gene transfers were found to be significantly associated with
-virulence genes (P-value of $<1.33 \times 10^{-11}$) whereas recombining gene
-transfers were not enriched for virulence genes (P-value of $>0.90$) using the
-parsimony-based approach.  This was also consistent with the fact that we could
-not find the association of recombining gene transfer with virulence genes in
-the model-based approach.  This supported more evidence that non-recombining
-gene transfer served as a more effective means by which SDE or SPY acquired
-virulence genes, compared to recombining gene transfer.  Families with high
-numbers of duplications were also significantly enriched for virulence genes
-(P-value of $<7.13 \times 10^{-3}$) whereas families with high losses showed no
-enrichment (P-value of $>0.54$).
-
-Roughly, half of the virulence genes were in the pan genome as mentioned above.
-We did not attempt to comprehensively list virulence genes that were inferred to
-be transferred horizontally or non-recombining fashion. We rather sort virulence
-genes by the number of horizontal gene transfer events to list the first several
-genes with high ranks.  Streptococcal C5a peptidase (\textit{scpB}, SDEG\_0933)
-was inferred to have six horizontal gene transfer.  Putative transposase were
-found to have a relatively large number of horizontal gene transfer events
-(SpyM3\_1016, SpyM3\_1053, SpyM3\_1054). Other genes with relatively large
-number of horizontal gene transfer events included putative response regulator
-(SpyM3\_0174), putative maltose operon transcriptional repressor (SpyM3\_0982),
-putative histidine kinase (SpyM3\_0171), and putative cell wall hydrolase
-(SpyM3\_0731, SpyM3\_0922, SpyM3\_1208).  Virulence genes encoding putative
-nucleases with a secretion signal peptide were also found to have had
-experienced horizontal gene transfer: genome-encoded streptodornase
-(SDEG\_0541), and prophage-associated deoxyribonuclease (SDEG\_1103).  Protein
-F2 like fibronectin-binding protein (SpyM3\_0104), an adhesin and invasin, was
-inferred to horizontally transferred.
-
-\subsection{Model-based approach or parsimony-based one for recombining gene
-transfer study}
-
-A full model-based approach would be desirable in analyzing data for testing
-relevant hypotheses more systematically. Model complexity might not allow easy
-access to the implementation of a model. Our study involved a model-based
-approach and a parsimony-based one for understanding evolution in the core
-genomes of several bacteria individuals. Because of inability of the model-based
-method for studying the pan genomes we had to resort to the parsimony-based
-method. However, the parsimony-based method was also applied to the core
-genomes. We wished to compare the two methods in their results of the core
-genome analyses.  Because of two independent analyses on the same data set, this
-would allow us to be more confident of the results if the two analyses performed
-similarly Generally, phylogenetic methods of detecting gene transfers often take
-advantage of discrepancy between a reference species tree and gene trees. The
-two approaches that we employed in this study were no exception. We found a few
-similar results in both methods; firstly, functional categories that were
-associated with recombining gene transfer were similar (Tables
-S\ref{tab:functional-all} and S\ref{tab:go-events-recombining}).  Secondly,
-either of them did not show association of virulence genes with recombining gene
-transfer.  To more directly compare the methods we used the recombination
-intensity and the number of recombining gene transfer events for each gene.  We
-found a weak positive correlation of 0.39 with a very small significant P-value
-of $2.2\times10^{-16}$ for testing a null hypothesis of the correlation being
-zero.  (Figure S\ref{fig:cmpcomowgli}).  Interestingly, the correlation (0.03
-with P-value of 0.18) disappeared in a comparison between recombination
-intensity and number of horizontal gene transfer events (Figure
-S\ref{fig:cmpcomowglihgt}).
-
 \subsection{UCSC genome browser tracks for bacterial gene transfer study}
@@
 loci from SpyM3\_1518 to SpyM3\_1525.
-
-\subsection{Transporters, transferases, aminoacyl tRNA synthetases, and
-transposases}
-
-Molecular evolutionary forces could act on genetic materials of organisms, and
-changes to the genotypes might lead to some phenotypes that could change
-biological functions of the species. So, the bacterial gene transfer process as
-an evolutionary force could be associated with some biological functions. We
-attempted to find which biological functions were associated with recombining or
-horizontal gene transfers.  Several functional categories that were found to be
-associated with recombining gene transfer were related with either
-membrane-bound molecules or aminoacyl tRNA synthetases (Table
-\ref{tab:functional}).  However, functional categories associated with
-horizontal gene transfer were different from those from recombining gene
-transfer analysis (Table \ref{tab:go-events}).  The top ranked functional
-category associated with horizontal gene transfer was related with transposases.
-These differences in the profile of functional categories between the two
-processes were interesting to us. An approach of dicriminating the two gene
-transfer processes would result in the differences.
-
-Because bacteria as a unicellular organism should be able to respond to changes
-to environment, some evolutionary pressures must have been acting on genes
-encoding membrane-bound molecules. Recombining gene transfer process seemed to
-be more associated with biological functions that had been critical of sensing
-queues of bacterial environmental changes.  Sugar phosphotransferase classes
-were ranked top in the list; phosphrylation could be important in signal
-transduction to recognize environmental changes.  The transport functional
-categories of carbohydrate transport and sodium ion transport also belonged to
-genes encoding membrane-bound molecules.  Also highly ranked were fatty acid and
-lipid biosyntheses that must be involved in producing membrane-bound molecules.
-Horizontal gene transfer process would allow bacteria to acquire foreign genes
-without the requirement of homologous regions in the bacteria.  Transposases
-would help foreign genes be excised from a donor genome and integrated into a
-recipient genome.
-
-In addition to these biological categories, we also found several aminoacyl tRNA
-synthetases (GO Term translation GO:0006412) in association with recombining
-gene transfer.  In the parsimony-based approach the term translation was top
-ranked (Table S\ref{tab:go-events-recombining}).  They had been a focus of gene
-transfer in bacterial species \citep{Woese2000}.  Some of the functional
-categories were unexpected; functional categories concerning ribosomes were
-found to be associated with recombining gene transfer.  Because genes related
-with ribosomes were considered to be housekeeping genes, gene transfer in which
-would be rare, it was unexpected that ribosome related genes were found to be
-associated with recombining gene transfer.  Because it was a bizarre result, we
-investigated the association by considering different recombinant edge types. We
-noted that two different classes of recombinant edge types: one that changes
-species tree topologies, and another that does not. Because recombinant edges
-that did not change species tree topology might be difficult to infer reliably,
-considering those recombinant edges in computing recombination intensity might
-lead to dubious results.  Ribosome related functional categories were found to
-be top ranked if we took those recombinant edges that did not change species
-tree topology (Tables S\ref{tab:functional-notopology} and
-S\ref{tab:functional-all}).  In Table \ref{tab:functional} ribosome related
-functional categories were not top ranked.  Therefore, high recombination
-intensity in ribosome related genes were likely to be attributed to
-topology-not-changing recombinant edges. We would be cautious of recombining
-transfer of genes related with ribosome related functions.
 
@@
 
+\section{Supplementary Information}
+
+\subsection{Summarization of posterior samples for local gene trees and recombination intensity}
+
+Because the procedure of summarizing posterior samples of \textit{recombinant
+tree}, which will be defined in a moment, was not detailed in the user guide at
+the time when we used \texttt{ClonalOrigin}, we elected to use Figure
+S\ref{fig:clonalorigin} to illustrate a recombinant tree in order to clarify
+summarization of \textit{recombinant history} posterior samples.  A
+recombination history is represented by a \textit{recombinant tree}, which is a
+species tree augmented with \textit{recombinant edges}.  A recombinant edge
+connect two species tree branches.  Each recombinant edge is defined by the time
+and branch of origin and destination, and the genomic interval that it affects.
+The branch of origin is referred to as the \textit{donor} and the destination
+branch is referred to as the \textit{recipient}. The type of a recombinant edge
+is determined by the donor and recipient species tree branches.  A recombinant
+tree can be viewed as a restricted version of an ancestral recombination graph
+(ARG), and as such it induces a local genealogy for each site in the genomic
+block it describes.  Didelot et al.\ \citet{Didelot2010} presented two procedures of summarizing
+recombinant trees by viewing trees from two different angles: by counting types
+of recombinant edges regardless of genomic positions affected by the edges, and
+by measuring recombinant edge changes to the species tree in site-wise manner.
+By similarly following their approaches we quantified additional aspects of
+recombinant trees; we counted gene tree topologies for each site, and also
+quantified recombination intensity along the genome.
+
+The recombination intensity allowed us to find genomic regions enriched for
+recombination events in general, or regions enriched for recombination events of
+certain types (e.g., between SPY and SDE).  We noted that the recombination
+intensity was not normalized to the interval of $[0,1]$. Rather than giving
+probabilistic meanings to the recombination intensity we interpreted it as an
+average number of recombinant edge types per site.  The measure appeared to act
+as an effective indicator of recombination intensity (see subsection of
+\textbf{Simulation Study}).  Our simulation study indicated that edges of
+certain types were better indicators of recombination events than others.  For
+instance, recombinant edges which resulted in a recombinant tree with the same
+topology as the species tree appeared to be more prone to spurious sampling.
+Such edges were ones where the donor was the same branch, sister branch, or
+parent branch of the recipient branch.  Therefore, in some cases we elected to
+remove such edges from consideration when measuring recombination intensity.  
+% The tree topologies would show more effectively regions where
+% recombination occurs along the genome (see subsection of \textbf{UCSC Genome
+% Browser}).
+
+\subsection{Simulation 1 - three population parameters}
+
+We describe a procedure of simulations for checking if the three
+population parameters were accurately recovered using the model-based approach.
+Ten recombinant trees were sampled given the three parameters (Table
+S\ref{tab:three}) and the species tree (Figure S\ref{fig:tree5}) set to the
+values inferred from the five bacterial genomes.  Each of the recombinant trees
+was used to simulate the Jukes-Cantor model for DNA sequence evolution to create
+274 blocks of alignments.  The block lengths configuration was the same as that
+of the blocks used in the analysis of the five genomes.
+
+% s15
+% FIXME: Newick tree in text in a file.
+% FIXME: Publish the result and the data set with the code.
+
+\subsection{Simulation 2 - numbers of recombinant edges from prior}
+
+We first sampled ten recombinant trees given the mutation rate, recombination
+rate, and tract length (Table S\ref{tab:three}) and the species tree (Figure
+S\ref{fig:tree5}), and then created ten independent simulated data sets with
+the Jukes-Cantor model for DNA sequence evolution using each of the recombinant
+trees. This consisted of 100 replicates of data sets.  The recombinant trees
+were created from the prior distribution and recombinant edges were added to the
+species tree with no particular directions of recombining gene transfers between
+species tree branches.  We computed the fraction of the mean number of recombinant
+edges for a pair of donor and recipient species tree branches over the simulated
+10 recombinant trees against the median of recombinant edges estimated from the
+10 simulated data sets, and then took the logarithm of the fraction to base 2. We
+also computed 5\% and 95\% quantiles to find the 90\% quantile range for the log
+base 2 fraction.  We expect that the log base 2 fraction would near zero. We also
+wished to see if how the recombinant edge count estimates from the real data set
+deviate from the prior. Instead of using the number of recombinant edges from
+the ten recombinant trees, we used the numbers of recombinant edges that were
+estimated from the real data.
+
+We tested if our measure of recombination intensity provided a reliable estimate
+of the true intensity. We artificially annotated our synthetic blocks to genes
+according to the real annotations from SPY1 genome.  For each gene in each of
+the simulated 100 data sets, we plotted the average estimated recombination
+intensity along that gene against the average number of different recombinant
+edge types from the ten simulated recombinant tree.
+
+% s16
+
+% s16
+% FIXME 1: IG: RESULTS FROM THIS POINT ON REQUIRE SOME REFINEMENT. SC: Let's see.
+% FIXME 2: ``TRUE" TO BE EXPECTED VALUE
+% RATHER THAN AVERAGE OVER THE 10 REC.  TREES. SC: OKAY. LET'S FIND THE PRIOR
+% EXPECTED NUMBER OF EVENTS. THIS MUST BE GIVEN SOMEWHERE}.  
+% FIXME 3: CONSIDER SHOWING DETAILED COMPARISON OF ESTIMATED VALUES TO ONES 
+% EXPECTED FROM THE ACTUAL REC. TREES. THESE DON'T LOOK AS NICE, BUT THEY MIGHT 
+% BE ABLE TO DEMONSTRATE OUR ABILITY TO DETECT DEPARTURE FROM EXPECTATION. SC:
+% THIS MAY BE ABOUT CHANGING AVERAGE TO PRIOR EXPECTED NUMBER OF EVENTS.
+% FIXME 4: WE SHOULD PROBABLY CONSIDER ALTERNATIVE WAYS OF USING THESE RESULTS TO 
+% ARGUE THAT OUR DIRECTIONAL SIGNAL IS REAL. SC: I MIGHT CONSIDER SIMULATIONS AND
+% ARTIFICIALLY DOUBLE OR TRIPLE THE RECOMBINATION EVENTS OF ONE OF TYPES.
+
+\subsection{Simulation 3 - number of recombinant edges from posterior}
+
+We used 100 recombinant trees sampled from the posterior distribution
+of the recombiant tree instead of using simulated recombinant trees.
+We created one simulated data set for each of the 100 recombinant trees with
+the Jukes-Cantor model for DNA sequence evolution 
+using the mutation rate estimate (Table S\ref{tab:three}).
+This consisted of 100 replicates of data sets.  
+We follow the same computation of the fraction as in the simulation 2 with 
+the numbers of recombinant edges estimated from the real data
+as numerators.
+
+% s17
+
 \end{document}
 
-\begin{figure}
-\includegraphics[scale=0.7]{figures/plot-number-recombination-within-blocks-1\lyxdot out\lyxdot recomb}
-\caption{\label{fig:recombination-boundaries}Distribution of number of 
-recombination
-events boundaries. The number of recombination event boundaries per
-site for each block along the \textit{S.\ dysgalactiae} ssp.\textit{\ equisimilis}
-ATCC 12394.}
-\end{figure}
-\clearpage{}%
-
-The mechanisms of horizontal gene transfer
-include conjugation by plasmid, transduction by bacteriophages, or
-transformation of naked DNA uptake \citep{Thomas2005}. Additionally,
-marine bacteria was reported to acquire foreign genes via gene transfer
-agents, i.e., viral-like particles produced by $\alpha$-Proteobacteria
-\citep{McDaniel2010}.
-
-Although
-recombinant edges between two species branches do not explicitly denote
-gene flow between the two species in their original model, \citet{Didelot2010}
-exploited their method to infer flows of genetic segment between \textit{Bacillus
-cereus} species. This model-based approach to detecting homologous
-recombining gene transfer was used to the core genomes of closely
-related \textit{Streptococcus} species.
-
-{[}WHERE SHOULD I PUT THESE UNIT OF GENE TRANSFER?{]}
-
-\textbf{Units of gene transfer:} \citet{Chan2009a} showed that protein
-domains were not gene transfer units. In another paper \citet{Chan2009}
-studied gene with and without breakpoints, and found that more often
-entire genes transfer in pathogens than in non-pathogens. In pathogenic
-bacteria they found more genes with breakpoints than genes without
-breakpoints.
-
-% Table and SDE and SPY
-Consistent
-with the number of recombinant edges were that gene transfer fragments
-with high posterior probability shown in Table S\ref{tab:genes-transfer}
-often appeared to be from SPY to SDE. While four genes were inferred
-to be transferred to SPY1, there were more genes transferred to SDE1
-or SDE2.
-
-\textbf{Evolutionary hypotheses about the ancestor of SPY and SDE:}
-\citet{Kalia2001} proposed two scenarios of the evolution of the
-two species: The first one was that the most recent common ancestor
-of SPY and SDE was a GAS-like pathogen, SDE was losing virulence factors
-since the divergence of the ancestor, and SDE regained virulence factors
-from SPY because of the same niche. The second scenario was that the
-most recent common ancestor of SPY and SDE was a GGS/GCS-like commensal,
-SPY evolved virulence factors since the divergence of the ancestor,
-and SDE gained virulence factors from SPY. Commensal bacteria could
-evolve by aquiring virulence genes from its closely related pathogenic
-bacteria. \citet{Kalia2001} suggested the divergence time of SDD
-to be around the time when human started to domesticate wildlife animals
-around 10,000 years ago. A demographic model could be made with two
-species to study the divergence time quantitatively. This would require
-ancestral recombination graphs embedded in a species tree.
-
-% From discussion
-This distinction of the two processes of
-gene transfer was envisioned a decade ago \citep[e.g.,][]{Ochman2001},
-and the methods of gene transfer detection by distinguishing the processes
-were not yet employed even if some of the methods were already available.
-We by no means provided any solutions to the problem of detecting
-gene transfer here. Yet, we hoped that we should be able to discriminate
-the two processes because each of the processes of gene transfer could
-affect biological functions that were different in nature.
-
-% From discussion
-For example, genes transferred via non-recombining
-gene transfer could be parts of functional units such
-as protein domains, and secondary structures of ribosomal RNAs. Genes
-transferred via homologous recombining gene transfer could be small
-fragments of genes as long as the flanking sequences of the fragments
-were more or less homologous.
-
-\textbf{Overlap of genes under positive selection and higher recombination:}
-As the uncertainty in gene tree topologies becomes larger, false positive
-rates of detecting sites under positive selection would be high (CITE).
-Genes could be divided into parts where we hoped no multiple gene
-tree topologies exist. Dividing genes into those fragments might be
-a way to avoid potentially high false positive rates of positively
-selected sites. Considering the fact that nucleotide substitutions
-seemed to be affected more by recombinations than by mutations we
-suspect that fragmenting genes might not be helpful.
-
-% Virulence gene and non-recombining gene transfer
-We investigated bacterial gene
-transfer between SPY and SDE using the genome sequences by comparing
-gene tree and species tree. We considered both core and pan genomes
-for detecting homologous recombining gene transfer and non-recombining
-gene transfer. 
-%
-The extent to which genes were transferred would vary on this
-mode of gene transfer. 
-While homologous
-recombination gene transfer would happen in the genes whose donor
-and recipient species were rather closely related, non-homologous
-gene transfer could happen virtually in any genes of donor
-and recipient species as long as the species shared their habitats.
-%
-We did not find an excess of virulence
-genes associated homologous recombining gene transfer. Actually, many important
-virulence genes were missing in the core genomes. We found association
-of virulence genes with non-recombining gene transfer.
-
-% Number of recombinant edges (Methods)
-This would indicate the degree to which the clonal
-species tree was disrupted by bacterial recombination. 
-%
-We used the method to infer which types of bacterial recombination
-were more prevalent than which other types in the lineages of the
-tree relating five streptococcus genomes.
-
-% Watterson's estimate
-in that the rate of mutation
-on the branches of the genealogy and branch lengths of the species
-tree $\mathcal{T}$ are reciprocally proportional.
-
-% simulation for ClonalOrigin stage 1
-We hoped that we could understand the power
-and limitations of \texttt{ClonalOrigin} under the parameter setup that we
-inferred from our data set.
-
-% Heat map for recombination from ClonalOrigin
-Because the
-lengths of SPY1 and SPY2 were larger than those of SDE1 and SDE2,
-recombinant edges could more often depart from one of SPY1 and SPY2
-lineages and arrive at one of SDE1 and SDE2 lineages than those for
-the reverse direction \textit{a priori} (Figure S\ref{fig:tree5}). 
-We took into account of the prior expected number of recombinant edges.
-
-\textbf{Gene transfer and functional categories:} We showed much clearer
-examples of lateral transfer of gene fragments.
-Modular structures of aaRS genes made them
-easy targets of gene transfer \citep{Woese2000}. The two transporter
-genes might have been under selective pressure from environment.
-
-% rRNA binding
-Of the genes in Table S\ref{tab:genes-transfer} we did not find ribosomal
-proteins that were heavily enriched in the gene ontology term, {}``rRNA
-binding.'' 
-Ribosomal proteins tended to cluster together in the genome,
-and they appeared to be affected by the recombinant edges whose departures
-were on the root branch. 
-Recombinant edges that connect the root branch
-are hard to be considered as gene flow. We ruled out those recombinant
-edges in Table S\ref{tab:genes-transfer} as we mentioned.
-
-% From Introduction
-It has been suspected that opportunistic pathogenicity
-of SDE may have been facilitated by the very genetic interaction between
-the two species. 
-Although virulence genes were suspected to be acquired via non-recombining gene
-transfer rather than recombining transfer \citep{Groisman1994}, systematic
-approaches to elucidating relative effects of the two types of gene transfer 
-rarely have been applied.
-
-This unfortunate situation motivated us to study the
-closely related bacteria as efforts of better understanding relative 
-association of gene transfer with virulence genes.
-We hoped that understanding their genetic exchanges
-would enhance our knowledge of mechanisms of their pathogenicity on
-humans.
-
-While compositional approaches were effective in analyzing
-genome-scale data, phylognetic approaches have allowed to understand
-evolutionary forces of gene exchange between species
-\citep[e.g.,][]{Bapteste2004,Simonson2005,Beiko2009} by comparing differences
-between gene trees and a reference species tree.  However, many of the methods
-did not have explicit models of homologous recombining gene transfer that was
-one of important evolutionary processes in horizontal gene transfer.  
-% Overall,
-% the development of methods for downstream analyses using bacterial genomes
-% appears to lag behind that of high-throughput sequencing technology
-% \citep[e.g.,][]{Green2011}.
-
-% Heat map
-% Another useful
-% summary suggested by \citet{Didelot2010} was to classify recombinant edges by
-% types defined according to donor/recipient branch pairs, and record the number
-% of edges sampled for each recombinant edge type. For instance, our species tree
-% has nine branches including rooting branch and 81 branch pairs, out of which 60
-% are possible donor/recipient pairs by prohibiting recombinant edges from going
-% backward in time. The posterior expected number of edges of each type can be
-% compared to the prior expected number determined by the three population
-% parameters fixed in the second stage of \texttt{ClonalOrigin}. This comparison
-% was displayed in a heat map of Figure \ref{fig:Heatmap-of-recombination}, which
-% indicated branch pairs that might experience an excess or depletion of
-% recombination events across all blocks.
-
-% \textbf{Functional categories associated with homologous recombination gene transfer:} 
-% Having established relative correlation of the surrogate measure of recombination
-% intensity in the simulation (Figure S\ref{fig:ri1}) we attempted to find
-% functional categories that were associated with recombination.
-% The nucleotide substitutions
-% due to recombination events might not be selectively neutral because
-% of large population of the bacteria, and accordingly natural selection
-% could be more stronger than random genetic drift. Reflecting on the
-% report of \citet{Lefebure2007} that genes that were likely to be
-% under positive selection tended to be under higher recombination,
-% we would expect that functional categories would be associated with higher
-% recombination, and consequently homologous recombining gene transfer
-% between species.
-
-% block 76: SDEG\_0641
-% ORTHOMCL43-1    SEQ\_1497        SpyM3\_0440      MGAS10750\_Spy0535 SDD27957\_03210  SDE12394\_03445  SDEG\_0641
-% Not found in Davies2007
-% putative calcium transporter
-% MgtA
-
-% block 97: SDEG\_0673
-% ORTHOMCL1244-1  SEQ\_0860        SDD27957\_03385  MGAS10750\_Spy0623 SpyM3\_0465      SDE12394\_03650  SDEG\_0673
-% Putative (put.) extracellular proteins: put. dipeptidase (SpyM3\_0465)
-% Dipeptiase for the same human host might be important for SDE to survive human
-% host environment. Scarvenge dipeptides.
-
-% block 101: SDEG\_0739
-% ORTHOMCL1207-1  SDD27957\_04880  SEQ\_0932        SDD27957\_03660  SDEG\_0739 SDE12394\_03975  MGAS10750\_Spy0677       SpyM3\_0511
-
-% block 151: SDEG\_1447
-% ORTHOMCL99-1	SEQ\_0680	SDD27957\_06980	SDEG\_1447	SDE12394\_07540	SpyM3\_1071	MGAS10750\_Spy1252
-% Not found in Davies2007
-% superoxide dismutase
-% sodM
-% Poyart2001
-% Contribution of Mn-Cofactored Superoxide Dismutase (SodA) to the Virulence of Streptococcus agalactiae.
-
-% block 244: SDEG\_2040
-% ORTHOMCL37-1	SEQ\_0209	SpyM3\_1765	MGAS10750\_Spy1855	SDD27957\_10625	SDE12394\_10325	SDEG\_2040
-% Not found in Davies2007
-% chaperonin GroEL, 
-% putative heat shock protein
-% What is the role of this GroEL?
-% Increase of GroEL in acid shock, which might be important for survival of the
-% bacteria in human hosts.
-
-% block 242: SDD27957\_00955
-% ORTHOMCL730-1	SDE12394\_10000	SEQ\_2112	SDD27957\_00955	SDE12394\_00680	SDEG\_0169	SpyM3\_1755	MGAS10750\_Spy1843
-% Not found in Davies2007
-% putative pyruvate formate-lyase activating enzyme
-% There are some upstream parts were replaced.
-
-% block 240: SDD27957\_00985
-% ORTHOMCL731-1	SDE12394\_10005	SEQ\_2106	SDD27957\_00985	SpyM3\_1749	MGAS10750\_Spy1837	SDE12394\_00710	SDEG\_0175
-% Not found in Davies2007
-% putative pyruvate formate-lyase
-% pflD
-% Yesilkaya2009
-% Pyruvate Formate Lyase Is Required for Pneumococcal Fermentative Metabolism and Virulence
-% SDE2 received from SPY1 via recombining gene transfer.
-
-% block 167: SDD27957\_06790
-% ORTHOMCL1139-1	SEQ\_0705	SpyM3\_1035	MGAS10750\_Spy1208	SDD27957\_06790	SDE12394\_07350	SDEG\_1372
-% Not found in Davies2007
-% putative internalin A precursor
-% Adhesins
-
-% block 255: MGAS10750\_Spy1922
-% ORTHOMCL158-1	SEQ\_2161	SDD27957\_10925	SDE12394\_10560	SDEG\_2091	SpyM3\_1806	MGAS10750\_Spy1922
-% Not found in Davies2007
-% DNA mismatch repair protein MutS
-% What makes MutS as a virulence genes?
-
-% block 67: SDD27957\_03085
-% ORTHOMCL121-1	SEQ\_1525	SDD27957\_03085	SpyM3\_0420	SDE12394\_03265	SDEG\_0610	MGAS10750\_Spy0514
-% Not found in Davies2007
-% lysyl-tRNA synthetase
-
-% block 133: SDD27957\_04505
-% ORTHOMCL58-1	SEQ\_1129	SDD27957\_04505	SDE12394\_04895	SDEG\_0866	SpyM3\_0764	MGAS10750\_Spy0977
-% Not found in Davies2007
-% putative spermidine/putrescine ABC transporter ATP-binding protein
-
-% List of virulence genes.
-% Genes for the Majority Of Group A Streptococcal Virulence Factors and
-% Extracellular Surface Proteins Do Not Confer an Increased Propensity to Cause
-% Invasive Disease
-
-% block 170: SDD27957\_07725
-% ORTHOMCL96-1	SEQ\_0613	SDD27957\_07725	SDEG\_1515	SDE12394\_07910	SpyM3\_1180	MGAS10750\_Spy1365
-% Not found in Davies2007
-% putative glucose kinase
-
-% Evidence for Reductive Genome Evolution and Lateral Acquisition of Virulence
-% Functions in Two Corynebacterium pseudotuberculosis Strains
-
-% block 186: MGAS10750\_Spy1444
-% ORTHOMCL93-1	SEQ\_1774	SDD27957\_08550	SDE12394\_08460	SDEG\_1691	SpyM3\_1369	MGAS10750\_Spy1444
-% Not found in Davies2007
-% putative serine/threonine kinase
-
-% Blocks with virulence genes.
-% Check virulence genes (they were combined) at 
-% SDD27957\_02890, SDD27957\_02895
-% MGAS10750\_Spy0455, MGAS10750\_Spy0456 (block 75)
-
-% HGT and virulene genes in detail
-% ORTHOMCL4-1	0	1	6	0	TRUE
-% ORTHOMCL4-1	SDD27957_02925	SEQ_0563	MGAS10750_Spy0339	SpyM3_0298	MGAS10750_Spy0340	SDEG_0469	SDD27957_02245	SDE12394_02260	SDD27957_05495	SDEG_0933	SpyM3_1726	MGAS10750_Spy1807
-% streptococcal C5a peptidase (scpB) (SDEG_0933) 
-% 
-% ORTHOMCL1826-1	0	2	6	0	TRUE
-% ORTHOMCL1826-1	SDEG_1261	SDE12394_05305	SDE12394_03405	SDEG_0634	SDD27957_00125	SDD27957_02625	SDD27957_04790	SDD27957_05475	SDE12394_02555
-% Nothing in Shimomura.
-% 
-% ORTHOMCL912-1	3	2	5	0	TRUE
-% ORTHOMCL912-1	SDD27957_04815	SDD27957_05095	SDD27957_02300	SDD27957_11185	SDD27957_05765	SDD27957_06935	SpyM3_1053	MGAS10750_Spy1189	MGAS10750_Spy1234	SpyM3_1016	SDD27957_02260	MGAS10750_Spy0401	SpyM3_1054	MGAS10750_Spy1235
-% Nothing in Shimomura.
-% SpyM3_1053 /product="putative transposase"
-% SpyM3_1054 /product="putative transposase"
-% SpyM3_1016 /product="IS1076 transposase"
-% 
-% # ORTHOMCL4-2	1	1	4	0	TRUE
-% 
-% # ORTHOMCL29-2	0	0	4	0	TRUE
-% ORTHOMCL29-1	1	1	3	1	TRUE
-% ORTHOMCL29-1	SEQ_1712	SEQ_0303	SpyM3_0174	MGAS10750_Spy0201	SDD27957_10325	SDE12394_09770	SDEG_1964	SEQ_1095	SDEG_0515	MGAS10750_Spy0403
-% Nothing in Shimomura.
-% SpyM3_0174 /product="putative response regulator"
-% 
-% # ORTHOMCL26-2	0	0	3	0	TRUE
-% ORTHOMCL26-1	0	0	3	0	TRUE
-% ORTHOMCL26-1	SpyM3_0982	MGAS10750_Spy1149	SEQ_1437	SDD27957_06480	SDEG_1294	SDE12394_06795	SEQ_2022	SDE12394_00945	SDEG_0221
-% Nothing in Shimomura.
-% SpyM3_0982 /product="putative maltose operon transcriptional repressor"
-% 
-% # ORTHOMCL1564-2	0	2	3	0	TRUE
-% 
-% #ORTHOMCL126-2	0	1	3	0	TRUE
-% ORTHOMCL126-1	0	1	3	0	TRUE
-% ORTHOMCL126-1	SEQ_0301	SDD27957_10335	SDEG_1966	SpyM3_0171	MGAS10750_Spy0199	SEQ_1711	MGAS10750_Spy0404	SDE12394_02570	SDEG_0516
-% Nothing in Shimomura.
-% SpyM3_0171 /product="putative histidine kinase"
-% 
-% ORTHOMCL1071-2	1	2	3	2	TRUE
-% % ORTHOMCL1071-2	SEQ_0845	SEQ_0190	SEQ_1729	MGAS10750_Spy0620	SpyM3_0731	SpyM3_0922	SpyM3_1208	SDEG_1106	SDD27957_07165
-% Nothing in Shimomura.
-% SpyM3_0731 /product="putative holin - phage associated"
-% SpyM3_0922 /product="putative cell wall hydrolase - phage associated"
-% SpyM3_1208 /product="putative cell wall hydrolase, lysin - phage associated"
-% 
-% ORTHOMCL1037-1	0	2	3	0	TRUE
-% ORTHOMCL1037-1	MGAS10750_Spy1697	SDE12394_04580	MGAS10750_Spy0948	SDD27957_05395	SDD27957_04295	SDE12394_07035
-% Nothing in Shimomura.
-% 
-% ORTHOMCL2721	0	4	2	0	TRUE
-% ORTHOMCL2721	SDE12394_10035	SDE12394_09945	SDE12394_09940
-% Nothing in Shimomura.
-% 
-% # ORTHOMCL1826-2	4	0	2	0	TRUE
-% # ORTHOMCL1826-2	SDEG_1261	SDE12394_05305	SDE12394_03405	SDEG_0634	SDD27957_00125	SDD27957_02625	SDD27957_04790	SDD27957_05475	SDE12394_02555
-% 
-% ORTHOMCL1564-1	0	2	2	0	TRUE
-% ORTHOMCL1564-1	SDEG_1103	SpyM3_1409	SDD27957_09010	SEQ_0696	SDD27957_02745	SDEG_0541
-% GGS_124 possesses 5 genes that encode putative nucleases with a secretion signal peptide: genome-encoded streptodornase (SDEG_0541), others,  prophage-associated deoxyribonuclease (SDEG_1103), all of which are predicted to code for a secretion signal peptide.
-% SpyM3_1409 /product="streptodornase (Sdn) - phage associated"
-% 
-% 
-% ORTHOMCL1439-1	0	1	2	0	TRUE
-% ORTHOMCL1439-1	SEQ_1606	SEQ_1999	SpyM3_0104	SDE12394_09890	SDD27957_10420	SDE12394_09885	SDEG_1984
-% Nothing in Shimomura.
-% SpyM3_0104 /product="protein F2 like fibronectin-binding protein"
-% 
-% ORTHOMCL1369-2	2	1	2	0	TRUE
-% ORTHOMCL1369-2	MGAS10750_Spy1683	SDEG_1338	SDEG_1346	SDE12394_07225	SDE12394_04610	SDD27957_04325	SDD27957_05365	SDD27957_08360
-% Nothing in Shimomura.
-% 
-% ORTHOMCL1283-1	3	0	2	0	TRUE
-% ORTHOMCL1283-1	SDD27957_00840	SDEG_1543	SDEG_1531	SDE12394_03300	SDEG_0983	SDE12394_05495	SDE12394_08020	SDE12394_07960
-% Nothing in Shimomura.
-% 
-% ORTHOMCL1232-2	0	1	2	0	TRUE
-% ORTHOMCL1232-2	SDE12394_10675	SDE12394_10850	SDEG_2142	MGAS10750_Spy1897	SDD27957_00535	SDD27957_02435
-% Nothing in Shimomura.
-% 
-% ORTHOMCL1229-2	0	2	2	0	TRUE
-% ORTHOMCL1229-2	SDD27957_03510	SDD27957_05295	SDD27957_08295	SEQ_0753	MGAS10750_Spy1709
-% Nothing in Shimomura.
-% 
-% --------------------------------
-% 
-% 
-% ORTHOMCL9-2	1	1	1	0	TRUE
-% ORTHOMCL9-2	SpyM3_0160	MGAS10750_Spy0188	SEQ_0271a	SEQ_0289a	SDD27957_10400	SDE12394_09855	SDEG_1980	SpyM3_1853
-% 
-% ORTHOMCL912-2	5	1	1	0	TRUE
-% ORTHOMCL9-1	1	0	1	0	TRUE
-% ORTHOMCL848-2	0	1	1	0	TRUE
-%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Functional category association
-Lateral transfer of transporter related genes are well known.  Functional
-categories associated with higher recombining gene transfer were found to
-include phosphotransferase systems and ATP-binding cassette (ABC) transporters,
-i.e., carbohydrate transport (GO:0008643) (Table \ref{tab:functional}).
-
-We also
-found several aminoacyl tRNA synthetases in gene ontology of translation
-(GO:0006412) to be significantly associated with higher recombining gene
-transfer.
-
-Functional categories concerning ribosomes were also
-associated with higher recombining gene transfer. These categories were not
-found significantly associated with higher recombining gene transfer when we
-considered only gene transfer between SPY and SDE genomes, which implied that
-many recombining gene transfer events should be attributed to those of internal
-species branches as donors.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% From the discussion about virulence genes and two gene transfer
-We suspected that virulence genes were exchanged between SPY and
-SDE, but did not find any directionality of the gene flow.
-This does not necessarily mean that
-virulence genes were not exchanged between SPY and SDE via non-recombining
-transfer.  Because the \texttt{Mowgli} analysis just did not reveal unbalanced
-non-recombining or recombinding gene transfer, we suspected that \texttt{Mowgli}'s
-parsimony might not be powerful in detecting unbalanced gene flow.  We
-conjecture the following hypothesis of the evolution of SPY and SDE to support
-the non-recombining virulence gene exchange between SPY and SDE.  Pathogenicity
-of SDE might have existed before the common ancestor of SPY and SDE.  We
-speculated that SPY might have been split off from the common ancestor that was
-SDE-like because the population size of SDE was larger than that of SPY (CITE).
-After the split, SPY might have been specialized in inhabting human hosts, and
-SDE in colonizing veterinary hosts. Recent surge in cases of human diseases
-attributed to SDE might be because of secondary contact of SPY and SDE since
-their split from their common ancestor.  When SPY thrives in its environment of
-human hosts and another relatively close species, SDE, starts to cohabit with
-the indigenous species SPY, genetic elements in one of the two species could
-benefit another species. If the two species do not compete with each other, then
-the increased genetic variation from multiple species would eventually bring
-advantage to both of the species.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Functional category association 
-\textbf{Two functional categories and recombining gene transfer:}
-Among the genes shown in Table S\ref{tab:functional} two classes of proteins
-attracted our attentions: aminoacyl tRNA synthetases (aaRS) and
-phosphotransferase system (PTS).  Aminoacyl tRNA synthetases charge tRNA with an
-amino acid so that amino acids are concatenated from charged tRNA onto a growing
-peptides through translation. The aaRS's are characterized by an amino acid that
-they charge tRNA with.  This gene family had been a target for horizontal gene
-transfer \citep[e.g.,][]{Lamour1994,Woese2000,Koonin2001}, which lent itself to
-an example of gene transfer in bacteria.  There were also recent reports of gene
-transfer in tRNA aminoacylation \citep{Andam2010,Wang2011}.  In our analysis,
-some of them appeared to be transferred from SPY to SDE: leucyl-tRNA synthetase,
-aspartyl-tRNA synthetase, and arginyl-tRNA synthetase. Half of phenylalanyl-tRNA
-synthetase gene appeared to be transferred in the reverse direction.
-
-Bacterial phosphotransferase system (PTS), also known as phosphoenolpyruvate
-(PEP) group translocation, transports sugars such as glucose into the cell where
-phosphorylation of the sugar happens. There have been studies of gene transfer
-of PTS in bacteria \citep[e.g.,][]{Barabote2005,Zuniga2005}. As the main source
-of energy for bacteria, sugars in the bacterial niches must have been acting as
-important selective pressure. It is not surprising that PTS sugars differ among
-bacterial groups because bacterial species are correlated with their environment
-and energy sources theirof.  Table S\ref{tab:functional} shows that two PTS
-genes appeared to be transferred from SPY to SDE: {}``PTS system,
-mannose-specific IIC component'' and {}``PTS system, fructose-specific IIABC
-component.'' Whereas almost half of the former gene was inferred to be affected
-by recombinant edges, only 3\% of the latter gene appeared to be affected.
-
-% This agreement
-% between the two different methods, applied on somewhat different parts of the
-% data, serves as validation for the signal picked up by these separate
-% association studies.
-
-% These topology-not-changing recombinant edges tend change only branch lengths.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Should we keep this section?}
-
-With the slight over-estimation of recombination events we were concerned with
-any potential biases hidden in \texttt{ClonalOrigin} implementation. Because the
-simulated data were generated with no directionality between SPY and SDE,  if we
-estimate recombination events from the simulated data sets using
-\texttt{ClonalOrigin} and compare those estimates from those from the real data
-set, this could remove biases in \texttt{ClonalOrigin} implementation. We could
-use inferred estimates of recombination from the simulated data under the prior
-model of \texttt{ClonalOrigin} as the prior expected number of recombination.
-Figure S\ref{fig:h2} shows a plot of numbers of recombination events from the
-real data against those from the simulated data sets. Note that the numbers of
-recombination events from simulated data sets appear to deviate from the line of
-slope being 1 and intercept being 0.  The numbers of recombination events for
-the recombinant edges from SPY to SDE increase by more than 2.5 fold whereas
-those for the recombinant edges from SDE to SPY increase by about 1.7 fold.
-This was not surprising because inference of number of recombinant edges was
-quite accurate. We considered that subtraction of effects of number of
-recombinant edges inferred from simulated data and subtraction of effects of the
-prior expected number of recombinant edges were effectively equivalent.
-However, this comparison between inferrence from simulated data and that of real
-data set suggested that the directional difference between SPY and SDE was due
-to neither lack of power of \texttt{ClonalOrigin} nor hidden bias thereof.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% 
-%(Table S\ref{tab:functional-sde-spy})
-%(Table S\ref{tab:functional-spy-sde})
-%(Table S\ref{tab:go-events-duplication})
-%(Table S\ref{tab:go-events-loss})
-%
-%
-%For horizontal gene transfer we found enrichment in very different categories
-%related to transposition, integration, and recombination (Table
-%\ref{tab:go-events}).  
-%
-%Gene duplication and loss events were enriched for
-%categories similar to those of horizontal gene transfer's enrichment (Tables
-%S\ref{tab:go-events-duplication} and S\ref{tab:go-events-loss}).
-%
-%Gene duplication and horizontal transfers showed enrichment in the same
-%categories perhaps because both mechanisms increase gene copy number.  
-%
-%Gene losses showed enrichment for transposition and integration, but less
-%significantly (P-value of $<10^{-6}$), and showed enrichment for several
-%additional categories such as transporter activity and transcription initiation.
-%
-%Comparing two types of recombinant edges, one from SDE to SPY and another from
-%SPY to SDE transport related categories were more associated for the direction
-%from SPY to SDE
-%(Table S\ref{tab:functional-sde-spy} and S\ref{tab:functional-spy-sde}).
-%
-%\texttt{More interpretations about Mawgli analysis of function categories?
-%Why transposases?}
-
-% \texttt{IG: NEXT PARAGRAPH CONTAINS SOME DISCUSSION ON THESE RESULTS. WE SHOULD 
-% GET BACK TO THIS ONCE WE DECIDE WHAT TO DO WITH THE RESULTS}
-
-%The parsimony-based approach also showed
-%enrichment of the recombining transfers in many of the same categories such as
-%translation, ribosome, and rRNA binding when using all types of recombinant
-%edges in computing recombination intensity (see Tables
-%S\ref{tab:functional-all} and S\ref{tab:go-events-recombining}).  
-%Because the
-%parsimony-based method did not take species tree branch lengths into account, 
-%the agreement between the two different methods might be likely.
+\subsection*{Recombination drived more substitutions than did point mutation}
+
+Five genomes except \textit{S.\ equi} ssp.\textit{\ equi} strain 4047 were
+chosen (Table S\ref{tab:genome}) to study recombining gene transfer using the
+model-based approach.  A genome alignment of the five genomes showed two
+distinct genome structures of SPY and SDE (Figure S\ref{fig:mauve}). The SDE and
+SDD genomes were larger in length than those of SPY. Although the graphical view
+of the genome alignment did not clearly show whether SDD was different from
+either of SDE1 and SDE2 in the genome organization, a spcies tree inferred from
+the genome alignment showed that the external SDD branch was attached to the
+internal clade branch of SDE1 and SDE2 (Figure S\ref{fig:tree5}). The
+distribution of the alignment block lengths was skewed to the right with mean of
+4188 base pairs long (Figure S\ref{fig:blocksize}), which supported that the
+model assumption of geometric block length distribution might be plausible in
+\texttt{ClonalOrigin} method (see section \textbf{Materials and Methods} and
+\citet{Didelot2010}).  The relative strength ($r/m$) of recombination compared
+to mutations indicated that recombination-driven substitutions affected more
+sites of the core genomes than did mutation-driven substitutions (Table
+S\ref{tab:clonalframe}). This could imply that in the \textit{Streptococcus}
+species bacterial recombination process appeared to accelerate molecular
+evolution of more sites than did point mutations process.
+
+\subsection*{Recombination hotspots were observed in genomes}
+
+Population mutation rate did not appear to change much across the genome of SDE1 
+(Figure S\ref{fig:scatter3}).  We found a few spots of genomic regions where the
+mutation rates were larger than twice the global estimate of mutation rate.  The
+recombinant tract length changed in a way that we could not note any particular
+positions with notable tract length values compared with other positions along
+the genome.  As recombinant tract lengths were smaller, recombination rates per
+site seemed larger (Figure S\ref{fig:rhologdelta}).  The recombination rate
+appeared to change more dynamically. Especially, a genomic region of SDE1 near
+200 kilobase pairs from the origin of replication contained alignment blocks with
+bell-shaped recombination rates.
+
+\subsection*{SPY and SDE were more closely related in about tenth of the core genome}
+
+There are 105 possible rooted labeled topologies relating five taxa.  We expected the
+dominant topology across sampled genealogies to be concordant with that of the
+species tree.  Indeed, two thirds of the local gene tree topologies (see section
+\textbf{Materials and Methods}) along the alignment blocks were the same as the
+species tree, which allowed us to be more confident of the species tree inferred
+using the alignment blocks (Table S\ref{tab:Gene-tree-topologies}).  Other
+frequently sampled topologies were expected to be potentially informative about
+common recombination events.  The second most frequent topology (9.3\%) was one
+where SDD was the outgroup, and SPY and SDE were sister clades. Therefore, SPY
+and SDE appeared to be more closely related with each other than each to SDD's
+genome in about tenth of the core genome. The other topologies had less than 4\%
+of proportions.
+
+\subsection*{Unbalaned recombining gene flow from SPY to SDE in genome-wide scale}
+
+The heat map of Figure \ref{fig:Heatmap-of-recombination} (see Tables
+S\ref{tab:heatmap} and S\ref{tab:obsheatmap} for the heat map values and the
+observed number of recombinant edges) shows the logarithm base 2 fractions of the
+\textit{a posteriori} number of recombinant edges for a pair of donor and
+recipient species tree branches and their corresponding expected values given
+the species tree (Figure S\ref{fig:tree5}) and the three population parameters
+(Table S\ref{tab:three}).  Because recombinant edges from any species tree
+branch to the root branch are impossible (see Figure S\ref{fig:clonalorigin}),
+the column for the root as recipient was colored green, indicating invalid
+recombinant edges in Figure \ref{fig:Heatmap-of-recombination}.  
+Similarly, all of the cells for which a recipient branch was
+older than the donor, and the donor and recipient branches did not share any contemporary time were
+colored green. The diagonal cells were colored gray, indicating no significant
+number of recombinant edges, because the observed and expected numbers of
+recombinant edges were similar. As recombinant edges that start from and end in
+the same species tree branch would not change the species tree topology, which
+might not
+change the likelihood much, the cells would be colored gray.  Because SDD is a
+strictly veterinary pathogenic bacteria, we did not expect that SDD received
+much gene transfer from either SPY1 or SPY2: the logarithm base 2 fraction from SPY1 to
+SDD was -2.20, and that from SPY2 to SDD was -1.97.  On the contrary, the
+recombing gene transfer from SDE1 to SDD was 0.756, and that from SDE2 to SDD
+was 0.698. This relatively higher recombining gene transfer from two SDE
+individuals to SDD compared with that from two SPY individuals to SDD reflected
+the fact
+that the SDD was more closely related to SDE than to SPY in most of the genomes,
+and the genomic organization between SDD and SDE1/SDE2 were much similar
+(Figure S\ref{fig:mauve}).  However, the gene transfer from SDD to either SDE1
+or SDE2 was less frequent than the case of reverse direction; SDD effectively
+served as a recipient rather than a donor. It was unclear why SDD received
+more genes from SDE1 or SDE2 than did SDE1 or SDE2 from SDD.
+
+Because it was expected that the recombining gene transfer within a clade would
+be more frequent than that between clades, not surprisingly the recombining gene
+transfer logarithm base 2 fractions between SDE1 and SDE2 (1.17 and 1.20) were about
+twice as large as those from SDE1 and SDE2 to SPY1 and SPY2 (0.67, 0.66, 0.60,
+and 0.39).  The intensities of recombining gene transfer from external branches
+of SDE1 and SDE2 to external branches of SPY1 and SPY2 was similar to those from
+SDE1 or SDE2 to SDD (0.76 and 0.70).  Therefore, either SPY1 or SPY2, and SDD as
+recipient appeared to be no different from the view point of SDE1 or SDE2 as
+donor since the similar values of recombining gene transfer intensity from SDE1
+or SDE2 to one of the other three external branches. Surprisingly, the recombining
+gene transfer logarithm base 2 fractions between SPY1 and SPY2 (1.06 and 1.07) was
+comparable with, yet even less than those from SPY1 and SPY2 to SDE1 and SDE2
+(1.25, 1.25, 1.17, and 1.29).  The recombining gene transfer from a clade SPY1 
+or SPY2 to a different clade SDE1 or SDE2 seemed to be effectively no different
+from that within the clade of SPY1 and SPY2.  The two external branches of SDE
+received genes from either SPY1 or SPY2 via recombining gene transfer as if SDE1
+and SDE2 were in the same clade as SPY1 and SPY2.  This was more surprising
+because the bacterial recombination between SPY and SDE seemed to happen more
+often than that between SDE and SDD even though the genome configurations of SDE
+and SDD were more similar than each to SPY.  However, this result could be
+concordant with the fact that SPY and SDE shared the same ecological niche
+despite their dissimilar genomic organization.  The gene transfer logarithm base 2
+fraction from SPY1 and SPY2 to SDE branch was also larger than any type of
+recombining gene transfer: 1.88 and 1.94, respectively.  Additionally we found
+58 genes for which a recombinant edge of certain types were sampled with
+probability at least 0.9 by focusing on specific types of recombinant edges
+(Table S\ref{tab:genes-transfer}).  Many of the genes in the list were
+transferred from SPY to SDE rather than the reverse direction overall.  All
+things considered, this epitomized that the recombining gene transfer was biased
+toward the net direction from SPY to SDE.
+
+We tried in vain to quantify differential horizontal gene transfer between SPY
+and SDE with the parsimony-based approach (Figure \ref{fig:hgt-heatmap}).  Among
+the clade-crossing transfers, we did not see any significant enrichment for a
+specific direction, as opposed to the result for recombining transfers from the
+model-based analysis.  SDE1 and SDE2 shared the most events (26 and 34) compared
+to other sample pairs.  SDD appeared to acquire many genes from the ancestral
+SDE branch.  Transfer events between the SPY and SDE clades were inferred in
+smaller numbers ($<6$) compared to those within each clade. We were puzzled by
+this finding because virulence genes were found to be associated with horizontal
+gene transfer, and yet we could not show net directionality of horizontal gene
+transfer between SPY and SDE.  However, we inferred relatively more recombining
+gene transfer events from the SPY branch to SDE branch even in the
+parsimony-based approach (Figure S\ref{fig:mowgli-recomb-heatmap}).
+
+\subsection*{Reduction of genes and divergence of SPY and SDE species}
+
+The parsimony-based approach estimated 304 duplications, 944 recombining gene
+transfers, and 294 horizontal gene transfers, comprising 23.7\% of total
+transfer events (Figure \ref{fig:Gene-duplication-loss}).  Among the 2314 gene
+families the largest family consisted of 59 genes, while nearly half of gene
+families (1066) had exactly one gene per species, or six genes total (Figure
+S\ref{fig:famsizes}).  Numbers of genes at the five ancestors were inferred to
+be smaller than those at the extant species.  SDE lineages from the root
+appeared to have increased in the number of genes while the common ancestor of
+SPY1 and SPY2 decreased in the number of genes.
+
+The four external lineages of SDE1, SDE2, SDD, and SPY1 had almost equal numbers
+of gene gain and loss events while the four lineages of SP, SD, SPY2, and SEE
+had larger numbers of gene gain events than those of gene loss events. In the
+two lineages of SP and SEE we might have inferred gene gain and loss events less
+confidently than events in other lineages because of lack of outgroup lineages
+to compare with the two lineages. The number of gene gain events in SPY2 lineage
+seemed to be large compared with those of gene gain events in the other
+lineages. On the contrary, two internal branches of SPY and SDE had smaller
+numbers of gene gain events than those of gene loss events.  This might indicate
+that SPY and SDE lineages had reduced number of genes to specialize in their
+specific niches after each split from their corresponding direct ancestors.
+
+Duplications in extant lineages were more frequent than those in the other
+lineages.  Although the number of duplication events along SPY2 branch seemed
+smaller than those of the other extant lineages, duplication events appeared to
+be most often species-specific.  Within clades we found that duplications made
+up only a small fraction of the events  (10.2\%), while horizontal transfers and
+losses were significantly more frequent (41.7\% and 48.0\%, respectively).  The
+extant species of SDD and two SDE's were the most expanded, both in terms of
+duplications and in-bound transfers. This was apparently related with the larger
+genome sizes of the two species than that of SPY.  
+The numbers of recombining gene transfer events tended to be larger than those
+of horizontal gene transfer events except SD lineage.  Numbers of recombining
+gene transfer in the extant lineages might have been larger than those in the
+internal branches because the number of recombining gene transfer events must
+had been affected by the availability of more closely related individual
+genomes.
+
+\subsection*{Virulence genes via recombining gene transfer}
+
+About 10\% of the genes comprised of virulence facters, about half of which were
+found in the core genome of the five genome alignment \citep{Suzuki2011}.  Using
+customized UCSC genome browser tracks available at
+\url{http://strep-genome.bscb.cornell.edu} we located virulence genes with
+notable posterior recombination probability.  These genes were by no means a
+comprehensive list of virulence factors that might have been transferred via
+homologous recombination.  We wished to showcase virulence genes using the
+result of our study of gene transfer.
+
+Stressful host conditions for pathogenic bacteria can include heat shock,
+starvation, and iron depletion, etc.  Iron uptake of bacteria was known to be
+essential for survival of bacteria under the hostile host condition. We found
+that parts of putative calcium transporter (MgtA) in SDE1 could have been
+transferred from SPY1 (SpyM3\_0440).  An extracellular protein, putative
+dipeptidase (SpyM3\_0465) that could be important for bacteria to scavenge
+dipeptides from their host, was inferred to be transferred from SPY to SDE: the
+whole gene was transferred, which was a case for the finding that virulence
+genes tended to be transferred as a whole unit \citep{Chan2009}.  Another
+example of whole gene transfer was superoxide dismutase that was reported to be
+associated with virulence of \textit{Streptococcus agalactiae}
+\citep{Poyart2001}: the whole gene including its 5' UTR, (SpyM3\_1071), was
+affected by recombining gene transfer from SPY to SDE.  Some of the genes were
+only affected in their parts.  The middle parts of an adhesins gene, putative
+internalin A precursor (SpyM3\_1035), was found to be affected by recombining
+gene transfer from SPY to SDE.  The middle part of a chaperonin GroEL, putative
+heat shock protein (SpyM3\_1765), was inferred to transfer from SPY to SDE1:
+increase of GroEL expression was known to allow for the bacteria to survive the
+hostile host environment of acid shock.  Parts of the gene (SpyM3\_1749),
+pyruvate formate lyase that was recently reported to have virulence
+\citep{Yesilkaya2009}, seemed to be affected by recombining gene transfer.
+Under starvation condition bacteria must use more effective energy metabolism: a
+putative glucose kinase, (SpyM3\_1180), were inferred to be transferred from SPY
+to SDE to adapt SDE's energy metabolism to survive such harsh condition.  We
+also found that an MLST locus, called DNA mismatch repair protein (MutS),
+appeared to be affected by recombining gene transfer.
+
+\subsection*{Association of virulence genes and horizontal gene transfer}
+
+Horizontal gene transfers were found to be significantly associated with
+virulence genes (P-value of $<1.33 \times 10^{-11}$) whereas recombining gene
+transfers were not enriched for virulence genes (P-value of $>0.90$) using the
+parsimony-based approach.  This was also consistent with the fact that we could
+not find the association of recombining gene transfer with virulence genes in
+the model-based approach.  This supported more evidence that non-recombining
+gene transfer served as a more effective means by which SDE or SPY acquired
+virulence genes, compared to recombining gene transfer.  Families with high
+numbers of duplications were also significantly enriched for virulence genes
+(P-value of $<7.13 \times 10^{-3}$) whereas families with high losses showed no
+enrichment (P-value of $>0.54$).
+
+Roughly, half of the virulence genes were in the pan genome as mentioned above.
+We did not attempt to comprehensively list virulence genes that were inferred to
+be transferred horizontally or non-recombining fashion. We rather sort virulence
+genes by the number of horizontal gene transfer events to list the first several
+genes with high ranks.  Streptococcal C5a peptidase (\textit{scpB}, SDEG\_0933)
+was inferred to have six horizontal gene transfer.  Putative transposase were
+found to have a relatively large number of horizontal gene transfer events
+(SpyM3\_1016, SpyM3\_1053, SpyM3\_1054). Other genes with relatively large
+number of horizontal gene transfer events included putative response regulator
+(SpyM3\_0174), putative maltose operon transcriptional repressor (SpyM3\_0982),
+putative histidine kinase (SpyM3\_0171), and putative cell wall hydrolase
+(SpyM3\_0731, SpyM3\_0922, SpyM3\_1208).  Virulence genes encoding putative
+nucleases with a secretion signal peptide were also found to have had
+experienced horizontal gene transfer: genome-encoded streptodornase
+(SDEG\_0541), and prophage-associated deoxyribonuclease (SDEG\_1103).  Protein
+F2 like fibronectin-binding protein (SpyM3\_0104), an adhesin and invasin, was
+inferred to horizontally transferred.
+
+\subsection*{Model-based approach or parsimony-based one for recombining gene
+transfer study}
+
+A full model-based approach would be desirable in analyzing data for testing
+relevant hypotheses more systematically. Model complexity might not allow easy
+access to the implementation of a model. Our study involved a model-based
+approach and a parsimony-based one for understanding evolution in the core
+genomes of several bacteria individuals. Because of inability of the model-based
+method for studying the pan genomes we had to resort to the parsimony-based
+method. However, the parsimony-based method was also applied to the core
+genomes. We wished to compare the two methods in their results of the core
+genome analyses.  Because of two independent analyses on the same data set, this
+would allow us to be more confident of the results if the two analyses performed
+similarly. Generally, phylogenetic methods of detecting gene transfers often take
+advantage of discrepancy between a reference species tree and gene trees. The
+two approaches that we employed in this study were no exception. We found a few
+similar results in both methods; firstly, functional categories that were
+associated with recombining gene transfer were similar (Tables
+S\ref{tab:functional-all} and S\ref{tab:go-events-recombining}).  Secondly,
+either of them did not show association of virulence genes with recombining gene
+transfer.  To more directly compare the methods we used the recombination
+intensity and the number of recombining gene transfer events for each gene.  We
+found a weak positive correlation of 0.39 with a significant P-value
+of $2.2\times10^{-16}$ for testing a null hypothesis of the correlation being
+zero.  (Figure S\ref{fig:cmpcomowgli}).  Interestingly, the correlation (0.03
+with P-value of 0.18) disappeared in a comparison between recombination
+intensity and number of horizontal gene transfer events (Figure
+S\ref{fig:cmpcomowglihgt}).
+
+\subsection*{UCSC genome browser tracks for bacterial gene transfer study}
+
+Often computational and statistical analyses on genomes could benefit other
+researchers when those resources were provided: e.g., UCSC Genome Browser, and
+Ensembl Genome Browser. We wished to provide research community especially in
+SPY and SDE bacteria with some of the results from our analysis of gene
+transfer. We used the UCSC genome browser for the purpose.  The resource shown
+in Figure S\ref{fig:ucsc} for gene transfer is available at
+\url{http://strep-genome.bscb.cornell.edu}.  For each of the five individual
+genomes as a reference genome several annotation tracks were displayed:
+alignment blocks, recombination rate per site for each block, local gene tree
+topologies, recombination intensity, locations of virulence genes, posterior
+probability of recombining gene transfer. The genome track of probability of
+recombining gene transfer allowed us to easily browse the bacterial genome to
+locate regions that we might be interested in. 
+% To Matt: what could we show in the pan-genome? or from Mowgli's analysis?
+
+\subsection*{Units of recombining gene transfer}
+
+The process of gene transfer may or may not move whole genes from one species to
+another \citep{Chan2009}. We observed that units of recombining gene
+transfer could be multiple genes (Table S\ref{tab:genes-transfer}). In some
+cases, parts of a gene were inferred to have been transferred, for instance,
+transfer from SPY1 to SDE1 was inferred for 29\% of the aspartyl-tRNA synthetase
+gene. Recombinant tracts were often shown to span two genes.  In other cases,
+the inferred transfer units spaned a set of as much as eight adjacent genes:
+loci from SpyM3\_1518 to SpyM3\_1525.
+
+\subsection*{Simulations}
+
+We conducted experiments on simulation data to check the accuracy of estimates
+obtained by the model-based approach in a scenario similar to the one from the
+real data set.  Population mutation and recombination parameters were accurately
+estimated (Table S\ref{tab:sim-three}).  Yet, the recombinant tract length was
+over-estimated; recombinant tract lengths were reported to be difficult to
+estimate from relatively short alignment blocks \citep{Didelot2010}.  Numbers of
+recombinant edges from the simulated recombinant trees of Simulation 2 (see
+\textbf{Supplementary Information} for the description of Simulation 2) were
+compared with the corresponding numbers of recombinant edges that were estimated
+from the simulated data set from Simulation 2 (see the solid-line intervals with
+open circles in Figure \ref{fig:h3}).  We noted that tree-topology-changing
+recombinant edges were more reliably recovered compared with
+tree-topology-not-changing ones (Figure \ref{fig:h3}). All of the
+tree-topology-not-changing recombinant edge types were over-estimated (Figure
+\ref{fig:h3}B). The heat map of Figure \ref{fig:Heatmap-of-recombination} did
+not show how confident of the values we could be. Comparison of numbers of
+recombinant edges estimated from the real data set and the corresponding numbers
+estimated from Simulation 2 showed the interval estimates for the heat map
+values.  Dotted-line intervals from the red and blue colored cells in the heat
+map were above and below the reference of the horizontal line at the zero on the
+Y-axis, respectively (Figure \ref{fig:h3}).
+
+For each recombinant edge type to assess the effect of posterior recombinant
+trees samples with respect to that of prior recombinant tree samples. The ten
+\textit{a priori} recombinant trees from Simulation 2 would have balanced
+recombining gene transfer events between SPY and SDE with respect to the prior
+expected number of recombinant edges. The posterior recombinant trees used in
+Simulation 3 would have unbalanced recombining gene transfer between the two
+species branches as we observed in the real data analysis. For significantly
+strong gene flows such as ones from SPY to SDE a solid-line interval with a
+closed circle would be different from its count part of the dotted-line
+interval.  Figure \ref{fig:h3} showed that the recombining gene flow from SPY to
+SDE was stronger enough to be nearer to the horizontal line at the zero on the
+Y-axis. However, the recombining gene transfer from SDE to SPY appeared to be
+too weak.  This comparison supported the biased directionality of recombining
+gene transfer from SPY to SDE.
+
+Recombination intensities that were measured on genes also showed rough
+correlation (Pearson's correlation coefficient of 0.63) of the estimated and
+true values (Figure S\ref{fig:ri1}).  Estimated values appeared to be driven
+somewhat by the prior implied by the population parameter setting, explaining
+why lower values were over-estimated and higher values were under-estimated.
+Nonetheless, the correlation indicated that the measure of recombination
+intensity used in our functional gene category served as an adequate indicator
+of actual intensity of recombination events along genes.

### 2011-10-19 doc/manuscript.tex
I am editting the Results section that had been modified by Ilan.

@@
 
-\section*{A list our major and minor findings or observations}
+\section*{A list of our major and minor findings or observations}
 \begin{enumerate}
@@
 section \textbf{Results}.
-\item Ilan suggested a different order of four parts in \textbf{Results}:
-preliminary, funtional category and virulence gene analyses, gene gain and loss,
-and directionality.
+\item Can selections after gene transfer be tested?
 \end{enumerate}
@@
 exotoxins. Davies et al.\ \citet{Davies2007} investigated phage 3396 from SDE as a vehicle of
-gene transfer from SDE to SPY. They also \citet{Davies2009} demonstrated that an
+gene transfer from SDE to SPY. They \citet{Davies2009} also demonstrated that an
 integrative conjugative element of SDE allowed gene transfer from SDE to other
@@
 \citep{Tettelin2009a}.  Studies on horizontal gene transfer in prokaryotes have
-used genes in very large sets of species \citep{Koonin2001} from a great deal of
-genomes available.  Although this type of study gave us insights into deep
+used genes in very large sets of species from a great deal of genomes available 
+\citep{Koonin2001}.  Although this type of study gave us insights into deep
 evolutionary histories of prokaryotes, with multiple individual genomes per
@@
 
-Recently, a method of inferring
-homologous recombination with whole bacterial genomes was developed
-\citep{Didelot2010}. The model of homologous recombination allowed them to study
-flows of genetic segments between bacteria individuals of \textit{Bacillus
-cereus} species in the core genomes.   
-% Descriptions of ClonalOrigin model.
-Other evolutionary processes acting on
-bacterial genomes include genome rearragement, translocation, inversion,
-duplication, gain, loss, etc.  These processes did not readily lend themselves
-to statistically more sound model-based approaches such as maximum likelihood.
+Recently, a method of inferring homologous recombination with whole bacterial
+genomes was developed \citep{Didelot2010}. The model of homologous recombination
+allowed them to study recombining gene flows between bacteria individuals of
+\textit{Bacillus cereus} species in the core genomes.   In the model, a species
+tree that relates individuals genomes is augmented by attaching
+\textit{recombinant edges} between species tree branches that can be used to
+measure gene transfers.  Other evolutionary processes acting on bacterial
+genomes include genome rearragement, translocation, inversion, duplication,
+gain, loss, etc.  These processes did not readily lend themselves to
+statistically more sound model-based approaches such as maximum likelihood.
 Instead, a parsimony-based approach to inferring gene duplication, loss, and
-transfer events was developed by Doyon et al.\ \citet{Doyon2011}.  
-% Descriptions of Mowgli analsysis.
-We noted that the first
-approach could be used to infer transfer of genomic segments due to homologous
-recombination, and the second could be useful in studying horizontal gene
-transfer.
-% Homologous recombining gene transfer could be studied more readily using the
-% core genomes of individual bacteria while horizontal gene transfer might be more
-% easily detected using the pan genomes.
+transfer events was developed by Doyon et al.\ \citet{Doyon2011}.  In the
+approach, gene trees are inferred for gene families that are constructed from
+all of the genomes, and are reconciled with a speices tree to infer
+parsimoniously events.  We noted that the first approach could be used to infer
+transfer of genomic segments due to homologous recombination, and the second
+could be useful in studying horizontal gene transfer.
 
@@
 recombination, and the parsimony-based one for detecting gene duplication, loss,
-and horizontal transfer. Virulence genes were found to be more associated
+and horizontal transfer.  We found that virulence genes were more associated
 non-recombining horizontal gene transfer than non-virulence genes.  We compared
@@
 genome but also in the pan genome, which made the analysis complete in terms of
-the amount of used genome data. Our analysis of the core genomes also showed for the
-first time that the net directionality of recombining gene transfer from SPY to
-SDE was stronger at the genome scale.
+the amount of used genome data. Our analysis of the core genomes also showed for
+the first time that the net directionality of recombining gene transfer from SPY
+to SDE was stronger at the genome scale.
 
@@
 
-\subsection*{Distinguishing two gene transfer processes shows different functional
-categories associations}
-
-Bacterial gene transfer processes as an evolutionary force would affect more
-frequently in some biological functions than in others.  In our attempt of
-finding which biological functions were associated with recombining or
-horizontal gene transfers, we devised a measure of recombination intensity for
-genes by using frequency of recombinant edges that affect those genes, and used
-frequency of horizontal gene transfer events (see section \textbf{Materials and
-Methods}).  Table \ref{tab:functional} shows a list of function categories
-that were found to be associated with recombining gene transfers. We noted three
-biological functions in the list: membrane-bound molecules, aminoacyl tRNA
-synthetases, and ribosomes.
-
-Firstly, we have sugar phosphotransferases ranked
-top in the list; phosphrylation could be important in signal transduction to
-recognize environmental changes.  Transport functional categories of
-carbohydrate transport and sodium ion transport would belong to genes encoding
-membrane-bound molecules.  Also highly ranked were fatty acid and lipid
-biosyntheses that must be involved in producing membrane-bound molecules. 
-Because bacteria as a unicellular organism should be able to respond to changes
-to environment, some evolutionary pressures must have been acting on genes
-encoding membrane-bound molecules. Therefoure, recombining gene transfer process
-appeared to be more associated with biological functions that had been critical
-of sensing queues of bacterial environmental changes.
-
-Secondly, we found several aminoacyl tRNA synthetases (GO Term translation
-GO:0006412) in association with recombining gene transfer.  In the
-parsimony-based approach the term translation was also top ranked (Table
-S\ref{tab:go-events-recombining}).  Because they had been a focus of gene
-transfer in bacterial species \citep{Woese2000}, we could confirmed that the
-recombination intensity and requency of recombining gene transfer events would
-be useful in finding genes associated with recombining gene transfer.
-
-Thirdly, functional categories concerning ribosomes were unexpectedly
-found to be associated with recombining gene transfer.  Because genes related
-with ribosomes were considered to be housekeeping genes, gene transfer in which
-would be rare, it was unexpected that ribosome related genes were found to be
-associated with recombining gene transfer.  Because it was a bizarre result, we
-investigated the association by considering different recombinant edge types. We
-noted that two different classes of recombinant edge types: one that changes
-species tree topologies, and another that does not
-(see section \textbf{Materials and Methods}).
-Because recombinant edges
-that did not change species tree topology might be difficult to infer reliably,
-considering those recombinant edges in computing recombination intensity might
-lead to dubious results.  Ribosome related functional categories were found to
-be top ranked if we took those recombinant edges that did not change species
-tree topology (Tables S\ref{tab:functional-notopology} and
-S\ref{tab:functional-all}).  In Table \ref{tab:functional} ribosome related
-functional categories were not top ranked.  Therefore, high recombination
-intensity in ribosome related genes were likely to be attributed to
-topology-not-changing recombinant edges. We would be cautious of recombining
-transfer of genes related with ribosome related functions.
-
-Table \ref{tab:go-events} shows a list of functional categories associated with
-horizontal gene transfer, which were different from those from recombining gene
-transfer analysis.  Ranked top in the list was transposases, which were known to
-help foreign genes be excised from a donor genome and integrated into a
-recipient genome. The biological role of transposases was concordant with
-horizontal gene transfer process because the process involved in transposases
-would allow bacteria to acquire foreign genes without the requirement of
-homologous regions in the bacteria.  
-These differences in the profile of functional categories association of the two
-gene transfer processes were because of the approach of dicriminating the two
-gene transfer processes.
-
-\subsection*{Recombining gene transfer from SPY to SDE at genome-wide scale}
-
-We show a net directionality of recombining gene transfers from SPY to SDE.
-
+\texttt{We show a net directionality of recombining gene transfers from SPY to SDE. 
 Describe our arguments of the net directionality. 
@@
 Is this number of recombinant edges good enough for the argument?
-
 Reflection of the previous analyses, 
 heat maps of real data analysis, 
-and supporting simulations.
-
-\subsection{Bayesian analysis of recombining gene transfer}
-
-{\bf IG: REMOVED A LOT OF DETAIL (MAINLY IN FIRST PART). PLEASE CHECK.}
-
-We followed the pipeline and methods suggested in \cite{Didelot2010} to pursue a model based study of recombining gene transfer.
-A preliminary step in this analysis involved obtaining an estimate of the species tree with branch lengths (via ClonalFrame), and estimates of the three population parameters: mutation rate, recombination rate, and average recombinant tract length (via a first phase of ClonalOrigin).
-Due to divergence of the SEE sample from the other five, we could not obtain suficiently long reliable multiple sequence alginments with it, and had to exclude it from the bayesian analysis.
-The spcies tree inferred by ClonalFrame over the five species has the expected topology (Fig.
-{\bf [ADD FIGURE 1]} (see Figure S\ref{fig:tree5} for branch length estimates).
-ClonalFrame also produces estimates for the relative strength of recombination compared
-to mutations ($r/m$; Table S\ref{tab:clonalframe}).
-The estimate of $r/m=4.66$ provides an initial indication to the important part recombination took in the evolutionary process of shaping these genmes: four times more sites have been affected by recombination-driven substitutions than have been affected by mutation-driven substitutions.
-Estimates of population mutation rate ($\theta$), recombination rate ($\rho$), and average tract length ($\delta$) were obtained for the different alignment blocks analyzed (Figure S\ref{fig:scatter3}).
-Global estimates of these three population parameters were obtained by a length-weighted average (Materials and Methods).
-The local estimates typically seem to be moderately fluctuating around the global average, with the exception of what apears to be a recombination hotspot at a region 200 kilobase pairs from the origin of replication of SDE1.
-
-The species tree and estimated branch lengths, together with the global estimates of the three population parameters are used in the second phase of ClonalOrigin, where recombinant trees are sampled from the posterior distriution.
-One way to summarize the posterior sample of recombinant trees provided by ClonalOrigin is to observe the topologies of the local genealogies implied by it.
-Out of the 105 possible rooted labeled topologies spanning five taxa, the dominant topology (sampled 66.7\% of the time) was that of the species tree (Materials and Methods; Table S\ref{tab:Gene-tree-topologies}).
-The second most frequently sampled topology (sampled 9.3\% of the time) was the one in which SPY and SDE are sister clades (with SDD as an outgroup).
-Other topologies were sampled with frequency less than 4\%.
-This gives initial indication to the extent of recombining gene transfer  between SPY and SDE.
-
-To gain a more fine-scale view of the main types of recombination scenarios sampled by ClonalOrigin, we looked at the recombinant edges present in the sampled recombinant trees.
-Following the approach of \cite{Didelot2010}, we classified these edges by types defined by distinct pairs of donor and recepient branches of the species tree (Fig.
-[ADD FIG 1]).
-For each edge type, we recorded the average number of edges of that type present in the sampled recombinant tree, and computed the ratio of this observed count and the prior expected count as implied by the species tree (considering branch lengths) and the three population parameters estimated in the preliminary round of the analysis.
-The base 2 logarithms of these ratios are depicted in the heatmap of Figure \ref{fig:Heatmap-of-recombination} (see Tables
+and supporting simulations.}
+
+\subsection*{Recombining gene transfer from SPY to SDE at genome-wide scale}
+% \subsection*{Bayesian analysis of recombining gene transfer}
+
+We followed the pipeline and methods recommended by Didelot et al.
+\cite{Didelot2010} to pursue a model-based study of recombining gene transfer. A
+preliminary step in the pipeline involved estimating the species
+tree with branch lengths (via \texttt{ClonalFrame}), and three population
+parameters of mutation rate, recombination rate, and average recombinant tract
+length (via a first phase of \texttt{ClonalOrigin}).  Because 
+\textit{S.\ equi} ssp.\textit{\ equi} strain 4047 (SEE)
+diverged significantly from the others among the six genomes (Table S\ref{tab:genome}), we
+could not obtain sufficiently long multiple sequence alignments with
+it, and had to exclude it from the model-based analysis.  The estimated species tree
+using the alignment of the five genomes (Figure
+S\ref{fig:mauve}) had the expected topology \texttt{The figure 1 must be nicer
+than one that now.} (see Figure \ref{fig:tree5}). \texttt{ClonalFrame} also
+estimated
+a relative strength of recombination compared to mutations ($r/m=4.66$), 
+indicating potentially significant roles of recombination in the evolutionary
+process of shaping these genomes; four times more sites had been affected by
+recombination-driven substitutions than by mutation-driven substitutions.
+($r/m$; Table S\ref{tab:clonalframe}).
+
+Estimates of population mutation rate ($\theta$), recombination rate ($\rho$),
+and average tract length ($\delta$) were obtained for the alignment blocks
+analyzed (Figure S\ref{fig:scatter3}), and 
+medians of each of the three parameters weighted over block lengths were
+obtained as global estimates of the three population
+parameters (Table S\ref{tab:three} and see \textbf{Materials and Methods}).  
+The local estimates typically seemed to be
+moderately fluctuating around the global average, with the exception of what
+appeared to be a recombination hotspot at a region 200 kilobase pairs from the
+origin of replication of SDE1.
+
+The species tree and estimated branch lengths, together with the global
+estimates of the three population parameters were used in the second phase of
+ClonalOrigin, where recombinant trees were sampled from the posterior
+distribution. One way to summarize the posterior sample of recombinant trees
+provided by ClonalOrigin was to observe the topologies of the local genealogies
+induced by it. Out of the 105 possible rooted labeled topologies spanning five
+taxa, the dominant topology (sampled 66.7\% of the time) was that of the species
+tree (Materials and Methods; Table S\ref{tab:Gene-tree-topologies}). The second
+most frequently sampled topology (sampled 9.3\% of the time) was the one in
+which SPY and SDE were sister clades (with SDD as an outgroup). Other topologies
+were sampled with frequency less than 4\%. This gave initial indication to the
+extent of recombining gene transfer  between SPY and SDE.
+
+To gain a more fine-scale view of the main types of recombination scenarios
+sampled by ClonalOrigin, we looked at the recombinant edges present in the
+sampled recombinant trees. Following the approach of Didelot et al. \cite{Didelot2010}, we
+classified these edges by types defined by distinct pairs of donor and recipient
+branches of the species tree (Figure \ref{fig:tree5}). For each edge type, we recorded
+the average number of edges of that type present in the sampled recombinant
+tree, and computed the ratio of this observed count and the prior expected count
+as implied by the species tree (considering branch lengths and their relative
+positions) and the three
+population parameters estimated in the preliminary round of the analysis. The
+base 2 logarithms of these ratios are depicted in the heatmap of Figure
+\ref{fig:Heatmap-of-recombination} (see also Tables
 S\ref{tab:heatmap} and S\ref{tab:obsheatmap}).
 
-Among the possible recombinant edge types, we distinguished between ones which preserve the species tree topology, and ones which alter it.
-Our main focus was on the topology-altering types, both because our main objective was to study cross-clade gene transfer, and because our simulation study indicates that the observed counts of topology-preserving recombinant edges are somewhat less reliable {\bf[IG: POINT TO RELEVANT SECTION]}.
-The heatmap allows us to find recombination scenarios that deviate from the expectation dictated by the probabilistic model of ClonalOrigin.
-{\bf [IG: CHECK THIS FOLLOWING LINE OF ARGUMENT]} We identify two main potential causes for such a deviation: (1) reduced or increased opportunities for species to interact and exchange genetic material; (2) selection acting on newly transferred genes.
-
-We observe a reduced count of recombinant edges sampled between SDD and the three branches of the SPY clade ($\sim$25\% of the expected count).
-This is consistent with the fact that these species have reduced opportunities to interact, SPY being a human pathogen and SDD being strictly veterinary.
-On the contrary, the number of sampled recombinant edges between SDD and branches in the SDE clade seems to be consistent with the prior, with a slight excess for edges from SDE1 and SDE2 to SDD {\bf [IG: IS THIS WORTH DWELLNG ON?]}.
-The greatest deviation of observed counts from expected ones was observed for recombinant edges between the SPY and SDE clades.
-This is consistent with the fact that these two species share an ecological niche (SDE being an opportunistic human pathogen).
-Interestingly, there appears to be much higher enrichment (up to four times more than expected) for recombinant edges originating in the SPY clade and ending up in the SDE clade.
-In a follow-up simulation study we show that this deviation from the expected count is indeed highly significant, and so is the observed asymmetry in pattern of recombination {\bf[IG: POINT TO RELEVANT SECTION]}.
-This bias in directionality of recombination pattern between SPY and SDE is especially intriguing in light of the hypothesis that SDE acquired human pathogenicity via gene transfer from SPY.
-
-\subsection{Parsimony-based reconciliation of gene trees with the species tree}
+Among the possible recombinant edge types, we distinguished between ones that
+preserve the species tree topology, and ones that alter it. We focused
+on the topology-altering types, both because our main objective was to study
+cross-clade gene transfer, and because our simulation study indicates that the
+observed counts of topology-preserving recombinant edges were 
+unreliably inferred {\bf[IG: POINT TO RELEVANT SECTION]}. The heatmap allowed to find
+recombination scenarios that deviated from the expectation dictated by the
+probabilistic model of ClonalOrigin. {\bf [IG: CHECK THIS FOLLOWING LINE OF
+ARGUMENT]} We envisioned two main potential causes for such a deviation: (1)
+reduced or increased opportunities for species to interact and exchange genetic
+material; (2) selection acting on newly transferred genes.
+
+We observed a reduced count of recombinant edges sampled between SDD and the
+three branches of the SPY clade ($\sim$25\% of the expected count). This is
+consistent with the fact that these species have reduced opportunities to
+interact with each other, SPY being a human pathogen and SDD being strictly veterinary. On the
+contrary, the number of sampled recombinant edges between SDD and branches in
+the SDE clade seemed to be consistent with the prior, with a slight excess for
+edges from SDE1 and SDE2 to SDD {\bf [IG: IS THIS WORTH DWELLNG ON?]}. The
+greatest deviation of observed counts from expected ones was observed for
+recombinant edges between the SPY and SDE clades. It was concordant with the
+fact that these two species share an ecological niche; SDE was an
+opportunistic human pathogen. Interestingly, there appeared to be much higher
+enrichment (up to four times more than expected) for recombinant edges
+originating in the SPY clade and ending up in the SDE clade. In a follow-up
+simulation study we show that this deviation from the expected count was indeed
+highly significant, and so was the observed asymmetry in pattern of recombination
+{\bf[IG: POINT TO RELEVANT SECTION]}. This bias in directionality of
+recombination pattern between SPY and SDE was especially intriguing in light of
+the hypothesis that SDE acquired human pathogenicity via gene transfer from SPY.
+
+\subsection*{Parsimony-based reconciliation of gene trees with the species tree}
 
 {\bf[IG: DID NOT TOUCH THE FIRST 3 PARAGRAPHS. REQUIRES ``CONDENSING".]}
-
 {\bf[IG: START WITH SOME GENERAL STATISTICS - NUMBER OF GENES, ETC?]}
-
-The parsimony-based approach estimated 304 duplications, 944 recombining gene
-transfers, and 294 horizontal gene transfers, comprising 23.7\% of total
-transfer events (Figure \ref{fig:Gene-duplication-loss}).  Among the 2314 gene
-families the largest family consisted of 59 genes, while nearly half of gene
-families (1066) had exactly one gene per species, or six genes total (Figure
-S\ref{fig:famsizes}).  Numbers of genes at the five ancestors were inferred to
-be smaller than those at the extant species.  SDE lineages from the root
+\texttt{SCC: Check this ``condensed'' version.}
+
+Figure \ref{fig:Gene-duplication-loss} shows gene gain, loss, duplication, and
+two different gene transfer events along the lineages of the phylogenetic tree
+that relates the six genomes. SPY had less genes than the other species. All of
+the genomes were used to construct 2314 gene families.  The largest family
+consisted of 59 genes while nearly half of gene families (1066) had exactly one
+gene per species, or six genes total (Figure S\ref{fig:famsizes}).  We estimated
+a larger number of recombining gene transfer events than that of horizontal gene
+transfer events; 944 and 294, respectively.  Duplications in extant lineages
+were more frequent than those in the other lineages.  SDE lineages from the root
 appeared to have increased in the number of genes while the common ancestor of
-SPY1 and SPY2 decreased in the number of genes.
-
-The four external lineages of SDE1, SDE2, SDD, and SPY1 had almost equal numbers
-of gene gain and loss events while the four lineages of SP, SD, SPY2, and SEE
-had larger numbers of gene gain events than those of gene loss events. In the
-two lineages of SP and SEE we might have inferred gene gain and loss events less
-confidently than events in other lineages because of lack of outgroup lineages
-to compare with the two lineages. The number of gene gain events in SPY2 lineage
-seemed to be large compared with those of gene gain events in the other
-lineages. On the contrary, two internal branches of SPY and SDE had smaller
-numbers of gene gain events than those of gene loss events.  This might indicate
-that SPY and SDE lineages had reduced number of genes to specialize in their
-specific niches after each split from their corresponding direct ancestors.
-
-Duplications in extant lineages were more frequent than those in the other
-lineages.  Although the number of duplication events along SPY2 branch seemed
-smaller than those of the other extant lineages, duplication events appeared to
-be most often species-specific.  Within clades we found that duplications made
-up only a small fraction of the events  (10.2\%), while horizontal transfers and
-losses were significantly more frequent (41.7\% and 48.0\%, respectively).  The
-extant species of SDD and two SDE's were the most expanded, both in terms of
-duplications and in-bound transfers. This was apparently related with the larger
-genome sizes of the two species than that of SPY.  
-The numbers of recombining gene transfer events tended to be larger than those
-of horizontal gene transfer events except SD lineage.  Numbers of recombining
-gene transfer in the extant lineages might have been larger than those in the
-internal branches because the number of recombining gene transfer events must
-had been affected by the availability of more closely related individual
-genomes.
-
-{\bf [IG: MODIFIED THIS PARAGRAPH]}
-
-As in the Bayesian analysis of ClonalOrigin, we attempted to see what types of transfer events (horizontal and recombining) were most common in our inference.
-We classified events into types defined by donor-recepient branches in the species tree.
-We recorded the number of inferred events of each type in Figures \ref{fig:hgt-heatmap} \&S\ref{fig:mowgli-recomb-heatmap} {\bf [IG: CONSIDER MOVING BOTH INTO SUPPLEMENT]}.
-Unlike in the Bayesian analysis, we cannot normalize these counts by their expectation, since we do not have any model-based expectation for them.
-This restrict our ability to interprest these counts.
-We did notice, however, a very high number of inferred recombining gene transfer events from the ancestral SPY branch to the ancestral SDE branch, consistent with our observations from the Bayesian analysis.
-With horizontal gene transfer, we do not see a similar pattern, but we do observe a high number of horizontal transfers from the ancesral SDE branch to SDD.
-{\bf [IG: IS THERE ANYTHING ELSE WE CAN SAY ABOUT THIS?
-MAYBE SHOW A SCATTER PLOT OF THE TWO TYPES OF EVENTS?
-]}
-
-\subsection{Gene categories associated with recombining and horizontal gene transfer}
-
-%\subsection{Association of virulence genes and horizontal gene transfer}
-
-We set out to seek cateories of genes that are significantly enriched for transfer events of either type (recombining and  horizontal).
-For horizontal transfer, we used a score based on the number of events inferred by the parsinomy-based method, and for recombining gene transfer, we used both the parsimony-based inferred counts and a recombination intensity score based on the Bayesian analysis (Materials and Methods).
-First, we checked association of transfers with virulence genes {\bf [IG: A SENTENCE ON WHERE WE TAKE THESE GENES FROM]}.
-This set of virulence genes was found to be significantly associated with horizontal gene transfers (P-value $<1.33 \times 10^{-11}$), but not with recombining gene transfers (P-value $>0.90$ for parsimony-based score and P-value $>??$ for recombnation intensity).
-We also found a marginally significant association between virulence genes and gene duplications (P-value $<7.13 \times 10^{-3}$).
-
-%\subsection{Transporters, transferases, aminoacyl tRNA synthetases, and transposases}
-
-We conducted similar tests for other functional categories of genes.
-Several functional categories that were found to be associated with recombination intensity were related with either membrane-bound molecules or aminoacyl tRNA synthetases (Table
-\ref{tab:functional}).
-Consistent results were obtained for the parsimony-based counts of recombining gene transfers (Table S\ref{tab:go-events-recombining}).
-Functional categories associated with horizontal gene transfer were clearly different from those associated with recombining gene transfer (Table \ref{tab:go-events}).
-The top ranked functional categories in that list were related with transposases.
-These differences add to our previous findings about the different roles these two types of transfer play in the evolution of these species.
-
-{\bf [IG: I THINK THE FOLLOWING 3 PARAGRAPHS SHOULD BE MOVED TO DISCUSSION]}
-
- Molecular evolutionary forces could act on genetic materials of organisms, and changes to the genotypes might lead to some phenotypes that could change biological functions of the species.
-So, the bacterial gene transfer process as an evolutionary force could be associated with some biological functions.
-We
-attempted to find which biological functions were associated with recombining or
-horizontal gene transfers.
-
-Because bacteria as a unicellular organism should be able to respond to changes to environment, some evolutionary pressures must have been acting on genes encoding membrane-bound molecules.
-Recombining gene transfer process seemed to be more associated with biological functions that had been critical of sensing queues of bacterial environmental changes.
-Sugar phosphotransferase classes
-were ranked top in the list; phosphrylation could be important in signal
-transduction to recognize environmental changes.  The transport functional
-categories of carbohydrate transport and sodium ion transport also belonged to
-genes encoding membrane-bound molecules.  Also highly ranked were fatty acid and
-lipid biosyntheses that must be involved in producing membrane-bound molecules.
-Horizontal gene transfer process would allow bacteria to acquire foreign genes
-without the requirement of homologous regions in the bacteria.  Transposases
-would help foreign genes be excised from a donor genome and integrated into a
-recipient genome.
-
-In addition to these biological categories, we also found several aminoacyl tRNA synthetases (GO Term translation GO:0006412) in association with recombining gene transfer.
-In the parsimony-based approach the term translation was top
-ranked (Table S\ref{tab:go-events-recombining}).  They had been a focus of gene
-transfer in bacterial species \citep{Woese2000}.
-
-{\bf [IG: DO WE NEED/WANT THE FOLLOWING DISCUSSION? MAYBE IN MATERIALS AND METHODS - JUSTIFYING WHY WE MEASURE INTENSITY USING TOPOLOGY-CHANGING REC EDGES.]}
-
- Some of the functional categories were unexpected; functional categories concerning ribosomes were found to be associated with recombining gene transfer.
-Because genes related with ribosomes were considered to be housekeeping genes, gene transfer in which
-would be rare, it was unexpected that ribosome related genes were found to be
-associated with recombining gene transfer.  Because it was a bizarre result, we
-investigated the association by considering different recombinant edge types. We
-noted that two different classes of recombinant edge types: one that changes
-species tree topologies, and another that does not. Because recombinant edges
-that did not change species tree topology might be difficult to infer reliably,
-considering those recombinant edges in computing recombination intensity might
-lead to dubious results.  Ribosome related functional categories were found to
-be top ranked if we took those recombinant edges that did not change species
-tree topology (Tables S\ref{tab:functional-notopology} and
-S\ref{tab:functional-all}).  In Table \ref{tab:functional} ribosome related
-functional categories were not top ranked.  Therefore, high recombination
-intensity in ribosome related genes were likely to be attributed to
-topology-not-changing recombinant edges. We would be cautious of recombining
-transfer of genes related with ribosome related functions.
-
-\subsection{Comparison of Bayesian and parsimony-based inferences of recombining gene transfer}
-
-The similarity between the two lists of gene categories found to be associated with recombining gene transfers in the separate analyses -- Bayesian and parsimony-based --  indirectly insdicate correlation between the two approaches.
-In order to obtain a more direct measure of correlation, we used the set of genes present in both analyses to compute the correlation between recombination intensity scores and the number of recombinng gene transfers inferred by the parsimony-based approach.
-We observed a positive correlation of $0.39$ (P-value $2.2\times10^{-16}$; Figure S\ref{fig:cmpcomowgli}) between the two measures.
-For comparison, the correlation between recombination intensity and the number of horizontal gene transfers was 0.03 (P-value 0.18;  \&S\ref{fig:cmpcomowglihgt}).
+SPY1 and SPY2 decreased in the number of genes.  Whereas gene loss events were
+either comparable to or less than gene gain ones along most of the lineages, it
+was not along two internal branches of SPY and SDE.  This indicated that SPY and
+SDE lineages had reduced number of genes to specialize in their specific niches
+after each split from their corresponding direct ancestors.
+
+As in the Bayesian analysis of ClonalOrigin, we attempted to see which type of
+horizontal or recombining transfer events was most common in our inference.  We
+classified events into types defined by donor-recipient branches in the species
+tree. We recorded the number of inferred events of each type in Figures
+S\ref{fig:hgt-heatmap} and S\ref{fig:mowgli-recomb-heatmap}. Unlike in the
+Bayesian analysis, we could not normalize these counts by their expectation
+because we did not have any model-based expectation for them. This restricted
+our ability to interprest these counts. We did notice, however, a large number
+of inferred recombining gene transfer events from the ancestral SPY branch to
+the ancestral SDE branch, consistent with our observations from the Bayesian
+analysis. With horizontal gene transfer, we did not observed a similar pattern,
+but we did observe a large number of horizontal transfers from the ancesral SDE
+branch to SDD.
+
+\subsection*{Gene categories associated with recombining and horizontal gene transfer}
+
+We set out to seek categories of genes that were significantly enriched for
+recombining or horizontal transfer events. For horizontal
+transfer, we used a score based on the number of events inferred by the
+parsinomy-based method, and for recombining gene transfer, we used both the
+parsimony-based inferred counts and a recombination intensity score based on the
+Bayesian analysis (see \textbf{Materials and Methods}). First, we checked association of
+transfers with virulence genes, which were provided from a previous study
+of Suzuki et al.\ \citet{Suzuki2011}.
+This set of virulence genes was found to be significantly associated
+with horizontal gene transfers (P-value $<1.33 \times 10^{-11}$), but not with
+recombining gene transfers. We also found a marginally
+significant association between virulence genes and gene duplications (P-value
+$<7.13 \times 10^{-3}$).
+
+We conducted similar tests for other functional categories of genes. Several
+functional categories that were found to be associated with recombination
+intensity were related with either membrane-bound molecules or aminoacyl tRNA
+synthetases (Table \ref{tab:functional}). Consistent results were obtained for
+the parsimony-based counts of recombining gene transfers (Table
+S\ref{tab:go-events-recombining}).  Functional categories associated with
+horizontal gene transfer were clearly different from those associated with
+recombining gene transfer (Table \ref{tab:go-events}). The top ranked functional
+categories in that list were related with transposases. These differences add to
+our previous findings about the different roles these two types of transfer play
+in the evolution of these species.
+
+\subsection*{Comparison of Bayesian and parsimony-based inferences of recombining gene transfer}
+
+The similarity between the two lists of gene categories found to be associated
+with recombining gene transfers in the separate analyses -- Bayesian and
+parsimony-based --  indirectly indicated concordance between the two independent
+approaches.  In order to obtain a more direct measure of correlation, we used
+the set of genes present in both analyses to compute the correlation between
+recombination intensity scores and the number of recombining gene transfers
+inferred by the parsimony-based approach. We observed a positive correlation of
+$0.39$ (P-value $2.2\times10^{-16}$; Figure S\ref{fig:cmpcomowgli}) between the
+two measures.  For comparison, the correlation between recombination intensity
+and the number of horizontal gene transfers was 0.03 (P-value 0.18;
+S\ref{fig:cmpcomowglihgt}).
+
 {\bf [IG: MAYBE THERE IS A BETTER WAY TO DEMONSTRATE THE CORRELATION BETWEEN METHODS]}.
 
-{\bf [IG: MOVE TO DISCUSSION:]}
-A full model-based approach would be desirable in analyzing data for testing
-relevant hypotheses more systematically. Model complexity might not allow easy
-access to the implementation of a model. Our study involved a model-based
-approach and a parsimony-based one for understanding evolution in the core
-genomes of several bacteria individuals. Because of inability of the model-based
-method for studying the pan genomes we had to resort to the parsimony-based
-method. However, the parsimony-based method was also applied to the core
-genomes. We wished to compare the two methods in their results of the core
-genome analyses.  Because of two independent analyses on the same data set, this
-would allow us to be more confident of the results if the two analyses performed
-similarly Generally, phylogenetic methods of detecting gene transfers often take
-advantage of discrepancy between a reference species tree and gene trees. The
-two approaches that we employed in this study were no exception.
-
-\subsection{A closer study of virulence genes}
-
-{\bf [IG: CONSIDER WHAT WE WANT TO SAY HERE]}
-
-Roughly, half of the virulence genes were in the pan genome as mentioned above.
-We did not attempt to comprehensively list virulence genes that were inferred to
-be transferred horizontally or non-recombining fashion. We rather sort virulence
-genes by the number of horizontal gene transfer events to list the first several
-genes with high ranks.  Streptococcal C5a peptidase (\textit{scpB}, SDEG\_0933)
-was inferred to have six horizontal gene transfer.  Putative transposase were
-found to have a relatively large number of horizontal gene transfer events
-(SpyM3\_1016, SpyM3\_1053, SpyM3\_1054). Other genes with relatively large
-number of horizontal gene transfer events included putative response regulator
-(SpyM3\_0174), putative maltose operon transcriptional repressor (SpyM3\_0982),
-putative histidine kinase (SpyM3\_0171), and putative cell wall hydrolase
-(SpyM3\_0731, SpyM3\_0922, SpyM3\_1208).  Virulence genes encoding putative
-nucleases with a secretion signal peptide were also found to have had
-experienced horizontal gene transfer: genome-encoded streptodornase
-(SDEG\_0541), and prophage-associated deoxyribonuclease (SDEG\_1103).  Protein
-F2 like fibronectin-binding protein (SpyM3\_0104), an adhesin and invasin, was
-inferred to horizontally transferred.
-
-%\subsection{Virulence genes via recombining gene transfer}
-
-About 10\% of the genes comprised of virulence facters, about half of which were
-found in the core genome of the five genome alignment \citep{Suzuki2011}.  Using
-customized UCSC genome browser tracks available at
-\url{http://strep-genome.bscb.cornell.edu} we located virulence genes with
-notable posterior recombination probability.  These genes were by no means a
-comprehensive list of virulence factors that might have been transferred via
-homologous recombination.  We wished to showcase virulence genes using the
-result of our strudy of gene transfer.
-
-Stressful host conditions for pathogenic bacteria can include heat shock,
-starvation, and iron depletion, etc.  Iron uptake of bacteria was known to be
-essential for survival of bacteria under the hostile host condition. We found
-that parts of putative calcium transporter (MgtA) in SDE1 could have been
-transferred from SPY1 (SpyM3\_0440).  An extracellular protein, putative
-dipeptidase (SpyM3\_0465) that could be important for bacteria to scavenge
-dipeptides from their host, was inferred to be transferred from SPY to SDE: the
-whole gene was transferred, which was a case for the finding that virulence
-genes tended to be transferred as a whole unit \citep{Chan2009}.  Another
-example of whole gene transfer was superoxide dismutase that was reported to be
-associated with virulence of \textit{Streptococcus agalactiae}
-\citep{Poyart2001}: the whole gene including its 5' UTR, (SpyM3\_1071), was
-affected by recombining gene transfer from SPY to SDE.  Some of the genes were
-only affected in their parts.  The middle parts of an adhesins gene, putative
-internalin A precursor (SpyM3\_1035), was found to be affected by recombining
-gene transfer from SPY to SDE.  The middle part of a chaperonin GroEL, putative
-heat shock protein (SpyM3\_1765), was inferred to transfer from SPY to SDE1:
-increase of GroEL expression was known to allow for the bacteria to survive the
-hostile host environment of acid shock.  Parts of the gene (SpyM3\_1749),
-pyruvate formate lyase that was recently reported to have virulence
-\citep{Yesilkaya2009}, seemed to be affected by recombining gene transfer.
-Under starvation condition bacteria must use more effective energy metabolism: a
-putative glucose kinase, (SpyM3\_1180), were inferred to be transferred from SPY
-to SDE to adapt SDE's energy metabolism to survive such harsh condition.  We
-also found that an MLST locus, called DNA mismatch repair protein (MutS),
-appeared to be affected by recombining gene transfer.
-
-\subsection{UCSC genome browser tracks for bacterial gene transfer study}
+\subsection*{A closer study of virulence genes}
 
 Often computational and statistical analyses on genomes could benefit other
-researchers when those resources were provided: e.g., UCSC Genome Browser, and
-Ensembl Genome Browser. We wished to provide research community especially in
-SPY and SDE bacteria with some of the results from our analysis of gene
-transfer. We used the UCSC genome browser for the purpose.  The resource shown
-in Figure S\ref{fig:ucsc} for gene transfer is available at
-\url{http://strep-genome.bscb.cornell.edu}.  For each of the five individual
-genomes as a reference genome several annotation tracks were displayed:
-alignment blocks, recombination rate per site for each block, local gene tree
-topologies, recombination intensity, locations of virulence genes, posterior
-probability of recombining gene transfer. The genome track of probability of
-recombining gene transfer allowed us to easily browse the bacterial genome to
-locate regions that we might be interested in. 
-% To Matt: what could we show in the pan-genome? or from Mowgli's analysis?
-
-\subsection{Units of recombining gene transfer}
+researchers when those resources were provided via Genome Browser.  We wished to
+provide research community especially in SPY and SDE bacteria with results from
+our analysis of gene transfer.  The resource shown in Figure S\ref{fig:ucsc} for
+gene transfer is available at \url{http://strep-genome.bscb.cornell.edu}.  For
+each of the five individual genomes as a reference genome several annotation
+tracks were displayed: alignment blocks, recombination rate per site for each
+block, local gene tree topologies, recombination intensity, locations of
+virulence genes, posterior probability of recombining gene transfer. The genome
+track of probability of recombining gene transfer allowed us to easily browse
+the bacterial genome to locate regions that we might be interested in.  Using
+the genome browser tracks we located virulence genes with notable posterior
+recombination probability (Table S\ref{tab:virrecomb}).  About 10\% of the genes
+comprised of virulence facters, about half of which were found in the core
+genome of the five genome alignment \citep{Suzuki2011}.  By sorting virulence
+genes by the number of horizontal gene transfer events we listed several genes
+with high ranks (Table S\ref{tab:virhgt}).
+% Events counts for genes.
 
 The process of gene transfer may or may not move whole genes from one species to
-another \citep[e.g.,][]{Chan2009}. We observed that units of recombining gene
+another \citep{Chan2009}. We observed that units of recombining gene
 transfer could be multiple genes (Table S\ref{tab:genes-transfer}). In some
@@
 
-\section*{Discussion}
+\section*{Discussions}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+\subsection*{Distinguishing two gene transfer processes shows different functional
+categories associations}
+
+Bacterial gene transfer processes as an evolutionary force would affect more
+frequently in some biological functions than in others.  In our attempt of
+finding which biological functions were associated with recombining or
+horizontal gene transfers, we devised a measure of recombination intensity for
+genes by using frequency of recombinant edges that affect those genes, and used
+frequency of horizontal gene transfer events (see section \textbf{Materials and
+Methods}).  Table \ref{tab:functional} shows a list of function categories
+that were found to be associated with recombining gene transfers. We noted three
+biological functions in the list: membrane-bound molecules, aminoacyl tRNA
+synthetases, and ribosomes.
+
+Firstly, we have sugar phosphotransferases ranked
+top in the list; phosphrylation could be important in signal transduction to
+recognize environmental changes.  Transport functional categories of
+carbohydrate transport and sodium ion transport would belong to genes encoding
+membrane-bound molecules.  Also highly ranked were fatty acid and lipid
+biosyntheses that must be involved in producing membrane-bound molecules. 
+Because bacteria as a unicellular organism should be able to respond to changes
+to environment, some evolutionary pressures must have been acting on genes
+encoding membrane-bound molecules. Therefoure, recombining gene transfer process
+appeared to be more associated with biological functions that had been critical
+of sensing queues of bacterial environmental changes.
+
+Secondly, we found several aminoacyl tRNA synthetases (GO Term translation
+GO:0006412) in association with recombining gene transfer.  In the
+parsimony-based approach the term translation was also top ranked (Table
+S\ref{tab:go-events-recombining}).  Because they had been a focus of gene
+transfer in bacterial species \citep{Woese2000}, we could confirmed that the
+recombination intensity and requency of recombining gene transfer events would
+be useful in finding genes associated with recombining gene transfer.
+
+Thirdly, functional categories concerning ribosomes were unexpectedly
+found to be associated with recombining gene transfer.  Because genes related
+with ribosomes were considered to be housekeeping genes, gene transfer in which
+would be rare, it was unexpected that ribosome related genes were found to be
+associated with recombining gene transfer.  Because it was a bizarre result, we
+investigated the association by considering different recombinant edge types. We
+noted that two different classes of recombinant edge types: one that changes
+species tree topologies, and another that does not
+(see section \textbf{Materials and Methods}).
+Because recombinant edges
+that did not change species tree topology might be difficult to infer reliably,
+considering those recombinant edges in computing recombination intensity might
+lead to dubious results.  Ribosome related functional categories were found to
+be top ranked if we took those recombinant edges that did not change species
+tree topology (Tables S\ref{tab:functional-notopology} and
+S\ref{tab:functional-all}).  In Table \ref{tab:functional} ribosome related
+functional categories were not top ranked.  Therefore, high recombination
+intensity in ribosome related genes were likely to be attributed to
+topology-not-changing recombinant edges. We would be cautious of recombining
+transfer of genes related with ribosome related functions.
+
+Table \ref{tab:go-events} shows a list of functional categories associated with
+horizontal gene transfer, which were different from those from recombining gene
+transfer analysis.  Ranked top in the list was transposases, which were known to
+help foreign genes be excised from a donor genome and integrated into a
+recipient genome. The biological role of transposases was concordant with
+horizontal gene transfer process because the process involved in transposases
+would allow bacteria to acquire foreign genes without the requirement of
+homologous regions in the bacteria.  
+These differences in the profile of functional categories association of the two
+gene transfer processes were because of the approach of dicriminating the two
+gene transfer processes.
+
+% The above was moved from Results of functional category analysis.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+{\bf [IG: MOVE TO DISCUSSION:]}
+A full model-based approach would be desirable in analyzing data for testing
+relevant hypotheses more systematically. Model complexity might not allow easy
+access to the implementation of a model. Our study involved a model-based
+approach and a parsimony-based one for understanding evolution in the core
+genomes of several bacteria individuals. Because of inability of the model-based
+method for studying the pan genomes we had to resort to the parsimony-based
+method. However, the parsimony-based method was also applied to the core
+genomes. We wished to compare the two methods in their results of the core
+genome analyses.  Because of two independent analyses on the same data set, this
+would allow us to be more confident of the results if the two analyses performed
+similarly Generally, phylogenetic methods of detecting gene transfers often take
+advantage of discrepancy between a reference species tree and gene trees. The
+two approaches that we employed in this study were no exception. 
 Another contribution of our study to bacterial gene transfer research is to
@@
 be beneficial to future studies. 
+%
+% The above was moved from Results of functional category analysis.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \begin{table}
-\caption{\label{tab:functional}Functional categories asscoiated with 
-recombination intensity based on topology changing recombinant edge types.}
+\caption{\label{tab:functional}
+\bf{Functional categories asscoiated with 
+recombination intensity based on topology-altering recombinant edges.}}
 \noindent \begin{centering}
@@
 %=============================================================================
+% Figure S2: Species tree
+\begin{figure}
+\includegraphics[scale=0.7]{figures/cornellf-3-tree}
+\caption{\label{fig:tree5}The species tree of the five genomes inferred using
+\texttt{ClonalFrame}.  SPY is the branch of a clade of \textit{Streptococcus
+pyogenes}, SDE is that for \textit{S.\ dysgalactiae} ssp.\textit{\ equisimilis},
+and SD is the branch for the three individual genomes of \textit{S.\
+dysgalactiae}. The scale is proportional to the expected number of mutations
+given the Watterson's estimate of mutation rate and the tree.}
+\end{figure}
+\clearpage{}%
+%=============================================================================
+
+%=============================================================================
 % Figure 1. Heat map of recombining gene transfer from ClonalOrigin
@@
 \clearpage{}
-%=============================================================================
-
-%=============================================================================
-% FIGURE 2. - HGT Transfer heatmap
-\begin{figure}
-\begin{center}
-% \includegraphics[width=5in]{figures/trans-hgt-heatmap}
-\includegraphics[scale=0.45]{figures/heatmap-hgtedge}
-\end{center}
-\caption{\label{fig:hgt-heatmap}Heat map of horizontal gene transfers using the
-six genomes.
-Each cell of the color matrix shows the number of inferred non-recombining
-horizontal gene transfer events between the different branches of the species
-tree. Black colored cells represent no inferred events.}
-\end{figure}
-\clearpage{}%
 %=============================================================================
@@
 %=============================================================================
+% Table S. ClonalOrigin's 1st stage result.
+\begin{table}
+\caption{\label{tab:three}
+\bf{Three population parameter estimates from the model-based approach.}
+The estimates of the three parameters were based on one of two independent
+MCMC chains.  We considered the similar 
+results of the two independent chains as confirmation of 
+the convergence of the first stage of MCMC.}
+\noindent \begin{centering}
+\begin{tabular}{ccc}
+\hline
+& Replicate 1 & Replicate 2\tabularnewline
+Parameter & Estimate (IQR$^a$) & Estimate (IQR)\tabularnewline
+\hline
+$\theta^b$ & 0.081 (0.067, 0.094) & 0.081 (0.067, 0.099)\tabularnewline
+$\rho^c$ & 0.012 (0.006, 0.019) & 0.012 (0.006, 0.020)\tabularnewline
+$\delta^d$ & 744 (346, 2848) & 723 (348, 2870)\tabularnewline
+\hline
+\end{tabular}
+\par\end{centering}
+\begin{flushleft}
+$^a$ IQR stands for interquartile range.
+
+$^b$ The median of mutation rate per site weighted over the lengths of blocks.
+
+$^c$ The median of recombination rate per site weighted over the lengths of blocks.
+
+$^d$ The median value for the recombinant tract length weighted over the lengths of blocks.
+\end{flushleft}
+
+\end{table}
+%\clearpage{}
+%=============================================================================
+
+%=============================================================================
 % Table S3. Gene Tree topologies
@@
 \end{tabular}
-\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Table S6. ClonalOrigin's 1st stage result.
-\begin{table}
-\caption{\label{tab:three}Three population parameter estimates from
-the model-based approach.
-The estimates of the three parameters were based on one of two independent
-MCMC chains.  We considered the similar 
-results of the two independent chains as confirmation of 
-the convergence of the first stage of MCMC.}
-\noindent \begin{centering}
-\begin{tabular}{ccc}
-\hline
-& Replicate 1 & Replicate 2\tabularnewline
-Parameter & Estimate (IQR$^a$) & Estimate (IQR)\tabularnewline
-\hline
-$\theta^b$ & 0.081 (0.067, 0.094) & 0.081 (0.067, 0.099)\tabularnewline
-$\rho^c$ & 0.012 (0.006, 0.019) & 0.012 (0.006, 0.020)\tabularnewline
-$\delta^d$ & 744 (346, 2848) & 723 (348, 2870)\tabularnewline
-\hline
-\end{tabular}
-\par\end{centering}
-
-$^a$ IQR stands for interquartile range.
-
-$^b$ The median of mutation rate per site weighted over the lengths of blocks.
-
-$^c$ The median of recombination rate per site weighted over the lengths of blocks.
-
-$^d$ The median value for the recombinant tract length weighted over the lengths of blocks.
-
 \end{table}
@@
 
+Virulence genes encoding putative
+nucleases with a secretion signal peptide were also found to have had
+experienced horizontal gene transfer:
+
+%=============================================================================
+% Table S. 
+\begin{table}
+\caption{\label{tab:virhgt}
+\bf{Virulence genes inferred to have experienced horizontal gene transfer
+events.}}
+\noindent \begin{centering}
+\begin{tabular}{ll}
+\hline 
+Genes & Gene Accessions\tabularnewline
+\hline 
+Streptococcal C5a peptidase & SDEG\_0933\\ 
+Putative transposase & SpyM3\_1016, SpyM3\_1053, SpyM3\_1054\\
+putative response regulator & SpyM3\_0174\\
+putative maltose operon transcriptional repressor & SpyM3\_0982\\ 
+putative histidine kinase & SpyM3\_0171\\
+putative cell wall hydrolase & SpyM3\_0731, SpyM3\_0922, SpyM3\_1208\\
+genome-encoded streptodornase & SDEG\_0541\\
+prophage-associated deoxyribonuclease & SDEG\_1103\\
+Protein F2 like fibronectin-binding protein (adhesin and invasin) & SpyM3\_0104\\
+\hline 
+\end{tabular}
+\par\end{centering}
+\end{table}
+%\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% Table S. 
+\begin{table}
+\caption{\label{tab:virrecomb}
+\bf{Virulence genes inferred to have experienced recombining gene transfer
+events.}}
+\noindent \begin{centering}
+\begin{tabular}{lll}
+\hline 
+Genes & Gene Accessions\tabularnewline
+\hline 
+putative calcium transporter (MgtA) & SpyM3\_0440 & SPY1 to SDE1\\
+putative dipeptidase & SpyM3\_0465 & SPY to SDE\\
+superoxide dismutase & SpyM3\_1071 & SPY to SDE\\
+putative internalin A precursor & SpyM3\_1035 & SPY to SDE\\
+putative heat shock protein (chaperonin GroEL) & SpyM3\_1765 & SPY to SDE1\\
+pyruvate formate lyase & SpyM3\_1749 & No specific directions\\
+putative glucose kinase & SpyM3\_1180 & SPY to SDE\\
+\hline 
+\end{tabular}
+\par\end{centering}
+\end{table}
+%\clearpage{}
+%=============================================================================
+
 %=============================================================================
@@
 %=============================================================================
-% Figure S2: Species tree
-\begin{figure}
-\includegraphics[scale=0.7]{figures/cornellf-3-tree}
-\caption{\label{fig:tree5}The species tree of the five genomes inferred using
-\texttt{ClonalFrame}.  SPY is the branch of a clade of \textit{Streptococcus
-pyogenes}, SDE is that for \textit{S.\ dysgalactiae} ssp.\textit{\ equisimilis},
-and SD is the branch for the three individual genomes of \textit{S.\
-dysgalactiae}. The scale is proportional to the expected number of mutations
-given the Watterson's estimate of mutation rate and the tree.}
-\end{figure}
-\clearpage{}%
-%=============================================================================
-
-%=============================================================================
 % Figure S3: Histogram of block sizes
-\begin{figure}
-\begin{center}
-\includegraphics[width=5in]{figures/blocksize}
-\end{center}
-\caption{\label{fig:blocksize}Length distribution of the 274 core alignment blocks.}
-\end{figure}
-\clearpage{}%
+%\begin{figure}
+%\begin{center}
+%\includegraphics[width=5in]{figures/blocksize}
+%\end{center}
+%\caption{\label{fig:blocksize}Length distribution of the 274 core alignment blocks.}
+%\end{figure}
+%\clearpage{}%
 %=============================================================================
@@
 \begin{figure}
-\includegraphics[scale=0.42]{figures/mauve-analysis-001}
+\includegraphics[scale=0.42]{figures/clonalorigin}
 \caption{\label{fig:clonalorigin}Recombinant tree in the \texttt{ClonalOrigin} model.
@@
 %=============================================================================
+% FIGURE 2. - HGT Transfer heatmap
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.45]{figures/heatmap-hgtedge}
+\end{center}
+\caption{\label{fig:hgt-heatmap}Heat map of horizontal gene transfers using the
+six genomes.
+Each cell of the color matrix shows the number of inferred non-recombining
+horizontal gene transfer events between the different branches of the species
+tree. Black colored cells represent no inferred events.}
+\end{figure}
+\clearpage{}%
+%=============================================================================
+
+%=============================================================================
 % Figure S7: recombination heatmap from mowgli
@@
 \begin{center}
-% \includegraphics[width=5in]{figures/trans-recomb-heatmap}
 \includegraphics[scale=0.45]{figures/heatmap-recedge-mowgli}

### 2011-10-20 doc/manuscript.tex
I am still merging Ilan's parts and mine.

@@
 \subsection*{Recombining gene transfer from SPY to SDE at genome-wide scale}
-% \subsection*{Bayesian analysis of recombining gene transfer}
 
@@
 \cite{Didelot2010} to pursue a model-based study of recombining gene transfer. A
-preliminary step in the pipeline involved estimating the species
-tree with branch lengths (via \texttt{ClonalFrame}), and three population
-parameters of mutation rate, recombination rate, and average recombinant tract
-length (via a first phase of \texttt{ClonalOrigin}).  Because 
-\textit{S.\ equi} ssp.\textit{\ equi} strain 4047 (SEE)
-diverged significantly from the others among the six genomes (Table S\ref{tab:genome}), we
-could not obtain sufficiently long multiple sequence alignments with
-it, and had to exclude it from the model-based analysis.  The estimated species tree
-using the alignment of the five genomes (Figure
-S\ref{fig:mauve}) had the expected topology \texttt{The figure 1 must be nicer
-than one that now.} (see Figure \ref{fig:tree5}). \texttt{ClonalFrame} also
-estimated
-a relative strength of recombination compared to mutations ($r/m=4.66$), 
-indicating potentially significant roles of recombination in the evolutionary
-process of shaping these genomes; four times more sites had been affected by
-recombination-driven substitutions than by mutation-driven substitutions.
-($r/m$; Table S\ref{tab:clonalframe}).
-
-Estimates of population mutation rate ($\theta$), recombination rate ($\rho$),
-and average tract length ($\delta$) were obtained for the alignment blocks
-analyzed (Figure S\ref{fig:scatter3}), and 
-medians of each of the three parameters weighted over block lengths were
-obtained as global estimates of the three population
-parameters (Table S\ref{tab:three} and see \textbf{Materials and Methods}).  
-The local estimates typically seemed to be
+preliminary step in the pipeline involved estimating the species tree with
+branch lengths (via \texttt{ClonalFrame}), and three population parameters of
+mutation rate, recombination rate, and average recombinant tract length (via a
+first phase of \texttt{ClonalOrigin}).  Because \textit{S.\ equi} ssp.\textit{\
+equi} strain 4047 (SEE) diverged significantly from the others among the six
+genomes (Table S\ref{tab:genome}), we could not obtain sufficiently long
+multiple sequence alignments with it, and had to exclude it from the model-based
+analysis.  The estimated species tree using the alignment of the five genomes
+(Figure S\ref{fig:mauve}) had the expected topology \texttt{The figure 1 must be
+nicer than one that now} (see Figure \ref{fig:tree5}). \texttt{ClonalFrame} also
+estimated a relative strength of recombination compared to mutations
+($r/m=4.66$), indicating potentially significant roles of recombination in the
+evolutionary process of shaping these genomes; four times more sites had been
+affected by recombination-driven substitutions than by mutation-driven
+substitutions.  (Table S\ref{tab:clonalframe}).  Estimates of population
+mutation rate ($\theta$), recombination rate ($\rho$), and average tract length
+($\delta$) were obtained for the alignment blocks analyzed (Figure
+S\ref{fig:scatter3}), and the median weighted over block lengths for each of the
+three parameters was obtained as a global estimate (Table S\ref{tab:three} and
+see \textbf{Materials and Methods}).  The local estimates typically seemed to be
 moderately fluctuating around the global average, with the exception of what
@@
 
-The species tree and estimated branch lengths, together with the global
-estimates of the three population parameters were used in the second phase of
-ClonalOrigin, where recombinant trees were sampled from the posterior
-distribution. One way to summarize the posterior sample of recombinant trees
-provided by ClonalOrigin was to observe the topologies of the local genealogies
-induced by it. Out of the 105 possible rooted labeled topologies spanning five
-taxa, the dominant topology (sampled 66.7\% of the time) was that of the species
-tree (Materials and Methods; Table S\ref{tab:Gene-tree-topologies}). The second
-most frequently sampled topology (sampled 9.3\% of the time) was the one in
-which SPY and SDE were sister clades (with SDD as an outgroup). Other topologies
-were sampled with frequency less than 4\%. This gave initial indication to the
-extent of recombining gene transfer  between SPY and SDE.
-
-To gain a more fine-scale view of the main types of recombination scenarios
-sampled by ClonalOrigin, we looked at the recombinant edges present in the
-sampled recombinant trees. Following the approach of Didelot et al. \cite{Didelot2010}, we
+The estimated species tree together with the global estimates of the three
+population parameters were used in the second phase of \texttt{ClonalOrigin},
+where recombinant trees were sampled from the posterior distribution. One way to
+summarize the posterior sample of recombinant trees was to count the topologies
+of the local genealogies induced by it. Of the 105 possible rooted topologies
+relatig five taxa, the dominant topology (66.7\%) was that of the species tree
+(see \textbf{Materials and Methods} and Table S\ref{tab:Gene-tree-topologies}).
+The second most frequently sampled topology (9.3\%) was the one in which SPY and
+SDE were sister clades to each other with SDD being as an outgroup. Other
+topologies were sampled with frequency less than 4\%. This gave initial
+indication to the extent of recombining gene transfer  between SPY and SDE.
+
+To gain a more fine-scale view of the main types of recombination scenarios, we
+looked at the recombinant edges present in the posterior sample of recombinant
+trees. Following the approach of Didelot et al. \cite{Didelot2010}, we
 classified these edges by types defined by distinct pairs of donor and recipient
-branches of the species tree (Figure \ref{fig:tree5}). For each edge type, we recorded
-the average number of edges of that type present in the sampled recombinant
-tree, and computed the ratio of this observed count and the prior expected count
-as implied by the species tree (considering branch lengths and their relative
-positions) and the three
-population parameters estimated in the preliminary round of the analysis. The
-base 2 logarithms of these ratios are depicted in the heatmap of Figure
-\ref{fig:Heatmap-of-recombination} (see also Tables
-S\ref{tab:heatmap} and S\ref{tab:obsheatmap}).
+branches of the species tree (Figure \ref{fig:tree5}). For each edge type, we
+recorded the average number of edges of that type present in the sampled
+recombinant tree, and computed the ratio of this observed frequency to the prior
+expected one as implied by the species tree (considering branch lengths and
+their relative positions) and the three population parameters estimated in the
+preliminary step of the model-based analysis.  Base 2 logarithms of these ratios
+are depicted in the heatmap of Figure \ref{fig:Heatmap-of-recombination} (see
+also Tables S\ref{tab:heatmap} and S\ref{tab:obsheatmap}).
 
 Among the possible recombinant edge types, we distinguished between ones that
-preserve the species tree topology, and ones that alter it. We focused
-on the topology-altering types, both because our main objective was to study
-cross-clade gene transfer, and because our simulation study indicates that the
-observed counts of topology-preserving recombinant edges were 
-unreliably inferred {\bf[IG: POINT TO RELEVANT SECTION]}. The heatmap allowed to find
-recombination scenarios that deviated from the expectation dictated by the
-probabilistic model of ClonalOrigin. {\bf [IG: CHECK THIS FOLLOWING LINE OF
-ARGUMENT]} We envisioned two main potential causes for such a deviation: (1)
-reduced or increased opportunities for species to interact and exchange genetic
+preserve the species tree topology, and ones that alter it. We focused on the
+topology-altering types, both because our main objective was to study
+cross-clade gene transfer, and because our simulation study indicated that the
+observed counts of topology-preserving recombinant edges were unreliably
+inferred (see \textbf{Simulations}). The heatmap allowed to find recombination
+scenarios that deviated from the expectation dictated by the probabilistic
+model.  We envisioned two main potential causes for such deviations: (1) reduced
+or increased opportunities for species to interact and exchange genetic
 material; (2) selection acting on newly transferred genes.
@@
 We observed a reduced count of recombinant edges sampled between SDD and the
-three branches of the SPY clade ($\sim$25\% of the expected count). This is
-consistent with the fact that these species have reduced opportunities to
-interact with each other, SPY being a human pathogen and SDD being strictly veterinary. On the
-contrary, the number of sampled recombinant edges between SDD and branches in
-the SDE clade seemed to be consistent with the prior, with a slight excess for
-edges from SDE1 and SDE2 to SDD {\bf [IG: IS THIS WORTH DWELLNG ON?]}. The
-greatest deviation of observed counts from expected ones was observed for
-recombinant edges between the SPY and SDE clades. It was concordant with the
-fact that these two species share an ecological niche; SDE was an
-opportunistic human pathogen. Interestingly, there appeared to be much higher
-enrichment (up to four times more than expected) for recombinant edges
-originating in the SPY clade and ending up in the SDE clade. In a follow-up
-simulation study we show that this deviation from the expected count was indeed
-highly significant, and so was the observed asymmetry in pattern of recombination
-{\bf[IG: POINT TO RELEVANT SECTION]}. This bias in directionality of
-recombination pattern between SPY and SDE was especially intriguing in light of
-the hypothesis that SDE acquired human pathogenicity via gene transfer from SPY.
+three branches of the SPY clade ($\sim$25\% of the expected count). This was
+consistent with the fact that these species had few opportunities to interact
+with each other, SPY being a human pathogen and SDD being strictly veterinary.
+On the contrary, the number of sampled recombinant edges between SDD and
+branches in the SDE clade seemed to be consistent with the prior, with a slight
+excess for edges from SDE1 and SDE2 to SDD.  The greatest deviation of observed
+counts from expected ones was observed for recombinant edges between the SPY and
+SDE clades. It was concordant with the fact that these two species shared an
+ecological niche; SDE was an opportunistic human pathogen. Interestingly, there
+appeared to be much higher enrichment (up to four times more than expected) for
+recombinant edges originating in the SPY clade and ending up in the SDE clade.
+In a follow-up simulation study we show that this deviation from the expected
+count was indeed highly significant, and so was the observed asymmetry in
+pattern of recombination (see \textbf{Simulations}).  This net directionality of
+recombination pattern from SPY to SDE was especially intriguing in light of the
+hypothesis that SDE acquired human pathogenicity via gene transfer from SPY.
 
@@
 
-{\bf[IG: DID NOT TOUCH THE FIRST 3 PARAGRAPHS. REQUIRES ``CONDENSING".]}
-{\bf[IG: START WITH SOME GENERAL STATISTICS - NUMBER OF GENES, ETC?]}
-\texttt{SCC: Check this ``condensed'' version.}
-
-Figure \ref{fig:Gene-duplication-loss} shows gene gain, loss, duplication, and
-two different gene transfer events along the lineages of the phylogenetic tree
-that relates the six genomes. SPY had less genes than the other species. All of
-the genomes were used to construct 2314 gene families.  The largest family
-consisted of 59 genes while nearly half of gene families (1066) had exactly one
-gene per species, or six genes total (Figure S\ref{fig:famsizes}).  We estimated
-a larger number of recombining gene transfer events than that of horizontal gene
-transfer events; 944 and 294, respectively.  Duplications in extant lineages
-were more frequent than those in the other lineages.  SDE lineages from the root
-appeared to have increased in the number of genes while the common ancestor of
-SPY1 and SPY2 decreased in the number of genes.  Whereas gene loss events were
-either comparable to or less than gene gain ones along most of the lineages, it
-was not along two internal branches of SPY and SDE.  This indicated that SPY and
-SDE lineages had reduced number of genes to specialize in their specific niches
-after each split from their corresponding direct ancestors.
-
-As in the Bayesian analysis of ClonalOrigin, we attempted to see which type of
-horizontal or recombining transfer events was most common in our inference.  We
-classified events into types defined by donor-recipient branches in the species
-tree. We recorded the number of inferred events of each type in Figures
-S\ref{fig:hgt-heatmap} and S\ref{fig:mowgli-recomb-heatmap}. Unlike in the
-Bayesian analysis, we could not normalize these counts by their expectation
-because we did not have any model-based expectation for them. This restricted
+Figure \ref{fig:Gene-duplication-loss} shows the inferred numbers of gene gain,
+loss, duplication, and two different kinds of gene transfer events along
+lineages of the tree that relates the six genomes. SPY had less genes than the
+other species. All of the genomes were used to construct 2314 gene families.
+The largest family consisted of 59 genes while nearly half of gene families
+(1066) had exactly one gene per species, or six genes total (Figure
+S\ref{fig:famsizes}).  We estimated a larger number of recombining gene transfer
+events than that of horizontal gene transfer events; 944 and 294, respectively.
+Duplications in extant lineages were more frequent than those in the other
+lineages.  SDE lineages from the root appeared to have increased in the number
+of genes while the common ancestor of SPY1 and SPY2 decreased in the number of
+genes.  Whereas gene loss events were either comparable to or less than gene
+gain ones along most of the lineages, it was not the case along the two internal
+branches of SPY and SDE.  This indicated that SPY and SDE lineages had reduced
+number of genes to specialize in their specific niches after each divergence
+from their corresponding direct ancestors.
+
+As in the model-based analysis, we attempted to see which type of horizontal or
+recombining transfer events was most common in our inference.  We classified
+events into types defined by donor-recipient branches in the species tree. We
+recorded the number of inferred events of each type (Figures
+S\ref{fig:hgt-heatmap} and S\ref{fig:mowgli-recomb-heatmap}). Unlike in the
+model-based analysis, we could not normalize these counts by their expectation
+because we did not have any model-based expectation for them, which restricted
 our ability to interprest these counts. We did notice, however, a large number
 of inferred recombining gene transfer events from the ancestral SPY branch to
-the ancestral SDE branch, consistent with our observations from the Bayesian
-analysis. With horizontal gene transfer, we did not observed a similar pattern,
-but we did observe a large number of horizontal transfers from the ancesral SDE
-branch to SDD.
+the ancestral SDE branch, consistent with our observations from the model-based
+analysis. Yet, we did not observed a similar pattern with horizontal gene
+transfer while we observed a large number of horizontal transfers from the
+ancesral SDE branch to SDD.
 
@@
 We set out to seek categories of genes that were significantly enriched for
-recombining or horizontal transfer events. For horizontal
-transfer, we used a score based on the number of events inferred by the
-parsinomy-based method, and for recombining gene transfer, we used both the
-parsimony-based inferred counts and a recombination intensity score based on the
-Bayesian analysis (see \textbf{Materials and Methods}). First, we checked association of
-transfers with virulence genes, which were provided from a previous study
-of Suzuki et al.\ \citet{Suzuki2011}.
-This set of virulence genes was found to be significantly associated
-with horizontal gene transfers (P-value $<1.33 \times 10^{-11}$), but not with
-recombining gene transfers. We also found a marginally
-significant association between virulence genes and gene duplications (P-value
-$<7.13 \times 10^{-3}$).
+recombining or horizontal transfer events. For horizontal transfer, we used a
+score based on a number of events inferred by the parsimony-based method, and
+for recombining gene transfer, we used both the parsimony-based inferred counts
+and a recombination intensity score based on the model-based analysis (see
+\textbf{Materials and Methods}). First, we checked association of transfers with
+virulence genes, which were provided from a previous study of Suzuki et al.\
+\citet{Suzuki2011}.  This set of virulence genes was found to be significantly
+associated with horizontal gene transfers (P-value $<1.33 \times 10^{-11}$), but
+not with recombining gene transfers. We also found a marginally significant
+association between virulence genes and gene duplications (P-value $<7.13 \times
+10^{-3}$).
 
@@
 
-\subsection*{Comparison of Bayesian and parsimony-based inferences of recombining gene transfer}
+\subsection*{Comparison of model-based and parsimony-based inferences of recombining gene transfer}
 
 The similarity between the two lists of gene categories found to be associated
-with recombining gene transfers in the separate analyses -- Bayesian and
+with recombining gene transfers in the separate analyses -- model-based and
 parsimony-based --  indirectly indicated concordance between the two independent
@@
 and the number of horizontal gene transfers was 0.03 (P-value 0.18;
-S\ref{fig:cmpcomowglihgt}).
+Figure S\ref{fig:cmpcomowglihgt}).
 
@@
 our analysis of gene transfer.  The resource shown in Figure S\ref{fig:ucsc} for
-gene transfer is available at \url{http://strep-genome.bscb.cornell.edu}.  For
+gene transfer was available at \url{http://strep-genome.bscb.cornell.edu}.  For
 each of the five individual genomes as a reference genome several annotation
@@
 with high ranks (Table S\ref{tab:virhgt}).
-% Events counts for genes.
-
-The process of gene transfer may or may not move whole genes from one species to
-another \citep{Chan2009}. We observed that units of recombining gene
-transfer could be multiple genes (Table S\ref{tab:genes-transfer}). In some
-cases, parts of a gene were inferred to have been transferred, for instance,
-transfer from SPY1 to SDE1 was inferred for 29\% of the aspartyl-tRNA synthetase
-gene. Recombinant tracts often were shown to span two genes.  In other cases,
-the inferred transfer units spaned a set of as much as eight adjacent genes:
-loci from SpyM3\_1518 to SpyM3\_1525.
-
-\subsection{Simulations}
+The whole gene of  putative dipeptidase (SpyM3\_0465) was transferred, which was
+a case for the finding that virulence genes tended to be transferred as a whole
+unit \citep{Chan2009}.  Another example of whole gene transfer was superoxide
+dismutase (SpyM3\_1071).  We investigated the unit of recombining gene transfer
+by finding 58 genes for which a recombinant edge of certain types were sampled
+with probability at least 0.9 by focusing on specific types of recombinant edges
+(Table S\ref{tab:genes-transfer}).  In some cases, parts of a gene were inferred
+to have been transferred, for instance, transfer from SPY1 to SDE1 was inferred
+for 29\% of the aspartyl-tRNA synthetase gene. Recombinant tracts often were
+shown to span two genes.  In other cases, the inferred transfer units spaned a
+set of as much as eight adjacent genes: loci from SpyM3\_1518 to SpyM3\_1525.
+
+\subsection*{Simulations}
 
@@
 
-\section*{Discussions}
+\section*{Discussion}
 
@@
 another facet of evolutionary forces.  Because the two gene transfer processes
-would act on bacterial genomes simultaneously, ignoring one of the two processes
+would act on closely related bacterial genomes simultaneously, ignoring one of the two processes
 in studying evolution of bacteria would lead to a narrowed point of view to the
 underlying evolutionary forces. Our study of SPY and SDE genomes undertook the
-work of discriminating recombining and non-recombining horizontal gene transfer
+work of discriminating recombining and horizontal gene transfer
 processes because the two species were very closely related.  We showed that the
-two gene transfer processes acted on different biological functions. Recombining
+two gene transfer processes acted on different biological functions. 
+Recombining
 gene transfer process in the analysis of SPY and SDE genomes was related with
@@
 human pathogen, we were interested in association of pathogenicity with the two
-gene transfer processes. We showed that virulence genes were more attributed to
-horizontal gene transfer process than to recombining gene transfer process.
+gene transfer processes. We showed that virulence genes were more associated
+with 
+horizontal gene transfer process than with recombining gene transfer process.
 With the current data of only SPY and SDE genomes we could not conclude that the
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\subsection*{Distinguishing two gene transfer processes shows different functional
-categories associations}
-
 Bacterial gene transfer processes as an evolutionary force would affect more
@@
 genes by using frequency of recombinant edges that affect those genes, and used
-frequency of horizontal gene transfer events (see section \textbf{Materials and
-Methods}).  Table \ref{tab:functional} shows a list of function categories
-that were found to be associated with recombining gene transfers. We noted three
-biological functions in the list: membrane-bound molecules, aminoacyl tRNA
-synthetases, and ribosomes.
-
-Firstly, we have sugar phosphotransferases ranked
-top in the list; phosphrylation could be important in signal transduction to
-recognize environmental changes.  Transport functional categories of
-carbohydrate transport and sodium ion transport would belong to genes encoding
-membrane-bound molecules.  Also highly ranked were fatty acid and lipid
-biosyntheses that must be involved in producing membrane-bound molecules. 
-Because bacteria as a unicellular organism should be able to respond to changes
-to environment, some evolutionary pressures must have been acting on genes
-encoding membrane-bound molecules. Therefoure, recombining gene transfer process
-appeared to be more associated with biological functions that had been critical
-of sensing queues of bacterial environmental changes.
-
-Secondly, we found several aminoacyl tRNA synthetases (GO Term translation
-GO:0006412) in association with recombining gene transfer.  In the
-parsimony-based approach the term translation was also top ranked (Table
-S\ref{tab:go-events-recombining}).  Because they had been a focus of gene
-transfer in bacterial species \citep{Woese2000}, we could confirmed that the
+frequency of horizontal gene transfer events.  This approach allowed us to be
+able to associate gene transfer with biological functions including
+membrane-bound molecules, aminoacyl tRNA synthetases, and ribosomes.  Sugar
+phosphotransferases could be important in signal transduction to recognize
+environmental changes.  Transporters, and fatty acid and lipid biosyntheses
+would be involved in membrane-bound molecules.  Because bacteria as a
+unicellular organism should be able to respond to changes to environment, some
+evolutionary pressures must have been acting on genes encoding membrane-bound
+molecules. Therefoure, recombining gene transfer process appeared to be more
+associated with biological functions that had been critical of sensing queues of
+bacterial environmental changes.  Because several aminoacyl tRNA synthetases
+belonging to gene ontology term of translation (GO:0006412) had been a focus of
+gene transfer in bacterial species \citep{Woese2000}, we could confirm that the
 recombination intensity and requency of recombining gene transfer events would
 be useful in finding genes associated with recombining gene transfer.
-
-Thirdly, functional categories concerning ribosomes were unexpectedly
-found to be associated with recombining gene transfer.  Because genes related
-with ribosomes were considered to be housekeeping genes, gene transfer in which
-would be rare, it was unexpected that ribosome related genes were found to be
-associated with recombining gene transfer.  Because it was a bizarre result, we
-investigated the association by considering different recombinant edge types. We
-noted that two different classes of recombinant edge types: one that changes
-species tree topologies, and another that does not
-(see section \textbf{Materials and Methods}).
-Because recombinant edges
-that did not change species tree topology might be difficult to infer reliably,
-considering those recombinant edges in computing recombination intensity might
-lead to dubious results.  Ribosome related functional categories were found to
-be top ranked if we took those recombinant edges that did not change species
-tree topology (Tables S\ref{tab:functional-notopology} and
-S\ref{tab:functional-all}).  In Table \ref{tab:functional} ribosome related
-functional categories were not top ranked.  Therefore, high recombination
-intensity in ribosome related genes were likely to be attributed to
-topology-not-changing recombinant edges. We would be cautious of recombining
-transfer of genes related with ribosome related functions.
-
-Table \ref{tab:go-events} shows a list of functional categories associated with
-horizontal gene transfer, which were different from those from recombining gene
-transfer analysis.  Ranked top in the list was transposases, which were known to
-help foreign genes be excised from a donor genome and integrated into a
-recipient genome. The biological role of transposases was concordant with
-horizontal gene transfer process because the process involved in transposases
-would allow bacteria to acquire foreign genes without the requirement of
-homologous regions in the bacteria.  
-These differences in the profile of functional categories association of the two
-gene transfer processes were because of the approach of dicriminating the two
-gene transfer processes.
-
-% The above was moved from Results of functional category analysis.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-{\bf [IG: MOVE TO DISCUSSION:]}
-A full model-based approach would be desirable in analyzing data for testing
-relevant hypotheses more systematically. Model complexity might not allow easy
-access to the implementation of a model. Our study involved a model-based
-approach and a parsimony-based one for understanding evolution in the core
-genomes of several bacteria individuals. Because of inability of the model-based
-method for studying the pan genomes we had to resort to the parsimony-based
-method. However, the parsimony-based method was also applied to the core
-genomes. We wished to compare the two methods in their results of the core
-genome analyses.  Because of two independent analyses on the same data set, this
-would allow us to be more confident of the results if the two analyses performed
-similarly Generally, phylogenetic methods of detecting gene transfers often take
-advantage of discrepancy between a reference species tree and gene trees. The
-two approaches that we employed in this study were no exception. 
+The biological role of transposases was concordant with horizontal gene transfer
+process because the process involved in transposases would allow bacteria to
+acquire foreign genes without the requirement of homologous regions in the
+bacteria.  These differences in the profile of functional categories association
+of the two gene transfer processes were because of the approach of dicriminating
+the two gene transfer processes.
+
+Functional categories concerning ribosomes were unexpectedly found to be
+associated with recombining gene transfer.  Because genes related with ribosomes
+were considered to be housekeeping genes, gene transfer in which would be rare,
+it was unexpected that ribosome related genes were found to be associated with
+recombining gene transfer.  Because it was a bizarre result, we investigated the
+association by considering different recombinant edge types.  Ribosome related
+functional categories were found to be top ranked if we took
+tree-topology-preserving recombinant edges into account (Tables
+S\ref{tab:functional-notopology} and S\ref{tab:functional-all}).  whereas
+ribosome related functional categories were not top ranked in Table
+\ref{tab:functional}.  Therefore, high recombination intensity in ribosome
+related genes were likely to be attributed to topology-preserving recombinant
+edges that were found to be difficult to infer reliably. So, we would have to be
+cautious of associating ribosome related functions with recombining gene
+transfer.
+
 Another contribution of our study to bacterial gene transfer research is to
@@
 be beneficial to future studies.
-%
-% The above was moved from Results of functional category analysis.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \textit{parC}, in which we also observed moderate recombination.
+
+A full model-based approach would be desirable in analyzing data for testing
+relevant hypotheses more systematically. Model complexity might not allow easy
+access to the implementation of a model. Our study involved a model-based
+approach and a parsimony-based one for understanding evolution in the core
+genomes of several bacteria individuals. Because of inability of the model-based
+method for studying the pan genomes we had to resort to the parsimony-based
+method. However, the parsimony-based method was also applied to the core
+genomes. We wished to compare the two methods in their results of the core
+genome analyses.  Because of two independent analyses on the same data set, this
+would allow us to be more confident of the results if the two analyses performed
+similarly Generally, phylogenetic methods of detecting gene transfers often take
+advantage of discrepancy between a reference species tree and gene trees. The
+two approaches that we employed in this study were no exception.
 
@@
 \end{tabular}
+\end{table}
+%\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% Table S. Mowgli - Recombining Transfer
+\begin{table}
+\caption{\label{tab:go-events-recombining}
+\bf{Functional categories enriched in 
+recombining transfer events that are inferred using the parsimony-based
+approach.}}
+\noindent \begin{centering}
+\begin{tabular}{cccl}
+\hline 
+P-value & Count & GO-term & Description \\
+\hline 
+7e-09 & 153 & GO:0006412 & translation\\
+1e-08 &  67 & GO:0030529 & ribonucleoprotein complex\\
+1e-07 &  45 & GO:0019843 & rRNA binding\\
+2e-06 &  73 & GO:0003735 & structural constituent of ribosome\\
+3e-05 & 157 & GO:0005840 & ribosome\\
+7e-05 & 221 & GO:0003723 & RNA binding\\
+2e-04 &  11 & GO:0004826 & phenylalanine-tRNA ligase activity\\
+3e-04 &  14 & GO:0015413 & nickel-transporting ATPase activity\\
+4e-04 &  93 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
+5e-04 & 104 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
+\hline 
+\end{tabular}
+\par\end{centering}
+\end{table}
+%\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% Table S. 
+\begin{table}
+\caption{\label{tab:virrecomb}
+\bf{Virulence genes inferred to have experienced recombining gene transfer
+events.}}
+\noindent \begin{centering}
+\begin{tabular}{lll}
+\hline 
+Genes & Gene Accessions\tabularnewline
+\hline 
+putative calcium transporter (MgtA) & SpyM3\_0440 & SPY1 to SDE1\\
+putative dipeptidase & SpyM3\_0465 & SPY to SDE\\
+superoxide dismutase & SpyM3\_1071 & SPY to SDE\\
+putative internalin A precursor & SpyM3\_1035 & SPY to SDE\\
+putative heat shock protein (chaperonin GroEL) & SpyM3\_1765 & SPY to SDE1\\
+pyruvate formate lyase & SpyM3\_1749 & No specific directions\\
+putative glucose kinase & SpyM3\_1180 & SPY to SDE\\
+\hline 
+\end{tabular}
+\par\end{centering}
+\end{table}
+%\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% Table S. 
+\begin{table}
+\caption{\label{tab:virhgt}
+\bf{Virulence genes inferred to have experienced horizontal gene transfer
+events.}}
+\noindent \begin{centering}
+\begin{tabular}{ll}
+\hline 
+Genes & Gene Accessions\tabularnewline
+\hline 
+Streptococcal C5a peptidase & SDEG\_0933\\ 
+Putative transposase & SpyM3\_1016, SpyM3\_1053, SpyM3\_1054\\
+putative response regulator & SpyM3\_0174\\
+putative maltose operon transcriptional repressor & SpyM3\_0982\\ 
+putative histidine kinase & SpyM3\_0171\\
+putative cell wall hydrolase & SpyM3\_0731, SpyM3\_0922, SpyM3\_1208\\
+genome-encoded streptodornase & SDEG\_0541\\
+prophage-associated deoxyribonuclease & SDEG\_1103\\
+Protein F2 like fibronectin-binding protein (adhesin and invasin) & SpyM3\_0104\\
+\hline 
+\end{tabular}
+\par\end{centering}
 \end{table}
@@
 %=============================================================================
-% Table S. Mowgli - Recombining Transfer
-\begin{table}
-\caption{\label{tab:go-events-recombining}Functional categories enriched in 
-recombining transfer events that are inferred using the parsimony-based approach.}
-\noindent \begin{centering}
-\begin{tabular}{cccl}
-\hline 
-P-value & Count & GO-term & Description \\
-\hline 
-7e-09 & 153 & GO:0006412 & translation\\
-1e-08 &  67 & GO:0030529 & ribonucleoprotein complex\\
-1e-07 &  45 & GO:0019843 & rRNA binding\\
-2e-06 &  73 & GO:0003735 & structural constituent of ribosome\\
-3e-05 & 157 & GO:0005840 & ribosome\\
-7e-05 & 221 & GO:0003723 & RNA binding\\
-2e-04 &  11 & GO:0004826 & phenylalanine-tRNA ligase activity\\
-3e-04 &  14 & GO:0015413 & nickel-transporting ATPase activity\\
-4e-04 &  93 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
-5e-04 & 104 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
-\hline 
-\end{tabular}
-\par\end{centering}
-\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
 % Table S. Simulation study for the three population parameters.
@@
 0.00052925 & 26 & GO:0019843 & rRNA binding\tabularnewline
-\hline 
-\end{tabular}
-\par\end{centering}
-\end{table}
-%\clearpage{}
-%=============================================================================
-
-Virulence genes encoding putative
-nucleases with a secretion signal peptide were also found to have had
-experienced horizontal gene transfer:
-
-%=============================================================================
-% Table S. 
-\begin{table}
-\caption{\label{tab:virhgt}
-\bf{Virulence genes inferred to have experienced horizontal gene transfer
-events.}}
-\noindent \begin{centering}
-\begin{tabular}{ll}
-\hline 
-Genes & Gene Accessions\tabularnewline
-\hline 
-Streptococcal C5a peptidase & SDEG\_0933\\ 
-Putative transposase & SpyM3\_1016, SpyM3\_1053, SpyM3\_1054\\
-putative response regulator & SpyM3\_0174\\
-putative maltose operon transcriptional repressor & SpyM3\_0982\\ 
-putative histidine kinase & SpyM3\_0171\\
-putative cell wall hydrolase & SpyM3\_0731, SpyM3\_0922, SpyM3\_1208\\
-genome-encoded streptodornase & SDEG\_0541\\
-prophage-associated deoxyribonuclease & SDEG\_1103\\
-Protein F2 like fibronectin-binding protein (adhesin and invasin) & SpyM3\_0104\\
-\hline 
-\end{tabular}
-\par\end{centering}
-\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Table S. 
-\begin{table}
-\caption{\label{tab:virrecomb}
-\bf{Virulence genes inferred to have experienced recombining gene transfer
-events.}}
-\noindent \begin{centering}
-\begin{tabular}{lll}
-\hline 
-Genes & Gene Accessions\tabularnewline
-\hline 
-putative calcium transporter (MgtA) & SpyM3\_0440 & SPY1 to SDE1\\
-putative dipeptidase & SpyM3\_0465 & SPY to SDE\\
-superoxide dismutase & SpyM3\_1071 & SPY to SDE\\
-putative internalin A precursor & SpyM3\_1035 & SPY to SDE\\
-putative heat shock protein (chaperonin GroEL) & SpyM3\_1765 & SPY to SDE1\\
-pyruvate formate lyase & SpyM3\_1749 & No specific directions\\
-putative glucose kinase & SpyM3\_1180 & SPY to SDE\\
 \hline 
@@
 %=============================================================================
+% Figure S8: gene family sizes
+% FIGURE - gene family sizes
+\begin{figure}
+\begin{center}
+\includegraphics[width=5in]{figures/famsizes}
+\end{center}
+\caption{\label{fig:famsizes}Distribution of gene family sizes.}
+\end{figure}
+\clearpage{}
+%=============================================================================
+
+%=============================================================================
 % FIGURE 2. - HGT Transfer heatmap
@@
 tree. Black colored cells represent no inferred events.}
-\end{figure}
-\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Figure S8: gene family sizes
-% FIGURE - gene family sizes
-\begin{figure}
-\begin{center}
-\includegraphics[width=5in]{figures/famsizes}
-\end{center}
-\caption{\label{fig:famsizes}Distribution of gene family sizes.}
 \end{figure}

### 2011-10-20 doc/manuscript.tex
I am still editting the Results sectioin with Ilan.

@@
 %% ** EDIT HERE **
+\markright{Gene Transfer using Streptococcus Genomes}
 
@@
 
-\section*{Things To Do for submitting this to PLoS Pathogens}
-\begin{enumerate}
-\item A paragraph for briefing the methods of ClonalOrigin and Mowgli before
-section \textbf{Results}.
-\item We can add at least 50 words in section \textbf{Abstract}.
-\item Figures, tables, and supporting information need to be ready.
-\item Publish source codes at public domain.
-\end{enumerate}
-
-\section*{A list of our major and minor findings or observations}
-\begin{enumerate}
-\item We showed that recombining gene transfer from SPY to SDE was strong
-enough.
-\item A few evidences for the net directionality from SPY to SDE:
-time-constrained gene transfer between SPY to SDE also showed the directionality
-from SPY to SDE.
-\item We showed that virulence genes were associated with horizontal gene
-transfer. 
-\item We found that recombining gene tranfer processes are involved in
-membrane-bound protein encoding genes and aminoacyl tRNA synthetases.
-\item We found that horizontal gene transfer processes are involved in
-transposases.
-\item We provided evidence that recombination process affected more sites along
-the core genome of SPY and SDE than mutation process.
-\item We found a recombination hotspot at 200 kb from the origin of replication
-along the genome of SDE1.
-\item We showed that species-tree-topology-changing recombinant edges were more
-reliably recovered than species-tree-topology-not-changing edges.
-\item We showed that recombination intensity might be useful in finding genes
-under high recombining gene transfer.
-\item We found that SPY and SDE are more closely related than SDD and SDE in
-about 10\% of the core genome.
-\item Recombination intensity and number of recombining gene transfer events
-were positively correlated.
-\item We found reduction of numbers of genes after divergence of SDE and SPY.
-\item We confirmed the species tree that relates the five genomes.
-\item We listed several virulence genes associated with recombining and
-horizontal gene transfer.
-\end{enumerate}
-
-\section*{Suggestions about the manuscript}
-\begin{enumerate}
-\item Ilan thought that a species tree must be shown at the outset of the
-section \textbf{Results}.
-\item Can selections after gene transfer be tested?
-\end{enumerate}
+% \section*{Things To Do for submitting this to PLoS Pathogens}
+% \begin{enumerate}
+% \item Legends for the main figures as a separate section
+% \item We can add at least 50 words in section \textbf{Abstract}.
+% \item Figures, tables, and supporting information need to be ready.
+% \item Merge Genome Browser tracks to the official one.
+% \item Publish source codes at public domain.
+% \end{enumerate}
+%
+% \section*{A list of our major and minor findings or observations}
+% \begin{enumerate}
+% \item We showed that recombining gene transfer from SPY to SDE was strong
+% enough.
+% \item A few evidences for the net directionality from SPY to SDE:
+% time-constrained gene transfer between SPY to SDE also showed the directionality
+% from SPY to SDE.
+% \item We showed that virulence genes were associated with horizontal gene
+% transfer. 
+% \item We found that recombining gene tranfer processes are involved in
+% membrane-bound protein encoding genes and aminoacyl tRNA synthetases.
+% \item We found that horizontal gene transfer processes are involved in
+% transposases.
+% \item We provided evidence that recombination process affected more sites along
+% the core genome of SPY and SDE than mutation process.
+% \item We found a recombination hotspot at 200 kb from the origin of replication
+% along the genome of SDE1.
+% \item We showed that species-tree-topology-changing recombinant edges were more
+% reliably recovered than species-tree-topology-not-changing edges.
+% \item We showed that recombination intensity might be useful in finding genes
+% under high recombining gene transfer.
+% \item We found that SPY and SDE are more closely related than SDD and SDE in
+% about 10\% of the core genome.
+% \item Recombination intensity and number of recombining gene transfer events
+% were positively correlated.
+% \item We found reduction of numbers of genes after divergence of SDE and SPY.
+% \item We confirmed the species tree that relates the five genomes.
+% \item We listed several virulence genes associated with recombining and
+% horizontal gene transfer.
+% \end{enumerate}
+% 
+% \section*{Suggestions about the manuscript}
+% \begin{enumerate}
+% \item Ilan thought that a species tree must be shown at the outset of the
+% section \textbf{Results}.
+% \item Can selections after gene transfer be tested?
+% \end{enumerate}
 
@@
 
-\texttt{We show a net directionality of recombining gene transfers from SPY to SDE. 
-Describe our arguments of the net directionality. 
-Number of recombinant edges.
-Confidence about the numbers of recombinant edges.
-Not all of genes experience recombining gene transfers from SPY to SDE.
-Gene transfer counts from the parsimony-based approach; the power of the
-parsimony-based approach may be limited.
-Is this number of recombinant edges good enough for the argument?
-Reflection of the previous analyses, 
-heat maps of real data analysis, 
-and supporting simulations.}
+% \texttt{We show a net directionality of recombining gene transfers from SPY to SDE. 
+% Describe our arguments of the net directionality. 
+% Number of recombinant edges.
+% Confidence about the numbers of recombinant edges.
+% Not all of genes experience recombining gene transfers from SPY to SDE.
+% Gene transfer counts from the parsimony-based approach; the power of the
+% parsimony-based approach may be limited.
+% Is this number of recombinant edges good enough for the argument?
+% Reflection of the previous analyses, 
+% heat maps of real data analysis, 
+% and supporting simulations.}
 
@@
 preliminary step in the pipeline involved estimating the species tree with
-branch lengths (via \texttt{ClonalFrame}), and three population parameters of
-mutation rate, recombination rate, and average recombinant tract length (via a
-first phase of \texttt{ClonalOrigin}).  Because \textit{S.\ equi} ssp.\textit{\
-equi} strain 4047 (SEE) diverged significantly from the others among the six
-genomes (Table S\ref{tab:genome}), we could not obtain sufficiently long
-multiple sequence alignments with it, and had to exclude it from the model-based
-analysis.  The estimated species tree using the alignment of the five genomes
-(Figure S\ref{fig:mauve}) had the expected topology \texttt{The figure 1 must be
-nicer than one that now} (see Figure \ref{fig:tree5}). \texttt{ClonalFrame} also
-estimated a relative strength of recombination compared to mutations
-($r/m=4.66$), indicating potentially significant roles of recombination in the
-evolutionary process of shaping these genomes; four times more sites had been
-affected by recombination-driven substitutions than by mutation-driven
-substitutions.  (Table S\ref{tab:clonalframe}).  Estimates of population
-mutation rate ($\theta$), recombination rate ($\rho$), and average tract length
-($\delta$) were obtained for the alignment blocks analyzed (Figure
-S\ref{fig:scatter3}), and the median weighted over block lengths for each of the
-three parameters was obtained as a global estimate (Table S\ref{tab:three} and
-see \textbf{Materials and Methods}).  The local estimates typically seemed to be
-moderately fluctuating around the global average, with the exception of what
-appeared to be a recombination hotspot at a region 200 kilobase pairs from the
-origin of replication of SDE1.
+branch lengths via \texttt{ClonalFrame} \cite{Didelot2007}, and three population
+parameters of mutation rate, recombination rate, and average recombinant tract
+length via a first phase of \texttt{ClonalOrigin} \cite{Didelot2010}.  Because
+\textit{S.\ equi} ssp.\textit{\ equi} strain 4047 (SEE) diverged significantly
+from the others among the six genomes (Table S\ref{tab:genome}), we could not
+obtain sufficiently long multiple sequence alignments with it, and had to
+exclude it from the model-based analysis.  The estimated species tree using the
+alignment of the five genomes (Figure S\ref{fig:mauve}) had the expected
+topology \texttt{SC: The figure 1 must be nicer than one that now} (see Figure
+\ref{fig:tree5}). \texttt{ClonalFrame} also estimated a relative strength of
+recombination compared to mutations as $r/m=4.66$ meaning that four times more
+sites had been affected by recombination-driven substitutions than by
+mutation-driven substitutions (Table S\ref{tab:clonalframe}).  This indicated
+potentially significant roles of recombination in the evolutionary process of
+shaping these genomes.  Estimates of population mutation rate ($\theta$),
+recombination rate ($\rho$), and average tract length ($\delta$) were obtained
+for the alignment blocks analyzed (Figure S\ref{fig:scatter3}), and a median of
+the estimates weighted over block lengths was obtained as a global estimate for
+each of the three parameters (Table S\ref{tab:three}).  The local estimates
+typically seemed to be moderately fluctuating around the global estimate, with
+the exception of what appeared to be a recombination hotspot at a region 200
+kilobase pairs from the origin of replication of SDE1.
 
 The estimated species tree together with the global estimates of the three
-population parameters were used in the second phase of \texttt{ClonalOrigin},
-where recombinant trees were sampled from the posterior distribution. One way to
-summarize the posterior sample of recombinant trees was to count the topologies
-of the local genealogies induced by it. Of the 105 possible rooted topologies
-relatig five taxa, the dominant topology (66.7\%) was that of the species tree
-(see \textbf{Materials and Methods} and Table S\ref{tab:Gene-tree-topologies}).
-The second most frequently sampled topology (9.3\%) was the one in which SPY and
-SDE were sister clades to each other with SDD being as an outgroup. Other
-topologies were sampled with frequency less than 4\%. This gave initial
-indication to the extent of recombining gene transfer  between SPY and SDE.
-
-To gain a more fine-scale view of the main types of recombination scenarios, we
-looked at the recombinant edges present in the posterior sample of recombinant
+population parameters were used to sample recombinant trees from the posterior
+distribution in the second phase of \texttt{ClonalOrigin} \cite{Didelot2010}.
+We summarized the posterior sample of recombinant trees by counting topologies
+of the local genealogies induced by those recombinant trees for each of the
+genomic positions (Table S\ref{tab:Gene-tree-topologies}).  Of the 105 possible
+rooted topologies relatig five taxa, the dominant topology (66.7\%) was that of
+the species tree.  The second most frequently sampled topology (9.3\%) was the
+one in which SPY and SDE were sister clades to each other with SDD being as an
+outgroup. Other topologies were sampled with frequency less than 4\%. This gave
+initial indication to the extent of recombining gene transfer  between SPY and
+SDE.
+
+To gain a fine-scale view of the main types of recombination scenarios, we
+looked at recombinant edges present in the posterior sample of recombinant
 trees. Following the approach of Didelot et al. \cite{Didelot2010}, we
@@
 branches of the species tree (Figure \ref{fig:tree5}). For each edge type, we
-recorded the average number of edges of that type present in the sampled
-recombinant tree, and computed the ratio of this observed frequency to the prior
-expected one as implied by the species tree (considering branch lengths and
-their relative positions) and the three population parameters estimated in the
-preliminary step of the model-based analysis.  Base 2 logarithms of these ratios
-are depicted in the heatmap of Figure \ref{fig:Heatmap-of-recombination} (see
-also Tables S\ref{tab:heatmap} and S\ref{tab:obsheatmap}).
-
-Among the possible recombinant edge types, we distinguished between ones that
-preserve the species tree topology, and ones that alter it. We focused on the
-topology-altering types, both because our main objective was to study
+averaged the number of edges present in the sampled recombinant tree, and
+computed the ratio of this observed average frequency to a prior expected
+frequency. The prior expected frequencies was computed by considering branch
+lengths and their relative positions of the species tree, and the three
+population parameter estimates in the preliminary step.  Base 2 logarithms of
+these ratios are depicted in the heatmap of Figure
+\ref{fig:Heatmap-of-recombination} (see also Tables S\ref{tab:heatmap} and
+S\ref{tab:obsheatmap}).
+
+Among the possible recombinant edge types, we can distinguish between ones that
+preserve the species tree topology, and ones that alter it. We focused on 
+topology-altering edge types both because our main objective was to study
 cross-clade gene transfer, and because our simulation study indicated that the
@@
 
-We observed a reduced count of recombinant edges sampled between SDD and the
-three branches of the SPY clade ($\sim$25\% of the expected count). This was
+We observed reduction in the number of recombinant edges sampled between SDD and
+the three branches of the SPY clade ($\sim$25\% of the expected count). This was
 consistent with the fact that these species had few opportunities to interact
-with each other, SPY being a human pathogen and SDD being strictly veterinary.
-On the contrary, the number of sampled recombinant edges between SDD and
-branches in the SDE clade seemed to be consistent with the prior, with a slight
-excess for edges from SDE1 and SDE2 to SDD.  The greatest deviation of observed
-counts from expected ones was observed for recombinant edges between the SPY and
-SDE clades. It was concordant with the fact that these two species shared an
-ecological niche; SDE was an opportunistic human pathogen. Interestingly, there
-appeared to be much higher enrichment (up to four times more than expected) for
-recombinant edges originating in the SPY clade and ending up in the SDE clade.
-In a follow-up simulation study we show that this deviation from the expected
-count was indeed highly significant, and so was the observed asymmetry in
-pattern of recombination (see \textbf{Simulations}).  This net directionality of
-recombination pattern from SPY to SDE was especially intriguing in light of the
-hypothesis that SDE acquired human pathogenicity via gene transfer from SPY.
+with each other, SPY being a human pathogen and SDD being a strictly veterinary
+one.  On the contrary, the number of sampled recombinant edges between SDD and
+branches in the SDE clade seemed to be consistent with the prior.  The greatest
+deviation of observed counts from expected ones was in recombinant edges between
+SPY and SDE clades. It was concordant with the fact that these two species
+shared an ecological niche; SDE was an opportunistic human pathogen.
+Interestingly, there appeared to be much higher enrichment (up to four times
+more than expected) for recombinant edges originating in the SPY clade and
+ending up in the SDE clade.  In a follow-up simulation study we concluded that
+this deviation from the expected count was indeed highly significant, and so was
+the observed asymmetry in pattern of recombination (see \textbf{Simulations}).
+This net directionality of recombination pattern from SPY to SDE was especially
+intriguing in light of the hypothesis that SDE acquired human pathogenicity via
+gene transfer from SPY.
 
@@
 
-Figure \ref{fig:Gene-duplication-loss} shows the inferred numbers of gene gain,
+Figure \ref{fig:Gene-duplication-loss} shows inferred numbers of gene gain,
 loss, duplication, and two different kinds of gene transfer events along
 lineages of the tree that relates the six genomes. SPY had less genes than the
-other species. All of the genomes were used to construct 2314 gene families.
+other species. All of the six genomes were used to construct 2314 gene families.
 The largest family consisted of 59 genes while nearly half of gene families
@@
 genes.  Whereas gene loss events were either comparable to or less than gene
-gain ones along most of the lineages, it was not the case along the two internal
+gain events along most of the lineages, it was not the case along the two internal
 branches of SPY and SDE.  This indicated that SPY and SDE lineages had reduced
@@
 because we did not have any model-based expectation for them, which restricted
-our ability to interprest these counts. We did notice, however, a large number
-of inferred recombining gene transfer events from the ancestral SPY branch to
-the ancestral SDE branch, consistent with our observations from the model-based
-analysis. Yet, we did not observed a similar pattern with horizontal gene
-transfer while we observed a large number of horizontal transfers from the
+our ability to interpret these counts. We did notice, however, a large number of
+inferred recombining gene transfer events from the ancestral SPY branch to the
+ancestral SDE branch, consistent with our observations from the model-based
+analysis. Yet, we did not observe any pattern of a net directional horizontal
+gene transfer while we observed a large number of horizontal transfers from the
 ancesral SDE branch to SDD.
@@
 \textbf{Materials and Methods}). First, we checked association of transfers with
-virulence genes, which were provided from a previous study of Suzuki et al.\
-\citet{Suzuki2011}.  This set of virulence genes was found to be significantly
+virulence genes, which were provided by a previous study of Suzuki et al.\
+\citet{Suzuki2011}.  The set of virulence genes was found to be significantly
 associated with horizontal gene transfers (P-value $<1.33 \times 10^{-11}$), but
@@
 recombining gene transfer (Table \ref{tab:go-events}). The top ranked functional
-categories in that list were related with transposases. These differences add to
-our previous findings about the different roles these two types of transfer play
-in the evolution of these species.
+categories in that list were related with transposases. These differences added
+to our previous findings about the different roles the two types of gene
+transfer had played in the evolution of these species.
 
@@
 
-{\bf [IG: MAYBE THERE IS A BETTER WAY TO DEMONSTRATE THE CORRELATION BETWEEN METHODS]}.
-
 \subsection*{A closer study of virulence genes}
@@
 Often computational and statistical analyses on genomes could benefit other
-researchers when those resources were provided via Genome Browser.  We wished to
+researchers when those resources were available in a form of genome browser.  We wished to
 provide research community especially in SPY and SDE bacteria with results from
@@
 track of probability of recombining gene transfer allowed us to easily browse
-the bacterial genome to locate regions that we might be interested in.  Using
+the bacterial genome to locate regions that we were interested in.  Using
 the genome browser tracks we located virulence genes with notable posterior
@@
 comprised of virulence facters, about half of which were found in the core
-genome of the five genome alignment \citep{Suzuki2011}.  By sorting virulence
+genome of the five genome alignment \citep{Suzuki2011}.  After sorting virulence
 genes by the number of horizontal gene transfer events we listed several genes
@@
 The whole gene of  putative dipeptidase (SpyM3\_0465) was transferred, which was
-a case for the finding that virulence genes tended to be transferred as a whole
-unit \citep{Chan2009}.  Another example of whole gene transfer was superoxide
+a exemplary case for the finding \citep{Chan2009} that virulence genes were transferred as a whole
+unit.  Another example of whole gene transfer was superoxide
 dismutase (SpyM3\_1071).  We investigated the unit of recombining gene transfer
@@
 
-\subsection*{Simulations}
-
-We conducted experiments on simulation data to check the accuracy of estimates
-obtained by the model-based approach in a scenario similar to the one from the
-real data set.  Population mutation and recombination parameters were accurately
-estimated (Table S\ref{tab:sim-three}).  Yet, the recombinant tract length was
-over-estimated; recombinant tract lengths were reported to be difficult to
-estimate from relatively short alignment blocks \citep{Didelot2010}.  Numbers of
+\subsection*{Simulation study and additional support for biased directionality in recombining gene transfer between SPY and SDE}
+
+We conducted a series of experiments on simulated data to check the reliability of estimates obtained by ClonalOrigin.
+We found that the population parameters obtained in the first phase of ClonalOrigin are rather reliable, with the exception of the average recombinant tract length, which appeared to be over-estimated (this was observed also in \citep{Didelot2010}).
+Our experiments also show that the recombination intensity used in our association tests and traced in the genome browser tracks are reliably recovered by ClonalOrigin (Figure S\ref{fig:ri1}).
+See Materials and Methods for a more detailed report of these results.
+One issue we
+
+ with a special focus on its ability to detect asymmetric patterns of recombination such as the one we found between SPY and SDE.
+
+Recombination intensities that were measured on genes also showed rough
+correlation (Pearson's correlation coefficient of 0.63) of the estimated and
+true values (Figure S\ref{fig:ri1}).  Estimated values appeared to be driven
+somewhat by the prior implied by the population parameter setting, explaining
+why lower values were over-estimated and higher values were under-estimated.
+Nonetheless, the correlation indicated that the measure of recombination
+intensity used in our functional gene category served as an adequate indicator
+of actual intensity of recombination events along genes.
+
+Numbers of 
 recombinant edges from the simulated recombinant trees of Simulation 2 (see
@@
 gene transfer from SPY to SDE.
-
-Recombination intensities that were measured on genes also showed rough
-correlation (Pearson's correlation coefficient of 0.63) of the estimated and
-true values (Figure S\ref{fig:ri1}).  Estimated values appeared to be driven
-somewhat by the prior implied by the population parameter setting, explaining
-why lower values were over-estimated and higher values were under-estimated.
-Nonetheless, the correlation indicated that the measure of recombination
-intensity used in our functional gene category served as an adequate indicator
-of actual intensity of recombination events along genes.
 
@@
 \section*{Figure Legends}
+
+\begin{flushleft}
+
+Figure \ref{fig:tree5}.\ \bf{The species tree of the five genomes inferred using
+\texttt{ClonalFrame}.}  SPY is the branch of a clade of \textit{Streptococcus
+pyogenes}, SDE is that for \textit{S.\ dysgalactiae} ssp.\textit{\ equisimilis},
+and SD is the branch for the three individual genomes of \textit{S.\
+dysgalactiae}. The scale is proportional to the expected number of mutations
+given the Watterson's estimate of mutation rate and the tree.
+
+Figure \ref{fig:Heatmap-of-recombination}.\ \bf{Heatmap of recombination events 
+using the five genomes.}
+Each cell of the heat map represents the logarithm base
+2 of the number of recombination events inferred relative to its expectation
+under the prior, for a donor and recipient pair of branches. The donor is given by
+the y-axis; the recipient on the x-axis.  The black colored cells are meaningless
+because the recipient species tree branch of the cells are older than the donor
+species branch.
+
+Figure \ref{fig:Gene-duplication-loss}.\ \bf{Distribution of inferred gene 
+duplication, loss, and transfer events across
+the phylogeny.}  Gene counts for
+each extant and ancestral species are given at each node of the phylogeny.  For
+each branch we denote the number of gains G*, duplications D*, losses L*,
+non-recombining transfers T*, and recombinations R*.  The non-recombining
+transfer and recombination counts are indicated on the recipient branch (donors
+are not indicated).
+
+Figure \ref{fig:h3}.\ \bf{Simulations of the model-based approach for assessing 
+the power of recovering recombinant edges.} 
+Pairs of species tree branches are shown at x-axis labels.
+Three intervals are displayed for each pair. Intervals with solid line with open
+circle in the middle are for the logarithm of the fraction to base 2 of the average
+number of recombinant edges from the ten simulated recombinant trees 
+to the numbers of recombinant edges estimated from Simulation 2. 
+Interverals with dotted line are for the logarithm of the fraction to base 2 of
+the number of recombinant edges estimated from the real data
+to the numbers of recombinant edges estimated from Simulation 2.
+Intervals with solid line with closed circle in the middle are for the logarithm
+of the fraction to base 2 of the number of recombinant edges estimated from the real data
+to the numbers of recombinant edges estimated from Simulation 3. 
+The horizontal lines are at the zero.
+The intervals range from 5\% quantile to 95\% of quantile.
+
+\end{flushleft}
+\clearpage{}
+
 %\begin{figure}[!ht]
@@
 
-%\section*{Figures\clearpage{}}
-
-%=============================================================================
-% Figure S2: Species tree
+%=============================================================================
+% Figure 1: Species tree
 \begin{figure}
@@
 %=============================================================================
-% Figure 1. Heat map of recombining gene transfer from ClonalOrigin
+% Figure 2. Heat map of recombining gene transfer from ClonalOrigin
 \begin{figure}
 \includegraphics[scale=0.45]{figures/heatmap-recedge}
-\caption{\label{fig:Heatmap-of-recombination}Heatmap of recombination events
-using the five genomes.  Each cell of the heat map represents the logarithm base
+\caption{\label{fig:Heatmap-of-recombination}
+\bf{Heatmap of recombination events using the five genomes.}
+Each cell of the heat map represents the logarithm base
 2 of the number of recombination events inferred relative to its expectation
 under the prior, for a donor and recipient pair of branches. The donor is given by
-the y-axis; the recipient on the x-axis.  The green colored cells are meaningless
+the y-axis; the recipient on the x-axis.  The black colored cells are meaningless
 because the recipient species tree branch of the cells are older than the donor
@@
 \includegraphics[width=6in]{figures/strep-events}
-\caption{\label{fig:Gene-duplication-loss} Distribution of inferred gene
-duplication, loss, and transfer events across the phylogeny.  Gene counts for
+\caption{\label{fig:Gene-duplication-loss} 
+\bf{Distribution of inferred gene duplication, loss, and transfer events across
+the phylogeny.}  Gene counts for
 each extant and ancestral species are given at each node of the phylogeny.  For
@@
 \caption{\label{fig:h3}
-Simulations of the model-based approach for assessing the power of recovering
-recombinant edges. Pairs of species tree branches are shown at x-axis labels.
+\bf{Simulations of the model-based approach for assessing the power of recovering
+recombinant edges.} Pairs of species tree branches are shown at x-axis labels.
 Three intervals are displayed for each pair. Intervals with solid line with open
@@
 \begin{table}
-\caption{\label{tab:genome}The six complete bacterial genomes. 
-Each bacteria and its genome is referred to as the given name.
-SDD genome was not yet available from NCBI.}
+\caption{
+\bf{The six complete bacterial genomes.}
+Each bacteria and its genome is referred to as the name at the first column.}
 \noindent \begin{centering}
@@
 SDE2 & NC\_012891 & 2106340 & Shimomura et al. (2011)\tabularnewline
-SDD & N/A & 2141837 & Suzuki et al. (2011)\tabularnewline
+SDD & N/A$^a$ & 2141837 & Suzuki et al. (2011)\tabularnewline
 SPY1 & NC\_004070 & 1900521 & Beres et al. (2002)\tabularnewline
@@
 \par\end{centering}
+\begin{flushleft}
+$^a$ SDD genome was not yet available from NCBI.
+\end{flushleft}
+\label{tab:genome}
 \end{table}
@@
 % Table S2. ClonalFrame Result.
-\begin{table}
-\caption{\label{tab:clonalframe}Parameter estimates from \texttt{ClonalFrame}. 
+\begin{table}[!ht]
+\caption{
+\bf{Parameter estimates from \texttt{ClonalFrame}.}
 $\theta$ is population mutation rate,
@@
 \end{tabular}
+\label{tab:clonalframe}
 \end{table}
@@
 % Table S. ClonalOrigin's 1st stage result.
-\begin{table}
-\caption{\label{tab:three}
+\begin{table}[!ht]
+\caption{
 \bf{Three population parameter estimates from the model-based approach.}
@@
 \end{flushleft}
-
+\label{tab:three}
 \end{table}
@@
 \begin{table}
-\caption{\label{tab:Gene-tree-topologies}Posterior probability of gene tree
-topologies. See Table S\ref{tab:genome} for the reference of the bacteria
+\caption{
+\bf{Posterior probability of gene tree topologies.}
+See Table S\ref{tab:genome} for the reference of the bacteria
 individual names.}
@@
 \par\end{centering}
+\label{tab:Gene-tree-topologies}
 \end{table}
@@
 % Table S4. The values of the heat map.
-\begin{table}
-\caption{\label{tab:heatmap}Ratios of logarithm base 2 of the observed
-recombinant edges to the expected number of recombinant edges.  Figure
+\begin{table}[!ht]
+\caption{
+\bf{Ratios of logarithm base 2 of the observed
+recombinant edges to the expected number of recombinant edges.}  Figure
 \ref{fig:Heatmap-of-recombination} shows the values graphically using heat map.
@@
 \end{tabular}
+\label{tab:heatmap}
 \end{table}
@@
 % Table S5. The values of the heat map of actual counts.
-\begin{table}
-\caption{\label{tab:obsheatmap}The observed number of recombinant edges.
+\begin{table}[!ht]
+\caption{
+\bf{The observed number of recombinant edges.}
 The zeros are for the branch pairs that are impossible under the model-based
@@
 \end{tabular}
+\label{tab:obsheatmap}
 \end{table}
@@
 %=============================================================================
-% Table S. Mowgli - Recombining Transfer
-\begin{table}
-\caption{\label{tab:go-events-recombining}
+% Table S7. Mowgli - Recombining Transfer
+\begin{table}[!ht]
+\caption{
 \bf{Functional categories enriched in 
@@
 \par\end{centering}
+\label{tab:go-events-recombining}
 \end{table}
@@
 % Table S. 
-\begin{table}
-\caption{\label{tab:virrecomb}
+\begin{table}[!ht]
+\caption{
 \bf{Virulence genes inferred to have experienced recombining gene transfer
@@
 \par\end{centering}
+\label{tab:virrecomb}
 \end{table}
@@
 % Table S. 
-\begin{table}
-\caption{\label{tab:virhgt}
+\begin{table}[!ht]
+\caption{
 \bf{Virulence genes inferred to have experienced horizontal gene transfer
@@
 \par\end{centering}
+\label{tab:virhgt}
 \end{table}
@@
 % Table S6. Genes with evident recombining gene transfer.
-\begin{table}
-\caption{\label{tab:genes-transfer}Genes and their parts transferred via 
-homologous
-recombining gene transfer. }
+\begin{table}[!ht]
+\caption{
+\bf{Genes and their parts transferred via homologous
+recombining gene transfer.}}
 {\footnotesize
@@
 
+\begin{flushleft}
 $^a$ Numbers represent a group of loci that are consecutive in gene order. Empty
@@
 the locus.
+\end{flushleft}
 }
-
+\label{tab:genes-transfer}
 \end{table}
@@
 \end{center}
-\caption{\label{fig:mauve}The genome alignment of the five genomes. The top
+\caption{\label{fig:mauve}
+\bf{The genome alignment of the five genomes.} The top
 first and second rows are SDE1 and SDE2 genomes, respectively. The third is SDD
@@
 %=============================================================================
-% Figure S3: Histogram of block sizes
+% Figure S: Histogram of block sizes
 %\begin{figure}
@@
 %=============================================================================
-% Figure S4. Three population parameters.
+% Figure S2. Three population parameters.
 \begin{figure}
@@
 %=============================================================================
-% Figure S5: Scatter plot of rho and log(delta)
-\begin{figure}
+% Figure S3: gene family sizes
+% FIGURE - gene family sizes
+\begin{figure}[!ht]
 \begin{center}
-\includegraphics[width=5in]{figures/rho-vs-delta}
+\includegraphics[width=5in]{figures/famsizes}
 \end{center}
-\caption{Plot of recombination rate and logarithm of average tract length
-of the posterior samples.}
-\label{fig:rhologdelta}
+\caption{\bf{Distribution of gene family sizes.}}
+\label{fig:famsizes}
+\end{figure}
+\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% FIGURE S4. - HGT Transfer heatmap
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[scale=0.45]{figures/heatmap-hgtedge}
+\end{center}
+\caption{
+\bf{Heat map of horizontal gene transfers using the six genomes.}
+Each cell of the color matrix shows the number of inferred non-recombining
+horizontal gene transfer events between the different branches of the species
+tree. Black colored cells represent no inferred events.}
+\label{fig:hgt-heatmap}
 \end{figure}
@@
 %=============================================================================
+% Figure S5: recombination heatmap from mowgli
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[scale=0.45]{figures/heatmap-recedge-mowgli}
+\end{center}
+\caption{
+\bf{Heat map of recombining gene transfers using the six genomes.}
+Each cell of the color matrix shows the number of inferred recombining
+gene transfer events between the different branches of the species
+tree. Black colored cells represent no inferred events.}
+\label{fig:mowgli-recomb-heatmap}
+\end{figure}
+\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% Figure S6: Comparison of ClonalOrigin and Mowgli analysis
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=5in]{figures/compareMowgliCoRecombining}
+\end{center}
+\caption{
+\bf{Plot of recombination intensity against number
+of recombining gene transfer.}}
+\label{fig:cmpcomowgli}
+\end{figure}
+\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% Figure S7: HGT Comparison of ClonalOrigin and Mowgli analysis
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=5in]{figures/compareMowgliCoHgt}
+\end{center}
+\caption{
+\bf{Plot of recombination intensity against number
+of horizontal gene transfer.}}
+\label{fig:cmpcomowglihgt}
+\end{figure}
+\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% Figure S5: Scatter plot of rho and log(delta)
+% \begin{figure}
+% \begin{center}
+% \includegraphics[width=5in]{figures/rho-vs-delta}
+% \end{center}
+% \caption{Plot of recombination rate and logarithm of average tract length
+% of the posterior samples.}
+% \label{fig:rhologdelta}
+% \end{figure}
+% \clearpage{}%
+%=============================================================================
+
+%=============================================================================
+% Figure S8: View of UCSC genome browser for bacterial gene transfer.
+\begin{figure}[!ht]
+\includegraphics[scale=0.8]{figures/ucsc}
+\caption{
+\bf{UCSC genome browser tracks for gene transfer study of
+SPY and SDE.}}
+\label{fig:ucsc}
+\end{figure}
+\clearpage{}%
+%=============================================================================
+
+%=============================================================================
 % Figure S6. A Recombinant Tree of ClonalOrigin Model
-\begin{figure}
+\begin{figure}[!ht]
 \includegraphics[scale=0.42]{figures/clonalorigin}
-\caption{\label{fig:clonalorigin}Recombinant tree in the \texttt{ClonalOrigin} model.
+\caption{
+\bf{Recombinant tree in the \texttt{ClonalOrigin} model.}
 (A) A recombinant tree relating the five genomes with four recombinant edges.
@@
 labeled topologies relating five taxa, we could count topologies.}
-\end{figure}
-\clearpage{}%
-%=============================================================================
-
-%=============================================================================
-% Figure S8: gene family sizes
-% FIGURE - gene family sizes
-\begin{figure}
-\begin{center}
-\includegraphics[width=5in]{figures/famsizes}
-\end{center}
-\caption{\label{fig:famsizes}Distribution of gene family sizes.}
-\end{figure}
-\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% FIGURE 2. - HGT Transfer heatmap
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.45]{figures/heatmap-hgtedge}
-\end{center}
-\caption{\label{fig:hgt-heatmap}Heat map of horizontal gene transfers using the
-six genomes.
-Each cell of the color matrix shows the number of inferred non-recombining
-horizontal gene transfer events between the different branches of the species
-tree. Black colored cells represent no inferred events.}
-\end{figure}
-\clearpage{}%
-%=============================================================================
-
-%=============================================================================
-% Figure S7: recombination heatmap from mowgli
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.45]{figures/heatmap-recedge-mowgli}
-\end{center}
-\caption{\label{fig:mowgli-recomb-heatmap}Heat map of recombining gene transfers using the
-six genomes.
-Each cell of the color matrix shows the number of inferred recombining
-gene transfer events between the different branches of the species
-tree. Black colored cells represent no inferred events.}
-\end{figure}
-\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Figure S9: Comparison of ClonalOrigin and Mowgli analysis
-\begin{figure}
-\begin{center}
-\includegraphics[width=5in]{figures/compareMowgliCoRecombining}
-\end{center}
-\caption{\label{fig:cmpcomowgli}Plot of recombination intensity against number
-of recombining gene transfer.}
-\end{figure}
-\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Figure S10: HGT Comparison of ClonalOrigin and Mowgli analysis
-\begin{figure}
-\begin{center}
-\includegraphics[width=5in]{figures/compareMowgliCoHgt}
-\end{center}
-\caption{\label{fig:cmpcomowglihgt}Plot of recombination intensity against number
-of horizontal gene transfer.}
-\end{figure}
-\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Figure S11: View of UCSC genome browser for bacterial gene transfer.
-\begin{figure}
-\includegraphics[scale=0.8]{figures/ucsc}
-\caption{\label{fig:ucsc}UCSC genome browser tracks for gene transfer study of
-SPY and SDE.}
+\label{fig:clonalorigin}
 \end{figure}
@@
 branch to the root branch are impossible (see Figure S\ref{fig:clonalorigin}),
-the column for the root as recipient was colored green, indicating invalid
+the column for the root as recipient was colored black, indicating invalid
 recombinant edges in Figure \ref{fig:Heatmap-of-recombination}.  
@@
 older than the donor, and the donor and recipient branches did not share any contemporary time were
-colored green. The diagonal cells were colored gray, indicating no significant
+colored black. The diagonal cells were colored gray, indicating no significant
 number of recombinant edges, because the observed and expected numbers of

### 2011-10-20 doc/manuscript.tex
Just an update.

@@
 \bibliography{siepel-strep}
+\clearpage{}
 
@@
 
-Figure \ref{fig:tree5}.\ \bf{The species tree of the five genomes inferred using
-\texttt{ClonalFrame}.}  SPY is the branch of a clade of \textit{Streptococcus
+Figure \ref{fig:tree5}.\ {\bf The species tree of the five genomes.}
+SPY is the branch of a clade of \textit{Streptococcus
 pyogenes}, SDE is that for \textit{S.\ dysgalactiae} ssp.\textit{\ equisimilis},
@@
 
-Figure \ref{fig:Heatmap-of-recombination}.\ \bf{Heatmap of recombination events 
+Figure \ref{fig:Heatmap-of-recombination}.\ {\bf Heatmap of recombination events 
 using the five genomes.}
@@
 
-Figure \ref{fig:Gene-duplication-loss}.\ \bf{Distribution of inferred gene 
+Figure \ref{fig:Gene-duplication-loss}.\ {\bf Distribution of inferred gene 
 duplication, loss, and transfer events across
@@
 
-Figure \ref{fig:h3}.\ \bf{Simulations of the model-based approach for assessing 
+Figure \ref{fig:h3}.\ {\bf Simulations of the model-based approach for assessing 
 the power of recovering recombinant edges.} 
@@
 % Table 1. 
-\begin{table}
-\caption{\label{tab:functional}
+\begin{table}[!ht]
+\caption{
 \bf{Functional categories asscoiated with 
@@
 \par\end{centering}
+\label{tab:functional}
 \end{table}
@@
 % Table 2. 
-\begin{table}
-\caption{\label{tab:go-events}Functional categories asscoiated with 
-non-recombining horizontal gene transfer.}
+\begin{table}[!ht]
+\caption{
+\bf{Functional categories asscoiated with 
+non-recombining horizontal gene transfer.}}
 \noindent \begin{centering}
@@
 \par\end{centering}
+\label{tab:go-events}
 \end{table}
@@
 
+\section*{Figures}
+
 %=============================================================================
 % Figure 1: Species tree
-\begin{figure}
+\begin{figure}[!ht]
 \includegraphics[scale=0.7]{figures/cornellf-3-tree}
-\caption{\label{fig:tree5}The species tree of the five genomes inferred using
-\texttt{ClonalFrame}.  SPY is the branch of a clade of \textit{Streptococcus
+\caption{
+{\bf The species tree of the five genomes.} 
+SPY is the branch of a clade of \textit{Streptococcus
 pyogenes}, SDE is that for \textit{S.\ dysgalactiae} ssp.\textit{\ equisimilis},
@@
 given the Watterson's estimate of mutation rate and the tree.}
+\label{fig:tree5}
 \end{figure}
@@
 \caption{\label{fig:Heatmap-of-recombination}
-\bf{Heatmap of recombination events using the five genomes.}
+{\bf Heatmap of recombination events using the five genomes.}
 Each cell of the heat map represents the logarithm base
@@
 \caption{\label{fig:Gene-duplication-loss} 
-\bf{Distribution of inferred gene duplication, loss, and transfer events across
+{\bf Distribution of inferred gene duplication, loss, and transfer events across
 the phylogeny.}  Gene counts for
@@
 
-\caption{\label{fig:h3}
-\bf{Simulations of the model-based approach for assessing the power of recovering
+\caption{
+{\bf Simulations of the model-based approach for assessing the power of recovering
 recombinant edges.} Pairs of species tree branches are shown at x-axis labels.
@@
 The intervals range from 5\% quantile to 95\% of quantile.}
+\label{fig:h3}
 \end{figure}
@@
 \bf{Parameter estimates from \texttt{ClonalFrame}.}
-$\theta$ is population mutation rate,
-$\delta$ is the average recombinant tract length,
-$r/m$ is the relative strength of recombination compared to mutations,
-TMRCA is the time to most recent common ancestor, and
-$T$ is the total branch length of the inferred species tree.}
 \noindent \centering{}\begin{tabular}{cc}
@@
 \hline
-$\theta$ & 92871.84 (N/A)\tabularnewline
-$\nu$ & 0.134 (0.132, 0.135)\tabularnewline
-$R$ & 9370 (8950, 9780)\tabularnewline
-$\delta$ & 362 (347, 377)\tabularnewline
-$r/m$ & 4.66 (4.45, 4.87)\tabularnewline
-$\rho/\theta$ & 0.101 (0.096, 0.105)\tabularnewline
-TMRCA & 0.331 (0.324, 0.337)\tabularnewline
-$T$ & 1.02 (1.01, 1.04)\tabularnewline
+$\theta^a$ & 92871.84 (N/A)\tabularnewline
+$\nu^b$ & 0.134 (0.132, 0.135)\tabularnewline
+$R^c$ & 9370 (8950, 9780)\tabularnewline
+$\delta^d$ & 362 (347, 377)\tabularnewline
+$r/m^e$ & 4.66 (4.45, 4.87)\tabularnewline
+$\rho/\theta^f$ & 0.101 (0.096, 0.105)\tabularnewline
+TMRCA$^g$ & 0.331 (0.324, 0.337)\tabularnewline
+$T^h$ & 1.02 (1.01, 1.04)\tabularnewline
 \hline
 \end{tabular}
+\begin{flushleft}
+$^a$ $\theta$ is population mutation rate,
+
+$^d$ $\delta$ is the average recombinant tract length,
+
+$^e$ $r/m$ is the relative strength of recombination compared to mutations,
+
+$^g$ TMRCA is the time to most recent common ancestor, and
+
+$^h$ $T$ is the total branch length of the inferred species tree.}
+\end{flushleft}
 \label{tab:clonalframe}

### 2011-10-20 pl/count-observed-recedge.pl

@@ -74,7 +74,7 @@ my $meanfile = "";
 my $append = 0;
 my $check = 0;
 my $xmlBasename = "core_co.phase3";
-my $endblockid = 0;
+my $endblockid = 1;
 my $outfile;
 my $isLowertime = 0; # Special case
 
@@ -171,7 +171,17 @@ elsif ($cmd eq "heatmap")
     &printError("you did not specify a number of blocks");
   }
 }
-
+elsif ($cmd eq "exponly")
+{
+  if ($heatDir eq "")
+  {
+    &printError("$cmd requires -e option");
+  }
+  unless (exists $params{n})
+  {
+    &printError("you did not specify a number of blocks");
+  }
+}
 ################################################################################
 ## DATA PROCESSING
 ################################################################################
@@ -274,16 +284,19 @@ if ($obsonly == 0)
     }
   }
 }
-if ($check == 1)
+if ($cmd eq "exponly")
 {
-  print "expMap:\n";
   for my $i ( 0 .. $#expMap ) {
-    print "  ";
     for my $j ( 0 .. $#{ $expMap[$i] } ) {
-      print "[$i][$j] $expMap[$i][$j] ";
+      print $outfile "$expMap[$i][$j]";
+      unless ($i == $#expMap and $j == $#{ $expMap[$i] })
+      {
+        print $outfile "\t";
+      }
     }
-    print "\n";
   }
+  print $outfile "\n";
+  exit;
 }
 
 ##############################################################

### 2011-10-20 sh/compute-heatmap-recedge.sh

@@ -43,6 +43,13 @@ function compute-heatmap-recedge {
       echo -e "The number of blocks is $NUMBER_BLOCK."
       echo -e "The number of species is $NUMBER_SPECIES."
       echo "NUMBER_BLOCK and NUMBER_SAMPLE must be checked"
+      echo perl pl/count-observed-recedge.pl exponly \
+        -d $RUNCLONALORIGIN/output2/${REPLICATE} \
+        -e $RUNCLONALORIGIN/output2/priorcount-${REPLICATE} \
+        -n $NUMBER_BLOCK \
+        -out $RUNANALYSIS/exponly-recedge-${REPLICATE}.txt
+
+      break
 
       perl pl/count-observed-recedge.pl \
         -d $RUNCLONALORIGIN/output2/${REPLICATE} \

### 2011-10-20 sh/run.sh

@@ -105,9 +105,9 @@ source sh/batch.sh
 #####################################################################
 # Main part of the script.
 #####################################################################
-batch
-exit
-short-notice
+#batch
+#exit
+#short-notice
 PS3="Select what you want to do with mauve-analysis: "
 CHOICES=( init-file-system \
           choose-simulation \
@@ -163,8 +163,8 @@ CHOICES=( init-file-system \
           ---RECOMBINATION-COUNT---\
           recombination-count \
           # count-observed-recedge \
-          # compute-prior-count-recedge \
-          # compute-heatmap-recedge \
+          compute-prior-count-recedge \
+          compute-heatmap-recedge \
           # prepare-run-compute-heatmap-recedge \
           ---RECOMBINATION-INTENSITY---\
           recombination-intensity1-map \

### 2011-10-20 doc/manuscript.tex
This is the version that Ilan spent time on rewriting section Results.

@@
 % Use the PLoS provided bibtex style
-\bibliographystyle{plos}
+% \bibliographystyle{plos}
 
@@
 
-\subsection*{Simulation study and additional support for biased directionality in recombining gene transfer between SPY and SDE}
-
-We conducted a series of experiments on simulated data to check the reliability of estimates obtained by ClonalOrigin.
-We found that the population parameters obtained in the first phase of ClonalOrigin are rather reliable, with the exception of the average recombinant tract length, which appeared to be over-estimated (this was observed also in \citep{Didelot2010}).
-Our experiments also show that the recombination intensity used in our association tests and traced in the genome browser tracks are reliably recovered by ClonalOrigin (Figure S\ref{fig:ri1}).
-See Materials and Methods for a more detailed report of these results.
-One issue we
-
- with a special focus on its ability to detect asymmetric patterns of recombination such as the one we found between SPY and SDE.
-
-Recombination intensities that were measured on genes also showed rough
-correlation (Pearson's correlation coefficient of 0.63) of the estimated and
-true values (Figure S\ref{fig:ri1}).  Estimated values appeared to be driven
-somewhat by the prior implied by the population parameter setting, explaining
-why lower values were over-estimated and higher values were under-estimated.
-Nonetheless, the correlation indicated that the measure of recombination
-intensity used in our functional gene category served as an adequate indicator
-of actual intensity of recombination events along genes.
-
-Numbers of 
-recombinant edges from the simulated recombinant trees of Simulation 2 (see
-\textbf{Supplementary Information} for the description of Simulation 2) were
-compared with the corresponding numbers of recombinant edges that were estimated
-from the simulated data set from Simulation 2 (see the solid-line intervals with
-open circles in Figure \ref{fig:h3}).  We noted that tree-topology-changing
-recombinant edges were more reliably recovered compared with
-tree-topology-not-changing ones (Figure \ref{fig:h3}). All of the
-tree-topology-not-changing recombinant edge types were over-estimated (Figure
-\ref{fig:h3}B). The heat map of Figure \ref{fig:Heatmap-of-recombination} did
-not show how confident of the values we could be. Comparison of numbers of
-recombinant edges estimated from the real data set and the corresponding numbers
-estimated from Simulation 2 showed the interval estimates for the heat map
-values.  Dotted-line intervals from the red and blue colored cells in the heat
-map were above and below the reference of the horizontal line at the zero on the
-Y-axis, respectively (Figure \ref{fig:h3}).
-
-for each recombinant edge type to assess the effect of posterior recombinant
-trees samples with respect to that of prior recombinant tree samples. The ten
-\textit{a priori} recombinant trees from Simulation 2 would have balanced
-recombining gene transfer events between SPY and SDE with respect to the prior
-expected number of recombinant edges. The posterior recombinant trees used in
-Simulation 3 would have unbalanced recombining gene transfer between the two
-species branches as we observed in the real data analysis. For significantly
-strong gene flows such as ones from SPY to SDE a solid-line interval with a
-closed circle would be different from its count part of the dotted-line
-interval.  Figure \ref{fig:h3} showed that the recombining gene flow from SPY to
-SDE was stronger enough to be nearer to the horizontal line at the zero on the
-Y-axis. However, the recombining gene transfer from SDE to SPY appeared to be
-too weak.  This comparison supported the biased directionality of recombining
-gene transfer from SPY to SDE.
+\subsection*{Simulation study and additional support for biased directionality
+in recombining gene transfer between SPY and SDE}
+
+We conducted a series of experiments on simulated data to check the reliability
+of estimates obtained by the model-based approach. We found that the population
+parameters obtained in the first phase were reliable, with the exception that
+the average recombinant tract length was over-estimated (this was observed also
+in \citep{Didelot2010}). Our experiments also showed that the recombination
+intensity used in our functional category association tests had fairly good
+correlation between the true value and the simulated ones.  (Figure
+S\ref{fig:ri1}). See \textbf{Materials and Methods} for a more detailed report
+of these results. One thing we specifically wished to test in these experiments
+was whether the model-based approach reliably detected deviations of recombinant
+edge counts from their expectation, as observed in the heatmap of Figure
+\ref{fig:Heatmap-of-recombination}. Our major concern was that these deviations
+were not directly accounted for by the probabilistic model of ClonalOrigin, but
+rather indirectly observed by summary of the posterior sample of recombinant
+trees.
+
+We simulated the model to generate a hundred simulated data sets using a
+collection of recombinant trees sampled from the prior distribution defined by
+the species tree and three population parameters, as estimated from our genomic
+data.  We then repeat the inference procedure of the second stage of
+\texttt{ClonalOrigin} on each data set and, as in the data analysis, and counted
+for each donor-recepient branch pair the average number of recombinant edges.
+By normalizing the counts by the prior expected count we obtained intervals of
+base 2 logarithms of the ratio of the observed number of recombinant edges to
+the prior expectionation (Figure \ref{fig:h3}). The comparison showed a good
+concordance of the estimated counts and the prior expected counts for all
+topology-altering edge types and a clear over-estimation for the counts of
+topology-preserving edges. Recall that we made sure to ignore these edges in our
+analysis. For each edge type, we also plotted in Figure \ref{fig:h3} the
+log-ratio of the count estimated for it in our data analysis and the expected
+count (the closed circles without interval bars); these are the values making up
+the heatmap of Figure \ref{fig:Heatmap-of-recombination}. Comparing the counts
+estimated in the data analysis and the counts estimated from the simulated data
+allowed to evaluate the significance of the divergence from expectation observed
+in the heatmap.  Recall the heatmap indicated an excess of recombinant edges
+between the SPY and SDE clades in both directions. Our simulations indicated
+that the excess observed in the direction SDE to SPY was much less significant,
+as the counts estimated from real data often falled very close to the
+corresponding counts estimated from simulated data. The excess observed in the
+direction SPY to SDE appeared not only to be larger, but also more significant.
+This further supported our claim for the net directionality in recombining gene
+transfer between these two species.
+
+We wished to use simulated data to test not only the ability of the ClonalOrigin
+sample to \textit{detect} deviations from expected counts of recombinant edges, but
+also its ability to reliably \textit{quantify} this deviation. For this purpose we
+generated a hundred simulated data sets using a collection of recombinant trees
+sampled from the posterior sample of ClonalOrigin on the genomic data. The
+difference between the two simulated batches was that in the first one, the
+recombinant trees used represent the prior distribution, and in the second one,
+they represented the posterior distribution, which we knew deviated from the
+prior. Again, we recorded the recombinant edge counts obtained by ClonalOrigin
+and normalized these counts by the prior expected counts (intervals with closed
+circles in in Figure \ref{fig:h3}). We expected these counts to be in better agreement with
+the counts observed in the data analysis. For recombinant edges that showed a
+defficiency in our data analysis (e.g., from the SPY clade to SDD), the counts
+appeared to fit well with the count estimated from genomic data. However, for
+recombinant edges that showed an excess in our data analysis (e.g., from the SPY
+clade to the SDE clade), the counts appeared to be significantly influenced by the
+prior. We noted that the influence of the prior appeared to be even stronger for
+topology-preserving recombinant edges. All things considered, we found that defficiency of
+a certain recombinant edge type was reliably estimated whereas excesses were 
+typically underestimated. Therefore, we concluded that the actual extent to which
+recombining transfers between SPY and SDE \textit{preferred} the direction from SPY to
+SDE was possibly somewhat underestimated in our analysis.
 
@@
 \caption{
-\bf{Functional categories asscoiated with 
+{\bf Functional categories associated with 
 recombination intensity based on topology-altering recombinant edges.}}
@@
 \caption{
-\bf{Functional categories asscoiated with 
+{\bf Functional categories associated with 
 non-recombining horizontal gene transfer.}}
@@
 
-% \section*{Supplementary Tables}
-% \clearpage{}
+\section*{Supplementary Tables}
 
@@
 % Table S1. List of six genomes.
-\begin{table}
-\caption{
-\bf{The six complete bacterial genomes.}
+\begin{table}[!ht]
+\caption{
+{\bf The six complete bacterial genomes.}
 Each bacteria and its genome is referred to as the name at the first column.}
@@
 \caption{
-\bf{Parameter estimates from \texttt{ClonalFrame}.}
+{\bf Parameter estimates from \texttt{ClonalFrame}.}}
 \noindent \centering{}\begin{tabular}{cc}
@@
 
-$^h$ $T$ is the total branch length of the inferred species tree.}
+$^h$ $T$ is the total branch length of the inferred species tree.
 \end{flushleft}
@@
 \caption{
-\bf{Three population parameter estimates from the model-based approach.}
+{\bf Three population parameter estimates from the model-based approach.}
 The estimates of the three parameters were based on one of two independent
@@
 %=============================================================================
-% Table S3. Gene Tree topologies
-\begin{table}
-\caption{
-\bf{Posterior probability of gene tree topologies.}
+% Table S4. Gene tree topologies
+\begin{table}[!ht]
+\caption{
+{\bf Posterior probability of gene tree topologies.}
 See Table S\ref{tab:genome} for the reference of the bacteria
@@
 \caption{
-\bf{Ratios of logarithm base 2 of the observed
+{\bf Ratios of logarithm base 2 of the observed
 recombinant edges to the expected number of recombinant edges.}  Figure
@@
 \caption{
-\bf{The observed number of recombinant edges.}
+{\bf The observed number of recombinant edges.}
 The zeros are for the branch pairs that are impossible under the model-based
@@
 \caption{
-\bf{Functional categories enriched in 
+{\bf Functional categories enriched in 
 recombining transfer events that are inferred using the parsimony-based
@@
 \caption{
-\bf{Virulence genes inferred to have experienced recombining gene transfer
+{\bf Virulence genes inferred to have experienced recombining gene transfer
 events.}}
@@
 \caption{
-\bf{Virulence genes inferred to have experienced horizontal gene transfer
+{\bf Virulence genes inferred to have experienced horizontal gene transfer
 events.}}
@@
 \caption{
-\bf{Genes and their parts transferred via homologous
+{\bf Genes and their parts transferred via homologous
 recombining gene transfer.}}
@@
 % Table S8. ClonalOrigin - All Topologies
-\begin{table}
-\caption{\label{tab:functional-all}Functional categories asscoiated with 
-recombination intensity based on all of the recombinant edge types.}
+\begin{table}[!ht]
+\caption{
+{\bf Functional categories associated with 
+recombination intensity based on all of the recombinant edge types.}}
 \noindent \begin{centering}
@@
 \par\end{centering}
+\label{tab:functional-all}
 \end{table}
@@
 % Table S. Simulation study for the three population parameters.
-\begin{table}
-\caption{\label{tab:functional-notopology}Functional categories asscoiated with 
-recombination intensity based on the topology-not-changing recombinant edge types.}
+\begin{table}[!ht]
+\caption{
+{\bf Functional categories asscoiated with 
+recombination intensity based on the topology-not-changing recombinant edge
+types.}}
 \noindent \begin{centering}
@@
 \par\end{centering}
+\label{tab:functional-notopology}
 \end{table}
@@
 % Table S15. Simulation study for the three population parameters.
-\begin{table}
-\caption{\label{tab:sim-three}
-Simulation study of the three population parameters.}
+\begin{table}[!ht]
+\caption{
+{\bf Simulation study of the three population parameters.}}
 \noindent \begin{centering}
@@
 \par\end{centering}
+\label{tab:sim-three}
 \end{table}
-%\clearpage{}
-%=============================================================================
-
-%\section*{Supplementary Figures}
+%=============================================================================
 
 \clearpage{}
+\section*{Supplementary Figures}
 
@@
 % Figure S1: Mauve alignment of the five genomes.
-\begin{figure}
+\begin{figure}[!ht]
 \begin{center}
@@
 \caption{\label{fig:mauve}
-\bf{The genome alignment of the five genomes.} The top
+{\bf The genome alignment of the five genomes.} The top
 first and second rows are SDE1 and SDE2 genomes, respectively. The third is SDD
@@
 % Figure S2. Three population parameters.
-\begin{figure}
+\begin{figure}[!ht]
 \begin{center}
@@
 \end{center}
-\caption{\label{fig:scatter3}Three scatter plots of mutation rate, recombination
-rate, and average tract length for the blocks. The plus signs are mean values
+\caption{
+{\bf Three scatter plots of mutation rate, recombination
+rate, and average tract length for the blocks.} The plus signs are mean values
 for each block. The density of each value for each block is depicted by black
@@
 \end{figure}
+\label{fig:scatter3}
 \clearpage{}%
@@
 \end{center}
-\caption{\bf{Distribution of gene family sizes.}}
+\caption{{\bf Distribution of gene family sizes.}}
 \label{fig:famsizes}
@@
 \caption{
-\bf{Heat map of horizontal gene transfers using the six genomes.}
+{\bf Heat map of horizontal gene transfers using the six genomes.}
 Each cell of the color matrix shows the number of inferred non-recombining
@@
 \caption{
-\bf{Heat map of recombining gene transfers using the six genomes.}
+{\bf Heat map of recombining gene transfers using the six genomes.}
 Each cell of the color matrix shows the number of inferred recombining
@@
 \caption{
-\bf{Plot of recombination intensity against number
+{\bf Plot of recombination intensity against number
 of recombining gene transfer.}}
@@
 \caption{
-\bf{Plot of recombination intensity against number
+{\bf Plot of recombination intensity against number
 of horizontal gene transfer.}}
@@
 \caption{
-\bf{UCSC genome browser tracks for gene transfer study of
+{\bf UCSC genome browser tracks for gene transfer study of
 SPY and SDE.}}
@@
 \caption{
-\bf{Recombinant tree in the \texttt{ClonalOrigin} model.}
+{\bf Recombinant tree in the \texttt{ClonalOrigin} model.}
 (A) A recombinant tree relating the five genomes with four recombinant edges.
@@
 % Figure S13:
-\begin{figure}
+\begin{figure}[!ht]
 \includegraphics{figures/ri1}
-\caption{\label{fig:ri1}The surrogate measure of recombination intensity
-of genes. Each point represents a pair of a true value of average number of
+\caption{
+{\bf The surrogate measure of recombination intensity
+of genes.} Each point represents a pair of a true value of average number of
 recombinant edges and its estimated value for a gene. The dashed line is with
 slope of 1 and intercept of 0.}
+\label{fig:ri1}
 \end{figure}
+
 \clearpage{}%

### 2011-10-21 doc/manuscript.tex
Just an update.

@@
 % Four potential reviewers whose e-mail addresses you can provide
+% Debra Bessen
 % Claus O. Wilke
-% Debra Bessen
+% Didelot Xavier
 % Bruce R. Levin
@@
 
-Molecular evolution of bacterial genomes is multifaceted. Studies of bacterial
-evolution often focused on one side of many evolutionary processes.  While
-understanding one aspect of bacterial evolutionary processes we could be
+Molecular evolution of bacterial genomes is multifaceted. Researches of
+bacterial evolution often focused on one side of many evolutionary processes.
+While understanding one aspect of bacterial evolutionary processes we could be
 ignorant of the other features.  Horizontal gene transfer process was often
 considered in macro-evolutionary scale using a great deal of bacterial species.
-As closely related species or groups of bacterial individuals are becoming a
-focus of evolutionary study, recombining gene transfer process would surface as
+As closely related species or groups of individual bacteria are becoming a focus
+of evolutionary study, recombining gene transfer process would surface as
 another facet of evolutionary forces.  Because the two gene transfer processes
-would act on closely related bacterial genomes simultaneously, ignoring one of the two processes
-in studying evolution of bacteria would lead to a narrowed point of view to the
-underlying evolutionary forces. Our study of SPY and SDE genomes undertook the
-work of discriminating recombining and horizontal gene transfer
-processes because the two species were very closely related.  We showed that the
-two gene transfer processes acted on different biological functions. 
-Recombining
-gene transfer process in the analysis of SPY and SDE genomes was related with
-genes encoding membrane-bound molecules and aminoacyl tRNA synthetases. On the
-contrary, horizontal gene transfer process was found to be associated with
-transposases, which would allow horizontal gene transfer of their neighboring
-genes.  Because SPY was a human-specific pathogen, and SDE was an opportunistic
-human pathogen, we were interested in association of pathogenicity with the two
-gene transfer processes. We showed that virulence genes were more associated
-with 
-horizontal gene transfer process than with recombining gene transfer process.
-With the current data of only SPY and SDE genomes we could not conclude that the
-association of virulence genes with horizontal gene transfer not recombining
-gene transfer was applicable to other species.  However, without the
-discrimination of the two processes it would have been difficult to find the
-disparate association of biological functions or virulence genes with the
-different gene transfer processes.
-
-Bacterial gene transfer processes as an evolutionary force would affect more
-frequently in some biological functions than in others.  In our attempt of
-finding which biological functions were associated with recombining or
-horizontal gene transfers, we devised a measure of recombination intensity for
-genes by using frequency of recombinant edges that affect those genes, and used
-frequency of horizontal gene transfer events.  This approach allowed us to be
-able to associate gene transfer with biological functions including
-membrane-bound molecules, aminoacyl tRNA synthetases, and ribosomes.  Sugar
-phosphotransferases could be important in signal transduction to recognize
-environmental changes.  Transporters, and fatty acid and lipid biosyntheses
-would be involved in membrane-bound molecules.  Because bacteria as a
-unicellular organism should be able to respond to changes to environment, some
-evolutionary pressures must have been acting on genes encoding membrane-bound
-molecules. Therefoure, recombining gene transfer process appeared to be more
-associated with biological functions that had been critical of sensing queues of
-bacterial environmental changes.  Because several aminoacyl tRNA synthetases
-belonging to gene ontology term of translation (GO:0006412) had been a focus of
-gene transfer in bacterial species \citep{Woese2000}, we could confirm that the
-recombination intensity and requency of recombining gene transfer events would
-be useful in finding genes associated with recombining gene transfer.
-The biological role of transposases was concordant with horizontal gene transfer
-process because the process involved in transposases would allow bacteria to
-acquire foreign genes without the requirement of homologous regions in the
-bacteria.  These differences in the profile of functional categories association
-of the two gene transfer processes were because of the approach of dicriminating
-the two gene transfer processes.
-
-Functional categories concerning ribosomes were unexpectedly found to be
-associated with recombining gene transfer.  Because genes related with ribosomes
-were considered to be housekeeping genes, gene transfer in which would be rare,
-it was unexpected that ribosome related genes were found to be associated with
-recombining gene transfer.  Because it was a bizarre result, we investigated the
-association by considering different recombinant edge types.  Ribosome related
-functional categories were found to be top ranked if we took
-tree-topology-preserving recombinant edges into account (Tables
-S\ref{tab:functional-notopology} and S\ref{tab:functional-all}).  whereas
-ribosome related functional categories were not top ranked in Table
-\ref{tab:functional}.  Therefore, high recombination intensity in ribosome
-related genes were likely to be attributed to topology-preserving recombinant
-edges that were found to be difficult to infer reliably. So, we would have to be
-cautious of associating ribosome related functions with recombining gene
-transfer.
+would act on closely related bacterial genomes simultaneously, ignoring one of
+the two processes in studying evolution of bacteria would lead to a narrowed
+point of view to the underlying evolutionary forces. Our study of SPY and SDE
+genomes undertook the work of discriminating recombining and horizontal gene
+transfer processes because the two species were very closely related.  In this
+study, we showed that the two gene transfer processes acted on different
+biological functions.  Recombining gene transfer process in the analysis of SPY
+and SDE genomes was related with genes encoding membrane-bound molecules and
+aminoacyl tRNA synthetases. On the contrary, horizontal gene transfer process
+was found to be associated with transposases, which would allow horizontal gene
+transfer of their neighboring genes.  Because SPY was a human-specific pathogen,
+and SDE was an opportunistic human pathogen, we were interested in association
+of pathogenicity with the two gene transfer processes. We also showed that
+virulence genes were more associated with horizontal gene transfer process than
+with recombining gene transfer process.  With the current data of only SPY and
+SDE genomes we could not conclude whether the association of virulence genes
+with horizontal gene transfer not recombining gene transfer was applicable
+generally to cases of other species.  However, without the discrimination of the
+two processes it would have been difficult to find the disparate association of
+biological functions or virulence genes with the different gene transfer
+processes.
+
+We showed differences in the profile of functional category association of the
+two gene transfer processes thanks to the approach of dicriminating the two gene
+transfer processes.  Sugar phosphotransferases could be important in signal
+transduction to recognize environmental changes.  Transporters, and fatty acid
+and lipid biosyntheses would be involved in membrane-bound molecules.  Because
+bacteria as a unicellular organism should be able to promptly respond to changes
+to environment, some evolutionary pressures must have been acting on genes
+encoding membrane-bound molecules.  We also found that the recombination
+intensity and requency of recombining gene transfer events would be useful in
+associating several aminoacyl tRNA synthetases genes belonging to gene ontology
+term of translation (GO:0006412) with recombining gene transfer. It had been a
+focus of gene transfer in bacterial species \citep{Woese2000}.  The biological
+role of transposases was concordant with horizontal gene transfer process
+because the process involved in transposases would allow bacteria to acquire
+foreign genes without the requirement of homologous regions in the bacteria.
 
@@
 happened in the pan genome. A model-based method with more refined models would
-be necessary for the reserach direction because the parsimony-based method we
-employed simply could not show any net directionality between SPY and SDE in the
-pan genome. Although we are unsure whether much larger number of genomes could
-be useful in finding net directionality, we consider that more SDE genomes would
-be beneficial to future studies.
+be necessary because the parsimony-based method we employed simply could not
+show any net directionality between SPY and SDE in the pan genome. Although we
+are unsure whether much larger number of genomes could be useful in finding net
+directionality, more SDE genomes might be beneficial to future studies.
 
@@
 were the barriers \citep{Thomas2005} to cause the unbalanced gene transfer in
-the species?  Kalia et al.\ \citet{Kalia2001} discussed restriction-modification system (RMS)
-as a possible mechanism of the unbalanced gene flow.  Restriction-modification
-system is a biological self-defense mechanism of bacteria, which is diverse in
-prokaryote world.  Biological roles of RMS might be to maintain and control
-species identity \citep{Jeltsch2003}.  Recently, Budroni et al.\ \citet{Budroni2011a} studied
-that \textit{Neisseria meningitidis} phylogenetic clades were associated with
-RMS, claiming that RMS modulated homologous recombination in the bacteria of
-\textit{N.\ meningitidis}.  Most of \textit{S.\ pyogenes} genomes contained
-restriction-modification type I system as well as type II system. The two
-genomes of SDE contained only type II system. More tight self-defense system of
-SPY may have made the species less capable of aquiring foreign genetic material
-from SDE, and accordingly relatively unbalanced gene flow between SPY and SDE
-may have been associated with different restriction-modification systems in the
-two species \citep{Kalia2001}.
-
-We reexamined the seven genes that Kalia et al.\ \citet{Kalia2001} had studied for net
-directionality of gene transfer between SPY and SDE.  The internal fragments of
-the six housekeeping genes (\textit{gki}, \textit{gtr}, \textit{murI},
-\textit{mutS}, \textit{recP}, \textit{xpt}) were located in the core genome that
-we investigated in this study.  Except for \textit{mutS} where Kalia et al.\ \citet{Kalia2001}
-did not find recombination, our results were consistent with that of
-Kalia et al.\ \citet{Kalia2001}.  Pinho et al.\ \citet{Pinho2010} reported recombination in the gene
-\textit{parC}, in which we also observed moderate recombination.
-
-A full model-based approach would be desirable in analyzing data for testing
-relevant hypotheses more systematically. Model complexity might not allow easy
-access to the implementation of a model. Our study involved a model-based
-approach and a parsimony-based one for understanding evolution in the core
-genomes of several bacteria individuals. Because of inability of the model-based
-method for studying the pan genomes we had to resort to the parsimony-based
-method. However, the parsimony-based method was also applied to the core
-genomes. We wished to compare the two methods in their results of the core
-genome analyses.  Because of two independent analyses on the same data set, this
-would allow us to be more confident of the results if the two analyses performed
-similarly Generally, phylogenetic methods of detecting gene transfers often take
-advantage of discrepancy between a reference species tree and gene trees. The
-two approaches that we employed in this study were no exception.
+the species?  Kalia et al.\ \citet{Kalia2001} discussed restriction-modification
+system (RMS) as a possible mechanism of the unbalanced gene flow.
+Restriction-modification system is a biological self-defense mechanism of
+bacteria, which is diverse in prokaryote world.  As biological roles of RMS
+might be to maintain and control species identity \citep{Jeltsch2003}, Budroni
+et al.\ \citet{Budroni2011a} recently studied that \textit{Neisseria
+meningitidis} phylogenetic clades were associated with RMS, claiming that RMS
+modulated homologous recombination in the bacteria of \textit{N.\ meningitidis}.
+Most of \textit{S.\ pyogenes} genomes contained restriction-modification type I
+system as well as type II system. The two genomes of SDE contained only type II
+system. More tight self-defense system of SPY may have made the species less
+capable of aquiring foreign genetic material from SDE \citep{Kalia2001}.
+
+We reexamined the seven genes that Kalia et al.\ \citet{Kalia2001} had studied
+for net directionality of gene transfer between SPY and SDE.  The internal
+fragments of the six housekeeping genes (\textit{gki}, \textit{gtr},
+\textit{murI}, \textit{mutS}, \textit{recP}, \textit{xpt}) were located in the
+core genome that we investigated in this study.  Except for \textit{mutS} where
+Kalia et al.\ \citet{Kalia2001} did not find recombination, our results were
+consistent with those of Kalia et al.\ \citet{Kalia2001}.  Pinho et al.\
+\citet{Pinho2010} reported recombination in the gene \textit{parC}, in which we
+also observed moderate recombination.
 
@@
 gene flow between the two species in a more statistically sound way.  We hope
-that we will see thie refinement of the model to better illustrate bacterial
+that we will see this refinement of the model to better illustrate bacterial
 evolution in forthcoming bacterial genomic era.
@@
 
-\section{Supplementary Information}
-
-\subsection{Summarization of posterior samples for local gene trees and recombination intensity}
+\clearpage{}
+\section*{Supplementary Information}
+
+\subsection*{Summarization of posterior samples for local gene trees and recombination intensity}
 
@@
 
-\subsection{Simulation 1 - three population parameters}
+\subsection*{Simulation 1 - three population parameters}
 
@@
 
-\subsection{Simulation 2 - numbers of recombinant edges from prior}
+\subsection*{Simulation 2 - numbers of recombinant edges from prior}
 
@@
 
-\subsection{Simulation 3 - number of recombinant edges from posterior}
+\subsection*{Simulation 3 - number of recombinant edges from posterior}
 

### 2011-10-28 doc/manuscript.tex
Just an update.

@@
 %=============================================================================
+% Table S4-1. The significance values of the heat map.
+\begin{table}[!ht]
+\caption{
+{\bf P-values for the significance of the ratios of logarithm base 2 of the observed
+recombinant edges to the expected number of recombinant edges.} 
+NnN's are for the branch pairs that are impossible under the model-based
+approached. Rows are for donor, and columns for recipient.}
+\noindent \centering{}\begin{tabular}{cccccccccc}
+\hline
+&         SDE1&     SDE&    SDE2&      SD&     SDD &ROOT&    SPY1&     SPY&    SPY2\\
+\hline
+SDE1 &3.1e-03 &    NaN &1.1e-29 &    NaN &8.8e-21  &NaN &4.4e-19 &    NaN &2.2e-23\\
+SDE  &4.9e-02 &2.3e-06 &7.5e-04 &    NaN &3.5e-32  &NaN &9.3e-01 &3.0e-08 &1.2e-25\\
+SDE2 &8.2e-32 &    NaN &6.0e-01 &    NaN &1.2e-23  &NaN &1.6e-17 &    NaN &3.3e-06\\
+SD   &6.4e-11 &8.0e-01 &1.4e-19 &3.2e-02 &7.3e-01  &NaN &2.2e-27 &1.5e-22 &1.1e-30\\
+SDD  &7.3e-33 &6.7e-32 &4.1e-04 &    NaN &9.1e-07  &NaN &8.7e-34 &1.8e-32 &8.7e-34\\
+ROOT &8.7e-34 &8.7e-34 &8.7e-34 &8.7e-34 &8.7e-34  &NaN &8.7e-34 &8.7e-34 &8.7e-34\\
+SPY1 &8.7e-34 &2.9e-32 &8.7e-34 &    NaN &8.7e-34  &NaN &9.3e-01 &    NaN &8.7e-34\\
+SPY  &2.6e-13 &2.2e-27 &5.8e-32 &7.5e-01 &8.7e-34  &NaN &1.3e-03 &3.9e-05 &6.5e-02\\
+SPY2 &8.7e-34 &2.1e-33 &8.7e-34 &    NaN &8.7e-34  &NaN &9.4e-34 &    NaN &9.3e-01\\
+\hline
+\end{tabular}
+\label{tab:heatmappvalue}
+\end{table}
+
+%=============================================================================
 % Table S5. The values of the heat map of actual counts.
@@
 \noindent \begin{centering}
-\begin{tabular}{lll}
-\hline 
-Genes & Gene Accessions\tabularnewline
-\hline 
-putative calcium transporter (MgtA) & SpyM3\_0440 & SPY1 to SDE1\\
-putative dipeptidase & SpyM3\_0465 & SPY to SDE\\
-superoxide dismutase & SpyM3\_1071 & SPY to SDE\\
-putative internalin A precursor & SpyM3\_1035 & SPY to SDE\\
-putative heat shock protein (chaperonin GroEL) & SpyM3\_1765 & SPY to SDE1\\
-pyruvate formate lyase & SpyM3\_1749 & No specific directions\\
-putative glucose kinase & SpyM3\_1180 & SPY to SDE\\
+\begin{tabular}{lrll}
+\hline 
+Direction & Fraction & Genes & Description\\ 
+\hline 
+SDE1-to-SDD & 0.97 & SpyM3\_1369 & putative serine threonine kinase\\
+ & 0.25 & SpyM3\_1370 & putative phosphoprotein phosphatase\\
+SDE2-to-SDD & 1 & SpyM3\_1522 & 3-oxoacyl-acyl carrier protein synthase II\\
+ & 1 & SpyM3\_1523 & 3-ketoacyl-acyl-carrier-protein reductase\\
+ & 1 & SpyM3\_1524 & acyl-carrier-protein S-malonyltransferase\\
+SDD-to-SDE1 & 0.16 & SpyM3\_0499 & F0F1 ATP synthase subunit beta\\
+SDD-to-SDE2 & 0.01 & SpyM3\_1767 & putative endopeptidase Clp ATP-binding chain C\\
+SDD-to-SDE & 0.49 & SpyM3\_1180 & putative glucose kinase\\
+SPY1-to-SDE1 & 0.13 & SpyM3\_0440 & putative calcium transporter\\
+SPY1-to-SDE2 & 0.28 & SpyM3\_1755 & putative pyruvate formate-lyase activating enzyme\\
+SPY2-to-SDE2 & 0.21 & SpyM3\_0420 & lysyl-tRNA synthetase\\
+SPY2-to-SDE & 0.57 & SpyM3\_1806 & DNA mismatch repair protein MutS\\
+SDE-to-SDD & 1 & SpyM3\_1188 & phosphopantetheine adenylyltransferase\\
+ & 0.31 & SpyM3\_0036 & bifunctional acetaldehyde-CoA alcohol dehydrogenase\\
+ & 0.26 & SpyM3\_0499 & F0F1 ATP synthase subunit beta\\
+ & 0.05 & SpyM3\_0497 & F0F1 ATP synthase subunit alpha\\
+SDE-to-SPY1 & 0.02 & SpyM3\_0440 & putative calcium transporter\\
+SDE-to-SPY2 & 0.47 & SpyM3\_1366 & putative two-component response regulator\\
+ & 0.19 & SpyM3\_1475 & N-acetylglucosamine-6-phosphate deacetylase\\
+ & 0.09 & SpyM3\_1201 & putative two-component sensor histidine kinase\\
+ & 0.02 & SpyM3\_1749 & putative pyruvate formate-lyase\\
+SDE-to-SPY & 0.58 & SpyM3\_0499 & F0F1 ATP synthase subunit beta\\
+ & 0.22 & SpyM3\_0497 & F0F1 ATP synthase subunit alpha\\
+ & 0.08 & SpyM3\_1370 & putative phosphoprotein phosphatase\\
+SPY-to-SDE1 & 0.86 & SpyM3\_0511 & ABC transporter ATP-binding protein\\
+ & 0.18 & SpyM3\_1765 & chaperonin GroEL\\
+SPY-to-SPY2 & 0.03 & SpyM3\_1749 & putative pyruvate formate-lyase\\
+SPY-to-SDE & 0.98 & SpyM3\_0465 & putative dipeptidase\\
+ & 0.84 & SpyM3\_1755 & putative pyruvate formate-lyase activating enzyme\\
+ & 0.54 & SpyM3\_1749 & putative pyruvate formate-lyase\\
+ & 0.3 & SpyM3\_0764 & putative spermidineputrescine ABC transporter ATP-binding protein\\
+ & 0.28 & SpyM3\_0420 & lysyl-tRNA synthetase\\
+ & 0.23 & SpyM3\_1182 & putative peroxide resistance protein\\
+ & 0.22 & SpyM3\_0440 & putative calcium transporter\\
+ & 0.07 & SpyM3\_1366 & putative two-component response regulator\\
+ & 0.06 & SpyM3\_1180 & putative glucose kinase\\
 \hline 

### 2011-10-28 pl/convert-gff-ingene.pl

@@ -25,6 +25,7 @@ GetOptions( \%params,
             'verbose',
             'version' => sub { print $VERSION."\n"; exit; },
             'gff=s',
+            'withdescription',
             'out=s'
             ) or pod2usage(2);
 pod2usage(1) if $help;
@@ -145,9 +146,20 @@ sub convert_gff_ingene($$) {
   open GFF, "$gffFilename" or die $!;
   while ($gffLine = <GFF>)
   {
-    if ($gffLine =~ /RefSeq\s+gene\s+(\d+)\s+(\d+).+([+-]).+locus_tag=(\w+);/)
+    if (exists $params{withdescription})
     {
-      print OUT "$4\t$1\t$2\t$3\n";
+      if ($gffLine =~ /RefSeq\s+CDS\s+(\d+)\s+(\d+).+([+-]).+locus_tag=(\w+);.+;product=(.+);/)
+      {
+        my @e = split (/;/, $5);
+        print OUT "$4\t$e[0]\n";
+      }
+    }
+    else
+    {
+      if ($gffLine =~ /RefSeq\s+gene\s+(\d+)\s+(\d+).+([+-]).+locus_tag=(\w+);/)
+      {
+        print OUT "$4\t$1\t$2\t$3\n";
+      }
     }
   }
   close GFF;

### 2011-10-28 pl/remove-blocks-from-core-alignment.pl

@@ -1,29 +1,35 @@
 #!/usr/bin/perl
-
-#===============================================================================
-#   Author: Sang Chul Choi, BSCB @ Cornell University, NY
+###############################################################################
+# Copyright (C) 2011 Sang Chul Choi
 #
-#   File: remove-blocks-from-core-alignment.pl
-#   Date: 2011-02-11
-#   Version: 0.1.0
+# This file is part of Mauve Analysis.
+# 
+# Mauve Analysis is free software: you can redistribute it and/or modify
+# it under the terms of the GNU General Public License as published by
+# the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
 #
-#   Usage:
-#      perl remove-blocks-from-core-alignment.pl [options]
+# Mauve Analysis is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
 #
-#      Try 'perl remove-blocks-from-core-alignment.pl -h' for more information.
-#
-#   Purpose: remove-blocks-from-core-alignment.pl help you to remove alignments
-#            from the core alignment file that is generated by stripSubsetLCBs.
-#===============================================================================
+# You should have received a copy of the GNU General Public License
+# along with Mauve Analysis.  If not, see <http://www.gnu.org/licenses/>.
+###############################################################################
 
 use strict;
 use warnings;
-
-#use Data::Dumper;
 use Getopt::Long;
 use Pod::Usage;
-use File::Temp qw(tempfile);
+require "pl/sub-error.pl";
+require "pl/sub-xmfa.pl";
 
+my $cmd = ""; 
+sub process {
+  my ($a) = @_; 
+  $cmd = $a; 
+}
 $| = 1; # Do not buffer output
 
 my $VERSION = 'remove-blocks-from-core-alignment.pl 0.1.0';
@@ -35,29 +41,144 @@ GetOptions( \%params,
             'help|h',
             'verbose',
             'version' => sub { print $VERSION."\n"; exit; },
-            'fasta=s',
-            'outfile=s',
-            'blocks=s'
+            'threshould=f',
+            'in=s',
[…]

### 2011-10-28 sh/choose-species.sh

@@ -33,6 +33,7 @@ function choose-species {
         mkdir $BASEDIR \
               $BASERUNANALYSIS \
               $NUMBERDIR \
+              $BASEDATADIR \
               $DATADIR \
               $RUNMAUVE \
               $RUNCLONALFRAME \
@@ -47,6 +48,7 @@ function choose-species {
       if [ "$WISH" == "y" ]; then
         ssh -x $CAC_USERHOST mkdir $CAC_BASEDIR \
                                    $CAC_NUMBERDIR \
+                                   $CAC_BASEDATADIR \
                                    $CAC_DATADIR \
                                    $CAC_RUNMAUVE \
                                    $CAC_RUNCLONALFRAME \

### 2011-10-28 sh/convert-gff-ingene.sh

@@ -35,6 +35,10 @@ function convert-gff-ingene {
       perl pl/$FUNCNAME.pl \
         -gff $GFF \
         -out $OUT
+      perl pl/$FUNCNAME.pl \
+        -gff $GFF \
+        -withdescription \
+        -out $OUT.description
       echo "File $OUT is created!"
       break
     fi

### 2011-10-28 sh/filter-blocks.sh

@@ -17,33 +17,6 @@
 # along with Mauve Analysis.  If not, see <http://www.gnu.org/licenses/>.
 ###############################################################################
 
-# 3. Find core alignment blocks.
-# ------------------------------
-# Core alignment blocks are generated from the mauve alignment. This is somewhat
-# iterative procedure. I often was faced with two difficulties. The program,
-# stripSubsetLCBs, is a part of progressiveMauve.  Core alignment blocks were
-# filtered. Often some allignments were a little bizarre: two many gaps are
-# still in some alignment. Another difficulty happens in ClonalOrigin analysis.
-# ClonalOrigin's runs with some alignment blocks did not finish within bearable
-# time, say a month. ClonalOrigin's run with some other alignment blocks were
-# finished relatively fast. It depends on the option of chain lengths. I
-# consider that runs are finished when multiple independent MCMC more or less
-# converged. Those blocks had to be excluded from the analysis. At the first run
-# of this whole procedure, I simply execute the stripSubsetLCBs to find core
-# alignment blocks. I manually check the core alignment blocks to remove any
-# weird alignments. I then proceed to ClonalFrame and the first stage of
-# ClonalOrigin. I find problematic alignment blocks with which ClonalOrigin runs
-# take too much computing time. Then, I come back to filter-blocks to remove
-# them. I had to be careful in ordering of alignment blocks because when I remove any
-# blocks that are not the last all the preceding block numbers have to change.
-#
-# Note that run-lcb alwasys use the output of Mauve alignment. In the 2nd stage
-# of filtering you need to consider proper numbering. Say, in the first
-# filtering of alignment with many gaps you removed the 3rd among 10 alignment
-# blocks. In the 2nd stage, you found 6th alignment needed too long computing
-# time. Then, you should remove 3rd and 7th, not 3rd and 6th because the 6th is
-# actually the 7th among the first 10 alignment blocks. The 6th alignment is 6th
-# among 9 alignment blocks that were from the 1st stage of filtering.
 function filter-blocks {
[…]

### 2011-10-28 sh/init-file-system.sh

@@ -1,15 +1,27 @@
 ###############################################################################
-# Functions: structuring file system (or creating directories)
+# Copyright (C) 2011 Sang Chul Choi
+#
+# This file is part of Mauve Analysis.
+# 
+# Mauve Analysis is free software: you can redistribute it and/or modify
+# it under the terms of the GNU General Public License as published by
+# the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# Mauve Analysis is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with Mauve Analysis.  If not, see <http://www.gnu.org/licenses/>.
 ###############################################################################
-# Create initial directories
-# --------------------------
-# After checking out the source code from the repository output directories need
-# to be created in the local, remote cluster, remote X11 machines.
-# input: nothing
-# output: 3 output directories
 function init-file-system {
   echo -n "Creating $MAUVEANALYSISDIR/output ..." 
   mkdir $MAUVEANALYSISDIR/output 
+  echo -e " done"
+  echo -n "Creating $MAUVEANALYSISDIR/input ..." 
+  mkdir $MAUVEANALYSISDIR/input 
   echo -e " done"
   echo -n "Creating $CAC_ROOT/output at $CAC_USERHOST ..."
   ssh -x $CAC_USERHOST mkdir -p $CAC_ROOT/output

### 2011-10-28 sh/prepare-mauve-alignment.sh

@@ -17,19 +17,6 @@
 # along with Mauve Analysis.  If not, see <http://www.gnu.org/licenses/>.
 ###############################################################################
 
-# 1. I make directories in CAC and copy genomes files to the data directory.
-# --------------------------------------------------------------------------
-# Users need to download genome files to a local directory named $GENOMEDATADIR.
-# They also need to prepare a species file that contains the actual Genbank
-# files. 
-# The first job that a user would want to do is to align the genomes. This would
-# be done in the cluster CAC. The procedure is as follows:
-# 1. Almost all bash variables are set in set-more-global-variable. See the bash function
-# for detail. 
-# 2. mkdir-species creates main file systems.
-# 3. copy-genomes-to-cac copies Genkbank genomes files to CAC.
-# 4. copy-batch-sh-run-mauve creates the batch file for mauve alignment, and
-# copies it to CAC cluster.
 function prepare-mauve-alignment {
   PS3="Choose the species for $FUNCNAME: "
   select SPECIES in ${SPECIESS[@]}; do 
@@ -39,21 +26,23 @@ function prepare-mauve-alignment {
     else  
       echo -n "What repetition do you wish to run? (e.g., 1) "
       read REPETITION
-      echo -e "Wait for muave-analysis file system preparation...\n"
+      echo "Wait for muave-analysis file system preparation..."
       set-more-global-variable $SPECIES $REPETITION
 
-      WALLTIME=$(grep REPETITION${REPETITION}-MAUVE-Walltime species/$SPECIES | cut -d":" -f2)
-      mkdir-species
+      WALLTIME=$(grep REPETITION${REPETITION}-MAUVE-WALLTIME species/$SPECIES | cut -d":" -f2)
+      # mkdir-species
       read-species-genbank-files data/$SPECIES copy-genomes-to-cac
+      read-species-genbank-files data/$SPECIES batch-copy-genome
       copy-batch-sh-run-mauve \
[…]

### 2011-10-28 sh/prepare-run-clonalframe.sh

@@ -17,30 +17,6 @@
 # along with Mauve Analysis.  If not, see <http://www.gnu.org/licenses/>.
 ###############################################################################
 
-# 4. Prepare clonalframe analysis.
-# --------------------------------
-# FIXME: Read the code and document it.
-#
-# NOTE: full_alignment.xmfa has input genome files full paths.
-#       These are the paths that were used in CAC not local machine.
-#       I have to replace those paths to the genome files paths
-#       of this local machine.
-# We could edit the xmfa file, but instead
-# %s/\/tmp\/1073978.scheduler.v4linux\/input/\/Users\/goshng\/Documents\/Projects\/mauve\/$SPECIES\/data/g
-# Also, change the backbone file name.
-# I make the same file system structure as the run-mauve.
-#
-# NOTE: One thing that I am not sure about is the mutation rate.
-#       Xavier said that I could fix the mutation rate to Watterson's estimate.
-#       I do not know how to do it with finite-sites data.
-#       McVean (2002) in Genetics.
-#       ln(L/(L-S))/\sum_{k=1}^{n-1}1/k.
-#       Just remove gaps and use the alignment without gaps.
-#       I may have to find this value from the core genome
-#       alignment: core_alignment.xmfa.
-# NOTE: I run clonalframe for a very short time to find a NJ tree.
-#       I had to run clonalframe twice.
-# NOTE: some of the alignments are removed from the analysis.
 function prepare-run-clonalframe {
   PS3="Choose the species for $FUNCNAME: "
   select SPECIES in ${SPECIESS[@]}; do 
@@ -52,6 +28,7 @@ function prepare-run-clonalframe {
       read REPETITION
       echo -e "  Preparing clonalframe analysis..."
       set-more-global-variable $SPECIES $REPETITION
+      WALLTIME=$(grep REPETITION${REPETITION}-CF-WALLTIME species/$SPECIES | cut -d":" -f2)
       echo -e "  Computing Wattersons's estimates..."
       echo -e "  Removing previous blocks..."
       rm -f $DATADIR/core_alignment.xmfa.*
@@ -61,11 +38,10 @@ function prepare-run-clonalframe {
[…]

### 2011-10-28 sh/read-species-file.sh

@@ -21,9 +21,9 @@ function copy-batch-sh-run-mauve-called {
   isLast=$2
   filename_gbk=`basename $line`
   if [ "$isLast" == "last" ]; then
-    echo "  \$INPUTDIR/$filename_gbk" >> $BATCH_SH_RUN_MAUVE 
+    echo "  \$DATADIR/$filename_gbk" >> $BATCH_SH_RUN_MAUVE 
   else
-    echo "  \$INPUTDIR/$filename_gbk \\" >> $BATCH_SH_RUN_MAUVE 
+    echo "  \$DATADIR/$filename_gbk \\" >> $BATCH_SH_RUN_MAUVE 
   fi
 }
 
@@ -43,10 +43,9 @@ function batch-copy-genome-called {
   cp $line output/$SPECIES/data
 }
 
-
 function copy-genomes-to-cac-called {
   line="$@" # get all args
-  scp -q $GENOMEDATADIR/$line $CAC_USERHOST:$CAC_DATADIR
+  scp -q $line $CAC_USERHOST:$CAC_BASEDATADIR
 }
 
 function processLine {

### 2011-10-28 sh/receive-mauve-alignment.sh

@@ -1,9 +1,24 @@
-# 2. Receive mauve-analysis.
-# --------------------------
-# I simply copy the alignment. 
-# I could copy alignment from other repetition.
+###############################################################################
+# Copyright (C) 2011 Sang Chul Choi
+#
+# This file is part of Mauve Analysis.
+# 
+# Mauve Analysis is free software: you can redistribute it and/or modify
+# it under the terms of the GNU General Public License as published by
+# the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# Mauve Analysis is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with Mauve Analysis.  If not, see <http://www.gnu.org/licenses/>.
+###############################################################################
+
 function receive-mauve-alignment {
-  PS3="Choose the species to analyze with mauve, clonalframe, and clonalorigin: "
+  PS3="Choose the species for $FUNCNAME: "
   select SPECIES in ${SPECIESS[@]}; do 
     if [ "$SPECIES" == "" ];  then
       echo -e "You need to enter something\n"
@@ -11,10 +26,10 @@ function receive-mauve-alignment {
     else  
       echo -n "What repetition do you wish to run? (e.g., 1) "
       read REPETITION
-      echo -e "  Receiving mauve-output...\n"
+      echo "  Receiving mauve-output..."
       set-more-global-variable $SPECIES $REPETITION
-      scp -r $CAC_USERHOST:$CAC_RUNMAUVE/output $RUNMAUVE/
-      echo -e "Now, find core blocks of the alignment.\n"
+      scp -qr $CAC_USERHOST:$CAC_RUNMAUVE/output $RUNMAUVE
+      echo "Now, find core blocks of the alignment."
       break
     fi
   done

### 2011-10-28 sh/run.sh

@@ -145,7 +145,6 @@ CHOICES=( init-file-system \
           ---REAL-DATA-ALIGNMENT---\
           choose-species \
           prepare-mauve-alignment \
-          copy-mauve-alignment \
           receive-mauve-alignment \
           ---REAL-DATA-CLONALFRAME---\
           filter-blocks \
@@ -188,6 +187,7 @@ CHOICES=( init-file-system \
           compute-block-length \
           compute-global-median \
           extract-species-tree \
+          copy-mauve-alignment \
           ---UCSC-GENOME-BROWSER---\
           ucsc-load-genome-not-yet-implemented \
           ---MANUSCRIPT---\

### 2011-10-28 sh/set-more-global-variable.sh

@@ -37,6 +37,7 @@ function set-more-global-variable {
   OUTPUTDIR=$MAUVEANALYSISDIR/output
   BASEDIR=$OUTPUTDIR/$SPECIES
   BASERUNANALYSIS=$BASEDIR/run-analysis
+  BASEDATADIR=$BASEDIR/data
   NUMBERDIR=$BASEDIR/$REPETITION_DIR
   DATADIR=$NUMBERDIR/data
   RUNMAUVE=$NUMBERDIR/run-mauve
@@ -54,6 +55,7 @@ function set-more-global-variable {
   # Note that the cluster base directory does not contain run-analysis. The
   # basic analysis is done in the local machine.
   CAC_BASEDIR=$CAC_OUTPUTDIR/$SPECIES
+  CAC_BASEDATADIR=$CAC_BASEDIR/data
   CAC_NUMBERDIR=$CAC_OUTPUTDIR/$SPECIES/$REPETITION_DIR
   CAC_DATADIR=$CAC_NUMBERDIR/data
   CAC_RUNMAUVE=$CAC_NUMBERDIR/run-mauve

### 2011-10-28 sh/sim6.sh

@@ -183,6 +183,21 @@ function sim6 {
         echo -e "e.g., output/cornellf/3/run-analysis/obsonly-recedge-1.txt"
       fi
 
+      echo -n "Do you wish to count recombination events across blocks (or obsiter) (y/n)? "
+      read WISH
+      if [ "$WISH" == "y" ]; then
+        for g in $(eval echo {1..$HOW_MANY_REPETITION}); do
+          RUNCLONALORIGIN=$BASEDIR/$g/run-clonalorigin
+          RUNANALYSIS=$BASEDIR/$g/run-analysis
+          for h in $(eval echo {1..$HOW_MANY_REPLICATE}); do
+            perl pl/count-observed-recedge.pl obsiter \
+              -d $RUNCLONALORIGIN/output2/${h} \
+              -n $NUMBER_BLOCK \
+              -out $RUNANALYSIS/obsiter-recedge-$h.txt
+          done
+        done
+        echo "Check file $RUNANALYSIS/obsiter-recedge-$REPLICATE.txt"
+      fi
       break
 
       #######################################################

### 2011-10-28 sh/summary-core-alignment.sh

@@ -1,5 +1,21 @@
-# Author: Sang Chul Choi
-# Date  : Sun Jun 19 17:50:19 EDT 2011
+###############################################################################
+# Copyright (C) 2011 Sang Chul Choi
+#
+# This file is part of Mauve Analysis.
+# 
+# Mauve Analysis is free software: you can redistribute it and/or modify
+# it under the terms of the GNU General Public License as published by
+# the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# Mauve Analysis is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with Mauve Analysis.  If not, see <http://www.gnu.org/licenses/>.
+###############################################################################
 
 function summary-core-alignment {
   PS3="Choose the species for $FUNCNAME: "
@@ -13,7 +29,7 @@ function summary-core-alignment {
       set-more-global-variable $SPECIES $REPETITION
 
       echo "  Summarizing the core_alignment.xmfa"
-      echo perl pl/$FUNCNAME.pl \
+      perl pl/$FUNCNAME.pl \
         -in $DATADIR/core_alignment.xmfa \
         -out $RUNANALYSIS/$FUNCNAME.txt
   
@@ -25,4 +41,3 @@ function summary-core-alignment {
     fi
   done
 }
-

### 2011-11-04 doc/manuscript.tex
Just an update.

@@
 \newcommand{\lyxdot}{.}
-\@ifundefined{showcaptionsetup}{}{%
- \PassOptionsToPackage{caption=false}{subfig}}
-\usepackage{subfig}
+
+% \@ifundefined{showcaptionsetup}{}{%
+% \PassOptionsToPackage{caption=false}{subfig}}
+% \usepackage{subfig}
 
@@
 
-\subfloat[]{\includegraphics[scale=0.4]{figures/h3-R-ratio-topology-changing}}
-
-\subfloat[]{\includegraphics[scale=0.4]{figures/h3-R-ratio-not-topology-changing}}
+\includegraphics[scale=0.4]{figures/h3-R-ratio-topology-changing}
+
+\includegraphics[scale=0.4]{figures/h3-R-ratio-not-topology-changing}
 
@@
 \begin{center}
-\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot theta}}
-
-\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot rho}}
-
-\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot delta}}
+\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot theta}
+
+\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot rho}
+
+\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot delta}
 \end{center}

### 2011-11-04 cac/sim/batch_task2.sh

@@ -6,6 +6,16 @@ LOCKFILE=$3
 STATUSDIR=$4
 MAXNODE=$5
 
+# Create a rank to note that the current job is running
+touch $STATUSDIR/$PMI_RANK
+BASESTATUSDIR=$(dirname $STATUSDIR)
+if [ $PBS_ARRAYID -eq 1 ]; then
+  # This means only when all of the jobs were finished, it would be finished.
+  MINIMUMJOB=2 
+else
+  MINIMUMJOB=5
+fi
+
 WHICHLINE=1
 JOBID=0
 LASTNODE=N
@@ -24,8 +34,6 @@ function hms
 
 function checkIfTheJobSuccessfullyFinished 
 {
-  # Check the exit code of the JOBID.
-  # If the job is kill, then put JOBID back to jobidfile.
   OUTPUTFILE=${JOBID##* }
   
   FINISHEDOUTPUT=N
@@ -42,21 +50,13 @@ function checkIfTheJobSuccessfullyFinished
     # echo end-$JOBID
     hms $ELAPSED
   else
-    # The job is not successfully finished.
-    # Put the job command back to the jobidfile.
     while [ "$JOBID" != "" ]; do
       lockfile=$LOCKFILE
       if ( set -o noclobber; echo "$$" > "$lockfile") 2> /dev/null; then
-        # BK: this will cause the lock file to be deleted 
-        # in case of other exit
         trap 'rm -f "$lockfile"; exit $?' INT TERM
-
-        # The critical section:
-        # Write a line, and delete it.
         cp $JOBIDFILE $JOBIDFILE.temp 
         echo $JOBID >> $JOBIDFILE.temp
         mv $JOBIDFILE.temp $JOBIDFILE
-
         rm -f "$lockfile"
         trap - INT TERM
         # Let the job be finished.
@@ -69,98 +69,48 @@ function checkIfTheJobSuccessfullyFinished
   fi
 }
 
-
 # Keep trying to read in jobidfile until the current node is the last one.
-while [ "$LASTNODE" == "N" ]; do
-  # Read the filelock
-  while [ "$JOBID" != "" ]; do
-    lockfile=$LOCKFILE
-    if ( set -o noclobber; echo "$$" > "$lockfile") 2> /dev/null; then
-      # BK: this will cause the lock file to be deleted 
-      # in case of other exit
[…]

### 2011-11-04 cac/sim/run2.sh

@@ -1,5 +1,6 @@
 #!/bin/bash
 rm -rf status
+mkdir status
 echo -n "How many computing nodes do you wish to use? (e.g., 3) "
 read HOW_MANY_NODE
 sed s/PBSARRAYSIZE/$HOW_MANY_NODE/g < batch.sh > tbatch.sh

### 2011-11-04 pl/report-clonalorigin-job.pl

@@ -25,6 +25,7 @@ use XML::Parser;
 use Getopt::Long;
 use Pod::Usage;
 use File::Temp qw(tempfile);
+require "pl/sub-error.pl";
 
 $| = 1; # Do not buffer output
 
@@ -38,6 +39,7 @@ GetOptions( \%params,
             'man',
             'verbose',
             'version' => sub { print $VERSION."\n"; exit; },
+            'samplesize=i',
             'xmlbase=s',
             'database=s'
             ) or pod2usage(2);
@@ -56,6 +58,10 @@ report-clonalorigin-job.pl 0.1.0
 
 perl report-clonalorigin-job.pl [-h] [-help] [-version] 
   [-xml xmlfile] 
+
+perl report-clonalorigin-job.pl -xmlbase dir/core_co.phase3.xml \
+  -database dir2/core_alignment.xmfa \
+  -samplesize 1001
 
 =head1 DESCRIPTION
 
@@ -186,16 +192,27 @@ for (my $blockid = 1; $blockid <= $number_xmfaFile; $blockid++)
     if ($@)
     {
       unlink $xmlFile; 
+      print STDERR "Error in $blockid block $xmlFilebasename:$itercount\n";
       print "$xmlFilebasename:NP\n";
     }
     else
     {
-      # print "$xmlFilebasename:$itercount\n";
+      if ($params{samplesize} == $itercount)
+      {
+        print STDERR "Checked $blockid block $xmlFilebasename:$itercount\r";
+      }
+      else
+      {
+        unlink $xmlFile; 
+        print STDERR "Error in $blockid block $xmlFilebasename:$itercount\n";
+        print "$xmlFilebasename:NP\n";
+      }
     }
   }
   else
   {
-    # print "$xmlFile:$xmlFilebasename:NE\n";
+    print STDERR "Error in $blockid block $xmlFilebasename:$itercount\n";
+    print "$xmlFilebasename:NE\n";
   }
 }
 
@@ -235,26 +252,4 @@ sub characterData {
 
 sub default {
 }
-
-##
-#################################################################################
-### MISC FUNCTIONS
-#################################################################################
-##
-
-sub printError {
-    my $msg = shift;
-    print STDERR "ERROR: ".$msg.".\n\nTry \'report-clonalorigin-job.pl -h\' for more information.\nExit program.\n";
-    exit(0);
-}
-
-sub getLineNumber {
[…]

### 2011-11-04 pl/simulate-data-clonalorigin1-analyze.pl

@@ -126,7 +126,7 @@ require "pl/sub-error.pl";
 ################################################################################
 
 my $in;
-my $out;
+my $outfile;
 my $numbersample;
 my $delta = 0;
 
@@ -136,10 +136,13 @@ if (exists $params{in}) {
   &printError("you did not specify an input file");
 }
 
-if (exists $params{out}) {
-  $out = $params{out};
-} else {
-  &printError("you did not specify an output file");
+if (exists $params{out})
+{
+  open ($outfile, ">", $params{out}) or die "cannot open > $params{out}: $!";
+}
+else
+{
+  $outfile = *STDOUT;   
 }
 
 if (exists $params{numbersample}) {
@@ -218,10 +221,16 @@ sub computeWeightedMedian ($)
                                                   \@meandelta,
                                                   \@lens,
                                                   \@blockIDs);
-    print "$wm[0]\t$wm[2]\t$wm[1]\n";
+    print $outfile "$wm[0]\t$wm[2]\t$wm[1]\n";
   }
   close IN;
 }
+
+if (exists $params{out})
+{
+  close $outfile;
+}
+exit;
 
 sub computeMedianWeighedOverBlockLength ($$$$$)
 {

### 2011-11-04 sh/compute-heatmap-recedge.sh

@@ -51,7 +51,7 @@ function compute-heatmap-recedge {
 
       break
 
-      perl pl/count-observed-recedge.pl \
+      perl pl/count-observed-recedge.pl heatmap \
         -d $RUNCLONALORIGIN/output2/${REPLICATE} \
         -e $RUNCLONALORIGIN/output2/priorcount-${REPLICATE} \
         -endblockid \

### 2011-11-04 sh/kill.sh

@@ -1,0 +1,2 @@
+
+ssh -x $1 ps ax | grep warg | awk '{print $1}'| ssh -x $1 xargs -i kill {} 2&>/dev/null

### 2011-11-04 sh/kill2.sh

@@ -1,0 +1,2 @@
+
+ssh -x $1 ps ax | grep batchjob | awk '{print $1}'| ssh -x $1 xargs -i kill {} 2&>/dev/null

### 2011-11-04 sh/plot-number-recombination-within-blocks.sh

@@ -19,11 +19,10 @@ function plot-number-recombination-within-blocks {
       echo -n "What repetition do you wish to run? (e.g., 1) "
       read REPETITION
       g=$REPETITION
-      echo -n "Which replicate set of output files? (e.g., 1) "
-      read REPLICATE
       set-more-global-variable $SPECIES $REPETITION
+      REPLICATE=$(grep ^REPETITION${REPETITION}-CO2-CO1ID $SPECIESFILE | cut -d":" -f2)
 
-      echo perl pl/$FUNCNAME.pl \
+      perl pl/$FUNCNAME.pl \
         -xmlbase $RUNCLONALORIGIN/output/$REPLICATE/core_co.phase2 \
         -xmfabase $DATADIR/core_alignment.xmfa \
         -thin 1 \
@@ -31,7 +30,7 @@ function plot-number-recombination-within-blocks {
       echo "$RUNANALYSIS/$FUNCNAME-$REPLICATE.out"
 
       NUMBER_BLOCK=$(echo `ls $DATADIR/core_alignment.xmfa.*|wc -l`)
-      NUMBER_SPECIES=$(echo `grep gbk $SPECIESFILE|wc -l`)
+      NUMBER_SPECIES=$(echo `grep gbk data/$SPECIES|wc -l`)
       NUMBER_SAMPLE=$(echo `grep number $RUNCLONALORIGIN/output/$REPLICATE/core_co.phase2.1.xml|wc -l`)
       echo -e "  The number of blocks is $NUMBER_BLOCK."
       echo -e "  The sample size per block is $NUMBER_SAMPLE."

### 2011-11-04 sh/prepare-run-2nd-clonalorigin.sh

@@ -50,7 +50,7 @@ function prepare-run-2nd-clonalorigin {
       NUMBER_BLOCK=$(trim $(echo `ls $DATADIR/$CORE_ALIGNMENT.*|wc -l`))
       JOBIDFILE=$RUNCLONALORIGIN/coii.jobidfile
       rm -f $JOBIDFILE
-      for h in $(eval echo {1..$NREPLICATE}); do
+      for h in $(eval echo {3..$NREPLICATE}); do
         for b in $(eval echo {1..$NUMBER_BLOCK}); do
           TREE=output/$SPECIES/$REPETITION/run-clonalorigin/clonaltree.nwk
           XMFA=output/$SPECIES/$REPETITION/data/core_alignment.xmfa.$b
@@ -106,7 +106,7 @@ function copy-data {
   mkdir \$CLONALORIGINDIR
   cp \$PBS_O_WORKDIR/$SPECIESTREE \$CLONALORIGINDIR
   # Create the status directory.
-  mkdir -p \$PBS_O_WORKDIR/status/\$PBS_ARRAYID
+  mkdir \$PBS_O_WORKDIR/status/\$PBS_ARRAYID
   for h in \$(eval echo {1..$NREPLICATE}); do
     mkdir -p \$CLONALORIGINDIR/output2/\$h
   done
@@ -129,7 +129,7 @@ function process-data {
       \$i \\
       \$PBS_O_WORKDIR/coii.jobidfile \\
       \$PBS_O_WORKDIR/coii.lockfile \\
-      \$PBS_O_WORKDIR/status \\
+      \$PBS_O_WORKDIR/status/\$PBS_ARRAYID \\
       PBSARRAYSIZE&
   done
 }
@@ -137,6 +137,8 @@ function process-data {
 copy-data
 process-data; wait
 retrieve-data
+cd \$PBS_O_WORKDIR
+rm -rf \$TMPDIR
 EOF
       scp -q $RUNCLONALORIGIN/batch.sh \
           $CAC_MAUVEANALYSISDIR/output/$SPECIES/$REPETITION/run-clonalorigin

### 2011-11-04 sh/prepare-run-clonalorigin.sh

@@ -149,7 +149,7 @@ function process-data {
       \$i \\
       \$PBS_O_WORKDIR/coi.jobidfile \\
       \$PBS_O_WORKDIR/coi.lockfile \\
-      \$PBS_O_WORKDIR/status \\
+      \$PBS_O_WORKDIR/status/\$PBS_ARRAYID \\
       PBSARRAYSIZE&
   done
 }
@@ -157,6 +157,8 @@ function process-data {
 copy-data
 process-data; wait
 retrieve-data
+cd \$PBS_O_WORKDIR
+rm -rf \$TMPDIR
 EOF
       scp -q $RUNCLONALORIGIN/batch.sh \
           $CAC_MAUVEANALYSISDIR/output/$SPECIES/$REPETITION/run-clonalorigin

### 2011-11-04 sh/prepare-run-compute-heatmap-recedge.sh

@@ -1,398 +1,0 @@
-# Author: Sang Chul Choi
-# Date  : Mon May  2 13:58:45 EDT 2011
-
-# Prepare runs of computing the heatmap of recombinant edge counts.
-# -----------------------------------------------------------------
-# 1. Copy a set of input data to the cluster. 
-#    (I could use the data in the cluster from run-clonalorigin2)
-# 2. Create scripts for the input data.
-# 3. Create a jobidfile.
-# I wish to extract recedge counts for each iteration.
-# I need to do this for computing sample variance as well.
-function prepare-run-compute-heatmap-recedge {
-  PS3="Choose the species to analyze with mauve, clonalframe, and clonalorigin: "
-  select SPECIES in `ls species`; do 
-    if [ "$SPECIES" == "" ];  then
-      echo -e "You need to enter something\n"
-      continue
-    else  
-      echo -n "What repetition do you wish to run? (e.g., 1) "
-      read REPETITION
-      g=$REPETITION
-      echo -n "Which replicate set of output files? (e.g., 1) "
-      read REPLICATE
-      set-more-global-variable $SPECIES $REPETITION
-
-      NUMBER_BLOCK=$(echo `ls $DATADIR/core_alignment.xmfa.*|wc -l`)
-      NUMBER_SPECIES=$(echo `grep gbk $SPECIESFILE|wc -l`)
-      echo -e "The number of blocks is $NUMBER_BLOCK."
-      echo -e "The number of species is $NUMBER_SPECIES."
-      echo "NUMBER_BLOCK and NUMBER_SAMPLE must be checked"
-
-      # perl pl/extractClonalOriginParameter12.pl \
-      perl pl/compute-heatmap-recedge.pl \
-        -d $RUNCLONALORIGIN/output2/${REPLICATE} \
-        -e $RUNCLONALORIGIN/output2/priorcount-${REPLICATE} \
-        -endblockid \
-        -n $NUMBER_BLOCK \
-        -s $NUMBER_SPECIES
[…]

### 2011-11-04 sh/read-species.sh

@@ -6,7 +6,7 @@ function read-species {
   echo " $HOW_MANY_REPETITION"
 
   echo -n "  Reading REPLICATE from $SPECIESFILE..."
-  HOW_MANY_REPLICATE=$(grep Replicate $SPECIESFILE | cut -d":" -f2)
+  HOW_MANY_REPLICATE=$(grep ^Replicate $SPECIESFILE | cut -d":" -f2)
   if [ "$HOW_MANY_REPLICATE" == "" ]; then
     HOW_MANY_REPLICATE=0
     echo " $HOW_MANY_REPLICATE"

### 2011-11-04 sh/receive-run-2nd-clonalorigin.sh

@@ -27,20 +27,48 @@ function receive-run-2nd-clonalorigin {
       echo -n "What repetition do you wish to run? (e.g., 1) "
       read REPETITION
       g=$REPETITION
-      echo -n "Which replicate set of output files? (e.g., 1) "
-      read REPLICATE
       set-more-global-variable $SPECIES $REPETITION
+      NREPLICATE=$(grep ^REPETITION${REPETITION}-CO2-NREPLICATE species/$SPECIES | cut -d":" -f2)
 
-      echo -e "  Receiving 2nd stage of clonalorigin-output..."
-      mkdir -p $RUNCLONALORIGIN/output2/${REPLICATE}
-      scp -q $CAC_USERHOST:$CAC_RUNCLONALORIGIN/output2/${REPLICATE}/* \
-        $RUNCLONALORIGIN/output2/${REPLICATE}/
+      echo -n "Do you wish to receive the ClonalOrigin's 1st stage MCMC? (y/n) " 
+      read WISH
+      if [ "$WISH" == "y" ]; then
+        echo -e "  Receiving 2nd stage of clonalorigin-output..."
+        for h in 3 4; do
+        #for h in $(eval echo {1..$NREPLICATE}); do
+          # mkdir -p $RUNCLONALORIGIN/output2/${h}
+          # Even if a numbered directory exists, the copy will be done.
+          scp -qr $CAC_USERHOST:$CAC_RUNCLONALORIGIN/output2/${h} \
+            $RUNCLONALORIGIN/output2
+        done
+      else
+        echo "  Skipping copy of the output2 files because I've already copied them ..."
+      fi
+
+      SAMPLESIZE=$(grep ^REPETITION${REPETITION}-CO2-SAMPLESIZE species/$SPECIES | cut -d":" -f2)
+      echo -n "Do you wish to find unfinished blocks in the ClonalOrigin's 2nd stage MCMC? (y/n) " 
+      read WISH
+      if [ "$WISH" == "y" ]; then
+        echo -e "  Finding unfinished blocks ..."
+        for h in $(eval echo {1..$NREPLICATE}); do
+          mkdir $RUNCLONALORIGIN/summary/${h}
+          perl pl/report-clonalorigin-job.pl \
+            -xmlbase $RUNCLONALORIGIN/output2/$h/core_co.phase3.xml \
+            -database $DATADIR/core_alignment.xmfa \
+            -samplesize $SAMPLESIZE \
+            > $RUNCLONALORIGIN/summary/${h}/unfinished2
+        done
+      else
[…]

### 2011-11-04 sh/receive-run-clonalorigin.sh

@@ -1,18 +1,24 @@
-# Author: Sang Chul Choi
-# Date  : Mon Apr 25 15:02:19 EDT 2011
-
-# 7. Receive clonalorigin-analysis.
-# ---------------------------------
-# Note that we can have multiple replicates of clonal origin. 
-# I might want to just compute global median estimates of the first stage of
-# ClonalOrigin.
+###############################################################################
+# Copyright (C) 2011 Sang Chul Choi
 #
-# NOTE: Some of jobs were not finished within reasonable time. It would be
-# better to have all jobs finished. Sometimes it is not possible partly because
-# I would not wait for those to be finished. It is desirable to handle partial
-# output of Clonal Origin. 
+# This file is part of Mauve Analysis.
+# 
+# Mauve Analysis is free software: you can redistribute it and/or modify
+# it under the terms of the GNU General Public License as published by
+# the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# Mauve Analysis is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with Mauve Analysis.  If not, see <http://www.gnu.org/licenses/>.
+###############################################################################
+
 function receive-run-clonalorigin {
-  PS3="Choose the species to analyze real data for clonalorigin1: "
+  PS3="Choose the species for $FUNCNAME: "
   select SPECIES in ${SPECIESS[@]}; do
     if [ "$SPECIES" == "" ];  then
       echo -e "You need to enter something\n"
@@ -22,41 +28,62 @@ function receive-run-clonalorigin {
       read REPETITION
       g=$REPETITION
 
-      echo -n "Which replicate set of output files? (e.g., 1) "
-      read REPLICATE
       set-more-global-variable $SPECIES $REPETITION
-      mkdir -p $RUNCLONALORIGIN/summary/${REPLICATE}
 
[…]

### 2011-11-04 sh/recombination-count.sh

@@ -27,9 +27,8 @@ function recombination-count {
       echo -n "What repetition do you wish to run? (e.g., 1) "
       read REPETITION
       g=$REPETITION
-      echo -n "Which replicate set of ClonalOrigin output2 files? (e.g., 1) "
-      read REPLICATE
       set-more-global-variable $SPECIES $REPETITION
+      NREPLICATE=$(grep ^REPETITION${REPETITION}-CO2-NREPLICATE species/$SPECIES | cut -d":" -f2)
 
       NUMBER_BLOCK=$(echo `ls $DATADIR/core_alignment.xmfa.*|wc -l`)
       NUMBER_SPECIES=$(echo `grep gbk data/$SPECIES|wc -l`)
@@ -40,69 +39,78 @@ function recombination-count {
       echo -n "Do you wish to count recombination events across blocks (or obsiter) (y/n)? "
       read WISH
       if [ "$WISH" == "y" ]; then
-        perl pl/count-observed-recedge.pl obsiter \
-          -d $RUNCLONALORIGIN/output2/${REPLICATE} \
-          -n $NUMBER_BLOCK \
-          -out $RUNANALYSIS/obsiter-recedge-$REPLICATE.txt
-        echo "Check file $RUNANALYSIS/obsiter-recedge-$REPLICATE.txt"
+        for h in $(eval echo {1..$NREPLICATE}); do
+          perl pl/count-observed-recedge.pl obsiter \
+            -d $RUNCLONALORIGIN/output2/${h} \
+            -n $NUMBER_BLOCK \
+            -out $RUNANALYSIS/obsiter-recedge-$h.txt
+          echo "Check file $RUNANALYSIS/obsiter-recedge-$h.txt"
+        done
       fi
 
       echo -n "Do you wish to count recombination events (y/n)? "
       read WISH
       if [ "$WISH" == "y" ]; then
-        perl pl/count-observed-recedge.pl obsonly \
-          -d $RUNCLONALORIGIN/output2/${REPLICATE} \
-          -n $NUMBER_BLOCK \
-          -endblockid \
-          -obsonly \
-          -out $RUNANALYSIS/obsonly-recedge-$REPLICATE.txt
-        echo "Check file $RUNANALYSIS/obsonly-recedge-$REPLICATE.txt"
+        for h in $(eval echo {1..$NREPLICATE}); do
+          perl pl/count-observed-recedge.pl obsonly \
[…]

### 2011-11-04 sh/recombination-intensity1-map.sh

@@ -27,24 +27,26 @@ function recombination-intensity1-map {
       echo -n "What repetition do you wish to run? (e.g., 1) "
       read REPETITION
       g=$REPETITION
-      echo -n "Which replicate set of ClonalOrigin output files? (e.g., 1) " 
-      read REPLICATE
       set-more-global-variable $SPECIES $REPETITION
-
+      NREPLICATE=$(grep ^REPETITION${REPETITION}-CO2-NREPLICATE species/$SPECIES | cut -d":" -f2)
       NUMBER_BLOCK=$(trim $(echo `ls $DATADIR/core_alignment.xmfa.*|wc -l`))
-      RIMAP=$RUNANALYSIS/rimap-$REPLICATE.txt
-      echo -n "Do you wish to generate rimap-$REPLICATE.txt? (e.g., y/n) "
+
+      echo -n "Do you wish to generate rimap-#REPLICATE.txt? (e.g., y/n) "
       read WISH
       if [ "$WISH" == "y" ]; then
-        perl pl/$FUNCNAME.pl \
-          -xml $RUNCLONALORIGIN/output2/${REPLICATE}/core_co.phase3.xml \
-          -xmfa $DATADIR/core_alignment.xmfa \
-          -numberblock $NUMBER_BLOCK \
-          -verbose \
-          -out $RIMAP
+        for h in $(eval echo {1..$); do
+          RIMAP=$RUNANALYSIS/rimap-$h.txt
+          perl pl/$FUNCNAME.pl \
+            -xml $RUNCLONALORIGIN/output2/${h}/core_co.phase3.xml \
+            -xmfa $DATADIR/core_alignment.xmfa \
+            -numberblock $NUMBER_BLOCK \
+            -verbose \
+            -out $RIMAP
+        done
       else
-        echo "  Skipping generating $RIMAP"
+        echo "  Skipping generating rimap files"
       fi
+      break
  
       # echo -e "  The number of blocks is $NUMBER_BLOCK."
       # echo "-------------------------------------"

### 2011-11-04 sh/run.sh

@@ -48,7 +48,6 @@ source sh/progressbar.sh
 source sh/init-file-system.sh
 source sh/choose-simulation.sh 
 source sh/choose-species.sh 
-source sh/simulate-data-clonalorigin1-prepare.sh
 source sh/simulate-data-clonalorigin1-receive.sh 
 source sh/simulate-data-clonalorigin1-analyze.sh
 source sh/receive-run-2nd-clonalorigin.sh
@@ -69,13 +68,14 @@ source sh/simulate-data-clonalorigin2-from-xml.sh
 source sh/probability-recombination.sh
 source sh/map-tree-topology.sh
 source sh/compute-heatmap-recedge.sh
-source sh/prepare-run-compute-heatmap-recedge.sh
 source sh/analyze-run-clonalorigin2-simulation2.sh 
 source sh/sim3-prepare.sh 
 source sh/sim3-receive.sh 
 source sh/sim3-analyze.sh 
-source sh/sim5-prepare.sh 
-source sh/sim6.sh
+source sh/sim1.sh # s15
+source sh/sim2.sh # s18
+source sh/sim3.sh # s16
+source sh/sim5.sh # s17
 source sh/create-ingene.sh
 source sh/convert-gff-ingene.sh 
 source sh/locate-gene-in-block.sh
@@ -112,36 +112,18 @@ source sh/batch.sh
 PS3="Select what you want to do with mauve-analysis: "
 CHOICES=( init-file-system \
           choose-simulation \
-          ---SIMULATION1---\
-          simulate-data-clonalorigin1 \
-          simulate-data-clonalorigin1-prepare \
-          simulate-data-clonalorigin1-receive \
-          simulate-data-clonalorigin1-analyze \
-          ---SIMULATION2---\
-          simulate-data-clonalorigin2 \
-          simulate-data-clonalorigin2-prepare \
-          sim2-receive \
+          ---SIMULATION1-s15,s2---\
+          sim1 \
+          ---SIMULATION2-s18---\
+          sim2 \
           sim2-analyze \
-          simulate-data-clonalorigin2-analyze \
-          analyze-run-clonalorigin2-simulation \
-          ---SIMULATION3---\
-          analyze-run-clonalorigin2-simulation2 \
+          ---SIMULATION3-s16,s3---\
+          sim3 \
           sim3-prepare \
           sim3-receive \
           sim3-analyze \
-          ---SIMULATION5---\
-          sim5-prepare \
[…]

### 2011-11-04 sh/scatter-plot-parameter.sh

@@ -17,12 +17,6 @@
 # along with Mauve Analysis.  If not, see <http://www.gnu.org/licenses/>.
 ###############################################################################
 
-# Plots the three main scalar parameters of clonal origin model.
-# --------------------------------------------------------------
-# The first stage of clonal origin run sample the three main parameters
-# including mutation rate, recombination rate, and average recombinant tract
-# legnth. I parse XML output files of clonal origin to find values of the
-# parameters. The values are plotted for all blocks.
 function scatter-plot-parameter {
   PS3="Choose the species for $FUNCNAME: "
   select SPECIES in ${SPECIESS[@]}; do 
@@ -33,9 +27,8 @@ function scatter-plot-parameter {
       echo -n "What repetition do you wish to run? (e.g., 1) "
       read REPETITION
       g=$REPETITION
-      echo -n "Which replicate set of output files? (e.g., 1) "
-      read REPLICATE
       set-more-global-variable $SPECIES $REPETITION
+      REPLICATE=$(grep ^REPETITION${REPETITION}-CO2-CO1ID $SPECIESFILE | cut -d":" -f2)
 
       perl pl/$FUNCNAME.pl three \
         -xmlbase $RUNCLONALORIGIN/output/$REPLICATE/core_co.phase2 \
@@ -44,7 +37,7 @@ function scatter-plot-parameter {
       echo "$RUNANALYSIS/$FUNCNAME-$REPLICATE.out"
 
       NUMBER_BLOCK=$(echo `ls $DATADIR/core_alignment.xmfa.*|wc -l`)
-      NUMBER_SPECIES=$(echo `grep gbk $SPECIESFILE|wc -l`)
+      NUMBER_SPECIES=$(echo `grep gbk data/$SPECIES|wc -l`)
       NUMBER_SAMPLE=$(echo `grep number $RUNCLONALORIGIN/output/$REPLICATE/core_co.phase2.xml.1|wc -l`)
       echo -e "  The number of blocks is $NUMBER_BLOCK."
       echo -e "  The sample size per block is $NUMBER_SAMPLE."
@@ -57,7 +50,7 @@ function scatter-plot-parameter {
       analyze-run-clonalorigin-scatter-plot-parameter-rscript \
         $RUNANALYSIS/$FUNCNAME-$REPLICATE-out 
  
-      #cat $RUNCLONALORIGIN/summary/${REPLICATE}/median.txt
[…]

### 2011-11-04 sh/sim1.sh

@@ -1,0 +1,270 @@
+###############################################################################
+# Copyright (C) 2011 Sang Chul Choi
+#
+# This file is part of Mauve Analysis.
+# 
+# Mauve Analysis is free software: you can redistribute it and/or modify
+# it under the terms of the GNU General Public License as published by
+# the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# Mauve Analysis is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with Mauve Analysis.  If not, see <http://www.gnu.org/licenses/>.
+###############################################################################
+
+function sim1 {
+  PS3="Choose the species for $FUNCNAME: "
+  select SPECIES in ${SIMULATIONS[@]}; do 
+    if [ "$SPECIES" == "" ]; then
+      echo -e "You need to enter something\n"
+      continue
+    else
+      read-species
+
+      OUTPUTBASE=output/$SPECIES
+      BASEDIR=$OUTPUTDIR/$SPECIES
+      BASERUNANALYSIS=$BASEDIR/run-analysis
+      THETA_PER_SITE=$(grep ThetaPerSite $SPECIESFILE | cut -d":" -f2)
+      RHO_PER_SITE=$(grep RhoPerSite $SPECIESFILE | cut -d":" -f2)
+      DELTA=$(grep Delta $SPECIESFILE | cut -d":" -f2)
+      HOW_MANY_REPETITION=$(grep Repetition $SPECIESFILE | cut -d":" -f2)
+      HOW_MANY_REPLICATE=$(grep Replicate $SPECIESFILE | cut -d":" -f2)
+      NREPLICATE=$HOW_MANY_REPLICATE
+      TREE=$(grep SpeicesTree $SPECIESFILE | cut -d":" -f2)
+      INBLOCK=$(grep InBlock $SPECIESFILE | cut -d":" -f2)
+      NUMBER_BLOCK=$(trim $(echo `cat data/$INBLOCK|wc -l`))
[…]

### 2011-11-04 sh/sim2-analyze.sh

@@ -8,6 +8,7 @@
 # 3. The recombination intensity.
 # I need to come back here because I need to change wargsim to simulate data
 # given recombinant edges.
+
 function sim2-analyze {
   PS3="Choose the simulation result of clonalorigin: "
   select SPECIES in ${SIMULATIONS[@]}; do 

### 2011-11-04 sh/sim2.sh

@@ -246,3 +246,4 @@ function sim6 {
     fi
   done
 }
+

### 2011-11-04 sh/sim3.sh

@@ -1,0 +1,339 @@
+###############################################################################
+# Copyright (C) 2011 Sang Chul Choi
+#
+# This file is part of Mauve Analysis.
+# 
+# Mauve Analysis is free software: you can redistribute it and/or modify
+# it under the terms of the GNU General Public License as published by
+# the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# Mauve Analysis is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with Mauve Analysis.  If not, see <http://www.gnu.org/licenses/>.
+###############################################################################
+
+function sim3 {
+  PS3="Choose the species for $FUNCNAME: "
+  select SPECIES in ${SIMULATIONS[@]}; do 
+    if [ "$SPECIES" == "" ]; then
+      echo -e "You need to enter something\n"
+      continue
+    else
+      read-species
+
+      OUTPUTBASE=output/$SPECIES
+      BASEDIR=$OUTPUTDIR/$SPECIES
+      BASERUNANALYSIS=$BASEDIR/run-analysis
+      THETA_PER_SITE=$(grep ThetaPerSite $SPECIESFILE | cut -d":" -f2)
+      RHO_PER_SITE=$(grep RhoPerSite $SPECIESFILE | cut -d":" -f2)
+      DELTA=$(grep Delta $SPECIESFILE | cut -d":" -f2)
+      HOW_MANY_REPETITION=$(grep Repetition $SPECIESFILE | cut -d":" -f2)
+      HOW_MANY_REPLICATE=$(grep Replicate $SPECIESFILE | cut -d":" -f2)
+      NREPLICATE=$HOW_MANY_REPLICATE
+      TREE=$(grep SpeicesTree $SPECIESFILE | cut -d":" -f2)
+      INBLOCK=$(grep InBlock $SPECIESFILE | cut -d":" -f2)
+      NUMBER_BLOCK=$(trim $(echo `cat data/$INBLOCK|wc -l`))
[…]

### 2011-11-04 sh/simulate-data-clonalorigin1-prepare.sh

@@ -1,149 +1,0 @@
-###############################################################################
-# Copyright (C) 2011 Sang Chul Choi
-#
-# This file is part of Mauve Analysis.
-# 
-# Mauve Analysis is free software: you can redistribute it and/or modify
-# it under the terms of the GNU General Public License as published by
-# the Free Software Foundation, either version 3 of the License, or
-# (at your option) any later version.
-#
-# Mauve Analysis is distributed in the hope that it will be useful,
-# but WITHOUT ANY WARRANTY; without even the implied warranty of
-# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-# GNU General Public License for more details.
-#
-# You should have received a copy of the GNU General Public License
-# along with Mauve Analysis.  If not, see <http://www.gnu.org/licenses/>.
-###############################################################################
-
-# 11. Prepare the first stage of clonalorigin for simulation.
-# -----------------------------------------------------------
-# This may change depending on what simulation setup I would go with.
-# The first simulation set is called c1 at
-# /Users/goshng/Documents/Projects/mauve/noweb/output/c1
-# where there is an input directory. I want to have an output directory for
-# that. I am not following the directory structure in the real data analysis.
-# Instead, I will use the output directory. Replicates are stored in the output
-# directory. I am concerned about that some duplicates of code can be generated.
-# I think that that might be necessary. Why is this the case?
-#
-# mkdir-simulation: a run of a simulation study is stored in a directory,
-# BASEDIR=$MAUVEANALYSISDIR/noweb/output/$CHOICE/output/${REPLICATE}
-# The $CHOICE is c1, and $REPLICATE goes from 1 to as many replicates as you
-# want. The same directories are created at the output directory in the cluster.
-# 
-# At the directory of 
-# BASECHOICEDIR=$MAUVEANALYSISDIR/noweb/output/$CHOICE
-# there is an input directory. A file with extension of fa is the alignment
-# file: e.g., input/c1_1.fa 
[…]

### 2011-11-04 sh/simulate-data-clonalorigin1.sh

@@ -1,55 +1,24 @@
-# Simulates data under ClonalOrigin model or ancestral recombination graph.
-# -------------------------------------------------------------------------
-# The run-lcb contains a list of core_alignment.xmfa.[NUMBER] files.
-# Two kinds of data are generated. I could simulate the ClonalOrigin model to
-# generate a single-block data set or a multiple-block data set. It would be
-# better to use the same bash script structure as that of real data sets. I need
-# an additional analysis step of comparing simulated results and their true
-# values. I would use the same file structure as real data analysis. Is there a
-# problem with that? I remember that there were some issues about this. There
-# was a subtle issue of file structure of the main output directory. I would
-# have a REPETITION under the SPECIES directory. REPETITION and REPLICATE would
-# be confusing. I did not have REPETITON, but I used it: e.g., cornell5 and
-# cornell5-1. Both of them use the same data set. One of them uses 419 or 415
-# blocks, and another uses 411. Their filtering steps were different. From the
-# view point of ClonalOrigin their data sets are different. Replicates are the
-# same analyses but different in time or random seeds. Repetitions are similar
-# analyses with differnt data.
-# 
-# After thinking over REPETITION and REPLICATE I decide to use a different
-# output file structure. The main output directory still contains SPECIES
-# directories. A SPECIES directory would contain directories named as numbers:
-# i.e., 1, 2, 3 and so on. A numbered directory would contain directories such
-# as run-mauve, run-clonalframe, run-clonalorigin, etc. This could change many
-# parts of this main run.sh script. One issue that I was concerned about was how
-# I could run multiple repeated analyses in a single batch script.
+###############################################################################
+# Copyright (C) 2011 Sang Chul Choi
 #
-# Each sub-directory of a species directory contains shell scripts, one of which
-# is called batch.sh. I used to execute it to submit jobs. Now, I wish to
-# control jobs in multiple repetitions. Shell scripts may well be placed at the
-# SPECIES directory. Doing so a batch script can let you submit jobs in REPETITON
-# directories. The SPECIES directory would contain numbered directories, a shell
-# script called run.sh, and a directory called sh that contains more scripts.
[…]

### 2011-11-04 sh/summarize-clonalorigin1.sh

@@ -18,7 +18,7 @@
 ###############################################################################
 
 function summarize-clonalorigin1 {
-  PS3="Choose the species to analyze real data for $FUNCNAME: "
+  PS3="Choose the species for $FUNCNAME: "
   select SPECIES in ${SPECIESS[@]}; do
     if [ "$SPECIES" == "" ];  then
       echo -e "You need to enter something\n"

### 2011-12-08 doc/manuscript.tex
Figures are changed. The manuscript is in the tex file.

--- 2011-10-21 doc/manuscript.tex
@@
-% Four potential reviewers whose e-mail addresses you can provide
-% Debra Bessen
-% Claus O. Wilke
-% Didelot Xavier
-% Bruce R. Levin
-% Michael Wessels
-
-% Template for PLoS
-% Version 1.0 January 2009
+% Potential reviewers and e-mail addresses 
+% DANIEL FALUSH: d.falush@ucc.ie
+% MATTHEW HAHN: mwh@indiana.edu 
+% Debra Bessen: debra_bessen@nymc.edu
+% Xavier Didelot: xavier.didelot@gmail.com
+% Brian Golding: Golding@McMaster.CA
+
+% TODO:
+% Figure S\label{fig:calling-transfers} was not referred in the text.(Delete this?) 
+% Figure S\label{fig:mowgli-sim} was not referred in the text. (Write comments)
+% Table S\label{tab:sim-three} was not referred in the text. (Delete this?)
 % 
-% To compile to pdf, run:
-% latex plos.template
-% bibtex plos.template
-% latex plos.template
-% latex plos.template
-% dvipdf plos.template
-
-\documentclass[10pt]{article}
+%
+% * Finalize section: Distribution of transfer events along the genome
+% * Finalize section: Supp. Materials.
+% Supp. Methods of Matt's modifications of Mowgli's method
+% 
+% Backup of the work
+% swiftgen:/usr/projects/strep/gainloss/recombination
+
+\documentclass[12pt]{article}
+
+\usepackage{times}
 
@@
 % cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
+%\usepackage{cite}
 
@@
 \usepackage{setspace}
-\doublespacing
+
+% To rotate tables using sidewaystable instead of just table.
+\usepackage{rotating}
 
@@
 
+\usepackage{natbib}
+\usepackage{../../latex/sty/genres}
+
 % Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
+\oddsidemargin 0in
+\evensidemargin 0in
+\topmargin -.5in
+\textwidth 6.5in
+\textheight 9in
+%\topmargin 0.0cm
+%\oddsidemargin 0.5cm
+%\evensidemargin 0.5cm
+%\textwidth 16cm 
+%\textheight 21cm
 
@@
 
-% Use the PLoS provided bibtex style
-% \bibliographystyle{plos}
-
 % Remove brackets from numbering in List of References
@@
 % FIXME: Remove the followings if you have to when submitting the manuscript.
-\let\citep\cite
-\let\citet\cite
 \providecommand{\tabularnewline}{\\}
 \newcommand{\lyxdot}{.}
-\@ifundefined{showcaptionsetup}{}{%
- \PassOptionsToPackage{caption=false}{subfig}}
-\usepackage{subfig}
+\newcommand{\comment}[1]{}
+
+% \@ifundefined{showcaptionsetup}{}{%
+% \PassOptionsToPackage{caption=false}{subfig}}
+% \usepackage{subfig}
 
@@
 %% ** EDIT HERE **
-\markright{Gene Transfer using Streptococcus Genomes}
-
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-%% END MACROS SECTION
+\markright{Horizontal gene transfer in Streptococcus}
 
@@
 
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf{Inferring homologous recombination and horizontal gene transfer 
-using whole genomes of \textit{Streptococcus pyogenes} 
-and\textit{ S.\ dysgalactiae} ssp.\textit{\ equisimilis}}
-}
-% Short Title (running head): Gene transfer using Streptococcus genomes
-% Insert Author names, affiliations and corresponding author email.
-\\
-Sang Chul Choi$^{1}$, Matthew D. Rasmussen$^{1}$, 
-Melissa Jane Hubisz$^{1}$, Ilan Gronau$^{1}$,
-Michael J. Stanhope$^{2}$, and Adam Siepel$^{1,\ast}$
-\\
-\bf{1} 
-Department of Biological Statistics and Computational Biology,
+\begin{titlepage}
+
+\title{Replacing and additive horizontal \\
+gene transfer in {\em Streptococcus} }
+
+\author{Sang Chul Choi$^{1}$, Matthew D.\ Rasmussen$^{1}$, 
+Melissa J.\ Hubisz$^{1}$, \\
+Ilan Gronau$^{1}$,
+Michael J.\ Stanhope$^{2}$, and Adam Siepel$^{1}$}
+
+\date{ }
+\maketitle
+
+\begin{footnotesize}
+\begin{center}
+$^1$Department of Biological Statistics and Computational Biology,\\
 Cornell University, Ithaca, NY 14853, USA
-\\
-\bf{2}
-Department of Population Medicine and Diagnostic Sciences,
+\\[1ex]
+$^2$Department of Population Medicine and Diagnostic Sciences,\\
 College of Veterinary Medicine, Cornell University, Ithaca, NY 14853, USA
 \\
-$\ast$ E-mail: acs4@cornell.edu, Phone: 607-254-1157, Fax: 607-255-4698
-\end{flushleft}
-
-% \section*{Things To Do for submitting this to PLoS Pathogens}
-% \begin{enumerate}
-% \item Legends for the main figures as a separate section
-% \item We can add at least 50 words in section \textbf{Abstract}.
-% \item Figures, tables, and supporting information need to be ready.
-% \item Merge Genome Browser tracks to the official one.
-% \item Publish source codes at public domain.
-% \end{enumerate}
+\end{center}
+\end{footnotesize}
+
+\vspace{1in}
+
+\begin{tabular}{lp{4.5in}}
+{\bf Submission type:}& Research Article
+\vspace{1ex}\\
+{\bf Keywords:}&Bacterial evolutionary
+genomics, recombination, {\em Streptococcus pyogenes}, {\em Streptococcus
+  dysgalactiae} 
+\vspace{1ex}\\
+{\bf Running Head:}&Horizontal gene transfer in {\em Streptococcus}
+\vspace{1ex}\\ 
+{\bf Corresponding Author:}&
+\begin{minipage}[t]{4in}
+ Adam Siepel\\
+ 102E Weill Hall, Cornell University\\
+ Ithaca, NY 14853\\
+ Phone: +1-607-254-1157\\
+ Fax: +1-607-255-4698\\
+ Email: acs4@cornell.edu
+\end{minipage}
+\end{tabular}
+
+\thispagestyle{empty}
+\end{titlepage}
+
+\doublespacing
+
+% MBE limit is 350 words
+\section*{Abstract}
+
+The prominent role of horizontal gene transfer (HGT) in the evolution of
+bacteria is now well documented, yet few studies have differentiated
+between the mechanisms of homologous recombination and additive integration.
+%as sources of HGT.  
+These processes leave distinct phylogenetic signatures, because homologous
+recombination causes a gene in one lineage to be replaced by a homolog from
+another lineage (``replacing HGT''), while additive integration causes
+addition without replacement (``additive HGT'').  Here we make use of these
+signatures in a genome-wide investigation of HGT in the important human
+pathogen {\em Streptococcus pyogenes} (SPY) and its close relatives {\em
+  S.\ dysgalactiae} subspecies {\em equisimilis} (SDE) and {\em S.\
+  dysgalactiae} subspecies {\em dysgalactiae} (SDD).  Using recently
+developed statistical models and computational methods, we find evidence
+for abundant gene flow of both kinds within the SPY and SDE clades, and of
+generally reduced levels of exchange between SPY/SDE and SDD.  In addition,
+our analysis strongly supports a previously reported, but unconfirmed,
+finding of a pronounced asymmetry in SPY-SDE gene flow, favoring the
+SPY-to-SDE direction, which may be associated with increasing virulence of
+pathogenic SDE.  We find that this enrichment is driven by replacing
+transfers, and likely reflects frequent homologous recombination between
+co-occurring SPY and SDE cells in human hosts.  We find that virulence
+genes are correlated with transfer events, but this correlation is driven
+by additive, not replacement, HGTs.  The genes affected by additive HGTs
+are enriched for transposition-related functions, while replacement HGTs
+appear to influence a more diverse set of genes.  These findings shed new
+light on the manner in which HGT has shaped pathogenic bacterial genomes
+during relatively recent evolutionary time.
+
+\thispagestyle{empty}
+\clearpage
+
+\section*{Introduction}
+
+During the last twenty years, it has become increasingly apparent that
+horizontal gene transfer (HGT) has played a major role in genomic
+evolution, particularly among bacteria and other microbes
+\citep{Smith1992,Smith1993,Ochman2000,Koonin2001}.  Indeed, it is likely
+that HGT has been sufficiently prevalent throughout evolutionary history
+that no present-day gene can trace an unbroken history of vertical descent
+to a common ancestor for all species \citep{Zhaxybayeva2011}.  This would
+mean that no single phylogenetic marker is a reliable indicator of the tree
+of life.
+
+For HGT to occur in bacteria, DNA must first be transferred from a donor to
+a recipient cell by transformation, transduction, or conjugation, then must
+be integrated into the recipient cell's genome (setting aside the case of
+replication-proficient plasmids).  This integration typically occurs in one
+of two major ways: either the new sequence replaces a homologous sequence
+through the process of homologous recombination (similar to ``gene
+conversion'' in sexually reproducing organisms), or it is acquired through
+an additive (non-replacing) integration process \citep{Thomas2005}.  These
+two types of transfers leave distinct molecular evolutionary signatures and
+will be referred to in this article as ``replacing'' and ``additive'' HGTs,
+respectively (fig.\ \ref{fig:hgt}).  The rates at which these processes
+occur depend on many factors, including the co-occurrence of bacterial
+species in the environment,
+%the propensity of donor cells to release intact DNA,
+the degree of competence of recipient cells, sequence similarity between
+the donor and recipient, and barriers to conjugation such as surface
+exclusion.  In addition, the persistence of horizontally transferred DNA
+segments in present-day species depends strongly on the subsequent effects
+of natural selection.  Not surprisingly, rates of HGT appear to vary
+considerably across groups of bacteria \citep[e.g.,][]{Feil2001}.
+
+HGT in the {\em Streptococcus} genus is of particular interest.  This group
+of Gram-positive bacteria contains the human pathogens {\em S.\ pyogenes}
+(the cause of Group A streptococcal infections, including pharyngitis,
+impetigo, cellulitis, necrotizing fasciitis, and rheumatic fever), {\em S.\
+  pneumoniae} (bacterial pneumonia), {\em S.\ agalactiae} (neonatal sepsis),
+{\em S.\ mutans} (dental caries), and {\em S.\ dysgalactiae} ssp.\ {\em
+  equisimilis} (cellulitis, peritonitis, pneumonia, and other infections),
+as well as the agricultural pathogens {\em S.\ dysgalactiae} ssp.\ {\em
+  dysgalactiae} (mastitis in cows, ewes, and goats), {\em S.\ equi} ssp.\
+{\em equi} (strangles in horses), and {\em S.\ canis} (various infections
+in dogs and other animals).  Many of these species are concentrated in the
+predominantly beta-hemolytic {\em Pyogenes} group.  Despite their
+relatively recent evolutionary divergence, these {\em Pyogenes}-group
+species have adapted to a remarkable variety of distinct ecological niches.
+Accordingly, they display high rates of recombination and HGT, as well as
+substantial evidence of positive selection
+\citep{Feil2001,Marri2006,Anisimova2007,Lefebure2007,Suzuki2011}.  Complete
+genome sequences are now available for several species in this group
+\citep[e.g.,][]{Ferretti2001,Holden2009,Shimomura2011,Suzuki2011},
+providing a rich resource for the study of the genomic evolution of
+pathogenic bacteria over relatively recent time scales.
+
+Within the {\em Pyogenes} group, several studies have focused on the issue
+of HGT between the human-colonizing species {\em S.\ pyogenes} (here
+denoted SPY) and {\em S.\ dysgalactiae} ssp.\ {\em equisimilis} (SDE).
+Until recently, SDE was considered to be primarily a commensal organism
+\citep{Vandamme1996}, unlike SPY, but it has become increasingly apparent
+that it also has an important pathogenetic role, with a disease spectrum
+similar to that of SPY \citep{Brandt2009}.  Shared virulence genes are well
+documented between the two species \citep[e.g.,][]{Davies2007a}, genetic
+exchange presumably having been enabled both by their close evolutionary
+relationship and shared ecological niches, but
+% Towers et al 2004
+% Maxted and Potter, 1967
+% Simpson et al., 1992
+% Sriprakash et al., 1996
+in a widely noted phylogenetic study of seven loci and more than 200
+isolates, \cite{Kalia2001} also found extensive evidence of SPY-SDE gene
+flow among housekeeping genes.  Interestingly, they observed a pronounced
+bias in the direction of gene flow, favoring the SPY-to-SDE direction.
+Gene flow has been confirmed in subsequent studies
+\citep{Kalia2004,Davies2005,Davies2007a,Davies2007}, but the evidence for
+directional asymmetry is mixed, with some indications of gene flow in the
+opposite (SDE-to-SPY) direction \citep{Sachse2002}.  This issue has been
+further complicated by the eventual retraction of Kalia et al.'s
+\citeyearpar{Kalia2001} paper, due to the authors' inability to replicate
+their findings and concerns about sample contamination.  The authors argued
+that whole-genome sequencing may be required to definitively resolve open
+questions about gene flow between these species.  These issues are of
+particular interest because of the possibility of increasing virulence of
+SDE due, at least in part, to gene flow from SPY.
+
+Most recent studies of gene flow in bacteria have been based on relatively
+simple phylogenetic methods that make use of incongruity of inferred gene
+trees across loci \citep[e.g.,][]{Kalia2001,Kalia2004,Ahmad2009}.  These
+methods are sensitive to errors in phylogenetic reconstruction, and ignore
+information from branch lengths and from phylogenetic correlation at
+adjacent loci.  Recently alternative, model-based approaches have been
+proposed for the case of homologous recombination (``replacing HGTs'')
+\citep{Didelot2007,Didelot2010}.  Using Bayesian principles and Markov
+chain Monte Carlo (MCMC) techniques, these methods allow for uncertainty in
+the phylogeny, make use of branch lengths and correlations across loci, and
+allow for inference not only of individual recombination events but also of
+parameters describing global rates and patterns of recombination.  In
+addition, several recent methods have been introduced to detect HGT in a
+phylogenetic framework, by parsimoniously reconciling reconstructed gene
+trees with a given species tree, allowing for duplication, transfer, and
+loss (DTL) events \citep{Merkle2010,David2011,Doyon2011,Doyon2011b}.  These
+new model- and parsimony-based methods are complementary in many respects,
+and could be particularly useful in combination.
+
+In this article, we take a fresh look at the issue of HGT in the {\em
+  Pyogenes} group, making use of newly available complete genome sequences
+for SPY, SDE, and the related species {\em S.\ dysgalactiae} ssp.\ {\em
+  dysgalactiae} (SDD), and new statistical and parsimony-based methods for
+analysis.  The combined use of these methods allows us to examine both
+replacing and additive HGTs, and compare their genome-wide effects.  We
+find evidence for abundant gene flow within SPY, within SDE, and between
+SPY and SDE, and for greatly reduced exchange between SPY/SDE and SDD.  In
+addition, our genome-wide analysis supports Kalia et al.'s
+\citeyearpar{Kalia2001} previously unconfirmed finding of a pronounced
+preference for the SPY-to-SDE direction in SPY-SDE gene flow.
+Interestingly, this property appears to hold for replacing but not for
+additive transfers.  More generally, we find evidence for an association
+between virulence genes and additive, but not replacing, HGTs.  Our results
+have been made available as tracks in our recently released {\em
+  Streptococcus} genome browser.
+
+\section*{Methods}
+
+\subsection*{Genome sequences and alignments}
+
+Our primary analysis was based on five complete genome sequences, including
+two representatives of {\em S.\ pyogenes} (accessions NC\_004070
+[\citealp{Beres2002}; SPY1] and NC\_008024 [\citealp{Beres2006}, SPY2]),
+two of {\em S.\ dysgalactiae} ssp.\ {\em equisimilis} (CP002215
+[\citealp{Suzuki2011}; SDE1] and NC\_012891 [\citealp{Shimomura2011};
+SDE2]), and one of {\em S.\ dysgalactiae} ssp.\ {\em dysgalactiae}
+(AEGO00000000 [\citealp{Suzuki2011}; SDD]).  In addition, we used 
+\textit{S.\ equi} ssp.\textit{\ equi} strain 4047 (NC\_012471
+[\citealp{Holden2009}; SEE]) as an outgroup in our parsimony-based
+analysis (supplementary table S\ref{tab:genome}).
+
+For the model-based analysis of replacing transfers, we obtained
+high-quality alignments of the five genomes (SPY1 \& 2, SDE1 \& 2, and
+SDD) using the 
+pipeline detailed by Didelot and colleagues
+\citep{Didelot2007,Didelot2010}.  Briefly, we first aligned the genomes
+using progressiveMauve v2.3.1 \citep{Darling2004,Darling2010} with default
+options (supplementary fig.\ S\ref{fig:mauve}), and then used stripSubsetLCBs (from the
+ClonalOrigin package; \citealp{Didelot2010}) to identify 276 blocks of
+sequence alignments that exceed a length threshold of 1500 bp.  This
+conservative threshold was selected to avoid biases from boundary effects
+in short alignment blocks, based on a preliminary analysis.  Two blocks
+containing long gaps were excluded from the analysis.
+The final set of 274
+alignment blocks was 1147632 nucleotide long.
+Of 1951 genes of SPY1 genome 1125 genes were covered by the blocks.
+Note that this alignment
+procedure effectively focused on the ``core genome,'' by discarding regions
+that were not conserved among the five species in question.  The 1125 genes
+contained 0.17\% of gaps on average.
+
+For the parsimony-based analysis, we began with 2314 gene family
+clusters representing the six genomes (including the SEE outgroup)
+from \cite{Suzuki2011}.  We aligned protein sequences corresponding to
+each family using MUSCLE \citep{Edgar2004a}, and then obtained
+nucleotide alignments by reverse translation, using the genomic
+sequences as a guide.  Next we identified likely intragenic
+recombinations using the Single Breakpoint Recombination (SBP) method
+from the HyPhy package \citep{KosakovskyPond2006}.  Based on the
+Akaike Information Criterion (AIC), SBP identified at least one
+topology-altering recombination in 47.5\% (1094) of gene
+families. These alignments were split into putatively non-recombining
+gene fragments.  The final set of alignments contained 3408 gene
+fragment families, 1794 of which contained one gene in each of the six
+species.  Families ranged as small as two genes (417 fragments) to as
+large as 59 (2 fragments) (supplementary fig.\ S\ref{fig:famsizes}).  Notably, in
+contrast to the genome-wide alignments produced by progressiveMauve,
+these alignments were representative of the ``pan genome,'' including
+regions that frequently turn over, rather than just the ``core
+genome'' \citep[e.g.,][]{Tettelin2005,Lefebure2007}.
+
+\subsection*{Model-based analysis of replacing HGTs}
+
+{\bf ClonalFrame and ClonalOrigin.}  To study replacing HGTs, we made use
+of the coalescent-based model of homologous recombination and the Markov
+chain Monte Carlo (MCMC) algorithm recently developed by
+\cite{Didelot2010}.  The method was implemented in a 
+program, called ClonalOrigin, that could produce samples from the posterior
+distribution of {\em recombinant trees} given a set of alignment blocks. A
+recombinant tree consists of a rooted tree with branch length, called a clonal
+frame, for the five bacterial genomes,
+which is estimated in a pre-processing step, augmented by a set of
+{\em recombinant edges}. Each recombinant edge is defined by the two points
+along the branches of clonal frame corresponding to the {\em donor} and
+{\em recipient} in the hypothesized recombination event, and by the start
+and end of the genomic interval that is affected by the event.  The donor
+must pre-date the recipient, and a delayed coalescence is possible
+(see \textbf{Supplementary Material} for details of summarizing recombinant
+trees and edges). 
+Another program, called ClonalFrame
+\citep{Didelot2007}, can be used to infer the phylogeny that serves as the
+basis of the recombinant trees.  ClonalFrame makes use of a similar, but
+somewhat simpler probabilistic model, which permits inference
+of the clonal frame. % and a set of model parameters.
+
+{\bf Fitting the model.}  We used
+ClonalFrame (v1.1) to estimate 
+a clonal frame for the five-way
+alignments, running the algorithm for $10^4$ burn-in iterations followed by
+another $10^4$ sampling iterations.  A series of seven additional sampling
+runs with 
+random initializations converged to an essentially identical clonal frame, % model parameters,
+indicating stable convergence.  We then used an initial run of
+ClonalOrigin (subversion r19) to estimate global parameters including
+the mutation rate, recombination rate, and recombinant tract length. We
+applied the method separately to each of the 274 blocks, in all cases using
+$10^6$ burn-in iterations, $10^7$ sampling iterations, and sub-sampling
+every $10^5$ iterations. A global estimate for each parameter was then
+obtained by taking the block-length weighted median of the block-specific
+posterior means, which down-weighed less reliable estimates obtained from
+shorter alignment blocks.  We generally followed \cite{Didelot2010} in 
+fitting the model. % , with some minor differences.
+
+{\bf Sampling recombinant trees}.  Next we ran a second stage of ClonalOrigin
+MCMC
+on all of the alignment blocks by fixing the clonal frame and global model
+parameters at their pre-estimated values, and sampled
+recombinant trees.  For each block, we ran the sampler for $10^7$ burn-in
+iterations and $10^8$ sampling iterations, sub-sampling every $10^5$
+iterations, resulting in 1001 recombinant trees per alignment block.  We
+performed a replicate of the two stages of the ClonalOrigin inference to
+ensure adequate convergence of the sampler.
+
+% A few alignment blocks were not successfully finished even after a month of
+% computation. Because those were not expected to affect the global estimates of
+% population parameters significantly, we used only finished blocks to estimate
+% mutation rate, recombination rate, and recombinant tract length.
+
+{\bf Summary statistics.}  We used various summary statistics to describe
+the sampled recombinant trees (see \textbf{Supplementary Material} for details). 
+First, we considered the distribution of
+tree topologies at each site in the alignments, by computing the site-wise
+relative frequencies of all 105 possible rooted tree topologies relating five
+taxa. Second, we measured frequencies in the posterior samples
+of all possible types of recombinant edge, where an edge
+type is defined by the branches of the clonal frame associated with its
+donor and recipient.
+%also searched the posterior sample for transfer scenarios that were
+%significantly enriched, compared to the prior expectation defined by the
+%clonal frame and the global estimates of the three population
+%parameters. We followed the approach of \cite{Didelot2010} and defined
+%transfer types by the branches of the clonal frame associated with each
+%recombinant edge. 
+%For each transfer type, we recorded the average number of
+%recombinant edges of that type sampled in the posterior sample of
+%recombinant trees, 
+We compared these posterior frequencies with 
+expected values under the prior distribution assumed by the model.
+%as well as the expected number of such edges implied by
+%the prior distribution assumed by the model. 
+%These counts were then used to
+%find types of transfer that showed a high deviation in our inference from
+%their expected counts. 
+To assess significance of the ratio of posterior to prior frequencies of
+recombinant edges we performed a simulation; 
+this empirical approach also 
+had the advantage of correcting for possible biases in the sampling
+process. 
+\textbf{CHECK:}
+We simulated the model to generate 100 data sets, for which we proceeded with
+the inference of the ratio of posterior to prior frequencies of recombinant
+edges. Comparing the obtained null distribution of the ratios from the
+simulation and the ratio from the real data analysis, we assessed the
+significance of the ratio.
+% (GIVE VERY BRIEF SUMMARY OF SIMULATION APPROACH)
+%Ilan: point to supplementary methods for the simulation tests?
+Third,
+we traced the posterior
+probabilities of transfer types in a position-specific manner. 
+%For each
+%distinct type of transfer (determined by a pair of donor and recipient
+%edges), and each site in the core genome alignment, we computed the
+%posterior probability that that site had been affected by a transfer of
+%that type. 
+Finally, we defined a general, branch-independent
+ {\em recombination intensity} at each site by 
+summing the posterior probabilities of recombination.
+In some cases, we also computed intensities for particular types of
+replacing transfers, such as 
+those between the SPY and SDE clades, or those producing topologies
+different from the clonal frame, or \textit{topology-altering} transfers.  
+% CHECK: Justify why we used only topology-altering edges for computing
+% recombination intensity.
+Note that \cite{Didelot2010} described the second and third summarization
+procedures.
+
+\subsection*{Parsimony-based analysis}
+
+% move stuff down about gene tree estimation
+
+{\bf Mowgli.} To study additive HGTs, we used a parsimony-based method
+for reconciling gene and species trees called Mowgli \citep{Doyon2011}.
+Given a gene tree and a species tree, Mowgli finds a reconciliation scenario 
+that minimizes the number of gene duplication, loss, and
+transfer (DLT) events by considering tree topologies with branching orders, also
+known as labeled histories \citep{Edwards1970}.  We estimated
+unrooted gene trees for the nucleotide alignments for each of our
+putatively non-recombining gene fragment using RAxML \citep{Stamatakis2006},
+then applied Mowgli to each of these trees together with the species
+tree estimated by ClonalFrame (augmented with the SEE outgroup).
+Since Mowgli required rooted gene trees, we ran Mowgli for all possible
+rootings of each gene tree, and chose the scenario that produced the
+minimum DLT score. In the case of multiple most parsimonious
+reconciliations, one was chosen uniformly at random.
+
+{\bf Distinguishing between replacing and additive HGTs.}  The transfer
+events inferred by Mowgli had to be classified as replacing or
+additive in a post-processing step because the program did not
+distinguish between these types of events.  A naive approach 
+would be to infer a replacing HGT whenever an inferred transfer
+coincided exactly with an inferred loss in the recipient lineage.  However,
+loss events were difficult to place accurately because they were frequently
+pushed toward the root of the tree due to reconstruction error or parallel
+losses in multiple lineages.  Therefore, we classified a transfer event as
+additive whenever a descendant of the recipient lineage in the species tree
+contained both a gene descended from the transferred copy and a gene not
+descended from that copy.  All other transfers were considered replacing
+HGTs.  This was conservative about calling additive HGTs, and would
+label any HGT that could plausibly be explained by replacing transfers
+as such (see fig.\ \ref{fig:hgt}).
+
+{\bf Statistical enrichment}. As with replacing
+HGTs, we classified transfer events by donor and recipient branch, 
+%sought to identify over-represented scenarios of additive HGT. 
+%and identified distinct types of additive
+%transfers by their donor and recipient branches in the species
+%tree, 
+and recorded the number of inferred additive HGTs of each type.
+%as well as the prior expected number of transfers
+%of that type. 
+We compared these numbers to expectations based on simulations that
+assumed 
+constant
+rates of duplication, loss, and transfer across the species tree, given 
+the total numbers of events inferred across all 2314 gene families
+divided by the total branch lengths of the trees
+(see \textbf{Supplementary Material}).
+
+%Ilan: add some detail about simulations, or point to supp
+
+\subsection*{Gene category associations and virulence genes}
+
+We performed a series of analyses to search for certain classes of genes
+enriched for replacing or additive of gene transfers. In this analysis,
+each gene, or gene family, was associated with a certain score. For
+replacing gene transfers we used the recombination intensity along
+a given gene, and for additive gene transfers we used the number of
+additive transfers inferred for a given family. Mann-Whitney tests were
+used to identify sets of genes that showed significantly elevated values in
+each score.  A 5\% False Discovery Rate (FDR) correction
+\citep{Benjamini1995} was used to determine an appropriate significance
+threshold for our p-values. We applied the association test for gene
+classes defined by Gene Ontology (GO) terms assigned  
+%using the method described in Suzuki et al.\ \cite{Suzuki2011}. In short, 
+by comparing the Streptococcus genes to bacterial proteins from the
+Uniref90 database using blastP, and then assigning the same GO
+classification as the target gene of the uniProt GOA database, if the
+match had an E-values less than $1.0\times10^{-5}$. We also classified a
+category of \textit{virulence genes} using the virulence profile obtained
+by \cite{Suzuki2011}. In both cases, a gene family was
+associsated with a certain classification if any of its genes were
+associated with that classification.
+
+\section*{Results}
+
+\subsection*{Clonal frame and global parameter estimates}
+% The 5-species tree was
+% (((1:0.045557,2:0.045557)6:0.195330,3:0.240887)8:0.089874,(4:0.074544,5:0.074544)7:0.256218)9:0.000000
+By applying ClonalFrame to our five-way genomic alignments 
+(supplementary fig.\ S\ref{fig:mauve}), we obtained a
+phylogeny in which the pairs of SPY and SDE samples were grouped together,
+as expected, and in which SDD and SDE formed a clade, with SPY as an
+outgroup (fig.\ \ref{fig:tree5}).  This tree was consistent with previous
+results from 16S rRNA sequences \citep{Facklam2002} and with genome-wide
+trends \citep{Suzuki2011}.  Despite the grouping of SDD and SDE, their
+estimated divergence was substantial, at $\sim$0.48 substitutions per site.  
+The estimated divergence between SPY and SDD/SDE was $\sim$0.66
+substitutions per site.  
+%These estimates correspond to average
+%genomic divergence times of roughly 660 and 900 thousand years,
+%respectively, assuming a mutation rate of $5 \times 10^{-10}$ substitutions
+%per site per generation \citep{Ochman2003} and two cell divisions per day
+%\citep{XX}.  
+An initial analysis with ClonalOrigin (see \textbf{Methods}) produced an
+estimated
+per-site population-scaled mutation rate of $\theta = $ 0.081
+[interquartile range across blocks: (0.067, 0.094)], a per-site
+population-scaled recombination rate of $\rho = $ 0.012 (0.006, 0.019), and
+an average recombinant tract length of $\delta = $ 744 (346, 2848) bp.  The
+full distributions of these quantities across alignment blocks are shown in
+supplementary
+figure S\ref{fig:scatter3}.  Our estimates implied a ratio of $\rho/\theta
+= 0.15$.  For comparison, \cite{Didelot2010} obtained estimates of $\theta
+= 0.044$, $\rho = 0.017$, $\delta = 236$, and $\rho/\theta = 0.405$ for the
+{\em Bacillus cereus} group, but differences of this magnitude were not
+surprising, given likely differences in effective population sizes,
+recombination rates, and mutation rates \citep[e.g.,][]{Feil2001}.  We
+found that estimates of the recombination tract length, $\delta$, were
+quite sensitive to our threshold for minimum block length, with the
+inclusion of short alignments producing much larger estimates due to edge
+effects.  We experimented with a range of thresholds and selected one (1500
+bp) at which the estimates stabilized.
+
+\subsection*{Model-based analysis of replacing HGTs}
+
+With a second stage of ClonalOrigin MCMC, we were able to obtain samples
+from the posterior distribution of recombinant trees (i.e., the clonal
+frame augmented by recombinant edges) along the genome, conditional on our
+five-way alignments and the previously estimated parameters (see \textbf{Methods}).
+This distribution indicated that most sites (66\%) were
+consistent with the topology of the clonal frame.
+% what does this mean?  most likely tree topology per site?  or expected value?
+Interestingly, the second most frequent tree topology (9.3\%) had
+SPY and SDE as sister clades and SDD as an outgroup (supplementary table
+S\ref{tab:Gene-tree-topologies}), providing an initial indication of gene
+transfer between SPY and SDE.  No other tree topology appeared with
+appreciable frequency.
+% This indicated that SPY and SDE appeared to be more closely related with each
+% other than each to SDD's genome in about tenth of the core genome. The other
+% topologies had less than 4\% of proportions.
+
+% eliminate some redundancy with methods
+To gain further insight into rates and patterns of HGT, we estimated the
+rates of occurrence of recombinant edges, grouping them by the associated
+donor and recipient edges in the clonal frame (see \textbf{Methods}).  Because donor
+edges must pre-date recipient edges under the model, 21 of the 81 edge
+pairs for the nine-branch rooted topology are prohibited, leaving 60
+possible recombinant edge types, each corresponding to
+a class of replacing HGTs.  
+% Because of the possibility of delayed
+% coalescence, these classes include not only transfers from the donor to the
+% recipient edge in the clonal frame, but also transfers from any descendant
+% of the donor to the recipient (with the implicit constraint of
+% contemporaneous donors and recipients).  Thus, recombinant edges are
+% informative about replacing HGTs, but with some ambiguity.  
+Because recombinant edges are not all equally likely \textit{a priori}, 
+we summarized the estimated
+rates by taking ratios with respect to their prior expectations under the
+model used by ClonalOrigin, which consider the branch lengths of the clonal
+frame.  Experiments conducted on simulated data demonstrated that our
+approach could accurately detect recombination scenarios that deviate from
+the prior, with some underestimation of enrichment levels 
+(see \textbf{Supplementary Material}).
+
+These normalized rates for our alignments can be summarized in a $9 \times
+9$ heatmap, consisting of 60 informative and 21 undefined cells (fig.\
+\ref{fig:Heatmap-of-transfers}A; see also supplementary tables
+S\ref{tab:heatmap} and S\ref{tab:obsheatmap}).  Most cells in this heatmap
+fell in the blue-to-white range, corresponding to rates of occurrence less
+than or equal to what was expected under the prior.  Some of these
+correspond to edges that were fairly strongly under-represented (blue),
+such as those between the SDD branch and the branches of the SPY clade
+(P-value of $<0.01$, based on an empirical test; see \textbf{Methods}),
+possibly reflecting ecological isolation of human and strictly veterinary
+pathogens.  By contrast, the recombination edges between the two SDE
+genomes, and between the two SPY genomes, were significantly
+over-represented (P-value of $<0.01$), consistent with previous evidence
+for intraspecies recombination in {\em Streptococcus} \citep{Feil2001}
+(CHECK).  In addition, a pronounced enrichment was observed in
+recombination edges between the SPY and SDE clades, as had been reported
+previously \citep{Kalia2001,Sachse2002,Kalia2004,Davies2005,Davies2007a}.
+While these edges occurred at elevated rates in both directions, we found a
+clear asymmetry, with edges in the SPY-to-SDE direction occuring at up to
+four times the expected rate, while those in the reverse direction occur at
+only about twice the expected rate.  \texttt{CHECK:} To directly compare
+the gene flow of SPY-to-SDE with that of SDE-to-SPY for each of 1001
+recombinant trees from the posterior distribution we summed numbers of
+recombinant edges from each of the three branches SPY1, SPY2, and SPY to
+each of the three SDE1, SDE2, and SDE, and did so for the reverse
+direction.  Taking the ratio of the first sum to the second resulted in a
+distribution of the relative strength of gene flow between the two clades.
+None of the ratio values was less than the corresponding ratio from the
+prior (95\% quantile range was $1.79 <$ posterior ratio $< 2.24$, and prior
+ratio was 1.41; P-value of $< 0.001$).  This provided genome-wide support
+for Kalia et al.'s \citeyearpar{Kalia2001} unconfirmed observations of
+directional asymmetry based on MLST data for seven loci.  We also observed
+a slight enrichment for SDE-to-SDD edges.
+
+% WORK IN DISCUSSION OF TABLE supplementary S\ref{tab:gene-counts-replacing}, SPECIFIC
+% NUMBERS OF SPY$\rightarrow$SDE and SDE$\rightarrow$SDD TRANSFERS.  
+\textbf{CHECK:} We counted genes with noticeable recombination probability 
+($>0.6$) for types of recombinant edges
+(supplementary table S\ref{tab:gene-counts-replacing}). There were more genes
+that underwent SPY-to-SDE replacing gene transfer than those of SDE-to-SPY
+transfer. This recapitulated the argument of asymetry of recplacing gene flow.
+In the subset of virulence genes, there were three genes that underwent
+SPY-to-SDE replacing transfer whereas there was one gene with the reverse
+direction. 
+Virulence genes and non-virulence genes did not seem to be different in the
+relative frequency of replacing gene transfer
+although the proportion of virulence genes with SPY-to-SDE replacing gene
+transfer decreased compared with that of the total genes including the
+virulence genes.
+
+We generated new tracks for our recently released {\em Streptococcus}
+Genome Browser (http://strep-genome.bscb.cornell.edu) that summarize the
+results of the ClonalOrigin analysis alongside known genes, alignments, and
+other annotations (fig.\ \ref{fig:ucsc}).  These tracks can be used to
+inspect loci of interest and to compare the results of our model- and
+parsimony-based analyses.  They can also be queried and intersected with
+other tracks using the UCSC Table Browser.
+
+% follow-up
+% first, give bayesian posterior prob of asymmetry
+% then something about effect of prior
+% then try Ilan's time-interval test
+
+\subsection*{Parsimony-based analysis}
+
+% bring out comparison with replacing (will have to explore more)
+% especially asymmetry, SPY-SDE
+
+To gain further insight into HGT in {\em Streptococcus},
+we estimated gene trees for 2314 gene families representing the six 
+genomes (including an SEE outgroup), used the Mowgli
+program \citep{Doyon2011} to find parsimonious 
+reconciliations of these gene trees with the estimated clonal frame
+(fig.\ \ref{fig:tree5}), and then, in a post-processing step, partitioned
+the inferred transfers into replacing and additive HGTs (see \textbf{Methods}).
+This analysis produced estimated numbers of five types of events on each
+branch of the phylogeny (gene duplications, losses, replacing and additive
+transfers, and ``appearances,'' i.e., transfers from phylogenetically
+distant species), as well as an estimated number of genes at each ancestral
+node of the tree (fig.\ \ref{fig:Gene-duplication-loss}).
+% total numbers of events:
+% appearance: 1957
+% duplication: 257
+% loss: 1293
+% additive: 290
+% replacement 928
+We observed large numbers of appearances on all branches of the tree,
+suggesting a steady influx of genes into the clade by HGT.  Indeed,
+appearance events accounted for 1957 of the 3432 (57\%) gene additions (by
+duplication or transfer).  Interestingly, the somewhat reduced numbers of
+genes in SPY (particularly in SPY1) were primarily explained by reduced rates
+of gene appearance, rather than by increased loss or other factors.  The
+estimated numbers of ancestral genes tended to decrease toward the root of
+the tree, but this likely reflects under-estimation due to parallel losses
+of some genes (especially on the branches beneath the root) rather than a
+true increase in gene number over evolutionary time.  An exception was the
+branch leading to the ancestor of the two SPY individuals, which was
+enriched for gene loss events, perhaps associated with niche adaptation.
+High rates of loss also occurred on the branches to SDE1, SDE2, and SDD.
+Duplications were relatively infrequent overall, but, as had been noted
+previously \citep{Marri2006}, they were substantially enriched on external
+branches of the phylogeny.
+
+%When normalized for branch length (Figure \ref{fig:XX})
+
+%The largest family consisted of 59 genes while nearly half of gene families
+%(1066) had exactly one gene per species, or six genes total (Figure
+%S\ref{fig:famsizes}).
+
+We conservatively identified 290 of the 1218 (23.8\%) transfers (excluding
+appearances) as additive HGTs. 
+As with the replacing HGTs evaluated in the model-based analysis, these
+predicted additive transfers were enriched within the SPY and SDE clades 
+(fig.\ \ref{fig:Heatmap-of-transfers}B).  
+They were also significantly enriched between
+SDE to SDD (with a preference for the SDE-to-SDD direction), and
+significantly depleted for SDD-to-SPY events.
+However, unlike the replacing HGTs above, these additive transfers were only
+slightly enriched between SPY and SDE, and did not exhibit a pronounced
+directional asymmetry.  This suggests that the apparent
+gene flow between SPY and SDE, particularly in the SPY-to-SDE
+direction, has been driven by homologous recombination rather than additive
+transfer.  However, our analyses of replacing and
+additive transfers were not directly comparable for various reasons---for
+example, the first considered the core genome only, while the second
+considered the pan genome; and the two types of analysis might have quite
+different power for different types of events.
+To examine this issue further, we.... (REPORT BRIEF FOLLOW-UP ANALYSIS)
+% follow-up analysis
+We also compared the replacing HGTs identified in the parsimony-based
+analysis with those from the model-based analysis, and found reasonable
+concordance, despite substantial
+differences between the data sets and methods 
+(supplementary fig.\ S\ref{fig:cmpcomowgli}).
+%Ilan: need to refer to appropriate supp figure
+
+\subsection*{Functional Categories of Transferred Genes}
+
+To gain insight into the functional impact of HGT, we assigned genes to
+functional classes and tested each class to see whether it was enriched for
+genes or gene families predicted to have experienced gene transfer events
+(see \textbf{Methods}).  First, we identified a subset of genes whose homologs
+in other
+genera of bacteria were known to exhibit virulence phenotypes (based on
+VFDB; see \citealp{Suzuki2011}) and considered all other genes to be
+non-virulence genes.  These virulence genes were indeed significantly
+enriched among gene families for which we inferred additive HGTs 
+(P-value of $1.33 \times 10^{-11}$), 
+possibly reflecting an increased rate of fixation of
+these transfers due to positive selection for virulence-related functions.
+% too much?
+Interestingly, SPY-to-SDE events were strongly over-represented (by
+$>$2x) among virulence genes that were associated with additive HGTs 
+(supplementary table S\ref{tab:gene-counts-additive}).  
+Transfers in the reverse direction
+(SDE-to-SPY) are also over-represented, but slightly less so
+(1.7x).  By contrast, virulence genes as a group showed no significant
+enrichment for replacing gene transfers, and virulence genes that exhibit
+strong evidence of replacing transfers were not enriched for SDE/SPY % SDE/SDD
+transfers (rather, they were enriched for SDE-to-SDD transfers;
+supplementary table S\ref{tab:gene-counts-replacing}).
+We also found a weak association between virulence genes and gene
+duplications (P-value of $7.13 \times 10^{-3}$).
+
+% (POSSIBLY HIGHLIGHT AN EXAMPLE OR TWO WITH A BROWSER SHOT)
+% Genome browser parts is moved to page 13.
+
+% reduce this to a couple of sentences
+We also 
+searched for gene ontology (GO) terms significantly associated with the
+various types of 
+transfer. Among the categories found to be associated with additive gene
+transfers (table \ref{tab:go-events}), we found several transposase-related 
+transposase.  
+%This can be explained by the role transposition plays in
+%integrating transfered genes into a new host.
+%Ilan: this explanation requires some work
+Replacing transfers appear to show quite different associations from the
+additive ones (table \ref{tab:functional}). (NEED TO EXPAND THIS A BIT)
+%We
+%compared associations found with replacing gene transfers inferred in our
+%model-based analysis with associations found with the number of inferred
+%replacing transfer events in the parsimony analysis, and overall found
+%consistent associations (Table S\ref{tab:go-events-recombining}).
+
+% can we still do something with examples?
+
+%Ilan: this part still under construction
+
+%THE FOLLOWING PARAGRAPH NEEDS SOME WORK
+
+%We used the genome browser tracks for posterior probability of
+%recombination to locate genes which showed an elevated signal of replacing
+%gene transfer. We constructed a list of genes that contain sites showing a
+%posterior probability greater than 0.8 for a given type of transfer
+%(indicated by donor and recipient branches). Each such gene is associated
+%with a transfer type (or several types). We find overall ?? genes fitting
+%these criteria, with ?? of them associated with replacing transfers from
+%SPY to SDE, and ?? associated with replacing transfers from SPY to
+%SDE... Focusing on virulence genes we see a similar distribution of transfer
+%types...
+
+%Ilan: do we want a table with the number of genes associated with each transfer type?
+
+% IS THERE SOMETHING WORTH KEEPING HERE?
+%\texttt{INFORMATIONAL / OPERATIONAL ANALYSIS:}
+%We grouped informational genes based on gene ontology to test if there was
+%differences in replacing gene transfer between informational genes and
+%operational genes. We found differences in replacing gene transfer from three
+%SDE branches to three SPY branches (p-value=0.04409 with Wilcoxon rank sum
+%test).
+%\\
+%\\
+
+\subsection*{Positive Selection and Gene Transfer}
+% selection analysis 
+We also tested for evidence of positive selection within the gene
+families and their association with various events.  We performed a
+likelihood ratio test using PAML's ``sites models'' (M1a and M2b) on
+each of the 2991 gene-fragment families with three or more genes.  Of
+those families tested, 31 showed evidence of positive selection
+(FDR$<$5\%).  For gene trees with high bootstrap ($>$90\% on every
+branch), we find that duplication events are slightly enriched with
+positive selection (P-value of $<0.046$; Mann-Whitney U) as well as
+families with additive transfers (P-value of $<0.002$; Mann-Whitney
+U).  Of the 31 families with significant positive selection, 13
+families were part of the core genome analyzed by ClonalOrigin.  We
+found that replacing transfers from SDE to SPY and from SPY to SDE to
+be slightly enriched within the positively selected families (P-value
+of $<0.005$ and $<0.01$; Mann-Whitney U).  For 11 of the 13
+(85\%) families the SPY-to-SDE direction had greater intensity than
+SDE-to-SPY, consistent with overall directional bias (83.3\%).  
+The same directional bias was not seen for additive
+transfers between the two clades.
+
+% we listed several virulence genes with high ranks (Table S\ref{tab:virhgt}).
+% The whole gene of  putative dipeptidase (SpyM3\_0465) was transferred, which was
+% a exemplary case for the finding \citep{Chan2009} that virulence genes were transferred as a whole
+% unit.  Another example of whole gene transfer was superoxide
+% dismutase (SpyM3\_1071).  We investigated the unit of recombining gene transfer
+% by finding 58 genes for which a recombinant edge of certain types were sampled
+% with probability at least 0.9 by focusing on specific types of recombinant edges
+% (Table S\ref{tab:genes-transfer}).  In some cases, parts of a gene were inferred
+% to have been transferred, for instance, transfer from SPY1 to SDE1 was inferred
+% for 29\% of the aspartyl-tRNA synthetase gene. Recombinant tracts often were
+% shown to span two genes.  In other cases, the inferred transfer units spaned a
+% set of as much as eight adjacent genes: loci from SpyM3\_1518 to SpyM3\_1525.
+
+\section*{Discussion}
+
+% accurate to say "recent"?  has been going on for some time
+There has been a great deal of recent interest in the process of horizontal
+gene transfer (HGT) and the manner in which it has influenced phylogenetic
+relationships, particularly among bacteria \citep{}.  Most previous studies,
+however, have either simply examined patterns of phylogenetic discordance,
+without differentiating between additive and replacing HGTs \citep{}, or
+have focused specifically on the process of homologous recombination,
+corresponding to replacing HGTs
+\citep{Didelot2007,Didelot2010}.
+% too strong?  neither are we
+% cite Feil?   Marri.  LEfebre.  recent review (Doolittle?)
+% another homologous recombination review -- see Awadalla
+To our knowledge, this is the first study to examine both
+processes on a genome-wide scale, using both model- and parsimony-based
+methods.
+
+We have focused our analysis on {\em Streptococcus pyogenes}
+(SPY) and two of its close relatives, {\em S.\ dysgalactiae} ssp.\ {\em
+  equisimilis} (SDE) and {\em S.\ dysgalactiae} ssp.\ {\em dysgalactiae}
+(SDD), a human commensal-like organism and a strict veterinary pathogen,
+respectively.  Our combined analysis reveals strong evidence of gene flow
+both within and between the SPY and SDE groups.  Interestingly, the SPY-SDE
+transfers are far more prominent among replacing than among additive
+transfers, suggesting that they are driven by homologous recombination.
+Moreover, we find strong support for an asymetry in replacing gene
+transfers between SPY and SDE, with a preference for the SPY-to-SDE
+direction.  The situation with SDD is more complex, with somewhat elevated
+rates of SDE-to-SDD transfer, more pronounced for additive than for
+replacing HGTs, mixed evidence for SDD-to-SDE transfer, and reduced rates
+of transfer between SDD and SPY, perhaps owing to a combination of genetic
+divergence and different ecological niches.
+% maybe revise the below after tuning up virulence section
+We also find that virulence genes are significantly enriched among
+additive, but not replacing, gene transfers, as are genes under positive
+selection.  Genes that have undergone replacing transfers between SPY and
+SDE are also enriched for positive selection.
+A central component of our study is the careful use of simulations to
+demonstrate that our methods have good power and accuracy in the
+detection of both types of transfer events.
+
+%We also confirm a species phylogeny (clonal frame) consistent with previous
+%results 
+%\citep{Facklam2002,Suzuki2011}
+
+%paragraph on Kalia et al. and the evidence for SDE/SPY gene flow
+%reasons for asymmetry -- barriers, etc.
+
+The finding of abundant, asymmetric gene flow between
+SPY and SDE raises a number of questions.
+% this is perhaps not so pressing; discuss briefly
+First, can Kalia et al.'s \citeyearpar{Kalia2001} observations be
+confirmed using our data, despite the authors' inability to reproduce them?
+\texttt{CHECK:} 
+Interestingly, we did find 
+that the six of their seven loci represented in our five-way alignments
+(\textit{gki}, \textit{gtr}, 
+\textit{murI}, \textit{mutS}, \textit{recP}, and \textit{xpt}) showed
+consistent results with those of \citet{Kalia2001}; high recombination in 
+\textit{recP}, and \textit{xpt} but not the others. 
+Recombination in \textit{gki} and \textit{mutS}, albeit weak, were also
+observed (despite that Kalia et al.\ found no evidence for {\em mutS}).
+Our visual inspection of the recombination probability on 
+\textit{recP} and \textit{xpt} indicated that 
+the replacing gene transfer appeared to be biased towards
+the directions from SPY to SDE.
+% In S. pyogenes MGAS315 genome track, we can check
+% gki -> not in the region but its neighboring
+% chr1:1207587-1208084
+% /locus_tag="SpyM3_1180"
 % 
-% \section*{A list of our major and minor findings or observations}
-% \begin{enumerate}
-% \item We showed that recombining gene transfer from SPY to SDE was strong
-% enough.
-% \item A few evidences for the net directionality from SPY to SDE:
-% time-constrained gene transfer between SPY to SDE also showed the directionality
-% from SPY to SDE.
-% \item We showed that virulence genes were associated with horizontal gene
-% transfer. 
-% \item We found that recombining gene tranfer processes are involved in
-% membrane-bound protein encoding genes and aminoacyl tRNA synthetases.
-% \item We found that horizontal gene transfer processes are involved in
-% transposases.
-% \item We provided evidence that recombination process affected more sites along
-% the core genome of SPY and SDE than mutation process.
-% \item We found a recombination hotspot at 200 kb from the origin of replication
-% along the genome of SDE1.
-% \item We showed that species-tree-topology-changing recombinant edges were more
-% reliably recovered than species-tree-topology-not-changing edges.
-% \item We showed that recombination intensity might be useful in finding genes
-% under high recombining gene transfer.
-% \item We found that SPY and SDE are more closely related than SDD and SDE in
-% about 10\% of the core genome.
-% \item Recombination intensity and number of recombining gene transfer events
-% were positively correlated.
-% \item We found reduction of numbers of genes after divergence of SDE and SPY.
-% \item We confirmed the species tree that relates the five genomes.
-% \item We listed several virulence genes associated with recombining and
-% horizontal gene transfer.
-% \end{enumerate}
+% gtr -> no recombination 
+% chr1: 1186522-1186073
+% /locus_tag="SpyM3_1160"
 %
-% \section*{Suggestions about the manuscript}
-% \begin{enumerate}
-% \item Ilan thought that a species tree must be shown at the outset of the
-% section \textbf{Results}.
-% \item Can selections after gene transfer be tested?
-% \end{enumerate}
-
-% Please keep the abstract between 250 and 300 words
-\section*{Abstract}
-
-Homologous recombination and horizontal gene transfer are investigated with
-complete genomes of closely related bacterial species of \textit{Streptococcus
-pyogenes} (SPY) and \textit{S.\ dysgalactiae} subspecies \textit{equisimilis}
-(SDE).  The former is a pervasive human-specific pathogen, and the latter an
-opportunistic human pathogen.  We aims at which parts of the genomic
-organization and human pathogenic virulence factors of SDE could have originated
-in those of SPY, or vice versa.  Two mechanisms of gene transfer in bacteria
-were known; recipient bacteria of transferred genes can substitute transferred
-genes for homologous parts of their genomes (Recombining Gene Transfer), or
-integrate transferred genes into their genomes without replacement (Horizontal
-Gene Transfer).  We discriminated the two processes by using a model-based
-approach for studying recombining gene transfer, and using a parsimony-based
-approach for both transfer processes.  Our genome-wide survey of gene transfer
-indicated that gene transfer via homologous recombination occurred more often
-for the net direction from SPY to SDE.  We also found evidence that virulence
-genes transferred more often via non-recombining fashion without replacing parts
-of destination genomes than did non-virulence genes.  
-A genome-wide resource of
-recombining and non-recombining gene transfer for the study of the SPY and SDE
-genomes is provided so that research community in the pathogenic SPY and SDE
-bacteria could take advantage of it.
-
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLoS ONE authors please skip this step. 
-% Author Summary not valid for PLoS ONE submissions.   
-\section*{Author Summary}
-
-\textit{Streptococcus pyogenes} (SPY) has afflicted people around the globe
-causing many diseases. Its closely related species, \textit{S.\ dysgalactiae}
-subspecies \textit{equisimilis} (SDE), used to be known as veterinary pathogens
-have been reported to share disease profiles with SPY.  While gene transfers
-between the two species were shown in small portions of genomic loci, no study
-on the gene transfer at genome-scale has been performed because of lack of
-methodologies implemented or genomic data available. Recent advances in both
-method development and genomic data allowed to study gene transfer between the
-two species in genome scale.  Recipient bacteria of transferred genes can
-substitute transferred genes for homologous parts of their genomes, or integrate
-transferred genes into their genomes without replacement.  The former is called
-recombining gene transfer, and the latter horizontal gene transfer. By
-discriminating the two processes, We found that virulence genes were more
-associated with horizontal gene transfer than recombining gene transfer.  We
-also found a signficant signal of recombining gene transfer for the net
-direction from SPY to SDE in the core genome.  Our study provides new insights
-into gene transfer between SPY snd SDE in genome scale.
-
-\section*{Introduction}
-
-Bacteria infect humans with a variety of diseases: e.g., food poisoning
-(\textit{Eschericha Coli}), strep throat (\textit{Streptococcus pyogenes}),
-tuberculosis (\textit{Mycobacterium tuberculosis}), cholera (\textit{Vibrio
-cholera}), anthrax (\textit{Bacillus anthracis}), pneumonia
-(\textit{Chlamydophila pneumoniae}), and influenza (\textit{Haemophilus
-influenzae}), to name a few.  \textit{Streptococcus pyogenes} (SPY) belonging to
-group A streptococci (GAS) primarily infect humans in the throat and skin
-\citep{Broyles2009}. It has high prevalence throughout the world causing
-diseases ranging from mild illnesses such as pharyngitis to severe invasive
-streptococcal diseases including necrotizing fasciitis and toxic shock syndrome
-\citep{Cunningham2000a}.  Its close relative, \textit{Streptococcus
-dysgalactiae} subspecies \textit{equisimilis} (SDE) belonging to group C or G
-streptococci (GCS/GGS), used to be generally regarded as a veterinary pathogen
-\citep{Vandamme1996}. However, it has been reported to cause serious and
-life-threatening streptococcal diseases to humans traditionally associated with
-GAS \citep{Brandt2009}.  Because of the overlap in disease profile and
-ecological niche of tissue sites betwen SPY and SDE, these bacteria may as
-well interact with each other by exchanging parts of their genomes.
-
-Therefore, the research topic on gene exchange between SPY and SDE has been
-performed in various ways. Sachse et al.\ \citet{Sachse2002} showed gene transfer from SDE to
-SPY for the downstream of the genomic region of streptococcal pyogenic
-exotoxins. Davies et al.\ \citet{Davies2007} investigated phage 3396 from SDE as a vehicle of
-gene transfer from SDE to SPY. They \citet{Davies2009} also demonstrated that an
-integrative conjugative element of SDE allowed gene transfer from SDE to other
-streptococci. Ahmad et al.\ \citet{Ahmad2009} did not conclusively deduce the net
-directionality of housekeeping gene transfer between SPY and SDE.  Notably,
-Kalia et al.\ \citet{Kalia2001} attempted to show that the net direction of gene transfer from
-SPY to SDE was more dominant than the other direction.  Because these researches
-on the controversial issue, net directionality of gene transfer between SPY and
-SDE, have been limited to relatively small regions in the genomes primarily on
-representative loci from the Multi Locus Sequence Typing (MLST) database, it
-remains to be elusive to determine the genome-wide degrees to which genomic
-segments flowed between the two species, and net directions of thereof.
-
-As sequencing costs plummet \citep{Mardis2011}, bacterial genome sequencing
-produces individual genomes of high quality even within a bacterial species
-\citep{Tettelin2009a}.  Studies on horizontal gene transfer in prokaryotes have
-used genes in very large sets of species from a great deal of genomes available 
-\citep{Koonin2001}.  Although this type of study gave us insights into deep
-evolutionary histories of prokaryotes, with multiple individual genomes per
-species we now can go beyond such rough broad study on gene transfer towards
-fine-scaled history of gene transfer between very closely related prokaryotes.
-For instance, Liu et al.\ \citet{Liu2009} studied gene transfer events in
-\textit{Lactobacillus bulgaricus} and \textit{Streptococcus thermophilus} using
-the compositional approach of Karlin \citet{Karlin2001}. Luo et al.\ \citet{Luo2011} investigated
-gene transfer events in \textit{Eschericha Coli} using the embedded quartet
-decomposition due to Zhaxybayeva et al.\ \citet{Zhaxybayeva2006}. Hamady et al.\ \citet{Hamady2006} combined
-phylogenetic and compositional approaches to detecting horizontal gene transfer.
-Caro-Quintero et al.\ \citet{Caro-Quintero2011} studied a number of \textit{Shewanella baltica}
-genomes to identify the genetic elements that enabled the species to adapt to
-redox gradients.  Although several complete genomes from SPY have been
-available, Suzuki et al.\ \citet{Suzuki2011} just recently added one more complete genome of
-SDE to the set of a previously unique SDE genome \citep{Shimomura2011}. These
-multiple SDE genomes prompted us to study genome-wide gene flows between SPY and
-SDE that afflicted human beings around the globe.
-
-Bacteria evolves new biological features through horizontally transferred genes
-from other lineages and vertically transmitted mutations on genes within the
-same lineage.  While mutations on a gene modifiy relatively small portions of the 
-gene through substitutions, insertions, and deletions of deoxyribonucleic acids,
-gene transfers can reshape the organization of a bacterial genome,
-which could revamp biological functions of the species.  Recipient bacteria of
-transferred genes can substitute transferred genes for homologous parts of their
-genomes, or integrate transferred genes into their genomes without replacement.
-These two processes are here in this manuscript discriminated by calling the
-former \textit{recombining gene transfer}, and the latter \textit{horizontal
-gene transfer} \citep{Ochman2001,Lawrence2009}.  Despite the two distinct
-evolutionary processes in bacteria, analyses of complete genome sequences in
-efforts of better understanding bacterial evolution have paid less attention to
-a combined approach to illuminating disparity of the two gene transfer
-processes. We attempted here to use and compare two approaches for
-distinguishing the recombining gene transfer process from the horizontal, or
-non-recombining, gene transfer, and tackled the two processes in association with
-virulence genes.
-
-Recently, a method of inferring homologous recombination with whole bacterial
-genomes was developed \citep{Didelot2010}. The model of homologous recombination
-allowed them to study recombining gene flows between bacteria individuals of
-\textit{Bacillus cereus} species in the core genomes.   In the model, a species
-tree that relates individuals genomes is augmented by attaching
-\textit{recombinant edges} between species tree branches that can be used to
-measure gene transfers.  Other evolutionary processes acting on bacterial
-genomes include genome rearragement, translocation, inversion, duplication,
-gain, loss, etc.  These processes did not readily lend themselves to
-statistically more sound model-based approaches such as maximum likelihood.
-Instead, a parsimony-based approach to inferring gene duplication, loss, and
-transfer events was developed by Doyon et al.\ \citet{Doyon2011}.  In the
-approach, gene trees are inferred for gene families that are constructed from
-all of the genomes, and are reconciled with a speices tree to infer
-parsimoniously events.  We noted that the first approach could be used to infer
-transfer of genomic segments due to homologous recombination, and the second
-could be useful in studying horizontal gene transfer.
-
-In short, the complete genomes of the two species of SPY and SDE presented us a
-unique opportunity to study gene transfer over relatively recent time scales.
-We took advantage of the model-based approach for inferring homologous
-recombination, and the parsimony-based one for detecting gene duplication, loss,
-and horizontal transfer.  We found that virulence genes were more associated
-non-recombining horizontal gene transfer than non-virulence genes.  We compared
-both of the results in recombining gene transfer events to show that the
-parsimony-based method inferred recombining gene transfer events in biological
-functional categories similar to those inferred with the model-based approach.
-This was the first attempt as far as we know of systematically analyzing SPY and
-SDE in genome-wide scale to detect gene transfer events not only in the core
-genome but also in the pan genome, which made the analysis complete in terms of
-the amount of used genome data. Our analysis of the core genomes also showed for
-the first time that the net directionality of recombining gene transfer from SPY
-to SDE was stronger at the genome scale.
-
-% Results and Discussion can be combined.
-\section*{Results}
-
-% \texttt{We show a net directionality of recombining gene transfers from SPY to SDE. 
-% Describe our arguments of the net directionality. 
-% Number of recombinant edges.
-% Confidence about the numbers of recombinant edges.
-% Not all of genes experience recombining gene transfers from SPY to SDE.
-% Gene transfer counts from the parsimony-based approach; the power of the
-% parsimony-based approach may be limited.
-% Is this number of recombinant edges good enough for the argument?
-% Reflection of the previous analyses, 
-% heat maps of real data analysis, 
-% and supporting simulations.}
-
-\subsection*{Recombining gene transfer from SPY to SDE at genome-wide scale}
-
-We followed the pipeline and methods recommended by Didelot et al.
-\cite{Didelot2010} to pursue a model-based study of recombining gene transfer. A
-preliminary step in the pipeline involved estimating the species tree with
-branch lengths via \texttt{ClonalFrame} \cite{Didelot2007}, and three population
-parameters of mutation rate, recombination rate, and average recombinant tract
-length via a first phase of \texttt{ClonalOrigin} \cite{Didelot2010}.  Because
-\textit{S.\ equi} ssp.\textit{\ equi} strain 4047 (SEE) diverged significantly
-from the others among the six genomes (Table S\ref{tab:genome}), we could not
-obtain sufficiently long multiple sequence alignments with it, and had to
-exclude it from the model-based analysis.  The estimated species tree using the
-alignment of the five genomes (Figure S\ref{fig:mauve}) had the expected
-topology \texttt{SC: The figure 1 must be nicer than one that now} (see Figure
-\ref{fig:tree5}). \texttt{ClonalFrame} also estimated a relative strength of
-recombination compared to mutations as $r/m=4.66$ meaning that four times more
-sites had been affected by recombination-driven substitutions than by
-mutation-driven substitutions (Table S\ref{tab:clonalframe}).  This indicated
-potentially significant roles of recombination in the evolutionary process of
-shaping these genomes.  Estimates of population mutation rate ($\theta$),
-recombination rate ($\rho$), and average tract length ($\delta$) were obtained
-for the alignment blocks analyzed (Figure S\ref{fig:scatter3}), and a median of
-the estimates weighted over block lengths was obtained as a global estimate for
-each of the three parameters (Table S\ref{tab:three}).  The local estimates
-typically seemed to be moderately fluctuating around the global estimate, with
-the exception of what appeared to be a recombination hotspot at a region 200
-kilobase pairs from the origin of replication of SDE1.
-
-The estimated species tree together with the global estimates of the three
-population parameters were used to sample recombinant trees from the posterior
-distribution in the second phase of \texttt{ClonalOrigin} \cite{Didelot2010}.
-We summarized the posterior sample of recombinant trees by counting topologies
-of the local genealogies induced by those recombinant trees for each of the
-genomic positions (Table S\ref{tab:Gene-tree-topologies}).  Of the 105 possible
-rooted topologies relatig five taxa, the dominant topology (66.7\%) was that of
-the species tree.  The second most frequently sampled topology (9.3\%) was the
-one in which SPY and SDE were sister clades to each other with SDD being as an
-outgroup. Other topologies were sampled with frequency less than 4\%. This gave
-initial indication to the extent of recombining gene transfer  between SPY and
-SDE.
-
-To gain a fine-scale view of the main types of recombination scenarios, we
-looked at recombinant edges present in the posterior sample of recombinant
-trees. Following the approach of Didelot et al. \cite{Didelot2010}, we
-classified these edges by types defined by distinct pairs of donor and recipient
-branches of the species tree (Figure \ref{fig:tree5}). For each edge type, we
-averaged the number of edges present in the sampled recombinant tree, and
-computed the ratio of this observed average frequency to a prior expected
-frequency. The prior expected frequencies was computed by considering branch
-lengths and their relative positions of the species tree, and the three
-population parameter estimates in the preliminary step.  Base 2 logarithms of
-these ratios are depicted in the heatmap of Figure
-\ref{fig:Heatmap-of-recombination} (see also Tables S\ref{tab:heatmap} and
-S\ref{tab:obsheatmap}).
-
-Among the possible recombinant edge types, we can distinguish between ones that
-preserve the species tree topology, and ones that alter it. We focused on 
-topology-altering edge types both because our main objective was to study
-cross-clade gene transfer, and because our simulation study indicated that the
-observed counts of topology-preserving recombinant edges were unreliably
-inferred (see \textbf{Simulations}). The heatmap allowed to find recombination
-scenarios that deviated from the expectation dictated by the probabilistic
-model.  We envisioned two main potential causes for such deviations: (1) reduced
-or increased opportunities for species to interact and exchange genetic
-material; (2) selection acting on newly transferred genes.
-
-We observed reduction in the number of recombinant edges sampled between SDD and
-the three branches of the SPY clade ($\sim$25\% of the expected count). This was
-consistent with the fact that these species had few opportunities to interact
-with each other, SPY being a human pathogen and SDD being a strictly veterinary
-one.  On the contrary, the number of sampled recombinant edges between SDD and
-branches in the SDE clade seemed to be consistent with the prior.  The greatest
-deviation of observed counts from expected ones was in recombinant edges between
-SPY and SDE clades. It was concordant with the fact that these two species
-shared an ecological niche; SDE was an opportunistic human pathogen.
-Interestingly, there appeared to be much higher enrichment (up to four times
-more than expected) for recombinant edges originating in the SPY clade and
-ending up in the SDE clade.  In a follow-up simulation study we concluded that
-this deviation from the expected count was indeed highly significant, and so was
-the observed asymmetry in pattern of recombination (see \textbf{Simulations}).
-This net directionality of recombination pattern from SPY to SDE was especially
-intriguing in light of the hypothesis that SDE acquired human pathogenicity via
-gene transfer from SPY.
-
-\subsection*{Parsimony-based reconciliation of gene trees with the species tree}
-
-Figure \ref{fig:Gene-duplication-loss} shows inferred numbers of gene gain,
-loss, duplication, and two different kinds of gene transfer events along
-lineages of the tree that relates the six genomes. SPY had less genes than the
-other species. All of the six genomes were used to construct 2314 gene families.
-The largest family consisted of 59 genes while nearly half of gene families
-(1066) had exactly one gene per species, or six genes total (Figure
-S\ref{fig:famsizes}).  We estimated a larger number of recombining gene transfer
-events than that of horizontal gene transfer events; 944 and 294, respectively.
-Duplications in extant lineages were more frequent than those in the other
-lineages.  SDE lineages from the root appeared to have increased in the number
-of genes while the common ancestor of SPY1 and SPY2 decreased in the number of
-genes.  Whereas gene loss events were either comparable to or less than gene
-gain events along most of the lineages, it was not the case along the two internal
-branches of SPY and SDE.  This indicated that SPY and SDE lineages had reduced
-number of genes to specialize in their specific niches after each divergence
-from their corresponding direct ancestors.
-
-As in the model-based analysis, we attempted to see which type of horizontal or
-recombining transfer events was most common in our inference.  We classified
-events into types defined by donor-recipient branches in the species tree. We
-recorded the number of inferred events of each type (Figures
-S\ref{fig:hgt-heatmap} and S\ref{fig:mowgli-recomb-heatmap}). Unlike in the
-model-based analysis, we could not normalize these counts by their expectation
-because we did not have any model-based expectation for them, which restricted
-our ability to interpret these counts. We did notice, however, a large number of
-inferred recombining gene transfer events from the ancestral SPY branch to the
-ancestral SDE branch, consistent with our observations from the model-based
-analysis. Yet, we did not observe any pattern of a net directional horizontal
-gene transfer while we observed a large number of horizontal transfers from the
-ancesral SDE branch to SDD.
-
-\subsection*{Gene categories associated with recombining and horizontal gene transfer}
-
-We set out to seek categories of genes that were significantly enriched for
-recombining or horizontal transfer events. For horizontal transfer, we used a
-score based on a number of events inferred by the parsimony-based method, and
-for recombining gene transfer, we used both the parsimony-based inferred counts
-and a recombination intensity score based on the model-based analysis (see
-\textbf{Materials and Methods}). First, we checked association of transfers with
-virulence genes, which were provided by a previous study of Suzuki et al.\
-\citet{Suzuki2011}.  The set of virulence genes was found to be significantly
-associated with horizontal gene transfers (P-value $<1.33 \times 10^{-11}$), but
-not with recombining gene transfers. We also found a marginally significant
-association between virulence genes and gene duplications (P-value $<7.13 \times
-10^{-3}$).
-
-We conducted similar tests for other functional categories of genes. Several
-functional categories that were found to be associated with recombination
-intensity were related with either membrane-bound molecules or aminoacyl tRNA
-synthetases (Table \ref{tab:functional}). Consistent results were obtained for
-the parsimony-based counts of recombining gene transfers (Table
-S\ref{tab:go-events-recombining}).  Functional categories associated with
-horizontal gene transfer were clearly different from those associated with
-recombining gene transfer (Table \ref{tab:go-events}). The top ranked functional
-categories in that list were related with transposases. These differences added
-to our previous findings about the different roles the two types of gene
-transfer had played in the evolution of these species.
-
-\subsection*{Comparison of model-based and parsimony-based inferences of recombining gene transfer}
-
-The similarity between the two lists of gene categories found to be associated
-with recombining gene transfers in the separate analyses -- model-based and
-parsimony-based --  indirectly indicated concordance between the two independent
-approaches.  In order to obtain a more direct measure of correlation, we used
-the set of genes present in both analyses to compute the correlation between
-recombination intensity scores and the number of recombining gene transfers
-inferred by the parsimony-based approach. We observed a positive correlation of
-$0.39$ (P-value $2.2\times10^{-16}$; Figure S\ref{fig:cmpcomowgli}) between the
-two measures.  For comparison, the correlation between recombination intensity
-and the number of horizontal gene transfers was 0.03 (P-value 0.18;
-Figure S\ref{fig:cmpcomowglihgt}).
-
-\subsection*{A closer study of virulence genes}
-
-Often computational and statistical analyses on genomes could benefit other
-researchers when those resources were available in a form of genome browser.  We wished to
-provide research community especially in SPY and SDE bacteria with results from
-our analysis of gene transfer.  The resource shown in Figure S\ref{fig:ucsc} for
-gene transfer was available at \url{http://strep-genome.bscb.cornell.edu}.  For
-each of the five individual genomes as a reference genome several annotation
-tracks were displayed: alignment blocks, recombination rate per site for each
-block, local gene tree topologies, recombination intensity, locations of
-virulence genes, posterior probability of recombining gene transfer. The genome
-track of probability of recombining gene transfer allowed us to easily browse
-the bacterial genome to locate regions that we were interested in.  Using
-the genome browser tracks we located virulence genes with notable posterior
-recombination probability (Table S\ref{tab:virrecomb}).  About 10\% of the genes
-comprised of virulence facters, about half of which were found in the core
-genome of the five genome alignment \citep{Suzuki2011}.  After sorting virulence
-genes by the number of horizontal gene transfer events we listed several genes
-with high ranks (Table S\ref{tab:virhgt}).
-The whole gene of  putative dipeptidase (SpyM3\_0465) was transferred, which was
-a exemplary case for the finding \citep{Chan2009} that virulence genes were transferred as a whole
-unit.  Another example of whole gene transfer was superoxide
-dismutase (SpyM3\_1071).  We investigated the unit of recombining gene transfer
-by finding 58 genes for which a recombinant edge of certain types were sampled
-with probability at least 0.9 by focusing on specific types of recombinant edges
-(Table S\ref{tab:genes-transfer}).  In some cases, parts of a gene were inferred
-to have been transferred, for instance, transfer from SPY1 to SDE1 was inferred
-for 29\% of the aspartyl-tRNA synthetase gene. Recombinant tracts often were
-shown to span two genes.  In other cases, the inferred transfer units spaned a
-set of as much as eight adjacent genes: loci from SpyM3\_1518 to SpyM3\_1525.
-
-\subsection*{Simulation study and additional support for biased directionality
-in recombining gene transfer between SPY and SDE}
-
-We conducted a series of experiments on simulated data to check the reliability
-of estimates obtained by the model-based approach. We found that the population
-parameters obtained in the first phase were reliable, with the exception that
-the average recombinant tract length was over-estimated (this was observed also
-in \citep{Didelot2010}). Our experiments also showed that the recombination
-intensity used in our functional category association tests had fairly good
-correlation between the true value and the simulated ones.  (Figure
-S\ref{fig:ri1}). See \textbf{Materials and Methods} for a more detailed report
-of these results. One thing we specifically wished to test in these experiments
-was whether the model-based approach reliably detected deviations of recombinant
-edge counts from their expectation, as observed in the heatmap of Figure
-\ref{fig:Heatmap-of-recombination}. Our major concern was that these deviations
-were not directly accounted for by the probabilistic model of ClonalOrigin, but
-rather indirectly observed by summary of the posterior sample of recombinant
-trees.
-
-We simulated the model to generate a hundred simulated data sets using a
-collection of recombinant trees sampled from the prior distribution defined by
-the species tree and three population parameters, as estimated from our genomic
-data.  We then repeat the inference procedure of the second stage of
-\texttt{ClonalOrigin} on each data set and, as in the data analysis, and counted
-for each donor-recepient branch pair the average number of recombinant edges.
-By normalizing the counts by the prior expected count we obtained intervals of
-base 2 logarithms of the ratio of the observed number of recombinant edges to
-the prior expectionation (Figure \ref{fig:h3}). The comparison showed a good
-concordance of the estimated counts and the prior expected counts for all
-topology-altering edge types and a clear over-estimation for the counts of
-topology-preserving edges. Recall that we made sure to ignore these edges in our
-analysis. For each edge type, we also plotted in Figure \ref{fig:h3} the
-log-ratio of the count estimated for it in our data analysis and the expected
-count (the closed circles without interval bars); these are the values making up
-the heatmap of Figure \ref{fig:Heatmap-of-recombination}. Comparing the counts
-estimated in the data analysis and the counts estimated from the simulated data
-allowed to evaluate the significance of the divergence from expectation observed
-in the heatmap.  Recall the heatmap indicated an excess of recombinant edges
-between the SPY and SDE clades in both directions. Our simulations indicated
-that the excess observed in the direction SDE to SPY was much less significant,
-as the counts estimated from real data often falled very close to the
-corresponding counts estimated from simulated data. The excess observed in the
-direction SPY to SDE appeared not only to be larger, but also more significant.
-This further supported our claim for the net directionality in recombining gene
-transfer between these two species.
-
-We wished to use simulated data to test not only the ability of the ClonalOrigin
-sample to \textit{detect} deviations from expected counts of recombinant edges, but
-also its ability to reliably \textit{quantify} this deviation. For this purpose we
-generated a hundred simulated data sets using a collection of recombinant trees
-sampled from the posterior sample of ClonalOrigin on the genomic data. The
-difference between the two simulated batches was that in the first one, the
-recombinant trees used represent the prior distribution, and in the second one,
-they represented the posterior distribution, which we knew deviated from the
-prior. Again, we recorded the recombinant edge counts obtained by ClonalOrigin
-and normalized these counts by the prior expected counts (intervals with closed
-circles in in Figure \ref{fig:h3}). We expected these counts to be in better agreement with
-the counts observed in the data analysis. For recombinant edges that showed a
-defficiency in our data analysis (e.g., from the SPY clade to SDD), the counts
-appeared to fit well with the count estimated from genomic data. However, for
-recombinant edges that showed an excess in our data analysis (e.g., from the SPY
-clade to the SDE clade), the counts appeared to be significantly influenced by the
-prior. We noted that the influence of the prior appeared to be even stronger for
-topology-preserving recombinant edges. All things considered, we found that defficiency of
-a certain recombinant edge type was reliably estimated whereas excesses were 
-typically underestimated. Therefore, we concluded that the actual extent to which
-recombining transfers between SPY and SDE \textit{preferred} the direction from SPY to
-SDE was possibly somewhat underestimated in our analysis.
-
-\section*{Discussion}
-
-Molecular evolution of bacterial genomes is multifaceted. Researches of
-bacterial evolution often focused on one side of many evolutionary processes.
-While understanding one aspect of bacterial evolutionary processes we could be
-ignorant of the other features.  Horizontal gene transfer process was often
-considered in macro-evolutionary scale using a great deal of bacterial species.
-As closely related species or groups of individual bacteria are becoming a focus
-of evolutionary study, recombining gene transfer process would surface as
-another facet of evolutionary forces.  Because the two gene transfer processes
-would act on closely related bacterial genomes simultaneously, ignoring one of
-the two processes in studying evolution of bacteria would lead to a narrowed
-point of view to the underlying evolutionary forces. Our study of SPY and SDE
-genomes undertook the work of discriminating recombining and horizontal gene
-transfer processes because the two species were very closely related.  In this
-study, we showed that the two gene transfer processes acted on different
-biological functions.  Recombining gene transfer process in the analysis of SPY
-and SDE genomes was related with genes encoding membrane-bound molecules and
-aminoacyl tRNA synthetases. On the contrary, horizontal gene transfer process
-was found to be associated with transposases, which would allow horizontal gene
-transfer of their neighboring genes.  Because SPY was a human-specific pathogen,
-and SDE was an opportunistic human pathogen, we were interested in association
-of pathogenicity with the two gene transfer processes. We also showed that
-virulence genes were more associated with horizontal gene transfer process than
-with recombining gene transfer process.  With the current data of only SPY and
-SDE genomes we could not conclude whether the association of virulence genes
-with horizontal gene transfer not recombining gene transfer was applicable
-generally to cases of other species.  However, without the discrimination of the
-two processes it would have been difficult to find the disparate association of
-biological functions or virulence genes with the different gene transfer
-processes.
-
-We showed differences in the profile of functional category association of the
-two gene transfer processes thanks to the approach of dicriminating the two gene
-transfer processes.  Sugar phosphotransferases could be important in signal
-transduction to recognize environmental changes.  Transporters, and fatty acid
-and lipid biosyntheses would be involved in membrane-bound molecules.  Because
-bacteria as a unicellular organism should be able to promptly respond to changes
-to environment, some evolutionary pressures must have been acting on genes
-encoding membrane-bound molecules.  We also found that the recombination
-intensity and requency of recombining gene transfer events would be useful in
-associating several aminoacyl tRNA synthetases genes belonging to gene ontology
-term of translation (GO:0006412) with recombining gene transfer. It had been a
-focus of gene transfer in bacterial species \citep{Woese2000}.  The biological
-role of transposases was concordant with horizontal gene transfer process
-because the process involved in transposases would allow bacteria to acquire
-foreign genes without the requirement of homologous regions in the bacteria.
-
-Another contribution of our study to bacterial gene transfer research is to
-elucidate the net directionality of gene flow from SPY to SDE in the core
-genomes. Although the issue of net directionality has been tackled, our study
-was the first attempt of resolving the issue in genome-wide scale. Yet, it still
-remains to be seen whether this net directionality of gene flow from SPY to SDE
-happened in the pan genome. A model-based method with more refined models would
-be necessary because the parsimony-based method we employed simply could not
-show any net directionality between SPY and SDE in the pan genome. Although we
-are unsure whether much larger number of genomes could be useful in finding net
-directionality, more SDE genomes might be beneficial to future studies.
-
-Were there unbalanced recombining gene transfer between the two species, what
-were the barriers \citep{Thomas2005} to cause the unbalanced gene transfer in
-the species?  Kalia et al.\ \citet{Kalia2001} discussed restriction-modification
-system (RMS) as a possible mechanism of the unbalanced gene flow.
-Restriction-modification system is a biological self-defense mechanism of
-bacteria, which is diverse in prokaryote world.  As biological roles of RMS
-might be to maintain and control species identity \citep{Jeltsch2003}, Budroni
-et al.\ \citet{Budroni2011a} recently studied that \textit{Neisseria
-meningitidis} phylogenetic clades were associated with RMS, claiming that RMS
-modulated homologous recombination in the bacteria of \textit{N.\ meningitidis}.
-Most of \textit{S.\ pyogenes} genomes contained restriction-modification type I
-system as well as type II system. The two genomes of SDE contained only type II
-system. More tight self-defense system of SPY may have made the species less
-capable of aquiring foreign genetic material from SDE \citep{Kalia2001}.
-
-We reexamined the seven genes that Kalia et al.\ \citet{Kalia2001} had studied
-for net directionality of gene transfer between SPY and SDE.  The internal
-fragments of the six housekeeping genes (\textit{gki}, \textit{gtr},
-\textit{murI}, \textit{mutS}, \textit{recP}, \textit{xpt}) were located in the
-core genome that we investigated in this study.  Except for \textit{mutS} where
-Kalia et al.\ \citet{Kalia2001} did not find recombination, our results were
-consistent with those of Kalia et al.\ \citet{Kalia2001}.  Pinho et al.\
-\citet{Pinho2010} reported recombination in the gene \textit{parC}, in which we
-also observed moderate recombination.
-
-It remains to be seen whether the inability of detecting net directionality
-between SPY and SDE in the pan genomes was because of lack of power of the
-parsimony-based approach that we employed. Considering gene duplication, loss,
-and horizontal gene transfer in a model-based approach would be desirable in the
-coming ages of abundant bacterial genomes. The model might consider genome
-alignments where each taxon does not necessarily have a unique gene; some taxa
-might have no gene due to gene loss, and other taxa multiple genes due to gene
-gain, duplication, or horizontal gene transfer. With these types of models will
-we be able to address more interesting evolutionary questions in bacteria. This
-need of more refined models would be also true even for the study of net
-directionality between species in the core genome.  In other words, one could
-envision a model where a population two-taxa (i.e., SPY and SD groups) tree is
-superimposed on the species tree shown in Figure S\ref{fig:clonalorigin}A. A
-splitting event of the population tree could happen somewhere between the root
-of the species tree and the internal node of SPY. Recombinant edges that connect
-branches within SPY or SD clades would be more frequent than those that connect
-branches between SPY and SDE clades.  With the reference to the recombination
-within a clade one parameter can dictate the degree of gene from from SPY to
-SDE, and another for the reverse direction.  If the two parameters could be
-incorporated in the model, then they might be used to compare the degrees of
-gene flow between the two species in a more statistically sound way.  We hope
-that we will see this refinement of the model to better illustrate bacterial
-evolution in forthcoming bacterial genomic era.
-
-% You may title this section "Methods" or "Models". 
-% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
-% authors may use "Analysis" 
-\section*{Materials and Methods}
-
-\subsection*{Complete bacterial genomes}
-
-Two genomes of \textit{S.\ pyogenes} (SPY), and two of \textit{S.\ dysgalactiae}
-ssp.\textit{\ equisimilis} (SDE) were downloaded from NCBI with the following
-accessions: NC\_004070 \citep{Beres2002}, NC\_008024 \citep{Beres2006},
-NC\_012891 \citep{Shimomura2011}, and CP002215 \citep{Suzuki2011}. We refer to
-these samples as SPY1, SPY2, SDE1, and SDE2, respectively. The genome of
-\textit{S.\ dysgalactiae} ssp.\textit{\ dysgalactiae} was provided by
-Suzuki et al.\ \citet{Suzuki2011} with accession identifiers of SddyATCC27957, referred to here
-as SDD. We also downloaded the genome (NC\_012471) of \textit{S.\ equi}
-ssp.\textit{\ equi} strain 4047 referred to as SEE \citep{Holden2009}.
-
-\subsection*{Model-based method for recombining gene transfer}
-
-We present a short summary of the statistical method developed by
-Didelot et al.\ \citet{Didelot2007} and \citet{Didelot2010} for studying bacterial
-recombination. A detailed account was found in the two papers and in the user
-guide available at \url{http://code.google.com/p/clonalorigin}.  We used
-\texttt{progressiveMauve} v2.3.1 \citep{Darling2004,Darling2010} with default
-options to attempt to align the six genomes, which led to too small proportion
-of core genomes. We had to remove the SEE sample from the genome alignment
-although we kept it in the parsimony-based approach.  We then used
-\texttt{stripSubsetLCBs}, a program of the \texttt{ClonalOrigin} package, and
-identified 276 blocks of sequence alignments with length threshold of 1500 base
-pairs.  We excluded two of these blocks in a subsequent analysis because they
-contained long consecutive gaps.  We applied \texttt{ClonalFrame} v1.1 available
-at \url{http://www.xavierdidelot.xtreemhost.com} to the remaining 274 blocks,
-and estimated the species tree relating the five genomes. \texttt{ClonalFrame}
-also estimates the fraction between recombination and mutation rates.
-We used a single Markov chain Monte Carlo (MCMC) with 10000 burn-in generations
-and 10000 sampling generations to obtain the tree estimate with branch lengths,
-and additionally seven independent MCMC runs to validate adequate convergence of
-the first run.
-
-Two stages of MCMC with \texttt{ClonalOrigin}, subversion r19 that had been
-downloaded from \url{http://clonalorigin.googlecode.com/svn/trunk}, were
-employed independently for each of the 274 alignment blocks as described in
-Didelot et al.\ \citet{Didelot2010}.   In the first MCMC we obtained estimates of mutation rate,
-recombination rate, and recombinant tract length.  Two independent chains were
-executed with $10^6$ burn-in generations and $10^7$ sampling generations. We
-sub-sampled the result every $10^5$-th generations.  A few alignment blocks were
-not successfully finished even after a month of computation.  Because those were
-not expected to affect the global estimates of population parameters
-significantly, we used only finished blocks to estimate mutation rate,
-recombination rate, and recombinant tract length.  The global estimators of the
-three population parameters were medians weighted over block lengths, which
-down-weighed less reliable estimates obtained from shorter alignment blocks.
-One of the two independent chains was used to obtain the block-wise mean
-estimates, and the other chain was used to validate convergence of the first
-stage of MCMC.  Two independent MCMC runs resulted in similar estimates of the
-three population parameters (Table S\ref{tab:three}).  The second stage of MCMC
-with \texttt{ClonalOrigin} was executed with the three population parameters
-fixed to the estimated values in the first stage.  Two independent MCMC chains
-were also executed with a total generation being $1.1\times10^8$. The first
-$10^7$ generations of each chain were discarded as burn-in. The remaining $10^8$
-generations were sub-sampled every $10^5$-th generation.
-
-\subsection*{Parsimony-based method for horizontal gene transfer}
-
-While \texttt{ClonalOrigin} could be used to understand recombining gene
-transfers within the conserved core genome, we sought to understand the history
-of the remainder of the genome where more complex evolutionary processes occur,
-such as gene duplications, losses, and horizontal transfers.  In recent years,
-several methods have been developed for inferring these evolutionary events in
-gene families by reconciling gene trees with a species tree
-\citep{David2011,Doyon2011,Tofigh2011}.  In this analysis, we complemented our
-study of recombination by using \texttt{Mowgli} \citep{Doyon2011} to
-characterize events that can change the copy number of a gene, such as
-duplications, losses, and non-recombining horizontal gene transfers.  In
-contrast to \texttt{ClonalOrigin}, \texttt{Mowgli} is parsimony-based for
-inferring events and uses pre-determined gene clusters where genes are treated
-as atomic units.  The advantage of using \texttt{Mowgli} was that it did not
-assume the sequence data contains a single copy of each homologous segment, and
-it attempted to explain different copy numbers by a series of gain and loss
-events.  We obtained gene family clusters defined in Suzuki et al.\ \citet{Suzuki2011} using
-the \texttt{OrthoMCL} method \citep{Li2003} over the full set of the six genomes
-including \textit{S.\ equi} subsp.\textit{\ equi} strain 4047 (SEE).  For each
-family cluster, the corresponding protein sequences were aligned using the
-\texttt{MUSCLE} program \citep{Edgar2004a} and nucleotide alignments were
-constructed by mapping nucleotide sequences onto each protein sequence
-alignment. We then identified recombination within gene alignments using the
-Single Breakpoint Recombination (\texttt{SBP}) method from the \texttt{HyPhy}
-package \citep{KosakovskyPond2006}.  Using the Akaike Information Criterion
-(AIC), \texttt{SBP} inferred that 47.5\% (1,094) of gene families contained at
-least one topology-changing recombination. Each of these families was then split
-into two gene fragments.  Gene trees were constructed from nucleotide alignments
-for each gene fragment using the \texttt{RAxML} program \citep{Stamatakis2006},
-and were rooted by minimizing inferred duplications and losses.  Each rooted
-gene tree was then reconciled using the \texttt{Mowgli} program
-\citep{Doyon2011}, which inferred the most parsimonious gene duplication, loss,
-and transfer events for each family.  In the case of multiple maximum
-parsimonious reconciliations, one was chosen uniformly at random.  Unless
-specified otherwise, default setings were used for all the programs.  Since the
-method of \texttt{Mowgli} did not directly model recombination events,
-recombination-based transfers were inferred in a post-processing step, by
-checking whether the transfer was coupled with a loss of a homologous gene in
-the recipient branch (indicating gene replacement rather than gain).  A transfer
-event was classified as non-recombining if the destination species of that
-transfer had an extant descendant that contained copies of that genes which were
-descendants of the transferred gene as well as copies which were
-non-descendants.  We noted that this parsimonious approach was imperfect and
-would tend to undercount events, especially recombining gene transfer events,
-which were modeled using two events: a transfer and a loss.  Nevertheless, this
-analysis allowed us to study non-recombining gene transfers, which were not
-modeled by \texttt{ClonalOrigin}, and to validate some of the
-\texttt{ClonalOrigin} results on recombining gene transfer.
-
-\subsection*{Functional categories association}
-
-We performed a series of analyses to check for functional gene categories
-enriched for certain types of gene transfer events. Functional categories were
-assigned in the same manner as Suzuki et al.\ \citet{Suzuki2011}. In short, \texttt{blastP} was
-used to compare Streptococcus genes to bacterial proteins from the Uniref90
-database.  Matches with E-values less than $1.0\times10^{-5}$ were assumed to
-have the same Gene Ontology (GO) classification as the target gene, as
-determined from the uniProt GOA database.  We also classified a category of
-\textit{virulent genes} using the virulence profile described in
-Suzuki et al.\ \citet{Suzuki2011}. Among 207 virulent genes at least one of the genes in the
-gene family was identified by that profile.  Each gene covered by our analysis
-was associated with a recombination intensity score from the
-\texttt{ClonalOrigin} analysis and scores from the \texttt{Mowgli} analysis
-corresponding to number of duplications, losses, horizontal gene transfers, and
-recombining gene transfer events.  Mann-Whitney tests were used to identify GO
-categories whose genes showed significantly elevated values in each score.  A
-5\% False Discovery Rate (FDR) correction \citep{Benjamini1995} was used to
-determine an appropriate significance threshold for our p-values.
-
-\subsection*{Simulation Study}
-
-We performed three sets of simulations.  In the first simulation, we wished to
-show that \texttt{ClonalOrigin} could infer population mutation rate, tract
-length, and recombination rate with reasonable accuracy. In the second
-simulation, we wished to check whether recombinant edges were recovered from
-simulated data with recombinant trees generated from the prior. The third
-simulation was conducted using simulated data with recombinant trees sampled the
-posterior distribution.  We used a simulation feature of \texttt{ClonalOrigin}
-to simulate the process of mutation and recombination across the five species
-(see section \textbf{Supplementary Information} for detail).
+% murI -> no recombination
+% chr1:303215-303652
+% /locus_tag="SpyM3_0262"
+%
+% mutS -> a little bit
+% chr1:1836841-1837245
+% /locus_tag="SpyM3_1806"
+%
+% recP -> somewhat strong (consistent with Kalia2001's conclusion)
+% chr1:1454877-1455335
+% /locus_tag="SpyM3_1462"
+%
+% xpt -> somewhat strong
+% chr1:846784-847233
+% /locus_tag="SpyM3_0794"
+%
+% yqil -> no data available
+% chr1:131266-131699
+% /locus_tag="SpyM3_0108"
+%
+%\texttt{SC: I am not sure, yet, because I wrote,
+%Except for \textit{mutS} where
+%they did not find recombination, our results were
+%consistent with those of \citet{Kalia2001}. Let me make sure this.}
+Of course, our study differs from theirs in numerous respects, but this
+suggests the two loci are prone to recombination.
+%Except for \textit{mutS} where
+%\cite{Kalia2001} did not find recombination, our results were
+%consistent with those of \cite{Kalia2001}.  
+%\cite{Pinho2010} reported recombination in the gene \textit{parC}, in which we
+%also observed moderate recombination.
+Second, are the observed patterns of gene flow truly driven by replacing
+transfers, or could additive transfers also be contributing?  Here, our
+evidence is weaker, because our parsimony-based method has limited ability
+to resolve the direction of transfer.  More genomes will be needed to
+evaluate the directionality of additive transfers between these groups.
+Third, what could explain the observed asymmetry in gene flow?  Possible
+mechanisms include differences in the sizes or demographic structure of
+co-mingling populations of cells, or differences in barriers to gene
+transfer \citep{Thomas2005}.  \cite{Kalia2001} speculated that 
+SDE might be inherently less capable than SPY to acquire DNA by homologous
+recombination, for example, due to differences in the 
+restriction-modification system (RMS), hetero-duplex formation, and/or mismatch
+repair.  
+%Restriction-modification
+%system is a biological self-defense mechanism of bacteria, which is diverse
+%in prokaryote world.  
+Interestingly,
+while most of \textit{S.\ pyogenes} genome contain both type I and type II
+RMSs, the two SDE genomes studied here contain
+only type II RMSs.  (CHECK: THIS NEEDS SOME WORK)
+
+% can we integrate some of this?  reread Kalia
+%More recently, 
+%As biological roles of RMS might be to maintain and
+%control species identity \citep{Jeltsch2003}, \cite{Budroni2011a} recently
+%studied that \textit{Neisseria meningitidis} 
+%phylogenetic clades were associated with RMS, claiming that RMS modulated
+%homologous recombination in the bacteria of \textit{N.\ meningitidis}.
+
+%Most of \textit{S.\ pyogenes} genomes contained restriction-modification
+%type I system as well as type II system. The two genomes of SDE contained
+%only type II system. A more robust self-defense system in SPY could have
+%made the species less capable of acquiring foreign genetic material from
+%SDE.
+% can we follow-up on this in some way?
+
+%More SDE genomes might be beneficial to future studies.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% be careful not to undercut our findings here.
+
+% maybe more than a paragraph of limitations, make it one about insight --
+% idea of the two events being identified with different parts of the genome
+
+In comparing replacing and additive HGTs, it is worth bearing in mind that
+these events are likely to be strongly correlated with the core and
+``dispensable'' portions of the genome (we take the dispensible genome to
+be the difference between the pan and core genomes).  Methods for detecting
+recombination (replacing transfer) tend to rely on large-scale alignments
+spanning multiple syntenic loci in order to gain power.  As a result, these
+methods are effectively limited to application in the core genome, even
+though homologous recombination probably also occurs (perhaps at a reduced
+rate) between genes that are less widely shared across species.  By
+contrast, phylogenetic reconciliation methods can be applied to the entire
+pan genome, as in this study.  Moreover, the additive transfer events that
+they detect will necessarily be enriched in the dispensible genome.  Thus,
+the replacing HGTs identified by recombination-based methods will tend to
+reflect the core genome and the additive HGTs identified by
+reconcilation-based methods will tend to reflect the dispensible genome.
+In our view, this is not necessarily a limitation of the comparison,
+because (with the exception of some replacing transfers in the dispensible
+genome) we expect the inferred transfer events to be fairly representative
+of true HGTs, but it is useful to keep in mind that some apparent
+differences between the two 
+types of transfer events may simply reflect general differences between the
+core and dispensible genome.
+
+%In addition, these methods are likely to have quite different power for various
+%kinds of events, because of factors such as
+%branch-length dependency
+%use of multiple loci or single loci
+%larger trees
+
+%paragraph on future methods
+As sequence data becomes available for many more species, including both
+closely and distantly related organisms, it will become increasingly
+important to devise improved methods that consider both additive and
+replacing HGT.  As noted above, the current statistical-models of
+recombination, such as ClonalOrigin, are effectively limited to considering
+orthologs, with one copy per species, in the core genome.  The phylogenetic
+reconciliation methods, on the other hand, fail to allow for multi-locus
+gene transfer events (which are known to occur), and also ignore
+information from branch lengths.  We have shown here that it is possible,
+to a degree, to extend them to distinguish between additive and replacement
+transfers but our methods are limited to a degree by their dependency on
+heuristic rules and parsimony assumptions (reference supplementary section
+on simulation experiments).  It may be possible to develop integrated
+statistical models that consider both types of gene transfers, and to use
+them to more directly compare the rates with which they occur.  Another
+extension worth considering is direct modeling of incomplete lineage
+sorting (ILS), which is likely to be increasingly important as larger
+phylogenies are considered, especially given the large effective population
+sizes of many bacteria.  ILS has recently been integrated into models for
+duplication and loss (Rasmussen and Kellis) but has yet to be considered in
+reconciliation-based methods that also allow for gene transfer.  It is
+permitted by ClonalOrigin, but only in an indirect and approximate manner,
+through the introduction of recombination edges to ancestral branches
+(which can be misinterpreted as evidence of recombination).  The
+combination of improved models and richer data sets could allow for a much
+more detailed understanding of horizontal gene transfer.
+
+%Molecular evolution of bacterial genomes is multifaceted. Researches of
+%bacterial evolution often focused on one side of many evolutionary processes.
+%While understanding one aspect of bacterial evolutionary processes we could be
+%ignorant of the other features.  Horizontal gene transfer process was often
+%considered in macro-evolutionary scale using a great deal of bacterial species.
+%As closely related species or groups of individual bacteria are becoming a focus
+%of evolutionary study, recombining gene transfer process would surface as
+%another facet of evolutionary forces.  Because the two gene transfer processes
+%would act on closely related bacterial genomes simultaneously, ignoring one of
+%the two processes in studying evolution of bacteria would lead to a narrowed
+%point of view to the underlying evolutionary forces. Our study of SPY and SDE
+%genomes undertook the work of discriminating recombining and horizontal gene
+%transfer processes because the two species were very closely related.  In this
+%study, we showed that the two gene transfer processes acted on different
+%biological functions.  Recombining gene transfer process in the analysis of SPY
+%and SDE genomes was related with genes encoding membrane-bound molecules and
+%aminoacyl tRNA synthetases. On the contrary, horizontal gene transfer process
+%was found to be associated with transposases, which would allow horizontal gene
+%transfer of their neighboring genes.  Because SPY was a human-specific pathogen,
+%and SDE was an opportunistic human pathogen, we were interested in association
+%of pathogenicity with the two gene transfer processes. We also showed that
+%virulence genes were more associated with horizontal gene transfer process than
+%with recombining gene transfer process.  With the current data of only SPY and
+%SDE genomes we could not conclude whether the association of virulence genes
+%with horizontal gene transfer not recombining gene transfer was applicable
+%generally to cases of other species.  However, without the discrimination of the
+%two processes it would have been difficult to find the disparate association of
+%biological functions or virulence genes with the different gene transfer
+%processes. 
+%
+%We showed differences in the profile of functional category association of the
+%two gene transfer processes thanks to the approach of dicriminating the two gene
+%transfer processes.  Sugar phosphotransferases could be important in signal
+%transduction to recognize environmental changes.  Transporters, and fatty acid
+%and lipid biosyntheses would be involved in membrane-bound molecules.  Because
+%bacteria as a unicellular organism should be able to promptly respond to changes
+%to environment, some evolutionary pressures must have been acting on genes
+%encoding membrane-bound molecules.  We also found that the recombination
+%intensity and requency of recombining gene transfer events would be useful in
+%associating several aminoacyl tRNA synthetases genes belonging to gene ontology
+%term of translation (GO:0006412) with recombining gene transfer. It had been a
+%focus of gene transfer in bacterial species \citep{Woese2000}.  The biological
+%role of transposases was concordant with horizontal gene transfer process
+%because the process involved in transposases would allow bacteria to acquire
+%foreign genes without the requirement of homologous regions in the bacteria.  
+%
+%Another contribution of our study to bacterial gene transfer research is to
+%elucidate the net directionality of gene flow from SPY to SDE in the core
+%genomes. Although the issue of net directionality has been tackled, our study
+%was the first attempt of resolving the issue in genome-wide scale. Yet, it still
+%remains to be seen whether this net directionality of gene flow from SPY to SDE
+%happened in the pan genome. A model-based method with more refined models would
+%be necessary because the parsimony-based method we employed simply could not
+%show any net directionality between SPY and SDE in the pan genome. Although we
+%are unsure whether much larger number of genomes could be useful in finding net
+%directionality, more SDE genomes might be beneficial to future studies. 
+%
+%Were there unbalanced recombining gene transfer between the two species, what
+%were the barriers \citep{Thomas2005} to cause the unbalanced gene transfer in
+%the species?  Kalia et al.\ \cite{Kalia2001} discussed restriction-modification
+%system (RMS) as a possible mechanism of the unbalanced gene flow.
+%Restriction-modification system is a biological self-defense mechanism of
+%bacteria, which is diverse in prokaryote world.  As biological roles of RMS
+%might be to maintain and control species identity \citep{Jeltsch2003}, Budroni
+%et al.\ \cite{Budroni2011a} recently studied that \textit{Neisseria
+%meningitidis} phylogenetic clades were associated with RMS, claiming that RMS
+%modulated homologous recombination in the bacteria of \textit{N.\ meningitidis}.
+%Most of \textit{S.\ pyogenes} genomes contained restriction-modification type I
+%system as well as type II system. The two genomes of SDE contained only type II
+%system. More tight self-defense system of SPY may have made the species less
+%capable of aquiring foreign genetic material from SDE \citep{Kalia2001}.
+%
+%We reexamined the seven genes that Kalia et al.\ \cite{Kalia2001} had studied
+%for net directionality of gene transfer between SPY and SDE.  The internal
+%fragments of the six housekeeping genes (\textit{gki}, \textit{gtr},
+%\textit{murI}, \textit{mutS}, \textit{recP}, \textit{xpt}) were located in the
+%core genome that we investigated in this study.  Except for \textit{mutS} where
+%\cite{Kalia2001} did not find recombination, our results were
+%consistent with those of \cite{Kalia2001}.  
+%\cite{Pinho2010} reported recombination in the gene \textit{parC}, in which we
+%also observed moderate recombination.
+%
+%It remains to be seen whether the inability of detecting net directionality
+%between SPY and SDE in the pan genomes was because of lack of power of the
+%parsimony-based approach that we employed. Considering gene duplication, loss,
+%and horizontal gene transfer in a model-based approach would be desirable in the
+%coming ages of abundant bacterial genomes. The model might consider genome
+%alignments where each taxon does not necessarily have a unique gene; some taxa
+%might have no gene due to gene loss, and other taxa multiple genes due to gene
+%gain, duplication, or horizontal gene transfer. With these types of models will
+%we be able to address more interesting evolutionary questions in bacteria. This
+%need of more refined models would be also true even for the study of net
+%directionality between species in the core genome.  In other words, one could
+%envision a model where a population two-taxa (i.e., SPY and SD groups) tree is
+%superimposed on the species tree shown in Figure S\ref{fig:clonalorigin}A. A
+%splitting event of the population tree could happen somewhere between the root
+%of the species tree and the internal node of SPY. Recombinant edges that connect
+%branches within SPY or SD clades would be more frequent than those that connect
+%branches between SPY and SDE clades.  With the reference to the recombination
+%within a clade one parameter can dictate the degree of gene from from SPY to
+%SDE, and another for the reverse direction.  If the two parameters could be
+%incorporated in the model, then they might be used to compare the degrees of
+%gene flow between the two species in a more statistically sound way.  We hope
+%that we will see this refinement of the model to better illustrate bacterial
+%evolution in forthcoming bacterial genomic era.
+
+% ACS: Let's let the journal do this 
+%\section*{Supplementary Material}
+%Supplementary figures S1-S10 and tables S1-S9 are available at Molecular Biology
+%and Evolution online (http://mbe.oxfordjournals.org/).
 
@@
 
-We are indebted to Xavier Didelot for his help in setting up \texttt{ClonalOrigin}
-analysis.  We are thankful to Haruo Suzuki for his help
-in using the data of SDD genome. This work was supported by the National
-Institute of Allergy and Infectious Disease, US National Institutes
-of Health, under grant number AI073368-01A2 awarded to A.S. and M.J.S.
-
-\section*{Author Contributions}
-
-Conceived and designed the experiments: SCC MDR MJS AS.
-Performed the experiments: SCC MDR MJH.
-Analyzed the data: SCC MDR MJH IG.
-Contributed reagents/materials/analysis tools: MJS.
-Wrote the paper: SCC MDR MJH IG AS.
-
-%\section*{References}
+We thank Xavier Didelot for assistance with the ClonalOrigin analysis, and
+Haruo Suzuki for assistance with the data for SDD. This work was primarily
+supported by the National Institute of Allergy and Infectious Disease, US
+National Institutes of Health, under grant number AI073368-01A2 (to
+M.J.S. and A.S).  Additional support was provided by National Science
+Foundation CAREER Award DBI-0644111 and a David and Lucile Packard
+Fellowship for Science and Engineering (to A.S.).
+
+%\section*{Author Contributions}
+
+%Conceived and designed the experiments: SCC MDR IG MJS AS.
+%Performed the experiments: SCC MDR MJH.
+%Analyzed the data: SCC MDR MJH IG.
+%Contributed reagents/materials/analysis tools: MJS.
+%Wrote the paper: SCC IG AS (with review and contributions from all authors).
+%\clearpage
+
 % The bibtex filename
-\bibliographystyle{plos}
+\renewcommand*{\refname}{Literature Cited}
+\bibliographystyle{../../latex/bst/mbe}
 \bibliography{siepel-strep}
 \clearpage{}
-
-\section*{Figure Legends}
-
-\begin{flushleft}
-
-Figure \ref{fig:tree5}.\ {\bf The species tree of the five genomes.}
-SPY is the branch of a clade of \textit{Streptococcus
-pyogenes}, SDE is that for \textit{S.\ dysgalactiae} ssp.\textit{\ equisimilis},
-and SD is the branch for the three individual genomes of \textit{S.\
-dysgalactiae}. The scale is proportional to the expected number of mutations
-given the Watterson's estimate of mutation rate and the tree.
-
-Figure \ref{fig:Heatmap-of-recombination}.\ {\bf Heatmap of recombination events 
-using the five genomes.}
-Each cell of the heat map represents the logarithm base
-2 of the number of recombination events inferred relative to its expectation
-under the prior, for a donor and recipient pair of branches. The donor is given by
-the y-axis; the recipient on the x-axis.  The black colored cells are meaningless
-because the recipient species tree branch of the cells are older than the donor
-species branch.
-
-Figure \ref{fig:Gene-duplication-loss}.\ {\bf Distribution of inferred gene 
-duplication, loss, and transfer events across
-the phylogeny.}  Gene counts for
-each extant and ancestral species are given at each node of the phylogeny.  For
-each branch we denote the number of gains G*, duplications D*, losses L*,
-non-recombining transfers T*, and recombinations R*.  The non-recombining
-transfer and recombination counts are indicated on the recipient branch (donors
-are not indicated).
-
-Figure \ref{fig:h3}.\ {\bf Simulations of the model-based approach for assessing 
-the power of recovering recombinant edges.} 
-Pairs of species tree branches are shown at x-axis labels.
-Three intervals are displayed for each pair. Intervals with solid line with open
-circle in the middle are for the logarithm of the fraction to base 2 of the average
-number of recombinant edges from the ten simulated recombinant trees 
-to the numbers of recombinant edges estimated from Simulation 2. 
-Interverals with dotted line are for the logarithm of the fraction to base 2 of
-the number of recombinant edges estimated from the real data
-to the numbers of recombinant edges estimated from Simulation 2.
-Intervals with solid line with closed circle in the middle are for the logarithm
-of the fraction to base 2 of the number of recombinant edges estimated from the real data
-to the numbers of recombinant edges estimated from Simulation 3. 
-The horizontal lines are at the zero.
-The intervals range from 5\% quantile to 95\% of quantile.
-
-\end{flushleft}
-\clearpage{}
-
-%\begin{figure}[!ht]
-%\begin{center}
-%%\includegraphics[width=4in]{figure_name.2.eps}
-%\end{center}
-%\caption{
-%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
-%should be left justified, as specified by the options to the caption 
-%package.
-%}
-%\label{Figure_label}
-%\end{figure}
 
@@
 %=============================================================================
+% Table 2. 
+\begin{table}[!ht]
+
+\caption{{\bf Functional category enrichments for additive gene
+  transfers}}
+\vspace{1ex}
+
+\noindent \begin{centering}
+\begin{tabular}{ccccl}
+\hline 
+$p^a$ & $q^b$ & Count$^c$ & GO term & Description \\
+\hline 
+7e-61 & 3e-58 & 67 & GO:0006313 & transposition, DNA-mediated\\
+3e-58 & 7e-56 & 59 & GO:0004803 & transposase activity\\
+1e-30 & 2e-28 & 52 & GO:0015074 & DNA integration\\
+1e-28 & 1e-26 & 23 & GO:0032196 & transposition\\
+1e-13 & 1e-11 & 527 & GO:0003677 & DNA binding\\
+2e-12 & 2e-10 & 86 & GO:0006310 & DNA recombination\\
+1e-11 & 7e-10 & 197 & GO:0003676 & nucleic acid binding\\
+3e-09 & 1e-07 & 36 & GO:0016987 & sigma factor activity\\
+3e-09 & 1e-07 & 36 & GO:0006352 & transcription initiation\\
+3e-08 & 1e-06 & 142 & GO:0043565 & sequence-specific DNA binding\\
+4e-07 & 1e-05 & 12 & GO:0008170 & N-methyltransferase activity\\
+1e-06 & 4e-05 & 12 & GO:0019867 & outer membrane\\
+2e-06 & 5e-05 & 25 & GO:0007059 & chromosome segregation\\
+1e-05 & 4e-04 & 15 & GO:0006306 & DNA methylation\\
+2e-05 & 6e-04 & 15 & GO:0006470 & protein amino acid dephosphorylation\\
+2e-04 & 4e-03 & 76 & GO:0005618 & cell wall\\
+3e-04 & 7e-03 & 25 & GO:0008236 & serine-type peptidase activity\\
+3e-04 & 7e-03 & 65 & GO:0009986 & cell surface\\
+5e-04 & 1e-02 & 10 & GO:0003872 & 6-phosphofructokinase activity\\
+1e-03 & 2e-02 & 11 & GO:0009307 & DNA restriction-modification system\\
+2e-03 & 5e-02 & 84 & GO:0006974 & response to DNA damage stimulus\\
+\hline 
+\end{tabular}
+\par\end{centering}
+\begin{flushleft}
+$^a$$P$-values based a Mann-Whitney $U$ test.  Shown are values that
+correspond to FDR$^b$  $<0.05$ based on the 
+Benjamini-Hochberg method (Methods).\\
+$^c$Number of genes in category.\\
+\end{flushleft}
+\label{tab:go-events}
+\end{table}
+%\clearpage{}
+%=============================================================================
+
+%=============================================================================
+% /Users/goshng/Documents/Projects/Mauve/output/cornellf/3/run-analysis/significant-1.txt
 % Table 1. 
@@
 \caption{
-{\bf Functional categories associated with 
-recombination intensity based on topology-altering recombinant edges.}}
+{\bf Functional category enrichments for replacing gene transfers$^a$}} 
+\vspace{1ex}
 \noindent \begin{centering}
-\begin{tabular}{cccl}
-\hline 
-P-Value & Count & GO term & Description\tabularnewline
-\hline 
-1.731e-06 & 38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
-2.775e-06 & 33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
-4.581e-05 & 22 & GO:0006633 & fatty acid biosynthetic process\tabularnewline
-0.0001248 & 20 & GO:0008610 & lipid biosynthetic process\tabularnewline
-0.0002675 & 45 & GO:0008643 & carbohydrate transport\tabularnewline
-0.000513 & 81 & GO:0006412 & translation\tabularnewline
-0.0005617 & 13 & GO:0006814 & sodium ion transport\tabularnewline
-0.001019 & 40 & GO:0030529 & ribonucleoprotein complex\tabularnewline
-0.001203 & 12 & GO:0031402 & sodium ion binding\tabularnewline
-0.00206 & 39 & GO:0003735 & structural constituent of ribosome\tabularnewline
-0.002115 & 237 & GO:0008152 & metabolic process\tabularnewline
-0.00215 & 26 & GO:0019843 & rRNA binding\tabularnewline
-0.00287 & 371 & GO:0016740 & transferase activity\tabularnewline
-0.003552 & 66 & GO:0005840 & ribosome\tabularnewline
+\begin{tabular}{ccccl}
+\hline 
+$p^b$ & $q^c$ & Count$^d$ & GO term & Description\\
+\hline 
+2e-06 & 2e-04 &  38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
+3e-06 & 2e-04 &  33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
+5e-05 & 3e-03 &  22 & GO:0006633 & fatty acid biosynthetic process\\
+9e-05 & 4e-03 &  21 & GO:0008610 & lipid biosynthetic process\\
+3e-04 & 9e-03 &  45 & GO:0008643 & carbohydrate transport\\
+5e-04 & 1e-02 &  82 & GO:0006412 & translation\\
+6e-04 & 1e-02 &  13 & GO:0006814 & sodium ion transport\\
+1e-03 & 2e-02 &  40 & GO:0030529 & ribonucleoprotein complex\\
+1e-03 & 2e-02 &  12 & GO:0031402 & sodium ion binding\\
+1e-03 & 2e-02 & 244 & GO:0008152 & metabolic process\\
+2e-03 & 3e-02 & 378 & GO:0016740 & transferase activity\\
+2e-03 & 3e-02 &  39 & GO:0003735 & structural constituent of ribosome\\
+2e-03 & 3e-02 &  26 & GO:0019843 & rRNA binding\\
+4e-03 & 4e-02 &  66 & GO:0005840 & ribosome\\
 \hline 
@@
 \par\end{centering}
+\begin{flushleft}
+$^a$Based on recombination intensities restricted to topology-altering
+recombinant edges (Methods).\\
+$^b$$P$-values based a Mann-Whitney $U$ test.  Shown are values that
+correspond to FDR$^c$  $<0.05$ based on the 
+Benjamini-Hochberg method (Methods).\\
+$^d$Number of genes in category.\\
+\end{flushleft}
 \label{tab:functional}
 \end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Table 2. 
-\begin{table}[!ht]
+\clearpage{}
+%=============================================================================
+
+\section*{Figures}
+
+%=============================================================================
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=3in]{figures/cells.pdf}
+\end{center}
+\caption{ {\bf Replacing and additive horizontal gene transfer.}  (A)
+  Foreign DNA can be integrated into a recipient genome by homologous
+  recombination (a {\em replacing transfer}, left) or additive integration
+  (an {\em additive transfer}, right). (B) These types of transfers produce
+  distinct phylogenetic signatures.  In homologous recombination a segment
+  of DNA is effectively overwritten by a homologous segment from another
+  species, which causes a lineage in the phylogeny to be replaced by a
+  transferred lineage.  This type of transfer can be identified in
+  gene-tree/species-tree reconciliation by the appearance of coinciding
+  transfer and loss events (left tree).  Additive integration, on the other
+  hand, leads to a transfer event that is not paired with a loss event
+  (right tree).  Of course, parallel losses can cause an additive transfer
+  to appear similar to a replacing transfer.  In our analysis, we
+  conservatively require that at least one descendant species (here, {\bf
+    B}) contains genes that both descend from ($b_1$) and do not descend
+  from ($b_2$) a lineage predicted to have been transferred when
+  inferring an additive transfer event.}
+\label{fig:hgt}
+\end{figure}
+\clearpage{}%
+%=============================================================================
+
+%=============================================================================
+\begin{figure}[!ht]
+\noindent \begin{centering}
+\includegraphics[scale=0.5]{figures/cornellf-3-tree}
+\par\end{centering}
 \caption{
-{\bf Functional categories associated with 
-non-recombining horizontal gene transfer.}}
-\noindent \begin{centering}
-\begin{tabular}{cccl}
-\hline 
-P-Value & Count & GO term & Description \\
-\hline 
-7e-61 &  67 & GO:0006313 & transposition, DNA-mediated\\
-3e-58 &  59 & GO:0004803 & transposase activity\\
-1e-30 &  52 & GO:0015074 & DNA integration\\
-1e-28 &  23 & GO:0032196 & transposition\\
-1e-13 & 527 & GO:0003677 & DNA binding\\
-2e-12 &  86 & GO:0006310 & DNA recombination\\
-1e-11 & 197 & GO:0003676 & nucleic acid binding\\
-3e-09 &  36 & GO:0016987 & sigma factor activity\\
-3e-09 &  36 & GO:0006352 & transcription initiation\\
-3e-08 & 142 & GO:0043565 & sequence-specific DNA binding\\
-\hline 
-\end{tabular}
-\par\end{centering}
-\label{tab:go-events}
-\end{table}
-\clearpage{}
-%=============================================================================
-
-\section*{Figures}
-
-%=============================================================================
-% Figure 1: Species tree
-\begin{figure}[!ht]
-\includegraphics[scale=0.7]{figures/cornellf-3-tree}
-\caption{
-{\bf The species tree of the five genomes.} 
-SPY is the branch of a clade of \textit{Streptococcus
-pyogenes}, SDE is that for \textit{S.\ dysgalactiae} ssp.\textit{\ equisimilis},
-and SD is the branch for the three individual genomes of \textit{S.\
-dysgalactiae}. The scale is proportional to the expected number of mutations
-given the Watterson's estimate of mutation rate and the tree.}
+{\bf Clonal frame inferred for the five genomes.}  The branch lengths are
+drawn to scale in the horizontal dimension.  Units are 
+expected substitutions per site given Watterson's estimate of the
+mutation rate. The branch length of the outgroup SEE is not scaled. 
+The labels for the ancestral nodes of the tree and the
+internal branches (SPY, SDE, and SD) are used throughout the paper.}
 \label{fig:tree5}
@@
 
-%=============================================================================
-% Figure 2. Heat map of recombining gene transfer from ClonalOrigin
+% A transfer is a additive if there exists an extant descendant species
+% (e.g. C and D) that contains both descendants (bold) and
+% non-descendants (normal weight) of the transferred gene (black
+% circle).
+
+%=============================================================================
 \begin{figure}
-\includegraphics[scale=0.45]{figures/heatmap-recedge}
-\caption{\label{fig:Heatmap-of-recombination}
-{\bf Heatmap of recombination events using the five genomes.}
-Each cell of the heat map represents the logarithm base
-2 of the number of recombination events inferred relative to its expectation
-under the prior, for a donor and recipient pair of branches. The donor is given by
-the y-axis; the recipient on the x-axis.  The black colored cells are meaningless
-because the recipient species tree branch of the cells are older than the donor
-species branch.}
+% \includegraphics[scale=0.45]{figures/heatmap-recedge}
+\includegraphics[scale=0.5]{figures/heatmap}
+\caption{\label{fig:Heatmap-of-transfers}
+{\bf Heatmaps showing rates of replacing and additive transfers.}
+Each cell of the heat map represents the base-2 logarithm 
+of the ratio of the estimated number recombination events to its expectation
+under the prior, for the corresponding donor ($y$ axis) and recipient ($x$
+axis) branches.  
+The black colored cells indicate prohibited transfer events.
+%
+(A) Replacing transfers, as inferred by the model-based approach. The
+plotted values represent 
+average values across sampled recombinant
+trees. The prior considers the clonal frame with branch lengths and the
+global estimates of the three population parameters. Prohibited transfer events
+are ones for which the recipient branch is strictly older than the donor
+branch. Significance levels are denoted by asterisks: one for $<0.01$.
+%
+(B) Additive transfers, as inferred by the parsimony-based
+approach. The plotted values represent
+numbers of inferred additive transfers across all gene families.
+The prior considers the clonal  frame (with branch lengths) and the total number
+of events of each type inferred by the analysis. Prohibited transfer events are
+ones for which the recipient and donor branch do not share a common time
+interval.
+branch. Significance levels are denoted by asterisks: one for $<0.01$, two for
+$<0.005$, and three for $<0.001$.}
 \end{figure}
@@
 %=============================================================================
-% Figure 3. Gene duplication, loss, and transfer
-\begin{figure}
-\includegraphics[width=6in]{figures/strep-events}
-\caption{\label{fig:Gene-duplication-loss} 
-{\bf Distribution of inferred gene duplication, loss, and transfer events across
-the phylogeny.}  Gene counts for
-each extant and ancestral species are given at each node of the phylogeny.  For
-each branch we denote the number of gains G*, duplications D*, losses L*,
-non-recombining transfers T*, and recombinations R*.  The non-recombining
-transfer and recombination counts are indicated on the recipient branch (donors
-are not indicated).}
+\begin{figure}[!ht]
+\includegraphics[scale=0.5]{figures/ucsc}
+% used genomic interval chr1:511,000-513,500
+\caption{ {\bf Genome Browser tracks.}  Our inferences of replacing and
+  additive horizontal gene transfers are summarized in new tracks in the
+  {\em Streptococcus} Genome Browser \citep{Suzuki2011}.  (A) The main
+  browser display, with gene annotations for the selected reference genome
+  ({\em S.\ pyogenes} strain MGAS315) shown in red and putative virulence
+  genes highlighted in blue. The third track from top (in black) shows the
+  putatively non-recombining gene fragments analyzed by Mowgli.  The next series
+  of tracks shows the results of the ClonalOrigin analysis of
+  replacing gene transfers.  Each track in this series corresponds to a
+  recipient lineage in the phylogeny (fig.\ \ref{fig:tree5}) and describes
+  the posterior probabilities along the genome of recombination edges from
+  all possible donor lineages (shown in different colors; see key at
+  right).  Here the putative virulence gene SpyM3\_0465, a dipeptidase,
+  shows strong evidence of a recombination edge from the SPY to the 
+  SDE lineage, as well as some evidence of a SPY$\rightarrow$SPY2 edge.
+  The genome-wide multiple alignment obtained with Mauve is shown at
+  bottom.  (B) The gene tree that is displayed after clicking on the
+  highlighted gene fragment.  The topology inferred by RAxML is consistent
+  with the combined influence of SPY$\rightarrow$SDE and
+  SPY$\rightarrow$SPY2 replacing transfers, as inferred by ClonalOrigin.}
+\label{fig:ucsc}
 \end{figure}
@@
 %=============================================================================
-% Figure 4. 
 \begin{figure}
-
-\subfloat[]{\includegraphics[scale=0.4]{figures/h3-R-ratio-topology-changing}}
-
-\subfloat[]{\includegraphics[scale=0.4]{figures/h3-R-ratio-not-topology-changing}}
-
-\caption{
-{\bf Simulations of the model-based approach for assessing the power of recovering
-recombinant edges.} Pairs of species tree branches are shown at x-axis labels.
-Three intervals are displayed for each pair. Intervals with solid line with open
-circle in the middle are for the logarithm of the fraction to base 2 of the average
-number of recombinant edges from the ten simulated recombinant trees 
-to the numbers of recombinant edges estimated from Simulation 2. 
-Interverals with dotted line are for the logarithm of the fraction to base 2 of
-the number of recombinant edges estimated from the real data
-to the numbers of recombinant edges estimated from Simulation 2.
-Intervals with solid line with closed circle in the middle are for the logarithm
-of the fraction to base 2 of the number of recombinant edges estimated from the real data
-to the numbers of recombinant edges estimated from Simulation 3. 
-The horizontal lines are at the zero.
-The intervals range from 5\% quantile to 95\% of quantile.}
-\label{fig:h3}
+\includegraphics[width=7in]{figures/strep-events}
+\caption{\label{fig:Gene-duplication-loss} 
+{\bf Distribution of inferred gene duplication, loss, and transfer events
+  for the six-species phylogeny.}  Gene counts for
+each extant and ancestral species are given at each node of the tree.  For
+each branch numbers of gene appearances (A*), duplications (D*), 
+losses (L*), additive transfers (T*), and replacing transfers (R*) are listed.
+Transfer events are recorded on the recipient branch (donors
+are not indicated).}
 \end{figure}
@@
 %=============================================================================
-% Table S1. List of six genomes.
 \begin{table}[!ht]
 \caption{
-{\bf The six complete bacterial genomes.}
-Each bacteria and its genome is referred to as the name at the first column.}
+{\bf The six bacterial genomes.}
+Each bacterial genome is referred to as the name at the first column.}
 \noindent \begin{centering}
@@
 %=============================================================================
-% Table S2. ClonalFrame Result.
 \begin{table}[!ht]
 \caption{
-{\bf Parameter estimates from \texttt{ClonalFrame}.}}
-\noindent \centering{}\begin{tabular}{cc}
-\hline
-Parameter & Estimates (95\% credibility interval) \tabularnewline
-\hline
-$\theta^a$ & 92871.84 (N/A)\tabularnewline
-$\nu^b$ & 0.134 (0.132, 0.135)\tabularnewline
-$R^c$ & 9370 (8950, 9780)\tabularnewline
-$\delta^d$ & 362 (347, 377)\tabularnewline
-$r/m^e$ & 4.66 (4.45, 4.87)\tabularnewline
-$\rho/\theta^f$ & 0.101 (0.096, 0.105)\tabularnewline
-TMRCA$^g$ & 0.331 (0.324, 0.337)\tabularnewline
-$T^h$ & 1.02 (1.01, 1.04)\tabularnewline
+{\bf Posterior probability of gene tree topologies.}
+See Table S\ref{tab:genome} for the reference of the bacteria
+individual names.}
+\noindent \begin{centering}
+\begin{tabular}{cc}
+\hline 
+Gene tree topologies & Posterior probability\tabularnewline
+\hline
+(((SDE1,SDE2),SDD),(SPY1,SPY2)) & 66.7\%\tabularnewline
+(((SDE1,SDE2),(SPY1,SPY2)),SDD) & 9.3\%\tabularnewline
+((SDE1,SDE2),(SDD,(SPY1,SPY2))) & 3.5\%\tabularnewline
+(((SDE1,(SDE2,SDD)),(SPY1,SPY2)) & 1.5\%\tabularnewline
+(((SDE1,SDD),SDE2),(SPY1,SPY2)) & 1.3\%\tabularnewline
+((SDE1,SDD),(SDE2,(SPY1,SPY2))) & 1.2\%\tabularnewline
+((((SDE1,SDE2),SPY1),SPY2),SDD) & 1.0\%\tabularnewline
+((SDE2,SDD),(SDE1,(SPY1,SPY2))) & 1.0\%\tabularnewline
 \hline
 \end{tabular}
-\begin{flushleft}
-$^a$ $\theta$ is population mutation rate,
-
-$^d$ $\delta$ is the average recombinant tract length,
-
-$^e$ $r/m$ is the relative strength of recombination compared to mutations,
-
-$^g$ TMRCA is the time to most recent common ancestor, and
-
-$^h$ $T$ is the total branch length of the inferred species tree.
-\end{flushleft}
-\label{tab:clonalframe}
+\par\end{centering}
+\label{tab:Gene-tree-topologies}
 \end{table}
@@
 %=============================================================================
-% Table S. ClonalOrigin's 1st stage result.
+\begin{table}[!ht]
+\caption{
+{\bf Ratios of logarithm base 2 of the observed
+recombinant edges to the expected number of recombinant edges.}  Figure
+\ref{fig:Heatmap-of-transfers}A shows the values graphically using heat map.
+NA's are for the branch pairs that are impossible under the model-based
+approached. Rows are for donor, and columns for recipient.}
+\noindent \centering{}\begin{tabular}{cccccccccc}
+\hline
+& SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2 \tabularnewline
+\hline
+SDE1& 0.176  &   NA  &1.170  &    NA & 0.7564  & NA & 0.67 &    NA & 0.663\tabularnewline
+SDE & -0.118 & 0.081 &-0.202 &     NA&  0.3395 &  NA& -0.15&  0.045& -0.323\tabularnewline
+SDE2& 1.201  &   NA  &0.092  &    NA & 0.6978  & NA & 0.60 &    NA & 0.387\tabularnewline
+SD  & -0.629 & 0.022 &-0.661 &0.00670&  0.0078 &  NA& -0.94& -0.299& -1.030\tabularnewline
+SDD & -0.317 & 0.371 &-0.197 &     NA&  0.0752 &  NA& -1.06& -0.525& -0.915\tabularnewline
+ROOT& -1.745 &-0.698 &-1.786 &0.23039& -0.4854 &  NA& -1.37&  0.497& -1.306\tabularnewline
+SPY1& 1.245  &1.883  &1.251  &    NA &-2.1997  & NA & 0.06 &    NA & 1.062\tabularnewline
+SPY & 0.089  &0.200  &0.298 &0.00098 &-0.7903  & NA &-0.19 &-0.058 &-0.159\tabularnewline
+SPY2& 1.174  &1.938  &1.287  &    NA &-1.9740  & NA & 1.07 &    NA & 0.063\tabularnewline
+\hline
+\end{tabular}
+\label{tab:heatmap}
+\end{table}
+%\clearpage{}
+%=============================================================================
+
+%=============================================================================
+\begin{table}[!ht]
+\caption{
+{\bf The observed number of recombinant edges.}
+The zeros are for the branch pairs that are impossible under the model-based
+approached. Rows are for donor, and columns for recipient.}
+\noindent \centering{}\begin{tabular}{cccccccccc}
+\hline
+& SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2 \tabularnewline
+\hline
+SDE1&8.896 &  0.00 &17.712  & 0.00 & 12.872   & 0 &12.59  &  0.00  &12.53\tabularnewline
+SDE &44.626& 136.49& 42.088 &  0.00& 227.061  &  0& 73.39 &  99.95 & 65.25\tabularnewline
+SDE2&18.099&   0.00&  8.393 &  0.00&  12.360  &  0& 11.97 &   0.00 & 10.36\tabularnewline
+SD  &9.671 &101.74 & 9.464  &31.26 &115.796   & 0 &13.68  & 98.38  &12.86\tabularnewline
+SDD &45.539& 168.58& 49.513 &  0.00& 194.706  &  0& 43.09 &  67.14 & 47.65\tabularnewline
+ROOT&43.553& 602.67& 42.337 &765.53& 802.813  &  0& 98.86 &2146.01 &103.60\tabularnewline
+SPY1&40.800&  11.46& 40.973 &  0.00&   4.371  &  0& 21.15 &   0.00 & 42.35\tabularnewline
+SPY &57.618& 262.08& 66.594 & 30.92& 163.518  &  0& 83.56 & 207.00 & 85.10\tabularnewline
+SPY2&38.837&  11.90& 42.000 &  0.00&   5.111  &  0& 42.45 &   0.00 & 21.19\tabularnewline
+\hline
+\end{tabular}
+\label{tab:obsheatmap}
+\end{table}
+%\clearpage{}
+%=============================================================================
+
+%=============================================================================
+\begin{table}[!ht]
+\caption{
+{\bf Number of genes showing significant signal of replacing transfer.}  For
+each type of recombinant edge (replacing gene transfer event in
+\texttt{ClonalOrigin} analysis), we record the number of genes with an average
+posterior probability greater than 0.6 for a recombinant edge of that type
+across their sites. Recombinant edges from the root branch are not considered
+here, and SPYclade and SDEclade each signify three branches in the clonal frame
+corresponding to the respective clades. Edge types not shown here have a zero
+count. The same is done for the subset of virulence genes. Absolute numbers are
+shown together with ratios relative to the total number of genes, as well as the
+total number of significant replacing transfers found in this analysis.}
+\noindent \centering{}\begin{tabular}{lll}
+\hline
+Recombinant edge type & Gene count & Virulence gene count \tabularnewline
+\hline
+SDEclade $\rightarrow$ SDD & 43 (33.1\%) & 5 (50\%) \tabularnewline
+SDEclade $\rightarrow$ SPYclade & 11 (8.5\%) & 1 (10\%) \tabularnewline
+SDD $\rightarrow$ SDEclade & 1 (0.8\%)  & 0 (0\%) \tabularnewline
+SPYclade $\rightarrow$ SDEclade & 59 (45.4\%) & 3 (30\%) \tabularnewline
+SPYclade $\rightarrow$ SPYclade & 16 (12.3\%) & 1 (10\%) \tabularnewline
+All types & 130  (100\%) & 10   ( 100\%)\tabularnewline
+% &  (11.8\% of total number of genes) & (13\% of virulence genes)\tabularnewline
+\hline
+\end{tabular}
+\label{tab:gene-counts-replacing}
+\end{table}
+%\clearpage{}
+%=============================================================================
+
+%=============================================================================
+\begin{table}[!ht]
+\caption{
+{\bf Number of inferred transfer events of different types.}  We record the number of inferred additive transfer events of each type.
+SPYclade and SDEclade each signify three branches in the clonal frame corresponding to the respective clades.
+Edge types not shown here have a zero count.
+The same is done for the subset of virulence genes.
+Absolute numbers are shown together with ratios relative to the total number of genes, as well as the total number of significant replacing transfers found in this analysis.}
+\noindent \centering{}\begin{tabular}{lll}
+\hline
+Recombinant edge type & Gene count & Virulence gene count \tabularnewline
+\hline
+SDEclade $\rightarrow$ SDEclade & 28 (10.6\%) &  5 (12.5\%)\tabularnewline
+SDEclade $\rightarrow$ SDD & 64 (24.2\%) & 2 (5.0\%)\tabularnewline
+SDEclade $\rightarrow$ SPYclade & 28 (10.6\%) &  7 (17.5\%)\tabularnewline
+SDD $\rightarrow$ SDEclade & 29 (10.9\%) &  5 (12.5\%)\tabularnewline
+SDD $\rightarrow$ SPYclade & 17 (6.4\%) &  2 (5.0\%)\tabularnewline
+SD $\rightarrow$  SPYclade &  4 (1.5\%) &  0 (0.0\%)\tabularnewline
+SPYclade $\rightarrow$ SDEclade & 32 (12.1\%) & 10 (25.0\%)\tabularnewline
+SPYclade $\rightarrow$ SDD & 33 (12.5\%) &  3 (7.5\%)\tabularnewline
+SPYclade $\rightarrow$ SD &   3 (1.1\%) &  0 (0.0\%)\tabularnewline
+SPYclade $\rightarrow$ SPYclade & 27 (10.2\%) &  6 (15.0\%)\tabularnewline
+All types  &  265 (100\%) & 40 (100\%)  \tabularnewline
+\hline
+\end{tabular}
+\label{tab:gene-counts-additive}
+\end{table}
+%\clearpage{}
+%=============================================================================
+
+%=============================================================================
+\begin{table}[!ht]
+\caption{
+{\bf Functional categories enriched in 
+recombining transfer events that are inferred using the parsimony-based
+approach.}}
+\noindent \begin{centering}
+\begin{tabular}{ccccl}
+\hline 
+$p^a$ & $q^b$ & Count$^c$ & GO term & Description \\
+\hline 
+7e-09 & 3e-06 & 153 & GO:0006412 & translation\\
+1e-08 & 3e-06 & 67 & GO:0030529 & ribonucleoprotein complex\\
+1e-07 & 2e-05 & 45 & GO:0019843 & rRNA binding\\
+2e-06 & 2e-04 & 73 & GO:0003735 & structural constituent of ribosome\\
+3e-05 & 3e-03 & 157 & GO:0005840 & ribosome\\
+7e-05 & 5e-03 & 221 & GO:0003723 & RNA binding\\
+2e-04 & 1e-02 & 11 & GO:0004826 & phenylalanine-tRNA ligase activity\\
+3e-04 & 2e-02 & 14 & GO:0015413 & nickel-transporting ATPase activity\\
+4e-04 & 2e-02 & 93 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
+5e-04 & 2e-02 & 104 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
+5e-04 & 2e-02 & 13 & GO:0006265 & DNA topological change\\
+5e-04 & 2e-02 & 13 & GO:0003916 & DNA topoisomerase activity\\
+5e-04 & 2e-02 & 733 & GO:0005737 & cytoplasm\\
+6e-04 & 2e-02 & 10 & GO:0046873 & metal ion transmembrane transporter activity\\
+7e-04 & 2e-02 & 12 & GO:0009435 & NAD biosynthetic process\\
+8e-04 & 2e-02 & 19 & GO:0016820 & hydrolase activity, acting on acid anhydrides, \\
+      &       &    &       &catalyzing transmembrane movement of substances \\
+1e-03 & 3e-02 & 20 & GO:0005694 & chromosome\\
+2e-03 & 4e-02 & 31 & GO:0006814 & sodium ion transport\\
+\hline 
+\end{tabular}
+\par\end{centering}
+\begin{flushleft}
+$^a$$P$-values based a Mann-Whitney $U$ test.  Shown are values that
+correspond to FDR$^b$  $<0.05$ based on the 
+Benjamini-Hochberg method (Methods).\\
+$^c$Number of genes in category.\\
+\end{flushleft}
+\label{tab:go-events-recombining}
+\end{table}
+\clearpage{}
+%=============================================================================
+
+%=============================================================================
 \begin{table}[!ht]
@@
 %=============================================================================
-% Table S4. Gene tree topologies
-\begin{table}[!ht]
-\caption{
-{\bf Posterior probability of gene tree topologies.}
-See Table S\ref{tab:genome} for the reference of the bacteria
-individual names.}
-\noindent \begin{centering}
-\begin{tabular}{cc}
-\hline 
-Gene tree topologies & Posterior probability\tabularnewline
-\hline
-(((SDE1,SDE2),SDD),(SPY1,SPY2)) & 66.7\%\tabularnewline
-(((SDE1,SDE2),(SPY1,SPY2)),SDD) & 9.3\%\tabularnewline
-((SDE1,SDE2),(SDD,(SPY1,SPY2))) & 3.5\%\tabularnewline
-(((SDE1,(SDE2,SDD)),(SPY1,SPY2)) & 1.5\%\tabularnewline
-(((SDE1,SDD),SDE2),(SPY1,SPY2)) & 1.3\%\tabularnewline
-((SDE1,SDD),(SDE2,(SPY1,SPY2))) & 1.2\%\tabularnewline
-((((SDE1,SDE2),SPY1),SPY2),SDD) & 1.0\%\tabularnewline
-((SDE2,SDD),(SDE1,(SPY1,SPY2))) & 1.0\%\tabularnewline
-\hline
-\end{tabular}
-\par\end{centering}
-\label{tab:Gene-tree-topologies}
-\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Table S4. The values of the heat map.
-\begin{table}[!ht]
-\caption{
-{\bf Ratios of logarithm base 2 of the observed
-recombinant edges to the expected number of recombinant edges.}  Figure
-\ref{fig:Heatmap-of-recombination} shows the values graphically using heat map.
-NA's are for the branch pairs that are impossible under the model-based
-approached. Rows are for donor, and columns for recipient.}
-\noindent \centering{}\begin{tabular}{cccccccccc}
-\hline
-& SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2 \tabularnewline
-\hline
-SDE1& 0.176  &   NA  &1.170  &    NA & 0.7564  & NA & 0.67 &    NA & 0.663\tabularnewline
-SDE & -0.118 & 0.081 &-0.202 &     NA&  0.3395 &  NA& -0.15&  0.045& -0.323\tabularnewline
-SDE2& 1.201  &   NA  &0.092  &    NA & 0.6978  & NA & 0.60 &    NA & 0.387\tabularnewline
-SD  & -0.629 & 0.022 &-0.661 &0.00670&  0.0078 &  NA& -0.94& -0.299& -1.030\tabularnewline
-SDD & -0.317 & 0.371 &-0.197 &     NA&  0.0752 &  NA& -1.06& -0.525& -0.915\tabularnewline
-ROOT& -1.745 &-0.698 &-1.786 &0.23039& -0.4854 &  NA& -1.37&  0.497& -1.306\tabularnewline
-SPY1& 1.245  &1.883  &1.251  &    NA &-2.1997  & NA & 0.06 &    NA & 1.062\tabularnewline
-SPY & 0.089  &0.200  &0.298 &0.00098 &-0.7903  & NA &-0.19 &-0.058 &-0.159\tabularnewline
-SPY2& 1.174  &1.938  &1.287  &    NA &-1.9740  & NA & 1.07 &    NA & 0.063\tabularnewline
-\hline
-\end{tabular}
-\label{tab:heatmap}
-\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Table S5. The values of the heat map of actual counts.
-\begin{table}[!ht]
-\caption{
-{\bf The observed number of recombinant edges.}
-The zeros are for the branch pairs that are impossible under the model-based
-approached. Rows are for donor, and columns for recipient.}
-\noindent \centering{}\begin{tabular}{cccccccccc}
-\hline
-& SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2 \tabularnewline
-\hline
-SDE1&8.896 &  0.00 &17.712  & 0.00 & 12.872   & 0 &12.59  &  0.00  &12.53\tabularnewline
-SDE &44.626& 136.49& 42.088 &  0.00& 227.061  &  0& 73.39 &  99.95 & 65.25\tabularnewline
-SDE2&18.099&   0.00&  8.393 &  0.00&  12.360  &  0& 11.97 &   0.00 & 10.36\tabularnewline
-SD  &9.671 &101.74 & 9.464  &31.26 &115.796   & 0 &13.68  & 98.38  &12.86\tabularnewline
-SDD &45.539& 168.58& 49.513 &  0.00& 194.706  &  0& 43.09 &  67.14 & 47.65\tabularnewline
-ROOT&43.553& 602.67& 42.337 &765.53& 802.813  &  0& 98.86 &2146.01 &103.60\tabularnewline
-SPY1&40.800&  11.46& 40.973 &  0.00&   4.371  &  0& 21.15 &   0.00 & 42.35\tabularnewline
-SPY &57.618& 262.08& 66.594 & 30.92& 163.518  &  0& 83.56 & 207.00 & 85.10\tabularnewline
-SPY2&38.837&  11.90& 42.000 &  0.00&   5.111  &  0& 42.45 &   0.00 & 21.19\tabularnewline
-\hline
-\end{tabular}
-\label{tab:obsheatmap}
-\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Table S7. Mowgli - Recombining Transfer
-\begin{table}[!ht]
-\caption{
-{\bf Functional categories enriched in 
-recombining transfer events that are inferred using the parsimony-based
-approach.}}
-\noindent \begin{centering}
-\begin{tabular}{cccl}
-\hline 
-P-value & Count & GO-term & Description \\
-\hline 
-7e-09 & 153 & GO:0006412 & translation\\
-1e-08 &  67 & GO:0030529 & ribonucleoprotein complex\\
-1e-07 &  45 & GO:0019843 & rRNA binding\\
-2e-06 &  73 & GO:0003735 & structural constituent of ribosome\\
-3e-05 & 157 & GO:0005840 & ribosome\\
-7e-05 & 221 & GO:0003723 & RNA binding\\
-2e-04 &  11 & GO:0004826 & phenylalanine-tRNA ligase activity\\
-3e-04 &  14 & GO:0015413 & nickel-transporting ATPase activity\\
-4e-04 &  93 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
-5e-04 & 104 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
-\hline 
-\end{tabular}
-\par\end{centering}
-\label{tab:go-events-recombining}
-\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Table S. 
-\begin{table}[!ht]
-\caption{
-{\bf Virulence genes inferred to have experienced recombining gene transfer
-events.}}
-\noindent \begin{centering}
-\begin{tabular}{lll}
-\hline 
-Genes & Gene Accessions\tabularnewline
-\hline 
-putative calcium transporter (MgtA) & SpyM3\_0440 & SPY1 to SDE1\\
-putative dipeptidase & SpyM3\_0465 & SPY to SDE\\
-superoxide dismutase & SpyM3\_1071 & SPY to SDE\\
-putative internalin A precursor & SpyM3\_1035 & SPY to SDE\\
-putative heat shock protein (chaperonin GroEL) & SpyM3\_1765 & SPY to SDE1\\
-pyruvate formate lyase & SpyM3\_1749 & No specific directions\\
-putative glucose kinase & SpyM3\_1180 & SPY to SDE\\
-\hline 
-\end{tabular}
-\par\end{centering}
-\label{tab:virrecomb}
-\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Table S. 
-\begin{table}[!ht]
-\caption{
-{\bf Virulence genes inferred to have experienced horizontal gene transfer
-events.}}
-\noindent \begin{centering}
-\begin{tabular}{ll}
-\hline 
-Genes & Gene Accessions\tabularnewline
-\hline 
-Streptococcal C5a peptidase & SDEG\_0933\\ 
-Putative transposase & SpyM3\_1016, SpyM3\_1053, SpyM3\_1054\\
-putative response regulator & SpyM3\_0174\\
-putative maltose operon transcriptional repressor & SpyM3\_0982\\ 
-putative histidine kinase & SpyM3\_0171\\
-putative cell wall hydrolase & SpyM3\_0731, SpyM3\_0922, SpyM3\_1208\\
-genome-encoded streptodornase & SDEG\_0541\\
-prophage-associated deoxyribonuclease & SDEG\_1103\\
-Protein F2 like fibronectin-binding protein (adhesin and invasin) & SpyM3\_0104\\
-\hline 
-\end{tabular}
-\par\end{centering}
-\label{tab:virhgt}
-\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Table S6. Genes with evident recombining gene transfer.
-\begin{table}[!ht]
-\caption{
-{\bf Genes and their parts transferred via homologous
-recombining gene transfer.}}
-{\footnotesize
-\begin{tabular}{lllllll}
-\hline 
-Locus  & N$^a$  & S$^b$  & D$^c$ & F$^d$(\%)  & C$^e$(\%)  & gene 
-product\tabularnewline
-\hline 
-SpyM3\_0074  &  & 4  & 0  & 20  & 100  & putative penicillin-binding protein 1b 
-\tabularnewline
-SpyM3\_0134  &  & 6  & 0  & 43, 0 & 79  & leucyl-tRNA synthetase \tabularnewline
-SpyM3\_0156  &  & 0  & 4  & 77  & 100  & glucose-6-phosphate isomerase 
-\tabularnewline
-SpyM3\_0186  &  & 5  & 4  & 83  & 94  & putative glucose kinase \tabularnewline
-SpyM3\_0192  & 1 & 6  & 5  & 14, 6 & 100  & ribulose-phosphate 3-epimerase 
-\tabularnewline
-SpyM3\_0193  & 1 & 6  & 5  & 99  & 100  & hypothetical protein\tabularnewline
-SpyM3\_0197  &  & 6  & 5  & 7  & 100  & putative surface exclusion 
-protein\tabularnewline
-SpyM3\_0204  &  & 6  & 5  & 45  & 100  & hypothetical protein \tabularnewline
-SpyM3\_0215  &  & 6  & 5  & 5, 26 & 100  & oligopeptide permease \tabularnewline
-SpyM3\_0238  &  & 5  & 3  & 18  & 100  & hypothetical protein \tabularnewline
-SpyM3\_0292  &  & 6  & 5  & 27  & 100  & putative tetrapyrrole methylase family 
-protein \tabularnewline
-SpyM3\_0429  &  & 2  & 1  & 0,22  & 100  & putative oligoendopeptidase 
-F\tabularnewline
-SpyM3\_0465  & 2 & 6  & 5  & 95  & 99  & putative dipeptidase \tabularnewline
-SpyM3\_0466  & 2 & 6  & 5  & 16  & 100  & putative adhesion protein 
-\tabularnewline
-SpyM3\_0506  &  & 5  & 3  & 52  & 100  & phenylalanyl-tRNA synthetase subunit 
-alpha \tabularnewline
-SpyM3\_0580  & 3 & 3  & 0  & 3  & 100  & PTS system, fructose-specific IIABC 
-component\tabularnewline
-SpyM3\_0581  & 3 & 3  & 0  & 18  & 100  & putative peptidoglycan hydrolase 
-\tabularnewline
-SpyM3\_0581  &  & 6  & 5  & 11  & 100  & putative peptidoglycan hydrolase 
-\tabularnewline
-SpyM3\_0745  & 4 & 6  & 1  & 14  & 100  & putative two-component sensor 
-histidine kinase \tabularnewline
-SpyM3\_0746  & 4 & 6  & 1  & 55  & 100  & putative two-component response 
-regulator \tabularnewline
-SpyM3\_0785  & 5 & 4  & 1  & 78  & 92  & inorganic polyphosphate/ATP-NAD 
-kinase\tabularnewline
-SpyM3\_0786  & 5 & 4  & 1  & 100  & 100  & putative ribosomal large subunit 
-pseudouridine synthase \tabularnewline
-SpyM3\_0787  & 5 & 4  & 1  & 12  & 100  & phosphotransacetylase \tabularnewline
-SpyM3\_1078  & 6 & 4  & 1  & 21  & 100  & putative cation/potassium uptake 
-protein \tabularnewline
-SpyM3\_1079  & 6 & 4  & 1  & 20  & 100  & putative ATP-dependent RNA helicase 
-\tabularnewline
-SpyM3\_1093  &  & 4  & 0  & 30  & 100  & putative heavy 
-metal/cadmium-transporting ATPase \tabularnewline
-SpyM3\_1151  &  & 3  & 1  & 34  & 100  & putative DNA repair and genetic 
-recombination protein \tabularnewline
-SpyM3\_1153  & 7 & 4  & 0  & 57  & 100  & putative hemolysin\tabularnewline
-SpyM3\_1154  & 7 & 4  & 0  & 4  & 100  & putative geranyltranstransferase 
-\tabularnewline
-SpyM3\_1201  &  & 6  & 5  & 21  & 100  & putative two-component sensor 
-histidine kinase \tabularnewline
-SpyM3\_1267  & 8 & 6  & 5  & 7  & 100  & bifunctional methionine sulfoxide 
-reductase A/B protein \tabularnewline
-SpyM3\_1268  & 8 & 6  & 5  & 8  & 100  & hypothetical protein \tabularnewline
-SpyM3\_1365  &  & 3  & 0  & 28  & 100  & putative Cof family 
-protein/peptidyl-prolyl cis-trans isomerase, cyclophilin
-type\tabularnewline
-SpyM3\_1380  & 9 & 6  & 1  & 98  & 100  & putative Acetyl-CoA:acetoacetyl-CoA 
-transferase b subunit \tabularnewline
-SpyM3\_1381  & 9 & 6  & 1  & 100  & 100  & putative oxidoreductase 
-\tabularnewline
-SpyM3\_1382  & 9 & 6  & 1  & 23  & 38  & 3-hydroxybutyrate dehydrogenase 
-\tabularnewline
-SpyM3\_1512  &  & 4  & 0  & 59  & 100  & PTS system, mannose-specific IIC 
-component \tabularnewline
-SpyM3\_1518  & 10 & 1  & 2  & 63  & 100  & acetyl-CoA carboxylase subunit 
-beta\tabularnewline
-SpyM3\_1519  & 10 & 1  & 2  & 100  & 100  & acetyl-CoA carboxylase biotin 
-carboxylase subunit \tabularnewline
-SpyM3\_1520  & 10 & 1  & 2  & 100  & 100  & (3R)-hydroxymyristoyl-ACP 
-dehydratase \tabularnewline
-SpyM3\_1521  & 10 & 1  & 2  & 100  & 100  & acetyl-CoA carboxylase biotin 
-carboxyl carrier protein subunit\tabularnewline
-SpyM3\_1522  & 10 & 1  & 2  & 100  & 100  & 3-oxoacyl-(acyl carrier protein) 
-synthase II \tabularnewline
-SpyM3\_1523  & 10 & 1  & 2  & 100  & 100  & 3-ketoacyl-(acyl-carrier-protein) 
-reductase \tabularnewline
-SpyM3\_1524  & 10 & 1  & 2  & 100  & 100  & acyl-carrier-protein 
-S-malonyltransferase\tabularnewline
-SpyM3\_1525  & 10 & 1  & 2  & 100  & 100  & putative trans-2-enoyl-ACP 
-reductase II \tabularnewline
-SpyM3\_1561  &  & 6  & 5  & 11  & 97  & hypothetical protein \tabularnewline
-SpyM3\_1577  &  & 4  & 1  & 11  & 100  & excinuclease ABC subunit A 
-\tabularnewline
-SpyM3\_1579  &  & 4  & 1  & 1  & 100  & hypothetical protein \tabularnewline
-SpyM3\_1591  &  & 6  & 5  & 5  & 100  & ribonuclease HIII \tabularnewline
-SpyM3\_1753  &  & 6  & 5  & 98  & 100  & putative transcriptional regulator 
-\tabularnewline
-SpyM3\_1753  & 11 & 3  & 1  & 53  & 100  & putative transcriptional regulator 
-\tabularnewline
-SpyM3\_1754  & 11 & 3  & 1  & 54  & 100  & putative transcriptional regulator 
-\tabularnewline
-SpyM3\_1754  &  & 6  & 5  & 95  & 100  & putative transcriptional regulator 
-\tabularnewline
-SpyM3\_1778  &  & 6  & 0  & 34  & 100  & putative cationic amino acid 
-transporter protein \tabularnewline
-SpyM3\_1809  &  & 4  & 0  & 5  & 100  & arginyl-tRNA synthetase \tabularnewline
-SpyM3\_1813  & 12 & 3  & 0  & 19  & 100  & hypothetical protein \tabularnewline
-SpyM3\_1814  & 12 & 3  & 0  & 29  & 100  & aspartyl-tRNA 
-synthetase\tabularnewline
-\hline
-\end{tabular}
-
-\begin{flushleft}
-$^a$ Numbers represent a group of loci that are consecutive in gene order. Empty
-cells denote stand alone loci,
-
-$^b$ Donor branches of homologous recombinant edges, 0 (SDE1), 1 (SDE2), 2
-(SDD), 3 (SPY1), 4 (SPY2), 5 (SDE), 6 (SPY), 7(SD), and 8(ROOT),
-
-$^c$ Recipient branches of recombinant edges,
-
-$^d$ Proportion of sites that the recombinant edge covers out of the sites in
-the core alignment block, and
-
-$^e$ Percentage of the sites in the core alignment block out of total sites of
-the locus.
-\end{flushleft}
-}
-\label{tab:genes-transfer}
-\end{table}
-\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Table S8. ClonalOrigin - All Topologies
-\begin{table}[!ht]
-\caption{
-{\bf Functional categories associated with 
-recombination intensity based on all of the recombinant edge types.}}
-\noindent \begin{centering}
-\begin{tabular}{cccl}
-\hline 
-P-Value & Count & GO term & Description\tabularnewline
-\hline 
-4.596e-06 & 26 & GO:0019843 & rRNA binding\tabularnewline
-5.395e-06 & 40 & GO:0030529 & ribonucleoprotein complex\tabularnewline
-7.935e-06 & 39 & GO:0003735 & structural constituent of ribosome\tabularnewline
-1.546e-05 & 81 & GO:0006412 & translation\tabularnewline
-3.518e-05 & 66 & GO:0005840 & ribosome\tabularnewline
-4.288e-05 & 22 & GO:0006633 & fatty acid biosynthetic process\tabularnewline
-9.098e-05 & 20 & GO:0008610 & lipid biosynthetic process\tabularnewline
-0.0003186 & 33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
-0.000357 & 45 & GO:0008643 & carbohydrate transport\tabularnewline
-0.0006318 & 38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
-0.0009886 & 105 & GO:0003723 & RNA binding\tabularnewline
-\hline 
-\end{tabular}
-\par\end{centering}
-\label{tab:functional-all}
-\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Table S. Simulation study for the three population parameters.
-\begin{table}[!ht]
-\caption{
-{\bf Functional categories asscoiated with 
-recombination intensity based on the topology-not-changing recombinant edge
-types.}}
-\noindent \begin{centering}
-\begin{tabular}{cccl}
-\hline 
-P-Value & Count & GO term & Description\tabularnewline
-\hline 
-6.515e-05 & 39 & GO:0003735 & structural constituent of ribosome\tabularnewline
-0.00020986 & 40 & GO:0030529 & ribonucleoprotein complex\tabularnewline
-0.00045049 & 66 & GO:0005840 & ribosome\tabularnewline
-0.00052925 & 26 & GO:0019843 & rRNA binding\tabularnewline
-\hline 
-\end{tabular}
-\par\end{centering}
-\label{tab:functional-notopology}
-\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Table S. ClonalOrigin - SDE-to-SPY
-%\begin{table}
-%\caption{\label{tab:functional-sde-spy}Functional categories asscoiated with 
-%recombination intensity based on SDE-to-SPY recombinant edge types.}
-%\noindent \begin{centering}
-%\begin{tabular}{cccl}
-%\hline 
-%P-Value & Count & GO term & Description\tabularnewline
-%\hline 
-%1.854e-05 & 38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
-%2.703e-05 & 81 & GO:0006412 & translation\tabularnewline
-%6.482e-05 & 33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
-%9.777e-05 & 105 & GO:0003723 & RNA binding\tabularnewline
-%0.00012793 & 40 & GO:0030529 & ribonucleoprotein complex\tabularnewline
-%0.00025176 & 26 & GO:0019843 & rRNA binding\tabularnewline
-%0.00031236 & 72 & GO:0005975 & carbohydrate metabolic process\tabularnewline
-%0.00044756 & 317 & GO:0005737 & cytoplasm\tabularnewline
-%0.00059675 & 45 & GO:0008643 & carbohydrate transport\tabularnewline
-%0.00066874 & 39 & GO:0003735 & structural constituent of ribosome\tabularnewline
-%0.00135304 & 10 & GO:0006098 & pentose-phosphate shunt\tabularnewline
-%\hline 
-%\end{tabular}
-%\par\end{centering}
-%\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Table S. ClonalOrigin - SPY-to-SDE
-%\begin{table}
-%\caption{\label{tab:functional-spy-sde}Functional categories asscoiated with 
-%recombination intensity based on SPY-to-SDE recombinant edge types.}
-%\noindent \begin{centering}
-%\begin{tabular}{cccl}
-%\hline 
-%P-Value & Count & GO term & Description\tabularnewline
-%\hline 
-%1.106e-06 & 45 & GO:0008643 & carbohydrate transport\tabularnewline
-%1.511e-06 & 33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
-%1.595e-06 & 38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
-%1.443e-05 & 187 & GO:0006810 & transport\tabularnewline
-%0.0001106 & 371 & GO:0016740 & transferase activity\tabularnewline
-%0.0003596 & 72 & GO:0005975 & carbohydrate metabolic process\tabularnewline
-%0.0004023 & 13 & GO:0006814 & sodium ion transport\tabularnewline
-%0.000823 & 317 & GO:0005737 & cytoplasm\tabularnewline
-%0.001142 & 12 & GO:0031402 & sodium ion binding\tabularnewline
-%\hline 
-%\end{tabular}
-%\par\end{centering}
-%\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Table S. Mowgli - Duplication 
-%\begin{table}
-%\caption{\label{tab:go-events-duplication}Functional categories enriched in 
-%gene duplication events that are inferred using the parsimony-based approach.}
-%\noindent \begin{centering}
-%\begin{tabular}{cccl}
-%\hline 
-%P-value & Count & GO-term & Description \\
-%\hline 
-%4e-81 &  67 & GO:0006313 & transposition, DNA-mediated\\
-%5e-80 &  59 & GO:0004803 & transposase activity\\
-%6e-70 &  23 & GO:0032196 & transposition\\
-%2e-58 &  52 & GO:0015074 & DNA integration\\
-%8e-25 &  86 & GO:0006310 & DNA recombination\\
-%1e-20 &  36 & GO:0016987 & sigma factor activity\\
-%1e-20 &  36 & GO:0006352 & transcription initiation\\
-%2e-19 & 197 & GO:0003676 & nucleic acid binding\\
-%2e-10 & 527 & GO:0003677 & DNA binding\\
-%4e-06 & 142 & GO:0043565 & sequence-specific DNA binding\\
-%\hline 
-%\end{tabular}
-%\par\end{centering}
-%\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Table S. Mowgli - Loss
-%\begin{table}
-%\caption{\label{tab:go-events-loss}Functional categories enriched in 
-%gene loss events that are inferred using the parsimony-based approach.}
-%\noindent \begin{centering}
-%\begin{tabular}{cccl}
-%\hline 
-%P-value & Count & GO-term & Description \\
-%\hline 
-%1e-06 &  67 & GO:0006313 & transposition, DNA-mediated\\
-%2e-05 &  59 & GO:0004803 & transposase activity\\
-%5e-05 &  52 & GO:0015074 & DNA integration\\
-%5e-04 &  23 & GO:0032196 & transposition\\
-%5e-04 &  10 & GO:0046873 & metal ion transmembrane transporter activity\\
-%7e-04 &  36 & GO:0016987 & sigma factor activity\\
-%7e-04 &  36 & GO:0006352 & transcription initiation\\
-%1e-03 &  31 & GO:0006814 & sodium ion transport\\
-%3e-03 &  86 & GO:0006310 & DNA recombination\\
-%6e-03 & 197 & GO:0003676 & nucleic acid binding\\
-%\hline 
-%\end{tabular}
-%\par\end{centering}
-%\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Table S15. Simulation study for the three population parameters.
 \begin{table}[!ht]
@@
 \hline
- & True & Estimates (Standard Deviation)\tabularnewline
-\hline
-Mutation rate per site & 0.081 & 0.081 (0.0011)\tabularnewline
-Recombination rate per site & 0.012 & 0.011 (0.0003)\tabularnewline
-Recombinant tract length & 744 bp & 919 bp (28)\tabularnewline
+Parameter & True & Estimates (Standard Deviation)\tabularnewline
+\hline
+$\theta^a$ & 0.081 & 0.081 (0.0011)\tabularnewline
+$\rho^b$   & 0.012 & 0.011 (0.0003)\tabularnewline
+$\delta^c$ & 744 bp & 919 bp (28)\tabularnewline
 \hline
@@
 \par\end{centering}
+\begin{flushleft}
+$^a$ Mutation rate per site.
+
+$^b$ Recombination rate per site.
+
+$^c$ Recombinant tract length.
+\end{flushleft}
 \label{tab:sim-three}
@@
 
-\clearpage{}
 \section*{Supplementary Figures}
@@
 %=============================================================================
-% Figure S1: Mauve alignment of the five genomes.
-\begin{figure}[!ht]
+\begin{sidewaysfigure}[!ht]
 \begin{center}
-\includegraphics[scale=0.25]{figures/mauve}
+\includegraphics[scale=0.3]{figures/mauve}
 \end{center}
@@
 genome. The fourth and fifth rows are SPY1 and SPY2 genomes, respectively.}
-\end{figure}
-\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Figure S: Histogram of block sizes
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=5in]{figures/blocksize}
-%\end{center}
-%\caption{\label{fig:blocksize}Length distribution of the 274 core alignment blocks.}
-%\end{figure}
-%\clearpage{}%
-%=============================================================================
-
-%=============================================================================
-% Figure S2. Three population parameters.
-\begin{figure}[!ht]
-\begin{center}
-\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot theta}}
-
-\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot rho}}
-
-\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot delta}}
-\end{center}
-\caption{
-{\bf Three scatter plots of mutation rate, recombination
-rate, and average tract length for the blocks.} The plus signs are mean values
-for each block. The density of each value for each block is depicted by black
-clouds.  The red dashed lines are the global block-length weighted median of the
-three parameter estimates.}
-\end{figure}
-\label{fig:scatter3}
-\clearpage{}%
-%=============================================================================
-
-%=============================================================================
-% Figure S3: gene family sizes
-% FIGURE - gene family sizes
+\end{sidewaysfigure}
+%=============================================================================
+
+%=============================================================================
 \begin{figure}[!ht]
@@
 %=============================================================================
-% FIGURE S4. - HGT Transfer heatmap
 \begin{figure}[!ht]
 \begin{center}
-\includegraphics[scale=0.45]{figures/heatmap-hgtedge}
+\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-theta}
+
+\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-rho}
+
+\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-delta}
 \end{center}
 \caption{
-{\bf Heat map of horizontal gene transfers using the six genomes.}
-Each cell of the color matrix shows the number of inferred non-recombining
-horizontal gene transfer events between the different branches of the species
-tree. Black colored cells represent no inferred events.}
-\label{fig:hgt-heatmap}
+{\bf Three scatter plots of mutation rate, recombination
+rate, and recombinant tract length for the blocks.} The plus signs are mean values
+for each block. The density of each value for each block is depicted by black
+clouds.  The red dashed lines are the global block-length weighted median of the
+three parameter estimates.}\label{fig:scatter3}
 \end{figure}
@@
 %=============================================================================
-% Figure S5: recombination heatmap from mowgli
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[scale=0.45]{figures/heatmap-recedge-mowgli}
-\end{center}
-\caption{
-{\bf Heat map of recombining gene transfers using the six genomes.}
-Each cell of the color matrix shows the number of inferred recombining
-gene transfer events between the different branches of the species
-tree. Black colored cells represent no inferred events.}
-\label{fig:mowgli-recomb-heatmap}
-\end{figure}
-\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Figure S6: Comparison of ClonalOrigin and Mowgli analysis
 \begin{figure}[!ht]
@@
 {\bf Plot of recombination intensity against number
-of recombining gene transfer.}}
+of replacing gene transfer.}}
 \label{fig:cmpcomowgli}
@@
 %=============================================================================
-% Figure S7: HGT Comparison of ClonalOrigin and Mowgli analysis
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[width=5in]{figures/compareMowgliCoHgt}
-\end{center}
-\caption{
-{\bf Plot of recombination intensity against number
-of horizontal gene transfer.}}
-\label{fig:cmpcomowglihgt}
-\end{figure}
-\clearpage{}
-%=============================================================================
-
-%=============================================================================
-% Figure S5: Scatter plot of rho and log(delta)
-% \begin{figure}
-% \begin{center}
-% \includegraphics[width=5in]{figures/rho-vs-delta}
-% \end{center}
-% \caption{Plot of recombination rate and logarithm of average tract length
-% of the posterior samples.}
-% \label{fig:rhologdelta}
-% \end{figure}
-% \clearpage{}%
-%=============================================================================
-
-%=============================================================================
-% Figure S8: View of UCSC genome browser for bacterial gene transfer.
-\begin{figure}[!ht]
-\includegraphics[scale=0.8]{figures/ucsc}
-\caption{
-{\bf UCSC genome browser tracks for gene transfer study of
-SPY and SDE.}}
-\label{fig:ucsc}
-\end{figure}
-\clearpage{}%
-%=============================================================================
-
-%=============================================================================
-% Figure S6. A Recombinant Tree of ClonalOrigin Model
 \begin{figure}[!ht]
@@
 \caption{
-{\bf Recombinant tree in the \texttt{ClonalOrigin} model.}
+{\bf Recombinant tree in the ClonalOrigin model.}
 (A) A recombinant tree relating the five genomes with four recombinant edges.
@@
 %=============================================================================
-% Figure S13:
-\begin{figure}[!ht]
-\includegraphics{figures/ri1}
+\begin{sidewaysfigure}
+
+\includegraphics[scale=0.6]{figures/sim2-1}
+
+\includegraphics[scale=0.6]{figures/sim2-2}
+
 \caption{
-{\bf The surrogate measure of recombination intensity
-of genes.} Each point represents a pair of a true value of average number of
-recombinant edges and its estimated value for a gene. The dashed line is with
-slope of 1 and intercept of 0.}
-\label{fig:ri1}
+{\bf Simulations of the model-based approach for assessing the power of
+recovering recombinant edges using prior recombinant trees.} 
+At X-axis labels are shown pairs of species tree branches, for each of which one
+interval with an open circle and a closed circle next to it
+are displayed.  Intervals with open circle are for 
+the ratio of the number of recombinant edges inferred from the 100
+simulated data sets under the prior with respect to the expected number of
+recombinant edges under the prior.  Closed circles are for 
+the ratio of the number of recombinant edges estimated
+from the real data with respect to the expected number of recombinant edges
+under the prior.  The horizontal lines are at the zero.  The intervals range
+from 5\% quantile to 95\% of quantile.}
+\label{fig:sim2}
+\end{sidewaysfigure}
+\clearpage{}%
+%=============================================================================
+
+%=============================================================================
+\begin{sidewaysfigure}
+
+\includegraphics[scale=0.6]{figures/sim3-1}
+
+\includegraphics[scale=0.6]{figures/sim3-2}
+
+\caption{
+{\bf Simulations of the model-based approach for assessing the power of
+recovering recombinant edges using posterior recombinant trees.} 
+At X-axis labels are shown pairs of species tree branches, for each of which one
+interval with an open circle and an closed circle next to it
+are displayed.  Intervals with open circle are for 
+the ratio of the number of recombinant edges inferred from the 100
+simulated data sets with the posterior samples of recombinant trees 
+with respect to the expected number of
+recombinant edges under the prior.  Closed circles are for 
+the ratio of the number of recombinant edges estimated
+from the real data with respect to the expected number of recombinant edges
+under the prior.  The horizontal lines are at the zero.  The intervals range
+from 5\% quantile to 95\% of quantile.}
+\label{fig:sim3}
+\end{sidewaysfigure}
+\clearpage{}%
+%=============================================================================
+
+%=============================================================================
+\begin{sidewaysfigure}
+
+% \includegraphics[width=6in]{figures/real_vs_null_additive.pdf}
+\includegraphics[scale=0.6]{figures/sim4-1}
+
+% \includegraphics[width=4in]{figures/real_vs_null_dup.pdf}
+\includegraphics[scale=0.6]{figures/sim4-2}
+
+\caption{
+{\bf The power of recovering additive transfers and duplications using simulations.} (top) For each pairs of species tree branches and clades SDE and SPY,
+we plot the number of reconstructed additive transfers in the real data 
+(circle) normalized
+by the true number of transfers in our simulations.  Additive transfers
+reconstructed from simulations are also normalized by the true number
+and are plotted with 90\% confidence intervals.
+(bottom) For each branch in the species tree, we plot the number of 
+reconstructed duplications in the real data (circles) normalized by the true
+number of duplications in our simulations.  Duplications
+reconstructed from simulations are also normalized by the true number
+and are plotted with 90\% confidence intervals.
+}
+\label{fig:mowgli-sim}
+\end{sidewaysfigure}
+\clearpage{}%
+%=============================================================================
+
+%=============================================================================
+\begin{figure}
+\begin{center}
+\includegraphics[width=5in]{figures/calling}
+\end{center}
+\caption{
+{\bf Distinguishing between replacing and additive horizontal gene transfers.}
+We consider a transfer ``additive'' if there exists an extant descendant species
+(e.g. C and D) that contains both descendants (bold) and
+non-descendants (non-bold) of the transferred gene (black circle).}
+\label{fig:calling-transfers}
 \end{figure}
-
 \clearpage{}%
@@
 
-%=============================================================================
-% Figure S13: 
-%\begin{figure}
-%\subfloat[]{\includegraphics[scale=0.5]{figures/h1-R}}
-%\subfloat[]{\includegraphics[scale=0.5]{figures/h1-R-inset}}
-%\caption{\label{fig:h1}Numer of recombination events. The diagonal dashed
-%lines are with slope of 1 and intercept of 0. For each order
-%pairs of branches of a reference species tree the true number of recombination
-%events is computed as the mean of the 10 replicates of recombinant
-%trees from the simulation. With ten estimates of each replicate total
-%number of 100 estimats are shown as intervals of 90\% of quantile
-%range, or the lower value of the interval is the 5\% quantile, and
-%the upper one is the 95\% quantile. }
-%\end{figure}
-%\clearpage{}%
-
-%=============================================================================
-%=============================================================================
-% Figure S:
-%\begin{figure}
-%\subfloat[]{\includegraphics[scale=0.5]{figures/h2-R}}
-%\subfloat[]{\includegraphics[scale=0.5]{figures/h2-R-inset}}
-
-%\subfloat[]{\includegraphics[scale=0.5]{figures/h2-R-inset2}}
-
-%\caption{\label{fig:h2}Comparison of number of recombinations between real
-%and simulated data sets. The diagonal dashed lines are with
-%slope of 1 and intercept of 0. For each order pairs of branches of
-%a reference species tree the number of recombination events is estimated
-%from the real data set. Total number of 100 estimats from a simulated
-%data set are shown as intervals of 90\% of quantile range, or the
-%lower value of the interval is the 5\% quantile, and the upper one
-%is the 95\% quantile. }
-%\end{figure}
-%\clearpage{}%
-%=============================================================================
-
 \clearpage{}
-\section*{Supplementary Information}
-
-\subsection*{Summarization of posterior samples for local gene trees and recombination intensity}
-
-Because the procedure of summarizing posterior samples of \textit{recombinant
-tree}, which will be defined in a moment, was not detailed in the user guide at
-the time when we used \texttt{ClonalOrigin}, we elected to use Figure
-S\ref{fig:clonalorigin} to illustrate a recombinant tree in order to clarify
-summarization of \textit{recombinant history} posterior samples.  A
-recombination history is represented by a \textit{recombinant tree}, which is a
-species tree augmented with \textit{recombinant edges}.  A recombinant edge
-connect two species tree branches.  Each recombinant edge is defined by the time
-and branch of origin and destination, and the genomic interval that it affects.
-The branch of origin is referred to as the \textit{donor} and the destination
-branch is referred to as the \textit{recipient}. The type of a recombinant edge
-is determined by the donor and recipient species tree branches.  A recombinant
-tree can be viewed as a restricted version of an ancestral recombination graph
-(ARG), and as such it induces a local genealogy for each site in the genomic
-block it describes.  Didelot et al.\ \citet{Didelot2010} presented two procedures of summarizing
-recombinant trees by viewing trees from two different angles: by counting types
-of recombinant edges regardless of genomic positions affected by the edges, and
-by measuring recombinant edge changes to the species tree in site-wise manner.
-By similarly following their approaches we quantified additional aspects of
-recombinant trees; we counted gene tree topologies for each site, and also
-quantified recombination intensity along the genome.
+\section*{Supplementary Material}
+
+% \subsection*{Preliminary ClonalFrame and ClonalOrigin analysis}
+% 
+% {\bf IG: we should write here details about the first part of the analysis: data
+% preparation, estimating species tree, and three population parameters. TBD. For
+% now we have a collection of details extracted from previous version of
+% results.}
+% \texttt{SC: Do we need this section because the method seems to describe everything enough
+% in detail?}
+% 
+% A genome alignment of the five genomes showed two
+% distinct genome structures of SPY and SDE (Figure S\ref{fig:mauve}).
+% 
+% ClonalFrame estimated a relative strength of
+% recombination compared to mutations as $r/m=4.66$ meaning that four times more
+% sites had been affected by recombination-driven substitutions than by
+% mutation-driven substitutions (Table S\ref{tab:clonalframe}).  This indicated
+% potentially significant roles of recombination in the evolutionary process of
+% shaping these genomes.  Estimates of population mutation rate ($\theta$),
+% recombination rate ($\rho$), and recombinant tract length ($\delta$) were obtained
+% for the alignment blocks analyzed (Figure S\ref{fig:scatter3}), and a median of
+% the estimates weighted over block lengths was obtained as a global estimate for
+% each of the three parameters (Table S\ref{tab:three}).  The local estimates
+% typically seemed to be moderately fluctuating around the global estimate, with
+% the exception of what appeared to be a recombination hotspot at a region 200
+% kilobase pairs from the origin of replication of SDE1.
+
+\subsection*{Summarization of recombinant tree samples}
+We describe the four summarizing procedures including tree topology along the
+genome, numbers of recombinant edges, probability of recombination, and
+recombination intensity using supplementary fig.\ S\ref{fig:clonalorigin}.  
+Supplementary figures S\ref{fig:clonalorigin}A 
+and S\ref{fig:clonalorigin}B fully specify a
+recombinant tree.  First, the recombinant tree induces  five gene tree
+topologies along the genome of size being 100 bp. The fraction of gene tree
+topology of supplementary figure S\ref{fig:clonalorigin}C is 0.15, and 
+that of supplementary figure S\ref{fig:clonalorigin}G is 0.05. 
+The fractions of the other three gene tree
+topologies of 
+supplementary figures S\ref{fig:clonalorigin}D, S\ref{fig:clonalorigin}E, 
+and S\ref{fig:clonalorigin}F are all 0.2.  Second, the heatmap values are computed
+by counting edges for each of 60 possible types and dividing the number of edges
+by prior expected number of recombinant edges given the species tree and the
+three global population parameters.  One edge for each of the four types of
+recombinant edges is observed in the recombinant tree of Figure
+S\ref{fig:clonalorigin}A.  Third, the probability of recombination at a site for
+a particular recombinant edge type is the fraction of recombinant trees, in
+which the recombinant edge spans the site.  
+Supplementary figure S\ref{fig:clonalorigin}B
+shows the probability of recombination along the 100 bp genomic positions.
+Branch SDE1 received the genomic segments ranging from 11 bp to 100 bp from
+branch SDE.  Similarly, branch SDE2 received the segments ranging from 31 bp to
+70 bp from branch SPY1.  Each of recombinant trees from a posterior sample would
+induce a map of recombination such as 
+supplementary figure S\ref{fig:clonalorigin}B.
+Recombination maps from samples of recombinant trees can be averaged as shown in
+figure \ref{fig:ucsc}.  
+Fourth, the recombination intensity of a site is the
+number of recombinant edge types at the site 
+(supplementary fig.\ S\ref{fig:clonalorigin}B).
+Note that it is not the number of recombinant edges.
 
@@
 recombination events in general, or regions enriched for recombination events of
-certain types (e.g., between SPY and SDE).  We noted that the recombination
-intensity was not normalized to the interval of $[0,1]$. Rather than giving
-probabilistic meanings to the recombination intensity we interpreted it as an
-average number of recombinant edge types per site.  The measure appeared to act
-as an effective indicator of recombination intensity (see subsection of
-\textbf{Simulation Study}).  Our simulation study indicated that edges of
+certain types (e.g., between SPY and SDE).  
+Our simulation study indicated that edges of
 certain types were better indicators of recombination events than others.  For
@@
 
-\subsection*{Simulation 1 - three population parameters}
-
+\subsection*{Comparison of model-based and parsimony-based inferences on
+replacing gene transfer}
+
+The similarity between the two lists of gene categories found to be associated
+with replacing gene transfers in the separate analyses -- model-based and
+parsimony-based --  indirectly indicated concordance between the two independent
+approaches.  In order to obtain a more direct measure of correlation, we used
+the set of genes present in both analyses to compute the correlation between
+recombination intensity scores and the number of replacing gene transfers
+inferred by the parsimony-based approach. We observed a positive correlation of
+$0.39$ (P-value $2.2\times10^{-16}$; supplementary fig.\ S\ref{fig:cmpcomowgli})
+between the
+two measures.  For comparison, the correlation between recombination intensity
+and the number of additive gene transfers was 0.03 (P-value 0.18).
+%Figure S\ref{fig:cmpcomowglihgt}).
+
+\subsection*{Simulation Studies for the model-based and parsimony-based approaches}
+We performed four sets of simulations.  In the first simulation, we 
+showed that ClonalOrigin could infer population mutation rate, tract
+length, and recombination rate with reasonable accuracy from our data sets of
+the five individual bacteria. In the second
+simulation, we checked whether recombinant edges were recovered from
+simulated data with recombinant trees generated from the prior under the
+parameters estimated from our data set. The third
+simulation was conducted using simulated data with recombinant trees sampled
+from the
+posterior distribution.  We used a simulation feature of ClonalOrigin
+to simulate the process of mutation and recombination across the five species.
+In the fourth simulation, we checked if the parsimony-based approach worked with
+our data set.
+
+\textbf{Simulation 1 - three population parameters.}
 We describe a procedure of simulations for checking if the three
 population parameters were accurately recovered using the model-based approach.
-Ten recombinant trees were sampled given the three parameters (Table
-S\ref{tab:three}) and the species tree (Figure S\ref{fig:tree5}) set to the
+Ten recombinant trees were sampled given the three parameters (table
+S\ref{tab:three}) and the species tree (fig.\ \ref{fig:tree5}) set to the
 values inferred from the five bacterial genomes.  Each of the recombinant trees
@@
 
-\subsection*{Simulation 2 - numbers of recombinant edges from prior}
-
-We first sampled ten recombinant trees given the mutation rate, recombination
-rate, and tract length (Table S\ref{tab:three}) and the species tree (Figure
-S\ref{fig:tree5}), and then created ten independent simulated data sets with
-the Jukes-Cantor model for DNA sequence evolution using each of the recombinant
-trees. This consisted of 100 replicates of data sets.  The recombinant trees
-were created from the prior distribution and recombinant edges were added to the
-species tree with no particular directions of recombining gene transfers between
-species tree branches.  We computed the fraction of the mean number of recombinant
-edges for a pair of donor and recipient species tree branches over the simulated
-10 recombinant trees against the median of recombinant edges estimated from the
-10 simulated data sets, and then took the logarithm of the fraction to base 2. We
-also computed 5\% and 95\% quantiles to find the 90\% quantile range for the log
-base 2 fraction.  We expect that the log base 2 fraction would near zero. We also
-wished to see if how the recombinant edge count estimates from the real data set
-deviate from the prior. Instead of using the number of recombinant edges from
-the ten recombinant trees, we used the numbers of recombinant edges that were
-estimated from the real data.
-
-We tested if our measure of recombination intensity provided a reliable estimate
-of the true intensity. We artificially annotated our synthetic blocks to genes
-according to the real annotations from SPY1 genome.  For each gene in each of
-the simulated 100 data sets, we plotted the average estimated recombination
-intensity along that gene against the average number of different recombinant
-edge types from the ten simulated recombinant tree.
+\textbf{Simulation 2 - numbers of recombinant edges from prior.}
+Because the number of statistics increases compared with Simulation 1, we
+increased the sample size from 10 to 100.  The recombinant trees were created
+from the prior distribution by adding Recombinant edges to the species tree with
+no particular directions of replacing gene transfers between species tree
+branches.  We computed the mean number of recombinant edges of each type from
+the inference with each of the 100 data sets. Dividing the mean number by the
+prior expected number of recombinant edges we obtained the heatmap value.  The
+distribution of the 100 heatmap values in the scale of logarithm to base 2 is
+shown at supplementary figure S\ref{fig:sim2}.  
+We compared this distribution from simulation
+with the heatmap values from the real data analysis.  We counted recombinant
+edges for the recombinant tree from each iteration of the posterior sample, and
+divided the number of edges by the prior expected number of edges. This allowed
+us to measure variation among recombinant trees in the posterior sample.  The
+distribution of the 1001 heatmap values in the scale of logarithm to base 2 is
+also shown next the simulation result at supplementary figure S\ref{fig:sim2}.
+
+% The following was used for checking recombination intensity of genes. We
+% decided not to describe it.
+% 
+% We first sampled ten recombinant trees given the mutation rate, recombination
+% rate, and tract length (table S\ref{tab:three}) and the species tree (Figure
+% \ref{fig:tree5}), and then created ten independent simulated data sets with
+% the Jukes-Cantor model for DNA sequence evolution using each of the recombinant
+% trees. This consisted of 100 replicates of data sets.  The recombinant trees
+% were created from the prior distribution and recombinant edges were added to the
+% species tree with no particular directions of replacing gene transfers between
+% species tree branches.  We computed the fraction of the mean number of recombinant
+% edges for a pair of donor and recipient species tree branches over the simulated
+% 10 recombinant trees against the median of recombinant edges estimated from the
+% 10 simulated data sets, and then took the logarithm of the fraction to base 2. We
+% also computed 5\% and 95\% quantiles to find the 90\% quantile range for the log
+% base 2 fraction.  We expect that the log base 2 fraction would near zero. We also
+% wished to see if how the recombinant edge count estimates from the real data set
+% deviate from the prior. Instead of using the number of recombinant edges from
+% the ten recombinant trees, we used the numbers of recombinant edges that were
+% estimated from the real data. 
+% 
+% We tested if our measure of recombination intensity provided a reliable estimate
+% of the true intensity. We artificially annotated our synthetic blocks to genes
+% according to the real annotations from SPY1 genome.  For each gene in each of
+% the simulated 100 data sets, we plotted the average estimated recombination
+% intensity along that gene against the average number of different recombinant
+% edge types from the ten simulated recombinant tree.
 
@@
 
-\subsection*{Simulation 3 - number of recombinant edges from posterior}
-
+\textbf{Simulation 3 - number of recombinant edges from posterior.}
 We used 100 recombinant trees sampled from the posterior distribution
@@
 the Jukes-Cantor model for DNA sequence evolution 
-using the mutation rate estimate (Table S\ref{tab:three}).
+using the mutation rate estimate (supplementary table S\ref{tab:three}).
 This consisted of 100 replicates of data sets.  
-We follow the same computation of the fraction as in the simulation 2 with 
-the numbers of recombinant edges estimated from the real data
-as numerators.
+We follow the same analysis as in the simulation 2 
+(supplementary fig.\ S\ref{fig:sim3}).
 
@@
 
+\textbf{Simulation 4 - Simulation of additive horizontal gene transfers.}
+We also performed a simulation analysis to understand the power of 
+the parsimony-based pipeline to recover events such as gene duplication 
+and additive transfer.  The simulations were also used as a null model
+for studying variations in the rate and biases in the direction of
+additive transfers.
+
+We implemented a new program that simulates all four events:
+duplication, loss, additive transfer, and replacing transfer.  The
+generative process begins with a single gene lineage located at the
+root of the species tree.  The gene lineage then grows down a species
+tree branch until one of the following events occur.
+
+\begin{itemize}
+\item If a species tree node is encountered (the root node is the
+  first encountered), the gene lineage undergoes {\em speciation} and
+  bifurcates into two lineages, each one evolving independently down a
+  child branch of the species tree node.  
+\item Using a Poisson process at constant rate $D$, a gene lineage
+  will undergo {\em duplication} and bifurcate into two lineages, both
+  growing independently within the same species branch.  
+\item At a constant rate $L$, a gene lineage can be {\em lost} and
+  terminate.  
+\item At a constant rate $T$, a gene lineage can undergo {\em additive
+  transfer}, bifurcating into two lineages, one lineage within
+  the current species branch and the second lineage within one
+  of the other contemporary species branches chosen uniformly.  Both
+  gene lineages continue to evolve independently of one another.  
+\item At a constant rate $R$, a gene lineage $a$ can undergo {\em replacing
+  transfer}.  From the other contemporary species branches, a second
+  gene lineage $b$ (if it exists) is chosen to be replaced.  The gene
+  lineage $b$ is terminated and lineage $a$ is bifurcated, one lineage $a_1$
+  remaining in the current species and the other $a_2$ transferring to the
+  species branch of lineage $b$.  Both lineages $a_1$ and $a_2$ continue
+  to evolve independently.
+\item Lastly, if a gene lineage encounters a species tree leaf, the
+  gene lineage terminats and is denoted an {\em extant gene copy}.
+\end{itemize}
+
+As a final step, any gene lineage that terminates before the species tree
+leaves, is declared extinct and is pruned from the gene tree.
+
+Given a gene tree generated by this process, we simulate sequence
+evolution using a continuous-time Markov process (e.g. HKY)
+parameterized to match the base frequency and transition/transversion
+ratio observed in the real data.  This results in a gapless simulated
+gene alignment.
+
+In our simulation analysis, we used our simulation program within a
+larger analysis pipeline that followed the same procedure we have used
+to analyzed the real dataset, thus giving an accurate view of the
+power of the entire pipeline.  In our simulation pipeline, we
+simulated 10,000 gene alignments.  For each alignment we reconstructed 
+gene trees using RAxML, reconciled them using Mowgli, and classified
+additive and replacing gene transfers using the procedure described in the
+main text.
+
+To create a genome-wide sample of gene trees we randomly chose 2314
+gene trees (the same number of real families) from the 10,000
+reconstructed.  To study the distribution of this tree set, we
+resampled with replacement 1000 times (This resampling strategy was
+chosen for computational efficiency).  This gave us the following
+samples: (1) a distribution of the ``true'' simulated gene trees along
+with their event counts, (2) a distribution of the reconstructed
+simulated gene trees and event counts, and (3) the original set of
+gene trees reconstructed from real data.  Differences between datasets
+(1) and (2) indicate the power of our pipeline to accurately infer
+each of the event types, and differences between datasets (2) and (3)
+indicate unusual deviations of the real data from the null model
+(i.e. variations in event rates or biases in transfer direction).
+
+\subsection*{Simulation-based support for biased directionality
+in replacing gene transfer between SPY and SDE}
+
+We conducted a series of experiments on the simulated data to check the reliability
+of estimates obtained by the model-based approach. We found that the population
+parameters obtained in the first phase were reliable, with the exception that
+the recombinant tract length was over-estimated as this was observed
+by \citealp{Didelot2010} (Simulation 1). Our experiments also showed that the recombination
+intensity used in our functional category association tests had fairly good
+correlation between the true value and the simulated ones (P-value of 0.79).  
+% (Figure S\ref{fig:ri1}). 
+One thing we specifically wished to test in these experiments
+was whether the model-based approach reliably detected deviations of recombinant
+edge counts from their expectation, as observed in the heatmap of Figure
+\ref{fig:Heatmap-of-transfers}. Our major concern was that these deviations
+were not directly accounted for by the probabilistic model of ClonalOrigin, but
+rather indirectly observed by summary of the posterior sample of recombinant
+trees.
+
+We simulated the model to generate a hundred simulated data sets using a
+collection of recombinant trees sampled from the prior distribution defined by
+the species tree and three population parameters, as estimated from our genomic
+data (Simulation 2).  We then repeated the inference procedure of the second stage of
+ClonalOrigin on each data set and, as in the data analysis, and counted
+for each donor-recipient branch pair the average number of recombinant edges.
+By normalizing the counts by the prior expected count we obtained intervals of
+base 2 logarithms of the ratio of the observed number of recombinant edges to
+the prior expectionation (supplementary fig.\ S\ref{fig:sim2}). 
+The comparison showed a good
+concordance of the estimated counts and the prior expected counts for 
+the topology-altering edge types.
+For each edge type, we also plotted in supplementary figure S\ref{fig:sim3} the
+log-ratio of the count estimated for it in our data analysis and the expected
+count; these are the values making up
+the heatmap of Figure \ref{fig:Heatmap-of-transfers}. Comparing the counts
+estimated in the data analysis and the counts estimated from the simulated data
+allowed to evaluate the significance of the divergence from expectation observed
+in the heatmap.  Recall the heatmap indicated an excess of recombinant edges
+between the SPY and SDE clades in both directions. Our simulations indicated
+that the excess observed in the direction SDE to SPY was much less significant,
+as the counts estimated from real data often fell very close to the
+corresponding counts estimated from simulated data. The excess observed in the
+direction SPY to SDE appeared not only to be larger, but also more significant.
+This further supported our claim for the biased directionality in replacing gene
+transfer between these two species.
+
+We used simulated data (Simulation 3) to test not only the ability of the ClonalOrigin
+sample to \textit{detect} deviations from expected counts of recombinant edges, but
+also its ability to reliably \textit{quantify} this deviation.  For this purpose we
+generated a hundred simulated data sets using a collection of recombinant trees
+sampled from the posterior sample of ClonalOrigin on the genomic data.  The
+difference between the two simulated batches was that in the first one, the
+recombinant trees used represent the prior distribution, and in the second one,
+they represented the posterior distribution, which we knew deviated from the
+prior. Again, we recorded the recombinant edge counts obtained by ClonalOrigin
+and normalized these counts by the prior expected counts (intervals with closed
+circles in supplementary figure S\ref{fig:sim3}). We expected these counts to be
+in better agreement with
+the counts observed in the data analysis. For recombinant edges that showed a
+defficiency in our data analysis (e.g., from the SPY clade to SDD), the counts
+appeared to fit well with the count estimated from genomic data. However, for
+recombinant edges that showed an excess in our data analysis (e.g., from the SPY
+clade to the SDE clade), the counts appeared to be significantly influenced by the
+prior. We noted that the influence of the prior appeared to be even stronger for
+topology-preserving recombinant edges. All things considered, we found that defficiency of
+a certain recombinant edge type was reliably estimated whereas excesses were 
+typically underestimated. Therefore, we concluded that the actual extent to which
+replacing transfers between SPY and SDE \textit{preferred} the direction from SPY to
+SDE was possibly somewhat underestimated in our analysis.
+
 \end{document}
-
-\subsection*{Recombination drived more substitutions than did point mutation}
-
-Five genomes except \textit{S.\ equi} ssp.\textit{\ equi} strain 4047 were
-chosen (Table S\ref{tab:genome}) to study recombining gene transfer using the
-model-based approach.  A genome alignment of the five genomes showed two
-distinct genome structures of SPY and SDE (Figure S\ref{fig:mauve}). The SDE and
-SDD genomes were larger in length than those of SPY. Although the graphical view
-of the genome alignment did not clearly show whether SDD was different from
-either of SDE1 and SDE2 in the genome organization, a spcies tree inferred from
-the genome alignment showed that the external SDD branch was attached to the
-internal clade branch of SDE1 and SDE2 (Figure S\ref{fig:tree5}). The
-distribution of the alignment block lengths was skewed to the right with mean of
-4188 base pairs long (Figure S\ref{fig:blocksize}), which supported that the
-model assumption of geometric block length distribution might be plausible in
-\texttt{ClonalOrigin} method (see section \textbf{Materials and Methods} and
-\citet{Didelot2010}).  The relative strength ($r/m$) of recombination compared
-to mutations indicated that recombination-driven substitutions affected more
-sites of the core genomes than did mutation-driven substitutions (Table
-S\ref{tab:clonalframe}). This could imply that in the \textit{Streptococcus}
-species bacterial recombination process appeared to accelerate molecular
-evolution of more sites than did point mutations process.
-
-\subsection*{Recombination hotspots were observed in genomes}
-
-Population mutation rate did not appear to change much across the genome of SDE1 
-(Figure S\ref{fig:scatter3}).  We found a few spots of genomic regions where the
-mutation rates were larger than twice the global estimate of mutation rate.  The
-recombinant tract length changed in a way that we could not note any particular
-positions with notable tract length values compared with other positions along
-the genome.  As recombinant tract lengths were smaller, recombination rates per
-site seemed larger (Figure S\ref{fig:rhologdelta}).  The recombination rate
-appeared to change more dynamically. Especially, a genomic region of SDE1 near
-200 kilobase pairs from the origin of replication contained alignment blocks with
-bell-shaped recombination rates.
-
-\subsection*{SPY and SDE were more closely related in about tenth of the core genome}
-
-There are 105 possible rooted labeled topologies relating five taxa.  We expected the
-dominant topology across sampled genealogies to be concordant with that of the
-species tree.  Indeed, two thirds of the local gene tree topologies (see section
-\textbf{Materials and Methods}) along the alignment blocks were the same as the
-species tree, which allowed us to be more confident of the species tree inferred
-using the alignment blocks (Table S\ref{tab:Gene-tree-topologies}).  Other
-frequently sampled topologies were expected to be potentially informative about
-common recombination events.  The second most frequent topology (9.3\%) was one
-where SDD was the outgroup, and SPY and SDE were sister clades. Therefore, SPY
-and SDE appeared to be more closely related with each other than each to SDD's
-genome in about tenth of the core genome. The other topologies had less than 4\%
-of proportions.
-
-\subsection*{Unbalaned recombining gene flow from SPY to SDE in genome-wide scale}
-
-The heat map of Figure \ref{fig:Heatmap-of-recombination} (see Tables
-S\ref{tab:heatmap} and S\ref{tab:obsheatmap} for the heat map values and the
-observed number of recombinant edges) shows the logarithm base 2 fractions of the
-\textit{a posteriori} number of recombinant edges for a pair of donor and
-recipient species tree branches and their corresponding expected values given
-the species tree (Figure S\ref{fig:tree5}) and the three population parameters
-(Table S\ref{tab:three}).  Because recombinant edges from any species tree
-branch to the root branch are impossible (see Figure S\ref{fig:clonalorigin}),
-the column for the root as recipient was colored black, indicating invalid
-recombinant edges in Figure \ref{fig:Heatmap-of-recombination}.  
-Similarly, all of the cells for which a recipient branch was
-older than the donor, and the donor and recipient branches did not share any contemporary time were
-colored black. The diagonal cells were colored gray, indicating no significant
-number of recombinant edges, because the observed and expected numbers of
-recombinant edges were similar. As recombinant edges that start from and end in
-the same species tree branch would not change the species tree topology, which
-might not
-change the likelihood much, the cells would be colored gray.  Because SDD is a
-strictly veterinary pathogenic bacteria, we did not expect that SDD received
-much gene transfer from either SPY1 or SPY2: the logarithm base 2 fraction from SPY1 to
-SDD was -2.20, and that from SPY2 to SDD was -1.97.  On the contrary, the
-recombing gene transfer from SDE1 to SDD was 0.756, and that from SDE2 to SDD
-was 0.698. This relatively higher recombining gene transfer from two SDE
-individuals to SDD compared with that from two SPY individuals to SDD reflected
-the fact
-that the SDD was more closely related to SDE than to SPY in most of the genomes,
-and the genomic organization between SDD and SDE1/SDE2 were much similar
-(Figure S\ref{fig:mauve}).  However, the gene transfer from SDD to either SDE1
-or SDE2 was less frequent than the case of reverse direction; SDD effectively
-served as a recipient rather than a donor. It was unclear why SDD received
-more genes from SDE1 or SDE2 than did SDE1 or SDE2 from SDD.
-
-Because it was expected that the recombining gene transfer within a clade would
-be more frequent than that between clades, not surprisingly the recombining gene
-transfer logarithm base 2 fractions between SDE1 and SDE2 (1.17 and 1.20) were about
-twice as large as those from SDE1 and SDE2 to SPY1 and SPY2 (0.67, 0.66, 0.60,
-and 0.39).  The intensities of recombining gene transfer from external branches
-of SDE1 and SDE2 to external branches of SPY1 and SPY2 was similar to those from
-SDE1 or SDE2 to SDD (0.76 and 0.70).  Therefore, either SPY1 or SPY2, and SDD as
-recipient appeared to be no different from the view point of SDE1 or SDE2 as
-donor since the similar values of recombining gene transfer intensity from SDE1
-or SDE2 to one of the other three external branches. Surprisingly, the recombining
-gene transfer logarithm base 2 fractions between SPY1 and SPY2 (1.06 and 1.07) was
-comparable with, yet even less than those from SPY1 and SPY2 to SDE1 and SDE2
-(1.25, 1.25, 1.17, and 1.29).  The recombining gene transfer from a clade SPY1 
-or SPY2 to a different clade SDE1 or SDE2 seemed to be effectively no different
-from that within the clade of SPY1 and SPY2.  The two external branches of SDE
-received genes from either SPY1 or SPY2 via recombining gene transfer as if SDE1
-and SDE2 were in the same clade as SPY1 and SPY2.  This was more surprising
-because the bacterial recombination between SPY and SDE seemed to happen more
-often than that between SDE and SDD even though the genome configurations of SDE
-and SDD were more similar than each to SPY.  However, this result could be
-concordant with the fact that SPY and SDE shared the same ecological niche
-despite their dissimilar genomic organization.  The gene transfer logarithm base 2
-fraction from SPY1 and SPY2 to SDE branch was also larger than any type of
-recombining gene transfer: 1.88 and 1.94, respectively.  Additionally we found
-58 genes for which a recombinant edge of certain types were sampled with
-probability at least 0.9 by focusing on specific types of recombinant edges
-(Table S\ref{tab:genes-transfer}).  Many of the genes in the list were
-transferred from SPY to SDE rather than the reverse direction overall.  All
-things considered, this epitomized that the recombining gene transfer was biased
-toward the net direction from SPY to SDE.
-
-We tried in vain to quantify differential horizontal gene transfer between SPY
-and SDE with the parsimony-based approach (Figure \ref{fig:hgt-heatmap}).  Among
-the clade-crossing transfers, we did not see any significant enrichment for a
-specific direction, as opposed to the result for recombining transfers from the
-model-based analysis.  SDE1 and SDE2 shared the most events (26 and 34) compared
-to other sample pairs.  SDD appeared to acquire many genes from the ancestral
-SDE branch.  Transfer events between the SPY and SDE clades were inferred in
-smaller numbers ($<6$) compared to those within each clade. We were puzzled by
-this finding because virulence genes were found to be associated with horizontal
-gene transfer, and yet we could not show net directionality of horizontal gene
-transfer between SPY and SDE.  However, we inferred relatively more recombining
-gene transfer events from the SPY branch to SDE branch even in the
-parsimony-based approach (Figure S\ref{fig:mowgli-recomb-heatmap}).
-
-\subsection*{Reduction of genes and divergence of SPY and SDE species}
-
-The parsimony-based approach estimated 304 duplications, 944 recombining gene
-transfers, and 294 horizontal gene transfers, comprising 23.7\% of total
-transfer events (Figure \ref{fig:Gene-duplication-loss}).  Among the 2314 gene
-families the largest family consisted of 59 genes, while nearly half of gene
-families (1066) had exactly one gene per species, or six genes total (Figure
-S\ref{fig:famsizes}).  Numbers of genes at the five ancestors were inferred to
-be smaller than those at the extant species.  SDE lineages from the root
-appeared to have increased in the number of genes while the common ancestor of
-SPY1 and SPY2 decreased in the number of genes.
-
-The four external lineages of SDE1, SDE2, SDD, and SPY1 had almost equal numbers
-of gene gain and loss events while the four lineages of SP, SD, SPY2, and SEE
-had larger numbers of gene gain events than those of gene loss events. In the
-two lineages of SP and SEE we might have inferred gene gain and loss events less
-confidently than events in other lineages because of lack of outgroup lineages
-to compare with the two lineages. The number of gene gain events in SPY2 lineage
-seemed to be large compared with those of gene gain events in the other
-lineages. On the contrary, two internal branches of SPY and SDE had smaller
-numbers of gene gain events than those of gene loss events.  This might indicate
-that SPY and SDE lineages had reduced number of genes to specialize in their
-specific niches after each split from their corresponding direct ancestors.
-
-Duplications in extant lineages were more frequent than those in the other
-lineages.  Although the number of duplication events along SPY2 branch seemed
-smaller than those of the other extant lineages, duplication events appeared to
-be most often species-specific.  Within clades we found that duplications made
-up only a small fraction of the events  (10.2\%), while horizontal transfers and
-losses were significantly more frequent (41.7\% and 48.0\%, respectively).  The
-extant species of SDD and two SDE's were the most expanded, both in terms of
-duplications and in-bound transfers. This was apparently related with the larger
-genome sizes of the two species than that of SPY.  
-The numbers of recombining gene transfer events tended to be larger than those
-of horizontal gene transfer events except SD lineage.  Numbers of recombining
-gene transfer in the extant lineages might have been larger than those in the
-internal branches because the number of recombining gene transfer events must
-had been affected by the availability of more closely related individual
-genomes.
-
-\subsection*{Virulence genes via recombining gene transfer}
-
-About 10\% of the genes comprised of virulence facters, about half of which were
-found in the core genome of the five genome alignment \citep{Suzuki2011}.  Using
-customized UCSC genome browser tracks available at
-\url{http://strep-genome.bscb.cornell.edu} we located virulence genes with
-notable posterior recombination probability.  These genes were by no means a
-comprehensive list of virulence factors that might have been transferred via
-homologous recombination.  We wished to showcase virulence genes using the
-result of our study of gene transfer.
-
-Stressful host conditions for pathogenic bacteria can include heat shock,
-starvation, and iron depletion, etc.  Iron uptake of bacteria was known to be
-essential for survival of bacteria under the hostile host condition. We found
-that parts of putative calcium transporter (MgtA) in SDE1 could have been
-transferred from SPY1 (SpyM3\_0440).  An extracellular protein, putative
-dipeptidase (SpyM3\_0465) that could be important for bacteria to scavenge
-dipeptides from their host, was inferred to be transferred from SPY to SDE: the
-whole gene was transferred, which was a case for the finding that virulence
-genes tended to be transferred as a whole unit \citep{Chan2009}.  Another
-example of whole gene transfer was superoxide dismutase that was reported to be
-associated with virulence of \textit{Streptococcus agalactiae}
-\citep{Poyart2001}: the whole gene including its 5' UTR, (SpyM3\_1071), was
-affected by recombining gene transfer from SPY to SDE.  Some of the genes were
-only affected in their parts.  The middle parts of an adhesins gene, putative
-internalin A precursor (SpyM3\_1035), was found to be affected by recombining
-gene transfer from SPY to SDE.  The middle part of a chaperonin GroEL, putative
-heat shock protein (SpyM3\_1765), was inferred to transfer from SPY to SDE1:
-increase of GroEL expression was known to allow for the bacteria to survive the
-hostile host environment of acid shock.  Parts of the gene (SpyM3\_1749),
-pyruvate formate lyase that was recently reported to have virulence
-\citep{Yesilkaya2009}, seemed to be affected by recombining gene transfer.
-Under starvation condition bacteria must use more effective energy metabolism: a
-putative glucose kinase, (SpyM3\_1180), were inferred to be transferred from SPY
-to SDE to adapt SDE's energy metabolism to survive such harsh condition.  We
-also found that an MLST locus, called DNA mismatch repair protein (MutS),
-appeared to be affected by recombining gene transfer.
-
-\subsection*{Association of virulence genes and horizontal gene transfer}
-
-Horizontal gene transfers were found to be significantly associated with
-virulence genes (P-value of $<1.33 \times 10^{-11}$) whereas recombining gene
-transfers were not enriched for virulence genes (P-value of $>0.90$) using the
-parsimony-based approach.  This was also consistent with the fact that we could
-not find the association of recombining gene transfer with virulence genes in
-the model-based approach.  This supported more evidence that non-recombining
-gene transfer served as a more effective means by which SDE or SPY acquired
-virulence genes, compared to recombining gene transfer.  Families with high
-numbers of duplications were also significantly enriched for virulence genes
-(P-value of $<7.13 \times 10^{-3}$) whereas families with high losses showed no
-enrichment (P-value of $>0.54$).
-
-Roughly, half of the virulence genes were in the pan genome as mentioned above.
-We did not attempt to comprehensively list virulence genes that were inferred to
-be transferred horizontally or non-recombining fashion. We rather sort virulence
-genes by the number of horizontal gene transfer events to list the first several
-genes with high ranks.  Streptococcal C5a peptidase (\textit{scpB}, SDEG\_0933)
-was inferred to have six horizontal gene transfer.  Putative transposase were
-found to have a relatively large number of horizontal gene transfer events
-(SpyM3\_1016, SpyM3\_1053, SpyM3\_1054). Other genes with relatively large
-number of horizontal gene transfer events included putative response regulator
-(SpyM3\_0174), putative maltose operon transcriptional repressor (SpyM3\_0982),
-putative histidine kinase (SpyM3\_0171), and putative cell wall hydrolase
-(SpyM3\_0731, SpyM3\_0922, SpyM3\_1208).  Virulence genes encoding putative
-nucleases with a secretion signal peptide were also found to have had
-experienced horizontal gene transfer: genome-encoded streptodornase
-(SDEG\_0541), and prophage-associated deoxyribonuclease (SDEG\_1103).  Protein
-F2 like fibronectin-binding protein (SpyM3\_0104), an adhesin and invasin, was
-inferred to horizontally transferred.
-
-\subsection*{Model-based approach or parsimony-based one for recombining gene
-transfer study}
-
-A full model-based approach would be desirable in analyzing data for testing
-relevant hypotheses more systematically. Model complexity might not allow easy
-access to the implementation of a model. Our study involved a model-based
-approach and a parsimony-based one for understanding evolution in the core
-genomes of several bacteria individuals. Because of inability of the model-based
-method for studying the pan genomes we had to resort to the parsimony-based
-method. However, the parsimony-based method was also applied to the core
-genomes. We wished to compare the two methods in their results of the core
-genome analyses.  Because of two independent analyses on the same data set, this
-would allow us to be more confident of the results if the two analyses performed
-similarly. Generally, phylogenetic methods of detecting gene transfers often take
-advantage of discrepancy between a reference species tree and gene trees. The
-two approaches that we employed in this study were no exception. We found a few
-similar results in both methods; firstly, functional categories that were
-associated with recombining gene transfer were similar (Tables
-S\ref{tab:functional-all} and S\ref{tab:go-events-recombining}).  Secondly,
-either of them did not show association of virulence genes with recombining gene
-transfer.  To more directly compare the methods we used the recombination
-intensity and the number of recombining gene transfer events for each gene.  We
-found a weak positive correlation of 0.39 with a significant P-value
-of $2.2\times10^{-16}$ for testing a null hypothesis of the correlation being
-zero.  (Figure S\ref{fig:cmpcomowgli}).  Interestingly, the correlation (0.03
-with P-value of 0.18) disappeared in a comparison between recombination
-intensity and number of horizontal gene transfer events (Figure
-S\ref{fig:cmpcomowglihgt}).
-
-\subsection*{UCSC genome browser tracks for bacterial gene transfer study}
-
-Often computational and statistical analyses on genomes could benefit other
-researchers when those resources were provided: e.g., UCSC Genome Browser, and
-Ensembl Genome Browser. We wished to provide research community especially in
-SPY and SDE bacteria with some of the results from our analysis of gene
-transfer. We used the UCSC genome browser for the purpose.  The resource shown
-in Figure S\ref{fig:ucsc} for gene transfer is available at
-\url{http://strep-genome.bscb.cornell.edu}.  For each of the five individual
-genomes as a reference genome several annotation tracks were displayed:
-alignment blocks, recombination rate per site for each block, local gene tree
-topologies, recombination intensity, locations of virulence genes, posterior
-probability of recombining gene transfer. The genome track of probability of
-recombining gene transfer allowed us to easily browse the bacterial genome to
-locate regions that we might be interested in. 
-% To Matt: what could we show in the pan-genome? or from Mowgli's analysis?
-
-\subsection*{Units of recombining gene transfer}
-
-The process of gene transfer may or may not move whole genes from one species to
-another \citep{Chan2009}. We observed that units of recombining gene
-transfer could be multiple genes (Table S\ref{tab:genes-transfer}). In some
-cases, parts of a gene were inferred to have been transferred, for instance,
-transfer from SPY1 to SDE1 was inferred for 29\% of the aspartyl-tRNA synthetase
-gene. Recombinant tracts were often shown to span two genes.  In other cases,
-the inferred transfer units spaned a set of as much as eight adjacent genes:
-loci from SpyM3\_1518 to SpyM3\_1525.
-
-\subsection*{Simulations}
-
-We conducted experiments on simulation data to check the accuracy of estimates
-obtained by the model-based approach in a scenario similar to the one from the
-real data set.  Population mutation and recombination parameters were accurately
-estimated (Table S\ref{tab:sim-three}).  Yet, the recombinant tract length was
-over-estimated; recombinant tract lengths were reported to be difficult to
-estimate from relatively short alignment blocks \citep{Didelot2010}.  Numbers of
-recombinant edges from the simulated recombinant trees of Simulation 2 (see
-\textbf{Supplementary Information} for the description of Simulation 2) were
-compared with the corresponding numbers of recombinant edges that were estimated
-from the simulated data set from Simulation 2 (see the solid-line intervals with
-open circles in Figure \ref{fig:h3}).  We noted that tree-topology-changing
-recombinant edges were more reliably recovered compared with
-tree-topology-not-changing ones (Figure \ref{fig:h3}). All of the
-tree-topology-not-changing recombinant edge types were over-estimated (Figure
-\ref{fig:h3}B). The heat map of Figure \ref{fig:Heatmap-of-recombination} did
-not show how confident of the values we could be. Comparison of numbers of
-recombinant edges estimated from the real data set and the corresponding numbers
-estimated from Simulation 2 showed the interval estimates for the heat map
-values.  Dotted-line intervals from the red and blue colored cells in the heat
-map were above and below the reference of the horizontal line at the zero on the
-Y-axis, respectively (Figure \ref{fig:h3}).
-
-For each recombinant edge type to assess the effect of posterior recombinant
-trees samples with respect to that of prior recombinant tree samples. The ten
-\textit{a priori} recombinant trees from Simulation 2 would have balanced
-recombining gene transfer events between SPY and SDE with respect to the prior
-expected number of recombinant edges. The posterior recombinant trees used in
-Simulation 3 would have unbalanced recombining gene transfer between the two
-species branches as we observed in the real data analysis. For significantly
-strong gene flows such as ones from SPY to SDE a solid-line interval with a
-closed circle would be different from its count part of the dotted-line
-interval.  Figure \ref{fig:h3} showed that the recombining gene flow from SPY to
-SDE was stronger enough to be nearer to the horizontal line at the zero on the
-Y-axis. However, the recombining gene transfer from SDE to SPY appeared to be
-too weak.  This comparison supported the biased directionality of recombining
-gene transfer from SPY to SDE.
-
-Recombination intensities that were measured on genes also showed rough
-correlation (Pearson's correlation coefficient of 0.63) of the estimated and
-true values (Figure S\ref{fig:ri1}).  Estimated values appeared to be driven
-somewhat by the prior implied by the population parameter setting, explaining
-why lower values were over-estimated and higher values were under-estimated.
-Nonetheless, the correlation indicated that the measure of recombination
-intensity used in our functional gene category served as an adequate indicator
-of actual intensity of recombination events along genes.

### 2011-12-08 R/obsiter.R

@@ -80,6 +80,7 @@ internal.branch.aswell <- function (spy.name, sde.name)
   cat("1001 values less than the ratio for the prior.\n")
   cat("The p-value is the fraction of them.\n")
   cat("Quantile (0.025, 0.5 or median, 0.975) is", quantile(x.v,  probs = c(0.025, 0.5, 0.975)), "\n")
+  cat("Mean is", mean(x.v), "\n")
   cat("The exp value is", x.exp.v, "\n")
   cat("The P value is", p.val, "\n\n")
   # x.v

### 2011-12-08 sh/output-backup.sh

@@ -1,7 +1,10 @@
 DIR1=output
 DIR2=output-backup
 
+cp -pr $DIR1/go $DIR2
+exit
 
+cp -p $DIR1/cornellf/3/run-analysis/* $DIR2/cornellf/3/run-analysis
 
 for i in matt virulence; do
   cp -pr $DIR1/$i $DIR2/
@@ -27,47 +30,9 @@ done
 for i in {1..4}; do
   cp -pr $DIR1/cornellf/3/run-clonalorigin/output2/$i $DIR2/cornellf/3/run-clonalorigin/output
 done
-
-exit
-
-
-
-
-
-
-
-
-
-
-
-for i in matt virulence; do
-  cp -pr $DIR1/$i $DIR2/
-done
-
-cp -p $DIR1/cornellf/3/data/core_alignment.maf.* $DIR2/cornellf/3/data
-cp -p $DIR1/cornellf/3/data/core_alignment.xmfa $DIR2/cornellf/3/data
-cp -p $DIR1/cornellf/3/data/core_alignment.xmfa.* $DIR2/cornellf/3/data
-
-cp -pr $DIR1/cornellf/3/run-mauve/output $DIR2/cornellf/3/run-mauve
-cp -pr $DIR1/cornellf/3/run-clonalframe/output $DIR2/cornellf/3/run-clonalframe
-cp -pr $DIR1/cornellf/3/run-clonalorigin/summary $DIR2/cornellf/3/run-clonalorigin
-cp -p $DIR1/cornellf/3/run-clonalorigin/clonaltree.nwk $DIR2/cornellf/3/run-clonalorigin
-cp -pr $DIR1/s15/run-analysis $DIR2/s15
-cp -pr $DIR1/s16/run-analysis $DIR2/s16
-cp -pr $DIR1/s17/run-analysis $DIR2/s17
-cp -pr $DIR1/s18/run-analysis $DIR2/s18
 
 for i in {1..4}; do
   rm -rf $DIR2/cornellf/3/run-analysis/recombprobwig-$i
   cp -pr $DIR1/cornellf/3/run-analysis/recombprobwig-$i $DIR2/cornellf/3/run-analysis
 done
-
-for i in {1..2}; do
-  cp -pr $DIR1/cornellf/3/run-clonalorigin/output/$i $DIR2/cornellf/3/run-clonalorigin/output
-done
-
-for i in {1..4}; do
[…]

### 2011-12-13 doc/manuscript.tex
A supplementary is added.

@@
 % TODO:
-% Figure S\label{fig:calling-transfers} was not referred in the text.(Delete this?) 
-% Figure S\label{fig:mowgli-sim} was not referred in the text. (Write comments)
-% Table S\label{tab:sim-three} was not referred in the text. (Delete this?)
-% 
-%
-% * Finalize section: Distribution of transfer events along the genome
-% * Finalize section: Supp. Materials.
-% Supp. Methods of Matt's modifications of Mowgli's method
+% 1. Not yet referenced figure - \label{fig:mowgli-sim}
+% 2. Not yet referenced figure - \label{fig:sim2}
+% 3. Not yet referenced figure - \label{fig:sim3}
 % 
@@
 \usepackage{times}
+% \usepackage[T1]{fontenc}
+% \usepackage{ae,aecompl}
 
@@
 
-% FIXME: Remove this package when submitting the manuscript.
-\usepackage{url}
-
 \usepackage{natbib}
 \usepackage{../../latex/sty/genres}
+
+\usepackage{xr}
+\externaldocument{supplementary}
 
@@
 \textheight 9in
-%\topmargin 0.0cm
-%\oddsidemargin 0.5cm
-%\evensidemargin 0.5cm
-%\textwidth 16cm 
-%\textheight 21cm
 
@@
 
-% \@ifundefined{showcaptionsetup}{}{%
-% \PassOptionsToPackage{caption=false}{subfig}}
-% \usepackage{subfig}
-
 \makeatother
-
-% Leave date blank
-\date{}
 
@@
 bacteria is now well documented, yet few studies have differentiated
-between the mechanisms of homologous recombination and additive integration.
-%as sources of HGT.  
-These processes leave distinct phylogenetic signatures, because homologous
-recombination causes a gene in one lineage to be replaced by a homolog from
-another lineage (``replacing HGT''), while additive integration causes
-addition without replacement (``additive HGT'').  Here we make use of these
-signatures in a genome-wide investigation of HGT in the important human
-pathogen {\em Streptococcus pyogenes} (SPY) and its close relatives {\em
-  S.\ dysgalactiae} subspecies {\em equisimilis} (SDE) and {\em S.\
-  dysgalactiae} subspecies {\em dysgalactiae} (SDD).  Using recently
-developed statistical models and computational methods, we find evidence
-for abundant gene flow of both kinds within the SPY and SDE clades, and of
-generally reduced levels of exchange between SPY/SDE and SDD.  In addition,
-our analysis strongly supports a previously reported, but unconfirmed,
-finding of a pronounced asymmetry in SPY-SDE gene flow, favoring the
-SPY-to-SDE direction, which may be associated with increasing virulence of
-pathogenic SDE.  We find that this enrichment is driven by replacing
-transfers, and likely reflects frequent homologous recombination between
-co-occurring SPY and SDE cells in human hosts.  We find that virulence
-genes are correlated with transfer events, but this correlation is driven
-by additive, not replacement, HGTs.  The genes affected by additive HGTs
-are enriched for transposition-related functions, while replacement HGTs
-appear to influence a more diverse set of genes.  These findings shed new
+between the mechanisms of homologous recombination and additive
+integration.  These processes leave distinct phylogenetic signatures,
+because homologous recombination causes a gene in one lineage to be
+replaced by a homolog from another lineage (``replacing HGT''), while
+additive integration causes addition without replacement (``additive
+HGT'').  Here we make use of these signatures in a genome-wide
+investigation of HGT in the important human pathogen {\em Streptococcus
+  pyogenes} (SPY) and its close relatives {\em S.\ dysgalactiae} subspecies
+{\em equisimilis} (SDE) and {\em S.\ dysgalactiae} subspecies {\em
+  dysgalactiae} (SDD).  Using recently developed statistical models and
+computational methods, we find evidence for abundant gene flow of both
+kinds within the SPY and SDE clades, and of reduced levels of exchange
+between SPY and SDD.  In addition, our analysis strongly supports a
+previously reported, but unconfirmed, finding of a pronounced asymmetry in
+SPY-SDE gene flow, favoring the SPY-to-SDE direction, which may be
+associated with increasing virulence of pathogenic SDE.  We find much
+stronger evidence for SPY-SDE gene flow among replacing rather than
+additive transfers, suggesting a primary influence from homologous
+recombination between co-occurring SPY and SDE cells in human hosts.
+Virulence genes are correlated with transfer events, but this correlation
+is driven by additive, not replacing, HGTs.  The genes affected by additive
+HGTs are enriched for functions having to do with transposition,
+recombination, and DNA integration, while replacing HGTs appear to
+influence a more diverse set of genes.  Additive transfers are also
+associated with evidence of positive selection.  These findings shed new
 light on the manner in which HGT has shaped pathogenic bacterial genomes
@@
 to a common ancestor for all species \citep{Zhaxybayeva2011}.  This would
-mean that no single phylogenetic marker is a reliable indicator of the tree
-of life.
+mean that no phylogenetic marker can be used to reconstruct a
+universal tree of life.
 
@@
 for SPY, SDE, and the related species {\em S.\ dysgalactiae} ssp.\ {\em
-  dysgalactiae} (SDD), and new statistical and parsimony-based methods for
+  dysgalactiae} (SDD), as well as new statistical and parsimony-based
+methods for 
 analysis.  The combined use of these methods allows us to examine both
@@
 find evidence for abundant gene flow within SPY, within SDE, and between
-SPY and SDE, and for greatly reduced exchange between SPY/SDE and SDD.  In
+SPY and SDE, and for greatly reduced exchange between SPY and SDD.  In
 addition, our genome-wide analysis supports Kalia et al.'s
@@
 preference for the SPY-to-SDE direction in SPY-SDE gene flow.
-Interestingly, this property appears to hold for replacing but not for
-additive transfers.  More generally, we find evidence for an association
-between virulence genes and additive, but not replacing, HGTs.  Our results
+Interestingly, this property is much more evident for replacing than for
+additive transfers.  We also examine correlations of gene transfer
+events with functional categories and positive selection.
+%find evidence for an association
+%between virulence genes and additive, but not replacing, HGTs.  
+Our results
 have been made available as tracks in our recently released {\em
@@
 SDE2]), and one of {\em S.\ dysgalactiae} ssp.\ {\em dysgalactiae}
-(AEGO00000000 [\citealp{Suzuki2011}; SDD]).  In addition, we used 
+(CM001076 [\citealp{Suzuki2011}; SDD]).  In addition, we used 
 \textit{S.\ equi} ssp.\textit{\ equi} strain 4047 (NC\_012471
 [\citealp{Holden2009}; SEE]) as an outgroup in our parsimony-based
-analysis (supplementary table S\ref{tab:genome}).
+analysis (supplementary table \ref{tab:genome}).
 
 For the model-based analysis of replacing transfers, we obtained
-high-quality alignments of the five genomes (SPY1 \& 2, SDE1 \& 2, and
-SDD) using the 
+high-quality alignments of SPY1, SPY2, SDE1, SDE2, and SDD using the
 pipeline detailed by Didelot and colleagues
-\citep{Didelot2007,Didelot2010}.  Briefly, we first aligned the genomes
-using progressiveMauve v2.3.1 \citep{Darling2004,Darling2010} with default
-options (supplementary fig.\ S\ref{fig:mauve}), and then used stripSubsetLCBs (from the
-ClonalOrigin package; \citealp{Didelot2010}) to identify 276 blocks of
-sequence alignments that exceed a length threshold of 1500 bp.  This
-conservative threshold was selected to avoid biases from boundary effects
-in short alignment blocks, based on a preliminary analysis.  Two blocks
-containing long gaps were excluded from the analysis.
-The final set of 274
-alignment blocks was 1147632 nucleotide long.
-Of 1951 genes of SPY1 genome 1125 genes were covered by the blocks.
-Note that this alignment
-procedure effectively focused on the ``core genome,'' by discarding regions
-that were not conserved among the five species in question.  The 1125 genes
-contained 0.17\% of gaps on average.
-
-For the parsimony-based analysis, we began with 2314 gene family
-clusters representing the six genomes (including the SEE outgroup)
-from \cite{Suzuki2011}.  We aligned protein sequences corresponding to
-each family using MUSCLE \citep{Edgar2004a}, and then obtained
-nucleotide alignments by reverse translation, using the genomic
-sequences as a guide.  Next we identified likely intragenic
-recombinations using the Single Breakpoint Recombination (SBP) method
-from the HyPhy package \citep{KosakovskyPond2006}.  Based on the
-Akaike Information Criterion (AIC), SBP identified at least one
-topology-altering recombination in 47.5\% (1094) of gene
-families. These alignments were split into putatively non-recombining
-gene fragments.  The final set of alignments contained 3408 gene
-fragment families, 1794 of which contained one gene in each of the six
-species.  Families ranged as small as two genes (417 fragments) to as
-large as 59 (2 fragments) (supplementary fig.\ S\ref{fig:famsizes}).  Notably, in
-contrast to the genome-wide alignments produced by progressiveMauve,
-these alignments were representative of the ``pan genome,'' including
-regions that frequently turn over, rather than just the ``core
-genome'' \citep[e.g.,][]{Tettelin2005,Lefebure2007}.
+\citep{Didelot2007,Didelot2010}.  Briefly, we first aligned these five
+genomes using progressiveMauve v2.3.1 \citep{Darling2004,Darling2010} with
+default options (supplementary fig.\ \ref{fig:mauve}), and then used
+stripSubsetLCBs (from the ClonalOrigin package; \citealp{Didelot2010}) to
+identify 276 blocks of sequence alignments that exceed a length threshold
+of 1,500 bp.  This conservative threshold was selected to avoid biases from
+boundary effects in short alignment blocks, based on a preliminary
+analysis.  Two blocks containing long gaps were excluded from the analysis.
+The final set of 274 alignment blocks contained 1,155,016 columns with
+relatively few gaps (0.2\% of characters) and covered 1,106 of the 1,951
+genes in the SPY1 genome.  Note that this alignment procedure effectively
+focused on the ``core genome,'' by discarding regions that were not
+conserved among the five bacterial individuals in question.
+
+For the parsimony-based analysis, we began with 2,314 gene family clusters
+representing all six genomes (including SEE), as described by
+\cite{Suzuki2011}.  We aligned protein sequences corresponding to each
+family using MUSCLE \citep{Edgar2004a}, and then obtained nucleotide
+alignments by reverse translation, using the genomic sequences as a guide.
+Next we identified likely intragenic recombinations using the Single
+Breakpoint Recombination (SBP) method from the HyPhy package
+\citep{KosakovskyPond2006}.  Based on the Akaike Information Criterion
+(AIC), SBP identified at least one topology-altering recombination in
+47.5\% (1094) of gene families. These alignments were split into putative
+non-recombining gene fragments.  The final set of alignments consisted of
+3,408 gene fragment families, about half (1,794) of which contained
+single gene from each of the six species.  The remaining families ranged
+from as few as two genes (417 fragments) to as many as 59 (2 fragments)
+(supplementary fig.\ \ref{fig:famsizes}).  These alignments were
+representative of the ``pan genome,'' including regions that frequently
+turn over (``dispensable'' regions), rather than just the ``core genome''
+\citep[e.g.,][]{Tettelin2005,Lefebure2007}.
 
@@
 {\bf ClonalFrame and ClonalOrigin.}  To study replacing HGTs, we made use
-of the coalescent-based model of homologous recombination and the Markov
-chain Monte Carlo (MCMC) algorithm recently developed by
-\cite{Didelot2010}.  The method was implemented in a 
-program, called ClonalOrigin, that could produce samples from the posterior
-distribution of {\em recombinant trees} given a set of alignment blocks. A
-recombinant tree consists of a rooted tree with branch length, called a clonal
-frame, for the five bacterial genomes,
-which is estimated in a pre-processing step, augmented by a set of
+of the ClonalOrigin program recently developed by \cite{Didelot2010}.
+Given a set of alignment blocks, ClonalOrigin samples from the posterior
+distribution of {\em recombinant trees} using a Markov chain Monte Carlo
+(MCMC) algorithm.  A recombinant tree consists of a pre-estimated rooted
+phylogeny with branch lengths (a ``clonal frame''), augmented by a set of
 {\em recombinant edges}. Each recombinant edge is defined by the two points
-along the branches of clonal frame corresponding to the {\em donor} and
-{\em recipient} in the hypothesized recombination event, and by the start
-and end of the genomic interval that is affected by the event.  The donor
-must pre-date the recipient, and a delayed coalescence is possible
-(see \textbf{Supplementary Material} for details of summarizing recombinant
-trees and edges). 
+along the branches of the clonal frame that correspond to the {\em donor}
+and {\em recipient} in the hypothesized recombination event, and by the
+start and end of the affected genomic interval.  The donor can pre-date the
+recipient, to allow for a delayed coalescence, but it cannot post-date it
+(supplementary fig.\ \ref{fig:clonalorigin}).
 Another program, called ClonalFrame
@@
 basis of the recombinant trees.  ClonalFrame makes use of a similar, but
-somewhat simpler probabilistic model, which permits inference
-of the clonal frame. % and a set of model parameters.
-
-{\bf Fitting the model.}  We used
-ClonalFrame (v1.1) to estimate 
-a clonal frame for the five-way
-alignments, running the algorithm for $10^4$ burn-in iterations followed by
-another $10^4$ sampling iterations.  A series of seven additional sampling
-runs with 
-random initializations converged to an essentially identical clonal frame, % model parameters,
-indicating stable convergence.  We then used an initial run of
-ClonalOrigin (subversion r19) to estimate global parameters including
-the mutation rate, recombination rate, and recombinant tract length. We
-applied the method separately to each of the 274 blocks, in all cases using
-$10^6$ burn-in iterations, $10^7$ sampling iterations, and sub-sampling
-every $10^5$ iterations. A global estimate for each parameter was then
-obtained by taking the block-length weighted median of the block-specific
-posterior means, which down-weighed less reliable estimates obtained from
-shorter alignment blocks.  We generally followed \cite{Didelot2010} in 
-fitting the model. % , with some minor differences.
-
-{\bf Sampling recombinant trees}.  Next we ran a second stage of ClonalOrigin
-MCMC
-on all of the alignment blocks by fixing the clonal frame and global model
-parameters at their pre-estimated values, and sampled
-recombinant trees.  For each block, we ran the sampler for $10^7$ burn-in
-iterations and $10^8$ sampling iterations, sub-sampling every $10^5$
-iterations, resulting in 1001 recombinant trees per alignment block.  We
-performed a replicate of the two stages of the ClonalOrigin inference to
-ensure adequate convergence of the sampler.
+somewhat simpler, probabilistic model.
+
+{\bf Fitting the model.}  Following \cite{Didelot2010}, we used ClonalFrame
+(v1.1) to estimate a clonal frame for our five-way alignments, running the
+algorithm for $10^4$ burn-in iterations followed by another $10^4$ sampling
+iterations.  A series of seven additional sampling runs with random
+initializations converged to essentially identical estimates, indicating
+stable convergence.  We then used an initial run of ClonalOrigin
+(subversion r19) to estimate global parameters including the mutation rate,
+recombination rate, and average recombinant tract length. We applied the method
+separately to each of the 274 blocks, in all cases using $10^6$ burn-in
+iterations, $10^7$ sampling iterations, and sub-sampling every $10^5$
+iterations. A global estimate for each parameter was then obtained by
+taking the block-length weighted median of the block-specific posterior
+means, which down-weighed less reliable estimates obtained from shorter
+alignment blocks.
+
+{\bf Sampling recombinant trees}.  We ran ClonalOrigin a second time on all
+alignment blocks, this time fixing the clonal frame and global model
+parameters at their pre-estimated values, and collected samples of
+recombinant trees across the genome.  For each block, we ran the sampler
+for $10^7$ burn-in iterations and $10^8$ sampling iterations, sub-sampling
+every $10^5$ iterations, resulting in 1,001 recombinant trees per alignment
+block.  We performed a replicate of the two stages of the ClonalOrigin
+inference to ensure adequate convergence of the sampler.
 
@@
 % population parameters significantly, we used only finished blocks to estimate
-% mutation rate, recombination rate, and recombinant tract length.
+% mutation rate, recombination rate, and average tract length.
 
 {\bf Summary statistics.}  We used various summary statistics to describe
-the sampled recombinant trees (see \textbf{Supplementary Material} for details). 
-First, we considered the distribution of
-tree topologies at each site in the alignments, by computing the site-wise
-relative frequencies of all 105 possible rooted tree topologies relating five
-taxa. Second, we measured frequencies in the posterior samples
-of all possible types of recombinant edge, where an edge
-type is defined by the branches of the clonal frame associated with its
-donor and recipient.
-%also searched the posterior sample for transfer scenarios that were
-%significantly enriched, compared to the prior expectation defined by the
-%clonal frame and the global estimates of the three population
-%parameters. We followed the approach of \cite{Didelot2010} and defined
-%transfer types by the branches of the clonal frame associated with each
-%recombinant edge. 
-%For each transfer type, we recorded the average number of
-%recombinant edges of that type sampled in the posterior sample of
-%recombinant trees, 
-We compared these posterior frequencies with 
-expected values under the prior distribution assumed by the model.
-%as well as the expected number of such edges implied by
-%the prior distribution assumed by the model. 
-%These counts were then used to
-%find types of transfer that showed a high deviation in our inference from
-%their expected counts. 
-To assess significance of the ratio of posterior to prior frequencies of
-recombinant edges we performed a simulation; 
-this empirical approach also 
-had the advantage of correcting for possible biases in the sampling
-process. 
-\textbf{CHECK:}
-We simulated the model to generate 100 data sets, for which we proceeded with
-the inference of the ratio of posterior to prior frequencies of recombinant
-edges. Comparing the obtained null distribution of the ratios from the
-simulation and the ratio from the real data analysis, we assessed the
-significance of the ratio.
-% (GIVE VERY BRIEF SUMMARY OF SIMULATION APPROACH)
-%Ilan: point to supplementary methods for the simulation tests?
-Third,
-we traced the posterior
-probabilities of transfer types in a position-specific manner. 
-%For each
-%distinct type of transfer (determined by a pair of donor and recipient
-%edges), and each site in the core genome alignment, we computed the
-%posterior probability that that site had been affected by a transfer of
-%that type. 
-Finally, we defined a general, branch-independent
- {\em recombination intensity} at each site by 
-summing the posterior probabilities of recombination.
-In some cases, we also computed intensities for particular types of
-replacing transfers, such as 
-those between the SPY and SDE clades, or those producing topologies
-different from the clonal frame, or \textit{topology-altering} transfers.  
-% CHECK: Justify why we used only topology-altering edges for computing
-% recombination intensity.
-Note that \cite{Didelot2010} described the second and third summarization
-procedures.
+the sampled recombinant trees.
+First, we recorded the relative frequencies of the 105 possible
+rooted trees at each site in the alignments.  Second, we recorded the
+frequencies of sampled recombinant edges, grouping them by the associated
+donor and recipient branches in the clonal frame \citep{Didelot2010}, and
+took ratios of these sampled counts with corresponding expected values
+under the assumed 
+prior distribution.  Third, we recorded the frequencies of these edges
+types per site.  Finally, we defined a general, branch-independent {\em
+  recombination intensity} at each site by summing the posterior
+probabilities of all recombination edges.  In some cases, we also computed
+intensities for particular types of replacing transfers, such as those
+between the SPY and SDE clades, or those producing topologies different
+from the clonal frame
+(see \textbf{Supplementary Material}).
+
+{\bf Significance testing}.  We used simulations to assess the significance
+of the
+prior-normalized counts of recombinant edges.  First,
+we generated 100 replicate data sets under the prior model, with the same
+block number and blocks lengths as the real data.  We then applied our
+entire inference procedure to these data sets, and computed the same ratios
+of posterior estimates to prior expectations as for the real data.  We then
+compared the ratios estimated from real data with this empirical null
+distribution.  Empirical one-sided $p$-values were computed as the fraction
+of null estimates greater or less than the values estimated from real data.
+This approach had the advantage of not only assessing the significance of
+departures from prior expectations, but also correcting for any systematic
+biases imposed by ClonalOrigin
+(see \textbf{Supplementary Material}).
+
+To compare the rates of SPY-to-SDE and SDE-to-SPY gene flow, we computed,
+for each of the 1,001 genome-wide collections of recombinant trees, the
+number of recombinant edges from any of the three SPY branches (SPY1, SPY2,
+and SPY) to any of the three SDE branches (SDE1, SDE2, and SDE), and the
+number of recombinant edges in the reverse direction.  We then took the
+ratio of these two counts.  This gave us 1,001 samples from the posterior
+distribution of the ratio of SPY-to-SDE and SDE-to-SPY recombinant edges,
+from which we could estimate a posterior mean (2.0) and 95\% credible
+interval (1.8--2.2).  None of the sampled values was as small as the prior
+estimate of the ratio (1.4), indicating strong support in the data for a
+preference for the SPY-to-SDE direction.
 
@@
 that minimizes the number of gene duplication, loss, and
-transfer (DLT) events by considering tree topologies with branching orders, also
-known as labeled histories \citep{Edwards1970}.  We estimated
+transfer (DLT) events, considering tree topologies with branching orders
+(``labeled histories''; \citealp{Edwards1970}).  We estimated
 unrooted gene trees for the nucleotide alignments for each of our
-putatively non-recombining gene fragment using RAxML \citep{Stamatakis2006},
+putative non-recombining gene fragment using RAxML \citep{Stamatakis2006},
 then applied Mowgli to each of these trees together with the species
@@
 {\bf Distinguishing between replacing and additive HGTs.}  The transfer
-events inferred by Mowgli had to be classified as replacing or
-additive in a post-processing step because the program did not
-distinguish between these types of events.  A naive approach 
-would be to infer a replacing HGT whenever an inferred transfer
-coincided exactly with an inferred loss in the recipient lineage.  However,
-loss events were difficult to place accurately because they were frequently
-pushed toward the root of the tree due to reconstruction error or parallel
-losses in multiple lineages.  Therefore, we classified a transfer event as
-additive whenever a descendant of the recipient lineage in the species tree
-contained both a gene descended from the transferred copy and a gene not
-descended from that copy.  All other transfers were considered replacing
-HGTs.  This was conservative about calling additive HGTs, and would
-label any HGT that could plausibly be explained by replacing transfers
-as such (see fig.\ \ref{fig:hgt}).
-
-{\bf Statistical enrichment}. As with replacing
-HGTs, we classified transfer events by donor and recipient branch, 
-%sought to identify over-represented scenarios of additive HGT. 
-%and identified distinct types of additive
-%transfers by their donor and recipient branches in the species
-%tree, 
-and recorded the number of inferred additive HGTs of each type.
-%as well as the prior expected number of transfers
-%of that type. 
-We compared these numbers to expectations based on simulations that
-assumed 
-constant
-rates of duplication, loss, and transfer across the species tree, given 
-the total numbers of events inferred across all 2314 gene families
-divided by the total branch lengths of the trees
-(see \textbf{Supplementary Material}).
+events inferred by Mowgli had to be classified as replacing or additive in
+a post-processing step because the program did not distinguish between
+these types of events.  A naive approach would be to infer a replacing HGT
+whenever an inferred transfer coincided exactly with an inferred loss in
+the recipient lineage.  However, we found that loss events were difficult
+to place accurately because they were frequently pushed toward the root of
+the tree due to reconstruction error or parallel losses in multiple
+lineages.  Therefore, we classified a transfer event as additive whenever a
+descendant of the recipient lineage in the species tree contained both a
+gene descended from the transferred copy and a gene not descended from that
+copy.  All other transfers were considered replacing HGTs.  This approach
+was conservative about calling additive HGTs, and labeled any HGT that
+could plausibly be explained by replacing transfers as such (see
+supplementary fig.\ \ref{fig:calling-transfers}).
+
+{\bf Statistical enrichment}. As with replacing HGTs, we classified
+transfer events by donor and recipient branch, and recorded the number of
+inferred additive HGTs of each type.  We compared these numbers to
+expectations based on simulations that assumed constant rates of
+duplication, loss, and transfer across the species tree, given the total
+numbers of events inferred across all 2,314 gene families divided by the
+total branch lengths of the trees (see \textbf{Supplementary Material})
+% (supplementary fig.\ \ref{fig:mowgli-sim}).
 
@@
 
-We performed a series of analyses to search for certain classes of genes
-enriched for replacing or additive of gene transfers. In this analysis,
-each gene, or gene family, was associated with a certain score. For
-replacing gene transfers we used the recombination intensity along
-a given gene, and for additive gene transfers we used the number of
-additive transfers inferred for a given family. Mann-Whitney tests were
-used to identify sets of genes that showed significantly elevated values in
-each score.  A 5\% False Discovery Rate (FDR) correction
-\citep{Benjamini1995} was used to determine an appropriate significance
-threshold for our p-values. We applied the association test for gene
-classes defined by Gene Ontology (GO) terms assigned  
-%using the method described in Suzuki et al.\ \cite{Suzuki2011}. In short, 
-by comparing the Streptococcus genes to bacterial proteins from the
-Uniref90 database using blastP, and then assigning the same GO
-classification as the target gene of the uniProt GOA database, if the
-match had an E-values less than $1.0\times10^{-5}$. We also classified a
-category of \textit{virulence genes} using the virulence profile obtained
-by \cite{Suzuki2011}. In both cases, a gene family was
-associsated with a certain classification if any of its genes were
-associated with that classification.
+We assigned genes to Gene Ontology (GO) categories by comparing the {\em
+  Streptococcus} genes to bacterial proteins from the Uniref90 database
+using BLASTP, and then assigning the same GO classification as the target
+gene of the uniProt GOA database if the match had an $E$-value of
+$<1.0\times10^{-5}.$  Virulence genes were identified using the methods
+described by \cite{Suzuki2011}. A gene family was assigned a given
+classification if any of its genes was assigned that classification.  To
+test for associations with replacing gene transfers we used the
+recombination intensity estimated for each gene, and for additive gene
+transfers we used the number of additive transfers inferred for each
+family.  To test for significance, we performed a Mann-Whitney $U$ test of
+the values (recombination intensities or numbers of transfers) associated
+with a given category vs.\ the values for the all other genes/families.  We
+used the \cite{Benjamini1995} method to correct for multiple comparisons.
 
@@
 By applying ClonalFrame to our five-way genomic alignments
-(supplementary fig.\ S\ref{fig:mauve}), we obtained a
-phylogeny in which the pairs of SPY and SDE samples were grouped together,
-as expected, and in which SDD and SDE formed a clade, with SPY as an
-outgroup (fig.\ \ref{fig:tree5}).  This tree was consistent with previous
-results from 16S rRNA sequences \citep{Facklam2002} and with genome-wide
-trends \citep{Suzuki2011}.  Despite the grouping of SDD and SDE, their
-estimated divergence was substantial, at $\sim$0.48 substitutions per site.  
-The estimated divergence between SPY and SDD/SDE was $\sim$0.66
-substitutions per site.  
-%These estimates correspond to average
-%genomic divergence times of roughly 660 and 900 thousand years,
-%respectively, assuming a mutation rate of $5 \times 10^{-10}$ substitutions
-%per site per generation \citep{Ochman2003} and two cell divisions per day
-%\citep{XX}.  
-An initial analysis with ClonalOrigin (see \textbf{Methods}) produced an
-estimated
-per-site population-scaled mutation rate of $\theta = $ 0.081
-[interquartile range across blocks: (0.067, 0.094)], a per-site
-population-scaled recombination rate of $\rho = $ 0.012 (0.006, 0.019), and
-an average recombinant tract length of $\delta = $ 744 (346, 2848) bp.  The
-full distributions of these quantities across alignment blocks are shown in
-supplementary
-figure S\ref{fig:scatter3}.  Our estimates implied a ratio of $\rho/\theta
-= 0.15$.  For comparison, \cite{Didelot2010} obtained estimates of $\theta
-= 0.044$, $\rho = 0.017$, $\delta = 236$, and $\rho/\theta = 0.405$ for the
-{\em Bacillus cereus} group, but differences of this magnitude were not
-surprising, given likely differences in effective population sizes,
-recombination rates, and mutation rates \citep[e.g.,][]{Feil2001}.  We
-found that estimates of the recombination tract length, $\delta$, were
-quite sensitive to our threshold for minimum block length, with the
-inclusion of short alignments producing much larger estimates due to edge
-effects.  We experimented with a range of thresholds and selected one (1500
-bp) at which the estimates stabilized.
+we obtained a phylogeny in which the two SPY
+samples grouped together, as did the two SDE samples, and in which SDD and
+SDE formed a clade, with SPY as an outgroup (fig.\ \ref{fig:tree5}).  This
+species tree was consistent with previous results from 16S rRNA sequences
+\citep{Facklam2002} and with general patterns observed genome-wide
+\citep{Suzuki2011}.  The estimated levels of inter-species divergence were
+substantial, at 0.48 substitution per site for SDD and SDE, and 0.66
+substitutions per site for SPY and SDD/SDE.  An initial analysis with
+ClonalOrigin (see \textbf{Methods}) produced an estimated per-site
+population-scaled mutation rate of $\theta = $ 0.081 [interquartile range
+across blocks: (0.067, 0.094)], a per-site population-scaled recombination
+rate of $\rho = $ 0.012 (0.006, 0.019), and an average recombinant tract
+length of $\delta = $ 744 (346, 2848) bp.  The full distributions of these
+quantities across alignment blocks are shown in supplementary figure
+\ref{fig:scatter3}.  A second run of the MCMC sampler produced nearly
+identical results.  For comparison,
+\cite{Didelot2010} obtained estimates of $\theta = 0.044$, $\rho = 0.017$,
+and $\delta = 236$ for the {\em Bacillus cereus} group.  Our estimates
+implied a ratio of $\rho/\theta = 0.15$, considerably lower than Didelot et
+al.'s \citeyearpar{Didelot2010} estimate of $\rho/\theta = 0.405$ for {\em
+  B.\ cereus}, suggesting increased rates of mutation and/or decreased
+rates of recombination after controlling for differences in effective
+population size.  We found that estimates of the average recombinant tract
+length, $\delta$, were quite sensitive to our threshold for minimum block
+length, with the inclusion of short alignments producing much larger
+estimates due to edge effects.  We experimented with a range of thresholds
+and selected a value (1500 bp) at which the estimates stabilized.
 
@@
 
-With a second stage of ClonalOrigin MCMC, we were able to obtain samples
-from the posterior distribution of recombinant trees (i.e., the clonal
-frame augmented by recombinant edges) along the genome, conditional on our
-five-way alignments and the previously estimated parameters (see \textbf{Methods}).
-This distribution indicated that most sites (66\%) were
-consistent with the topology of the clonal frame.
-% what does this mean?  most likely tree topology per site?  or expected value?
-Interestingly, the second most frequent tree topology (9.3\%) had
-SPY and SDE as sister clades and SDD as an outgroup (supplementary table
-S\ref{tab:Gene-tree-topologies}), providing an initial indication of gene
-transfer between SPY and SDE.  No other tree topology appeared with
-appreciable frequency.
-% This indicated that SPY and SDE appeared to be more closely related with each
-% other than each to SDD's genome in about tenth of the core genome. The other
-% topologies had less than 4\% of proportions.
-
-% eliminate some redundancy with methods
+In a second round of analysis with ClonalOrigin, we obtained samples from an
+approximate posterior distribution of recombinant trees along the genome,
+conditional on our five-way alignments and the previously estimated
+parameters. (A recombinant tree is modeled by ClonalOrigin as the clonal
+frame augmented by zero or more recombinant edges; see \textbf{Methods}.)
+This distribution indicated that a majority of sites in the genome (66\%)
+were topologically consistent with the clonal frame, but that there was
+also substantial support for various alternative tree topologies (supplementary
+table \ref{tab:Gene-tree-topologies}).  Interestingly, the second most
+frequent tree topology (9.3\% of sites) had SPY and SDE as sister clades
+and SDD as an outgroup, providing an initial indication of gene transfer
+between SPY and SDE.  No other single tree topology appeared with a
+frequency of $\geq$4\%.
+
 To gain further insight into rates and patterns of HGT, we estimated the
-rates of occurrence of recombinant edges, grouping them by the associated
-donor and recipient edges in the clonal frame (see \textbf{Methods}).  Because donor
-edges must pre-date recipient edges under the model, 21 of the 81 edge
-pairs for the nine-branch rooted topology are prohibited, leaving 60
-possible recombinant edge types, each corresponding to
-a class of replacing HGTs.  
-% Because of the possibility of delayed
-% coalescence, these classes include not only transfers from the donor to the
-% recipient edge in the clonal frame, but also transfers from any descendant
-% of the donor to the recipient (with the implicit constraint of
-% contemporaneous donors and recipients).  Thus, recombinant edges are
-% informative about replacing HGTs, but with some ambiguity.  
-Because recombinant edges are not all equally likely \textit{a priori}, 
-we summarized the estimated
-rates by taking ratios with respect to their prior expectations under the
-model used by ClonalOrigin, which consider the branch lengths of the clonal
-frame.  Experiments conducted on simulated data demonstrated that our
-approach could accurately detect recombination scenarios that deviate from
-the prior, with some underestimation of enrichment levels 
-(see \textbf{Supplementary Material}).
-
-These normalized rates for our alignments can be summarized in a $9 \times
-9$ heatmap, consisting of 60 informative and 21 undefined cells (fig.\
+rates of occurrence of various types of recombinant edges, grouping them by
+their donor and recipient edges (see \textbf{Methods}).  There are 81
+possible edge pairs associated with our nine-branch rooted phylogeny, but
+because donor edges cannot post-date recipient edges under the model, 21 of
+these pairs are prohibited, leaving 60 possible recombinant edge types.
+Each of these viable edge types corresponds to a class of replacing HGTs,
+including not only transfers from the donor to the recipient edge, but also
+transfers from any (sufficiently old) descendant of the donor to the
+recipient (due to the possibility of delayed coalescence).  Because all
+recombinant edges are not equally likely \textit{a priori}, we summarized
+the estimated rates by taking ratios with respect to their prior
+expectations under the ClonalOrigin model.  Experiments conducted on
+simulated data demonstrated that our approach had reasonable power to
+detect recombination scenarios that deviated from the prior, with some
+underestimation of the degree of deviation due to the use of the prior in
+the inference procedure 
+(see \textbf{Supplementary Material};
+supplementary figures\ \ref{fig:sim2} and \ref{fig:sim3}).
+
+We summarized these normalized rates in a heatmap (fig.\
 \ref{fig:Heatmap-of-transfers}A; see also supplementary tables
-S\ref{tab:heatmap} and S\ref{tab:obsheatmap}).  Most cells in this heatmap
+\ref{tab:obsheatmap} and \ref{tab:heatmap}), and found that most cells
 fell in the blue-to-white range, corresponding to rates of occurrence less
-than or equal to what was expected under the prior.  Some of these
-correspond to edges that were fairly strongly under-represented (blue),
-such as those between the SDD branch and the branches of the SPY clade
-(P-value of $<0.01$, based on an empirical test; see \textbf{Methods}),
-possibly reflecting ecological isolation of human and strictly veterinary
-pathogens.  By contrast, the recombination edges between the two SDE
-genomes, and between the two SPY genomes, were significantly
-over-represented (P-value of $<0.01$), consistent with previous evidence
-for intraspecies recombination in {\em Streptococcus} \citep{Feil2001}
-(CHECK).  In addition, a pronounced enrichment was observed in
-recombination edges between the SPY and SDE clades, as had been reported
-previously \citep{Kalia2001,Sachse2002,Kalia2004,Davies2005,Davies2007a}.
-While these edges occurred at elevated rates in both directions, we found a
-clear asymmetry, with edges in the SPY-to-SDE direction occuring at up to
-four times the expected rate, while those in the reverse direction occur at
-only about twice the expected rate.  \texttt{CHECK:} To directly compare
-the gene flow of SPY-to-SDE with that of SDE-to-SPY for each of 1001
-recombinant trees from the posterior distribution we summed numbers of
-recombinant edges from each of the three branches SPY1, SPY2, and SPY to
-each of the three SDE1, SDE2, and SDE, and did so for the reverse
-direction.  Taking the ratio of the first sum to the second resulted in a
-distribution of the relative strength of gene flow between the two clades.
-None of the ratio values was less than the corresponding ratio from the
-prior (95\% quantile range was $1.79 <$ posterior ratio $< 2.24$, and prior
-ratio was 1.41; P-value of $< 0.001$).  This provided genome-wide support
-for Kalia et al.'s \citeyearpar{Kalia2001} unconfirmed observations of
-directional asymmetry based on MLST data for seven loci.  We also observed
-a slight enrichment for SDE-to-SDD edges.
-
-% WORK IN DISCUSSION OF TABLE supplementary S\ref{tab:gene-counts-replacing}, SPECIFIC
-% NUMBERS OF SPY$\rightarrow$SDE and SDE$\rightarrow$SDD TRANSFERS.  
-\textbf{CHECK:} We counted genes with noticeable recombination probability 
-($>0.6$) for types of recombinant edges
-(supplementary table S\ref{tab:gene-counts-replacing}). There were more genes
-that underwent SPY-to-SDE replacing gene transfer than those of SDE-to-SPY
-transfer. This recapitulated the argument of asymetry of recplacing gene flow.
-In the subset of virulence genes, there were three genes that underwent
-SPY-to-SDE replacing transfer whereas there was one gene with the reverse
-direction. 
-Virulence genes and non-virulence genes did not seem to be different in the
-relative frequency of replacing gene transfer
-although the proportion of virulence genes with SPY-to-SDE replacing gene
-transfer decreased compared with that of the total genes including the
-virulence genes.
+than or equal to what was expected under the prior.  A few edge types were
+strongly under-represented (blue), such as those between the SDD branch and
+the branches of the SPY clade ($p<0.01$ based on simulations; see
+\textbf{Methods}), probably due to ecological isolation of human and
+strictly veterinary pathogens.  Edges from the branch at the root were also
+under-represented, suggesting a deficiency of long delays in coalescence
+relative to the prior.  By contrast, the recombination edges between the
+two SDE genomes, and between the two SPY genomes, were significantly
+over-represented ($p<0.01$), consistent with previous evidence for abundant
+intraspecies recombination in {\em Streptococcus} \citep{Feil2001}.  In
+addition, we found a pronounced enrichment for recombinant edges between
+the SPY and SDE clades, as has been reported previously
+\citep{Kalia2001,Sachse2002,Kalia2004,Davies2005,Davies2007a}.  We also
+observed a slight enrichment for SDE-to-SDD edges.
+
+While SPY-SDE edges occurred at elevated rates in both directions, they
+showed a pronounced directional asymmetry, with edges from SPY to SDE
+occurring at up to four times the expected rate, while those in the reverse
+direction occurred at only about twice the expected rate.  By reanalyzing
+the sampled recombinant trees, we were able to obtain a Bayesian posterior
+mean estimate of 2.0 (95\% credible interval: 1.8--2.2) for the ratio of the
+numbers of SPY-to-SDE to SDE-to-SPY recombinant edges, compared with a ratio of 1.4
+implied by the prior (the prior ratio is larger than one because of
+differences in branch lengths).  In addition, we counted genes having high
+probability of recombination ($>$0.6) for various types of recombinant
+edges (supplementary table \ref{tab:gene-counts-replacing}), and found
+that many more genes showed strong evidence of SPY-to-SDE than of SDE-to-SPY
+recombinant edges (for example, 59 vs.\ 11 genes, when the entire SPY and
+SDE clades are considered).  These findings provided genome-wide support
+for Kalia et al.'s \citeyearpar{Kalia2001} previously unconfirmed
+observations of directional asymmetry based on MLST data for seven loci.
 
@@
 Genome Browser (http://strep-genome.bscb.cornell.edu) that summarize the
-results of the ClonalOrigin analysis alongside known genes, alignments, and
+results of this model-based analysis alongside known genes, alignments, and
 other annotations (fig.\ \ref{fig:ucsc}).  These tracks can be used to
 inspect loci of interest and to compare the results of our model- and
-parsimony-based analyses.  They can also be queried and intersected with
-other tracks using the UCSC Table Browser.
-
-% follow-up
-% first, give bayesian posterior prob of asymmetry
-% then something about effect of prior
-% then try Ilan's time-interval test
+parsimony-based analyses (below).  They can also be queried and intersected
+with other tracks using the UCSC Table Browser.
 
@@
 
-% bring out comparison with replacing (will have to explore more)
-% especially asymmetry, SPY-SDE
-
-To gain further insight into HGT in {\em Streptococcus},
-we estimated gene trees for 2314 gene families representing the six 
-genomes (including an SEE outgroup), used the Mowgli
-program \citep{Doyon2011} to find parsimonious 
-reconciliations of these gene trees with the estimated clonal frame
-(fig.\ \ref{fig:tree5}), and then, in a post-processing step, partitioned
-the inferred transfers into replacing and additive HGTs (see \textbf{Methods}).
-This analysis produced estimated numbers of five types of events on each
-branch of the phylogeny (gene duplications, losses, replacing and additive
-transfers, and ``appearances,'' i.e., transfers from phylogenetically
-distant species), as well as an estimated number of genes at each ancestral
-node of the tree (fig.\ \ref{fig:Gene-duplication-loss}).
+To gain further insight into HGT in {\em Streptococcus}, we estimated gene
+trees for 2,314 gene families from six genomes (including the SEE outgroup),
+obtained parsimonious reconciliations of these gene trees with the clonal
+frame (fig.\ \ref{fig:tree5}) using the Mowgli program \citep{Doyon2011},
+and partitioned the inferred transfers into replacing and additive HGTs
+(see \textbf{Methods}).  This analysis produced estimated numbers of five
+types of events for each branch of the phylogeny (gene duplications,
+losses, replacing and additive transfers, and ``appearances,'' most of
+which are probably transfers from phylogenetically distant species), as
+well as an estimated number of genes at each ancestral node 
+(fig.\ \ref{fig:Gene-duplication-loss}).  We observed large numbers of
+appearances on all branches of the tree, suggesting a steady influx of
+genes into the clade by HGT.  Indeed, appearance events accounted for 1,957
+of the 3,432 (57\%) gene additions (by duplication, transfer, or appearance;
+loss events are excluded here).  Interestingly, the somewhat reduced
+numbers of genes in SPY (particularly in SPY1) were primarily explained by
+reduced rates of gene appearance, rather than by increased loss or other
+factors.  The estimated numbers of ancestral genes tended to decrease
+toward the root of the tree, but this likely reflects under-estimation due
+to parallel losses of some genes (especially on the branches beneath the
+root) rather than a true increase in gene number over evolutionary time.
+An exception was the branch leading to the ancestor of the two SPY
+individuals, which was enriched for gene loss events, perhaps associated
+with niche adaptation.  High rates of loss also occurred on the branches to
+SDE1, SDE2, and SDD.  Duplications were relatively infrequent overall, but,
+as has been noted previously \citep{Marri2006}, they were substantially
+enriched on external branches of the phylogeny.
+
 % total numbers of events:
@@
 % replacement 928
-We observed large numbers of appearances on all branches of the tree,
-suggesting a steady influx of genes into the clade by HGT.  Indeed,
-appearance events accounted for 1957 of the 3432 (57\%) gene additions (by
-duplication or transfer).  Interestingly, the somewhat reduced numbers of
-genes in SPY (particularly in SPY1) were primarily explained by reduced rates
-of gene appearance, rather than by increased loss or other factors.  The
-estimated numbers of ancestral genes tended to decrease toward the root of
-the tree, but this likely reflects under-estimation due to parallel losses
-of some genes (especially on the branches beneath the root) rather than a
-true increase in gene number over evolutionary time.  An exception was the
-branch leading to the ancestor of the two SPY individuals, which was
-enriched for gene loss events, perhaps associated with niche adaptation.
-High rates of loss also occurred on the branches to SDE1, SDE2, and SDD.
-Duplications were relatively infrequent overall, but, as had been noted
-previously \citep{Marri2006}, they were substantially enriched on external
-branches of the phylogeny.
-
-%When normalized for branch length (Figure \ref{fig:XX})
-
-%The largest family consisted of 59 genes while nearly half of gene families
-%(1066) had exactly one gene per species, or six genes total (Figure
-%S\ref{fig:famsizes}).
-
-We conservatively identified 290 of the 1218 (23.8\%) transfers (excluding
-appearances) as additive HGTs. 
-As with the replacing HGTs evaluated in the model-based analysis, these
-predicted additive transfers were enriched within the SPY and SDE clades 
-(fig.\ \ref{fig:Heatmap-of-transfers}B).  
-They were also significantly enriched between
-SDE to SDD (with a preference for the SDE-to-SDD direction), and
-significantly depleted for SDD-to-SPY events.
-However, unlike the replacing HGTs above, these additive transfers were only
-slightly enriched between SPY and SDE, and did not exhibit a pronounced
-directional asymmetry.  This suggests that the apparent
-gene flow between SPY and SDE, particularly in the SPY-to-SDE
-direction, has been driven by homologous recombination rather than additive
-transfer.  However, our analyses of replacing and
-additive transfers were not directly comparable for various reasons---for
-example, the first considered the core genome only, while the second
-considered the pan genome; and the two types of analysis might have quite
-different power for different types of events.
-To examine this issue further, we.... (REPORT BRIEF FOLLOW-UP ANALYSIS)
-% follow-up analysis
-We also compared the replacing HGTs identified in the parsimony-based
-analysis with those from the model-based analysis, and found reasonable
-concordance, despite substantial
-differences between the data sets and methods 
-(supplementary fig.\ S\ref{fig:cmpcomowgli}).
-%Ilan: need to refer to appropriate supp figure
-
-\subsection*{Functional Categories of Transferred Genes}
+
+We conservatively identified 290 of the 1,218 (non-appearance) transfers
+(23.8\%) as additive HGTs.  As with the replacing HGTs evaluated in the
+model-based analysis, these predicted additive transfers were significantly
+enriched within the SDE clade, and they were weakly enriched within the SPY
+clade (fig.\ \ref{fig:Heatmap-of-transfers}B and
+supplementary fig.\ \ref{fig:mowgli-sim}).
+They were also
+significantly enriched between SDE and SDD, with a preference for the
+SDE-to-SDD direction, and significantly depleted between SDD and SPY.
+However, unlike the replacing HGTs from the previous section, these
+additive transfers were not significantly enriched between SPY and SDE.
+They also did not exhibit a pronounced directional asymmetry between SPY
+and SDE, except between the ancestral SPY and SDE branches, where they were
+significantly depleted in both directions, but more strongly from
+SDE-to-SPY.  It is worth noting, however, that this apparent asymmetry in
+depletion could in fact reflect an asymmetry in enrichment, as the
+normalization in this case reflects average rates across the tree, and an
+inflated average could cause a global shift in the heatmap toward depletion
+(blue).  Other differences also make it difficult to compare the model- and
+parsimony-based analyses---for example, the first considered the core
+genome only, while the second considered the pan genome; and the two types
+of analysis might have quite different power for events affecting different
+portions of the phylogeny.  Nevertheless, we find much clearer evidence
+that homologous recombination, rather than additive transfer, has driven
+the apparent gene flow between SPY and SDE, particularly in the SPY-to-SDE
+direction.  Notably, we compared the replacing HGTs identified in the
+parsimony-based analysis with those from the model-based analysis, and
+found reasonable concordance, despite substantial differences between the
+data sets and methods
+(see \textbf{Supplementary Material}).
+
+\subsection*{Functional categories of transferred genes}
 
 To gain insight into the functional impact of HGT, we assigned genes to
-functional classes and tested each class to see whether it was enriched for
-genes or gene families predicted to have experienced gene transfer events
-(see \textbf{Methods}).  First, we identified a subset of genes whose homologs
-in other
-genera of bacteria were known to exhibit virulence phenotypes (based on
-VFDB; see \citealp{Suzuki2011}) and considered all other genes to be
-non-virulence genes.  These virulence genes were indeed significantly
-enriched among gene families for which we inferred additive HGTs 
-(P-value of $1.33 \times 10^{-11}$), 
-possibly reflecting an increased rate of fixation of
-these transfers due to positive selection for virulence-related functions.
-% too much?
-Interestingly, SPY-to-SDE events were strongly over-represented (by
-$>$2x) among virulence genes that were associated with additive HGTs 
-(supplementary table S\ref{tab:gene-counts-additive}).  
-Transfers in the reverse direction
-(SDE-to-SPY) are also over-represented, but slightly less so
-(1.7x).  By contrast, virulence genes as a group showed no significant
-enrichment for replacing gene transfers, and virulence genes that exhibit
-strong evidence of replacing transfers were not enriched for SDE/SPY % SDE/SDD
-transfers (rather, they were enriched for SDE-to-SDD transfers;
-supplementary table S\ref{tab:gene-counts-replacing}).
-We also found a weak association between virulence genes and gene
-duplications (P-value of $7.13 \times 10^{-3}$).
-
-% (POSSIBLY HIGHLIGHT AN EXAMPLE OR TWO WITH A BROWSER SHOT)
-% Genome browser parts is moved to page 13.
-
-% reduce this to a couple of sentences
-We also 
-searched for gene ontology (GO) terms significantly associated with the
-various types of 
-transfer. Among the categories found to be associated with additive gene
-transfers (table \ref{tab:go-events}), we found several transposase-related 
-transposase.  
-%This can be explained by the role transposition plays in
-%integrating transfered genes into a new host.
-%Ilan: this explanation requires some work
-Replacing transfers appear to show quite different associations from the
-additive ones (table \ref{tab:functional}). (NEED TO EXPAND THIS A BIT)
-%We
-%compared associations found with replacing gene transfers inferred in our
-%model-based analysis with associations found with the number of inferred
-%replacing transfer events in the parsimony analysis, and overall found
-%consistent associations (Table S\ref{tab:go-events-recombining}).
-
-% can we still do something with examples?
-
-%Ilan: this part still under construction
-
-%THE FOLLOWING PARAGRAPH NEEDS SOME WORK
-
-%We used the genome browser tracks for posterior probability of
-%recombination to locate genes which showed an elevated signal of replacing
-%gene transfer. We constructed a list of genes that contain sites showing a
-%posterior probability greater than 0.8 for a given type of transfer
-%(indicated by donor and recipient branches). Each such gene is associated
-%with a transfer type (or several types). We find overall ?? genes fitting
-%these criteria, with ?? of them associated with replacing transfers from
-%SPY to SDE, and ?? associated with replacing transfers from SPY to
-%SDE... Focusing on virulence genes we see a similar distribution of transfer
-%types...
-
-%Ilan: do we want a table with the number of genes associated with each transfer type?
-
-% IS THERE SOMETHING WORTH KEEPING HERE?
-%\texttt{INFORMATIONAL / OPERATIONAL ANALYSIS:}
-%We grouped informational genes based on gene ontology to test if there was
-%differences in replacing gene transfer between informational genes and
-%operational genes. We found differences in replacing gene transfer from three
-%SDE branches to three SPY branches (p-value=0.04409 with Wilcoxon rank sum
-%test).
-%\\
-%\\
-
-\subsection*{Positive Selection and Gene Transfer}
-% selection analysis 
-We also tested for evidence of positive selection within the gene
-families and their association with various events.  We performed a
-likelihood ratio test using PAML's ``sites models'' (M1a and M2b) on
-each of the 2991 gene-fragment families with three or more genes.  Of
-those families tested, 31 showed evidence of positive selection
-(FDR$<$5\%).  For gene trees with high bootstrap ($>$90\% on every
-branch), we find that duplication events are slightly enriched with
-positive selection (P-value of $<0.046$; Mann-Whitney U) as well as
-families with additive transfers (P-value of $<0.002$; Mann-Whitney
-U).  Of the 31 families with significant positive selection, 13
-families were part of the core genome analyzed by ClonalOrigin.  We
-found that replacing transfers from SDE to SPY and from SPY to SDE to
-be slightly enriched within the positively selected families (P-value
-of $<0.005$ and $<0.01$; Mann-Whitney U).  For 11 of the 13
-(85\%) families the SPY-to-SDE direction had greater intensity than
-SDE-to-SPY, consistent with overall directional bias (83.3\%).  
-The same directional bias was not seen for additive
-transfers between the two clades.
+various functional categories and looked for statistical associations
+between these categories and predicted gene transfer events (see
+\textbf{Methods}).  First, we partitioned all genes into ``virulence'' and
+``non-virulence'' categories, considering genes whose homologs in other
+genera of bacteria exhibit virulence phenotypes (based on VFDB) to be
+virulence genes \citep[see][]{Suzuki2011}.  Among gene families for which
+we inferred additive HGTs, virulence genes were significantly enriched
+($p=1.33 \times 10^{-11}$), possibly due to selection favoring
+virulence-related functions.  Interestingly, virulence genes associated
+with additive HGTs showed a pronounced enrichment for SPY-to-SDE events
+(25.0\% of events compared with 12.1\% for all genes at the level of the
+entire SPY and SDE clades, $p=0.01$, Fisher's exact test; supplementary
+table \ref{tab:gene-counts-additive}). 
+% this is based on a one-sided test of the follow 2x2 table
+%     [,1] [,2]
+%[1,]   10   30
+%[2,]   22  203
+Transfers in the reverse direction (SDE-to-SPY) were also over-represented
+(17.5\% vs.\ 10.6\%), but not significantly ($p=0.11$).  Replacing gene
+transfers, on the other hand, were not significantly enriched for virulence
+genes, and those virulence genes that did exhibit strong evidence of
+replacing transfers were not enriched for SDE/SPY transfers (supplementary
+table \ref{tab:gene-counts-replacing}).  We also found a significant
+association between virulence genes and gene duplications ($p=7.13 \times
+10^{-3}$).
+
+Next, we assigned gene ontology (GO) terms to all genes, and searched for
+significant associations with transfer events.  Several GO categories
+having to do with transposition, recombination, and DNA integration were
+significantly associated with additive gene transfers (table
+\ref{tab:go-events}), consistent with observations in other bacterial
+groups \citep[e.g.,][]{Liu2009}.  Replacing transfers, on the other hand,
+showed much weaker, and quite different, functional associations compared with
+additive transfers.  These associations were similar for replacing
+transfers inferred by the model- and parsimony-based methods (supplementary
+tables \ref{tab:functional} and \ref{tab:go-events-recombining}).
+
+\subsection*{Positive selection and gene transfer}
+
+% some of this should really go in "Methods".  Should also explain exactly
+% how the MWU tests are done
+
+Finally, we tested for evidence of positive selection within the gene
+families, and for associations between positive selection and various types
+of events.  We performed a likelihood ratio test using PAML's ``sites
+models'' of M1a and M2b \citep{Yang2007}
+on each of the 2,991 gene-fragment families with three or more genes.
+After controlling for multiple comparisons, 31
+families showed evidence of positive selection (FDR$<$5\%).  Among
+high-confidence gene trees (bootstrap values of $>$90\% on every branch),
+families with duplication events were slightly enriched for evidence of
+positive selection ($p<0.046$, Mann-Whitney $U$ test [MWU]) as were families
+with additive transfers ($p<0.002$, MWU).  Of the 31 families with
+significant positive selection, 13 families were part of the core genome
+analyzed by ClonalOrigin.  Replacing transfers from SDE to SPY and from SPY
+to SDE were slightly enriched for positive selection ($p<0.005$ and
+$p<0.01$, MWU).  For 11 of the 13 (85\%) families showing evidence of
+positive selection, the SPY-to-SDE direction had greater intensity than the
+SDE-to-SPY direction, consistent with overall directional bias (83.3\%).
+The same directional bias was not seen for additive transfers between the
+two clades.
 
@@
 
-% accurate to say "recent"?  has been going on for some time
-There has been a great deal of recent interest in the process of horizontal
-gene transfer (HGT) and the manner in which it has influenced phylogenetic
-relationships, particularly among bacteria \citep{}.  Most previous studies,
-however, have either simply examined patterns of phylogenetic discordance,
-without differentiating between additive and replacing HGTs \citep{}, or
-have focused specifically on the process of homologous recombination,
-corresponding to replacing HGTs
-\citep{Didelot2007,Didelot2010}.
-% too strong?  neither are we
-% cite Feil?   Marri.  LEfebre.  recent review (Doolittle?)
-% another homologous recombination review -- see Awadalla
-To our knowledge, this is the first study to examine both
-processes on a genome-wide scale, using both model- and parsimony-based
-methods.
-
-We have focused our analysis on {\em Streptococcus pyogenes}
-(SPY) and two of its close relatives, {\em S.\ dysgalactiae} ssp.\ {\em
-  equisimilis} (SDE) and {\em S.\ dysgalactiae} ssp.\ {\em dysgalactiae}
-(SDD), a human commensal-like organism and a strict veterinary pathogen,
-respectively.  Our combined analysis reveals strong evidence of gene flow
-both within and between the SPY and SDE groups.  Interestingly, the SPY-SDE
-transfers are far more prominent among replacing than among additive
-transfers, suggesting that they are driven by homologous recombination.
-Moreover, we find strong support for an asymetry in replacing gene
-transfers between SPY and SDE, with a preference for the SPY-to-SDE
-direction.  The situation with SDD is more complex, with somewhat elevated
-rates of SDE-to-SDD transfer, more pronounced for additive than for
-replacing HGTs, mixed evidence for SDD-to-SDE transfer, and reduced rates
-of transfer between SDD and SPY, perhaps owing to a combination of genetic
-divergence and different ecological niches.
-% maybe revise the below after tuning up virulence section
-We also find that virulence genes are significantly enriched among
-additive, but not replacing, gene transfers, as are genes under positive
-selection.  Genes that have undergone replacing transfers between SPY and
-SDE are also enriched for positive selection.
-A central component of our study is the careful use of simulations to
-demonstrate that our methods have good power and accuracy in the
-detection of both types of transfer events.
-
-%We also confirm a species phylogeny (clonal frame) consistent with previous
-%results 
-%\citep{Facklam2002,Suzuki2011}
-
-%paragraph on Kalia et al. and the evidence for SDE/SPY gene flow
-%reasons for asymmetry -- barriers, etc.
+There has been a great deal of interest for decades in the process of
+horizontal gene transfer (HGT) and the manner in which it has influenced
+phylogenetic relationships, particularly among bacteria \citep{Koonin2001}.
+Most previous studies, however, have either simply examined patterns of
+phylogenetic discordance, without differentiating between additive and
+replacing HGTs \citep[e.g.,][]{Lerat2005}, or have focused specifically on
+the process of homologous recombination, corresponding to replacing HGTs
+\citep{Didelot2007,Didelot2010}.  To our knowledge, this is the first study
+to examine both processes on a genome-wide scale, using both model- and
+parsimony-based methods.
+
+We have focused our analysis on {\em Streptococcus pyogenes} (SPY) and two
+of its close relatives, {\em S.\ dysgalactiae} ssp.\ {\em equisimilis}
+(SDE) and {\em S.\ dysgalactiae} ssp.\ {\em dysgalactiae} (SDD), a human
+commensal-like organism and a strict veterinary pathogen, respectively.
+Our combined analysis reveals strong evidence of gene flow both within and
+between the SPY and SDE groups.  Interestingly, events between SPY and SDE
+are more prominent among inferred replacing than among inferred additive
+transfers, suggesting that they may be driven by homologous recombination,
+although we acknowledge several challenges in comparing inferences from the
+model- and parsimony-based analyses.  We also find strong support for an
+asymmetry in replacing gene transfers between SPY and SDE, with a
+preference for the SPY-to-SDE direction.  The situation with SDD is more
+complex, with somewhat elevated rates of SDE-to-SDD transfer, more
+pronounced for additive than for replacing HGTs, mixed evidence for
+SDD-to-SDE transfer, and generally reduced rates of transfer between SDD
+and SPY, perhaps owing to a combination of genetic divergence and different
+ecological niches.  We also find that virulence genes are significantly
+associated with additive, but not replacing, gene transfers, as are genes
+under positive selection.  Genes that have undergone replacing transfers
+between SPY and SDE are also enriched for positive selection.  A central
+component of our study is the use of simulations to demonstrate that our
+methods have good power and accuracy in the detection of both types of
+transfer events.
 
@@
 SPY and SDE raises a number of questions.
-% this is perhaps not so pressing; discuss briefly
 First, can Kalia et al.'s \citeyearpar{Kalia2001} observations be
 confirmed using our data, despite the authors' inability to reproduce them?
-\texttt{CHECK:} 
-Interestingly, we did find 
-that the six of their seven loci represented in our five-way alignments
+%\texttt{CHECK:} 
+Six of their seven loci were represented in our five-way alignments
 (\textit{gki}, \textit{gtr}, 
-\textit{murI}, \textit{mutS}, \textit{recP}, and \textit{xpt}) showed
-consistent results with those of \citet{Kalia2001}; high recombination in 
-\textit{recP}, and \textit{xpt} but not the others. 
-Recombination in \textit{gki} and \textit{mutS}, albeit weak, were also
-observed (despite that Kalia et al.\ found no evidence for {\em mutS}).
-Our visual inspection of the recombination probability on 
-\textit{recP} and \textit{xpt} indicated that 
-the replacing gene transfer appeared to be biased towards
-the directions from SPY to SDE.
+\textit{murI}, \textit{mutS}, \textit{recP}, and \textit{xpt}) 
+and we did find 
+strong evidence of recombination two of them 
+(\textit{recP} and \textit{xpt}) and weak evidence in two others
+(\textit{gki} and \textit{mutS}).  
+These results were generally consistent those of \citet{Kalia2001}, except
+that they found no recombination in {\em mutS}.
+%Our visual inspection of the recombination probability on 
+%\textit{recP} and \textit{xpt} indicated that 
+%the replacing gene transfer appeared to be biased towards
+%the directions from SPY to SDE.
+%that the six of their seven loci represented in our five-way alignments
+%(\textit{gki}, \textit{gtr}, 
+%\textit{murI}, \textit{mutS}, \textit{recP}, and \textit{xpt}) showed
+%consistent results with those of \citet{Kalia2001}
 % In S. pyogenes MGAS315 genome track, we can check
@@
 %
-%\texttt{SC: I am not sure, yet, because I wrote,
-%Except for \textit{mutS} where
-%they did not find recombination, our results were
-%consistent with those of \citet{Kalia2001}. Let me make sure this.}
-Of course, our study differs from theirs in numerous respects, but this
-suggests the two loci are prone to recombination.
-%Except for \textit{mutS} where
-%\cite{Kalia2001} did not find recombination, our results were
-%consistent with those of \cite{Kalia2001}.  
-%\cite{Pinho2010} reported recombination in the gene \textit{parC}, in which we
-%also observed moderate recombination.
+%Of course, our study differs from theirs in numerous respects, but this
+%suggests the two loci are prone to recombination.
+%
 Second, are the observed patterns of gene flow truly driven by replacing
@@
 evidence is weaker, because our parsimony-based method has limited ability
-to resolve the direction of transfer.  More genomes will be needed to
+to resolve the direction of transfer.  
+% FIXME: THIS NEEDS WORK
+More genomes will be needed to
 evaluate the directionality of additive transfers between these groups.
@@
 repair.  
-%Restriction-modification
-%system is a biological self-defense mechanism of bacteria, which is diverse
-%in prokaryote world.  
 Interestingly,
@@
 RMSs, the two SDE genomes studied here contain
-only type II RMSs.  (CHECK: THIS NEEDS SOME WORK)
+only type II RMSs \citep{Roberts2010}.
+%  (THIS STILL NEEDS SOME WORK)
 
@@
 
-% be careful not to undercut our findings here.
-
-% maybe more than a paragraph of limitations, make it one about insight --
-% idea of the two events being identified with different parts of the genome
-
 In comparing replacing and additive HGTs, it is worth bearing in mind that
 these events are likely to be strongly correlated with the core and
-``dispensable'' portions of the genome (we take the dispensible genome to
-be the difference between the pan and core genomes).  Methods for detecting
-recombination (replacing transfer) tend to rely on large-scale alignments
-spanning multiple syntenic loci in order to gain power.  As a result, these
-methods are effectively limited to application in the core genome, even
-though homologous recombination probably also occurs (perhaps at a reduced
-rate) between genes that are less widely shared across species.  By
+dispensable portions of the genome.  In part, this reflects an
+ascertainment bias in the inference procedures used to detect these events.
+Methods for detecting recombination (replacing transfers) tend to rely on
+large-scale alignments spanning multiple syntenic loci in order to gain
+power.  As a result, these methods are effectively limited to application
+in the core genome, even though homologous recombination probably also
+occurs between genes that are less widely shared across species.  By
 contrast, phylogenetic reconciliation methods can be applied to the entire
-pan genome, as in this study.  Moreover, the additive transfer events that
-they detect will necessarily be enriched in the dispensible genome.  Thus,
-the replacing HGTs identified by recombination-based methods will tend to
-reflect the core genome and the additive HGTs identified by
-reconcilation-based methods will tend to reflect the dispensible genome.
-In our view, this is not necessarily a limitation of the comparison,
-because (with the exception of some replacing transfers in the dispensible
-genome) we expect the inferred transfer events to be fairly representative
-of true HGTs, but it is useful to keep in mind that some apparent
-differences between the two 
-types of transfer events may simply reflect general differences between the
-core and dispensible genome.
-
-%In addition, these methods are likely to have quite different power for various
-%kinds of events, because of factors such as
-%branch-length dependency
-%use of multiple loci or single loci
-%larger trees
-
-%paragraph on future methods
-As sequence data becomes available for many more species, including both
-closely and distantly related organisms, it will become increasingly
-important to devise improved methods that consider both additive and
-replacing HGT.  As noted above, the current statistical-models of
-recombination, such as ClonalOrigin, are effectively limited to considering
-orthologs, with one copy per species, in the core genome.  The phylogenetic
-reconciliation methods, on the other hand, fail to allow for multi-locus
-gene transfer events (which are known to occur), and also ignore
-information from branch lengths.  We have shown here that it is possible,
-to a degree, to extend them to distinguish between additive and replacement
-transfers but our methods are limited to a degree by their dependency on
-heuristic rules and parsimony assumptions (reference supplementary section
-on simulation experiments).  It may be possible to develop integrated
-statistical models that consider both types of gene transfers, and to use
-them to more directly compare the rates with which they occur.  Another
-extension worth considering is direct modeling of incomplete lineage
-sorting (ILS), which is likely to be increasingly important as larger
-phylogenies are considered, especially given the large effective population
-sizes of many bacteria.  ILS has recently been integrated into models for
-duplication and loss (Rasmussen and Kellis) but has yet to be considered in
-reconciliation-based methods that also allow for gene transfer.  It is
-permitted by ClonalOrigin, but only in an indirect and approximate manner,
-through the introduction of recombination edges to ancestral branches
-(which can be misinterpreted as evidence of recombination).  The
-combination of improved models and richer data sets could allow for a much
-more detailed understanding of horizontal gene transfer.
+pan genome, as in this study.  However, this correlation is also likely to
+hold for true events.  True additive transfer events will be enriched in
+the dispensable genome, by definition, and it is likely that homologous
+recombination occurs at higher rates in the core genome (which is enriched
+for close homologs between species).  For all of these reasons, some
+apparent differences between the two types of transfer events are likely to
+reflect general differences between the core and dispensable portions of
+the genome.
+
+As sequence data becomes available for many more closely and distantly
+related organisms, it will become increasingly important to devise improved
+methods that consider both additive and replacing HGT.  As noted above, the
+current statistical models of recombination, such as ClonalOrigin, are
+effectively limited to considering orthologs, with one copy per species, in
+the core genome.  The phylogenetic reconciliation methods, on the other
+hand, fail to allow for multi-locus gene transfer events, and ignore
+information from branch lengths.  We have shown that it is possible, to a
+degree, to extend these methods to distinguish between additive and
+replacing transfers, but our methods are limited by their dependency on
+heuristic rules and parsimony assumptions.  It may be possible to develop
+integrated statistical models that consider both types of gene transfers,
+and produce unbiased estimates of the relative rates at which these events
+occur.  Another extension worth considering is direct modeling of
+incomplete lineage sorting (ILS), which is likely to be increasingly
+important as larger phylogenies are considered, especially given the large
+effective population sizes of many bacteria.  ILS has recently been
+integrated into models for duplication and loss \citep{Rasmussen2012} but
+has yet to be considered together with gene transfer.  
+%It is permitted by
+%ClonalOrigin, but only in an 
+%indirect and approximate manner, through the introduction of recombination
+%edges to ancestral branches (which can be misinterpreted as evidence of
+%recombination).  
+We expect that the combination of improved models and richer data sets
+will allow for a much more detailed understanding of the important
+process of horizontal gene transfer.
 
@@
 
-We thank Xavier Didelot for assistance with the ClonalOrigin analysis, and
-Haruo Suzuki for assistance with the data for SDD. This work was primarily
-supported by the National Institute of Allergy and Infectious Disease, US
-National Institutes of Health, under grant number AI073368-01A2 (to
-M.J.S. and A.S).  Additional support was provided by National Science
+This study was funded by the National Institute of Allergy and Infectious
+Disease, US National Institutes of Health, under grant number AI073368-01A2
+(to M.J.S.\ and A.S).  Additional support was provided by National Science
 Foundation CAREER Award DBI-0644111 and a David and Lucile Packard
-Fellowship for Science and Engineering (to A.S.).
+Fellowship for Science and Engineering (to A.S.).  We thank Xavier Didelot
+for assistance with ClonalOrigin and Haruo Suzuki for providing data for
+SDD.
 
@@
 %=============================================================================
-% Table 2. 
-\begin{table}[!ht]
-
-\caption{{\bf Functional category enrichments for additive gene
+\begin{table}[!h]
+
+\caption{{\bf Gene Ontology (GO) enrichments for additive gene
   transfers}}
@@
 \begin{flushleft}
-$^a$$P$-values based a Mann-Whitney $U$ test.  Shown are values that
-correspond to FDR$^b$  $<0.05$ based on the 
-Benjamini-Hochberg method (Methods).\\
-$^c$Number of genes in category.\\
+$^a$$P$-value based on a Mann-Whitney $U$ test (Methods).\\
+$^b$Corresponding false discovery rate estimated by the
+Benjamini-Hochberg method.  All categories having at least ten genes and
+$q\leq 0.05$ are displayed.\\ 
+$^c$Number of genes assigned to category.\\
 \end{flushleft}
@@
 \end{table}
-%\clearpage{}
+\clearpage{}
+
 %=============================================================================
 
+\section*{Figures}
+
 %=============================================================================
-% /Users/goshng/Documents/Projects/Mauve/output/cornellf/3/run-analysis/significant-1.txt
-% Table 1. 
-\begin{table}[!ht]
-\caption{
-{\bf Functional category enrichments for replacing gene transfers$^a$}} 
-\vspace{1ex}
-\noindent \begin{centering}
-\begin{tabular}{ccccl}
-\hline 
-$p^b$ & $q^c$ & Count$^d$ & GO term & Description\\
-\hline 
-2e-06 & 2e-04 &  38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
-3e-06 & 2e-04 &  33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
-5e-05 & 3e-03 &  22 & GO:0006633 & fatty acid biosynthetic process\\
-9e-05 & 4e-03 &  21 & GO:0008610 & lipid biosynthetic process\\
-3e-04 & 9e-03 &  45 & GO:0008643 & carbohydrate transport\\
-5e-04 & 1e-02 &  82 & GO:0006412 & translation\\
-6e-04 & 1e-02 &  13 & GO:0006814 & sodium ion transport\\
-1e-03 & 2e-02 &  40 & GO:0030529 & ribonucleoprotein complex\\
-1e-03 & 2e-02 &  12 & GO:0031402 & sodium ion binding\\
-1e-03 & 2e-02 & 244 & GO:0008152 & metabolic process\\
-2e-03 & 3e-02 & 378 & GO:0016740 & transferase activity\\
-2e-03 & 3e-02 &  39 & GO:0003735 & structural constituent of ribosome\\
-2e-03 & 3e-02 &  26 & GO:0019843 & rRNA binding\\
-4e-03 & 4e-02 &  66 & GO:0005840 & ribosome\\
-\hline 
-\end{tabular}
-\par\end{centering}
-\begin{flushleft}
-$^a$Based on recombination intensities restricted to topology-altering
-recombinant edges (Methods).\\
-$^b$$P$-values based a Mann-Whitney $U$ test.  Shown are values that
-correspond to FDR$^c$  $<0.05$ based on the 
-Benjamini-Hochberg method (Methods).\\
-$^d$Number of genes in category.\\
-\end{flushleft}
-\label{tab:functional}
-\end{table}
-\clearpage{}
-%=============================================================================
-
-\section*{Figures}
-
-%=============================================================================
-\begin{figure}[!ht]
+\begin{figure}[!h]
 \begin{center}
@@
   to appear similar to a replacing transfer.  In our analysis, we
-  conservatively require that at least one descendant species (here, {\bf
-    B}) contains genes that both descend from ($b_1$) and do not descend
+  conservatively require that at least one descendant species (here, 
+    $B$) contains genes that both descend from ($b_1$) and do not descend
   from ($b_2$) a lineage predicted to have been transferred when
@@
 %=============================================================================
-\begin{figure}[!ht]
+\begin{figure}%[!ht]
 \noindent \begin{centering}
@@
 \par\end{centering}
-\caption{
-{\bf Clonal frame inferred for the five genomes.}  The branch lengths are
-drawn to scale in the horizontal dimension.  Units are 
-expected substitutions per site given Watterson's estimate of the
-mutation rate. The branch length of the outgroup SEE is not scaled. 
-The labels for the ancestral nodes of the tree and the
-internal branches (SPY, SDE, and SD) are used throughout the paper.}
+\caption{ {\bf Clonal frame inferred for the five genomes.}  This phylogeny
+  was inferred using the ClonalFrame program \citep{Didelot2007}.  Branch
+  lengths are in units of expected substitutions per site and are drawn to
+  scale in the horizontal dimension.  The labels for the ancestral nodes of
+  the tree and the branches immediately ancestral to them (SPY, SDE, and
+  SD) are used throughout the paper.  The outgroup species, SEE, was not
+  part of the estimated clonal frame but was used in the parsimony-based
+  analysis.}
 \label{fig:tree5}
 \end{figure}
-\clearpage{}%
 %=============================================================================
@@
 % \includegraphics[scale=0.45]{figures/heatmap-recedge}
-\includegraphics[scale=0.5]{figures/heatmap}
+\begin{center}
+\includegraphics[scale=0.45]{figures/heatmap}
 \caption{\label{fig:Heatmap-of-transfers}
@@
 Each cell of the heat map represents the base-2 logarithm 
-of the ratio of the estimated number recombination events to its expectation
-under the prior, for the corresponding donor ($y$ axis) and recipient ($x$
+of the ratio of the estimated number recombination events to its prior
+expectation, for the corresponding donor ($y$ axis) and recipient ($x$
 axis) branches.  
-The black colored cells indicate prohibited transfer events.
+Cells in black indicate prohibited transfer events.
 %
@@
 average values across sampled recombinant
-trees. The prior considers the clonal frame with branch lengths and the
+trees. The prior considers the clonal frame with branch lengths (fig.\
+\ref{fig:tree5}) and the 
 global estimates of the three population parameters. Prohibited transfer events
 are ones for which the recipient branch is strictly older than the donor
-branch. Significance levels are denoted by asterisks: one for $<0.01$.
+branch. An asterisks indicates statistical significance ($p<0.01$).
 %
 (B) Additive transfers, as inferred by the parsimony-based
-approach. The plotted values represent
-numbers of inferred additive transfers across all gene families.
+approach. The plotted log ratios reflect
+total numbers of inferred additive transfers across all gene families.
 The prior considers the clonal frame (with branch lengths) and the total number
@@
 interval.
-branch. Significance levels are denoted by asterisks: one for $<0.01$, two for
-$<0.005$, and three for $<0.001$.}
+branch. Significance levels are denoted by asterisks: one for $p<0.01$, two for
+$p<0.005$, and three for $p<0.001$.  See {\bf Methods} for details.}
+\end{center}
 \end{figure}
@@
 %=============================================================================
-\begin{figure}[!ht]
+\begin{figure}%[!ht]
+\begin{center}
 \includegraphics[scale=0.5]{figures/ucsc}
@@
   genes highlighted in blue. The third track from top (in black) shows the
-  putatively non-recombining gene fragments analyzed by Mowgli.  The next series
+  putative non-recombining gene fragments analyzed by Mowgli.  The next series
   of tracks shows the results of the ClonalOrigin analysis of
@@
   recipient lineage in the phylogeny (fig.\ \ref{fig:tree5}) and describes
-  the posterior probabilities along the genome of recombination edges from
+  the posterior probabilities along the genome of recombinant edges from
   all possible donor lineages (shown in different colors; see key at
   right).  Here the putative virulence gene SpyM3\_0465, a dipeptidase,
-  shows strong evidence of a recombination edge from the SPY to the 
+  shows strong evidence of a recombinant edge from the SPY to the 
   SDE lineage, as well as some evidence of a SPY$\rightarrow$SPY2 edge.
@@
   bottom.  (B) The gene tree that is displayed after clicking on the
-  highlighted gene fragment.  The topology inferred by RAxML is consistent
+  highlighted gene fragment.  The tree estimated by RAxML is consistent
   with the combined influence of SPY$\rightarrow$SDE and
@@
 \label{fig:ucsc}
+\end{center}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=7in]{figures/strep-events}
-\caption{\label{fig:Gene-duplication-loss} 
-{\bf Distribution of inferred gene duplication, loss, and transfer events
-  for the six-species phylogeny.}  Gene counts for
-each extant and ancestral species are given at each node of the tree.  For
-each branch numbers of gene appearances (A*), duplications (D*), 
-losses (L*), additive transfers (T*), and replacing transfers (R*) are listed.
-Transfer events are recorded on the recipient branch (donors
-are not indicated).}
+\begin{center}
+\includegraphics[width=5in]{figures/strep-events}
+\caption{\label{fig:Gene-duplication-loss} {\bf Distribution of inferred
+    gene duplication, loss, and transfer events across the six-species
+    phylogeny.}  Numbers at nodes represent gene counts for extant species,
+  and estimated counts for ancestral species.  Numbers on branches indicate
+  inferred gene appearances (A*), duplications (D*), losses (L*), additive
+  transfers (T*), and replacing transfers (R*).  Transfer events are
+  recorded on recipient branches (donors are not indicated).}
+\end{center}
 \end{figure}
@@
 
-\clearpage{}\setcounter{figure}{0}
-\setcounter{table}{0}
-\renewcommand{\figurename}{Supplementary Figure}
-\renewcommand{\tablename}{Supplementary Table}
-
-\section*{Supplementary Tables}
-
-%=============================================================================
-\begin{table}[!ht]
-\caption{
-{\bf The six bacterial genomes.}
-Each bacterial genome is referred to as the name at the first column.}
-\noindent \begin{centering}
-\begin{tabular}{cccc}
-\hline 
-Name & NCBI accession & Size (base pairs) & Reference\tabularnewline
-\hline
-SDE1 & CP002215 & 2159491 & Suzuki et al. (2011)\tabularnewline
-SDE2 & NC\_012891 & 2106340 & Shimomura et al. (2011)\tabularnewline
-SDD & N/A$^a$ & 2141837 & Suzuki et al. (2011)\tabularnewline
-SPY1 & NC\_004070 & 1900521 & Beres et al. (2002)\tabularnewline
-SPY2 & NC\_008024 & 1937111 & Beres et al. (2006)\tabularnewline
-SEE & NC\_012471 & 2253793 & Holden et al. (2009)\tabularnewline
-\hline
-\end{tabular}
-\par\end{centering}
-\begin{flushleft}
-$^a$ SDD genome was not yet available from NCBI.
-\end{flushleft}
-\label{tab:genome}
-\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-\begin{table}[!ht]
-\caption{
-{\bf Posterior probability of gene tree topologies.}
-See Table S\ref{tab:genome} for the reference of the bacteria
-individual names.}
-\noindent \begin{centering}
-\begin{tabular}{cc}
-\hline 
-Gene tree topologies & Posterior probability\tabularnewline
-\hline
-(((SDE1,SDE2),SDD),(SPY1,SPY2)) & 66.7\%\tabularnewline
-(((SDE1,SDE2),(SPY1,SPY2)),SDD) & 9.3\%\tabularnewline
-((SDE1,SDE2),(SDD,(SPY1,SPY2))) & 3.5\%\tabularnewline
-(((SDE1,(SDE2,SDD)),(SPY1,SPY2)) & 1.5\%\tabularnewline
-(((SDE1,SDD),SDE2),(SPY1,SPY2)) & 1.3\%\tabularnewline
-((SDE1,SDD),(SDE2,(SPY1,SPY2))) & 1.2\%\tabularnewline
-((((SDE1,SDE2),SPY1),SPY2),SDD) & 1.0\%\tabularnewline
-((SDE2,SDD),(SDE1,(SPY1,SPY2))) & 1.0\%\tabularnewline
-\hline
-\end{tabular}
-\par\end{centering}
-\label{tab:Gene-tree-topologies}
-\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-\begin{table}[!ht]
-\caption{
-{\bf Ratios of logarithm base 2 of the observed
-recombinant edges to the expected number of recombinant edges.}  Figure
-\ref{fig:Heatmap-of-transfers}A shows the values graphically using heat map.
-NA's are for the branch pairs that are impossible under the model-based
-approached. Rows are for donor, and columns for recipient.}
-\noindent \centering{}\begin{tabular}{cccccccccc}
-\hline
-& SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2 \tabularnewline
-\hline
-SDE1& 0.176  &   NA  &1.170  &    NA & 0.7564  & NA & 0.67 &    NA & 0.663\tabularnewline
-SDE & -0.118 & 0.081 &-0.202 &     NA&  0.3395 &  NA& -0.15&  0.045& -0.323\tabularnewline
-SDE2& 1.201  &   NA  &0.092  &    NA & 0.6978  & NA & 0.60 &    NA & 0.387\tabularnewline
-SD  & -0.629 & 0.022 &-0.661 &0.00670&  0.0078 &  NA& -0.94& -0.299& -1.030\tabularnewline
-SDD & -0.317 & 0.371 &-0.197 &     NA&  0.0752 &  NA& -1.06& -0.525& -0.915\tabularnewline
-ROOT& -1.745 &-0.698 &-1.786 &0.23039& -0.4854 &  NA& -1.37&  0.497& -1.306\tabularnewline
-SPY1& 1.245  &1.883  &1.251  &    NA &-2.1997  & NA & 0.06 &    NA & 1.062\tabularnewline
-SPY & 0.089  &0.200  &0.298 &0.00098 &-0.7903  & NA &-0.19 &-0.058 &-0.159\tabularnewline
-SPY2& 1.174  &1.938  &1.287  &    NA &-1.9740  & NA & 1.07 &    NA & 0.063\tabularnewline
-\hline
-\end{tabular}
-\label{tab:heatmap}
-\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-\begin{table}[!ht]
-\caption{
-{\bf The observed number of recombinant edges.}
-The zeros are for the branch pairs that are impossible under the model-based
-approached. Rows are for donor, and columns for recipient.}
-\noindent \centering{}\begin{tabular}{cccccccccc}
-\hline
-& SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2 \tabularnewline
-\hline
-SDE1&8.896 &  0.00 &17.712  & 0.00 & 12.872   & 0 &12.59  &  0.00  &12.53\tabularnewline
-SDE &44.626& 136.49& 42.088 &  0.00& 227.061  &  0& 73.39 &  99.95 & 65.25\tabularnewline
-SDE2&18.099&   0.00&  8.393 &  0.00&  12.360  &  0& 11.97 &   0.00 & 10.36\tabularnewline
-SD  &9.671 &101.74 & 9.464  &31.26 &115.796   & 0 &13.68  & 98.38  &12.86\tabularnewline
-SDD &45.539& 168.58& 49.513 &  0.00& 194.706  &  0& 43.09 &  67.14 & 47.65\tabularnewline
-ROOT&43.553& 602.67& 42.337 &765.53& 802.813  &  0& 98.86 &2146.01 &103.60\tabularnewline
-SPY1&40.800&  11.46& 40.973 &  0.00&   4.371  &  0& 21.15 &   0.00 & 42.35\tabularnewline
-SPY &57.618& 262.08& 66.594 & 30.92& 163.518  &  0& 83.56 & 207.00 & 85.10\tabularnewline
-SPY2&38.837&  11.90& 42.000 &  0.00&   5.111  &  0& 42.45 &   0.00 & 21.19\tabularnewline
-\hline
-\end{tabular}
-\label{tab:obsheatmap}
-\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-\begin{table}[!ht]
-\caption{
-{\bf Number of genes showing significant signal of replacing transfer.}  For
-each type of recombinant edge (replacing gene transfer event in
-\texttt{ClonalOrigin} analysis), we record the number of genes with an average
-posterior probability greater than 0.6 for a recombinant edge of that type
-across their sites. Recombinant edges from the root branch are not considered
-here, and SPYclade and SDEclade each signify three branches in the clonal frame
-corresponding to the respective clades. Edge types not shown here have a zero
-count. The same is done for the subset of virulence genes. Absolute numbers are
-shown together with ratios relative to the total number of genes, as well as the
-total number of significant replacing transfers found in this analysis.}
-\noindent \centering{}\begin{tabular}{lll}
-\hline
-Recombinant edge type & Gene count & Virulence gene count \tabularnewline
-\hline
-SDEclade $\rightarrow$ SDD & 43 (33.1\%) & 5 (50\%) \tabularnewline
-SDEclade $\rightarrow$ SPYclade & 11 (8.5\%) & 1 (10\%) \tabularnewline
-SDD $\rightarrow$ SDEclade & 1 (0.8\%)  & 0 (0\%) \tabularnewline
-SPYclade $\rightarrow$ SDEclade & 59 (45.4\%) & 3 (30\%) \tabularnewline
-SPYclade $\rightarrow$ SPYclade & 16 (12.3\%) & 1 (10\%) \tabularnewline
-All types & 130  (100\%) & 10   ( 100\%)\tabularnewline
-% &  (11.8\% of total number of genes) & (13\% of virulence genes)\tabularnewline
-\hline
-\end{tabular}
-\label{tab:gene-counts-replacing}
-\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-\begin{table}[!ht]
-\caption{
-{\bf Number of inferred transfer events of different types.}  We record the number of inferred additive transfer events of each type.
-SPYclade and SDEclade each signify three branches in the clonal frame corresponding to the respective clades.
-Edge types not shown here have a zero count.
-The same is done for the subset of virulence genes.
-Absolute numbers are shown together with ratios relative to the total number of genes, as well as the total number of significant replacing transfers found in this analysis.}
-\noindent \centering{}\begin{tabular}{lll}
-\hline
-Recombinant edge type & Gene count & Virulence gene count \tabularnewline
-\hline
-SDEclade $\rightarrow$ SDEclade & 28 (10.6\%) &  5 (12.5\%)\tabularnewline
-SDEclade $\rightarrow$ SDD & 64 (24.2\%) & 2 (5.0\%)\tabularnewline
-SDEclade $\rightarrow$ SPYclade & 28 (10.6\%) &  7 (17.5\%)\tabularnewline
-SDD $\rightarrow$ SDEclade & 29 (10.9\%) &  5 (12.5\%)\tabularnewline
-SDD $\rightarrow$ SPYclade & 17 (6.4\%) &  2 (5.0\%)\tabularnewline
-SD $\rightarrow$  SPYclade &  4 (1.5\%) &  0 (0.0\%)\tabularnewline
-SPYclade $\rightarrow$ SDEclade & 32 (12.1\%) & 10 (25.0\%)\tabularnewline
-SPYclade $\rightarrow$ SDD & 33 (12.5\%) &  3 (7.5\%)\tabularnewline
-SPYclade $\rightarrow$ SD &   3 (1.1\%) &  0 (0.0\%)\tabularnewline
-SPYclade $\rightarrow$ SPYclade & 27 (10.2\%) &  6 (15.0\%)\tabularnewline
-All types  &  265 (100\%) & 40 (100\%)  \tabularnewline
-\hline
-\end{tabular}
-\label{tab:gene-counts-additive}
-\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-\begin{table}[!ht]
-\caption{
-{\bf Functional categories enriched in 
-recombining transfer events that are inferred using the parsimony-based
-approach.}}
-\noindent \begin{centering}
-\begin{tabular}{ccccl}
-\hline 
-$p^a$ & $q^b$ & Count$^c$ & GO term & Description \\
-\hline 
-7e-09 & 3e-06 & 153 & GO:0006412 & translation\\
-1e-08 & 3e-06 & 67 & GO:0030529 & ribonucleoprotein complex\\
-1e-07 & 2e-05 & 45 & GO:0019843 & rRNA binding\\
-2e-06 & 2e-04 & 73 & GO:0003735 & structural constituent of ribosome\\
-3e-05 & 3e-03 & 157 & GO:0005840 & ribosome\\
-7e-05 & 5e-03 & 221 & GO:0003723 & RNA binding\\
-2e-04 & 1e-02 & 11 & GO:0004826 & phenylalanine-tRNA ligase activity\\
-3e-04 & 2e-02 & 14 & GO:0015413 & nickel-transporting ATPase activity\\
-4e-04 & 2e-02 & 93 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
-5e-04 & 2e-02 & 104 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
-5e-04 & 2e-02 & 13 & GO:0006265 & DNA topological change\\
-5e-04 & 2e-02 & 13 & GO:0003916 & DNA topoisomerase activity\\
-5e-04 & 2e-02 & 733 & GO:0005737 & cytoplasm\\
-6e-04 & 2e-02 & 10 & GO:0046873 & metal ion transmembrane transporter activity\\
-7e-04 & 2e-02 & 12 & GO:0009435 & NAD biosynthetic process\\
-8e-04 & 2e-02 & 19 & GO:0016820 & hydrolase activity, acting on acid anhydrides, \\
-      &       &    &       &catalyzing transmembrane movement of substances \\
-1e-03 & 3e-02 & 20 & GO:0005694 & chromosome\\
-2e-03 & 4e-02 & 31 & GO:0006814 & sodium ion transport\\
-\hline 
-\end{tabular}
-\par\end{centering}
-\begin{flushleft}
-$^a$$P$-values based a Mann-Whitney $U$ test.  Shown are values that
-correspond to FDR$^b$  $<0.05$ based on the 
-Benjamini-Hochberg method (Methods).\\
-$^c$Number of genes in category.\\
-\end{flushleft}
-\label{tab:go-events-recombining}
-\end{table}
-\clearpage{}
-%=============================================================================
-
-%=============================================================================
-\begin{table}[!ht]
-\caption{
-{\bf Three population parameter estimates from the model-based approach.}
-The estimates of the three parameters were based on one of two independent
-MCMC chains.  We considered the similar 
-results of the two independent chains as confirmation of 
-the convergence of the first stage of MCMC.}
-\noindent \begin{centering}
-\begin{tabular}{ccc}
-\hline
-& Replicate 1 & Replicate 2\tabularnewline
-Parameter & Estimate (IQR$^a$) & Estimate (IQR)\tabularnewline
-\hline
-$\theta^b$ & 0.081 (0.067, 0.094) & 0.081 (0.067, 0.099)\tabularnewline
-$\rho^c$ & 0.012 (0.006, 0.019) & 0.012 (0.006, 0.020)\tabularnewline
-$\delta^d$ & 744 (346, 2848) & 723 (348, 2870)\tabularnewline
-\hline
-\end{tabular}
-\par\end{centering}
-\begin{flushleft}
-$^a$ IQR stands for interquartile range.
-
-$^b$ The median of mutation rate per site weighted over the lengths of blocks.
-
-$^c$ The median of recombination rate per site weighted over the lengths of blocks.
-
-$^d$ The median value for the recombinant tract length weighted over the lengths of blocks.
-\end{flushleft}
-\label{tab:three}
-\end{table}
-%\clearpage{}
-%=============================================================================
-
-%=============================================================================
-\begin{table}[!ht]
-\caption{
-{\bf Simulation study of the three population parameters.}}
-\noindent \begin{centering}
-\begin{tabular}{ccc}
-\hline
-Parameter & True & Estimates (Standard Deviation)\tabularnewline
-\hline
-$\theta^a$ & 0.081 & 0.081 (0.0011)\tabularnewline
-$\rho^b$   & 0.012 & 0.011 (0.0003)\tabularnewline
-$\delta^c$ & 744 bp & 919 bp (28)\tabularnewline
-\hline
-\end{tabular}
-\par\end{centering}
-\begin{flushleft}
-$^a$ Mutation rate per site.
-
-$^b$ Recombination rate per site.
-
-$^c$ Recombinant tract length.
-\end{flushleft}
-\label{tab:sim-three}
-\end{table}
-%=============================================================================
-
-\section*{Supplementary Figures}
-
-%=============================================================================
-\begin{sidewaysfigure}[!ht]
-\begin{center}
-\includegraphics[scale=0.3]{figures/mauve}
-\end{center}
-\caption{\label{fig:mauve}
-{\bf The genome alignment of the five genomes.} The top
-first and second rows are SDE1 and SDE2 genomes, respectively. The third is SDD
-genome. The fourth and fifth rows are SPY1 and SPY2 genomes, respectively.}
-\end{sidewaysfigure}
-%=============================================================================
-
-%=============================================================================
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[width=5in]{figures/famsizes}
-\end{center}
-\caption{{\bf Distribution of gene family sizes.}}
-\label{fig:famsizes}
-\end{figure}
-\clearpage{}
-%=============================================================================
-
-%=============================================================================
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-theta}
-
-\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-rho}
-
-\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-delta}
-\end{center}
-\caption{
-{\bf Three scatter plots of mutation rate, recombination
-rate, and recombinant tract length for the blocks.} The plus signs are mean values
-for each block. The density of each value for each block is depicted by black
-clouds.  The red dashed lines are the global block-length weighted median of the
-three parameter estimates.}\label{fig:scatter3}
-\end{figure}
-\clearpage{}%
-%=============================================================================
-
-%=============================================================================
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[width=5in]{figures/compareMowgliCoRecombining}
-\end{center}
-\caption{
-{\bf Plot of recombination intensity against number
-of replacing gene transfer.}}
-\label{fig:cmpcomowgli}
-\end{figure}
-\clearpage{}
-%=============================================================================
-
-%=============================================================================
-\begin{figure}[!ht]
-\includegraphics[scale=0.42]{figures/clonalorigin}
-\caption{
-{\bf Recombinant tree in the ClonalOrigin model.}
-(A) A recombinant tree relating the five genomes with four recombinant edges.
-(B) A genomic region of size 100 base pairs with subregions affected by the
-four recombinant edges. Genomic positions are evenly marked by ticks every 10-th
-base pair.  For each edge type and each site, we recorded the sampling frequency
-of a recombinant edge of the type at that site.  The posterior probability that
-a site experiences influx from a donor branch using a single recombinant tree.
-Summing these posteriors over a certain set of edge types, we obtain a measure
-of recombination intensity along the genome.  Each segment from b to g are
-scored by the number of recombinant edges per site.  (C-G) Local gene trees
-induced by the recombinant tree are illustrated for each region covered by
-combinations of recombinant edges along the genomic region. For example, the
-local gene tree (E) is created by three recombinant edges in the subregion
-between 51 base pairs and 70 base pairs. The local gene tree (C) is the same as
-the species tree (A) in their topology.  Because there are 105 possible rooted
-labeled topologies relating five taxa, we could count topologies.}
-\label{fig:clonalorigin}
-\end{figure}
-\clearpage{}%
-%=============================================================================
-
-%=============================================================================
-\begin{sidewaysfigure}
-
-\includegraphics[scale=0.6]{figures/sim2-1}
-
-\includegraphics[scale=0.6]{figures/sim2-2}
-
-\caption{
-{\bf Simulations of the model-based approach for assessing the power of
-recovering recombinant edges using prior recombinant trees.} 
-At X-axis labels are shown pairs of species tree branches, for each of which one
-interval with an open circle and a closed circle next to it
-are displayed.  Intervals with open circle are for 
-the ratio of the number of recombinant edges inferred from the 100
-simulated data sets under the prior with respect to the expected number of
-recombinant edges under the prior.  Closed circles are for 
-the ratio of the number of recombinant edges estimated
-from the real data with respect to the expected number of recombinant edges
-under the prior.  The horizontal lines are at the zero.  The intervals range
-from 5\% quantile to 95\% of quantile.}
-\label{fig:sim2}
-\end{sidewaysfigure}
-\clearpage{}%
-%=============================================================================
-
-%=============================================================================
-\begin{sidewaysfigure}
-
-\includegraphics[scale=0.6]{figures/sim3-1}
-
-\includegraphics[scale=0.6]{figures/sim3-2}
-
-\caption{
-{\bf Simulations of the model-based approach for assessing the power of
-recovering recombinant edges using posterior recombinant trees.} 
-At X-axis labels are shown pairs of species tree branches, for each of which one
-interval with an open circle and an closed circle next to it
-are displayed.  Intervals with open circle are for 
-the ratio of the number of recombinant edges inferred from the 100
-simulated data sets with the posterior samples of recombinant trees 
-with respect to the expected number of
-recombinant edges under the prior.  Closed circles are for 
-the ratio of the number of recombinant edges estimated
-from the real data with respect to the expected number of recombinant edges
-under the prior.  The horizontal lines are at the zero.  The intervals range
-from 5\% quantile to 95\% of quantile.}
-\label{fig:sim3}
-\end{sidewaysfigure}
-\clearpage{}%
-%=============================================================================
-
-%=============================================================================
-\begin{sidewaysfigure}
-
-% \includegraphics[width=6in]{figures/real_vs_null_additive.pdf}
-\includegraphics[scale=0.6]{figures/sim4-1}
-
-% \includegraphics[width=4in]{figures/real_vs_null_dup.pdf}
-\includegraphics[scale=0.6]{figures/sim4-2}
-
-\caption{
-{\bf The power of recovering additive transfers and duplications using simulations.} (top) For each pairs of species tree branches and clades SDE and SPY,
-we plot the number of reconstructed additive transfers in the real data 
-(circle) normalized
-by the true number of transfers in our simulations.  Additive transfers
-reconstructed from simulations are also normalized by the true number
-and are plotted with 90\% confidence intervals.
-(bottom) For each branch in the species tree, we plot the number of 
-reconstructed duplications in the real data (circles) normalized by the true
-number of duplications in our simulations.  Duplications
-reconstructed from simulations are also normalized by the true number
-and are plotted with 90\% confidence intervals.
-}
-\label{fig:mowgli-sim}
-\end{sidewaysfigure}
-\clearpage{}%
-%=============================================================================
-
-%=============================================================================
-\begin{figure}
-\begin{center}
-\includegraphics[width=5in]{figures/calling}
-\end{center}
-\caption{
-{\bf Distinguishing between replacing and additive horizontal gene transfers.}
-We consider a transfer ``additive'' if there exists an extant descendant species
-(e.g. C and D) that contains both descendants (bold) and
-non-descendants (non-bold) of the transferred gene (black circle).}
-\label{fig:calling-transfers}
-\end{figure}
-\clearpage{}%
-%=============================================================================
-
-\clearpage{}
-\section*{Supplementary Material}
-
-% \subsection*{Preliminary ClonalFrame and ClonalOrigin analysis}
-% 
-% {\bf IG: we should write here details about the first part of the analysis: data
-% preparation, estimating species tree, and three population parameters. TBD. For
-% now we have a collection of details extracted from previous version of
-% results.}
-% \texttt{SC: Do we need this section because the method seems to describe everything enough
-% in detail?}
-% 
-% A genome alignment of the five genomes showed two
-% distinct genome structures of SPY and SDE (Figure S\ref{fig:mauve}).
-% 
-% ClonalFrame estimated a relative strength of
-% recombination compared to mutations as $r/m=4.66$ meaning that four times more
-% sites had been affected by recombination-driven substitutions than by
-% mutation-driven substitutions (Table S\ref{tab:clonalframe}).  This indicated
-% potentially significant roles of recombination in the evolutionary process of
-% shaping these genomes.  Estimates of population mutation rate ($\theta$),
-% recombination rate ($\rho$), and recombinant tract length ($\delta$) were obtained
-% for the alignment blocks analyzed (Figure S\ref{fig:scatter3}), and a median of
-% the estimates weighted over block lengths was obtained as a global estimate for
-% each of the three parameters (Table S\ref{tab:three}).  The local estimates
-% typically seemed to be moderately fluctuating around the global estimate, with
-% the exception of what appeared to be a recombination hotspot at a region 200
-% kilobase pairs from the origin of replication of SDE1.
-
-\subsection*{Summarization of recombinant tree samples}
-We describe the four summarizing procedures including tree topology along the
-genome, numbers of recombinant edges, probability of recombination, and
-recombination intensity using supplementary fig.\ S\ref{fig:clonalorigin}.  
-Supplementary figures S\ref{fig:clonalorigin}A 
-and S\ref{fig:clonalorigin}B fully specify a
-recombinant tree.  First, the recombinant tree induces  five gene tree
-topologies along the genome of size being 100 bp. The fraction of gene tree
-topology of supplementary figure S\ref{fig:clonalorigin}C is 0.15, and 
-that of supplementary figure S\ref{fig:clonalorigin}G is 0.05. 
-The fractions of the other three gene tree
-topologies of 
-supplementary figures S\ref{fig:clonalorigin}D, S\ref{fig:clonalorigin}E, 
-and S\ref{fig:clonalorigin}F are all 0.2.  Second, the heatmap values are computed
-by counting edges for each of 60 possible types and dividing the number of edges
-by prior expected number of recombinant edges given the species tree and the
-three global population parameters.  One edge for each of the four types of
-recombinant edges is observed in the recombinant tree of Figure
-S\ref{fig:clonalorigin}A.  Third, the probability of recombination at a site for
-a particular recombinant edge type is the fraction of recombinant trees, in
-which the recombinant edge spans the site.  
-Supplementary figure S\ref{fig:clonalorigin}B
-shows the probability of recombination along the 100 bp genomic positions.
-Branch SDE1 received the genomic segments ranging from 11 bp to 100 bp from
-branch SDE.  Similarly, branch SDE2 received the segments ranging from 31 bp to
-70 bp from branch SPY1.  Each of recombinant trees from a posterior sample would
-induce a map of recombination such as 
-supplementary figure S\ref{fig:clonalorigin}B.
-Recombination maps from samples of recombinant trees can be averaged as shown in
-figure \ref{fig:ucsc}.  
-Fourth, the recombination intensity of a site is the
-number of recombinant edge types at the site 
-(supplementary fig.\ S\ref{fig:clonalorigin}B).
-Note that it is not the number of recombinant edges.
-
-The recombination intensity allowed us to find genomic regions enriched for
-recombination events in general, or regions enriched for recombination events of
-certain types (e.g., between SPY and SDE).  
-Our simulation study indicated that edges of
-certain types were better indicators of recombination events than others.  For
-instance, recombinant edges which resulted in a recombinant tree with the same
-topology as the species tree appeared to be more prone to spurious sampling.
-Such edges were ones where the donor was the same branch, sister branch, or
-parent branch of the recipient branch.  Therefore, in some cases we elected to
-remove such edges from consideration when measuring recombination intensity.  
-% The tree topologies would show more effectively regions where
-% recombination occurs along the genome (see subsection of \textbf{UCSC Genome
-% Browser}).
-
-\subsection*{Comparison of model-based and parsimony-based inferences on
-replacing gene transfer}
-
-The similarity between the two lists of gene categories found to be associated
-with replacing gene transfers in the separate analyses -- model-based and
-parsimony-based --  indirectly indicated concordance between the two independent
-approaches.  In order to obtain a more direct measure of correlation, we used
-the set of genes present in both analyses to compute the correlation between
-recombination intensity scores and the number of replacing gene transfers
-inferred by the parsimony-based approach. We observed a positive correlation of
-$0.39$ (P-value $2.2\times10^{-16}$; supplementary fig.\ S\ref{fig:cmpcomowgli})
-between the
-two measures.  For comparison, the correlation between recombination intensity
-and the number of additive gene transfers was 0.03 (P-value 0.18).
-%Figure S\ref{fig:cmpcomowglihgt}).
-
-\subsection*{Simulation Studies for the model-based and parsimony-based approaches}
-We performed four sets of simulations.  In the first simulation, we 
-showed that ClonalOrigin could infer population mutation rate, tract
-length, and recombination rate with reasonable accuracy from our data sets of
-the five individual bacteria. In the second
-simulation, we checked whether recombinant edges were recovered from
-simulated data with recombinant trees generated from the prior under the
-parameters estimated from our data set. The third
-simulation was conducted using simulated data with recombinant trees sampled
-from the
-posterior distribution.  We used a simulation feature of ClonalOrigin
-to simulate the process of mutation and recombination across the five species.
-In the fourth simulation, we checked if the parsimony-based approach worked with
-our data set.
-
-\textbf{Simulation 1 - three population parameters.}
-We describe a procedure of simulations for checking if the three
-population parameters were accurately recovered using the model-based approach.
-Ten recombinant trees were sampled given the three parameters (table
-S\ref{tab:three}) and the species tree (fig.\ \ref{fig:tree5}) set to the
-values inferred from the five bacterial genomes.  Each of the recombinant trees
-was used to simulate the Jukes-Cantor model for DNA sequence evolution to create
-274 blocks of alignments.  The block lengths configuration was the same as that
-of the blocks used in the analysis of the five genomes.
-
-% s15
-% FIXME: Newick tree in text in a file.
-% FIXME: Publish the result and the data set with the code.
-
-\textbf{Simulation 2 - numbers of recombinant edges from prior.}
-Because the number of statistics increases compared with Simulation 1, we
-increased the sample size from 10 to 100.  The recombinant trees were created
-from the prior distribution by adding Recombinant edges to the species tree with
-no particular directions of replacing gene transfers between species tree
-branches.  We computed the mean number of recombinant edges of each type from
-the inference with each of the 100 data sets. Dividing the mean number by the
-prior expected number of recombinant edges we obtained the heatmap value.  The
-distribution of the 100 heatmap values in the scale of logarithm to base 2 is
-shown at supplementary figure S\ref{fig:sim2}.  
-We compared this distribution from simulation
-with the heatmap values from the real data analysis.  We counted recombinant
-edges for the recombinant tree from each iteration of the posterior sample, and
-divided the number of edges by the prior expected number of edges. This allowed
-us to measure variation among recombinant trees in the posterior sample.  The
-distribution of the 1001 heatmap values in the scale of logarithm to base 2 is
-also shown next the simulation result at supplementary figure S\ref{fig:sim2}.
-
-% The following was used for checking recombination intensity of genes. We
-% decided not to describe it.
-% 
-% We first sampled ten recombinant trees given the mutation rate, recombination
-% rate, and tract length (table S\ref{tab:three}) and the species tree (Figure
-% \ref{fig:tree5}), and then created ten independent simulated data sets with
-% the Jukes-Cantor model for DNA sequence evolution using each of the recombinant
-% trees. This consisted of 100 replicates of data sets.  The recombinant trees
-% were created from the prior distribution and recombinant edges were added to the
-% species tree with no particular directions of replacing gene transfers between
-% species tree branches.  We computed the fraction of the mean number of recombinant
-% edges for a pair of donor and recipient species tree branches over the simulated
-% 10 recombinant trees against the median of recombinant edges estimated from the
-% 10 simulated data sets, and then took the logarithm of the fraction to base 2. We
-% also computed 5\% and 95\% quantiles to find the 90\% quantile range for the log
-% base 2 fraction.  We expect that the log base 2 fraction would near zero. We also
-% wished to see if how the recombinant edge count estimates from the real data set
-% deviate from the prior. Instead of using the number of recombinant edges from
-% the ten recombinant trees, we used the numbers of recombinant edges that were
-% estimated from the real data. 
-% 
-% We tested if our measure of recombination intensity provided a reliable estimate
-% of the true intensity. We artificially annotated our synthetic blocks to genes
-% according to the real annotations from SPY1 genome.  For each gene in each of
-% the simulated 100 data sets, we plotted the average estimated recombination
-% intensity along that gene against the average number of different recombinant
-% edge types from the ten simulated recombinant tree.
-
-% s16
-
-% s16
-% FIXME 1: IG: RESULTS FROM THIS POINT ON REQUIRE SOME REFINEMENT. SC: Let's see.
-% FIXME 2: ``TRUE" TO BE EXPECTED VALUE
-% RATHER THAN AVERAGE OVER THE 10 REC.  TREES. SC: OKAY. LET'S FIND THE PRIOR
-% EXPECTED NUMBER OF EVENTS. THIS MUST BE GIVEN SOMEWHERE}.  
-% FIXME 3: CONSIDER SHOWING DETAILED COMPARISON OF ESTIMATED VALUES TO ONES 
-% EXPECTED FROM THE ACTUAL REC. TREES. THESE DON'T LOOK AS NICE, BUT THEY MIGHT 
-% BE ABLE TO DEMONSTRATE OUR ABILITY TO DETECT DEPARTURE FROM EXPECTATION. SC:
-% THIS MAY BE ABOUT CHANGING AVERAGE TO PRIOR EXPECTED NUMBER OF EVENTS.
-% FIXME 4: WE SHOULD PROBABLY CONSIDER ALTERNATIVE WAYS OF USING THESE RESULTS TO 
-% ARGUE THAT OUR DIRECTIONAL SIGNAL IS REAL. SC: I MIGHT CONSIDER SIMULATIONS AND
-% ARTIFICIALLY DOUBLE OR TRIPLE THE RECOMBINATION EVENTS OF ONE OF TYPES.
-
-\textbf{Simulation 3 - number of recombinant edges from posterior.}
-We used 100 recombinant trees sampled from the posterior distribution
-of the recombiant tree instead of using simulated recombinant trees.
-We created one simulated data set for each of the 100 recombinant trees with
-the Jukes-Cantor model for DNA sequence evolution 
-using the mutation rate estimate (supplementary table S\ref{tab:three}).
-This consisted of 100 replicates of data sets.  
-We follow the same analysis as in the simulation 2 
-(supplementary fig.\ S\ref{fig:sim3}).
-
-% s17
-
-\textbf{Simulation 4 - Simulation of additive horizontal gene transfers.}
-We also performed a simulation analysis to understand the power of 
-the parsimony-based pipeline to recover events such as gene duplication 
-and additive transfer.  The simulations were also used as a null model
-for studying variations in the rate and biases in the direction of
-additive transfers.
-
-We implemented a new program that simulates all four events:
-duplication, loss, additive transfer, and replacing transfer.  The
-generative process begins with a single gene lineage located at the
-root of the species tree.  The gene lineage then grows down a species
-tree branch until one of the following events occur.
-
-\begin{itemize}
-\item If a species tree node is encountered (the root node is the
-  first encountered), the gene lineage undergoes {\em speciation} and
-  bifurcates into two lineages, each one evolving independently down a
-  child branch of the species tree node.  
-\item Using a Poisson process at constant rate $D$, a gene lineage
-  will undergo {\em duplication} and bifurcate into two lineages, both
-  growing independently within the same species branch.  
-\item At a constant rate $L$, a gene lineage can be {\em lost} and
-  terminate.  
-\item At a constant rate $T$, a gene lineage can undergo {\em additive
-  transfer}, bifurcating into two lineages, one lineage within
-  the current species branch and the second lineage within one
-  of the other contemporary species branches chosen uniformly.  Both
-  gene lineages continue to evolve independently of one another.  
-\item At a constant rate $R$, a gene lineage $a$ can undergo {\em replacing
-  transfer}.  From the other contemporary species branches, a second
-  gene lineage $b$ (if it exists) is chosen to be replaced.  The gene
-  lineage $b$ is terminated and lineage $a$ is bifurcated, one lineage $a_1$
-  remaining in the current species and the other $a_2$ transferring to the
-  species branch of lineage $b$.  Both lineages $a_1$ and $a_2$ continue
-  to evolve independently.
-\item Lastly, if a gene lineage encounters a species tree leaf, the
-  gene lineage terminats and is denoted an {\em extant gene copy}.
-\end{itemize}
-
-As a final step, any gene lineage that terminates before the species tree
-leaves, is declared extinct and is pruned from the gene tree.
-
-Given a gene tree generated by this process, we simulate sequence
-evolution using a continuous-time Markov process (e.g. HKY)
-parameterized to match the base frequency and transition/transversion
-ratio observed in the real data.  This results in a gapless simulated
-gene alignment.
-
-In our simulation analysis, we used our simulation program within a
-larger analysis pipeline that followed the same procedure we have used
-to analyzed the real dataset, thus giving an accurate view of the
-power of the entire pipeline.  In our simulation pipeline, we
-simulated 10,000 gene alignments.  For each alignment we reconstructed 
-gene trees using RAxML, reconciled them using Mowgli, and classified
-additive and replacing gene transfers using the procedure described in the
-main text.
-
-To create a genome-wide sample of gene trees we randomly chose 2314
-gene trees (the same number of real families) from the 10,000
-reconstructed.  To study the distribution of this tree set, we
-resampled with replacement 1000 times (This resampling strategy was
-chosen for computational efficiency).  This gave us the following
-samples: (1) a distribution of the ``true'' simulated gene trees along
-with their event counts, (2) a distribution of the reconstructed
-simulated gene trees and event counts, and (3) the original set of
-gene trees reconstructed from real data.  Differences between datasets
-(1) and (2) indicate the power of our pipeline to accurately infer
-each of the event types, and differences between datasets (2) and (3)
-indicate unusual deviations of the real data from the null model
-(i.e. variations in event rates or biases in transfer direction).
-
-\subsection*{Simulation-based support for biased directionality
-in replacing gene transfer between SPY and SDE}
-
-We conducted a series of experiments on the simulated data to check the reliability
-of estimates obtained by the model-based approach. We found that the population
-parameters obtained in the first phase were reliable, with the exception that
-the recombinant tract length was over-estimated as this was observed
-by \citealp{Didelot2010} (Simulation 1). Our experiments also showed that the recombination
-intensity used in our functional category association tests had fairly good
-correlation between the true value and the simulated ones (P-value of 0.79).  
-% (Figure S\ref{fig:ri1}). 
-One thing we specifically wished to test in these experiments
-was whether the model-based approach reliably detected deviations of recombinant
-edge counts from their expectation, as observed in the heatmap of Figure
-\ref{fig:Heatmap-of-transfers}. Our major concern was that these deviations
-were not directly accounted for by the probabilistic model of ClonalOrigin, but
-rather indirectly observed by summary of the posterior sample of recombinant
-trees.
-
-We simulated the model to generate a hundred simulated data sets using a
-collection of recombinant trees sampled from the prior distribution defined by
-the species tree and three population parameters, as estimated from our genomic
-data (Simulation 2).  We then repeated the inference procedure of the second stage of
-ClonalOrigin on each data set and, as in the data analysis, and counted
-for each donor-recipient branch pair the average number of recombinant edges.
-By normalizing the counts by the prior expected count we obtained intervals of
-base 2 logarithms of the ratio of the observed number of recombinant edges to
-the prior expectionation (supplementary fig.\ S\ref{fig:sim2}). 
-The comparison showed a good
-concordance of the estimated counts and the prior expected counts for 
-the topology-altering edge types.
-For each edge type, we also plotted in supplementary figure S\ref{fig:sim3} the
-log-ratio of the count estimated for it in our data analysis and the expected
-count; these are the values making up
-the heatmap of Figure \ref{fig:Heatmap-of-transfers}. Comparing the counts
-estimated in the data analysis and the counts estimated from the simulated data
-allowed to evaluate the significance of the divergence from expectation observed
-in the heatmap.  Recall the heatmap indicated an excess of recombinant edges
-between the SPY and SDE clades in both directions. Our simulations indicated
-that the excess observed in the direction SDE to SPY was much less significant,
-as the counts estimated from real data often fell very close to the
-corresponding counts estimated from simulated data. The excess observed in the
-direction SPY to SDE appeared not only to be larger, but also more significant.
-This further supported our claim for the biased directionality in replacing gene
-transfer between these two species.
-
-We used simulated data (Simulation 3) to test not only the ability of the ClonalOrigin
-sample to \textit{detect} deviations from expected counts of recombinant edges, but
-also its ability to reliably \textit{quantify} this deviation.  For this purpose we
-generated a hundred simulated data sets using a collection of recombinant trees
-sampled from the posterior sample of ClonalOrigin on the genomic data.  The
-difference between the two simulated batches was that in the first one, the
-recombinant trees used represent the prior distribution, and in the second one,
-they represented the posterior distribution, which we knew deviated from the
-prior. Again, we recorded the recombinant edge counts obtained by ClonalOrigin
-and normalized these counts by the prior expected counts (intervals with closed
-circles in supplementary figure S\ref{fig:sim3}). We expected these counts to be
-in better agreement with
-the counts observed in the data analysis. For recombinant edges that showed a
-defficiency in our data analysis (e.g., from the SPY clade to SDD), the counts
-appeared to fit well with the count estimated from genomic data. However, for
-recombinant edges that showed an excess in our data analysis (e.g., from the SPY
-clade to the SDE clade), the counts appeared to be significantly influenced by the
-prior. We noted that the influence of the prior appeared to be even stronger for
-topology-preserving recombinant edges. All things considered, we found that defficiency of
-a certain recombinant edge type was reliably estimated whereas excesses were 
-typically underestimated. Therefore, we concluded that the actual extent to which
-replacing transfers between SPY and SDE \textit{preferred} the direction from SPY to
-SDE was possibly somewhat underestimated in our analysis.
-
 \end{document}

### 2011-12-13 doc/supplementary.tex
A supplementary is added.

\documentclass[12pt]{article}

\usepackage{times}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
%\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
\usepackage{setspace}

% To rotate tables using sidewaystable instead of just table.
\usepackage{rotating}

% FIXME: Remove this package when submitting the manuscript.
\usepackage{url}
\usepackage{xr}

\usepackage{natbib}
\usepackage{../../latex/sty/genres}

\externaldocument{manuscript}
% Text layout
\oddsidemargin 0in
\evensidemargin 0in
\topmargin -.5in
\textwidth 6.5in
\textheight 9in
%\topmargin 0.0cm
%\oddsidemargin 0.5cm
%\evensidemargin 0.5cm
%\textwidth 16cm 
%\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}

\providecommand{\\}{\\}
\newcommand{\lyxdot}{.}
\newcommand{\comment}[1]{}

\makeatother

\renewcommand{\thefootnote}{\textit{\alph{footnote}}}

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **
\markright{Horizontal gene transfer in Streptococcus}

\begin{document}

\setcounter{secnumdepth}{-1}

\noindent \begin{LARGE}{\bf Replacing and additive horizontal 
    \\[0.7ex] 
gene transfer in {\em Streptococcus}}\end{LARGE}\\[3ex]
{\bf Sang Chul Choi, Matthew D.\ Rasmussen, Melissa J.\ Hubisz, \\
Ilan Gronau, Michael J.\ Stanhope, and Adam Siepel}\\[0.5ex]

\tableofcontents

\listoftables

\listoffigures

\doublespacing

\section{Supplementary Text}

\subsection{Recombinant tree summaries}

We use a recombinant tree portrayed in supplementary figure
\ref{fig:clonalorigin} to demonstrate the various summaries we use in our
analysis of recombinant trees sampled by ClonalOrigin. This recombinant tree is
augmented by four recombinant edges denoted e1 to e4 of four different edge
types (supplementary fig.\ \ref{fig:clonalorigin}A): SDE-to-SDE1, SPY1-to-SDE2,
SPY-to-SDD, and ROOT-to-SPY2. Edges e2, e3 and e4 are topology-altering ones
whereas edge e1 is topology-preserving. Each recombinant edge is associated with
a contiguous genomic segment  (supplementary fig.\ \ref{fig:clonalorigin}B),
inducing a series of six different gene trees, T1 through T6, along the genome
(supplementary fig.\ \ref{fig:clonalorigin}C).  Gene trees T1 and T2 are of the
same tree topology, but the other trees are of distinct topologies with the
following genomic frequencies: 0.3 (T1 and T2), 0.2 (T3, T4, and T5), and 0.1
(T6).  The recombination
intensity at a given site is defined by the number of recombinant edge types
present at that site; it can be computed from the map of recombinations along
the genome (bottom of supplementary fig.\ \ref{fig:clonalorigin}B). In this
example the recombination intensity ends up being the number of recombinant
edges present at the site.
% , but that does not generally have to be the case
% (there might be several edges of the same type across a given site).

\subsection{Simulations performed for the model-based analysis}

We performed several simulation suites in order to validate the performance of
ClonalOrigin on our data, and found significant trends. We used the simulation
component of ClonalOrigin to produce these simulations. Each simulated data set
was constructed to be of the same length distribution as used in our data
analysis. In each simulation, sequence data was generated according to the
Jukes-Cantor substitution model 
\citep{Jukes1969}
using a designated recombinant tree defined over
a clonal frame identical to the one inferred for our data (fig.\
\ref{fig:tree5}).

We started by testing accuracy of the estimates of the three population
parameters: population mutation rate ($\theta$), recombination rate ($\rho$),
and average recombinant tract length ($\delta$). We generated ten simulated sets
using ten different recombinant trees sampled from the prior distribution
given the clonal frame and the values of the three population parameters
estimated in data analysis: $\theta=0.081$, $\rho=0.012$, and $\delta=744$ bp.
Estimates of the mutation and recombination rates appeared to be accurately
inferred with average global estimates of $0.081\pm 0.0011$, and $0.011\pm
0.0003$, respectively. The average recombinant tract length were over-estimated at
$919\pm28$ bp as previously observed by \citealp{Didelot2010}.

In order to evaluate significant trends in our data analysis, we generated a
more extensive suite with 100 data sets simulated using 100 recombinant trees
generated from the model prior as the previous simulation setup.  As in the real
data analysis, we compared the average number of recombinant edges of each type
to the corresponding prior expected number (supplementary fig.\
\ref{fig:sim2}). The ratios between inferred and expected edge counts ranged
from 0.64 to 1.6 for all edge types across all simulations. We observed no
notable bias in the estimates. The counts estimated from the 100 simulated sets
were also used to assign empirical $p$-values to the deviations from the prior
portrayed in the heatmap of figure \ref{fig:Heatmap-of-transfers}. We also used
simulations under the prior model to assess the accuracy of our estimates for
recombination intensity. We found a high correlation of 0.79 between inferred
intensities and true ones when restricting the intensity measure to
topology-altering recombinant edges, and a lower correlation of 0.63 for the
total intensity. Therefore, we used the topology-altering intensity for our GO
enrichment tests.

Additionally, we constructed 100 simulated data sets using recombinant trees
sampled from the posterior distribution (selecting every tenth tree in the
posterior sample generated by ClonalOrigin in the real data analysis). Using
these data we were able to assess the effect the prior had on the estimates
produced by ClonalOrigin.
% when the data does not fit the prior. 
Summarizing average recombinant edge counts across
the 100 data sets we observed that, as
expected, these counts were in better agreement with the counts inferred in our
data analysis than were counts inferred from data simulated under the
prior (supplementary fig.\ \ref{fig:sim3}). 
The agreement was notably better for edge types where the inferred
count was less than the prior expected count (e.g., edges from the SPY clade to
SDD). The prior distribution did appear to have an effect on estimated counts
associated with enriched edge types, leading to an under-estimation of the
enrichment level. We concluded, therefore, that the enrichment observed for
SPY/SDE replacing gene transfer was likely under-estimated in our analysis.

\subsection{Simulations performed for the parsimony-based analysis}

We performed separate simulations to understand the power of 
the parsimony-based pipeline to recover events such as gene duplication 
and additive transfer.  The simulations were also used as a null model
for assessing significance of the deviation from prior expectation.

We implemented a new program that simulated all of the four events including
duplication, loss, additive transfer, and replacing transfer.  The program
assumes that each of these events occurs at a (different) constant rate, 
generating gene trees given a species tree and specified rates (see description
below). Using this simulation program, we generated $10^4$ gene trees, setting
the rates of the four types of events to the average rates inferred in our data
analysis (i.e., number of events divided by total tree length). We then
simulated sequence evolution using the HKY substitution model 
\citep{Hasegawa1985}
with base
frequencies and transition/transversion ratio as observed in the real data,
resulting in gapless alignments.
For each of the generated alignments, we performed the entire analysis pipeline,
inferring the gene tree using RAxML, reconciling it using Mowgli, and classifying
additive transfers as described in the main text. This allowed
us to assess the power of the entire pipeline. To create a genome-wide sample of
gene trees we wanted to generate sets of 2,314 gene trees (the same number of
families in our data analysis). Hence, for computational efficiency, we
resampled with replacement 1,000 subsets of 2,314 trees from the original set of
10,000. Comparison of event counts in the true (simulated) gene trees with the
ones reconstructed from the simulated data 
allowed us to assess the accuracy of our pipeline of parsimony-based approach
(supplementary fig.\ \ref{fig:mowgli-sim}).
Comparison of the counts inferred in our data analysis with the
ones inferred from simulated data allowed us to assess significance of deviation
from the null model, as portrayed in figure \ref{fig:Heatmap-of-transfers}B.

\subsection{Program simulating duplication, loss, transfer, and recombination}

The program takes as input a species tree and four rate parameters $D,L,T,$ 
and $R.$
It generates gene trees by starting with a single gene lineage located at the
root of the species tree, and then traversing down branches of the species
tree, allowing for the following events: 
\begin{itemize}
\item If a species tree node is encountered (the root node is the
  first encountered), the gene lineage undergoes {\em speciation} and
  bifurcates into two lineages, each one evolving independently down a
  child branch of the species tree node.  
\item Using a Poisson process at constant rate $D$, a gene lineage
  undergoes {\em duplication} and bifurcates into two lineages, both
  growing independently within the same species branch.  
\item At a constant rate $L$, a gene lineage can be {\em lost} and
  terminate.  
\item At a constant rate $T$, a gene lineage can undergo {\em additive
  transfer}, bifurcating into two lineages, one lineage within
  the current species branch and the second lineage within one
  of the other contemporary species branches chosen uniformly.  Both
  gene lineages continue to evolve independently of one another.  
\item At a constant rate $R$, a gene lineage $a$ can undergo {\em replacing
  transfer}.  From the other contemporary species branches, a second
  gene lineage $b$ (if it exists) is chosen to be replaced.  The gene
  lineage $b$ is terminated and lineage $a$ is bifurcated, one lineage $a_1$
  remaining in the current species and the other $a_2$ transferring to the
  species branch of lineage $b$.  Both lineages $a_1$ and $a_2$ continue
  to evolve independently.
\end{itemize}

If a gene lineage encounters a species tree leaf, the gene lineage terminates
and is denoted an {\em extant gene copy}. Any gene lineage that terminates
before the species tree leaves, is declared extinct and is pruned from the gene
tree.

\subsection{Replacing gene transfers inferred by the parsimony-based method.}

We studied transfers inferred by Mowgli that were not identified as additive
transfers by our criterion. Majority of these inferred
transfer events were expected to be of replacing type. We set out to compare the
inference with our model-based analysis of replacing gene transfers. First, we
compared the set of Gene Ontology (GO) categories associated with the two sets
of inferred transfers. We found that 8 of the 14 categories associated with
replacing gene transfer in the model-based analysis were also associated with
the parsimony-based inferred replacing gene transfer (supplementary tables
\ref{tab:functional} and \ref{tab:go-events-recombining}).  Some disagreement
was expected because of inherent differences of the two inference methods.  We
also computed the correlation between recombination intensity scores and the
number of replacing gene transfers inferred by the parsimony-based approach over
the set of genes present in both analyses.  We found a significant positive
correlation of $0.39$ ($p=2.2\times10^{-16}$) between the two measures.

\renewcommand*{\refname}{Literature Cited}
\bibliographystyle{../../latex/bst/mbe}
\bibliography{siepel-strep}

\clearpage{}\setcounter{figure}{0}
\setcounter{table}{0}

\makeatletter 
\renewcommand{\thefigure}{S\@arabic\c@figure} 
\renewcommand{\thetable}{S\@arabic\c@table} 
\renewcommand{\figurename}{Supplementary Figure}
\renewcommand{\tablename}{Supplementary Table}
\makeatother

\section{Supplementary Tables}

%=============================================================================
\begin{table}[!ht]
\caption[The six bacterial genomes]{
{\bf The six bacterial genomes.}
Each bacterial genome is referred to as the name at the first column.}
\noindent \begin{centering}
\begin{tabular}{cccc}
\hline 
Name & NCBI accession & Size (base pairs) & Reference\\
\hline
SDE1 & CP002215 & 2159491 & Suzuki et al. (2011)\\
SDE2 & NC\_012891 & 2106340 & Shimomura et al. (2011)\\
SDD & CM001076 & 2141837 & Suzuki et al. (2011)\\
SPY1 & NC\_004070 & 1900521 & Beres et al. (2002)\\
SPY2 & NC\_008024 & 1937111 & Beres et al. (2006)\\
SEE & NC\_012471 & 2253793 & Holden et al. (2009)\\
\hline
\end{tabular}
\par\end{centering}
\label{tab:genome}
\end{table}
%=============================================================================

%=============================================================================
\begin{table}[!ht]
\caption[Posterior probability of gene tree topologies]{
{\bf Posterior probability of gene tree topologies.} Frequencies of the eight
most frequent topologies in Newick format are shown.}
\noindent \begin{centering}
\begin{tabular}{cc}
\hline 
Gene tree topologies & Posterior probability\\
\hline
(((SDE1,SDE2),SDD),(SPY1,SPY2)) & 66.7\%\\
(((SDE1,SDE2),(SPY1,SPY2)),SDD) & 9.3\%\\
((SDE1,SDE2),(SDD,(SPY1,SPY2))) & 3.5\%\\
(((SDE1,(SDE2,SDD)),(SPY1,SPY2)) & 1.5\%\\
(((SDE1,SDD),SDE2),(SPY1,SPY2)) & 1.3\%\\
((SDE1,SDD),(SDE2,(SPY1,SPY2))) & 1.2\%\\
((((SDE1,SDE2),SPY1),SPY2),SDD) & 1.0\%\\
((SDE2,SDD),(SDE1,(SPY1,SPY2))) & 1.0\%\\
\hline
\end{tabular}
\par\end{centering}
\label{tab:Gene-tree-topologies}
\end{table}
%=============================================================================

%=============================================================================
\begin{table}[!ht]
\caption[The observed values of recombinant edges]{
{\bf The observed values of recombinant edges.}}  
\noindent \begin{centering}
\begin{tabular}{cccccccccc}
\hline
Edge$^a$ & SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2 \\
\hline
SDE1&8.896 &0.00$^b$ &17.712  & 0.00 & 12.872   & 0 &12.59  &  0.00  &12.53\\
SDE &44.626& 136.49& 42.088 &  0.00& 227.061  &  0& 73.39 &  99.95 & 65.25\\
SDE2&18.099&   0.00&  8.393 &  0.00&  12.360  &  0& 11.97 &   0.00 & 10.36\\
SD  &9.671 &101.74 & 9.464  &31.26 &115.796   & 0 &13.68  & 98.38  &12.86\\
SDD &45.539& 168.58& 49.513 &  0.00& 194.706  &  0& 43.09 &  67.14 & 47.65\\
ROOT&43.553& 602.67& 42.337 &765.53& 802.813  &  0& 98.86 &2146.01 &103.60\\
SPY1&40.800&  11.46& 40.973 &  0.00&   4.371  &  0& 21.15 &   0.00 & 42.35\\
SPY &57.618& 262.08& 66.594 & 30.92& 163.518  &  0& 83.56 & 207.00 & 85.10\\
SPY2&38.837&  11.90& 42.000 &  0.00&   5.111  &  0& 42.45 &   0.00 & 21.19\\
\hline
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$ Rows and columns indicate the donor and the recipient branches,
respectively.\\
$^b$ Zeros are given for branch pairs that are impossible under the model-based
approach.\\
\end{flushleft}
\label{tab:obsheatmap}
\end{table}
%=============================================================================

%=============================================================================
\begin{table}[!ht]
\caption[Heatmap values for replacing gene transfer]{
{\bf Logarithm base 2 of the ratios of the inferred to the expected number of
recombinant edges.} This table lists the values graphically displayed in the
heatmap of figure \ref{fig:Heatmap-of-transfers}A.}
\noindent \begin{centering}
\begin{tabular}{cccccccccc}
\hline
Edge$^a$ & SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2 \\
\hline
SDE1& 0.176  &   NA$^b$  &1.170  &    NA & 0.7564  & NA & 0.67 &    NA & 0.663\\
SDE & -0.118 & 0.081 &-0.202 &     NA&  0.3395 &  NA& -0.15&  0.045& -0.323\\
SDE2& 1.201  &   NA  &0.092  &    NA & 0.6978  & NA & 0.60 &    NA & 0.387\\
SD  & -0.629 & 0.022 &-0.661 &0.00670&  0.0078 &  NA& -0.94& -0.299& -1.030\\
SDD & -0.317 & 0.371 &-0.197 &     NA&  0.0752 &  NA& -1.06& -0.525& -0.915\\
ROOT& -1.745 &-0.698 &-1.786 &0.23039& -0.4854 &  NA& -1.37&  0.497& -1.306\\
SPY1& 1.245  &1.883  &1.251  &    NA &-2.1997  & NA & 0.06 &    NA & 1.062\\
SPY & 0.089  &0.200  &0.298 &0.00098 &-0.7903  & NA &-0.19 &-0.058 &-0.159\\
SPY2& 1.174  &1.938  &1.287  &    NA &-1.9740  & NA & 1.07 &    NA & 0.063\\
\hline
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$ Rows and columns indicate the donor and the recipient branches,
respectively.\\
$^b$ NA's are given for the branch pairs that are impossible under the model.\\
\end{flushleft}
\label{tab:heatmap}
\end{table}
%=============================================================================

%=============================================================================
\begin{table}[!ht]
\caption[Number of genes experiencing replacing transfer]{
{\bf Number of genes experiencing replacing transfer inferred from ClonalOrigin.} 
Numbers of genes with an average posterior probability
greater than 0.6 for recombinant edge types across their sites are 
shown together with
ratios relative to the total number of genes across all edge types.
The same is done for the subset of virulence genes.}
\label{tab:gene-counts-replacing}
\noindent \begin{centering}
\begin{tabular}{lll}
\hline
Transfer type$^a$ & Gene count & Subset of virulence genes
\\
\hline
SDEclade$^b$ $\rightarrow$ SDD & 43 (33.1\%) & 5 (50\%) \\
SDEclade $\rightarrow$ SPYclade$^c$ & 11 (8.5\%) & 1 (10\%) \\
SDD $\rightarrow$ SDEclade & 1 (0.8\%)  & 0 (0\%) \\
SPYclade $\rightarrow$ SDEclade & 59 (45.4\%) & 3 (30\%) \\
SPYclade $\rightarrow$ SPYclade & 16 (12.3\%) & 1 (10\%) \\
\hline
All types & 130  (100\%) & 10   ( 100\%)\\
% &  (11.8\% of total number of genes) & (13\% of virulence genes)\\
\hline
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$Edge types not shown have a zero count. 
Recombinant edges from the root branch are not considered.\\
$^b$SDEclade denotes three branches of SDE1, SDE2 and SDE in the clonal frame.\\
$^c$SPYclade denotes three branches of SPY1, SPY2 and SPY in the clonal frame.\\
\end{flushleft}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
\begin{table}[!ht]
\caption[Number of inferred additive transfer events]{
{\bf Number of inferred additive transfer events of different types.}  
Numbers of inferred additive transfer events of each type across
all genes and subset of virulence genes 
are
shown together with ratios relative to the total number of inferred
additive transfers.}
\label{tab:gene-counts-additive}
%
\noindent \begin{centering}
\begin{tabular}{lll}
\hline
Transfer type$^a$ & Number of events & Subset of virulence genes \\
\hline
SDEclade$^b$ $\rightarrow$ SDEclade & 28 (10.6\%) &  5 (12.5\%)\\
SDEclade $\rightarrow$ SDD & 64 (24.2\%) & 2 (5.0\%)\\
SDEclade $\rightarrow$ SPYclade$^c$ & 28 (10.6\%) &  7 (17.5\%)\\
SDD $\rightarrow$ SDEclade & 29 (10.9\%) &  5 (12.5\%)\\
SDD $\rightarrow$ SPYclade & 17 (6.4\%) &  2 (5.0\%)\\
SD $\rightarrow$  SPYclade &  4 (1.5\%) &  0 (0.0\%)\\
SPYclade $\rightarrow$ SDEclade & 32 (12.1\%) & 10 (25.0\%)\\
SPYclade $\rightarrow$ SDD & 33 (12.5\%) &  3 (7.5\%)\\
SPYclade $\rightarrow$ SD &   3 (1.1\%) &  0 (0.0\%)\\
SPYclade $\rightarrow$ SPYclade & 27 (10.2\%) &  6 (15.0\%)\\
\hline
All types  &  265 (100\%) & 40 (100\%)  \\
\hline
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$Edge types not shown have a zero count.\\
$^b$SDEclade denotes three branches of SDE1, SDE2 and SDE in the clonal frame.\\
$^c$SPYclade denotes three branches of SPY1, SPY2 and SPY in the clonal frame.\\
\end{flushleft}
\end{table}
%=============================================================================

%=============================================================================
% /Users/goshng/Documents/Projects/Mauve/output/cornellf/3/run-analysis/significant-1.txt
\begin{table}[!ht]
\caption[Functional category enrichments for replacing transfers with ClonalOrigin]{
{\bf Functional category enrichments for replacing gene transfers
inferred by the model-based approach$^a$}}
\vspace{1ex}
\noindent \begin{centering}
\begin{tabular}{ccccp{3.5in}}
\hline 
$p^b$ & $q^c$ & Count$^d$ & GO term & Description\\
\hline 
2e-06 & 2e-04 &  38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
3e-06 & 2e-04 &  33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
5e-05 & 3e-03 &  22 & GO:0006633 & fatty acid biosynthetic process\\
9e-05 & 4e-03 &  21 & GO:0008610 & lipid biosynthetic process\\
3e-04 & 9e-03 &  45 & GO:0008643 & carbohydrate transport\\
5e-04 & 1e-02 &  82 & GO:0006412 & translation\\
6e-04 & 1e-02 &  13 & GO:0006814 & sodium ion transport\\
1e-03 & 2e-02 &  40 & GO:0030529 & ribonucleoprotein complex\\
1e-03 & 2e-02 &  12 & GO:0031402 & sodium ion binding\\
1e-03 & 2e-02 & 244 & GO:0008152 & metabolic process\\
2e-03 & 3e-02 & 378 & GO:0016740 & transferase activity\\
2e-03 & 3e-02 &  39 & GO:0003735 & structural constituent of ribosome\\
2e-03 & 3e-02 &  26 & GO:0019843 & rRNA binding\\
4e-03 & 4e-02 &  66 & GO:0005840 & ribosome\\
\hline 
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$Based on recombination intensities restricted to topology-altering
recombinant edges (Methods).\\
$^b$$P$-values based a Mann-Whitney $U$ test.\\
$^c$Values corresponding to FDR$<0.05$ based on the Benjamini-Hochberg method (Methods).\\
$^d$Number of genes in category.\\
\end{flushleft}
\label{tab:functional}
\end{table}
%=============================================================================

%=============================================================================
\begin{table}[!ht]
\caption[Functional categories enriched in replacing transfers with Mowgli]{
{\bf Functional categories enriched in 
replacing transfer events inferred by the parsimony-based approach$^a$}}
\noindent \begin{centering}
\begin{tabular}{ccccp{3.5in}}
\hline 
$p^b$ & $q^c$ & Count$^d$ & GO term & Description \\
\hline 
7e-09 & 3e-06 & 153 & GO:0006412 & translation\\
1e-08 & 3e-06 & 67 & GO:0030529 & ribonucleoprotein complex\\
1e-07 & 2e-05 & 45 & GO:0019843 & rRNA binding\\
2e-06 & 2e-04 & 73 & GO:0003735 & structural constituent of ribosome\\
3e-05 & 3e-03 & 157 & GO:0005840 & ribosome\\
7e-05 & 5e-03 & 221 & GO:0003723 & RNA binding\\
2e-04 & 1e-02 & 11 & GO:0004826 & phenylalanine-tRNA ligase activity\\
3e-04 & 2e-02 & 14 & GO:0015413 & nickel-transporting ATPase activity\\
4e-04 & 2e-02 & 93 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
5e-04 & 2e-02 & 104 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
5e-04 & 2e-02 & 13 & GO:0006265 & DNA topological change\\
5e-04 & 2e-02 & 13 & GO:0003916 & DNA topoisomerase activity\\
5e-04 & 2e-02 & 733 & GO:0005737 & cytoplasm\\
6e-04 & 2e-02 & 10 & GO:0046873 & metal ion transmembrane transporter activity\\
7e-04 & 2e-02 & 12 & GO:0009435 & NAD biosynthetic process\\
8e-04 & 2e-02 & 19 & GO:0016820 & hydrolase activity, acting on acid anhydrides, \\
      &       &    &       &catalyzing transmembrane movement of substances \\
1e-03 & 3e-02 & 20 & GO:0005694 & chromosome\\
2e-03 & 4e-02 & 31 & GO:0006814 & sodium ion transport\\
\hline 
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$Transfer events inferred by Mowgli are considered to be replacing if they do
not pass our criterion for additive transfers (see supplementary fig.\
\ref{fig:calling-transfers}). \\
$^b$$P$-values based a Mann-Whitney $U$ test. \\
$^c$Values corresponding to FDR$<0.05$ based on the Benjamini-Hochberg method (Methods).\\
$^d$Number of genes in category.\\
\end{flushleft}
\label{tab:go-events-recombining}
\end{table}
\clearpage{}
%=============================================================================

\clearpage{}
\section{Supplementary Figures}

%=============================================================================
\begin{sidewaysfigure}[!ht]
\begin{center}
\includegraphics[scale=0.3]{figures/mauve}
\end{center}
\caption[The genome alignment of the five genomes]{\label{fig:mauve}
{\bf The genome alignment of the five genomes.} The top
first and second rows are SDE1 and SDE2 genomes, respectively. The third is SDD
genome. The fourth and fifth rows are SPY1 and SPY2 genomes, respectively.}
\end{sidewaysfigure}
%=============================================================================

%=============================================================================
\begin{figure}[!ht]
\begin{center}
\includegraphics[width=5in]{figures/famsizes}
\end{center}
\caption[Distribution of gene family sizes]{
{\bf Distribution of gene family sizes.}}
\label{fig:famsizes}
\end{figure}
\clearpage{}
%=============================================================================

%=============================================================================
\begin{figure}[!ht]
\includegraphics[scale=0.42]{figures/clonalorigin}
\caption[Recombinant tree in the ClonalOrigin model]{
{\bf Recombinant tree in the ClonalOrigin model.}
(A) A recombinant tree with a clonal frame over our five samples, and four
recombinant edges.  (B) Each recombinant edge is mapped to a genomic region
(each dotted vertical line corresponds to 20\% of the genome). Recombination
intensities (number of edge types present) and gene tree labels are plotted at
the bottom of the map. (C) The local gene trees induced by the recombinant tree. 
Bold edges portray branches of the clonal frame, and dashed edges portray
recombinant edges. See section \textbf{Recombinant tree summaries} in
\textbf{Supplementary Text}.}
\label{fig:clonalorigin}
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
\begin{figure}
\begin{center}
\includegraphics[width=5in]{figures/calling}
\end{center}
\caption[Distinguishing between replacing and additive gene HGT]{
{\bf Distinguishing between replacing and additive horizontal gene transfers.}
We consider a transfer ``additive'' if there exists an extant descendant species
(e.g. C and D) that contains both descendants (bold) and
non-descendants (non-bold) of the transferred gene (black circle).}
\label{fig:calling-transfers}
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
\begin{figure}[!ht]
\begin{center}
\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-theta}

\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-rho}

\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-delta}
\end{center}
\caption[Three scatter plots of population parameters]{
{\bf Three scatter plots of mutation rate, recombination
rate, and average recombinant tract length for the blocks.} The plus signs are mean values
for each block. The density of each value for each block is depicted by black
clouds.  The red dashed lines are the global block-length weighted median of the
three parameter estimates.}\label{fig:scatter3}
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
\begin{sidewaysfigure}

\includegraphics[scale=0.6]{figures/sim2-1}

\includegraphics[scale=0.6]{figures/sim2-2}

\caption[ClonalOrigin simulation using prior of recombinant trees]{
{\bf Simulations of the model-based approach using recombinant trees sampled from the prior distribution.} 
For each recombinant edge type (X-axis), we record the ratio between the average
count of edges as inferred in our data analysis (closed circle; see also
supplementary tables \ref{tab:obsheatmap} \& \ref{tab:heatmap}), and as inferred
across 100 data sets simulated according to the prior distribution (intervals
with open circles). The open circle corresponds to the median across the
different sets, and the intervals correspond to the range between the  5\% and
95\% quantiles.}
\label{fig:sim2}
\end{sidewaysfigure}
\clearpage{}%
%=============================================================================

%=============================================================================
\begin{sidewaysfigure}

\includegraphics[scale=0.6]{figures/sim3-1}

\includegraphics[scale=0.6]{figures/sim3-2}

\caption[ClonalOrigin simulation using posterior of recombinant trees]{
{\bf Simulations of the model-based approach using recombinant trees sampled from the posterior distribution.} 
For each recombinant edge type (X-axis), we record the ratio between the average
count of edges as inferred in our data analysis (closed circle; see also
supplementary tables \ref{tab:obsheatmap} \& \ref{tab:heatmap}), and as inferred
across 100 data sets simulated using posterior trees sampled from the posterior
distribution (intervals with open circles). The open circle corresponds to the
median across the different sets, and the intervals correspond to the range
between the  5\% and 95\% quantiles.}
\label{fig:sim3}
\end{sidewaysfigure}
\clearpage{}%
%=============================================================================

%=============================================================================
\begin{sidewaysfigure}

\includegraphics[scale=0.6]{figures/sim4-1}

\includegraphics[scale=0.6]{figures/sim4-2}

\caption[Mowgli simulation for additive transfers and duplications]{
{\bf The power of recovering additive transfers and duplications using
simulations.} (top) For each pair of species tree branches, we plot the number
of inferred additive transfers in the real data (closed circle) normalized by
the expected number of transfers (computed assuming a constant rate of transfer
along the species tree). The same is done for each of the 1,000 simulated data
sets. The open circle corresponds to the median across the different sets, and
the intervals correspond to the range between the  5\% and 95\% quantiles.
(bottom) For each branch in the species tree, we plot the number of inferred
duplications in the real data (closed circles) normalized by the expected number
of duplications in our simulations (computed assuming a constant rate of
duplications along the species tree).  The same is done for each of the 1,000
simulated data sets. The open circle corresponds to the median across the
different sets, and the intervals correspond to the range between the  5\% and
95\% quantiles.}
\label{fig:mowgli-sim}
\end{sidewaysfigure}
\clearpage{}%
%=============================================================================

\end{document}

### 2011-12-13 sh/scatter-plot-parameter.sh

@@ -95,7 +95,7 @@ plotThreeParameter <- function (f, xlab, ylab, m, logscale) {
 }
 plotThreeParameter ("$S2OUT-theta", "Genomic position on SDE1", "Mutation rate per site", $MEDIAN_THETA, FALSE)
 plotThreeParameter ("$S2OUT-rho", "Genomic position on SDE1", "Recombination rate per site", $MEDIAN_RHO, FALSE)
-plotThreeParameter ("$S2OUT-delta", "Genomic position on SDE1", "Log of tract length", $MEDIAN_DELTA, TRUE)
+plotThreeParameter ("$S2OUT-delta", "Genomic position on SDE1", "Log of average tract length", $MEDIAN_DELTA, TRUE)
 EOF
   Rscript $BATCH_R > $BATCH_R.out 
 }

### 2011-12-13 test/famsize/famsizes.pl

@@ -1,0 +1,40 @@
+open IN, "famsizes.txt" or die "cannot open famsizes.txt $!";
+my %famsize;
+while (<IN>)
+{
+  chomp;
+  my @e = split /\s+/;
+  my $n = $e[1];
+  my $g = $e[2];
+  $g =~ s/(-\d)//; 
+  if (exists $famsize{$g})
+  {
+    unless ($famsize{$g} == $n)
+    {
+      print STDERR "The family has two fragments with different sizes $g - $famsize{$g} vs $n\n" 
+    }
+    # die "The family has two fragments with different sizes $g" 
+      # unless $famsize{$g} == $n;
+  }
+  else
+  {
+    $famsize{$g} = $n;
+  }
+}
+close IN;
+
+open OUT, ">famsizes-only.txt" or die "cannot open famsizes-only.txt $!";
+for my $i (keys %famsize)
+{
+  print OUT "$i\t$famsize{$i}\n";
+}
+close OUt;
+
+
+=comment
+     59 ORTHOMCL0-1
+     59 ORTHOMCL0-2
+      5 ORTHOMCL1000
+      6 ORTHOMCL100-1
+      6 ORTHOMCL1001-1
+=comment
